### 2017-06-12 notes/2017-02-10-single-molecule/report/report.tex
Add notes history.

\documentclass[11pt]{article}

%%%%%%%%%%%%
% Packages %
%%%%%%%%%%%%
\hyphenpenalty=10000
\usepackage{tocloft}
\renewcommand\cftsecleader{\cftdotfill{\cftdotsep}}

\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{epstopdf}
\usepackage{braket}
\usepackage{upgreek}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{amssymb,latexsym,amsmath,gensymb}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{float}
\usepackage{enumitem}
\usepackage{pdflscape}
\usepackage{url}
\usepackage{tikz, calc}
\usetikzlibrary{shapes.geometric, arrows, calc}
\tikzstyle{norm} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{arrow} = [thick, ->, >=stealth]

\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
\providecommand{\mb}[1]{\mathbf{#1}}
\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
\providecommand{\bs}[1]{\boldsymbol{#1}} 
\providecommand{\intinf}{\int_{-\infty}^{\infty}}
\providecommand{\fig}[4]{
  % filename, width, caption, label
\begin{figure}[h]
 \captionsetup{width=1.0\linewidth}
 \centering
 \includegraphics[width = #2\textwidth]{#1}
 \caption{#3}
 \label{fig:#4}
\end{figure}
}

\newcommand{\tensor}[1]{\overset{\text{\tiny$\leftrightarrow$}}{\mb{#1}}}
\newcommand{\tunderbrace}[2]{\underbrace{#1}_{\textstyle#2}}
\providecommand{\figs}[7]{
  % filename1, filename2, caption1, caption2, label1, label2, shift
\begin{figure}[H]
\centering
\begin{minipage}[b]{.4\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#1}
  \captionsetup{justification=justified, singlelinecheck=true}
  \caption{#3}
  \label{fig:#5}
\end{minipage}
\hspace{2em}
\begin{minipage}[b]{.4\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#2}
  \vspace{#7em}
  \captionsetup{justification=justified}
  \caption{#4}
  \label{fig:#6}
\end{minipage}
\end{figure}
}
\makeatletter

\providecommand{\code}[1]{
\begin{center}
\lstinputlisting{#1}
\end{center}
}

%%%%%%%%%%%
% Spacing %
%%%%%%%%%%%
% Margins
\usepackage[
top    = 1.5cm,
bottom = 1.5cm,
left   = 1.5cm,
right  = 1.5cm]{geometry}

% Indents, paragraph space
\usepackage{parskip}

% Section spacing
\usepackage{titlesec}
\titlespacing*{\title}
{0pt}{0ex}{0ex}
\titlespacing*{\section}
{0pt}{0ex}{0ex}
\titlespacing*{\subsection}
{0pt}{0ex}{0ex}
\titlespacing*{\subsubsection}
{0pt}{0ex}{0ex}

% Line spacing
\linespread{1.1}

%%%%%%%%%%%%
% Document %
%%%%%%%%%%%%
\begin{document}
\title{\vspace{-2.5em}Single Molecule Fluorescence Microscopy Notes\vspace{-1em}}
\author{Talon Chandler}% and Patrick La Rivi\`ere}
\date{\vspace{-1em}February 10, 2017\vspace{-1em}}
\maketitle

\section{Introduction}
These notes develop the forward model for a wide class of single molecule
fluorescence microscopes. We split the imaging process into four
parts---illumination, excitation, emission, and detection---and analyze each
part individually. For each part of the imaging process, we will perform a
general analysis, relate the analysis to current techniques, and mention
possible improvements.

\subsection{Notation}
We use boldface for vectors $\mathbf{r}$, hats for unit vectors
$\mathbf{\hat{r}}$, normal face for vector lengths $r = |\mathbf{r}|$, tildes
for matrices $\mb{\tilde{O}}$, and double arrows for tensors $\tensor{\mb{G}}$.

\subsection{Coordinates}
First, we place the origin at the geometric focus of the microscope. If there
are multiple objectives we assume that the foci are perfectly aligned. Next, we
choose a fixed laboratory coordinate system using Cartesian unit vectors
$\mb{r} = x\mh{x} + y\mh{y} + z\mh{z}$.

\section{Illumination}
\subsection{General Analysis}
We view illumination as the process of establishing an electric field
distribution $\mb{E}(\mb{r}, t)$ near the origin. In these notes we
will use the complex spatial electric field $\mb{E}(\mb{r})$ related to the
time dependent electric field by
\begin{align*}
  \mb{E}(\mb{r}, t) = \text{Re}\{\mb{E}(\mb{r})e^{-j\omega t}\}. 
\end{align*}
We will refer to the complex spatial electric field as the electric field.
Analysis of the illumination pattern consists of finding the input electric field
$\mb{E}_{\text{in}}(\mb{r})$ near the focal point.

\subsection{Current Techniques}
\subsubsection{Laser K\"{o}hler Illumination}
Laser K\"{o}hler illumination is accomplished by focusing a laser beam onto the
center of the back focal plane of the objective as shown in Figure 2(a) of
\cite{backlund}. Under laser K\"{o}hler illumination a constant electric field
is established in the focal region. That is
\begin{align*}
  \mb{E}_{\text{in}}(\mb{r}) = \mb{P}
\end{align*}
where $\mb{P}$ is the 3D vector pointing along the polarization direction of the
laser. Under laser K\"{o}hler illumination the electric field in the focal
region is always transverse, and the electric field is constant throughout the
field of view.

If we use a modulated polarizing element on the illumination side, its pass
direction will change $\mb{P}$.

Notice that both $\mb{E}$ and $\mb{P}$ are 3D vectors, not 2D Jones vectors. We
will continue using 3D vectors throughout these notes so that we can consider
illumination patterns from any angle.

\subsubsection{Laser Critical Illumination}
Laser critical illumination is accomplished by illuminating the entire back
focal plane of the objective with a laser as shown in shown in Figure 2b) of
\cite{backlund}. Under laser critical illumination, the electric field varies
near the focal plane. The electric field at every point can be calculated using
the Richards-Wolf diffraction integral \cite{richards}. To calculate the
electric field at a point we calculate the electric field at that point due to a
single point on the lens, then integrate the contributions from each point while
keeping track of the phase to allow for interference. See Chapter 3 of
\cite{nov} for a complete calculation of the focal fields created by a spherical
lens including consideration of under filling the back aperture and higher order
laser modes.

If the back aperture receives uniform illumination, then the electric fields in
the focal plane will be transverse and the electric fields away from the focal
plane will have a transverse component. See Figures 8-10 of \cite{youngworth} for
images of the fields.

Laser critical illumination only creates an appreciable electric field near the
focal spot, so an image must be built up by scanning.

We can also illuminate the sample with a cylindrical lens. The electric field
pattern created by a cylindrical lens can be found in \cite{youngworth}.

Under laser critical illumination the electric field near the focal plane can be
modified by modulating the intensity and phase in the back focal plane.
Axial patterning, radial patterning, and ``bone'' patterning are some modulation
patterns that have been explored \cite{debarre}.

\subsection{Possible Techniques}
\subsubsection{Lamp Illumination}
Lamp illumination could be used to provide inexpensive and novel illumination
patterns. Under laser illumination we can establish a fixed electric field
distribution, but under lamp illumination we need to consider the electric field
at each point as a vector stochastic process. Analyzing the vector stochastic
process requires a statistical optics approach that includes polarization
effects.

The complete theory of coherence and polarization was completed in 2003
\cite{wolf}, and has not, to my knowledge, been applied to single molecule
microscopy problems.

\section{Excitation}
\subsection{General Analysis}
Consider a two-level quantum system in the dipole approximation---an excellent
model for many fluorophores. Each transition can be described by a
\textit{transition dipole moment}, a vector that points in a fixed direction
with respect to the molecule's molecular structure. We denote the absorption
dipole moment by $\bs{\mu}_{\text{abs}}$ (called $\bs{\mu}_{21}$ in quantum
mechanics) and the emission dipole moment by $\bs{\mu}_{\text{em}}$
($\bs{\mu}_{12}$). The transition dipole moments can be calculated analytically
for simple systems using quantum mechanics, calculated numerically for complex
systems using Hartree-Fock theory or density functional theory, or measured
experimentally.

When a fluorophore with transition dipole moments $\bs{\mu}_{\text{em}}$ and
$\bs{\mu}_{\text{abs}}$ is exposed to a weak electric field $\mb{E}$ that is
constant near the fluorophore, the electric field creates an
$\textit{induced dipole moment}$$\ \bs{\mu}$ where
\begin{align*}
  \bs{\mu} \propto \bs{\mu}_{\text{em}}\bs{\mu}_{\text{abs}}^{\dagger}\mb{E}_{\text{in}}. 
\end{align*}
The relationship between the incident electric field and the induced dipole
moment is summarized by the \textit{polarizability tensor}
$\tensor{\bs{\alpha}}$ where
\begin{align*}
  \bs{\mu} &= \tensor{\bs{\alpha}}\mb{E}_{\text{in}}\\
  \tensor{\bs{\alpha}} &\propto \bs{\mu}_{\text{em}}\bs{\mu}_{\text{abs}}^{\dagger}
\end{align*}

If we assume $\bs{\mu}_{\text{em}}
=\bs{\mu}_{\text{abs}}$ then the induced dipole can only be created along a
single axis relative to the molecule's structure. This assumption is invalid for
many fluorophores---even GFP's transition dipole moments are a few degrees apart
\cite{rosell2003}. Some crystals have excitation and emission transition dipole
moments that are orthogonal \cite{koberling}!

General reference for this section: Appendix A of \cite{nov}.

\subsection{Current Techniques}
Many single molecule orientation and location estimation experiments make the
unstated assumption that $\bs{\mu}_{\text{em}} =\bs{\mu}_{\text{abs}}$. Under
this assumption, the induced dipole direction is identical to the directions of
the transition dipole moments.

Many groups approach the excitation analysis from a ``power absorbed'' or a
``probability'' viewpoint. In the their 2014 review \cite{backer}, the Moerner
group says that ``the probability of absorption is proportional to
$|\bs{\mu}_{\text{abs}}\cdot\mb{E}|^2$''. This approach is identical to the
induced dipole approach above. We deal with classical electric fields and
induced dipole moments, while they deal with probabilities of photon absorption. We
will take the modulus squared of the electric field during detection, so the
approaches are identical.

\subsection{Possible Techniques}
\subsubsection{Estimating The Induced Dipole Moment}
With the analysis above, we can attempt to estimate the complete induced dipole
moment, not just the orientation of the induced dipole moment.

\subsubsection{Simultaneous Analysis of Illumination and Detection}
Estimating the induced dipole moment allows us to simultaneously analyze the
illumination and detection paths. To my knowledge, the illumination and
detection paths have only been considered individually. A simultaneous analysis
could allow us to find unique ways of modulating the illumination and detection
paths together.

\subsubsection{Estimating Both Transition Dipole Moments}
If $\bs{\mu}_{\text{em}} \neq\bs{\mu}_{\text{abs}}$, then we may find techniques
for simultaneously measuring both $\bs{\mu}_{\text{em}}$ and
$\bs{\mu}_{\text{abs}}$. Measured both transition dipole moments could be useful
for characterizing fluorophore and for measuring the effects of an environment on
a fluorophore.

\section{Emission}
\subsection{General Analysis}
\fig{../figures/locations/locations.pdf}{.27}{Position and orientation
  vectors.}{locations} Consider an induced dipole moment $\bs{\mu}$ at
$\mathbf{r}'$ in a medium with index of refraction $n_1$ with an as shown in
Figure \ref{fig:locations}. The electric field distribution $\mathbf{E}_{\text{obj}}(\mb{r})$ is related to $\mathbf{r}'$ and $\bs{\mu}$ by the Green's
tensor \cite{nov}
\begin{align}
  \mb{E}_{\text{obj}}(\mb{r}) &= \tensor{\mb{G}}(\mb{r}, \mb{r'})\bs{\mu}\\
  \tensor{\mb{G}}(\mb{r}, \mb{r'}) &= \left[\tensor{\mb{I}} + \frac{1}{k^2}\nabla\nabla\right]\frac{e^{ikn_1R}}{4\pi R}.
\end{align}
Notice that $\mb{E}_{\text{obj}}$ and $\mb{E}_{\text{in}}$ are field with different
frequencies, so they effectively don't interfere.

After evaluating the gradients and only keeping the terms that persist in the
far field ($R\gg\lambda$), the Green's tensor becomes
 \begin{align}
   \tensor{\mb{G}}_{FF}(\mb{r}, \mb{r'}) = \left[\tensor{\mb{I}} - \frac{\mb{R}\mb{R}^{\dag}}{R^2}\right]\frac{e^{ikn_1R}}{4\pi R}. \label{eq:ff}
 \end{align}
 When the dipole is near the origin and we are considering the electric fields
 at points far from the origin ($r\gg r'$), we can use the approximations
 $\mb{R} \approx \mb{r}$ and $R \approx r - \mb{r'}\cdot\hat{\mb{r}}$. Using
 these approximations in Equation~\ref{eq:ff} with
 $R = r - \mb{r'}\cdot\hat{\mb{r}}$ in the phase term and $R = r$ everywhere
 else gives
 \begin{align}
   \tensor{\mb{G}}_{FF}(\mb{r}, \mb{r'}) = \left[\tensor{\mb{I}} - \mb{\hat{r}}\mb{\hat{r}}^{\dag}\right]\frac{e^{ikn_1(r - \mb{r'}\cdot\hat{\mb{r}})}}{4\pi r}. \label{eq:approx}
 \end{align}

\subsubsection{Aside: Interpreting The Green's Tensor}
The Green's tensor acts on a vector (the dipole moment) and outputs a
vector field (the electric field). When we choose a particular point (say $r=1$,
$\theta_z=0$, $\phi_{xy}=0$), then the Green's tensor becomes a matrix and we
can find the electric field at that point by matrix multiplication.

Alternatively, the Green's tensor can be interpreted as a mathematical shorthand
for the $\mb{r}\times\bs{\mu}\times\mb{r}$ method \cite{fourkas} of calculating
the fields created by a dipole moment.

\subsection{Current Techniques}
 Compare equation \ref{eq:approx} with equations 7 and 9 in \cite{backer}. While they
 consider a dipole at the origin and multiply by a phase term to consider
 defocus, the formulation above considers a dipole at an arbitrary location with
 respect to the origin. This generalization allows us to consider localizing the
 dipole from orthogonal views.

 Notice that equation~\ref{eq:approx} can be evaluated using any convenient coordinate
 system. For example, if we express $\mb{r}$ in spherical coordinates
 [$\mb{r} = r\cos(\phi)\sin(\theta)\mb{\hat{x}} +
 r\sin(\phi)\sin(\theta){\mb{\hat{y}} + r\cos(\theta)\mb{\hat{z}}}$]
 and $\mb{r'}$ in Cartesian coordinates
 [$\mb{r'} = x'\mb{\hat{x}} + y'\mb{\hat{y}} + z'\mb{\hat{z}}$] then:
 \begin{align}
   \tensor{\mb{G}}_{FF}(\mb{r}, \mb{r'}) &=
   \begin{bmatrix}
     1 - \cos^2(\phi)\sin^2(\theta) & -\cos(\phi)\sin(\phi)\sin^2(\theta) & -\cos(\phi)\cos(\theta)\sin(\theta)\\
     -\cos(\phi)\sin(\phi)\sin^2(\theta) & 1 - \sin^2(\phi)\sin^2(\theta)  & -\sin(\phi)\cos(\theta)\sin(\theta) \\
     -\cos(\phi)\cos(\theta)\sin(\theta) & -\sin(\phi)\cos(\theta)\sin(\theta) & 1- \cos^2(\theta)
   \end{bmatrix}\frac{\Pi(\mb{r}, \mb{r'})}{4\pi r}\label{eq:gt}\\
   \Pi(\mb{r}, \mb{r'}) &= \text{exp}\left[ikn_1(r - \{x'\sin(\theta)\cos(\phi) + y'\sin(\theta)\sin(\phi) + z'\cos(\theta)\})\right] \label{eq:phase}
 \end{align}
 Compare equation~\ref{eq:phase} with equation 3 in \cite{agrawal}. We have
 derived the same phase expression except we don't ignore the $e^{ikn_1r}$ term, and we
 correct the sign errors on the $x'$ and $y'$ terms.

 \subsection{Possible Techniques}
 \subsubsection{Consideration of Dipoles At Any Position}
 The general analysis allows us to consider an induced dipole at any position. Existing
 analyses only consider dipoles along the optic axis \cite{backer}.

\section{Detection}
\subsection{General Analysis}
\subsubsection{Coordinates}
When we are analyzing a microscope with a single optic axis, we typically choose
$\mh{z}$ parallel to the optic axis and define cylindrical and spherical
coordinates with respect to $\mh{z}$. In these notes, we want to consider
arbitrary optic axes, so we will define spherical and cylindrical coordinates
with respect to the optic axis.

To specify a set of spherical coordinates called the $\mb{\hat{s}}\mb{\hat{a}}$
spherical coordinates, we choose a zenith unit vector $\mb{\hat{s}}$ along the
optic axis and an orthogonal unit vector $\mb{\hat{a}}$ then parameterize any
vector $\mb{r}$ using the length of the vector ($r$), the angle between
$\mb{\hat{s}}$ and $\mb{r}$ ($\theta_{\mb{\hat{s}}}$), and the angle between
$\mb{\hat{a}}$ and the projection of $\mb{r}$ into the plane spanned by
$\mb{\hat{a}}$ and $\mb{\hat{s}} \times \mb{\hat{a}}$ in the right-handed
$\mb{\hat{s}}$ direction ($\phi_{\mb{\hat{a}},\mb{\hat{s}}\times\mb{\hat{a}}}$).
For example, the $\mb{\hat{z}}\mb{\hat{x}}$ and $\mb{\hat{x}}\mb{\hat{y}}$
spherical coordinates are shown in Figures \ref{fig:frames}(a) and
\ref{fig:frames}(b).

Similarly, to specify a set of cylindrical coordinates called the
$\mb{\hat{s}}\mb{\hat{a}}$ cylindrical coordinates, we choose a zenith unit
vector $\mb{\hat{s}}$ along the optic axis and an orthogonal unit vector $\mb{\hat{a}}$ then
parameterize any vector $\mb{r}$ using the length of the vector projected into
the $\mb{\hat{s}}$ direction ($s_{\mh{s}}$), the length of the vector projected into the
plane spanned by $\mb{\hat{a}}$ and $\mb{\hat{s}} \times \mb{\hat{a}}$
($\rho_{\mb{\hat{a}},\mb{\hat{s}}\times\mb{\hat{a}}}$), and the angle between
$\mb{\hat{a}}$ and the projection of $\mb{r}$ into the plane spanned by
$\mb{\hat{a}}$ and $\mb{\hat{s}} \times \mb{\hat{a}}$ in the right-handed
$\mb{\hat{s}}$ direction ($\phi_{\mb{\hat{a}},\mb{\hat{s}}\times\mb{\hat{a}}}$).
For example, the $\mb{\hat{z}}\mb{\hat{x}}$ and $\mb{\hat{x}}\mb{\hat{y}}$
spherical coordinates are shown in Figures \ref{fig:frames}(c) and
\ref{fig:frames}(d).

% Equation \ref{eq:frames} shows the relationships between the Cartesian
% coordinates and the coordinate systems show in Figure \ref{fig:frames}.

% \begin{align}
%   \begin{bmatrix} x\\y\\z \end{bmatrix} =
%   \begin{bmatrix} r\cos(\phi_{\mh{x}\mh{y}})\sin(\theta_{\mh{z}})\\r\sin(\phi_{\mh{x}\mh{y}})\sin(\theta_{\mh{z}})\\r\cos(\theta_{\mh{z}}) \end{bmatrix} =
%   \begin{bmatrix} r\cos(\theta_{\mh{x}})\\r\cos(\phi_{\mh{y}\mh{z}})\sin(\theta_{\mh{x}})\\r\sin(\phi_{\mh{y}\mh{z}})\sin(\theta_{\mh{x}}) \end{bmatrix} =
%   \begin{bmatrix} \rho_{\mh{x}\mh{y}}\cos(\phi_{\mh{x}\mh{y}})\\\rho_{\mh{x}\mh{y}}\sin(\phi_{\mh{x}\mh{y}})\\s_{\mh{z}} \end{bmatrix} =
%   \begin{bmatrix} s_{\mh{x}}\\\rho_{\mh{y}\mh{z}}\cos(\phi_{\mh{y}\mh{z}})\\\rho_{\mh{y}\mh{z}}\sin(\phi_{\mh{y}\mh{z}}) \end{bmatrix}
%   \label{eq:frames}
% \end{align} 
\begin{figure*}[h]
 \captionsetup{width=1.0\linewidth}
 \centering
 \begin{subfigure}[t]{0.25\textwidth}
   \centering
   \includegraphics[width = 1.0\textwidth]{../figures/frames/frames_a.pdf}
   \caption{$\hat{\mathbf{z}}\mb{\hat{x}}$ spherical coordinates.}
   \label{fig:frames_a}
 \end{subfigure}%
 ~
 \begin{subfigure}[t]{0.25\textwidth}
   \centering
   \includegraphics[width = 1.0\textwidth]{../figures/frames/frames_b.pdf}
   \caption{$\hat{\mathbf{x}}\mb{\hat{y}}$ spherical coordinates.}
   \label{fig:frames_b}
 \end{subfigure}%
~
 \begin{subfigure}[t]{0.25\textwidth}
   \centering
   \includegraphics[width = 1.0\textwidth]{../figures/frames/frames_d.pdf}
   \caption{$\hat{\mathbf{z}}\mb{\hat{x}}$ cylindrical coordinates.}
   \label{fig:frames_c}
 \end{subfigure}%
  \begin{subfigure}[t]{0.25\textwidth}
   \centering
   \includegraphics[width = 1.0\textwidth]{../figures/frames/frames_c.pdf}
   \caption{$\hat{\mathbf{x}}\mb{\hat{y}}$ cylindrical coordinates.}
   \label{fig:frames_c}
 \end{subfigure}%

 \caption{Coordinate systems.}
 \label{fig:frames}
\end{figure*}

\subsubsection{Electric Field In The Back Focal Plane}
We can model the action of the objective on the electric fields by a position
dependent rotation matrix.

A right handed rotation about a vector $\mh{u} = u_x\mh{x} + u_y\mh{y} + u_z\mh{z}$
by angle $\theta$ can be represented by the rotation matrix
\begin{align}
\tilde{\mb{R}}_{\mh{u}}(\theta) &= \begin{bmatrix}
  \cos\theta + u_x^2(1 - \cos\theta) & u_xu_y(1 - \cos\theta) - u_z\sin\theta & u_xu_z(1 - \cos\theta) + u_y\sin\theta\\
  u_yu_x(1-\cos\theta) + u_z\sin\theta & \cos\theta + u_y^2(1 - \cos\theta) & u_yu_z(1 - \cos\theta) - u_x\sin\theta\\
  u_zu_x(1 - \cos\theta) - u_y\sin\theta & u_zu_y(1 - \cos\theta) + u_x\sin\theta &  \cos\theta + u_z^2(1 - \cos\theta)
\end{bmatrix}\\
\tilde{\mb{R}}_{\mh{u}}(\theta) &= \cos\theta\tilde{\mb{I}} + \sin\theta\left[\mh{u}\right]_{\times} + (1 - \cos\theta)\mh{u}\mh{u}^{\dagger}
\end{align}
where $\left[\mh{u}\right]_{\times}$ is the cross product matrix of $\mh{u}$ \cite{rotation}.

A lens rotates electric fields into the plane perpendicular to the optic
axis (equivalently, a lens rotates each ray parallel to the optic axis). The
rotation matrix that describes the rotation of the electric field at $\mh{r}$ by a
lens with optic axis $\mh{s}$ is
\begin{align}
  \tilde{\mb{R}}_{\mh{r}\times\mh{s}}(\text{acos}(\mh{r}\cdot\mh{s})) = (\mh{r}\cdot\mh{s})\tilde{\mb{I}} + \sqrt{1 - (\mh{r}\cdot\mh{s})^2}\left[\mh{r}\times\mh{s}\right]_{\times} + (1 - \mh{r}\cdot\mh{s})(\mh{r}\times\mh{s})(\mh{r}\times\mh{s})^{\dagger}
\end{align}

To conserve energy, we need to multiply the rotation matrix by a factor of $\sqrt{\frac{n_1}{n_0(\mh{r}\cdot\mh{s})}}$ \cite{nov} where $n_0$ is the index of refraction of the lens. Finally, we truncate the electric fields to simulate the finite numerical aperture of the lens. The final matrix that represents the lens is given by
\begin{align}
  \mb{\tilde{O}}_{\mh{s}} = \sqrt{\frac{n_1}{n_o(\mh{r}\cdot\mh{s})}}\Pi\left(\frac{|\mb{r} - (\mb{r}\cdot\mh{s})\mh{s}|}{\rho_{\text{max}}}\right)  \tilde{\mb{R}}_{\mh{r}\times\mh{s}}(\text{acos}(\mh{r}\cdot\mh{s}))\label{eq:2}
\end{align}
  where $|\mb{r} - (\mb{r}\cdot\mh{s})\mh{s}|$ is the shortest distance from $r$ to the optic axis, $\rho_{\text{max}}$ is the
  radius of the objective lens
  ($\rho_{\text{max}} =
  f_{\text{obj}}\tan\left(\arcsin\left(\frac{\text{NA}}{n_1}\right)\right)$),
  and
\begin{align}
  \Pi(x) = \begin{cases} 1, |x| < 1\\ 0, |x| > 1\end{cases}.
\end{align}

Now that we can compute the effect of the objective lens, we can compute the
electric field in the back focal plane
\begin{align}
  \mb{E}_{\text{bfp}} &= \mb{\tilde{O}}_{\text{obj}}\tensor{\mathbf{G}}_{FF}\bs{\mu}
  \end{align}
  If we place a polarizing element in the back focal plane, $\mb{E}_{\text{bfp}}$ will
  be projected onto the pass direction of the polarizing element.

The intensity in the back focal plane is given by
\begin{align}
  I_{\text{bfp}}(\mb{r}) = \mb{E}_{\text{bfp}}^{\dag}\mb{E}_{\text{bfp}}
  = |\mb{E}_{\text{bfp}}|^2.
\end{align}

\subsubsection{Electric Field In The Image Plane}
If $f_{\text{obj}} \ll f_{\text{tube}}$, then the paraxial approximation applies
and
\begin{align}
  \mb{E}_{\text{img}}(\mb{r''}) = \mathcal{F}_{3D}\{\mb{E}_{\text{bfp}}\}|_{\lambda f_{\text{tube}}} = \int_{\mathbb{R}^3}\mb{E}_{\text{bfp}}e^{-i\frac{k}{f_{\text{tube}}}\mb{r}\cdot\mb{r''}}d\mb{r}.
\end{align}
where $\mb{r''}$ is the position in the image plane. If we place a polarizing
element in the imaging plane, $\mb{E}_{\text{img}}$ will be projected onto
the pass direction of the polarizing element.

Finally, the intensity in the image plane is
\begin{align}
  I_{\text{img}}(\mb{r''}) = \mb{E}_{\text{img}}^{\dag}\mb{E}_{\text{img}}
  = |\mb{E}_{\text{img}}|^2.
\end{align}

Notice that we take the 3D Fourier transform instead of the usual 2D. The
objective lens rotates the electric fields to the transverse plane, so the
usual 2D FT is equivalent to the 3D FT evaluated in any transverse plane. I've
used the 3D FT to simplify the notation and show that we can use the same
operation for any optic axis.

Also note that the Fourier transform is acting on a vector field instead of the
usual scalar field. The FT of a vector field is just the usual scalar FT applied
to each component of the vector.

\subsection{Current Techniques}
\subsubsection{Single View}
If the optic axis is along the $\mh{z}$ axis and we have an infinite numerical aperture, the objective lens matrix reduces to
\begin{align}
  \mb{\tilde{O}}_{\mb{\hat{z}},\text{obj}} &= 
  \sqrt{\frac{n_1}{n_o\cos(\theta_z)}}\begin{bmatrix}
    \cos(\theta_z)\cos^2(\phi_{xy}) + \sin^2(\phi_{xy}) & (\cos(\theta_z) - 1)\sin(\phi_{xy})\cos(\phi_{xy}) & -\sin(\theta_z)\cos(\phi_{xy})\\
    (\cos(\theta_z) - 1)\sin(\phi_{xy})\cos(\phi_{xy}) & \cos(\theta_z)\sin^2(\phi_{xy}) + \cos^2(\phi_{xy}) & \sin(\theta_z)\sin(\phi_{xy})\\
    \sin(\theta_z)\cos(\phi_{xy}) & \sin(\theta_z)\sin(\phi_{xy}) & \cos(\theta_z)
  \end{bmatrix}. \label{eq:1}
\end{align}
Equation \ref{eq:1} is identical to equation 11 in \cite{backer} and equation 2
in Fourkas' paper \cite{fourkas}. Fourkas sums the intensity over the entire
image plane, while we consider the complete imaging model. In our formulation,
we combine the effects of NA, the apodization, and the rotation of the lens into
a single matrix $\tilde{\mb{O}}$ (Equation \ref{eq:2}).

\subsubsection{Back Focal Plane Manipulation}
The Moerner group has used a variety of phase and amplitude masks in the back
focal plane of the detection path \cite{backer}. These techniques effectively change
the point spread function of the detection system.

\subsection{Possible Techniques}
\subsubsection{Arbitrary Views Of A Dipole Near The Focal Point}
The work in the general analysis section allows us to calculate the images created by
multiple views along arbitrary optic axes. This allows us to analyze new geometries
like diSPIM.

\section{Complete Forward Model}
Combining all of the previous sections gives the complete forward
model:
\begin{align}
  I_{\text{img}}(\mb{r''}) = \left|\mathcal{F}_{3D}\left\{
    \mb{\tilde{O}}_{\text{obj}}\tensor{\mathbf{G}}_{FF}\tensor{\bs{\alpha}}\mb{E}_{\text{in}}
  \right\}
  \right|^2.\label{eq:5}
\end{align}
where
\begin{align*}
  \mb{E}_{\text{in}}\ & \text{is the incident electric field pattern established by the illumination optics. It is a function of the input light}\\ & \text{source, input light pattern (phase, amplitude, polarization modulation), and input light optics (spherical,}\\ & \text{cylindrical lens). }\\
  \tensor{\bs{\alpha}}\ & \text{is the polarizability tensor. It is a function of
                          the fluorophore type and the fluorophore's orientation.}\\
  \tensor{\mb{G}}_{FF}\ & \text{is the Green's tensor that generates the electric fields
                          far from the induced dipole moment. It is a function}\\ & \text{of the
                          dipole moment's position and the refractive index of the
                          medium.}\\
  \mb{\tilde{O}}_{\text{obj}}\ & \text{is the matrix that describes the action of the
                                 objective lens. It is a function of the objective's
                                 focal length,}\\ &\text{NA, and optic axis.}
\end{align*}

\bibliography{report}{}
\bibliographystyle{unsrt}

\end{document}

### 2017-06-12 notes/2017-03-24-illumination-update2/report/report.tex
Add notes history.

\documentclass[11pt]{article}

%%%%%%%%%%%%
% Packages %
%%%%%%%%%%%%
\hyphenpenalty=10000
\usepackage{tocloft}
\renewcommand\cftsecleader{\cftdotfill{\cftdotsep}}
\def\undertilde#1{\mathord{\vtop{\ialign{##\crcr
$\hfil\displaystyle{#1}\hfil$\crcr\noalign{\kern1.5pt\nointerlineskip}
$\hfil\tilde{}\hfil$\crcr\noalign{\kern1.5pt}}}}}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{epstopdf}
\usepackage{braket}
\usepackage{upgreek}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{amssymb,latexsym,amsmath,gensymb}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{float}
\usepackage{enumitem}
\usepackage{pdflscape}
\usepackage{url}
\usepackage{tikz, calc}
\usetikzlibrary{shapes.geometric, arrows, calc}
\tikzstyle{norm} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{arrow} = [thick, ->, >=stealth]

\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
\providecommand{\mb}[1]{\mathbf{#1}}
\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
\providecommand{\bs}[1]{\boldsymbol{#1}} 
\providecommand{\intinf}{\int_{-\infty}^{\infty}}
\providecommand{\fig}[4]{
  % filename, width, caption, label
\begin{figure}[h]
 \captionsetup{width=1.0\linewidth}
 \centering
 \includegraphics[width = #2\textwidth]{#1}
 \caption{#3}
 \label{fig:#4}
\end{figure}
}

\newcommand{\tensor}[1]{\overset{\text{\tiny$\leftrightarrow$}}{\mb{#1}}}
\newcommand{\tunderbrace}[2]{\underbrace{#1}_{\textstyle#2}}
\providecommand{\figs}[7]{
  % filename1, filename2, caption1, caption2, label1, label2, shift
\begin{figure}[H]
\centering
\begin{minipage}[b]{.45\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#1}
  \captionsetup{justification=justified, singlelinecheck=true}
  \caption{#3}
  \label{fig:#5}
\end{minipage}
\hspace{2em}
\begin{minipage}[b]{.45\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#2}
  \vspace{#7em}
  \captionsetup{justification=justified}
  \caption{#4}
  \label{fig:#6}
\end{minipage}
\end{figure}
}
\makeatletter

\providecommand{\code}[1]{
\begin{center}
\lstinputlisting{#1}
\end{center}
}

%%%%%%%%%%%
% Spacing %
%%%%%%%%%%%
% Margins
\usepackage[
top    = 1.5cm,
bottom = 1.5cm,
left   = 1.5cm,
right  = 1.5cm]{geometry}

% Indents, paragraph space
\usepackage{parskip}

% Section spacing
\usepackage{titlesec}
\titlespacing*{\title}
{0pt}{0ex}{0ex}
\titlespacing*{\section}
{0pt}{0ex}{0ex}
\titlespacing*{\subsection}
{0pt}{0ex}{0ex}
\titlespacing*{\subsubsection}
{0pt}{0ex}{0ex}

% Line spacing
\linespread{1.1}

%%%%%%%%%%%%
% Document %
%%%%%%%%%%%%
\begin{document}
\title{\vspace{-2.5em}Dipole Illumination Model Correction\vspace{-1em}}
\author{Talon Chandler}% and Patrick La Rivi\`ere}
\date{\vspace{-1em}March 24, 2017\vspace{-1em}}
\maketitle

\section{Summary}
Our current model for the induced dipole moment created by illuminating a single
fluorophore with a monochromatic plane wave is
\begin{align}
  \bs{\mu}_{\text{ind}} \propto \hat{\bs{\mu}}_{\text{em}} \left[\hat{\bs{\mu}}_{\text{abs}}\cdot \mb{E}\right].
\end{align}

In these notes I will re-derive this model and correct several errors. The corrected
model is
\begin{align}
  \bs{\mu}_{\text{ind}} \propto \hat{\bs{\mu}}_{\text{em}} \left|\hat{\bs{\mu}}_{\text{abs}}^{\dagger}\mb{A}\right|
\end{align}
where $\mb{A}$ is the complex envelope (also known as the generalized Jones
vector).

\section{Math Preliminaries}
\subsection{Set Definitions}
\begin{align*}
  \mathbb{R}:&\ \text{the set of real numbers}\\
  \mathbb{R}^n:&\ \text{the set of $n$ dimensional vectors with real components}\\
  \mathbb{C}:&\ \text{the set of complex numbers}\\
  \mathbb{C}^n:&\ \text{the set of $n$ dimensional vectors with complex components}\\
  \mathbb{S}^n &= \left\{x\in \mathbb{R}^{n}: |x| = 1 \right\}:\ \text{the set of real unit vectors in $n$ dimensions.}\\
  \mathbb{T}^n &= \left\{x\in \mathbb{C}^{n}: |x| = 1 \right\}:\ \text{the set of complex unit vectors in $n$ dimensions.}\\
  [a, b) &= \left\{x\in \mathbb{R}: a\leq x < b\right\}:\ \text{shorthand for subsets of the real numbers. }
\end{align*}
\subsection{Useful Operations With Complex Vectors}
\label{math}
Consider $\mb{x}, \mb{y} \in \mathbb{C}^n$.
\begin{enumerate}
\item The conjugate transpose of $\mb{x}$ is denoted by a dagger ($\mb{x}^{\dagger}$).
\item $|\mb{x}|^2 = \mb{x}^{\dagger}\mb{x}$.
\item $\mb{x}^{\dagger}\mb{y} = \mb{x}\cdot \mb{y}$ if $\mb{x}, \mb{y}\in \mathbb{R}^n$
\item $(\mb{x}\mb{y}^{\dagger})^{\dagger} = \mb{y}\mb{x}^{\dagger}$
\end{enumerate}

\section{Representations of Polarized Plane Waves}
The time dependent electric field $\mathcal{E}$ of a polarized monochromatic
plane wave is given by
\begin{equation}
  \mathcal{E}(\mb{r},t) = E_{o,x}\cos(k_{x}r_{x} - \omega t)\mh{x}\ + \\
                     E_{o,y}\cos(k_{y}r_{y} - \omega t - \varphi_y)\mh{y}\ +\\
                     E_{o,z}\cos(k_{z}r_{z} - \omega t - \varphi_z)\mh{z}
                       \label{eq:plane}
\end{equation}
where $\mb{E}_{o} = [E_{o,x}, E_{o,y}, E_{o,z}]^T \in \mathbb{R}^3$ is the real
electric field amplitude, $\mb{k} = [k_x, k_y, k_z]^T \in \mathbb{R}^3$ is the
wave vector which points in the direction of propagation,
$\mb{r} = [r_x, r_y, r_z]^T \in \mathbb{R}^3$ is the position,
$\omega \in \mathbb{R}$ is the angular frequency, and $\varphi_y$ ($\varphi_z$)
$\in [0, 2\pi)$ is the phase shift of the y (z) component. For equation
\ref{eq:plane} to represent a physically realizable polarized plane wave, we
require that $\mb{E}\cdot \mb{k} = 0$ so that the plane wave is transverse, and
$|\mb{k}| = \frac{\omega}{c}$ so that the wave can propagate.

The complex spatial electric field $\mb{E}$ for a monochromatic polarized plane wave is given by
\begin{equation}
  \mb{E}(\mb{r}) = \mb{A}e^{j\mb{k}\cdot\mb{r}}
  \label{eq:factor}
\end{equation}
where $\mb{E} \in \mathbb{C}^3$, and $\mb{A} \in \mathbb{C}^3$ is the complex
envelope. We can recover the time dependent electric field from the complex spatial
electric field using
\begin{equation}
  \mathcal{E}(\mb{r},t) = \text{Re}\left\{\mb{E}(\mb{r})e^{-j\omega t}\right\}
\end{equation}

We can view equation \ref{eq:factor} as a factorization of $\mb{E}$ into a
spatially dependent part ($e^{j\mb{k}\cdot\mb{r}}$) and a spatially independent
part ($\mb{A}$). $\mb{A}$ is called the complex envelope \cite{saleh} or the
generalized Jones vector \cite{azzam}. The conventional Jones vector $\mb{J}$
only considers polarized plane waves traveling along the $z$ axis, so
$\mb{J}\in\mathbb{C}^2$ suffices to describe the polarization state. $\mb{A}$
describes the polarization state for plane waves traveling in any direction, so
$\mb{A} \in \mathbb{C}^3$ is required.

\section{Power Absorbed By A Dipole}
The power absorbed by a single dipole with absorption dipole moment
$\bs{\mu}_{\text{abs}}$ illuminated by a polarized plane wave is \cite{nov}
\begin{equation}
  P_{\text{abs}} \propto |\hat{\bs{\mu}}_{\text{abs}}^{\dagger} \mb{E}|^2
  \label{eq:power}
\end{equation}
where $\hat{\bs{\mu}}_{\text{abs}} \in \mathbb{T}^3$ is the absorption dipole
moment. Notice that we've allowed $\hat{\bs{\mu}}_{\text{abs}}$ to be a unit
complex vector instead of restricting it to be a unit real vector. Although
we'll focus on the case of $\hat{\bs{\mu}}_{\text{abs}} \in \mathbb{S}^3$
experimentally, allowing $\hat{\bs{\mu}}_{\text{abs}} \in \mathbb{T}^3$
simplifies the notation and allows us to consider complex transition moments
(transitions that are excited by circularly polarized light) if required.

\subsection{Absorbed Power Is Independent of Phase}
Equation \ref{eq:power} extends to generalized Jones vectors. Plugging
\ref{eq:factor} into \ref{eq:power} and using the operations in section \ref{math} gives
\begin{align}
  P_{\text{abs}} &\propto |\hat{\bs{\mu}}_{\text{abs}}^{\dagger} \mb{A}e^{j\mb{k}\cdot\mb{r}}|^2\\
  P_{\text{abs}} &\propto (\hat{\bs{\mu}}_{\text{abs}}^{\dagger} \mb{A}e^{j\mb{k}\cdot\mb{r}})^{\dagger}(\hat{\bs{\mu}}_{\text{abs}}^{\dagger} \mb{A}e^{j\mb{k}\cdot\mb{r}})\\
  P_{\text{abs}} &\propto (\hat{\bs{\mu}}_{\text{abs}}^{\dagger} \mb{A})^{\dagger}(\hat{\bs{\mu}}_{\text{abs}}^{\dagger} \mb{A})e^{j\mb{k}\cdot\mb{r}}e^{-j\mb{k}\cdot\mb{r}}\\
  P_{\text{abs}} &\propto |\hat{\bs{\mu}}_{\text{abs}}^{\dagger} \mb{A}|^2
  \label{eq:power2}                   
\end{align}
This confirms Rudolph and Shalin's claim that absorbed power is independent of
phase.

\subsection{Sum Over Source}
Finally, we show that if a dipole is illuminated by two plane waves with
independent phase (say by two lasers that are not phase locked, or by
illuminating two points on the back focal plane of a condenser with a thermal
source) then the total absorbed power is the sum of the absorbed powers
due to the individual plane waves. Let $\mb{E}_1$ and
$\mb{E}_2\text{exp}\{j\undertilde{\phi}\}$ be the complex spatial electric
fields created by the first and second illuminating plane waves where
$\undertilde{\phi} \sim U(0, 2\pi)$ is a uniformly distributed phase difference
between the two plane waves. The expected value of the absorbed power is
\begin{align}
  E[\undertilde{P}{\tiny{\text{abs}}}] &\propto E\left[|\hat{\bs{\mu}}_{\text{abs}}^{\dagger} (\mb{E}_1 + \mb{E}_2\text{exp}\{j\undertilde{\phi}\})|^2\right]\\
  E[\undertilde{P}{\tiny{\text{abs}}}] &\propto E\left[\left[\hat{\bs{\mu}}_{\text{abs}}^{\dagger} (\mb{E}_1 + \mb{E}_2\text{exp}\{j\undertilde{\phi}\})\right]^{\dagger}\left[\hat{\bs{\mu}}_{\text{abs}}^{\dagger} (\mb{E}_1 + \mb{E}_2\text{exp}\{j\undertilde{\phi}\})\right]\right]\\
  E[\undertilde{P}{\tiny{\text{abs}}}] &\propto E\left[\left[\hat{\bs{\mu}}_{\text{abs}}^{\dagger} (\mb{E}_1 + \mb{E}_2\text{exp}\{j\undertilde{\phi}\})(\mb{E}_1^{\dagger} + \mb{E}_2^{\dagger}\text{exp}\{-j\undertilde{\phi}\})\hat{\bs{\mu}}_{\text{abs}}\right]^{\dagger}\right]\\
  E[\undertilde{P}{\tiny{\text{abs}}}] &\propto E\left[\left[\hat{\bs{\mu}}_{\text{abs}}^{\dagger} (\mb{E}_1\mb{E}_1^{\dagger} + \mb{E}_2\mb{E}_1^{\dagger}\text{exp}\{j\undertilde{\phi}\} + \mb{E}_1^{\dagger}\mb{E}_2\text{exp}\{-j\undertilde{\phi}\} + \mb{E}_2\mb{E}_2^{\dagger})\hat{\bs{\mu}}_{\text{abs}}\right]^{\dagger}\right]
\intertext{Pulling the expectation inside and using $E[\text{exp}\{j\undertilde{\phi}\}] = 0$ gives}
  E[\undertilde{P}{\tiny{\text{abs}}}] &\propto \left[\hat{\bs{\mu}}_{\text{abs}}^{\dagger} (\mb{E}_1\mb{E}_1^{\dagger} + \mb{E}_2\mb{E}_2^{\dagger})\hat{\bs{\mu}}_{\text{abs}}\right]^{\dagger}\\
  E[\undertilde{P}{\tiny{\text{abs}}}] &\propto \left[\hat{\bs{\mu}}_{\text{abs}}^{\dagger}\mb{E}_1\mb{E}_1^{\dagger}\hat{\bs{\mu}}_{\text{abs}} + \hat{\bs{\mu}}_{\text{abs}}^{\dagger}\mb{E}_2\mb{E}_2^{\dagger}\hat{\bs{\mu}}_{\text{abs}}\right]^{\dagger}\\
  E[\undertilde{P}{\tiny{\text{abs}}}] &\propto |\hat{\bs{\mu}}_{\text{abs}}^{\dagger}\mb{E}_1|^2 + |\hat{\bs{\mu}}_{\text{abs}}^{\dagger}\mb{E}_2|^2\\
  E[\undertilde{P}{\tiny{\text{abs}}}] &\propto P_{\text{abs},1} + P_{\text{abs},2}
\end{align}
This confirms Rudolph and Shalin's claim that independent plane waves excite a
fluorophore independently. We can extend the argument to an arbitrary number of
independent plane waves which allows us to integrate over all plane waves
incident on the fluorophore to find the total absorbed power.

\section{Dipole Emission Pattern}
The emitted electric field pattern is proportional to the Green's
tensor multiplied by the emission dipole moment of the fluorophore
\begin{align}
  \mb{E}_{\text{em}}(\mb{r}) \propto \mb{G}\hat{\bs{\mu}}_{\text{em}}.\label{eq:eight}
\end{align}
where $\hat{\bs{\mu}}_{\text{em}} \in \mathbb{T}^3$.
The power emitted by a fluorophore is proportional to the power absorbed
\begin{align}
  P_{\text{em}} &\propto P_{\text{abs}}. 
\end{align}
Using equation $P_{\text{em}} \propto |\mb{E}_{\text{em}}(\mb{r})|^2$ and \ref{eq:power2} gives
\begin{align}
  |\mb{E}_{\text{em}}(\mb{r})|^2 &\propto |\hat{\bs{\mu}}_{\text{abs}}^{\dagger}\mb{A}|^2.\label{eq:twenty}
\end{align}
Taking the square root of \ref{eq:twenty} and combining with \ref{eq:eight} gives the final result
\begin{align}
  \mb{E}_{\text{em}}(\mb{r}) \propto \mb{G}\hat{\bs{\mu}}_{\text{em}}|\hat{\bs{\mu}}^{\dagger}_{\text{abs}}\mb{A}|. \label{eq:final}
\end{align}

\section{Induced Dipole Moment}
We can split equation \ref{eq:final} into two parts using the induced dipole
moment
\begin{align}
  \bs{\mu}_{\text{ind}} \equiv \hat{\bs{\mu}}_{\text{em}}|\hat{\bs{\mu}}^{\dagger}_{\text{abs}}\mb{A}|\\
  \mb{E}_{\text{em}}(\mb{r}) \propto \mb{G}\bs{\mu}_{\text{ind}}
\end{align}
where $\bs{\mu}_{\text{ind}} \in \mathbb{C}^3$.

\section{K\"{o}hler Illumination}
The work in these notes extends to the previous note set. The induced dipole
moment created by K\"{o}hler illumination is given by
\begin{align}
  \bs{\mu}_{\text{ind}} \propto
  \bs{\mu}_{\text{em}} \int d\mb{r'}
  \left|\bs{\mu}_{\text{abs}}^{\dagger} \mb{R}(\mb{r'})\mb{A}_{\text{bfp}}\right| \label{eq:test2}
\end{align}
Shalin suggested that we separate the model into three parts by moving
$\bs{\mu}_{\text{em}}$ and $\bs{\mu}_{\text{abs}}$ outside of the integral. This
would allow us to calculate the ``effective electric field'' or ``polarization resolved
electric field''. The addition of the absolute value bars prevent us from making this simplification.

With equation \ref{eq:test2} in view, we can see why the absolute value bars are
required. Without the absolute value bars,
$\bs{\mu}_{\text{abs}}^{\dagger}(\mb{r})
\mb{R}_{\mb{\hat{s}}}(\mb{r'})\mb{A}_{\text{bfp}} \in [-1, 1]$, so the integral
can evaluate to zero. With the absolute value bars,
$|\bs{\mu}_{\text{abs}}^{\dagger}(\mb{r})
\mb{R}_{\mb{\hat{s}}}(\mb{r'})\mb{A}_{\text{bfp}}| \in [0, 1]$ and each point in
the back focal plane contributes to the induced dipole moment.

\section{Total Power Emitted By A Dipole}
\figs{../figures/scene1}{../figures/intensity1}{Scene schematic. A single dipole (see arrow in the
  front focal plane) under $x$-polarized K\"{o}hler illumination (see arrow in
  the back focal plane).\\}{Total power emitted by a single dipole
  ($|\bs{\mu}_{\text{ind}}|^2$) as a function of direction under the illumination
  geometry in Figure \ref{fig:scene}. Dots indicate where the Cartesian axes
  intersect the sphere. Arbitrary units.}{scene}{int}{-0.2}

Figure \ref{fig:scene} shows a simple K\"{o}hler illumination geometry where
$x$-polarized light illuminates a dipole with the optical axis along $z$. Figure
\ref{fig:int} shows the total power emitted by the dipole as a function of its
direction. In practice, we will only collect a fraction of the total power
emitted, but Figure \ref{fig:int} shows the total power---as if we had
detectors on all sides of the dipole. We can see that dipoles
aligned along the $x$ axis will emit (and absorb) the most power, while dipoles
aligned along the $y$ or $z$ axes will emit very little power.

Figure \ref{fig:illum} shows the total power emitted by a single dipole under
different back aperture radii, $\rho$, with constant illumination power. Our
model assumes that the lens is large enough to map each point in the back focal
plane to a plane wave in the front focal plane, so the only parameter of
interest is $\frac{\rho}{f}$, the ratio of the back aperture radius to the focal
length of the lens.

Figure \ref{fig:illum} also shows how the results change when we neglect (second
row) and keep (third row) the absolute value bars inside the integral in
equation \ref{eq:test2}. When we neglect the absolute value bars, the model
predicts that as we open the back aperture the emitted power pattern is not
mirror symmetric about the $x-y$ plane. When we keep the absolute value bars,
the absorbed power pattern is symmetric about the $x-y$ plane. Also, notice that
$+z$ and $-z$ aligned dipoles emit more power as we open the back aperture (you
may have to zoom in on the bottom row to see the difference---the $+z$ axis in
the bottom right plot is $\sim 300$ compared to $\sim 0$ in the bottom left
plot).

\fig{../figures/compare_illum}{1.0}{Top row---schematics of the illumination
  geometries. Middle row---\textbf{incorrect} emitted power pattern if the absolute value bars are neglected. Bottom row---\textbf{correct} emitted power pattern if the absolute value bars are retained. Columns---the back aperture radius increases from left to right.}{illum}

\section{Aside: Generalized Jones Vectors}
The generalized Jones vector in equation \ref{eq:factor} is going to be extremely
useful for analyzing light fields traveling through crystals. Ortega
\cite{ortega} and Azzam \cite{azzam} have started developing the generalized
Jones calculus, but they haven't applied the results to stacks of
crystals. Foreman and Torok \cite{foreman} have used generalized Jones vectors
to analyze high-NA imaging systems, but they haven't applied their results to
transmission microscopes like the LC-Polscope. More progress on this front soon!

\bibliography{report}{}
\bibliographystyle{unsrt}

\end{document}

### 2017-06-12 notes/2017-04-25-fourkas-correction/report/report.tex
Add notes history.

\documentclass[11pt]{article}

%%%%%%%%%%%%
% Packages %
%%%%%%%%%%%%
\hyphenpenalty=10000
\usepackage{tocloft}
\renewcommand\cftsecleader{\cftdotfill{\cftdotsep}}
\def\undertilde#1{\mathord{\vtop{\ialign{##\crcr
$\hfil\displaystyle{#1}\hfil$\crcr\noalign{\kern1.5pt\nointerlineskip}
$\hfil\tilde{}\hfil$\crcr\noalign{\kern1.5pt}}}}}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{epstopdf}
\usepackage{braket}
\usepackage{upgreek}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{amssymb,latexsym,amsmath,gensymb}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{float}
\usepackage{enumitem}
\usepackage{pdflscape}
\usepackage{url}
\usepackage{tikz, calc}
\usetikzlibrary{shapes.geometric, arrows, calc}
\tikzstyle{norm} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{arrow} = [thick, ->, >=stealth]

\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
\providecommand{\mb}[1]{\mathbf{#1}}
\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
\providecommand{\bs}[1]{\boldsymbol{#1}} 
\providecommand{\intinf}{\int_{-\infty}^{\infty}}
\providecommand{\fig}[4]{
  % filename, width, caption, label
\begin{figure}[h]
 \captionsetup{width=1.0\linewidth}
 \centering
 \includegraphics[width = #2\textwidth]{#1}
 \caption{#3}
 \label{fig:#4}
\end{figure}
}

\newcommand{\tensor}[1]{\overset{\text{\tiny$\leftrightarrow$}}{\mb{#1}}}
\newcommand{\tunderbrace}[2]{\underbrace{#1}_{\textstyle#2}}
\providecommand{\figs}[7]{
  % filename1, filename2, caption1, caption2, label1, label2, shift
\begin{figure}[H]
\centering
\begin{minipage}[b]{.45\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#1}
  \captionsetup{justification=justified, singlelinecheck=true}
  \caption{#3}
  \label{fig:#5}
\end{minipage}
\hspace{2em}
\begin{minipage}[b]{.45\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#2}
  \vspace{#7em}
  \captionsetup{justification=justified}
  \caption{#4}
  \label{fig:#6}
\end{minipage}
\end{figure}
}
\makeatletter

\providecommand{\code}[1]{
\begin{center}
\lstinputlisting{#1}
\end{center}
}

%%%%%%%%%%%
% Spacing %
%%%%%%%%%%%
% Margins
\usepackage[
top    = 1.5cm,
bottom = 1.5cm,
left   = 1.5cm,
right  = 1.5cm]{geometry}

% Indents, paragraph space
\usepackage{parskip}

% Section spacing
\usepackage{titlesec}
\titlespacing*{\title}
{0pt}{0ex}{0ex}
\titlespacing*{\section}
{0pt}{0ex}{0ex}
\titlespacing*{\subsection}
{0pt}{0ex}{0ex}
\titlespacing*{\subsubsection}
{0pt}{0ex}{0ex}

% Line spacing
\linespread{1.1}

%%%%%%%%%%%%
% Document %
%%%%%%%%%%%%
\begin{document}
\title{\vspace{-2.5em} Inconsistency In ``Rapid determination of the three-dimensional orientation of single molecules''\vspace{-1em}}
\author{Talon Chandler}% and Patrick La Rivi\`ere}
\date{\vspace{-1em}April 25, 2017\vspace{-1em}}
\maketitle

\section{Introduction}
There is an inconsistency in John T. Fourkas' paper titled ``Rapid determination
of the three dimensional orientation of single molecules''\cite{fourkas}. In
section 2 I describe the inconsistency. In section 3 I follow Fourkas'
calculation in detail and correct it. In section 4 I show the results of the
corrected calculation and show that the new results are consistent. Finally, in
section 5 I discuss how these notes will affect other papers and the
applicability of these results.

\section{Inconsistency}
Equations 4 and 5 of \cite{fourkas} model the fraction of the total intensity
from a single fluorophore collected by an objective with a polarizer in the back
focal plane as a function of polarizer orientation, dipole orientation
($\Theta$, $\Phi$), and half angle of the light collection cone
($\alpha = \sin^{-1}(\text{NA}/n)$). I expect that as $\alpha$ approaches
$\frac{\pi}{2}$ (or a NA approaches $n$) the fraction of the total power
radiated by the dipole that is collected by the lens should approach
$\frac{1}{2}$
\begin{align}
  \lim_{\alpha\rightarrow \frac{\pi}{2}} \frac{I_{0}(\Theta, \Phi, \alpha) + I_{90}(\Theta, \Phi, \alpha)}{I_{\text{tot}}(\Theta, \Phi)} = \frac{1}{2}. \label{eq:prop}
\end{align}

First, I reproduce equations 4 and 5 from \cite{fourkas}
\begin{subequations}
\begin{align}
    I_{0}(\Theta, \Phi, \alpha) &= I_{\text{tot}}(t, t+\tau)(A + B\sin^{2}{\Theta} + C\sin^{2}{\Theta} \cos{2 \Phi})\\
    I_{45}(\Theta, \Phi, \alpha) &= I_{\text{tot}}(t, t+\tau)(A + B\sin^{2}{\Theta} + C\sin^{2}{\Theta}\sin{2 \Phi})\\
    I_{90}(\Theta, \Phi, \alpha) &= I_{\text{tot}}(t, t+\tau)(A + B\sin^{2}{\Theta} - C\sin^{2}{\Theta} \cos{2 \Phi})\\
  I_{135}(\Theta, \Phi, \alpha) &= I_{\text{tot}}(t, t+\tau)(A + B\sin^{2}{\Theta} - C\sin^{2}{\Theta} \sin{2 \Phi}),
\end{align}\label{eq:int1}
\end{subequations}
and
\vspace{-1em}
\begin{subequations}
\begin{align}
  A &= \frac{1}{6} - \frac{1}{4}\cos\alpha + \frac{1}{12}\cos^3\alpha\\
  B &= \frac{1}{8}\cos\alpha - \frac{1}{8}\cos^3\alpha\\
  C &= \frac{7}{48} - \frac{1}{16} \cos{\alpha } - \frac{1}{16} \cos^{2}{\alpha } - \frac{1}{48} \cos^{3}{\alpha }.
\end{align}\label{eq:coeff1}
\end{subequations}
Next, I take the limit proposed in equation \ref{eq:prop}
\begin{align}
  \lim_{\alpha\rightarrow \frac{\pi}{2}} \frac{I_{0}(\Theta, \Phi, \alpha) + I_{90}(\Theta, \Phi, \alpha)}{I_{\text{tot}}(\Theta, \Phi)} &= \lim_{\alpha\rightarrow \frac{\pi}{2}} 2\left(A(\alpha) + B(\alpha)\sin^2\Theta\right) =2\left(\frac{1}{6} + 0\right) = \frac{1}{3} \neq \frac{1}{2}.
\end{align}
Fourkas' model predicts that only $\frac{1}{3}$ of the total power radiated by the
fluorophore is collected by an objective with a collection half angle of $\frac{\pi}{2}$.

\section{Corrected Calculation}
I start by following Fourkas and define the $z$ axis as the optical axis of the
objective. I express a directional unit vector $\hat{\mb{r}}$ and the
fluorescence emission dipole $\hat{\bs{\mu}}_{\text{em}}$ in spherical
coordinates as follows
\begin{align}
  \hat{\mb{r}} &= \sin\theta\cos\phi\hat{\mb{i}} + \sin\theta\sin\phi\hat{\mb{j}} + \cos\theta\hat{\mb{k}}\label{eq:r_coords}\\
  \hat{\bs{\mu}}_{\text{em}} &= \sin\Theta\cos\Phi\hat{\mb{i}} + \sin\Theta\sin\Phi\hat{\mb{j}} + \cos\Theta\hat{\mb{k}}.\label{eq:mu_coords}
\end{align}
Next, I calculate the generalized Jones vector $\mb{A}$ (GJV or the complex
envelope) along a direction $\hat{\mb{r}}$ in the far field
\begin{align}
\mb{\hat{A}}_{\text{ff}}(\hat{\mb{r}}) &= \mh{r}\times\hat{\bs{\mu}}_{\text{em}}\times\mh{r} \label{eq:fourkasa}\\ &= (\tilde{\mb{I}} - \hat{\mb{r}}\hat{\mb{r}}^{\dagger})\hat{\bs{\mu}}_{\text{em}} \label{eq:greena} 
\end{align}
where $\mb{I}$ is the identity matrix and ${}^{\dagger}$ is the adjoint
operator. Notice that $(\tilde{\mb{I}} - \hat{\mb{r}}\hat{\mb{r}}^{\dagger})$ is
the Green's tensor with the phase excluded. Also notice that $\mb{\hat{A}}_{\text{ff}}(\mh{r})$ is
not the electric field---it is the GJV because I have stopped keeping track of
the phase. I will follow Fourkas and use equation \ref{eq:fourkasa}, but I have
written equation \ref{eq:greena} to relate this work to future work with the
Green's tensor.

Next, I model the action of an ideal, infinity corrected, polarization
preserving objective on the GJV. The lens rotates the GJV so that it is
perpendicular to the $z$ axis at every point on the unit sphere---we can model this by rotating the ray by $-\phi$ about the $z$ axis, by $-\theta$ about the $y$ axis, then by $\phi$ about $z$ axis. In matrix form the position-dependent rotation matrix is
\begin{align}  
  \tilde{\mb{R}}(\mh{r}) = \begin{bmatrix} \cos\theta\cos^2\phi + \sin^2\phi & (\cos\theta -1)\sin\phi\cos\phi & -\sin\theta\cos\phi\\ (\cos\theta - 1)\sin\phi\cos\phi & \cos\theta\sin^2\phi + \cos^2\phi & -\sin\theta\sin\phi \\ \sin\theta\cos\phi& \sin\theta\sin\phi & \cos\theta \end{bmatrix},\label{eq:matrix}
\end{align}
which is identical to the equation 2 in \cite{fourkas}.

Next, I express the transmission axis of the polarizer $\mh{P}$ using
\begin{align}
  \hat{\mb{P}} = \cos\phi_{\text{pol}}\hat{\mb{i}} + \sin\phi_{\text{pol}}\hat{\mb{j}}. \label{eq:polarizer}
\end{align}
I model the action of the polarizer in the back focal plane by taking the
dot product of the generalized Jones vector with the polarizer axis $\mh{P}$ and multiplying by the polarizer axis direction. The GJV in the back focal plane of the objective is given by
\begin{align}
  \mb{A}_{\text{bfp}}(\mh{r}) = \mh{P}\left[\mh{P}\cdot\tilde{\mb{R}}(\mh{r})(\mh{r}\times\hat{\bs{\mu}}_{\text{em}}\times\mh{r})\right].\label{eq:approx}
\end{align}
Notice that I am following Fourkas and mapping the GJV directly to the back focal plane
without considering the change of coordinates. I will discuss this approximation in section \ref{discussion}.

Next, I model the detection process by squaring the GJV and integrating over
the cap of the sphere collected by the objective, $\Omega$. The complete model
of the detected intensity is
\begin{align}
  I_{\phi_{\text{pol}}}(\Theta, \Phi) \propto \int_{\Omega}d\mh{r}\ \left|\mh{P}\left[\mh{P}\cdot\tilde{\mb{R}}(\mh{r})(\mh{r}\times\hat{\bs{\mu}}_{\text{em}}\times\mh{r})\right]\right|^2. \label{eq:model1}
\end{align}
Next, I calculate the total power emitted by a unit dipole
\begin{align}
  I_{\text{tot}} = \int_{\mathbb{S}^2}d\mh{r}\ \left|\mh{r}\times\hat{\bs{\mu}}_{\text{em}}\times\mh{r}\right|^2 = \int_0^{2\pi}d\phi\int_0^{\pi}d\theta\sin^3\theta = \frac{8\pi}{3}
\end{align}
Finally, I calculate the fraction of the total power detected
\begin{align}
  \frac{I_{\phi_{\text{pol}}}(\Theta, \Phi)}{I_{\text{tot}}} = \frac{3}{8\pi}\int_{\Omega}d\mh{r}\ \left|\mh{P}\left[\mh{P}\cdot\tilde{\mb{R}}(\mh{r})(\mh{r}\times\hat{\bs{\mu}}_{\text{em}}\times\mh{r})\right]\right|^2\label{eq:final}
\end{align}
where the total radiated power $I_{\text{tot}}$ was found using equation \ref{eq:model1} with no polarizer or objective and by integrating over all directions.

Notice that I normalized using the total radiated intensity. Fourkas incorrectly normalized the GJV before calculating the intensity as follows 
\begin{align}
  \frac{I_{\phi_{\text{pol}}}(\Theta, \Phi)}{I_{\text{tot}}} &\neq \int_{\Omega}d\mh{r}\left|\left(\frac{\mb{A}_{\phi_{\text{pol}}}}{\mb{A}_{\text{tot}}}\right)\right|^2\label{eq:incorrect}\\
  \intertext{It's not physically clear what $\mb{A}_{\text{tot}}$ represents, but I can make a good guess about how Fourkas calculated it}
  \mb{A}_{\text{tot}} &= \int_{\mathbb{S}^2} d\mh{r} (\mh{r}\times\hat{\bs{\mu}}_{\text{em}}\times\mh{r}) = \int_0^{2\pi}d\phi\int_0^{\pi}d\theta\sin\theta = 4\pi\label{eq:atot}\\
  \intertext{Plugging \ref{eq:atot} into \ref{eq:incorrect} gives}
  \frac{I_{\phi_{\text{pol}}}(\Theta, \Phi)}{I_{\text{tot}}} &\neq \frac{1}{16\pi}\int_{\Omega}d\mh{r}\ \left|\mh{P}\left[\mh{P}\cdot\tilde{\mb{R}}(\mh{r})(\mh{r}\times\hat{\bs{\mu}}_{\text{em}}\times\mh{r})\right]\right|^2\label{eq:four_final}
\end{align}
Notice that equation \ref{eq:four_final} and equation \ref{eq:final} differ by a factor of $\frac{3}{2}$.

\section{Results}
I evaluated equations \ref{eq:final} and \ref{eq:four_final} by substituting equations \ref{eq:r_coords}, \ref{eq:mu_coords}, \ref{eq:matrix}, and \ref{eq:polarizer} and simplifying using SymPy, an open-source computer algebra system. Equation \ref{eq:four_final} yields Fourkas' published expressions. Equation \ref{eq:final} yields Fourkas' expressions multiplied by a factor of $\frac{3}{2}$. We can make this correction by changing the
coefficients of $A$, $B$, and $C$ to 
\begin{subequations}
\begin{align}
  A_{\text{new}} &= \frac{1}{4} - \frac{3}{8} \cos{\alpha } + \frac{1}{8} \cos^{3}{\alpha }\\
  B_{\text{new}} &= \frac{3}{16} \cos{\alpha } - \frac{3}{16} \cos^{3}{\alpha }\\
  C_{\text{new}} &= \frac{7}{32} - \frac{3}{32} \cos{\alpha } - \frac{3}{32} \cos^{2}{\alpha } - \frac{1}{32} \cos^{3}{\alpha}.
\end{align}\label{eq:coeff2}
\end{subequations}

Fourkas' published inversion expressions still hold if we use $A_{\text{new}}$,
$B_{\text{new}}$, and $C_{\text{new}}$.

Finally, I confirm that taking the limit with the new expressions approaches $\frac{1}{2}$
\begin{align}
  \lim_{\alpha\rightarrow \frac{\pi}{2}} \frac{I_{0}(\Theta, \Phi) + I_{90}(\Theta, \Phi)}{I_{\text{tot}}(\Theta, \Phi)} &= \lim_{\alpha\rightarrow \frac{\pi}{2}} 2\left(A_{\text{new}} + B_{\text{new}}\sin^2\Theta\right) =2\left(\frac{1}{4} + 0\right) = \frac{1}{2}  
\end{align}

\section{Discussion} \label{discussion} Fourkas' expressions underpredict the
fraction of the total power collected by an objective with a polarizer by a
constant fraction of $\frac{3}{2}$. Therefore, using Fourkas' expressions to
find $I_{\text{tot}}$ will result in overpredictions of $I_{\text{tot}}$, but
the incorrect expressions will not cause any error in the predictions of molecular
orientation.

Lu and Bout \cite{bout2008} study the effect of noise on the reconstruction of
single fluorophores using Fourkas' expressions. Their main results are
unaffected by the correction introduced in these notes, but because these notes
predict a higher intensity than Fourkas, Lu and Bout overpredict the effect of
noise in their model.

As mentioned after equation \ref{eq:approx}, Fourkas uses an approximation when
he directly maps the GJV in the back focal plane to the GJV in the detector
plane. To conserve energy, the power carried by a ray that makes an angle
$\theta$ with the optical axis needs to be attenuated by a factor of
$\sqrt{\cos\theta}$ \cite{nov, backer}. This attenuation factor can be ignored
under the paraxial approximation ($\cos\theta \approx 1$), which means that all
of the results in Fourkas and the correction in these notes only apply to low NA
lenses. Analytically extending these results to high NA lenses could be the
subject of future work.

As noted in \cite{lieb}, Fourkas' model only considers single molecules in
homogeneous environments. With an extension of the Green's tensor (equation
\ref{eq:greena}) \cite{nov, backer}, we could consider single molecules near
glass interfaces.

\bibliography{report}{}
\bibliographystyle{unsrt}

\end{document}

### 2017-06-12 notes/2017-05-02-illumination-update3/report/report.tex
Add notes history.

\documentclass[11pt]{article}

%%%%%%%%%%%%
% Packages %
%%%%%%%%%%%%
\hyphenpenalty=10000
\usepackage{tocloft}
\renewcommand\cftsecleader{\cftdotfill{\cftdotsep}}
\def\undertilde#1{\mathord{\vtop{\ialign{##\crcr
$\hfil\displaystyle{#1}\hfil$\crcr\noalign{\kern1.5pt\nointerlineskip}
$\hfil\tilde{}\hfil$\crcr\noalign{\kern1.5pt}}}}}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{epstopdf}
\usepackage{braket}
\usepackage{upgreek}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{amssymb,latexsym,amsmath,gensymb}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{float}
\usepackage{enumitem}
\usepackage{pdflscape}
\usepackage{url}
\usepackage{tikz, calc}
\usetikzlibrary{shapes.geometric, arrows, calc}
\tikzstyle{norm} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{arrow} = [thick, ->, >=stealth]

\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
\providecommand{\mb}[1]{\mathbf{#1}}
\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
\providecommand{\bs}[1]{\boldsymbol{#1}} 
\providecommand{\intinf}{\int_{-\infty}^{\infty}}
\providecommand{\fig}[4]{
  % filename, width, caption, label
\begin{figure}[h]
 \captionsetup{width=1.0\linewidth}
 \centering
 \includegraphics[width = #2\textwidth]{#1}
 \caption{#3}
 \label{fig:#4}
\end{figure}
}

\newcommand{\tensor}[1]{\overset{\text{\tiny$\leftrightarrow$}}{\mb{#1}}}
\newcommand{\tunderbrace}[2]{\underbrace{#1}_{\textstyle#2}}
\providecommand{\figs}[7]{
  % filename1, filename2, caption1, caption2, label1, label2, shift
\begin{figure}[H]
\centering
\begin{minipage}[b]{.45\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#1}
  \captionsetup{justification=justified, singlelinecheck=true}
  \caption{#3}
  \label{fig:#5}
\end{minipage}
\hspace{2em}
\begin{minipage}[b]{.45\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#2}
  \vspace{#7em}
  \captionsetup{justification=justified}
  \caption{#4}
  \label{fig:#6}
\end{minipage}
\end{figure}
}
\makeatletter

\providecommand{\code}[1]{
\begin{center}
\lstinputlisting{#1}
\end{center}
}

%%%%%%%%%%%
% Spacing %
%%%%%%%%%%%
% Margins
\usepackage[
top    = 1.5cm,
bottom = 1.5cm,
left   = 1.5cm,
right  = 1.5cm]{geometry}

% Indents, paragraph space
\usepackage{parskip}

% Section spacing
\usepackage{titlesec}
\titlespacing*{\title}
{0pt}{0ex}{0ex}
\titlespacing*{\section}
{0pt}{0ex}{0ex}
\titlespacing*{\subsection}
{0pt}{0ex}{0ex}
\titlespacing*{\subsubsection}
{0pt}{0ex}{0ex}

% Line spacing
\linespread{1.1}

%%%%%%%%%%%%
% Document %
%%%%%%%%%%%%
\begin{document}
\title{\vspace{-2.5em}Estimating A Single Molecule's Orientation With Polarized
  Illumination Microscopy\vspace{-2em}} \author{}% and Patrick La Rivi\`ere}
\date{\vspace{-1em}May 2, 2017\vspace{-1em}}
\maketitle

\section{Introduction}
In this work we evaluate wide-field polarized microscope designs for determining
the orientation of single molecules. In section 2 we model the relationship
between a molecule's orientation, the microscope geometry, and the measured
intensity. In section 3 we present simulation results to show how these
microscopes behave, and we compare the ability of several microscope designs to
reconstruct the orientation of single dipoles using techniques from estimation
theory. Finally, in sections 4 and 5 we summarize our results and present plans
for future work.

\section{Methods}
In this section we model the relationship between a single molecule's
orientation, the microscope's geometry, and the measured intensity. We split the
model into two parts. First, we develop the illumination model and calculate the
excitation efficiency, $\eta_{\text{exc}}$---the fraction of the maximum power
absorbed by the molecule. Next, we develop the detection model and calculate the
detection efficiency, $\eta_{\text{det}}$---the fraction of the power emitted by
the molecule that we detect. Finally, we multiply the excitation efficiency and
the detection efficiency to find the total efficiency, $\eta_{\text{tot}}$---a
quantity that is proportional to the measured intensity.

\subsection{Illumination Model}
In this section we model the excitation efficiency of a single molecule with absorption
dipole moment $\hat{\bs{\mu}}_{\text{abs}}$ at the focal point of a
condenser. We excite the molecule with K\"{o}hler illumination with a linear
polarizer in the back focal plane of the condenser. Each point in the back focal
plane illuminates the molecule with a polarized plane wave rotated
by the condenser, and each polarized plane wave acts on the molecule independently, so
we can find the excitation efficiency by integrating over the back focal plane of the condenser. Therefore, the excitation efficiency is
\begin{align}
  \eta_{\text{exc}} &= \frac{\int_{\text{bfp}} d\mathbf{r}'|\hat{\bs{\mu}}_{\text{abs}}^{\dagger}\mathbf{\hat{A}}_{\text{ffp}}|^2}{\int_{\text{bfp}}d\mathbf{r}'}\label{eq:illum}\\
\mathbf{\hat{A}}_{\text{ffp}}(\mh{r}') &= \tilde{\mb{R}}(\mh{r}')\mathbf{A}_{\text{bfp}} 
\end{align}
where $\mh{A}_{\text{bfp}}$ is the direction of the polarizer in the back focal
plane, $\mh{r}'$ is the position coordinate in the back focal plane,
$\tilde{\mb{R}}(\mh{r}')$ is the position-dependent rotation matrix that rotates
the generalized Jones vector (GJV) in the back focal to a GJV in the front focal
plane, $\mh{A}_{\text{ffp}}$ is the GJV in the front focal plane due to the
point $\mh{r}'$ in the back focal plane, $\hat{\bs{\mu}}_{\text{abs}}$ is the
molecule's absorption dipole moment, and ${}^{\dagger}$ denotes the adjoint
operator.

$\eta_{\text{exc}}$ is normalized by the area of the back focal plane so the it
can be interpreted as the fraction of the maximum power absorbed by the
molecule. If the back focal plane is a pinhole then the molecule is illuminated
by a single plane wave. If the plane wave is aligned with the dipole
($\mathbf{\hat{A}}_{\text{bfp}} = \hat{\bs{\mu}}_{\text{abs}}$) then
$\eta_{\text{exc}}=1$, the most power the dipole can absorb. Opening the
aperture or rotating the molecule will decrease the excitation efficiency.

Equation \ref{eq:illum} is too computationally expensive to calculate for each
illumination and dipole orientation. Instead, we follow \cite{backer} and factor
the integral so that we can calculate an illumination basis once for each
illumination geometry. The factored form of equation \ref{eq:illum} is
\begin{align}
  \eta_{\text{exc}} &= \frac{1}{\int_{\text{bfp}}d\mathbf{r}'}[\mu_x^2, \mu_y^2, \mu_z^2, \mu_x\mu_y, \mu_x\mu_z, \mu_y\mu_z]\cdot\int_{\text{bfp}} d\mathbf{r}'[|A_x|^2, |A_y|^2, |A_z|^2, 2\text{Re}\{A_x^*A_y\}, 2\text{Re}\{A_x^*A_z\}, 2\text{Re}\{A_y^*A_z\}]^T \label{eq:factored}
\end{align}
where
\begin{align}
  \mb{\hat{A}}_{\text{ffp}} &= A_x\mh{i} + A_y\mh{j} + A_z\mh{k}\\
  \hat{\bs{\mu}}_{\text{abs}} &= \mu_x\mh{i} + \mu_y\mh{j} + \mu_z\mh{k}
\end{align}
We plot the illumination basis elements in figure \ref{fig:illum_basis}.

\fig{../figures/illum_basis.png}{1.0}{Illumination basis elements as a function of
  polarization orientation and back focal plane radius. Row 1: schematics of the
  illumination geometries---a polarizer is placed in the back focal plane of the
  condenser lens. Row 2 and 3: plots of the illumination basis elements as a
  function of $\rho/f$---the ratio of the back focal plane radius to the focal
  length. The illumination basis elements should be interpreted with
  equation \ref{eq:factored} in view---they do not have a direct physical
  interpretation. Columns: varying polarization orientations. }{illum_basis}

\subsection{Detection Model}
In this section we model the detection efficiency---the fraction of the power
emitted by a single molecule that we detect. We follow Fourkas \cite{fourkas} and find that the
detection efficiency is
\begin{align}
  \eta_{\text{det}} &= 2\left(A + B\sin^2\theta\right)\\
  A &= \frac{1}{4} - \frac{3}{8} \cos\alpha + \frac{1}{8} \cos^{3}\alpha\\
  B &= \frac{3}{16} \cos\alpha - \frac{3}{16} \cos^{3}\alpha
\end{align}
where $\theta$ is the angle between the molecule's emission dipole moment
$\hat{\bs{\mu}}_{\text{em}}$ and the detection arm's optical axis, and $\alpha$
is the half angle of the collection cone. See the note set ``Inconsistency in
rapid determination of the three-dimensional orientation of single molecules''
for the derivation of this model. Notice that we're using the corrected versions
of Fourkas' variables $A$ and $B$.

For this work we assume that $\hat{\bs{\mu}}_{\text{abs}} = \hat{\bs{\mu}}_{\text{em}}$.

\section{Results}
\subsection{Excitation, Detection, and Total Efficiency}
Figure \ref{fig:efficiency} shows representative results of the excitation,
detection, and total efficiencies as a function of dipole direction and
microscope geometry. Notice that a single frame does not give us enough
information to recover a dipole's direction. In the next section we will
consider using the information from combinations of single frame microscopes
like the ones shown in Figure \ref{fig:efficiency}.

\fig{../figures/geometry.png}{1.0}{Representative examples of single frame microscopes. Row
  1: schematics of the illumination and detection geometries. Rows 2-4: excitation, detection, and total efficiency as a function of dipole direction. The total efficiency is the product of the excitation efficiency and the detection efficiency.}{efficiency}

\subsection{Dipole Orientation Estimates With Multiple Frame Experiments}
In this section we combine the results from several single frame microscopes to
create multiple frame experiments, and we investigate the possibility of using
the information from multiple frames to reconstruct a single molecule's
orientation. To assess our ability to reconstruct a molecule's orientation from
intensity measurements, we parameterize the molecule's orientation with
spherical coordinates
($\hat{\bs{\mu}}_{\text{em}} = \sin\Theta\cos\Phi\ \hat{\mb{i}} +
\sin\Theta\sin\Phi\ \hat{\mb{j}} + \cos\Theta\ \hat{\mb{k}}$), find the minimum
variance of an unbiased estimator for these parameters ($\sigma^2_{\Theta}$ and
$\sigma^2_{\Phi}$) using the Cramer Rao lower bound, and calculate a metric we
call the solid-angle uncertainty---
$\sigma_{\Omega} = \sin\Theta\sigma_{\Theta}\sigma_{\Phi}$. We use
$\sigma_{\Omega}$ as our figure of merit because it is easy to interpret---given
a single molecule's orientation and a set of measurements, $\sigma_{\Omega}$ is
the solid angle of the cone of uncertainty about the molecule's reconstructed
orientation.

\subsection{Single-View Microscope Results}
Figure \ref{fig:one_arm} shows the solid-angle uncertainty for multiple-frame,
single-view experiments. Each experiment consists of four frames with different
polarization orientations and fixed illumination and detection geometry. We used
a Poisson noise model with 1000 illuminating photons per frame, a back focal
plane radius of 3/10 the focal length of the condenser, and a detection NA of
1.3.

\fig{../figures/singlearm3.png}{0.9}{Solid-angle uncertainty for single view multiple frame
  microscopes. Row 1: schematics of the microscope geometries---we detect four
  frames with different illumination polarizations. Row 2: solid-angle
  uncertainty as a function of dipole orientation. A small $\sigma_{\Omega}$
  indicates that the orientation of a dipole oriented along this direction can
  be precisely recovered from the data collected with this microscope
  geometry. Columns: varying angles between the illumination and detection
  arms. }{one_arm}

\subsection{Dual-View Microscope Results}

Figure \ref{fig:dual_arm} shows the solid-angle uncertainty for multiple-frame,
dual-view experiments. We've repeated the experiments in the previous section,
but now we collect four polarization frames with one view, swap the illumination
and detection arms, then collect another four polarization frames. We halved the
number of illuminating photons to 500 per frame to allow for a fair comparison
between the single arm and dual arm results.

\fig{../figures/dualarm3.png}{0.9}{Solid-angle uncertainty for dual view multiple frame
  microscopes. Same caption as Figure \ref{fig:one_arm} except here we
  illuminate and detect along both arms. Notice that the 45${}^{\circ}$
  detection scheme provides the most uniform solid-angle uncertainty of the
  designs considered here.}{dual_arm}

\section{Discussion}
An ideal microscope geometry would allow us to reconstruct the orientation of a
single molecule with a constant solid-angle uncertainty for every orientation of
the molecule. We can't accomplish this goal without a large number of detectors
at all positions around the dipole, so instead we choose microscope geometries
that are experimentally feasible and give us the most uniform solid-angle
uncertainty.

Figure \ref{fig:one_arm} shows that single-view microscopes cannot be used to
reconstruct single molecules with uniform solid-angle uncertainty. All
single-view geometries are degenerate near a plane of molecule orientations. In
the epi-detection case the degeneracy occurs near the plane orthogonal to the
optical axis. A molecule lying near this plane will give rise to data that
cannot be used to determine whether the molecule lies above or below the
plane. Lu discussed this degeneracy in detail and developed techniques to avoid
reconstruction issues \cite{lux}. When the detector is rotated the plane of
degeneracy moves but does not disappear. Notice that moving the detector away
from epi-detection always decreases the detection efficiency, so there is no
advantage to using oblique or orthogonal detection geometries with only one
view.

Figure \ref{fig:dual_arm} shows that oblique and orthogonal geometries can be
advantageous for dual-view microscopes. The first column of Figure
\ref{fig:dual_arm} shows the epi-detection case with the plane of degeneracy
orthogonal to the detection axis. The second and third columns show that
rotating the second arm away from the first arm creates a dramatically more
uniform solid-angle uncertainty. Between the epi-, 45${}^{\circ}$-, and
ortho-detection geometries, we can see that the 45${}^{\circ}$ detection
geometry creates the most uniform solid-angle uncertainty. We note that the
results extend symmetrically past angles larger than 90${}^{\circ}$. For
example, 135${}^{\circ}$ detection gives equivalent results to 45${}^{\circ}$
detection, and 135${}^{\circ}$ detection may be more experimentally feasible.

\section{Next Steps}
Limitations and/or next steps:
\begin{enumerate}
\item Both the illumination and detection models use the paraxial approximation. I know how to extend to the high-NA case, but the conclusions in this work do not depend
  on high-NA lenses. Do you think I should extend now or later?
\item I'm currently only using a Poisson noise model, but I've implemented a
  Poisson + Gaussian model as well if we'd like results with read-out noise. Read-out
  noise won't cause any major difference---it will just raise the solid-angle
  uncertainty baseline.
\item Currently I'm calculating Equation \ref{eq:factored} numerically, and I'm
  getting reasonable results (see Figure \ref{fig:illum_basis} and Figure
  \ref{fig:efficiency}). I can write \ref{eq:factored} in terms of six
  integrals, and I think that those integrals will simplify under the paraxial
  assumption. This may allow us to consider the effect of illumination in closed
  form. Still investigating.
\item The 45${}^{\circ}$ and 135${}^{\circ}$ cases are identical if the molecule
  is in a homogeneous environment. I have the tools to consider the case where
  the molecule is near an interface \cite{backer}, but I haven't implemented
  this yet.
\end{enumerate}

Questions:
\begin{enumerate}
\item I've parameterized the radius of the back focal plane with $\rho/f$, the
  ratio the back focal plane radius to the focal length. Would NA${}_{\text{illum}}$
  be more appropriate?
\item Is 500/1000 illuminating photons per frame appropriate? What are typical
  read out noise levels?
\end{enumerate}

\bibliography{report}{}
\bibliographystyle{unsrt}

\end{document}

### 2017-06-12 notes/2017-06-09-rotated-frames/report/report.tex
Add notes history.

\documentclass[11pt]{article}

%%%%%%%%%%%%
% Packages %
%%%%%%%%%%%%
\hyphenpenalty=10000
\usepackage{tocloft}
\renewcommand\cftsecleader{\cftdotfill{\cftdotsep}}
\usepackage{xcolor}
\usepackage{empheq}
\usepackage{hyperref}
\usepackage{epstopdf}
\usepackage{braket}
\usepackage{upgreek}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{amssymb,latexsym,amsmath,gensymb}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{float}
\usepackage{enumitem}
\usepackage{pdflscape}
\usepackage{url}
\usepackage{tikz, calc}
\usetikzlibrary{shapes.geometric, arrows, calc}
\tikzstyle{norm} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{arrow} = [thick, ->, >=stealth]

\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
\providecommand{\mb}[1]{\mathbf{#1}}
\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
\providecommand{\bs}[1]{\boldsymbol{#1}} 
\providecommand{\intinf}{\int_{-\infty}^{\infty}}
\providecommand{\fig}[4]{
  % filename, width, caption, label
\begin{figure}[h]
 \captionsetup{width=1.0\linewidth}
 \centering
 \includegraphics[width = #2\textwidth]{#1}
 \caption{#3}
 \label{fig:#4}
\end{figure}
}

\newcommand{\tensor}[1]{\overset{\text{\tiny$\leftrightarrow$}}{\mb{#1}}}
\newcommand{\tunderbrace}[2]{\underbrace{#1}_{\textstyle#2}}
\providecommand{\figs}[7]{
  % filename1, filename2, caption1, caption2, label1, label2, shift
\begin{figure}[H]
\centering
\begin{minipage}[b]{.45\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#1}
  \captionsetup{justification=justified, singlelinecheck=true}
  \caption{#3}
  \label{fig:#5}
\end{minipage}
\hspace{2em}
\begin{minipage}[b]{.45\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#2}
  \vspace{#7em}
  \captionsetup{justification=justified}
  \caption{#4}
  \label{fig:#6}
\end{minipage}
\end{figure}
}
\makeatletter

\providecommand{\code}[1]{
\begin{center}
\lstinputlisting{#1}
\end{center}
}
\setlength{\parindent}{0pt}
%%%%%%%%%%%
% Spacing %
%%%%%%%%%%%
% Margins
\usepackage[
top    = 1.5cm,
bottom = 1.5cm,
left   = 1.5cm,
right  = 1.5cm]{geometry}

% Indents, paragraph space
\usepackage{parskip}

% Section spacing
\usepackage{titlesec}
\titlespacing*{\title}
{0pt}{0ex}{0ex}
\titlespacing*{\section}
{0pt}{0ex}{0ex}
\titlespacing*{\subsection}
{0pt}{0ex}{0ex}
\titlespacing*{\subsubsection}
{0pt}{0ex}{0ex}

% Line spacing
\linespread{1.1}

%%%%%%%%%%%%
% Document %
%%%%%%%%%%%%
\begin{document}
\title{\vspace{-2.5em} Relating Spherical Coordinates In Rotated Frames\vspace{-1em}} \author{Talon Chandler}% and Patrick La Rivi\`ere}
\date{\vspace{-1em}June 9, 2017\vspace{-1em}}
\maketitle

\section{Introduction}
In these notes we will relate typical spherical coordinates (Figure
\ref{fig:frames_a}) to obliquely rotated (Figure \ref{fig:frames_b}) and
orthogonal (Figure \ref{fig:frames_c}) spherical coordinates.
\begin{figure*}[h]
 \captionsetup{width=1.0\linewidth}
 \centering
 \begin{subfigure}[t]{0.33\textwidth}
   \centering
   \includegraphics[width = 1.0\textwidth]{../figures/frame_a.pdf}
   \caption{Typical spherical coordinates}
   \label{fig:frames_a}
 \end{subfigure}%
 ~
 \begin{subfigure}[t]{0.33\textwidth}
   \centering
   \includegraphics[width = 1.0\textwidth]{../figures/frame_b.pdf}
   \caption{$\psi$-rotated spherical coordinates}
   \label{fig:frames_b}
 \end{subfigure}%
 ~
 \begin{subfigure}[t]{0.33\textwidth}
   \centering
   \includegraphics[width = 1.0\textwidth]{../figures/frame_c.pdf}
   \caption{Orthogonal spherical coordinates}
   \label{fig:frames_c}
 \end{subfigure}%
 \caption{Possible spherical coordinate definitions.}
 \label{fig:frames}
\end{figure*}

\section{Relating Spherical Coordinates In Two Oblique Frames}
First, we expand a single unit vector $\mh{r}$ in two different coordinate systems
\begin{align}
  \mh{r} = x\mh{x} + y\mh{y} + z\mh{z} &= x'\mh{x}' + y'\mh{y}' + z'\mh{z}'.
\intertext{Next, we rewrite the vector in spherical coordinates}
  \mh{r} = \cos\phi\sin\theta\mh{x} + \sin\phi\sin\theta\mh{y} + \cos\theta\mh{z} &= \cos\phi'\sin\theta'\mh{x}' + \sin\phi'\sin\theta'\mh{y}' + \cos\theta'\mh{z}'.\label{eq:sphere}
\end{align}
Finally, we relate the coordinate systems. We will only consider coordinate
systems that are related by a right handed rotation by angle $\psi$ about the
$\mh{y}$ axis, so the coordinate systems are related by
\begin{subequations}
\begin{align}
  \mh{x} &= \cos\psi\mh{x}' +\sin\psi\mh{z}'\\
  \mh{y} &= \mh{y}'\\
  \mh{z} &= -\sin\psi\mh{x}' + \cos\psi\mh{z}'.
 \end{align}\label{eq:coords}
\end{subequations}%
Plugging equation \ref{eq:coords} into equation \ref{eq:sphere} and equating
each component gives
\begin{subequations}
  \begin{align}
    \cos\psi\cos\phi\sin\phi - \sin\psi\cos\theta\label{eq:preva} &=\cos\phi'\sin\theta'\\
     \sin\phi\sin\theta&=\sin\phi'\sin\theta' \label{eq:prevb}\\
    \sin\psi\cos\phi\sin\theta + \cos\psi\cos\theta&=\cos\theta' .\label{eq:prevc}
  \end{align}
\end{subequations}
Solving for $\theta'$ and $\phi'$ in terms of $\theta$, $\phi$, and $\psi$ gives
\begin{subequations}
\begin{align}
  \theta' &= \arccos\left(\sin\psi\cos\phi\sin\theta + \cos\psi\cos\theta\right)\label{eq:thetap}\\
  \phi' &= \arccos\left(\frac{\cos\psi\cos\phi\sin\phi - \sin\psi\cos\theta}{\sqrt{1 - (\sin\psi\cos\phi\sin\theta + \cos\psi\cos\theta)^2}}\right).\label{eq:phip}
\end{align}\label{eq:solution}% 
\end{subequations}
Equation \ref{eq:thetap} follows directly from equation \ref{eq:prevc}. Equation
\ref{eq:phip} follows from plugging equation \ref{eq:thetap} into equation \ref{eq:prevc} and using the identity $\sin(\arccos(x)) = \sqrt{1 - x^2}$.

The inverse equations can be found by the substituting $\psi \rightarrow -\psi$:
\begin{subequations}
\begin{align}
  \theta &= \arccos\left(-\sin\psi\cos\phi'\sin\theta' + \cos\psi\cos\theta'\right)\\
  \phi &= \arccos\left(\frac{\cos\psi\cos\phi'\sin\phi' + \sin\psi\cos\theta'}{\sqrt{1 - (\sin\psi\cos\phi'\sin\theta' + \cos\psi\cos\theta')^2}}\right).
\end{align}\label{eq:solutionp}% 
\end{subequations}

\section{Relating Spherical Coordinates In Two Orthogonal Frames}
Setting $\psi = \pi/2$ in equation \ref{eq:solution} gives
\begin{subequations}
  \begin{align}
  \theta' &= \arccos\left(\cos\phi\sin\theta\right)\\
  \phi' &= \arccos\left(\frac{-\cos\theta}{\sqrt{1 - \cos^2\phi\sin^2\theta}}\right).
\end{align} \label{eq:solution2}%
\end{subequations}
The inverse equations are
\begin{subequations}
\begin{align}
  \theta &= \arccos\left(-\cos\phi'\sin\theta'\right)\\
  \phi &= \arccos\left(\frac{\cos\theta'}{\sqrt{1 - \cos^2\phi'\sin^2\theta'}}\right).
\end{align} \label{eq:solution2p}% 
\end{subequations}
We'll test these equations with $\theta = \pi/4$ and $\phi=\pi/3$ as illustrated in
Figure \ref{fig:frames_a}. Plugging into equation \ref{eq:solution2} gives
\begin{subequations}
\begin{align}
  \theta' &= \arccos\left(\cos(\pi/3)\sin(\pi/4)\right) = 1.21 = 69.0^{\circ}\\
  \phi' &= \arccos\left(\frac{-\cos(\pi/4)}{\sqrt{1 - \cos^2(\pi/3)\sin^2(\pi/4)}}\right) = 2.43 = 139.1^{\circ}
\end{align}\label{eq:check1}%
\end{subequations}
which is consistent with Figure \ref{fig:frames_c}. We'll test the inverse equation as
well. Plugging equation \ref{eq:check1} into equation \ref{eq:solution2p} gives
\begin{subequations}
\begin{align}
  \theta &= \arccos\left(-\cos(2.43)\sin(1.21)\right) = \pi/4\\
  \phi &= \arccos\left(\frac{\cos(1.21)}{\sqrt{1 - \cos^2(2.43)\sin^2(1.21)}}\right) = \pi/3
\end{align}\label{eq:check1}% 
\end{subequations}
as expected.

\section{Discussion}
Equations \ref{eq:solution2} and \ref{eq:solution2p} will be useful for
analyzing data collected with the polarization diSPIM. If we measure $\theta$
and $\phi$ in both views with the optical axis along the $\mh{z}$ axis, then we
can use the equations \ref{eq:solution2} and \ref{eq:solution2p} to convert the
measurements to the same coordinate system. Note that these equations will only
work if the $\mh{y}$ axis is defined as the normal to the plane containing the
optical axes.
\end{document}

### 2017-06-12 notes/2017-04-25-fourkas-correction/calculations/check.py

@@ -1,0 +1,10 @@
+import numpy as np
+
+alpha = np.pi/2
+
+A = (1.0/6.0) - 0.25*np.cos(alpha) + (1.0/12.0)*(np.cos(alpha)**3)
+B = 0.125*np.cos(alpha) - 0.125*(np.cos(alpha)**3)
+
+print(A)
+print(B)
+print(2*A + 2*B)

### 2017-06-12 notes/2017-04-25-fourkas-correction/calculations/intensities.py

@@ -1,0 +1,86 @@
+import numpy as np
+from mpmath import *
+from sympy import *
+from sympy.matrices.dense import *
+import functools
+import util
+
+print("Working...")
+
+# Symbols
+Theta = Symbol('Theta')
+Phi = Symbol('Phi')
+theta = Symbol('theta')
+phi = Symbol('phi')
+n0 = Symbol('n0')
+n1 = Symbol('n1')
+phi_pol = Symbol('phi_pol')
+alpha = Symbol('alpha')
+
+# Set up integral
+mu = Matrix([sin(Theta)*cos(Phi), sin(Theta)*sin(Phi), cos(Theta)]) # Dipole 
+r = Matrix([sin(theta)*cos(phi), sin(theta)*sin(phi), cos(theta)]) # Position
+R = rot_axis3(-phi)*rot_axis2(theta)*rot_axis3(phi) # Rotation matrix
+apod = 1 # sqrt(n1)/(sqrt(n0*cos(theta))) # Energy conservation
+P = Matrix([cos(phi_pol), sin(phi_pol), 0]) # Polarizer
+J = sin(theta) # Jacobian
+integrand = J*(P.dot(apod*R*r.cross(mu).cross(r))**2) # Integrand
+tot_integrand = J*(r.cross(mu).cross(r)).dot((r.cross(mu).cross(r))) 
+
+# Calculate I integral
+int1 = integrate(expand(integrand), (phi, 0, 2*pi)) # phi integral
+I = integrate(expand(int1), (theta, 0, alpha)) # theta integral
+
+# Calculate I_tot integral
+int1 = integrate(expand(tot_integrand), (phi, 0, 2*pi)) # phi integral
+I_tot = integrate(expand(int1), (theta, 0, pi)) # theta integral
+
+# Calculate fraction of total power detected
+I_frac = I/I_tot
+
+# Print expressions for each polarization in Fourkas' form
+pols = [0, pi/4, pi/2, 3*pi/4]
+for pol in pols:
+    I_pol = trigsimp(I_frac.subs([(phi_pol, pol)]))
+    B_term = sin(Theta)**2
+    if pol == 0 or pol == pi/2:
+        sub_in = sin(Phi)**2
+        sub_out = (1 - cos(2*Phi))/2
+        C_term = sin(Theta)**2*cos(2*Phi)
+    else:
+        sub_in = sin(Phi)*cos(Phi)
+        sub_out = sin(2*Phi)/2
+        C_term = sin(Theta)**2*sin(2*Phi)
[…]

### 2017-06-12 notes/2017-03-02-illumination-update1/report.tex
Cleaning notes history.

\documentclass[11pt]{article}

%%%%%%%%%%%%
% Packages %
%%%%%%%%%%%%
\hyphenpenalty=10000
\usepackage{tocloft}
\renewcommand\cftsecleader{\cftdotfill{\cftdotsep}}

\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{epstopdf}
\usepackage{braket}
\usepackage{upgreek}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{amssymb,latexsym,amsmath,gensymb}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{float}
\usepackage{enumitem}
\usepackage{pdflscape}
\usepackage{url}
\usepackage{tikz, calc}
\usetikzlibrary{shapes.geometric, arrows, calc}
\tikzstyle{norm} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{arrow} = [thick, ->, >=stealth]

\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
\providecommand{\mb}[1]{\mathbf{#1}}
\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
\providecommand{\bs}[1]{\boldsymbol{#1}} 
\providecommand{\intinf}{\int_{-\infty}^{\infty}}
\providecommand{\fig}[4]{
  % filename, width, caption, label
\begin{figure}[h]
 \captionsetup{width=1.0\linewidth}
 \centering
 \includegraphics[width = #2\textwidth]{#1}
 \caption{#3}
 \label{fig:#4}
\end{figure}
}

\newcommand{\tensor}[1]{\overset{\text{\tiny$\leftrightarrow$}}{\mb{#1}}}
\newcommand{\tunderbrace}[2]{\underbrace{#1}_{\textstyle#2}}
\providecommand{\figs}[7]{
  % filename1, filename2, caption1, caption2, label1, label2, shift
\begin{figure}[H]
\centering
\begin{minipage}[b]{.4\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#1}
  \captionsetup{justification=justified, singlelinecheck=true}
  \caption{#3}
  \label{fig:#5}
\end{minipage}
\hspace{2em}
\begin{minipage}[b]{.4\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#2}
  \vspace{#7em}
  \captionsetup{justification=justified}
  \caption{#4}
  \label{fig:#6}
\end{minipage}
\end{figure}
}
\makeatletter

\providecommand{\code}[1]{
\begin{center}
\lstinputlisting{#1}
\end{center}
}

%%%%%%%%%%%
% Spacing %
%%%%%%%%%%%
% Margins
\usepackage[
top    = 1.5cm,
bottom = 1.5cm,
left   = 1.5cm,
right  = 1.5cm]{geometry}

% Indents, paragraph space
\usepackage{parskip}

% Section spacing
\usepackage{titlesec}
\titlespacing*{\title}
{0pt}{0ex}{0ex}
\titlespacing*{\section}
{0pt}{0ex}{0ex}
\titlespacing*{\subsection}
{0pt}{0ex}{0ex}
\titlespacing*{\subsubsection}
{0pt}{0ex}{0ex}

% Line spacing
\linespread{1.1}

%%%%%%%%%%%%
% Document %
%%%%%%%%%%%%
\begin{document}
\title{\vspace{-2.5em}Single Molecule Fluorescence Microscopy Update\vspace{-1em}}
\author{Talon Chandler}% and Patrick La Rivi\`ere}
\date{\vspace{-1em}March 3, 2017\vspace{-1em}}
\maketitle

\section{Current Model and Limitations}
In the previous note set, we developed the following forward model
\begin{align}
  I_{\text{img}}(\mb{r''}) = \left|\mathcal{F}_{3D}\left\{
    \mb{\tilde{O}}_{\text{obj}}\tensor{\mathbf{G}}_{FF}\tensor{\bs{\alpha}}\mb{E}_{\text{in}}
  \right\}
  \right|^2.\label{eq:5}
\end{align}

Rudolph and Shalin identified several limitations of the model:
\begin{itemize}
\item The model needlessly tracks phase from start to finish. Rudolph and Shalin
  correctly pointed out that phase is irrelevant on the illumination side
  because we measure the intensity over long time periods compared to the period
  of the electric fields.
\item The model only considers monochromatic illumination by a single plane wave
  and focused laser illumination, not K\"{o}hler illumination by a broadband
  source.
\item The model only considers excitation of a single fluorophore. 
\end{itemize}
These notes address these limitations.

\subsection{Single Fluorophore, Monochromatic Plane Wave}
Consider a monochromatic plane wave with polarization axis $\mb{E}$
incident on a single fluorophore with absorption dipole moment axis
$\hat{\bs{\mu}}_{\text{abs}}$ and emission dipole moment
$\hat{\bs{\mu}}_{\text{em}}$. The incident plane wave induces a dipole moment given by
\begin{align}
  \bs{\mu}_{\text{ind}} \propto \hat{\bs{\mu}}_{\text{em}} \left[\hat{\bs{\mu}}_{\text{abs}}\cdot \mb{E}\right].
\end{align}

\subsection{Many Fluorophores, Monochromatic Plane Wave}
Now consider $N$ fluorophores in the specimen volume with absorption dipole
moments $\hat{\bs{\mu}}_{\text{abs},i}$, emission dipole moments
$\hat{\bs{\mu}}_{\text{em},i}$, and positions $\mb{r}_i$. We define the
\textit{emission dipole moment map} as
$\bs{\mu}_{\text{em}}(\mb{r}) \equiv \sum_{i=0}^N \hat{\bs{\mu}}_{\text{em},i}
\delta (\mb{r} - \mb{r}_i)$. Similarly, we define the \textit{absorption dipole
  moment map} as
$\bs{\mu}_{\text{abs}}(\mb{r}) \equiv \sum_{i=0}^N \hat{\bs{\mu}}_{\text{abs},i}
\delta (\mb{r} - \mb{r}_i)$. When we illuminate the fluorophores with a
monochromatic plane wave we create the \textit{induced dipole map} given by
\begin{align}
    \bs{\mu}_{\text{ind}}(\mb{r}) \propto \bs{\mu}_{\text{em}}(\mb{r}) \left[\bs{\mu}_{\text{abs}}(\mb{r})\cdot \mb{E}\right].
\end{align}

\subsection{Many Fluorophores,  Focused Laser Illumination}
Under focused laser illumination the electric field in the specimen volume
is position dependent. If $\mb{E}(\mb{r})$ is the real 3D electric field
direction at every point in the specimen volume, then the induced dipole map is given
by
\begin{align}
    \bs{\mu}_{\text{ind}}(\mb{r}) \propto \bs{\mu}_{\text{em}}(\mb{r}) \left[\bs{\mu}_{\text{abs}}(\mb{r})\cdot \mb{E}(\mb{r})\right].
\end{align}

\subsection{Many Fluorophores, Broadband Plane Wave}
If we illuminate a fluorophore with a broadband plane wave and measure the
induced dipole moment over a period of time much longer than the coherence time,
then the effective induced dipole moment will be the sum of the induced dipole
moments created by each frequency component. In other words, each frequency
component acts independently on the fluorophore to create an induced dipole
moment.

If we want to compare illumination sources with different spectra, we will need
to consider the \textit{excitation efficiency}, $\eta(\omega)$, of each
frequency $\omega$. For a two level system, $\eta(\omega) = 1$ at the resonance
frequency and drops to 0 far from the resonance frequency following a Lorentzian
function. $\eta(\omega)$ is a more complicated function for real
fluorophores. The induced dipole map is given by
\begin{align}
    \bs{\mu}_{\text{ind}}(\mb{r}) \propto \intinf d\omega\ \eta(\omega) \bs{\mu}_{\text{em}}(\mb{r}) \left[\bs{\mu}_{\text{abs}}(\mb{r})\cdot \mb{E}(\omega)\right].
\end{align}

\subsection{Many Fluorophores, K\"{o}hler Illumination}
If we illuminate a fluorophore with two plane waves with a random phase
difference traveling in different directions and measure the induced dipole over
a long time compared with the coherence time of the two waves, the induced
dipole is the sum of the induced dipole created by each plane wave
independently. In other words, each plane wave creates an induced dipole
independently. (Thank you Rudolph and Shalin for steering me in this direction).

Under K\"{o}hler illumination with a polarizer in the back focal plane, each
paraxial point in the back focal plane of the condenser creates a polarized
plane wave that is constant throughout the specimen volume. Therefore, to find
the induced dipole moment map, we can integrate over the back focal plane of the condenser
\begin{align}
  \bs{\mu}_{\text{ind}}(\mb{r}) \propto
  \intinf d\omega
  \int d\mb{r'}\ \eta(\omega)\bs{\mu}_{\text{em}}(\mb{r})
  \left[\bs{\mu}_{\text{abs}}(\mb{r})\cdot \mb{R}_{\mb{\hat{s}}}(\mb{r'})\mb{E}_{\text{bfp}}(\omega)\right] \label{eq:test}
\end{align}
where $\mb{r'}$ is the position in the back focal plane, $\mb{E}_{\text{bfp}}$
is the electric field direction in the back focal plane set by a linear
polarizer, and $\mb{R}_{\mb{\hat{s}}}(\mb{r'})$ is a rotation matrix that rotates the electric field in the back focal plane to the electric field in the specimen volume.

Note that equation \ref{eq:test} depends on the paraxial approximation. Points
in the back focal plane that are far from the optical axis do not create perfect
plane waves in the specimen volume. We can explore more accurate models for high
NA illumination if necessary.

\section{Revised Model Summary}
On the detection side, the model remains the same. The intensity in the image plane
is given by
\begin{align}
  I_{\text{img}}(\mb{r''}) \propto \left|\mathcal{F}_{3D}\left\{
  \mb{\tilde{O}}_{\text{obj}}\tensor{\mathbf{G}}_{FF}\bs{\mu}_{\text{ind}}(\mb{r})
  \right\}
  \right|^2.
\end{align}

Under K\"{ohler} illumination the induced dipole moment map is
\begin{align}
  \bs{\mu}_{\text{ind}}(\mb{r}) \propto
  \intinf d\omega
  \int d\mb{r'}\ \eta(\omega)\bs{\mu}_{\text{em}}(\mb{r})
  \left[\bs{\mu}_{\text{abs}}(\mb{r})\cdot \mb{R}_{\mb{\hat{s}}}(\mb{r'})\mb{E}_{\text{bfp}}(\omega)\right] \label{eq:test2}
\end{align}
Under laser illumination the induced dipole moment is\begin{align}
  \bs{\mu}_{\text{ind}}(\mb{r}) \propto
  \bs{\mu}_{\text{em}}(\mb{r})
  \left[\bs{\mu}_{\text{abs}}(\mb{r})\cdot \mb{E}(\mb{r})\right] \label{eq:test3}
\end{align}
where $\mb{E}(\mb{r})$ is the electric field in the specimen volume.

\section{Matrix Form of the Model}
Rudolph and Shalin suggested simplifying the model by defining a
\textit{polarization resolved intensity},
$\mb{I}_{\text{in}} = \left[I_{\text{in}}^x, I_{\text{in}}^y,
  I_{\text{in}}^z\right]$ that we could use to create a matrix form of the
model. Unfortunately, both the emission and absorption dipole maps appear in
equations \ref{eq:test2} and \ref{eq:test3}. This means that even if the
emission and absorption dipoles point in the same direction, the induced dipole
moments depend on the cross terms of the dipole direction. In other words, the
model is linear in
$\left[\mu_x^2, \mu_y^2, \mu_z^2, \mu_{x}\mu_y, \mu_x\mu_z, \mu_y\mu_z\right]$,
not just $\left[\mu_x^2, \mu_y^2, \mu_z^2\right]$. If we define a polarization resolved intensity vector it will need six terms, not three.

Backer and Moerner made a similar simplification on the detection side to speed
up their deconvolutions \cite{backer}.

\bibliography{report}{}
\bibliographystyle{unsrt}

\end{document}

### 2017-06-14 notes/2017-06-14-illumination-calculation/report/report.tex
Add closed form wide field notes.

\documentclass[11pt]{article}

%%%%%%%%%%%%
% Packages %
%%%%%%%%%%%%
\hyphenpenalty=10000
\usepackage{tocloft}
\renewcommand\cftsecleader{\cftdotfill{\cftdotsep}}
\def\undertilde#1{\mathord{\vtop{\ialign{##\crcr
$\hfil\displaystyle{#1}\hfil$\crcr\noalign{\kern1.5pt\nointerlineskip}
$\hfil\tilde{}\hfil$\crcr\noalign{\kern1.5pt}}}}}
\usepackage{cleveref}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{epstopdf}
\usepackage{braket}
\usepackage{upgreek}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{amssymb,latexsym,amsmath,gensymb}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{float}
\usepackage{enumitem}
\usepackage{pdflscape}
\usepackage{url}
\usepackage{tikz, calc}
\usetikzlibrary{shapes.geometric, arrows, calc}
\tikzstyle{norm} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{arrow} = [thick, ->, >=stealth]

\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
\providecommand{\mb}[1]{\mathbf{#1}}
\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
\providecommand{\bs}[1]{\boldsymbol{#1}} 
\providecommand{\intinf}{\int_{-\infty}^{\infty}}
\providecommand{\fig}[4]{
  % filename, width, caption, label
\begin{figure}[h]
 \captionsetup{width=1.0\linewidth}
 \centering
 \includegraphics[width = #2\textwidth]{#1}
 \caption{#3}
 \label{fig:#4}
\end{figure}
}

\newcommand{\tensor}[1]{\overset{\text{\tiny$\leftrightarrow$}}{\mb{#1}}}
\newcommand{\tunderbrace}[2]{\underbrace{#1}_{\textstyle#2}}
\providecommand{\figs}[7]{
  % filename1, filename2, caption1, caption2, label1, label2, shift
\begin{figure}[H]
\centering
\begin{minipage}[b]{.45\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#1}
  \captionsetup{justification=justified, singlelinecheck=true}
  \caption{#3}
  \label{fig:#5}
\end{minipage}
\hspace{2em}
\begin{minipage}[b]{.45\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#2}
  \vspace{#7em}
  \captionsetup{justification=justified}
  \caption{#4}
  \label{fig:#6}
\end{minipage}
\end{figure}
}
\makeatletter

\providecommand{\code}[1]{
\begin{center}
\lstinputlisting{#1}
\end{center}
}

\newcommand{\crefrangeconjunction}{--}
%%%%%%%%%%%
% Spacing %
%%%%%%%%%%%
% Margins
\usepackage[
top    = 1.5cm,
bottom = 1.5cm,
left   = 1.5cm,
right  = 1.5cm]{geometry}

% Indents, paragraph space
\usepackage{parskip}

% Section spacing
\usepackage{titlesec}
\titlespacing*{\title}
{0pt}{0ex}{0ex}
\titlespacing*{\section}
{0pt}{0ex}{0ex}
\titlespacing*{\subsection}
{0pt}{0ex}{0ex}
\titlespacing*{\subsubsection}
{0pt}{0ex}{0ex}

% Line spacing
\linespread{1.1}

%%%%%%%%%%%%
% Document %
%%%%%%%%%%%%
\begin{document}
\title{\vspace{-2.5em} Closed Form Expressions for the Excitation and Detection Efficiency of Single Dipoles Under Polarized Wide-Field Illumination \vspace{-1em}}
\author{Talon Chandler}% and Patrick La Rivi\`ere}
\date{\vspace{-1em}\today\vspace{-1em}}
\maketitle
\section{Introduction}
In these notes I will develop the relationships between the excitation/detection
efficiencies and the dipole orientation, polarizer orientation, and microscope
geometry for wide-field illumination microscopes. First, I will write out the
relationships for the epi-illumination and epi-detection case, then I will
generalize the relationships for oblique or orthogonal geometries.

\section{Detection Efficiency}
In the 2017-04-25 notes I followed Fourkas and showed that the detection
efficiency (the fraction of emitted power that we collect) for a polarized
wide-field microscope is
\begin{align}
  \eta_{\text{det},\phi_{\text{det}}}  = \frac{\int_{\Omega}d\mh{r}\ \left|\mh{P}_{\text{det}}\cdot\tilde{\mb{R}}(\mh{r})(\mh{r}\times\hat{\bs{\mu}}_{\text{em}}\times\mh{r})\right|^2}{\int_{\mathbb{S}^2}d\mh{r}\left|\mh{r}\times\hat{\bs{\mu}}_{\text{em}}\times\mh{r}\right|^2}\label{eq:final}
\end{align}
where
\begin{align}
  \hat{\mb{r}} &= \sin\theta\cos\phi\hat{\mb{i}} + \sin\theta\sin\phi\hat{\mb{j}} + \cos\theta\hat{\mb{k}}\label{eq:r_coords}\ \ \ \text{is the dummy integration vector,}\\
  \hat{\bs{\mu}}_{\text{em}} &= \sin\Theta\cos\Phi\hat{\mb{i}} + \sin\Theta\sin\Phi\hat{\mb{j}} + \cos\Theta\hat{\mb{k}}\ \ \ \text{is the emission dipole moment,}\label{eq:mu_coords}\\
  \tilde{\mb{R}}(\mh{r}) &= \begin{bmatrix} \cos\theta\cos^2\phi + \sin^2\phi & (\cos\theta -1)\sin\phi\cos\phi & -\sin\theta\cos\phi\\ (\cos\theta - 1)\sin\phi\cos\phi & \cos\theta\sin^2\phi + \cos^2\phi & -\sin\theta\sin\phi \\ \sin\theta\cos\phi& \sin\theta\sin\phi & \cos\theta \end{bmatrix}\label{eq:matrix}\\&\text{is the rotation matrix that models the objective lens,}\\
  \hat{\mb{P}}_{\text{det}} &= \cos\phi_{\text{det}}\hat{\mb{i}} + \sin\phi_{\text{det}}\hat{\mb{j}}\ \ \ \text{is the pass axis of the polarizer,}\\
  \Omega &= \{(\phi, \theta)|\phi \in (0,2\pi]\ \text{and}\ \theta\in (0, \alpha]\}\ \ \ \text{is the cap of the sphere collected by the objective.}\label{eq:omega}
\end{align}
Plugging equations \ref{eq:r_coords}--\ref{eq:omega} into equation \ref{eq:final} and
simplifying gives Fourkas' result with an extra factor of $\frac{3}{2}$
\begin{subequations}
\begin{align}
  \eta_{\text{det},0}(\Theta, \Phi, \alpha) &= A + B\sin^{2}{\Theta} + C\sin^{2}{\Theta} \cos{2 \Phi}\\
  \eta_{\text{det},45}(\Theta, \Phi, \alpha) &= A + B\sin^{2}{\Theta} + C\sin^{2}{\Theta} \sin{2 \Phi}\\
  \eta_{\text{det},90}(\Theta, \Phi, \alpha) &= A + B\sin^{2}{\Theta} - C\sin^{2}{\Theta} \cos{2 \Phi}\\
  \eta_{\text{det},135}(\Theta, \Phi, \alpha) &= A + B\sin^{2}{\Theta} - C\sin^{2}{\Theta} \sin{2 \Phi}
\end{align}\label{eq:int1}
\end{subequations}
where
\vspace{-1em}
\begin{subequations}
\begin{align}
  A &= \frac{1}{4} - \frac{3}{8} \cos{\alpha } + \frac{1}{8} \cos^{3}{\alpha }\\
  B &= \frac{3}{16} \cos{\alpha } - \frac{3}{16} \cos^{3}{\alpha }\\
  C &= \frac{7}{32} - \frac{3}{32} \cos{\alpha } - \frac{3}{32} \cos^{2}{\alpha } - \frac{1}{32} \cos^{3}{\alpha}.
\end{align}\label{eq:coeff2}%
\end{subequations}
If we remove the polarizer from the detection arm the detection efficiency is
\begin{align}
  \eta_{\text{det},\times}(\Theta, \alpha) = \eta_{\text{det},0} + \eta_{\text{det},90} = 2A + 2B\sin^2\Theta
\end{align}
where $\times$ denotes ``no polarizer''.

\section{Excitation Efficiency}
Here we calculate the excitation efficiency (the fraction of incident power that
excites the fluorophore). We calculate the power that excites the dipole by
taking the 3D Jones vector that is incident on the condenser
$\mh{P}_{\text{exc}}$, pass it through the condenser using a rotation matrix
$\tilde{\mb{R}}$, take the dot product with the absorption dipole moment
$\hat{\bs{\mu}}_{\text{abs}}$, take the modulus squared to find the intensity,
then integrate over the cone of illumination. Finally, we find the excitation
efficiency by dividing by the total power incident on the fluorophore. The final
expression is
\begin{align}
  \eta_{\text{exc},\phi_{\text{exc}}} = \frac{\int_{\Omega}d\mh{r}|\hat{\bs{\mu}}_{\text{abs}}\cdot\tilde{\mb{R}}(\mh{r})\mh{P}_{\text{exc}}|^2}{\int_{\Omega}d\mh{r}}\label{eq:finalexc}
\end{align}
where
\begin{align}
  \hat{\mb{r}} &= \sin\theta\cos\phi\hat{\mb{i}} + \sin\theta\sin\phi\hat{\mb{j}} + \cos\theta\hat{\mb{k}}\label{eq:r_coords2}\ \ \ \text{is the dummy integration vector,}\\
  \hat{\bs{\mu}}_{\text{abs}} &= \sin\Theta\cos\Phi\hat{\mb{i}} + \sin\Theta\sin\Phi\hat{\mb{j}} + \cos\Theta\hat{\mb{k}}\ \ \ \text{is the absorption dipole moment,}\label{eq:mu_coords}\\
  \tilde{\mb{R}}(\mh{r}) &= \begin{bmatrix} \cos\theta\cos^2\phi + \sin^2\phi & (\cos\theta -1)\sin\phi\cos\phi & -\sin\theta\cos\phi\\ (\cos\theta - 1)\sin\phi\cos\phi & \cos\theta\sin^2\phi + \cos^2\phi & -\sin\theta\sin\phi \\ \sin\theta\cos\phi& \sin\theta\sin\phi & \cos\theta \end{bmatrix}\label{eq:matrix}\\&\text{is the rotation matrix that models the condenser lens,}\\
  \hat{\mb{P}}_{\text{exc}} &= \cos\phi_{\text{exc}}\hat{\mb{i}} + \sin\phi_{\text{det}}\hat{\mb{j}}\ \ \ \text{is the pass axis of the excitation polarizer,}\\
  \Omega &= \{(\phi, \theta)|\phi \in (0,2\pi]\ \text{and}\ \theta\in (0, \alpha]\}\ \ \ \text{is the cap of the sphere illuminated by the condenser.}\label{eq:omega2}
\end{align}
Plugging equations \ref{eq:r_coords2}--\ref{eq:omega2} into equation
\ref{eq:finalexc} and simplifying gives
\begin{subequations}
\begin{align}
  \eta_{\text{exc},0}(\Theta, \Phi, \alpha) &= D(A + B\sin^{2}{\Theta} + C\sin^{2}{\Theta} \cos{2 \Phi})\\
  \eta_{\text{exc},45}(\Theta, \Phi, \alpha) &= D(A + B\sin^{2}{\Theta} + C\sin^{2}{\Theta} \sin{2 \Phi})\\
  \eta_{\text{exc},90}(\Theta, \Phi, \alpha) &= D(A + B\sin^{2}{\Theta} - C\sin^{2}{\Theta} \cos{2 \Phi})\\
  \eta_{\text{exc},135}(\Theta, \Phi, \alpha) &= D(A + B\sin^{2}{\Theta} - C\sin^{2}{\Theta} \sin{2 \Phi})
\end{align}\label{eq:int2}
\end{subequations}
where
\vspace{-1em}
\begin{align}
  D = \frac{4}{3(1 - \cos\alpha)}
\end{align}\label{eq:coeff2}%
The excitation efficiency is the same as the detection efficiency with an extra
factor $D$.

Finally, we'll confirm that the excitation efficiency reduces to Malus' law when $\Theta = \pi/2$ and as $\alpha \rightarrow 0$.
\begin{align*}
  &=\lim_{\alpha\rightarrow 0} \eta_{\text{exc},0}(\pi/2, \Phi, \alpha)\\
  &= \lim_{\alpha\rightarrow 0}D(A + B + C\cos{2 \Phi})\\
  &= \lim_{\alpha\rightarrow 0} \left(\frac{4}{3(1-\cos\alpha)}\right)\left(\frac{1}{4} - \frac{3}{16} \cos{\alpha } - \frac{1}{16} \cos^{3}{\alpha } + \left(\frac{7}{32} - \frac{3}{32} \cos{\alpha } - \frac{3}{32} \cos^{2}{\alpha } - \frac{1}{32} \cos^{3}{\alpha}\right)\cos 2\Phi\right)\\
  &= \lim_{\alpha\rightarrow 0} \frac{\frac{1}{4} - \frac{1}{4} \cos{\alpha } - \frac{1}{12} \cos^{3}{\alpha } + \left(\frac{7}{24} - \frac{1}{8} \cos{\alpha } - \frac{1}{8} \cos^{2}{\alpha } - \frac{1}{24} \cos^{3}{\alpha}\right)\cos 2\Phi}{1 - \cos\alpha}\\
  \intertext{Applying L'Hospital's rule gives}
  &= \lim_{\alpha\rightarrow 0} \frac{\frac{1}{4} \sin{\alpha } + \frac{1}{4}\sin\alpha\cos^{2}{\alpha } + \left(\frac{1}{8} \sin{\alpha } + \frac{1}{4}\sin \alpha \cos{\alpha } + \frac{1}{8} \sin\alpha\cos^{2}{\alpha}\right)\cos 2\Phi}{\sin\alpha}\\
  &= \frac{1}{2} + \frac{1}{2}\cos2\Phi = \boxed{\cos^2{\Phi}}.
\end{align*}

\section{Epi-illumination and Epi-detection Forward Model}
If we excite with polarized illumination from above and detect without a
polarizer from the same direction the intensity we expect to collect is
\begin{align}
  I_{\phi_{\text{exc}}} &= I_{\text{tot}}\eta_{\text{exc},\phi_{\text{exc}}}\eta_{\text{det}, \times}
\end{align}
where $I_{\text{tot}}$ is the intensity collected from the fluorophore if we had
an experimental setup that had an excitation and detection efficiency of one.

If we collect frames with four polarization settings we get the following expressions
\begin{subequations}
\begin{align}
  I_{0}(\Theta, \Phi, \alpha) &= I_{\text{tot}}D(A + B\sin^{2}{\Theta} + C\sin^{2}{\Theta} \cos{2 \Phi})(2A + 2B\sin^2\Theta)\\
  I_{45}(\Theta, \Phi, \alpha) &= I_{\text{tot}}D(A + B\sin^{2}{\Theta} + C\sin^{2}{\Theta} \sin{2 \Phi})(2A + 2B\sin^2\Theta)\\
  I_{90}(\Theta, \Phi, \alpha) &= I_{\text{tot}}D(A + B\sin^{2}{\Theta} - C\sin^{2}{\Theta} \cos{2 \Phi})(2A + 2B\sin^2\Theta)\\
  I_{135}(\Theta, \Phi, \alpha) &= I_{\text{tot}}D(A + B\sin^{2}{\Theta} - C\sin^{2}{\Theta} \sin{2 \Phi})(2A + 2B\sin^2\Theta).
\end{align}\label{eq:int1}
\end{subequations}

\section{Oblique or Orthogonal Arms}
The expressions above assume that the optical axis is aligned with the $\mh{z}$
axis. If the optical axis is along a different axis, we could modify the
expressions by (1) changing the limits of integration in equations
\ref{eq:final} and \ref{eq:finalexc} and recalculating or (2) performing a
change of coordinates in equation \ref{eq:int2}. Approach (2) is much easier and
uses the change of spherical coordinates derived in the 2017-06-09 notes. The
main results are reproduced here
\begin{subequations}
\begin{align}
  \Theta' &= \arccos\left(\sin\psi\cos\Phi\sin\Theta + \cos\psi\cos\Theta\right)\label{eq:thetap}\\
  \Phi' &= \arccos\left(\frac{\cos\psi\cos\Phi\sin\Phi - \sin\psi\cos\Theta}{\sqrt{1 - (\sin\psi\cos\Phi\sin\Theta + \cos\psi\cos\Theta)^2}}\right)\label{eq:phip}
\end{align}\label{eq:solution}% 
\end{subequations}
\begin{subequations}
\begin{align}
  \Theta &= \arccos\left(-\sin\psi\cos\Phi'\sin\Theta' + \cos\psi\cos\Theta'\right)\\
  \Phi &= \arccos\left(\frac{\cos\psi\cos\Phi'\sin\Phi' + \sin\psi\cos\Theta'}{\sqrt{1 - (\sin\psi\cos\Phi'\sin\Theta' + \cos\psi\cos\Theta')^2}}\right)
\end{align}\label{eq:solutionp}% 
\end{subequations}

The forward model for an oblique detection or excitation arm can be found by plugging
equation \ref{eq:solutionp} into equation \ref{eq:int1}.

\end{document}

### 2017-06-14 notes/2017-06-14-illumination-calculation/calculations/intensities.py

@@ -1,0 +1,114 @@
+import numpy as np
+from mpmath import *
+from sympy import *
+from sympy.matrices.dense import *
+import functools
+import util
+
+print("Working...")
+
+# Symbols
+Theta = Symbol('Theta')
+Phi = Symbol('Phi')
+theta = Symbol('theta')
+phi = Symbol('phi')
+phi_pol = Symbol('phi_pol')
+alpha = Symbol('alpha')
+psi = Symbol('psi')
+
+# Set up integral
+mu = Matrix([sin(Theta)*cos(Phi), sin(Theta)*sin(Phi), cos(Theta)]) # Dipole
+mu_star = rot_axis2(psi)*mu # Dipole in rotated frame
+r = Matrix([sin(theta)*cos(phi), sin(theta)*sin(phi), cos(theta)]) # Position
+R = rot_axis3(-phi)*rot_axis2(theta)*rot_axis3(phi) # Rotation matrix
+P = Matrix([cos(phi_pol), sin(phi_pol), 0]) # Polarizer
+J = sin(theta) # Jacobian
+
+integrand = J*((mu.dot(R*P))**2)
+tot_integrand = J
+
+# Calculate I integral
+int1 = integrate(expand(integrand), (phi, 0, 2*pi)) # phi integral
+exc_power = integrate(expand(int1), (theta, 0, alpha)) # theta integral
+
+# Calculate I_tot integral
+int1 = integrate(expand(tot_integrand), (phi, 0, 2*pi)) # phi integral
+in_power = integrate(expand(int1), (theta, 0, alpha)) # theta integral
+
+# Calculate fraction of total power that excitest the molecule
+exc_eff = exc_power/in_power
+
+# Print expressions for each polarization in Fourkas' form
+pols = [0, pi/4, pi/2, 3*pi/4]
+for pol in pols:
+    # Factor out constant D from front    
+    D_exc_pol = trigsimp(exc_eff.subs([(phi_pol, pol)]))
+    D = -4/(3*(cos(alpha) - 1))
+    exc_pol = simplify(D_exc_pol/D)
+
+    B_term = sin(Theta)**2    
+    if pol == 0 or pol == pi/2:
+        sub_in = sin(Phi)**2
+        sub_out = (1 - cos(2*Phi))/2
+        C_term = sin(Theta)**2*cos(2*Phi)
+    else:
+        sub_in = sin(Phi)*cos(Phi)
[…]

### 2017-06-14 notes/2017-06-14-illumination-calculation/calculations/util.py

@@ -1,0 +1,90 @@
+from dipsim import visuals
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+import vispy
+
+def fibonacci_sphere(n):
+    # Returns "equally" spaced points on a unit sphere in spherical coordinates.
+    # http://stackoverflow.com/a/26127012/5854689
+    z = np.linspace(1 - 1/n, -1 + 1/n, num=n) 
+    theta = np.arccos(z)
+    phi = np.mod((np.pi*(3.0 - np.sqrt(5.0)))*np.arange(n), 2*np.pi)
+    return np.vstack((theta, phi)).T
+
+def plot_sphere(filename, directions=None, data=None, interact=False,
+                color_norm='linear', color_min=0, color_max=4*np.pi,
+                gamma=0.25, color_map='coolwarm',
+                my_ax=None, my_cax=None, dpi=500, vis_px=1000,
+                save_file=False):
+
+    # Setup viewing window
+    vispy.use('glfw')
+    canvas = vispy.scene.SceneCanvas(keys='interactive', bgcolor='white',
+                               size=(vis_px, vis_px), show=interact)
+    my_cam = vispy.scene.cameras.turntable.TurntableCamera(fov=0, azimuth=135,
+                                                           scale_factor=2.05)
+    view = canvas.central_widget.add_view(camera=my_cam)
+
+    # Plot dots
+    dots = vispy.scene.visuals.Markers(parent=view.scene)
+    dots.antialias = False
+    dots.set_data(pos=np.array([[1.01,0,0],[0,1.01,0],[0,0,1.01]]),
+                  edge_color='black', face_color='black', size=vis_px/50)
+
+    # Calculate colors
+    if color_norm == 'linear':
+        norm = matplotlib.colors.Normalize(vmin=color_min, vmax=color_max)
+    elif color_norm == 'log':
+        norm = matplotlib.colors.LogNorm(vmin=color_min, vmax=color_max)
+    elif color_norm == 'linlog':
+        norm = matplotlib.colors.SymLogNorm(linthresh=1e-3, vmin=data.min(), vmax=data.max())
+    elif color_norm == 'power':
+        norm = matplotlib.colors.PowerNorm(gamma=gamma, vmin=data.min(), vmax=data.max())
+
+    norm_data = norm(data).data
+    norm_data = np.expand_dims(norm_data, 1)    
[…]

### 2017-06-21 notes/2017-06-21-light-sheet-illumination/report.tex
Light sheet illumination notes.

\documentclass[11pt]{article}

\usepackage{myreport}
\usepackage[american]{babel}
\usepackage[backend=biber, sorting=none]{biblatex}
\addbibresource{report.bib}

\begin{document}
\title{\vspace{-2.5em} Excitation and Detection Efficiency Under Scanned Light
  Sheet Illumination \vspace{-2.5em}} \author{Talon Chandler}
\date{\vspace{-1em}\today\vspace{-1em}}
\maketitle
\section{Introduction}
In these notes I will calculate the excitation and detection efficiency of a
single fluorophore under polarized scanned light sheet illumination. I will
start by finding the transverse and longitudinal fields at all positions in a
Gaussian beam using reasonable approximations. Next I will calculate the
excitation efficiency of a single fluorophore as the beam is scanned across the
fluorophore. I will combine these results with the results from previous notes
to write the complete forward model. Finally I will discuss the approximations
we may need to simplify reconstruction.

\section{Transverse and Longitudinal Fields In A Scanned Gaussian Beam}
The complex spatial electric field of a paraxial Gaussian beam propagating along the
$\mh{z}$ axis is \cite{nov}
\begin{align}
  \mb{E}(x,y,z) &= \mb{A}\frac{w_0}{w(z)}\text{exp}\left\{{-\frac{(x^2+y^2)}{w^2(z)}}+i\left[kz - \eta(z) + \frac{k(x^2+y^2)}{2R(z)}\right]\right\}\label{eq:gauss}
\end{align}
where
\begin{alignat}{2}  
  \mb{A} &= \cos\phi_{\text{pol}}\mh{x} + \sin\phi_{\text{pol}}\mh{y}\qquad &&\text{is the input Jones vector},\\
  w_0 &\approx \frac{2n}{k(\text{NA})}\qquad &&\text{is the beam waist radius,}\\
  z_0 &= \frac{kw_0^2}{2} &&\text{is the Rayleigh range,}\\
  k &= \frac{2\pi n}{\lambda}\qquad &&\text{is the wave number,}\\
  w(z) &= w_0\sqrt{1+\frac{z^2}{z_0^2}}\qquad &&\text{is the beam radius,}\\  
  R(z) &= z\left(1+\frac{z_0^2}{z^2}\right)\qquad &&\text{is the wavefront radius,}\\
  \eta(z) &= \text{arctan}\left(\frac{z}{z_0}\right)\qquad &&\text{is the phase correction.}
\end{alignat}
Equation \ref{eq:gauss} uses the paraxial approximation, so it can only be used
for beams with a waist that is significantly larger than the reduced wavelength
($w_0 \gg \lambda/n$). Notice that under the paraxial approximation the beam is uniformly polarized in the transverse plane.

We would like to calculate the longitudinal component of a Gaussian beam when
the beam waist approaches the reduced wavelength ($w_0 > \lambda/n$). One
approach is to numerically evaluate the Richards-Wolf diffraction integral
\cite{richards, biobeam}. This approach is time consuming and too
accurate for our needs---we only want to model weak longitudinal
fields. Instead, I will follow Novotny et. al. \cite{nov} and use a longitudinal
correction to equation \ref{eq:gauss}.

As written, equation \ref{eq:gauss} doesn't satisfy Gauss' law
($\nabla \cdot \mb{E} = 0$), so we will add a longitudinal field to correct
it. If the input beam is polarized along the $\mh{x}$ axis
($\phi_{\text{pol}}=0$), then we can rearrange Gauss' law to relate the
longitudinal and transverse fields with
\begin{align}
  E_z(x, y, z) = - \int \left[\frac{\partial}{\partial x} E_x(x,y,z)\right]dz. 
\end{align}
Carnicer et al. \cite{carnicer} worked through this integral using the angular
spectrum representation and found that
\begin{align}
  E_z(x,y,z) = -i\frac{2x}{kw_0^2}E_x(x,y,z) = -i\frac{x}{z_0}E_x(x,y,z)\label{eq:longit}. 
\end{align}

Equation \ref{eq:longit} means that:
\begin{itemize}
\item There is no longitudinal polarization in the $\mh{y}-\mh{z}$ plane
  because $E_z(0,y,z) = 0$.
\item There are longitudinal fields lobes on both sides of the optical axis
  along the transverse polarization direction.
\item The longitudinal fields are $90^{\circ}$ out of phase with the transverse
  fields which means that the total field is elliptically polarized off axis.
\item The longitudinal field strength is proportional to
  $\lambda/w_0^2$---highly focused beams have the strongest longitudinal fields.
\end{itemize}

If the input is polarized along the $\mh{x}$ axis, the corrected 3D Jones vector is
\begin{align}
\mb{A}(x,y,z) = \mh{x}- i\frac{x}{z_0}\mh{z}.
\end{align}

If the input polarization is arbitrary then the corrected 3D Jones vector is
\begin{align}
  \mb{A}(x,y,z) = \cos\phi_{\text{pol}}\mh{x} + \sin\phi_{\text{pol}}\mh{y} - i\frac{x\cos\phi_{\text{pol}} + y\sin\phi_{\text{pol}}}{z_0}\mh{z}.
\end{align}

If the beam is scanned along the $\mh{y}$ axis with velecity $v$ then the time
dependent 3D Jones vector is
\begin{align}
  \mb{A}(x,y,z,t) = \cos\phi_{\text{pol}}\mh{x} + \sin\phi_{\text{pol}}\mh{y} - i\frac{x\cos\phi_{\text{pol}} + (y - vt)\sin\phi_{\text{pol}}}{z_0}\mh{z}\label{eq:scanned_j}
\end{align}
and the time dependent electric field is
\begin{align}
  \mb{E}(x,y,z,t) &= \mb{A}(x,y,z,t)\frac{w_0}{w(z)}\text{exp}\left\{{-\frac{(x^2+(y-vt)^2)}{w^2(z)}}+i\left[kz - \eta(z) + \frac{k(x^2+(y-vt)^2)}{2R(z)}\right]\right\}\label{eq:scanned_e}.
\end{align}
We will use equations \ref{eq:scanned_j} and \ref{eq:scanned_e} to calculate
the excitation efficiency of a single fluorophore. Notice that as the Rayleigh
range $z_0$ increases the longitudinal electric field decreases, so we can ignore the
longitudinal component when the beam is weakly focused.

\section{Scanned Beam Excitation Efficiency}
We define the excitation efficiency of a fluorophore as the fraction of 
incident power that excites the fluorophore. If a fluorophore with absorption
dipole moment $\bs{\mu}_{\text{abs}}$ is placed in a time-independent complex
electric field $\mb{E}$, then the excitation efficiency is given by
\begin{align}
  \eta_{\text{exc}}(x,y,z) = \frac{|\bs{\mu}_{\text{abs}} \cdot \mb{E}(x,y,z)|^2}{|\mb{E}(x,y,z) |^2}.
\end{align}
If the fluorophore is placed in the path of a scanned laser then the electric
field becomes time dependent. If the laser is scanned extremely fast we would
need to consider the coherence of the electric field, but we will only consider
slow scanning here to simplify the calculation. Specifically, we require that
$v \ll w_0/\tau_c$---the scan velocity must be much less than the beam width
divided by the coherence time. In this case the excitation efficiency is
\begin{align}
  \eta_{\text{exc}}(x,y,z) = \frac{\intinf|\bs{\mu}_{\text{abs}} \cdot \mb{E}(x,y,z,t)|^2dt}{\intinf|\mb{E}(x,y,z,t)|^2dt}\label{eq:excitation}
\end{align}
We plug equation \ref{eq:scanned} into equation \ref{eq:excitation}, express
$\bs{\mu}_{\text{abs}}$ in spherical coordinates
\begin{align}
  \bs{\mu}_{\text{abs}} = \cos\Phi\sin\Theta\mh{x} + \sin\Phi\sin\Theta\mh{y} + \cos\Theta\mh{z},
\end{align}
and evaluate the integrals (see Appendix for details) to write the
excitation efficiency as
\begin{align}
  \eta_{\text{exc}}(x,y,z) = \frac{\sin^2\Theta\cos^2(\Phi-\phi_{\text{pol}}) + \cos^2\Theta\frac{x^2\cos^2\phi_{\text{pol}} + \frac{1}{4}w^2(z)\sin^2\phi_{\text{pol}}}{z_0^2}}{1+\frac{x^2\cos^2\phi_{\text{pol}} + \frac{1}{4}w^2(z)\sin^2\phi_{\text{pol}}}{z_0^2}}.\label{eq:strong}
\end{align}
If the beam is weakly focused then we can ignore the longitudinal excitation and
the excitation efficiency simplifies to
\begin{align}
  \eta_{\text{exc}} = \sin^2\Theta\cos^2(\Phi-\phi_{\text{pol}}).
\end{align}

\section{Detection Efficiency}
If we detect fluorescence in wide-field mode with an orthogonal arm and no
polarizer then the detection efficiency is
\begin{align}
  \eta_{\text{det}} = 2A + 2B\sin^2\Theta'
\end{align}
where
\begin{align}
  A &= \frac{1}{4} - \frac{3}{8}\cos\alpha + \frac{1}{8}\cos^3\alpha,\\
  B &= \frac{3}{16}\cos\alpha - \frac{3}{16}\cos^3\alpha, 
\end{align}
$\alpha = \text{arcsin(NA}/n)$ is the detection cone half angle, and $\Theta'$
is the angle between the dipole axis and the detection optical axis. See the
2017-06-09 notes for the relationship between $\Theta'$ and $\Theta,\Phi$. See
\cite{fourkas} and the 2017-04-25 notes for the derivation of the detection
efficiencies and additional expression for detection arms that use a polarizer.

\section{Orientation Forward Model}
The detected intensity is proportional to the the product of the excitation and
detection efficiencies. Using a weakly focused excitation beam and an
unpolarized detection arm gives us the following model
\begin{align}
  I_{\phi_{\text{pol}}} &= I_{\text{tot}}\sin^2\Theta\cos^2(\Phi - \phi_{\text{pol}})(2A+2B\sin^2\Theta')\label{eq:forward}
\end{align}
where $I_{\text{tot}}$ is the intensity we would collect if we had an excitation and detection efficiency of 1.

\section{Discussion}
Equation \ref{eq:strong} shows that for strongly focused beams the excitation
efficiency is a function of position. This couples the orientation and location
of the fluorophore and complicates our reconstruction. For now we'll use only
weakly focused beams so that we can ignore the longitudinal component.

To ignore the longitudinal component, we require that the ratio of the beam
waist to the Rayleigh range be much less than 2 ($w/z_0 \ll 2$).  The excitation
beam in \cite{wu2013} satisfies this criteria
$w/z_0 = 1.2\ \mu\text{m}/9\ \mu\text{m} = 0.13$, but we'll need to be careful if
we want to use beams that are more strongly focused.

Under the weak focusing approximation the orientation and position of
fluorophores are decoupled. This will allow us to split the reconstruction into
two steps (1) estimate the position of the fluorophores using unpolarized frames
(or the sum of orthogonally polarized frames) with established reconstruction
techniques then (2) estimate the orientation or the fluorophores using polarized
frames and equation \ref{eq:forward}.

Note that we are working in a different regime than Agrawal et. al. \cite{agrawal}. They
are considering imaging systems with better resolution than ours, so the
position and orientation are coupled and must be estimated together. At our
resolution, the position and orientation are decoupled so we can estimate them
separately.

\section{References}
\setlength\biblabelsep{0.025\textwidth}
\printbibliography[heading=none]
\pagebreak

\section{Appendix}
We will evaluate the following integrals to find the excitation efficiency
\begin{align}
  \eta_{\text{exc}}(x,y,z) = \frac{\intinf|\bs{\mu}_{\text{abs}} \cdot \mb{E}(x,y,z,t)|^2dt}{\intinf|\mb{E}(x,y,z,t)|^2dt}\label{eq:excitation}
\end{align}
where
\begin{align}
  \bs{\mu}_{\text{abs}} &= \cos\Phi\sin\Theta\mh{x} + \sin\Phi\sin\Theta\mh{y} + \cos\Theta\mh{z}\\
  \mb{E}(x,y,z,t) &= \mb{A}(x,y,z,t)\frac{w_0}{w(z)}\text{exp}\left\{{-\frac{(x^2+(y-vt)^2)}{w^2(z)}}+i\left[kz - \eta(z) + \frac{k(x^2+(y-vt)^2)}{2R(z)}\right]\right\}\\
  \mb{A}(x,y,z,t) &= \cos\phi_{\text{pol}}\mh{x} + \sin\phi_{\text{pol}}\mh{y} - i\frac{x\cos\phi_{\text{pol}} + (y - vt)\sin\phi_{\text{pol}}}{z_0}\mh{z}.\label{eq:scanned_jones}
\end{align}

We'll need the following facts
\begin{align}
  \intinf e^{-ax^2}dx &= \sqrt{\frac{\pi}{a}}\\
  \intinf xe^{-ax^2}dx &= 0 \label{eq:odd}\\ 
  \intinf x^2e^{-ax^2}dx &= \frac{1}{2a}\sqrt{\frac{\pi}{a}}.
\end{align}

The numerator is
\begin{align}
  = &\intinf|\bs{\mu}_{\text{abs}} \cdot \mb{E}(x,y,z,t)|^2dt\\
  \begin{split}
  = &\intinf\Bigg |\left[\cos\Phi\sin\Theta\cos\phi_{\text{pol}} + \sin\Phi\sin\Theta\sin\phi_{\text{pol}} - i\cos\Theta\frac{x\cos\phi_{\text{pol}} + (y - vt)\sin\phi_{\text{pol}}}{z_0}\right]\\
    &\ \ \ \ \ \ \ \ \frac{w_0}{w(z)}\text{exp}\left\{{-\frac{(x^2+(y-vt)^2)}{w^2(z)}}+i\left[kz - \eta(z) + \frac{k(x^2+(y-vt)^2)}{2R(z)}\right]\right\}\Bigg |^2dt.
  \end{split}\\
  \intertext{After changing variables $y' = y - vt$ and moving constants outside the integral we get}
  = &\frac{w_0^2}{w^2(z)}e^{-\frac{2x^2}{w^2(z)}}\intinf\left|\left[\cos\Phi\sin\Theta\cos\phi_{\text{pol}} + \sin\Phi\sin\Theta\sin\phi_{\text{pol}} - i\cos\Theta\frac{x\cos\phi_{\text{pol}} + y'\sin\phi_{\text{pol}}}{z_0}\right]\right|^2e^{-\frac{2y'^2}{w^2(z)}}dy'.
 \intertext{After expanding the square brackets we get}
   = &\frac{w_0^2}{w^2(z)}e^{-\frac{2x^2}{w^2(z)}}\intinf \Bigg [(\cos\Phi\sin\Theta\cos\phi_{\text{pol}} + \sin\Phi\sin\Theta\sin\phi_{\text{pol}})^2 +\cos^2\Theta\frac{(x\cos\phi_{\text{pol}} +y'\sin\phi_{\text{pol}})^2}{z_0^2} \Bigg ]e^{-\frac{2y'^2}{w^2(z)}}dy'\\
   = &\frac{w_0^2}{w(z)}e^{-\frac{2x^2}{w^2(z)}}\sqrt{\frac{\pi}{2}}\Bigg [\sin^2\Theta\cos^2(\Phi - \phi_{\text{pol}})  + \cos^2\Theta\frac{x^2\cos^2\phi_{\text{pol}} + \frac{1}{4}w^2(z)\sin^2\phi_{\text{pol}}}{z_0^2}  \Bigg ].
\end{align}
The denominator is
\begin{align}
  = &\intinf|\mb{E}(x,y,z,t)|^2dt\\
  = &\frac{w_0^2}{w^2(z)}e^{-\frac{2x^2}{w^2(z)}}\intinf\left|\left[\cos\phi_{\text{pol}}\mh{x} + \sin\phi_{\text{pol}}\mh{y} - i\frac{x\cos\phi_{\text{pol}} + y'\sin\phi_{\text{pol}}}{z_0}\mh{z}\right]\right|^2e^{-\frac{2y'^2}{w^2(z)}}dy'\\
  = &\frac{w_0^2}{w^2(z)}e^{-\frac{2x^2}{w^2(z)}}\intinf\left[1 + \frac{(x\cos\phi_{\text{pol}} + y'\sin\phi_{\text{pol}})^2}{z^2_0}\right]e^{-\frac{2y'^2}{w^2(z)}}dy'\\
  = &\frac{w_0^2}{w(z)}e^{-\frac{2x^2}{w^2(z)}}\sqrt{\frac{\pi}{2}}\left[1 + \frac{x^2\cos^2\phi_{\text{pol}} + \frac{1}{4}w^2(z)\sin^2\phi_{\text{pol}}}{z^2_0}\right].
\end{align}
The final excitation efficiency is
\begin{align*}
\eta_{\text{exc}}(x,y,z) = \frac{\sin^2\Theta\cos^2(\Phi - \phi_{\text{\text{pol}}}) + \cos^2\Theta\frac{x^2\cos^2\phi_{\text{pol}} + \frac{1}{4}w^2(z)\sin^2\phi_{\text{pol}}}{z_0^2}}{1 + \frac{x^2\cos^2\phi_{\text{pol}} + \frac{1}{4}w^2(z)\sin^2\phi_{\text{pol}}}{z^2_0}}.
\end{align*}

\end{document}

### 2017-06-26 notes/2017-06-21-light-sheet-illumination/report/report.tex
Revised light sheet illumination notes with quantified error.

--- 2017-06-21 notes/2017-06-21-light-sheet-illumination/report.tex
@@
 \begin{align}
-  \eta_{\text{exc}}(x,y,z) = \frac{|\bs{\mu}_{\text{abs}} \cdot \mb{E}(x,y,z)|^2}{|\mb{E}(x,y,z) |^2}.
+  \eta_{\text{exc}} = \frac{|\bs{\mu}_{\text{abs}} \cdot \mb{E}(x,y,z)|^2}{|\mb{E}(x,y,z) |^2}.
 \end{align}
 If the fluorophore is placed in the path of a scanned laser then the electric
-field becomes time dependent. If the laser is scanned extremely fast we would
-need to consider the coherence of the electric field, but we will only consider
-slow scanning here to simplify the calculation. Specifically, we require that
+field becomes time dependent. If the laser is scanned quickly we would need to
+consider the coherence of the electric field, but we will only consider slow
+scanning here to simplify the calculation. Specifically, we require that
 $v \ll w_0/\tau_c$---the scan velocity must be much less than the beam width
@@
 \begin{align}
-  \eta_{\text{exc}}(x,y,z) = \frac{\intinf|\bs{\mu}_{\text{abs}} \cdot \mb{E}(x,y,z,t)|^2dt}{\intinf|\mb{E}(x,y,z,t)|^2dt}\label{eq:excitation}
+  \eta_{\text{exc}} = \frac{\intinf|\bs{\mu}_{\text{abs}} \cdot \mb{E}(x,y,z,t)|^2dt}{\intinf|\mb{E}(x,y,z,t)|^2dt}\label{eq:excitationx}
 \end{align}
-We plug equation \ref{eq:scanned} into equation \ref{eq:excitation}, express
+We plug equation \ref{eq:scanned_e} into equation \ref{eq:excitationx}, express
 $\bs{\mu}_{\text{abs}}$ in spherical coordinates
@@
 \begin{align}
-  \eta_{\text{exc}}(x,y,z) = \frac{\sin^2\Theta\cos^2(\Phi-\phi_{\text{pol}}) + \cos^2\Theta\frac{x^2\cos^2\phi_{\text{pol}} + \frac{1}{4}w^2(z)\sin^2\phi_{\text{pol}}}{z_0^2}}{1+\frac{x^2\cos^2\phi_{\text{pol}} + \frac{1}{4}w^2(z)\sin^2\phi_{\text{pol}}}{z_0^2}}.\label{eq:strong}
+  \eta_{\text{exc}} = \frac{\sin^2\Theta\cos^2(\Phi-\phi_{\text{pol}}) + \cos^2\Theta\frac{x^2\cos^2\phi_{\text{pol}} + \frac{1}{4}w^2(z)\sin^2\phi_{\text{pol}}}{z_0^2}}{1+\frac{x^2\cos^2\phi_{\text{pol}} + \frac{1}{4}w^2(z)\sin^2\phi_{\text{pol}}}{z_0^2}}.\label{eq:strong}
 \end{align}
@@
 \begin{align}
-  \eta_{\text{exc}} = \sin^2\Theta\cos^2(\Phi-\phi_{\text{pol}}).
+  \eta_{\text{exc}} \approx \sin^2\Theta\cos^2(\Phi-\phi_{\text{pol}}).\label{eq:weak_approx}
 \end{align}
+
+\textbf{How good is the approximation in equation \ref{eq:weak_approx}?}
+
+For longitudinal fluorophores equation \ref{eq:weak_approx} is a very bad
+approximation---it predicts that a longitudinal fluorophore will not be excited
+at all. We can't use a percentage error because the approximation completely
+ignores the excitation of longitudinal dipoles.
+
+Instead, we can compare the size of signals from longitudinal and transverse
+fluorophores. If the signal from longitudinal fluorophores is less than the
+signal from noise and background, then we can ignore the signal from
+longitudinal fluorophores.
+\begin{align}
+  \text{Excitation Ratio} &= \frac{\text{Max Longitudinal Excitation}}{\text{Max Transverse Excitation}}\\
+                          &= \frac{\eta_{\text{exc}}(\Theta = 0)}{\eta_{\text{exc}}(\Theta = \pi/2, \Phi = \phi_{\text{pol}})}\\
+                          &= \frac{x^2\cos^2\phi_{\text{pol}} + \frac{1}{4}w^2(z)\sin^2\phi_{\text{pol}}}{z_0^2}\label{eq:ratio}
+\end{align}
+
+At first glance equation \ref{eq:ratio} looks bleak---the excitation ratio grows
+without bound in the $x$ direction which means that the longitudinal excitation
+becomes a larger fraction of the total excitation as we move away from the plane
+of the light sheet. Fortunately, we only care about regions of the beam with a
+high intensity, so we look at the intensity weighted excitation ratio instead.
+\begin{align*}
+    \text{Intensity Weighted Excitation Ratio} &= \frac{w_0}{w(z)}e^{-\frac{2x^2}{w^2(z)}}\frac{x^2\cos^2\phi_{\text{pol}} + \frac{1}{4}w^2(z)\sin^2\phi_{\text{pol}}}{z_0^2}\label{eq:intratio} \\
+\end{align*}
+We can interpret the intensity weighted excitation ratio as the fraction of the
+maximum signal (created by a transverse fluorophore at the origin) that we
+ignore by ignoring longitudinal excitation. Figure \ref{fig:error} shows that
+the intensity weighted excitation ratio is <2\% for the imaging parameters used
+in Wu et. al. \cite{wu2013}.
+
+\fig{figures/error.pdf}{1.0}{Intensity weighted excitation ratio as a
+  function of position using the parameters in Wu et. al. \cite{wu2013}:
+  $w_0 = 1.2\ \mu$m, $z_0 = 9\ \mu$m, $\lambda = 488$ nm, FOV$\approx$ 80$\times$80
+  $\mu$m${}^2$. The maximum values are at the edge of the FOV and are <2\%.}{error}
+
+As a rough heuristic, if the fraction of the signal from noise and background is
+greater than the intensity weighted excitation ratio we can justifiably ignore
+the longitudinal component.
+
+Figure 2 shows the maximum intensity weighted excitation ratio for an
+80$\times$80 $\mu$m${}^2$ FOV as a function of $w_0$. The maximum intensity
+weighted excitation ratio decreases as $w_0$ increases because wider beams
+have smaller longitudinal components.
+
+\fig{figures/max-error.pdf}{0.5}{Maximum intensity weighted excitation ration (see equation \ref{eq:intratio}) for an 80$\times$80 $\mu$m${}^2$ FOV with $\lambda = 488$ nm.}{max-error}
 
@@
 
-To ignore the longitudinal component, we require that the ratio of the beam
-waist to the Rayleigh range be much less than 2 ($w/z_0 \ll 2$).  The excitation
-beam in \cite{wu2013} satisfies this criteria
-$w/z_0 = 1.2\ \mu\text{m}/9\ \mu\text{m} = 0.13$, but we'll need to be careful if
-we want to use beams that are more strongly focused.
+To ignore the longitudinal component, we require that the fraction of the signal
+from noise and background is greater than the intensity weighted excitation
+ratio (<2\% with current imaging parameters). We'll need to be careful about
+longitudinal excitation if we want to use beams that are more strongly focused.
 

### 2017-06-26 examples/detection.py

@@ -5,7 +5,7 @@ import matplotlib.pyplot as plt
 import os; import time; start = time.time(); print('Running...')
 
 # Main input parameters
-n_pts = 2500 # 50000
+n_pts = 250 # 50000
 pols = [np.array([-1,1,0]), np.array([1,1,0]), np.array([0,1,0]),np.array([1,0,0])]
 ill_axes = [np.array([0, 0, 1]), np.array([0, 0, 1]), np.array([0, 0, 1]), np.array([0, 0, 1])]
 det_axes = [np.array([0, 0, 1]), np.array([0, 0, 1]), np.array([-1.0/np.sqrt(2), 0, 1.0/np.sqrt(2)]), np.array([-1.0/np.sqrt(2), 0, 1.0/np.sqrt(2)])]
@@ -14,8 +14,6 @@ n_cols = len(det_axes)
 inch_fig = 5
 vis_px = 2000
 dpi = 250
-#row_labels = [r'$x$-polarized, epi-detection', r'$y$-polarized, 45-detection', r'$x$-polarized, ortho-detection', r'$y$-polarized, 135-detection', ]
-#col_labels = [r'$0$-polarized, 45-detection', r'$45$-polarized, 45-detection']#, r'$90$-polarized, 45-detection', r'$135$-polarized, 45-detection']
 col_labels = 4*['']
 row_labels = ['Scene', r'Excitation Efficiency $\eta_{\text{exc}}$', r'Detection Efficiency $\eta_{\text{det}}$', r'Total Efficiency $\eta_{\text{tot}}$']
 
@@ -33,14 +31,14 @@ for i, (pol, det_axis, ill_axis) in enumerate(zip(pols, det_axes, ill_axes)):
     print('Computing microscope: ' + str(i))
 
     # Create microscope
-    det = detector.Detector(optical_axis=np.array(det_axis),
-                            na=0.8,
-                            n=1.5)
-    
     ill = illuminator.Illuminator(illum_type='kohler',
                                   optical_axis=ill_axis,
                                   na=0.8,
                                   bfp_pol_dir=pol)
+    
+    det = detector.Detector(optical_axis=np.array(det_axis),
+                            na=0.8,
+                            n=1.5)
                                   
     m = microscope.Microscope(illuminator=ill, detector=det, max_photons=1)
     

### 2017-06-26 notes/2017-06-21-light-sheet-illumination/figures/plot-error.py

@@ -1,0 +1,74 @@
+from dipsim import util
+import numpy as np
+import matplotlib.pyplot as plt
+from functools import partial
+
+import matplotlib
+matplotlib.rcParams.update({'font.size': 14})
+
+# All lengths in um
+
+# Define functions
+def w(z, z0, w0):
+    return w0*np.sqrt(1+(z/z0)**2)
+
+def z_excite(x, z, w0, phi_pol, lamb=0.488):
+    k = 2*np.pi/lamb
+    z0 = k*(w0**2)/2.0
+    a = w0/w(z, z0, w0)
+    b = np.exp(-(2*x**2)/(w(z, z0, w0)**2))
+    c = ((x*np.cos(phi_pol))**2 + (0.5*w(z, z0, w0)*np.sin(phi_pol))**2)/(z0**2)
+    return a*b*c
+
+# Setup x-z plots
+n_rows = 1
+n_cols = 2
+inch_fig = 5
+size = (inch_fig*n_cols, inch_fig*n_rows)
+fig, axs = plt.subplots(n_rows, n_cols, figsize=size)
+plt.subplots_adjust(wspace=0.5, hspace=0)
+caxs = util.generate_caxs(axs)
+
+x = np.arange(-40, 40, 0.05)
+z = np.arange(-40, 40, 0.05)
+X, Z = np.meshgrid(x, z)
+
+phi_pols = [0, np.pi/2]
+for i, phi_pol in enumerate(phi_pols):
+    I = z_excite(X, Z, w0=1.2, phi_pol=phi_pol, lamb=0.488)
+    extent = (x[0], x[-1], z[0], z[-1])
+    cmap = axs[i].imshow(I, interpolation='none', cmap='Greys', extent=extent)
+    fig.colorbar(cmap, cax=caxs[i], orientation='vertical')
+
+    axs[i].set_xlabel('$x$ ($\mu$m)')
+    axs[i].set_ylabel('$z$ ($\mu$m)')
+    if i == 0:
+        axs[i].set_title('$\phi_{\mathrm{pol}} = 0$', y=1.05)
+    if i == 1:
+        axs[i].set_title('$\phi_{\mathrm{pol}} = \pi/2$', y=1.05)
+
+fig.savefig('error.pdf')
+
+# Setup max error plots
+def find_max_error(waist, FOV):
+    x = np.arange(-(FOV/2.0), (FOV/2.0), 0.1)
[…]

### 2017-06-26 notes/2017-06-21-light-sheet-illumination/report/report.tex
Patrick's June 26 edits.

@@
 \begin{document}
-\title{\vspace{-2.5em} Excitation and Detection Efficiency Under Scanned Light
-  Sheet Illumination \vspace{-2.5em}} \author{Talon Chandler}
-\date{\vspace{-1em}\today\vspace{-1em}}
+\title{\vspace{-2.5em} Excitation and Detection Efficiency Under Light-Sheet
+  Illumination \vspace{-1.0em}} \author{Talon Chandler}
+\date{\vspace{-1em}June 21, 2017\\ (Updated: June 26, 2017)\vspace{-1em}}
 \maketitle
@@
 In these notes I will calculate the excitation and detection efficiency of a
-single fluorophore under polarized scanned light sheet illumination. I will
+single fluorophore under polarized scanned light-sheet illumination. I will
 start by finding the transverse and longitudinal fields at all positions in a
@@
 \begin{alignat}{2}  
-  \mb{A} &= \cos\phi_{\text{pol}}\mh{x} + \sin\phi_{\text{pol}}\mh{y}\qquad &&\text{is the input Jones vector},\\
+  \mb{A} &= \cos\phi_{\text{pol}}\mh{x} + \sin\phi_{\text{pol}}\mh{y}\qquad &&\text{is the input Jones vector in 3D},\\
   w_0 &\approx \frac{2n}{k(\text{NA})}\qquad &&\text{is the beam waist radius,}\\
@@
   because $E_z(0,y,z) = 0$.
-\item There are longitudinal fields lobes on both sides of the optical axis
+\item There are longitudinal field lobes on both sides of the optical axis
   along the transverse polarization direction.
-\item The longitudinal fields are $90^{\circ}$ out of phase with the transverse
-  fields which means that the total field is elliptically polarized off axis.
+\item The factor of $i$ means that the longitudinal fields are $90^{\circ}$ out
+  of phase with the transverse fields which means that the total field is
+  elliptically polarized off axis.
 \item The longitudinal field strength is proportional to
@@
 
-If the beam is scanned along the $\mh{y}$ axis with velecity $v$ then the time
+If the beam is scanned along the $\mh{y}$ axis with velocity $v$ then the time
 dependent 3D Jones vector is
@@
 \end{align}
-and the time dependent electric field is
+and the time-dependent electric field is
 \begin{align}
@@
 becomes a larger fraction of the total excitation as we move away from the plane
-of the light sheet. Fortunately, we only care about regions of the beam with a
-high intensity, so we look at the intensity weighted excitation ratio instead.
+of the light-sheet. Fortunately, we only care about regions of the beam with a
+high intensity, so we look at the intensity-weighted excitation ratio instead.
 \begin{align*}
-    \text{Intensity Weighted Excitation Ratio} &= \frac{w_0}{w(z)}e^{-\frac{2x^2}{w^2(z)}}\frac{x^2\cos^2\phi_{\text{pol}} + \frac{1}{4}w^2(z)\sin^2\phi_{\text{pol}}}{z_0^2}\label{eq:intratio} \\
+    \text{Intensity-Weighted Excitation Ratio} &= \frac{w_0}{w(z)}e^{-\frac{2x^2}{w^2(z)}}\frac{x^2\cos^2\phi_{\text{pol}} + \frac{1}{4}w^2(z)\sin^2\phi_{\text{pol}}}{z_0^2}\label{eq:intratio} \\
 \end{align*}
-We can interpret the intensity weighted excitation ratio as the fraction of the
+We can interpret the intensity-weighted excitation ratio as the fraction of the
 maximum signal (created by a transverse fluorophore at the origin) that we
 ignore by ignoring longitudinal excitation. Figure \ref{fig:error} shows that
-the intensity weighted excitation ratio is <2\% for the imaging parameters used
+the intensity weighted excitation ratio is \textless 2\% for the imaging parameters used
 in Wu et. al. \cite{wu2013}.
 
-\fig{figures/error.pdf}{1.0}{Intensity weighted excitation ratio as a
+\fig{../figures/error.pdf}{1.0}{Intensity-weighted excitation ratio as a
   function of position using the parameters in Wu et. al. \cite{wu2013}:
   $w_0 = 1.2\ \mu$m, $z_0 = 9\ \mu$m, $\lambda = 488$ nm, FOV$\approx$ 80$\times$80
-  $\mu$m${}^2$. The maximum values are at the edge of the FOV and are <2\%.}{error}
+  $\mu$m${}^2$. The maximum values are at the edge of the FOV and are \textless 2\%.}{error}
 
 As a rough heuristic, if the fraction of the signal from noise and background is
-greater than the intensity weighted excitation ratio we can justifiably ignore
+greater than the intensity-weighted excitation ratio we can justifiably ignore
 the longitudinal component.
 
-Figure 2 shows the maximum intensity weighted excitation ratio for an
-80$\times$80 $\mu$m${}^2$ FOV as a function of $w_0$. The maximum intensity
-weighted excitation ratio decreases as $w_0$ increases because wider beams
-have smaller longitudinal components.
+Figure 2 shows the maximum intensity-weighted excitation ratio for an
+80$\times$80 $\mu$m${}^2$ FOV as a function of $w_0$. The maximum
+intensity-weighted excitation ratio decreases as $w_0$ increases because wider
+beams have smaller longitudinal components.
 
-\fig{figures/max-error.pdf}{0.5}{Maximum intensity weighted excitation ration (see equation \ref{eq:intratio}) for an 80$\times$80 $\mu$m${}^2$ FOV with $\lambda = 488$ nm.}{max-error}
+\fig{../figures/max-error.pdf}{0.5}{Maximum intensity-weighted excitation ration (see equation \ref{eq:intratio}) for an 80$\times$80 $\mu$m${}^2$ FOV with $\lambda = 488$ nm.}{max-error}
 
@@
 from noise and background is greater than the intensity weighted excitation
-ratio (<2\% with current imaging parameters). We'll need to be careful about
+ratio (\textless 2\% with current imaging parameters). We'll need to be careful about
 longitudinal excitation if we want to use beams that are more strongly focused.
 
-Under the weak focusing approximation the orientation and position of
+Under the weak-focusing approximation the orientation and position of
 fluorophores are decoupled. This will allow us to split the reconstruction into

### 2017-06-26 notes/2017-06-21-light-sheet-illumination/figures/plot-error.py

@@ -70,5 +70,14 @@ for FOV in FOVs:
 ax.set_yscale('log')
 ax.set_xlim([1,5])
 ax.set_xlabel('$w_0$ ($\mu$m)')
-ax.set_ylabel('Maximum Intensity Weighted Excitation Ratio')
+ax.set_ylabel('Maximum Intensity-Weighted Excitation Ratio')
+
+from matplotlib.ticker import FuncFormatter
+def log_10_product(x, pos):
+    return str(100*x) + '\%'
+formatter = FuncFormatter(log_10_product)
+ax.set_yticks([0.001, 0.01, 0.04])
+ax.yaxis.set_major_formatter(formatter)
+
+
 fig.savefig('max-error.pdf')

### 2017-06-27 dipsim/detector.py
Use spherical coordinates everywhere.

@@ -1,12 +1,13 @@
 import numpy as np
+import dipsim.util as util
 
 class Detector:
     """A detection path is specified by its optical axis, numerical aperture, and
     index of refraction of the medium.
 
     """
-    def __init__(self, optical_axis, det_type='lens', na=0.8, n=1.33):
-        self.optical_axis = optical_axis
+    def __init__(self, theta_optical_axis, det_type='lens', na=0.8, n=1.33):
+        self.theta_optical_axis = theta_optical_axis
         self.det_type = det_type
         self.na = na # Numerical aperture
         self.n = n
@@ -18,7 +19,7 @@ class Detector:
         if self.det_type=='lens':
             A = (1.0/4.0) - (3.0/8.0)*np.cos(self.alpha) + (1.0/8.0)*(np.cos(self.alpha)**3)
             B = (3.0/16.0)*np.cos(self.alpha) - (3.0/16.0)*(np.cos(self.alpha)**3)
-            theta = np.arccos(np.dot(fluorophore.mu_em, self.optical_axis))
+            theta = util.theta_prime(fluorophore.mu_em[0], fluorophore.mu_em[1], self.theta_optical_axis)
             return A + B*(np.sin(theta)**2)
         elif self.det_type=='4pi':
             return 1.0

### 2017-06-27 dipsim/illuminator.py

@@ -10,48 +10,21 @@ class Illuminator:
     rotated. Be careful with oblique illumination. ***
 
     """
-    def __init__(self, illum_type, optical_axis, na=0.8, n=1.33,
-                 bfp_pol_dir=None, bfp_apod=None):
-        
+    def __init__(self, illum_type, theta_optical_axis, na=0.8, n=1.33, phi_pol=0):
         self.illum_type = illum_type
-        
-        if np.linalg.norm(optical_axis) - 1.0 > 1e-3:
-            print("Warning: optical axis is not a unit vector. Normalizing.")
-        self.optical_axis = optical_axis/np.linalg.norm(optical_axis)
-
+        self.theta_optical_axis = theta_optical_axis
         self.na = na
         self.n = n
         self.alpha = np.arcsin(self.na/self.n)
         self.f = 10 # Arbitrary
         self.bfp_rad = self.f*np.tan(self.alpha)
-
-        if np.dot(bfp_pol_dir, np.array([0, 0, 1])) != 0:
-            print("Warning: polarization must be specified in x-y plane.")
-        elif np.linalg.norm(bfp_pol_dir) - 1 >= 1e-10:
-            print("Warning: bfp_pol_dir is not a unit vector. Normalizing.")
-        self.bfp_pol_dir = bfp_pol_dir/np.linalg.norm(bfp_pol_dir)
-
-        # Rotate polarization state so that it is perp to optical axis
-        self.bfp_pol = np.dot(util.rot_map(self.optical_axis), self.bfp_pol_dir)
-
-        # bfp_apod is an apodization function. If no argument is supplied, there
-        # is a sharp cutoff at bfp_rad.
-        if bfp_apod == None:
-            def default_apod(r):
-                if r <= self.bfp_rad:
-                    return 1.0
-                else:
-                    return 0
-            self.bfp_apod = default_apod
-        else:
-            self.bfp_apod = bfp_apod
+        self.phi_pol = phi_pol
 
     def calc_excitation_efficiency(self, fluorophore):
         A = (1.0/4.0) - (3.0/8.0)*np.cos(self.alpha) + (1.0/8.0)*(np.cos(self.alpha)**3)
[…]

### 2017-06-27 dipsim/microscope.py

@@ -24,8 +24,7 @@ class Microscope:
         return self.max_photons*self.calc_sensitivity(args)
     
     def calc_sensitivity(self, args):
-        flu_dir = util.tp2xyz(args)
-        flu = fluorophore.Fluorophore(mu_abs=flu_dir, mu_em=flu_dir)
+        flu = fluorophore.Fluorophore(mu_abs=args, mu_em=args)
         excite = self.illuminator.calc_excitation_efficiency(flu)
         collect = self.detector.calc_collection_efficiency(flu)        
         return excite*collect
@@ -38,8 +37,7 @@ class Microscope:
         util.plot_sphere(filename, directions=directions, data=I, **kwargs)
 
     def calc_excitation_efficiency(self, args):
-        flu_dir = util.tp2xyz(args)
-        flu = fluorophore.Fluorophore(mu_abs=flu_dir, mu_em=flu_dir)        
+        flu = fluorophore.Fluorophore(mu_abs=args, mu_em=args)
         I = self.illuminator.calc_excitation_efficiency(flu)
         return I
     
@@ -52,8 +50,7 @@ class Microscope:
         util.plot_sphere(filename, directions=directions, data=I, **kwargs)
 
     def calc_collection_efficiency(self, args):
-        flu_dir = util.tp2xyz(args)
-        flu = fluorophore.Fluorophore(mu_abs=flu_dir, mu_em=flu_dir)        
+        flu = fluorophore.Fluorophore(mu_abs=args, mu_em=args)        
         I = self.detector.calc_collection_efficiency(flu)
         return I
     
@@ -81,16 +78,16 @@ class Microscope:
         view = canvas.central_widget.add_view(camera=my_cam)
 
         if dual_arm:
-            det_axes = [self.illuminator.optical_axis, self.detector.optical_axis]
-            ill_axes = [self.detector.optical_axis, self.illuminator.optical_axis]
+            det_axes = [self.illuminator.theta_optical_axis, self.detector.theta_optical_axis]
+            ill_axes = [self.detector.theta_optical_axis, self.illuminator.theta_optical_axis]
         else:
-            det_axes = [self.detector.optical_axis]
-            ill_axes = [self.illuminator.optical_axis]
+            det_axes = [self.detector.theta_optical_axis]
+            ill_axes = [self.illuminator.theta_optical_axis]
 
         for idx, (det_axis, ill_axis) in enumerate(zip(det_axes, ill_axes)):
[…]

### 2017-06-27 dipsim/util.py

@@ -69,6 +69,19 @@ def fibonacci_sphere(n):
     phi = np.mod((np.pi*(3.0 - np.sqrt(5.0)))*np.arange(n), 2*np.pi) - np.pi
     return np.vstack((theta, phi)).T
 
+def theta_prime(theta, phi, psi):
+    return np.arccos(np.sin(psi)*np.cos(phi)*np.sin(theta) + np.cos(psi)*np.cos(theta))
+
+def phi_prime(theta, phi, psi):
+    num = np.cos(psi)*np.cos(phi)*np.sin(theta) - np.sin(psi)*np.cos(theta)
+    den = np.sqrt(1 - (np.sin(psi)*np.cos(phi)*np.sin(theta) + np.cos(psi)*np.cos(theta))**2)
+    if phi < np.pi and phi > 0:
+        return np.arccos(num/den)
+    elif phi < 0 and phi > -np.pi:
+        return -np.arccos(num/den)
+    else:
+        return np.nan
+    
 # Three coordinate conversion functions. Use R to use theta-phi coordinates that
 # are measured from axes other than the typical z and x axes.
 def tp2xyz(tp, R=np.eye(3,3)):

### 2017-06-27 notes/2017-06-09-rotated-frames/report/report.tex
Fixed incorrect reference.
Correct error in spherical coordinates equations. Fix phi cases.

@@
 \title{\vspace{-2.5em} Relating Spherical Coordinates In Rotated Frames\vspace{-1em}} \author{Talon Chandler}% and Patrick La Rivi\`ere}
-\date{\vspace{-1em}June 9, 2017\vspace{-1em}}
+\date{\vspace{-1em}June 9, 2017\\ (Updated: June 27, 2017)\vspace{-1em}}
 \maketitle
@@
   \theta' &= \arccos\left(\sin\psi\cos\phi\sin\theta + \cos\psi\cos\theta\right)\label{eq:thetap}\\
-  \phi' &= \arccos\left(\frac{\cos\psi\cos\phi\sin\phi - \sin\psi\cos\theta}{\sqrt{1 - (\sin\psi\cos\phi\sin\theta + \cos\psi\cos\theta)^2}}\right).\label{eq:phip}
+  \phi' &= \arccos\left(\frac{\cos\psi\cos\phi\sin\theta - \sin\psi\cos\theta}{\sqrt{1 - (\sin\psi\cos\phi\sin\theta + \cos\psi\cos\theta)^2}}\right).\label{eq:phip}
 \end{align}\label{eq:solution}% 
@@
 Equation \ref{eq:thetap} follows directly from equation \ref{eq:prevc}. Equation
-\ref{eq:phip} follows from plugging equation \ref{eq:thetap} into equation \ref{eq:prevc} and using the identity $\sin(\arccos(x)) = \sqrt{1 - x^2}$.
+\ref{eq:phip} follows from plugging equation \ref{eq:thetap} into equation \ref{eq:preva} and using the identity $\sin(\arccos(x)) = \sqrt{1 - x^2}$.
 
@@
   \theta &= \arccos\left(-\sin\psi\cos\phi'\sin\theta' + \cos\psi\cos\theta'\right)\\
-  \phi &= \arccos\left(\frac{\cos\psi\cos\phi'\sin\phi' + \sin\psi\cos\theta'}{\sqrt{1 - (\sin\psi\cos\phi'\sin\theta' + \cos\psi\cos\theta')^2}}\right).
+  \phi &= \arccos\left(\frac{\cos\psi\cos\phi'\sin\theta' + \sin\psi\cos\theta'}{\sqrt{1 - (\sin\psi\cos\phi'\sin\theta' + \cos\psi\cos\theta')^2}}\right).
 \end{align}\label{eq:solutionp}% 
 \end{subequations}
+
+\textbf{Note:} The arccos in equations \ref{eq:solution} and \ref{eq:solutionp}
+depends the range of $\phi$. If $\phi \in [-\pi, \pi]$ then use
+\begin{align}
+  \phi' &=
+          \begin{cases}
+            \arccos\left(\frac{\cos\psi\cos\phi\sin\theta - \sin\psi\cos\theta}{\sqrt{1 - (\sin\psi\cos\phi\sin\theta + \cos\psi\cos\theta)^2}}\right) \ \ \ &0 \leq \phi < \pi  \\
+            -\arccos\left(\frac{\cos\psi\cos\phi\sin\theta - \sin\psi\cos\theta}{\sqrt{1 - (\sin\psi\cos\phi\sin\theta + \cos\psi\cos\theta)^2}}\right) \ \ \ &-\pi \leq \phi < 0\\
+          \end{cases}
+\end{align}
 

### 2017-07-06 paper/Express_style.tex
Added paper template and moved examples into figures folder.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                File: express_style.tex              %
%             Created: 2 September 2009               %
%              Updated: 14 June 2017                  %
%                                                     %
%           LaTeX template file for use with          %
%           OSA's journals Optics Express,            %
%             Biomedical Optics Express,              %
%            and Optical Materials Express            %
%                                                     %
%  send comments to Theresa Miller, tmiller@osa.org   %
%                                                     %
%       (c) 2017 Optical Society of America           %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[10pt]{article}
%% Specify the Express journal you are submitting to
%\usepackage[OME]{express}
\usepackage[OE]{express}
%\usepackage[BOE]{express}

\begin{document}

\title{Template and style guide for authors submitting to OSA Express Journals}

\author{Author One,\authormark{1} Author Two,\authormark{2,*} and Author Three\authormark{2,3}}

\address{\authormark{1}Peer Review, Publications Department, The Optical Society, 2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
\authormark{2}Publications Department, The Optical Society, 2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
\authormark{3}Currently with the Department of Electronic Journals, The Optical Society, 2010 Massachusetts Avenue NW, Washington, DC 20036, USA}

\email{\authormark{*}opex@osa.org} %% email address is required

% \homepage{http:...} %% author's URL, if desired

%%%%%%%%%%%%%%%%%%% abstract and OCIS codes %%%%%%%%%%%%%%%%
%% [use \begin{abstract*}...\end{abstract*} if exempt from copyright]

\begin{abstract}
Updated 14 June 2017. Detailed instructions and formatting guidelines for preparing \textit{Optics Express}, \textit{Biomedical Optics Express}, and \textit{Optical Materials Express} manuscripts in \LaTeX. For a simple outline and template, open the simple template file \texttt{Express\_temp.tex}. The Express journal simple and extended templates are also available on \href{http://www.overleaf.com/gallery/tagged/osa}{Overleaf}. OSA encourages the use of this free online collaborative tool for writing your OSA article.
\end{abstract}

\ocis{(000.0000) General; (000.2700) General science.} % REPLACE WITH CORRECT OCIS CODES FOR YOUR ARTICLE, MINIMUM OF TWO; Avoid using the OCIS codes for “General” or “General science” whenever possible.
%For a complete list of OCIS codes, visit: https://www.osapublishing.org/oe/submit/ocis/

%%%%%%%%%%%%%%%%%%%%%%% References %%%%%%%%%%%%%%%%%%%%%%%%%
\begin{thebibliography}{99}
\bibitem{bib1}P. J. Harshman, T. K. Gustafson, and P. Kelley, ``Title of paper," J. Chem. Phys. {\bfseries 3}, (to be published).

\bibitem{gallo99} K. Gallo and G. Assanto, ``All-optical diode based on second-harmonic generation in an asymmetric waveguide,'' \josab {\bfseries 16}(2), 267--269 (1999).

\bibitem{Masters98a} B. R. Masters, ``Three-dimensional microscopic tomographic imagings of the cataract in a human lens in vivo,'' \opex {\bfseries 3}(9), 332--338 (1998).

\bibitem{Oron03} D. Yelin,  D. Oron,  S. Thiberge,  E. Moses, and Y. Silberberg, ``Multiphoton plasmon-resonance microscopy,'' \opex {\bfseries 11}(12), 1385--1391 (2003).

\bibitem{samplefig}
B.~N.~Behnken, G.~Karunasiri, D.~R.~Chamberlin, P.~R.~Robrish, and J.~Faist,
``Real-time imaging using a 2.8~THz quantum cascade laser and uncooled infrared microbolometer camera,'' \ol \textbf{33}(5), 440--442 (2008).

\end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%  body  %%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Adherence to the specifications listed in this template is essential for efficient review and publication of submissions. Since OSA does not routinely perform copyediting and typesetting for this journal, use of the template is critical to providing a consistent appearance. Proper reference format is especially important (see Section \ref{sec:refs}).

\section{\texttt{express.sty} and required \LaTeX{} packages}
Page layout is set with the \texttt{geometry} package for US Letter paper. \texttt{express.sty} uses the following package files:

\begin{itemize}
\item \texttt{geometry} \ (page layout)
\item \texttt{color, graphicx} \ (replaces \texttt{graphics}; has preset options)
\item \texttt{mathptmx, courier, helvet} \ (Times, Courier, and Helvetica fonts)
\end{itemize}

The latest versions of these standard package files can be obtained at CTAN: the Comprehensive TeX Archive Network, \url{http://www.ctan.org}.

The command \verb+\usepackage{ae}+ can be invoked to revert font to Computer Modern, although we prefer to publish with Times (with \texttt{mathptmx.sty}) for consistency.

\section{Multiple corresponding authors}

There are two options for indicating multiple corresponding authorship, and they are formatted quite differently. The first format would be as follows, still using the asterisk to denote one of the authors:

\begin{verbatim}
\author{Author One\authormark{1,3} and Author Two\authormark{2,4}}

\address{\authormark{1}Peer Review, Publications Department,
Optical Society of America, 2010 Massachusetts Avenue NW,
Washington, DC 20036, USA\\
\authormark{2}Publications Department, Optical Society of America,
2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
\authormark{3}xyz@osa.org}

\email{\authormark{*}opex@osa.org}
\end{verbatim}

This format will generate the following appearance:\\

\author{Author One\authormark{1,3} and Author Two\authormark{2,4}}

\address{\authormark{1}Peer Review, Publications Department,
Optical Society of America, 2010 Massachusetts Avenue NW,
Washington, DC 20036, USA\\
\authormark{2}Publications Department, Optical Society of America,
2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
\authormark{3}xyz@osa.org}

\email{\authormark{*}opex@osa.org}
\medskip
The second format forgoes the asterisk and sets all email addresses equally within the affiliations. Please note that this format does not use the \verb+\email{}+ field at all.
\begin{verbatim}
\author{Author One\authormark{1,3} and Author Two\authormark{2,4}}

\address{\authormark{1}Peer Review, Publications Department,
Optical Society of America, 2010 Massachusetts Avenue NW,
Washington, DC 20036, USA\\
\authormark{2}Publications Department, Optical Society of America,
2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
\authormark{3}xyz@osa.org\\
\authormark{4}opex@osa.org}
\end{verbatim}

This format will generate the following appearance:\\

\author{Author One\authormark{1,3} and Author Two\authormark{2,4}}

\address{\authormark{1}Peer Review, Publications Department,
Optical Society of America, 2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
\authormark{2}Publications Department, Optical Society of America, 2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
\authormark{3}xyz@osa.org\\
\authormark{4}opex@osa.org}
\medskip
These are the preferred express journal formats for multiple corresponding authorship, and either may be used.

\section{Abstract}
The abstract should be limited to approximately 100 words. It should be an explicit summary of the paper that states the problem, the methods used, and the major results and conclusions. It also should contain the relevant key words that would allow it to be found in a cursory computerized search. If the work of another author is cited in the abstract, that citation should be written out without a number, (e.g., journal, volume, first page, and year in square brackets [Opt. Express {\bfseries 22}, 1234 (2014)]), and a separate citation should be included in the body of the text. The first reference cited in the main text must be [1]. Do not include numbers, bullets, or lists inside the abstract.

\section{Figures, tables, and supplemental materials}

\subsection{Figures and tables}

OSA express journals encourage authors to submit color figures with their manuscripts. Figures and tables should be placed in the body of the manuscript.

\begin{figure}[ht!]
\centering\includegraphics[width=7cm]{expressfig1}
\caption{Sample caption (Ref. \cite{Oron03}, Fig. 2).}
\end{figure}

\noindent Standard \LaTeX{} environments should be used to place tables and figures:
\begin{verbatim}
\begin{figure}[htbp]
\centering\includegraphics[width=7cm]{expressfig1}
\caption{Sample caption (\cite{Oron03}, Fig. 2).}
\end{figure}
\end{verbatim}

\subsection{Supplementary materials in OSA express journals}

Most OSA journals allow authors to include supplementary materials as integral parts of a manuscript. Such materials are subject to the same editorial standards and peer review procedures along with the rest of the paper and should be uploaded and described using OSA's Prism manuscript system.

Authors can submit appropriate visualizations or small data files (see details below) for OSA to host. Large datasets and code or simulation files can be included but must be placed in an appropriate archival repository and cited as described here.

\begin{table}[ht!]
\centering
\caption{Supplementary Materials Supported in OSA Express Journals}
\begin{tabular}{|l|l|}
\hline
Visualization & 2D image, 3D image, video \\ \hline
Data File     & Small data file such as data underlying a plot in a figure \\ \hline
Dataset       & Dataset stored in an appropriate external repository \\ \hline
Code          & Code or simulation files stored in an appropriate external repository \\
\hline
\end{tabular}
\end{table}

Video visualizations are the most commonly submitted type of supplementary materials for the express journals. They typically illustrate a synopsis of research results. They are integral and as such should be included only when they convey essential information beyond what can be presented within the article's PDF representation. Video visualizations should be uploaded upon submission and peer-reviewed along with the manuscript. Video files must use open compression standards for display on broadly available applications such as VLC or Windows Media Player. MOV, AVI, MPG, and MP4 video containers are accepted. The following video guidelines will help with the submission process:

\begin{itemize}
\item 15 MB is the recommended maximum video file size.
\item 720 x 480 pixels (width by height) is the recommended screen size.
\item If appropriate, insert a representative frame from the video in the manuscript as a figure.
\item Minimize file size by using an acceptable codec such as x264 or XviD. \href{https://handbrake.fr/}{HandBrake} is an open source tool for converting video to common codecs.
\item Video files must use open compression standards for display on broadly available applications such as VLC.
\item Animations must be formatted into a standard video container.
\end{itemize}

Visualizations must be associated with a figure, table, or equation OR be referenced in the results section of the manuscript. Use the label ``Visualization" and the item number to identify the visualization. Please note that to create text color for supplementary materials links, use of the color.sty package and the command \verb|\textcolor{blue}{Visualization}| is preferred to using the command \verb|\url{Visualization}|.

\newpage

\begin{figure}[ht!]
\centering\includegraphics{expressfig2}
\caption{Three traps create three rings of magnetic nanoparticles. The rings interact with one another (see \textcolor{blue}{Visualization 3}). [From Masajada \textit{et al.}, Opt. Lett. \textbf{38}, 3910 (2013)].}
\end{figure}

\begin{verbatim}
\begin{figure}[h]
\centering\includegraphics{opexfig2}
\caption{Three traps create three rings of magnetic
nanoparticles. The rings interact with one another (see
\textcolor{blue}{Visualization 3}). [From Masajada
\textit{et al.}, Opt. Lett. \textbf{38}, 3910 (2013)].}
\end{figure}
\end{verbatim}

Please refer to the \href{http://www.osapublishing.org/submit/style/multimedia.cfm}{Author Guidelines for Supplementary Materials} for more detailed instructions on acceptable multimedia formats for audio and tabular data.

\section{Mathematical and scientific notation}

\subsection{Displayed equations} Displayed equations should be centered.
Equation numbers should appear at the right-hand margin, in
parentheses:
\begin{equation}
H = \frac{1}{2m}(p_x^2 + p_y^2) + \frac{1}{2} M{\Omega}^2
     (x^2 + y^2) + \omega (x p_y - y p_x).
\end{equation}

All equations should be numbered in the order in which they appear
and should be referenced  from within the main text as Eq. (1),
Eq. (2), and so on [or as inequality (1), etc., as appropriate].

\subsection{Inline math} To help with conversion, place all math in a proper math environment. For example, expression \mbox{$3\times 4 = 12$} should be set this way, \texttt{\$3$\backslash$times 4=12\$}, not this way, \texttt{3 \$$\backslash$times\$4=12}. Simple fractions for inline math
should use parentheses when necessary to avoid ambiguity, for
example, to distinguish between $1/(n-1)$ and $1/n-1$.  Exceptions
to this are the proper fractions such as $\frac{1}{2}$, which are
better left in this form. Summations and integrals that appear
within text such as $\frac{1}{2}{\sum }_{n=1}^{n=\infty} (n^2 -
2n)^{-1}$ should have limits placed to the right of the symbol to
reduce white space.

\subsection{General guidelines on notation} Notation must be
legible, clear, compact, and consistent with standard usage. In
general, acronyms should be defined at first use. Adherence to the
following guidelines will greatly assist the production process:

\paragraph*{\bfseries Radical signs.}
When possible, avoid oversized radical signs
by using the notation of a superscript $1/2$. For example, change
$\sqrt{(a + b)(a - c)}$ to $[(a + b)(a - c)]^{1/2}$.

\paragraph*{\bfseries Exponentials.} Avoid tiny superscripts of exponential $e$ (e.g.,
$e^{jkl})$ by using the alternative \verb+\exp+ notation,
$\exp(jkl)$.

\paragraph*{\bfseries Variables and vectors.}
Set single-letter variables in italics $(k)$. Set three-vectors in
boldface $(\mathbf{k})$. Functions, derivative ``d,''
abbreviations, and multiletter identifiers should be set in roman
(plain) type  ($\alpha \cos, \int\!\dots{d}x, k^{out}$).

\paragraph*{\bfseries Multiplication.}
In general, close up multiplied terms $(p_yp_x)$;
use $\times$ if multiplication sign is essential $(2 \times
10^{-2})$ or for continuation in displayed equations. Use raised dot only for scalar product $(\mathbf{k \cdot k})$.

\paragraph*{\bfseries Fences.}
For simple bracketing the usual order of parentheses and brackets
is $\{ \, [  \, (  \,  \{  \, [  \, (  \, |  \, )  \, ]  \, \} \,
)  \, ]  \, \}$.

\paragraph*{\bfseries Metric system.}
The metric system is used in OSA journals. If nonmetric units are
essential (e.g., for parts specifications), conversion should be
given at first mention:  ``. . . a $\frac{1}{4}$\,-in. bolt \mbox{(1 in.
= 2.54 cm).''}

\section*{Funding}
Please identify all appropriate funding sources by name and contract number. Funding information should be listed in a separate block preceding any acknowledgments. List only the funding agencies and any associated grants or project numbers, as shown in the example below:\\
\\
National Science Foundation (NSF) (1253236, 0868895, 1222301); Program 973 (2014AA014402); Natural National Science Foundation (NSFC) (123456).\\
\\
OSA participates in \href{http://www.crossref.org/fundingdata/}{Crossref's Funding Data}, a service that provides a standard way to report funding sources for published scholarly research. To ensure consistency, please enter any funding agencies and contract numbers from the Funding section in Prism during submission or revisions.

\section*{Acknowledgments}
Acknowledgments, if included, should appear at the end of the document. The section title should not follow the numbering scheme of the paper.

\section*{Disclosures}
For \textit{Biomedical Optics Express} submissions only, disclosures should be listed in a separate nonnumbered section at the end of the manuscript. List the Disclosures codes identified on OSA's \href{http://www.osapublishing.org/submit/review/conflicts-interest-policy.cfm}{Conflict of Interest policy page}, as shown in the examples below:\\
\\
ABC: 123 Corporation (I,E,P), DEF: 456 Corporation (R,S). GHI: 789 Corporation (C).\\
\\
If there are no disclosures, then list ``The authors declare that there are no conflicts of interest related to this article.''

\section{References}
\label{sec:refs}
Proper formatting of references is extremely important, not only for consistent appearance but also for accurate electronic tagging. Please follow the guidelines provided below on formatting, callouts, and use of Bib\TeX.

\subsection{Formatting reference items}
Each source must have its own reference number. Footnotes (notes at the bottom of text pages) are not used in OSA journals. References require all author names, full titles, and inclusive pagination. Here are some examples of how to set the most common reference types:

\begin{description}
\item[Journal paper] \hfill\\
{\itshape Do not include web addresses in journal citations.}

C. van Trigt, ``Visual system-response functions and estimating reflectance,'' %\josaa
   J. Opt. Soc. Am. A {\bfseries 14}(4), 741--755 (1997).

S. Yerolatsitis, I. Gris-S\'anchez, and T. A. Birks,
``Adiabatically-tapered fiber mode multiplexers,'' Opt. Express {\bfseries 22}(1), 608--617 (2014).

\item[Journal paper identified by paper number] \hfill\\
{\itshape The paper number is sufficient. There is no need to give the number of pages.}

L. Rippe, B. Julsgaard, A. Walther, Y. Ying, and S. Kr\"oll,
``Experimental quantum-state tomography of a solid-state qubit,'' Phys. Rev. A {\bfseries 77}, 022307 (2008).

\item[Book] \hfill\\
T. Masters, {\itshape Practical Neural Network Recipes in C++} (Academic, 1993).

F. Ladouceur and J. D. Love, \textit{Silica-Based Buried Channel Waveguides and Devices} (Chapman \& Hall, 1995), Chap. 8.

\item[Article in a book] \hfill\\
D. F. Edwards, ``Silicon (Si),'' in {\itshape Handbook of Optical Constants of Solids,} E. D. Palik, ed. (Academic, 1985).

\item[Paper in a published conference proceedings] \hfill\\
R. E. Kalman, ``Algebraic aspects of the generalized inverse of a
rectangular matrix,'' in {\itshape Proceedings of Advanced Seminar on
Generalized Inverse and Applications,} M. Z. Nashed, ed. (Academic, 1976), pp. 111--124.

\item[Paper published in an OSA conference proceedings] \hfill\\
R. Craig and B. Gignac, ``High-power 980-nm pump lasers,'' in
{\itshape Optical Fiber Communication Conference}, Vol. 2 of 1996 OSA Technical Digest Series
(Optical Society of America, 1996), paper ThG1.

\item[Paper presented at a meeting/from an unpublished conference proceeding] \hfill\\
D. Steup and J. Weinzierl, ``Resonant THz-meshes,'' presented at the
Fourth International Workshop on THz Electronics,
Erlangen-Tennenlohe, Germany, 5--6 Sept. 1996.

\item[SPIE proceedings] \hfill\\
S. K. Griebel, M. Richardson, K. E. Devenport, and H. S. Hinton,
``Experimental performance of an ATM-based buffered hyperplane
CMOS-SEED smart pixel array,'' Proc. SPIE {\bfseries 3005}, 254--256 (1997).

{\itshape For later SPIE proceedings with a paper number, cite just the number and not any page information.}

S. Gu, F. Shao, G. Jiang, F. Li, and M. Yu,
``An objective visibility threshold measurement method for asymmetric stereoscopic images,''
Proc. SPIE {\bfseries 8205}, 820505 (2011).

\item[IEEE proceedings] \hfill\\
T. Darrel and K. Wohn, ``Pyramid based depth from focus,'' in
{\itshape Proceedings of IEEE Conference on Computer Vision and Pattern
Recognition} (IEEE, 1988), pp. 504--509.

\item[Paper accepted for publication] \hfill\\
D. Piao, ``Cancelation of coherent artifacts in optical coherence tomography imaging,'' Appl. Opt. (to be published).

D. W. Diehl and T. D. Visser,
"Phase singularities of the longitudinal field components in the focal region of a high-aperture optical system,"
J. Opt. Soc. Am. A, doc. ID 56789 (posted 11 November 2005, in press).

\item[Manuscript in preparation] \hfill\\
J. Q. Smith, Laboratory for Laser Energetics, University of Rochester, 250 East River Road, Rochester, New York 14623, USA, and K. Marshall are preparing a manuscript to be called ``Optical aspects in liquid crystals.''

\item[Personal communication] \hfill\\
T. Miller, Publications Department, Optical Society of
America, 2010 Massa\-chusetts Avenue, N.W., Washington, D.C.,
20036 (personal communication, 2010).

\item[Internet links] \hfill\\
Extreme Networks white paper, ``Virtual metropolitan area networks,'' (Extreme Networks, 2001), \texttt{http://www.extremenetworks.com/technology/
whitepapers/vMAN.asp}.

A. G. Ramm, ``Invisible obstacles,'' \texttt{http://www.arxiv.org/abs/math-ph
/0608034}.

\end{description}

The commands \verb+\begin{thebibliography}{}+ and \verb+\end{thebibliography}+ format the section according to standard style, showing the title {\bfseries References and links}.  Use the \verb+\bibitem{label}+ command to start each reference.

\subsection{Formatting reference citations}
References should be numbered consecutively in the order in which they are referenced in the body of the paper. Set reference callouts with standard \verb+\cite{}+ command or set manually inside square brackets [1].

To reference multiple articles at once, simply use the cite command with a comma separating the reference labels, e.g. \verb+\cite{gallo99,Masters98a,Oron03}+, produces \cite{gallo99,Masters98a,Oron03}. Using the \texttt{cite.sty} package will make these citations appear like so: [2--4].

\subsection{Bib\TeX}
\label{sec:bibtex}
Bib\TeX{} may be used to create a file containing the references, whose contents (i.e., contents of \texttt{.bbl} file) can then be pasted into the bibliography section of the \texttt{.tex} file. A new Bib\TeX{} style file, \texttt{osajnl.bst}, is provided.

To assist authors with journal abbreviations in references, standard abbreviations for some commonly cited journals have been included as macros within opex3.sty.  The abbreviations are shown in Table 2 below.

\begin{table}[htbp]
\centering\caption{Standard abbreviations
 for commonly cited journals.}
\begin{tabular}{lp{1.7in}|lp{1.7in}}
\hline
Macro        & Abbreviation                & Macro        & Abbreviation          \\ \hline
\verb+\ao+   & Appl.\  Opt.\               & \verb+\jpp+  & J. Phys.              \\
\verb+\aop+  & Adv. Opt. Photon.           & \verb+\nat+  & Nature                \\
\verb+\ap+   & Appl.\  Phys.\              & \verb+\oc+   & Opt.\ Commun.\        \\
\verb+\apl+  & Appl.\ Phys.\ Lett.\        & \verb+\opex+ & Opt.\ Express         \\
\verb+\apj+  & Astrophys.\ J.\             & \verb+\ol+   & Opt.\ Lett.\          \\
\verb+\bell+ & Bell Syst.\ Tech.\ J.\      & \verb+\ome+  & Opt.\ Mater.\ Express \\
\verb+\boe+  & Biomed.\ Opt.\ Express      & \verb+\opn+  & Opt.\ Photon.\ News   \\
\verb+\jqe+ & IEEE J.\ Quantum Electron.\  & \verb+\pl+   & Phys.\ Lett.\         \\
\verb+\assp+ & IEEE Trans.\ Acoust.\ Speech Signal Process.\  & \verb+\pr+ & Photon.\ Res.\ \\
\verb+\aprop+ & IEEE Trans.\  Antennas Propag.\    & \verb+\pra+ & Phys.\ Rev.\ A   \\
\verb+\mtt+ & IEEE Trans.\ Microwave Theory Tech.\ & \verb+\prb+ & Phys.\ Rev.\ B   \\
\verb+\iovs+ & Invest.\ Ophthalmol.\ Vis.\ Sci.\    & \verb+\prc+ & Phys.\ Rev.\ C   \\
\verb+\jcp+ & J.\ Chem.\ Phys.\            & \verb+\prd+ & Phys.\ Rev.\ D   \\
\verb+\jmo+ & J.\ Mod.\ Opt.\              & \verb+\pre+ & Phys.\ Rev.\ E   \\
\verb+\jocn+ & J.\ Opt.\ Commun.\ Netw.\   & \verb+\prl+ & Phys.\ Rev.\ Lett.\    \\
\verb+\jon+ & J.\ Opt.\ Netw.\             & \verb+\rmp+ & Rev.\ Mod.\ Phys.\    \\
\verb+\josa+ & J.\ Opt.\ Soc.\ Am.\        & \verb+\pspie+ & Proc.\ Soc.\ Photo-Opt.\ Instrum.\ Eng.\   \\
\verb+\josaa+ & J.\ Opt.\ Soc.\ Am.\ A     & \verb+\sjqe+ & Sov.\ J.\ Quantum Electron.\   \\
\verb+\josab+ & J.\ Opt.\ Soc.\ Am.\ B     & \verb+\vr+ & Vision Res.\   \\ \hline
\end{tabular}
\end{table}

\section{Conclusion}
After proofreading the manuscript, compress your .TEX manuscript file and all figures (which should be in EPS format, or PDF format if you are using PDF-\LaTeX) in a ZIP, TAR or TAR-GZIP package. Prism, OSA's article tracking system, will process in \LaTeX mode by default but will use PDF-\LaTeX if PDF figure files are detected. Note: TAR or TAR-GZIP is no longer required. All files must be referenced at the root level (e.g., file \texttt{figure-1.eps}, not \texttt{/myfigs/figure-1.eps}). If there is video or other supplementary materials, the associated files should be uploaded separately.

\end{document}

### 2017-07-06 paper/paper/paper.tex
Added paper template and moved examples into figures folder.

--- 2017-07-06 paper/Express_style.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%                File: express_style.tex              %
-%             Created: 2 September 2009               %
-%              Updated: 14 June 2017                  %
-%                                                     %
-%           LaTeX template file for use with          %
-%           OSA's journals Optics Express,            %
-%             Biomedical Optics Express,              %
-%            and Optical Materials Express            %
-%                                                     %
-%  send comments to Theresa Miller, tmiller@osa.org   %
-%                                                     %
-%       (c) 2017 Optical Society of America           %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \documentclass[10pt]{article}
-%% Specify the Express journal you are submitting to
-%\usepackage[OME]{express}
 \usepackage[OE]{express}
-%\usepackage[BOE]{express}
 
 \begin{document}
+\title{Determining Single-Molecule Orientation With Polarized Illumination:
+  Instrument Design and Forward Model}
 
-\title{Template and style guide for authors submitting to OSA Express Journals}
-
-\author{Author One,\authormark{1} Author Two,\authormark{2,*} and Author Three\authormark{2,3}}
+\author{Talon Chandler,\authormark{1} Author Two,\authormark{2,*} and Author Three\authormark{2,3}}
 
@@
 
-\email{\authormark{*}opex@osa.org} %% email address is required
-
-% \homepage{http:...} %% author's URL, if desired
+\email{talonchandler@talonchandler.com} %% email address is required
 
 %%%%%%%%%%%%%%%%%%% abstract and OCIS codes %%%%%%%%%%%%%%%%
-%% [use \begin{abstract*}...\end{abstract*} if exempt from copyright]
+\begin{abstract*}
+Abstract here. 
+\end{abstract*}
 
-\begin{abstract}
-Updated 14 June 2017.
-Detailed instructions and formatting guidelines for preparing \textit{Optics Express}, \textit{Biomedical Optics Express}, and \textit{Optical Materials Express} manuscripts in \LaTeX.
-For a simple outline and template, open the simple template file \texttt{Express\_temp.tex}.
-The Express journal simple and extended templates are also available on \href{http://www.overleaf.com/gallery/tagged/osa}{Overleaf}.
-OSA encourages the use of this free online collaborative tool for writing your OSA article.
-\end{abstract}
-
-\ocis{(000.0000) General; (000.2700) General science.} % REPLACE WITH CORRECT OCIS CODES FOR YOUR ARTICLE, MINIMUM OF TWO; Avoid using the OCIS codes for “General” or “General science” whenever possible.
-%For a complete list of OCIS codes, visit: https://www.osapublishing.org/oe/submit/ocis/
+\ocis{(180.0180) Microscopy; (260.5430) Polarization; (110.0110) Imaging systems; (180.2520) Fluorescence microscopy; (180.6900) Three-dimensional microscopy.}
 
@@
 \begin{thebibliography}{99}
-\bibitem{bib1}P. J. Harshman, T. K. Gustafson, and P. Kelley, ``Title of paper," J. Chem. Phys. {\bfseries 3}, (to be published).
-
 \bibitem{gallo99} K. Gallo and G. Assanto, ``All-optical diode based on second-harmonic generation in an asymmetric waveguide,'' \josab {\bfseries 16}(2), 267--269 (1999).
-
-\bibitem{Masters98a} B. R. Masters, ``Three-dimensional microscopic tomographic imagings of the cataract in a human lens in vivo,'' \opex {\bfseries 3}(9), 332--338 (1998).
-
-\bibitem{Oron03} D. Yelin,  D. Oron,  S. Thiberge,  E. Moses, and Y. Silberberg, ``Multiphoton plasmon-resonance microscopy,'' \opex {\bfseries 11}(12), 1385--1391 (2003).
-
-\bibitem{samplefig}
-B.~N.~Behnken, G.~Karunasiri, D.~R.~Chamberlin, P.~R.~Robrish, and J.~Faist,
-``Real-time imaging using a 2.8~THz quantum cascade laser and uncooled infrared microbolometer camera,'' \ol \textbf{33}(5), 440--442 (2008).
-
 \end{thebibliography}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%  body  %%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
-Adherence to the specifications listed in this template is essential for efficient review and publication of submissions.
-Since OSA does not routinely perform copyediting and typesetting for this journal, use of the template is critical to providing a consistent appearance.
-Proper reference format is especially important (see Section \ref{sec:refs}).
 
-\section{\texttt{express.sty} and required \LaTeX{} packages}
-Page layout is set with the \texttt{geometry} package for US Letter paper. \texttt{express.sty} uses the following package files:
+\section{Methods}
+Illumination forward model, CRLB, dispersion index
 
-\begin{itemize}
-\item \texttt{geometry} \ (page layout)
-\item \texttt{color, graphicx} \ (replaces \texttt{graphics}; has preset options)
-\item \texttt{mathptmx, courier, helvet} \ (Times, Courier, and Helvetica fonts)
-\end{itemize}
-
-The latest versions of these standard package files can be obtained at CTAN: the Comprehensive TeX Archive Network, \url{http://www.ctan.org}.
-
-The command \verb+\usepackage{ae}+ can be invoked to revert font to Computer Modern, although we prefer to publish with Times (with \texttt{mathptmx.sty}) for consistency.
-
-\section{Multiple corresponding authors}
-
-There are two options for indicating multiple corresponding authorship, and they are formatted quite differently.
-The first format would be as follows, still using the asterisk to denote one of the authors:
-
-\begin{verbatim}
-\author{Author One\authormark{1,3} and Author Two\authormark{2,4}}
-
-\address{\authormark{1}Peer Review, Publications Department,
-Optical Society of America, 2010 Massachusetts Avenue NW,
-Washington, DC 20036, USA\\
-\authormark{2}Publications Department, Optical Society of America,
-2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
-\authormark{3}xyz@osa.org}
-
-\email{\authormark{*}opex@osa.org}
-\end{verbatim}
-
-This format will generate the following appearance:\\
-
-\author{Author One\authormark{1,3} and Author Two\authormark{2,4}}
-
-\address{\authormark{1}Peer Review, Publications Department,
-Optical Society of America, 2010 Massachusetts Avenue NW,
-Washington, DC 20036, USA\\
-\authormark{2}Publications Department, Optical Society of America,
-2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
-\authormark{3}xyz@osa.org}
-
-\email{\authormark{*}opex@osa.org}
-\medskip
-The second format forgoes the asterisk and sets all email addresses equally within the affiliations. Please note that this format does not use the \verb+\email{}+ field at all.
-\begin{verbatim}
-\author{Author One\authormark{1,3} and Author Two\authormark{2,4}}
-
-\address{\authormark{1}Peer Review, Publications Department,
-Optical Society of America, 2010 Massachusetts Avenue NW,
-Washington, DC 20036, USA\\
-\authormark{2}Publications Department, Optical Society of America,
-2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
-\authormark{3}xyz@osa.org\\
-\authormark{4}opex@osa.org}
-\end{verbatim}
-
-This format will generate the following appearance:\\
-
-\author{Author One\authormark{1,3} and Author Two\authormark{2,4}}
-
-\address{\authormark{1}Peer Review, Publications Department,
-Optical Society of America, 2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
-\authormark{2}Publications Department, Optical Society of America, 2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
-\authormark{3}xyz@osa.org\\
-\authormark{4}opex@osa.org}
-\medskip
-These are the preferred express journal formats for multiple corresponding authorship, and either may be used.
-
-\section{Abstract}
-The abstract should be limited to approximately 100 words.
-It should be an explicit summary of the paper that states the problem, the methods used, and the major results and conclusions.
-It also should contain the relevant key words that would allow it to be found in a cursory computerized search.
-If the work of another author is cited in the abstract, that citation should be written out without a number, (e.g., journal, volume, first page, and year in square brackets [Opt.
-Express {\bfseries 22}, 1234 (2014)]), and a separate citation should be included in the body of the text.
-The first reference cited in the main text must be [1].
-Do not include numbers, bullets, or lists inside the abstract.
-
-\section{Figures, tables, and supplemental materials}
-
-\subsection{Figures and tables}
-
-OSA express journals encourage authors to submit color figures with their manuscripts. Figures and tables should be placed in the body of the manuscript.
-
-\begin{figure}[ht!]
-\centering\includegraphics[width=7cm]{expressfig1}
-\caption{Sample caption (Ref. \cite{Oron03}, Fig. 2).}
+\section{Results}
+\subsection{One-Arm Designs}
+\begin{figure}[htbp]
+\centering\includegraphics[width=\textwidth]{single-frame}
+\caption{Caption 123}
 \end{figure}
 
-\noindent Standard \LaTeX{} environments should be used to place tables and figures:
-\begin{verbatim}
-\begin{figure}[htbp]
-\centering\includegraphics[width=7cm]{expressfig1}
-\caption{Sample caption (\cite{Oron03}, Fig. 2).}
-\end{figure}
-\end{verbatim}
+\subsection{Two-Arm Designs}
+\subsection{Three-Arm Designs}
 
-\subsection{Supplementary materials in OSA express journals}
+\section{Discussion}
 
-Most OSA journals allow authors to include supplementary materials as integral parts of a manuscript.
-Such materials are subject to the same editorial standards and peer review procedures along with the rest of the paper and should be uploaded and described using OSA's Prism manuscript system.
-
-Authors can submit appropriate visualizations or small data files (see details below) for OSA to host.
-Large datasets and code or simulation files can be included but must be placed in an appropriate archival repository and cited as described here.
-
-\begin{table}[ht!]
-\centering
-\caption{Supplementary Materials Supported in OSA Express Journals}
-\begin{tabular}{|l|l|}
-\hline
-Visualization & 2D image, 3D image, video \\ \hline
-Data File     & Small data file such as data underlying a plot in a figure \\ \hline
-Dataset       & Dataset stored in an appropriate external repository \\ \hline
-Code          & Code or simulation files stored in an appropriate external repository \\
-\hline
-\end{tabular}
-\end{table}
-
-Video visualizations are the most commonly submitted type of supplementary materials for the express journals.
-They typically illustrate a synopsis of research results.
-They are integral and as such should be included only when they convey essential information beyond what can be presented within the article's PDF representation.
-Video visualizations should be uploaded upon submission and peer-reviewed along with the manuscript.
-Video files must use open compression standards for display on broadly available applications such as VLC or Windows Media Player.
-MOV, AVI, MPG, and MP4 video containers are accepted.
-The following video guidelines will help with the submission process:
-
-\begin{itemize}
-\item 15 MB is the recommended maximum video file size.
-\item 720 x 480 pixels (width by height) is the recommended screen size.
-\item If appropriate, insert a representative frame from the video in the manuscript as a figure.
-\item Minimize file size by using an acceptable codec such as x264 or XviD. \href{https://handbrake.fr/}{HandBrake} is an open source tool for converting video to common codecs.
-\item Video files must use open compression standards for display on broadly available applications such as VLC.
-\item Animations must be formatted into a standard video container.
-\end{itemize}
-
-Visualizations must be associated with a figure, table, or equation OR be referenced in the results section of the manuscript.
-Use the label ``Visualization" and the item number to identify the visualization.
-Please note that to create text color for supplementary materials links, use of the color.sty package and the command \verb|\textcolor{blue}{Visualization}| is preferred to using the command \verb|\url{Visualization}|.
-
-\newpage
-
-\begin{figure}[ht!]
-\centering\includegraphics{expressfig2}
-\caption{Three traps create three rings of magnetic nanoparticles.
-The rings interact with one another (see \textcolor{blue}{Visualization 3}).
-[From Masajada \textit{et al.}, Opt.
-Lett.
-\textbf{38}, 3910 (2013)].}
-\end{figure}
-
-\begin{verbatim}
-\begin{figure}[h]
-\centering\includegraphics{opexfig2}
-\caption{Three traps create three rings of magnetic
-nanoparticles. The rings interact with one another (see
-\textcolor{blue}{Visualization 3}). [From Masajada
-\textit{et al.}, Opt. Lett. \textbf{38}, 3910 (2013)].}
-\end{figure}
-\end{verbatim}
-
-Please refer to the \href{http://www.osapublishing.org/submit/style/multimedia.cfm}{Author Guidelines for Supplementary Materials} for more detailed instructions on acceptable multimedia formats for audio and tabular data.
-
-\section{Mathematical and scientific notation}
-
-\subsection{Displayed equations} Displayed equations should be centered.
-Equation numbers should appear at the right-hand margin, in
-parentheses:
-\begin{equation}
-H = \frac{1}{2m}(p_x^2 + p_y^2) + \frac{1}{2} M{\Omega}^2
-     (x^2 + y^2) + \omega (x p_y - y p_x).
-\end{equation}
-
-All equations should be numbered in the order in which they appear
-and should be referenced  from within the main text as Eq. (1),
-Eq. (2), and so on [or as inequality (1), etc., as appropriate].
-
-\subsection{Inline math} To help with conversion, place all math in a proper math environment.
-For example, expression \mbox{$3\times 4 = 12$} should be set this way, \texttt{\$3$\backslash$times 4=12\$}, not this way, \texttt{3 \$$\backslash$times\$4=12}.
-Simple fractions for inline math
-should use parentheses when necessary to avoid ambiguity, for
-example, to distinguish between $1/(n-1)$ and $1/n-1$.  Exceptions
-to this are the proper fractions such as $\frac{1}{2}$, which are
-better left in this form. Summations and integrals that appear
-within text such as $\frac{1}{2}{\sum }_{n=1}^{n=\infty} (n^2 -
-2n)^{-1}$ should have limits placed to the right of the symbol to
-reduce white space.
-
-\subsection{General guidelines on notation} Notation must be
-legible, clear, compact, and consistent with standard usage. In
-general, acronyms should be defined at first use. Adherence to the
-following guidelines will greatly assist the production process:
-
-\paragraph*{\bfseries Radical signs.}
-When possible, avoid oversized radical signs
-by using the notation of a superscript $1/2$. For example, change
-$\sqrt{(a + b)(a - c)}$ to $[(a + b)(a - c)]^{1/2}$.
-
-\paragraph*{\bfseries Exponentials.} Avoid tiny superscripts of exponential $e$ (e.g.,
-$e^{jkl})$ by using the alternative \verb+\exp+ notation,
-$\exp(jkl)$.
-
-\paragraph*{\bfseries Variables and vectors.}
-Set single-letter variables in italics $(k)$. Set three-vectors in
-boldface $(\mathbf{k})$. Functions, derivative ``d,''
-abbreviations, and multiletter identifiers should be set in roman
-(plain) type  ($\alpha \cos, \int\!\dots{d}x, k^{out}$).
-
-\paragraph*{\bfseries Multiplication.}
-In general, close up multiplied terms $(p_yp_x)$;
-use $\times$ if multiplication sign is essential $(2 \times
-10^{-2})$ or for continuation in displayed equations. Use raised dot only for scalar product $(\mathbf{k \cdot k})$.
-
-\paragraph*{\bfseries Fences.}
-For simple bracketing the usual order of parentheses and brackets
-is $\{ \, [  \, (  \,  \{  \, [  \, (  \, |  \, )  \, ]  \, \} \,
-)  \, ]  \, \}$.
-
-\paragraph*{\bfseries Metric system.}
-The metric system is used in OSA journals. If nonmetric units are
-essential (e.g., for parts specifications), conversion should be
-given at first mention:  ``. . . a $\frac{1}{4}$\,-in. bolt \mbox{(1 in.
-= 2.54 cm).''}
+\section{Conclusion}
+After proofreading the manuscript, compress your .TEX manuscript file and all
+figures (which should be in EPS format, or PDF format if you are using
+PDF-\LaTeX) in a ZIP, TAR or TAR-GZIP package. Prism, OSA's article tracking
+system, will process in \LaTeX mode by default but will use PDF-\LaTeX if PDF
+figure files are detected. Note: TAR or TAR-GZIP is no longer required. All
+files must be referenced at the root level (e.g., file \texttt{figure-1.eps},
+not \texttt{/myfigs/figure-1.eps}). If there is video or other multimedia, the
+associated files should be uploaded separately.
 
 \section*{Funding}
-Please identify all appropriate funding sources by name and contract number.
-Funding information should be listed in a separate block preceding any acknowledgments.
-List only the funding agencies and any associated grants or project numbers, as shown in the example below:\\
-\\
-National Science Foundation (NSF) (1253236, 0868895, 1222301); Program 973 (2014AA014402); Natural National Science Foundation (NSFC) (123456).\\
-\\
-OSA participates in \href{http://www.crossref.org/fundingdata/}{Crossref's Funding Data}, a service that provides a standard way to report funding sources for published scholarly research.
-To ensure consistency, please enter any funding agencies and contract numbers from the Funding section in Prism during submission or revisions.
+TODO
 
 \section*{Acknowledgments}
-Acknowledgments, if included, should appear at the end of the document. The section title should not follow the numbering scheme of the paper.
+TODO
 
 \section*{Disclosures}
-For \textit{Biomedical Optics Express} submissions only, disclosures should be listed in a separate nonnumbered section at the end of the manuscript.
-List the Disclosures codes identified on OSA's \href{http://www.osapublishing.org/submit/review/conflicts-interest-policy.cfm}{Conflict of Interest policy page}, as shown in the examples below:\\
-\\
-ABC: 123 Corporation (I,E,P), DEF: 456 Corporation (R,S). GHI: 789 Corporation (C).\\
-\\
-If there are no disclosures, then list ``The authors declare that there are no conflicts of interest related to this article.''
-
-\section{References}
-\label{sec:refs}
-Proper formatting of references is extremely important, not only for consistent appearance but also for accurate electronic tagging.
-Please follow the guidelines provided below on formatting, callouts, and use of Bib\TeX.
-
-\subsection{Formatting reference items}
-Each source must have its own reference number.
-Footnotes (notes at the bottom of text pages) are not used in OSA journals.
-References require all author names, full titles, and inclusive pagination.
-Here are some examples of how to set the most common reference types:
-
-\begin{description}
-\item[Journal paper] \hfill\\
-{\itshape Do not include web addresses in journal citations.}
-
-C. van Trigt, ``Visual system-response functions and estimating reflectance,'' %\josaa
-   J. Opt. Soc. Am. A {\bfseries 14}(4), 741--755 (1997).
-
-S. Yerolatsitis, I. Gris-S\'anchez, and T. A. Birks,
-``Adiabatically-tapered fiber mode multiplexers,'' Opt. Express {\bfseries 22}(1), 608--617 (2014).
-
-\item[Journal paper identified by paper number] \hfill\\
-{\itshape The paper number is sufficient. There is no need to give the number of pages.}
-
-L. Rippe, B. Julsgaard, A. Walther, Y. Ying, and S. Kr\"oll,
-``Experimental quantum-state tomography of a solid-state qubit,'' Phys. Rev. A {\bfseries 77}, 022307 (2008).
-
-\item[Book] \hfill\\
-T. Masters, {\itshape Practical Neural Network Recipes in C++} (Academic, 1993).
-
-F. Ladouceur and J. D. Love, \textit{Silica-Based Buried Channel Waveguides and Devices} (Chapman \& Hall, 1995), Chap. 8.
-
-\item[Article in a book] \hfill\\
-D. F. Edwards, ``Silicon (Si),'' in {\itshape Handbook of Optical Constants of Solids,} E. D. Palik, ed. (Academic, 1985).
-
-\item[Paper in a published conference proceedings] \hfill\\
-R. E. Kalman, ``Algebraic aspects of the generalized inverse of a
-rectangular matrix,'' in {\itshape Proceedings of Advanced Seminar on
-Generalized Inverse and Applications,} M. Z. Nashed, ed. (Academic, 1976), pp. 111--124.
-
-\item[Paper published in an OSA conference proceedings] \hfill\\
-R. Craig and B. Gignac, ``High-power 980-nm pump lasers,'' in
-{\itshape Optical Fiber Communication Conference}, Vol. 2 of 1996 OSA Technical Digest Series
-(Optical Society of America, 1996), paper ThG1.
-
-\item[Paper presented at a meeting/from an unpublished conference proceeding] \hfill\\
-D. Steup and J. Weinzierl, ``Resonant THz-meshes,'' presented at the
-Fourth International Workshop on THz Electronics,
-Erlangen-Tennenlohe, Germany, 5--6 Sept. 1996.
-
-\item[SPIE proceedings] \hfill\\
-S. K. Griebel, M. Richardson, K. E. Devenport, and H. S. Hinton,
-``Experimental performance of an ATM-based buffered hyperplane
-CMOS-SEED smart pixel array,'' Proc. SPIE {\bfseries 3005}, 254--256 (1997).
-
-{\itshape For later SPIE proceedings with a paper number, cite just the number and not any page information.}
-
-S. Gu, F. Shao, G. Jiang, F. Li, and M. Yu,
-``An objective visibility threshold measurement method for asymmetric stereoscopic images,''
-Proc. SPIE {\bfseries 8205}, 820505 (2011).
-
-\item[IEEE proceedings] \hfill\\
-T. Darrel and K. Wohn, ``Pyramid based depth from focus,'' in
-{\itshape Proceedings of IEEE Conference on Computer Vision and Pattern
-Recognition} (IEEE, 1988), pp. 504--509.
-
-\item[Paper accepted for publication] \hfill\\
-D. Piao, ``Cancelation of coherent artifacts in optical coherence tomography imaging,'' Appl. Opt. (to be published).
-
-D. W. Diehl and T. D. Visser,
-"Phase singularities of the longitudinal field components in the focal region of a high-aperture optical system,"
-J. Opt. Soc. Am. A, doc. ID 56789 (posted 11 November 2005, in press).
-
-\item[Manuscript in preparation] \hfill\\
-J.
-Q.
-Smith, Laboratory for Laser Energetics, University of Rochester, 250 East River Road, Rochester, New York 14623, USA, and K.
-Marshall are preparing a manuscript to be called ``Optical aspects in liquid crystals.''
-
-\item[Personal communication] \hfill\\
-T. Miller, Publications Department, Optical Society of
-America, 2010 Massa\-chusetts Avenue, N.W., Washington, D.C.,
-20036 (personal communication, 2010).
-
-\item[Internet links] \hfill\\
-Extreme Networks white paper, ``Virtual metropolitan area networks,'' (Extreme Networks, 2001), \texttt{http://www.extremenetworks.com/technology/
-whitepapers/vMAN.asp}.
-
-A. G. Ramm, ``Invisible obstacles,'' \texttt{http://www.arxiv.org/abs/math-ph
-/0608034}.
-
-\end{description}
-
-The commands \verb+\begin{thebibliography}{}+ and \verb+\end{thebibliography}+ format the section according to standard style, showing the title {\bfseries References and links}.
-Use the \verb+\bibitem{label}+ command to start each reference.
-
-\subsection{Formatting reference citations}
-References should be numbered consecutively in the order in which they are referenced in the body of the paper.
-Set reference callouts with standard \verb+\cite{}+ command or set manually inside square brackets [1].
-
-To reference multiple articles at once, simply use the cite command with a comma separating the reference labels, e.g.
-\verb+\cite{gallo99,Masters98a,Oron03}+, produces \cite{gallo99,Masters98a,Oron03}.
-Using the \texttt{cite.sty} package will make these citations appear like so: [2--4].
-
-\subsection{Bib\TeX}
-\label{sec:bibtex}
-Bib\TeX{} may be used to create a file containing the references, whose contents (i.e., contents of \texttt{.bbl} file) can then be pasted into the bibliography section of the \texttt{.tex} file.
-A new Bib\TeX{} style file, \texttt{osajnl.bst}, is provided.
-
-To assist authors with journal abbreviations in references, standard abbreviations for some commonly cited journals have been included as macros within opex3.sty.
-The abbreviations are shown in Table 2 below.
-
-\begin{table}[htbp]
-\centering\caption{Standard abbreviations
- for commonly cited journals.}
-\begin{tabular}{lp{1.7in}|lp{1.7in}}
-\hline
-Macro        & Abbreviation                & Macro        & Abbreviation          \\ \hline
-\verb+\ao+   & Appl.\  Opt.\               & \verb+\jpp+  & J. Phys.              \\
-\verb+\aop+  & Adv. Opt. Photon.           & \verb+\nat+  & Nature                \\
-\verb+\ap+   & Appl.\  Phys.\              & \verb+\oc+   & Opt.\ Commun.\        \\
-\verb+\apl+  & Appl.\ Phys.\ Lett.\        & \verb+\opex+ & Opt.\ Express         \\
-\verb+\apj+  & Astrophys.\ J.\             & \verb+\ol+   & Opt.\ Lett.\          \\
-\verb+\bell+ & Bell Syst.\ Tech.\ J.\      & \verb+\ome+  & Opt.\ Mater.\ Express \\
-\verb+\boe+  & Biomed.\ Opt.\ Express      & \verb+\opn+  & Opt.\ Photon.\ News   \\
-\verb+\jqe+ & IEEE J.\ Quantum Electron.\  & \verb+\pl+   & Phys.\ Lett.\         \\
-\verb+\assp+ & IEEE Trans.\ Acoust.\ Speech Signal Process.\  & \verb+\pr+ & Photon.\ Res.\ \\
-\verb+\aprop+ & IEEE Trans.\  Antennas Propag.\    & \verb+\pra+ & Phys.\ Rev.\ A   \\
-\verb+\mtt+ & IEEE Trans.\ Microwave Theory Tech.\ & \verb+\prb+ & Phys.\ Rev.\ B   \\
-\verb+\iovs+ & Invest.\ Ophthalmol.\ Vis.\ Sci.\    & \verb+\prc+ & Phys.\ Rev.\ C   \\
-\verb+\jcp+ & J.\ Chem.\ Phys.\            & \verb+\prd+ & Phys.\ Rev.\ D   \\
-\verb+\jmo+ & J.\ Mod.\ Opt.\              & \verb+\pre+ & Phys.\ Rev.\ E   \\
-\verb+\jocn+ & J.\ Opt.\ Commun.\ Netw.\   & \verb+\prl+ & Phys.\ Rev.\ Lett.\    \\
-\verb+\jon+ & J.\ Opt.\ Netw.\             & \verb+\rmp+ & Rev.\ Mod.\ Phys.\    \\
-\verb+\josa+ & J.\ Opt.\ Soc.\ Am.\        & \verb+\pspie+ & Proc.\ Soc.\ Photo-Opt.\ Instrum.\ Eng.\   \\
-\verb+\josaa+ & J.\ Opt.\ Soc.\ Am.\ A     & \verb+\sjqe+ & Sov.\ J.\ Quantum Electron.\   \\
-\verb+\josab+ & J.\ Opt.\ Soc.\ Am.\ B     & \verb+\vr+ & Vision Res.\   \\ \hline
-\end{tabular}
-\end{table}
-
-\section{Conclusion}
-After proofreading the manuscript, compress your .TEX manuscript file and all figures (which should be in EPS format, or PDF format if you are using PDF-\LaTeX) in a ZIP, TAR or TAR-GZIP package.
-Prism, OSA's article tracking system, will process in \LaTeX mode by default but will use PDF-\LaTeX if PDF figure files are detected.
-Note: TAR or TAR-GZIP is no longer required.
-All files must be referenced at the root level (e.g., file \texttt{figure-1.eps}, not \texttt{/myfigs/figure-1.eps}).
-If there is video or other supplementary materials, the associated files should be uploaded separately.
+The authors declare that there are no conflicts of interest related to this article.
 

### 2017-07-06 examples/compare-microscopes.py

@@ -1,94 +1,0 @@
-from dipsim import util
-from dipsim import multiframe
-import numpy as np
-import matplotlib.pyplot as plt
-import os; import time; start = time.time(); print('Running...')
-
-# Main input parameters
-row_labels = ['Geometry', r'$\sigma_{\Omega} = \sqrt{\textrm{det}\{I^{-1}\}}\sin\Theta$', r'Fractional Histogram']
-col_labels = ['One-View Wide-Field \n0.8 NA', 'Orthogonal Two-View Wide-Field \n0.8/0.8 NA', 'Two-View Wide-Field \n0.8/0.8 NA', 'Symmetric Two-View Light-Sheet \n0.8/0.8 NA', 'Asymmetric Two-View Light-Sheet \n0.71/1.1 NA', 'Three-View Wide-Field \n0.8/0.8/1.2 NA', 'Three-View Light-Sheet \n0.8/0.8/1.2 NA']
-
-n_pts = 50000
-n_cols = len(col_labels)
-n_rows = 3
-inch_fig = 5
-dpi = 400
-
-m1 = multiframe.MultiFrameMicroscope(ill_thetas=[0], det_thetas=[0],
-                                    ill_nas=[0.8], det_nas=[0.8],
-                                    ill_types=['wide'], det_types=['lens'],
-                                    colors=['(1,0,0)'], n_frames=4,
-                                    n_pts=n_pts, max_photons=1000, n_samp=1.33)
-
-m2 = multiframe.MultiFrameMicroscope(ill_thetas=[np.pi/4, -np.pi/4], det_thetas=[-np.pi/4, np.pi/4],
-                                    ill_nas=2*[0.8], det_nas=2*[0.8],
-                                    ill_types=2*['wide'], det_types=2*['lens'],
-                                    colors=['(1,0,0)','(0,0,1)'], n_frames=4,
-                                    n_pts=n_pts, max_photons=500, n_samp=1.33)
-
-m3 = multiframe.MultiFrameMicroscope(ill_thetas=[np.pi/4, -np.pi/4], det_thetas=[np.pi/4, -np.pi/4],
-                                    ill_nas=2*[0.8], det_nas=2*[0.8],
-                                    ill_types=2*['wide'], det_types=2*['lens'],
-                                    colors=['(1,0,0)','(0,0,1)'], n_frames=4,
-                                    n_pts=n_pts, max_photons=500, n_samp=1.33)
-
[…]

### 2017-07-06 examples/single-arm.py

@@ -1,43 +1,0 @@
-from dipsim import util
-from dipsim import multiframe
-import numpy as np
-import matplotlib.pyplot as plt
-import os; import time; start = time.time(); print('Running...')
-
-# Main input parameters
-n_pts = [100, 500, 1000]
-nas = np.linspace(0.01, 1.33, num=50)
-inch_fig = 5
-dpi = 400
-
-# Compute and plot on axes
-fig, ax = plt.subplots(1, 1, figsize=(5,5))
-for n_pt in n_pts:
-    d = []
-    for na in nas:
-        exp = multiframe.MultiFrameMicroscope(ill_thetas=[0], det_thetas=[0],
-                                        ill_nas=[na], det_nas=[na],
-                                        ill_types=['wide'], det_types=['lens'],
-                                        colors=['(1,0,0)'], n_frames=4,
-                                        n_pts=n_pt, max_photons=1000, n_samp=1.33)
-
-        print('Computing microscope: ' + str(n_pt) + ', ' + str(na))
-        exp.calc_estimation_stats()
-        d.append(util.dispersion_index(exp.root_det_sin))
-    print('Optimal NA is: ' + str(nas[np.argmin(d)]))
-    ax.plot(nas, d, '-', label='N = '+str(n_pt))
-    if n_pt == n_pts[-1]:
-        x = nas[np.argmin(d)]
-        y = np.min(d)
-        ax.arrow(x, y+0.05, 0, -0.04)
-        ax.annotate('NA* = {:.3}'.format(x), xy=(x, y+0.08), xytext=(x, y+0.08), horizontalalignment='center')
-
-ax.set_xlabel('NA')
-ax.set_ylabel('Dispersion Index')
-ax.set_xlim([0, 1.33])
-ax.set_ylim([0, 1])
-ax.legend(loc='upper right')
-fig.savefig('single-arm.png', dpi=dpi)
-
-print('Total time: '+str(np.round(time.time() - start, 2)))
-os.system('say "done"')

### 2017-07-06 examples/single-frame.py

@@ -1,63 +1,0 @@
-from dipsim import multiframe, util, detector, illuminator, microscope, util
-import numpy as np
-import matplotlib.pyplot as plt
-
-import os; import time; start = time.time(); print('Running...')
-
-# Main input parameters
-n_pts = 10000
-
-pols = 3*[0]
-ill_axes = 3*[0]
-det_axes = 3*[0]
-ill_nas = [0.9, 1.1, 1.3]
-det_nas = [0.9, 1.1, 1.3]
-n_rows = 4
-n_cols = len(det_axes)
-inch_fig = 5
-dpi = 250
-col_labels = ['NA = 0.9', 'NA = 1.1', 'NA = 1.3']
-row_labels = ['Geometry', r'Excitation Efficiency $\eta_{\text{exc}}$', r'Detection Efficiency $\eta_{\text{det}}$', r'Total Efficiency $\eta_{\text{tot}}$']
-
-# Generate axes
-size = (inch_fig*n_cols, inch_fig*n_rows)
-fig, axs = plt.subplots(n_rows, n_cols, figsize=size)
-plt.subplots_adjust(wspace=0.2, hspace=-0.3)
-if len(pols) == 1:
-    axs = np.expand_dims(axs, 1)
-caxs = util.generate_caxs(axs)
-
-# Compute and plot on axes
-for i, (pol, det_axis, ill_axis, det_na, ill_na) in enumerate(zip(pols, det_axes, ill_axes, det_nas, ill_nas)):
-    print('Computing microscope: ' + str(i))
-
-    # Create microscope
-    ill = illuminator.Illuminator(illum_type='wide',
-                                  theta_optical_axis=ill_axis,
-                                  na=ill_na,
-                                  phi_pol=pol)
-    
-    det = detector.Detector(theta_optical_axis=np.array(det_axis),
-                            na=det_na)
-                                  
-    m = microscope.Microscope(illuminator=ill, detector=det, max_photons=1)
-    
-    # Plot scene and efficiencies
-    m.plot_excitation_efficiency(n=n_pts, my_ax=axs[1,i], my_cax=caxs[1,i],
-                                 color_min=0, color_max=1.0)
-    m.plot_collection_efficiency(n=n_pts, my_ax=axs[2,i], my_cax=caxs[2,i],
-                                 color_min=0, color_max=0.5)
-    m.plot_sensitivity(n=n_pts, my_ax=axs[3,i], my_cax=caxs[3,i],
[…]

### 2017-07-06 paper/figures/single-frame.py

@@ -1,0 +1,63 @@
+from dipsim import multiframe, util, detector, illuminator, microscope, util
+import numpy as np
+import matplotlib.pyplot as plt
+
+import os; import time; start = time.time(); print('Running...')
+
+# Main input parameters
+n_pts = 10000
+
+pols = 3*[0]
+ill_axes = 3*[0]
+det_axes = 3*[0]
+ill_nas = [0.9, 1.1, 1.3]
+det_nas = [0.9, 1.1, 1.3]
+n_rows = 4
+n_cols = len(det_axes)
+inch_fig = 5
+dpi = 250
+col_labels = ['NA = 0.9', 'NA = 1.1', 'NA = 1.3']
+row_labels = ['Geometry', r'Excitation Efficiency $\eta_{\text{exc}}$', r'Detection Efficiency $\eta_{\text{det}}$', r'Total Efficiency $\eta_{\text{tot}}$']
+
+# Generate axes
+size = (inch_fig*n_cols, inch_fig*n_rows)
+fig, axs = plt.subplots(n_rows, n_cols, figsize=size)
+plt.subplots_adjust(wspace=0.2, hspace=-0.3)
+if len(pols) == 1:
+    axs = np.expand_dims(axs, 1)
+caxs = util.generate_caxs(axs)
+
+# Compute and plot on axes
+for i, (pol, det_axis, ill_axis, det_na, ill_na) in enumerate(zip(pols, det_axes, ill_axes, det_nas, ill_nas)):
+    print('Computing microscope: ' + str(i))
+
+    # Create microscope
+    ill = illuminator.Illuminator(illum_type='wide',
+                                  theta_optical_axis=ill_axis,
+                                  na=ill_na,
+                                  phi_pol=pol)
+    
+    det = detector.Detector(theta_optical_axis=np.array(det_axis),
+                            na=det_na)
+                                  
+    m = microscope.Microscope(illuminator=ill, detector=det, max_photons=1)
+    
+    # Plot scene and efficiencies
+    m.plot_excitation_efficiency(n=n_pts, my_ax=axs[1,i], my_cax=caxs[1,i],
+                                 color_min=0, color_max=1.0)
+    m.plot_collection_efficiency(n=n_pts, my_ax=axs[2,i], my_cax=caxs[2,i],
+                                 color_min=0, color_max=0.5)
+    m.plot_sensitivity(n=n_pts, my_ax=axs[3,i], my_cax=caxs[3,i],
[…]

### 2017-07-06 paper/figures/sweep-na.py

@@ -1,0 +1,43 @@
+from dipsim import util
+from dipsim import multiframe
+import numpy as np
+import matplotlib.pyplot as plt
+import os; import time; start = time.time(); print('Running...')
+
+# Main input parameters
+n_pts = [100, 500, 1000]
+nas = np.linspace(0.01, 1.33, num=50)
+inch_fig = 5
+dpi = 400
+
+# Compute and plot on axes
+fig, ax = plt.subplots(1, 1, figsize=(5,5))
+for n_pt in n_pts:
+    d = []
+    for na in nas:
+        exp = multiframe.MultiFrameMicroscope(ill_thetas=[0], det_thetas=[0],
+                                        ill_nas=[na], det_nas=[na],
+                                        ill_types=['wide'], det_types=['lens'],
+                                        colors=['(1,0,0)'], n_frames=4,
+                                        n_pts=n_pt, max_photons=1000, n_samp=1.33)
+
+        print('Computing microscope: ' + str(n_pt) + ', ' + str(na))
+        exp.calc_estimation_stats()
+        d.append(util.dispersion_index(exp.root_det_sin))
+    print('Optimal NA is: ' + str(nas[np.argmin(d)]))
+    ax.plot(nas, d, '-', label='N = '+str(n_pt))
+    if n_pt == n_pts[-1]:
+        x = nas[np.argmin(d)]
+        y = np.min(d)
+        ax.arrow(x, y+0.05, 0, -0.04)
+        ax.annotate('NA* = {:.3}'.format(x), xy=(x, y+0.08), xytext=(x, y+0.08), horizontalalignment='center')
+
+ax.set_xlabel('NA')
+ax.set_ylabel('Dispersion Index')
+ax.set_xlim([0, 1.33])
+ax.set_ylim([0, 1])
+ax.legend(loc='upper right')
+fig.savefig('sweep-na.png', dpi=dpi)
+
+print('Total time: '+str(np.round(time.time() - start, 2)))
+os.system('say "done"')

### 2017-07-07 paper/paper/paper.tex
Generate and caption single arm figure.

@@
 \begin{document}
-\title{Determining Single-Molecule Orientation With Polarized Illumination:
-  Instrument Design and Forward Model}
+\title{Polarized Illumination For Determining Single-Molecule Orientation:
+  Modeling and Instrument Design}
 
@@
 \section{Introduction}
+Single molecule orientation determination is important because
+
+A variety of techniques have been tested to determine the orientation of
+single molecules:
+
+polarized detection [Fourkas], shape analysis [Moerner], other things from
+Moerner review.
+
+We propose polarized illumination as an complementary method to polarized
+detection. We get the speed (no shape analysis required).
+
+In this work we make three contributions:
+\begin{itemize}
+\item we develop the polarized illumination forward model for a broad class of
+  microscopes (section 2.X-2.X)
+\item we develop metrics to compare the performance of polarized illumination
+  microscopes 
+\item we use these metrics to compare microscopes design and make recommendations
+  for polarized illumination microscopes. 
+\end{itemize}
 
 \section{Methods}
-Illumination forward model, CRLB, dispersion index
+\subsection{Illumination Model}
+
+\subsection{Detection Model}
+
+\subsection{Complete Forward Model}
+\begin{align}
+  I = I_{\text{tot}}\eta_{\text{exc}}\eta_{\text{det}}
+\end{align}
+
+\subsection{Microscope Designs}
+
+\subsection{Evaluation Metrics}
+\begin{align}
+  F &= \sum_{i=0}^N \frac{1}{I}
+  \begin{bmatrix}
+    \frac{\partial^2 I}{\partial \Theta^2}&\frac{\partial I}{\partial \Theta}\frac{\partial I}{\partial \Phi}\\
+    \frac{\partial I}{\partial \Theta}\frac{\partial I}{\partial \Phi}&\frac{\partial^2 I}{\partial \Phi^2}\\    
+  \end{bmatrix}\\
+  \sigma_{\Omega} &= \frac{\sin\Theta}{\sqrt{\text{det}\{F\}}}
+\end{align}
+We call $\sigma_{\Omega}$ the \emph{solid-angle uncertainty}. 
+We sampled $\sigma_{\Omega}$ on 100,000 approximately equally
+spaced points on the unit sphere and calculated the \emph{coefficient of variation}
+$c_v$ of these points where
+\begin{align}
+  c_v = \frac{\sigma_{\sigma_{\Omega}}}{\mu_{\sigma_{\Omega}}}
+\end{align}
 
@@
 \begin{figure}[htbp]
-\centering\includegraphics[width=\textwidth]{single-frame}
-\caption{Caption 123}
+\centering\includegraphics[width=\textwidth]{single-arm}
+\caption{\textbf{Left:} Schematic of a single-arm four-frame epi-illumination
+  microscope with NA = 0.8 and $n = 1.33$. The solid and dashed lines enclose
+  the illumination and detection solid angles, respectively, and the arrows
+  indicate the illumination polarization orientations. \textbf{Right:} Solid
+  angle uncertainty for the same microscope when $I_{\text{tot}} = 1000$
+  photons. Note the regions of high uncertainty near the equator and poles
+  as noted in XXX.}
 \end{figure}
@@
 \section{Conclusion}
-After proofreading the manuscript, compress your .TEX manuscript file and all
-figures (which should be in EPS format, or PDF format if you are using
-PDF-\LaTeX) in a ZIP, TAR or TAR-GZIP package. Prism, OSA's article tracking
-system, will process in \LaTeX mode by default but will use PDF-\LaTeX if PDF
-figure files are detected. Note: TAR or TAR-GZIP is no longer required. All
-files must be referenced at the root level (e.g., file \texttt{figure-1.eps},
-not \texttt{/myfigs/figure-1.eps}). If there is video or other multimedia, the
-associated files should be uploaded separately.
 

### 2017-07-07 dipsim/microscope.py

@@ -27,7 +27,7 @@ class Microscope:
         collect = self.detector.calc_collection_efficiency(flu)        
         return excite*collect
 
-    def plot_sensitivity(self, filename='out.png', n=50, **kwargs):
+    def plot_sensitivity(self, filename='', n=50, **kwargs):
         directions = util.fibonacci_sphere(n)
         print('Generating data for microscope: '+filename)
         I = np.apply_along_axis(self.calc_sensitivity, 1, directions)
@@ -47,7 +47,7 @@ class Microscope:
         util.plot_sphere(filename, directions=directions, data=I, **kwargs)
 
     def calc_collection_efficiency(self, direction):
-        flu = fluorophore.Fluorophore(mu_abs=direction, mu_em=args)        
+        flu = fluorophore.Fluorophore(mu_abs=direction, mu_em=direction)
         I = self.detector.calc_collection_efficiency(flu)
         return I
     

### 2017-07-07 paper/figures/single-arm.py

@@ -1,0 +1,53 @@
+from dipsim import util
+from dipsim import multiframe
+import numpy as np
+import matplotlib.pyplot as plt
+import os; import time; start = time.time(); print('Running...')
+
+# Main input parameters
+row_labels = ['']
+col_labels = ['Geometry', r'$\sigma_{\Omega}$']
+
+n_pts = 100000
+n_cols = len(col_labels)
+n_rows = 1
+inch_fig = 5
+dpi = 300
+
+# Generate axes
+size = (inch_fig*n_cols, inch_fig*n_rows)
+fig, axs = plt.subplots(n_rows, n_cols, figsize=size)
+plt.subplots_adjust(wspace=0, hspace=0)
+if len(col_labels) == 1:
+    axs = np.expand_dims(axs, 1)
+if len(row_labels) == 1:
+    axs = np.expand_dims(axs, 0)
+caxs = util.generate_caxs(axs)
+
+# Compute and plot on axes
+exp = multiframe.MultiFrameMicroscope(ill_thetas=[0], det_thetas=[0],
+                                      ill_nas=[0.8], det_nas=[0.8],
+                                      ill_types=['wide'], det_types=['lens'],
+                                      colors=['(1,0,0)'], n_frames=4,
+                                      n_pts=n_pts, max_photons=1000, n_samp=1.33)
+
+print('Computing microscope')
+exp.calc_estimation_stats()
+
+scene_string = exp.scene_string()
+util.draw_scene(scene_string, my_ax=axs[0,0], dpi=dpi)
+
+util.plot_sphere(directions=exp.directions, data=exp.sa_uncert,
+                 color_norm='log', linthresh=1e-4,
+                 color_min=8e-4, color_max=1e0,
+                 my_ax=axs[0,1], my_cax=caxs[0,1])
+
+caxs[0,0].axis('off')
+
+# Label axes and save
+util.label_rows_and_cols(axs, row_labels, col_labels, row_pos=(-0.2, 0.5))
+print('Saving final figure.')    
+fig.savefig('../paper/single-arm.pdf', dpi=dpi)
+
+print('Total time: '+str(np.round(time.time() - start, 2)))
+os.system('say "done"')

### 2017-07-07 paper/figures/single-frame.py

@@ -5,8 +5,7 @@ import matplotlib.pyplot as plt
 import os; import time; start = time.time(); print('Running...')
 
 # Main input parameters
-n_pts = 10000
-
+n_pts = 1000
 pols = 3*[0]
 ill_axes = 3*[0]
 det_axes = 3*[0]

### 2017-07-07 paper/paper/paper.tex
Generate and caption single frame figure.

@@
 \usepackage[OE]{express}
+\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
+\providecommand{\mb}[1]{\mathbf{#1}}
+\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
+\providecommand{\bs}[1]{\boldsymbol{#1}} 
+\providecommand{\intinf}{\int_{-\infty}^{\infty}}
 
@@
 \title{Polarized Illumination For Determining Single-Molecule Orientation:
-  Modeling and Instrument Design}
+  Modeling and Design Comparison}
 
-\author{Talon Chandler,\authormark{1} Author Two,\authormark{2,*} and Author Three\authormark{2,3}}
+\author{Talon Chandler,\authormark{1,*} Shalin Mehta,\authormark{2} Rudolf Oldenbourg,\authormark{2, 3} and Patrick J. La Rivi\`ere\authormark{1}}
 
-\address{\authormark{1}Peer Review, Publications Department, The Optical Society, 2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
-\authormark{2}Publications Department, The Optical Society, 2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
-\authormark{3}Currently with the Department of Electronic Journals, The Optical Society, 2010 Massachusetts Avenue NW, Washington, DC 20036, USA}
-
-\email{talonchandler@talonchandler.com} %% email address is required
+\address{\authormark{1}University of Chicago, Department of Radiology, Chicago, IL, USA.\\
+\authormark{2}Marine Biological Laboratory, Bell Center for Regenerative Medicine, Woods Hole, MA, USA.\\
+\authormark{3}Brown University, Department of Physics, Providence, RI, USA.}
+\email{*talonchandler@talonchandler.com} %% email address is required
 
@@
 We propose polarized illumination as an complementary method to polarized
-detection. We get the speed (no shape analysis required).
+detection. We get the speed (no shape analysis required). We don't waste any
+dose (polarized light throws away light dose).
 
@@
 \subsection{Illumination Model}
+\begin{align}
+  \eta_{\text{det},\phi_{\text{det}}}  = \frac{\int_{\Omega}d\mh{r}\ \left|\mh{P}_{\text{det}}\cdot\tilde{\mb{R}}(\mh{r})(\mh{r}\times\hat{\bs{\mu}}_{\text{em}}\times\mh{r})\right|^2}{\int_{\mathbb{S}^2}d\mh{r}\left|\mh{r}\times\hat{\bs{\mu}}_{\text{em}}\times\mh{r}\right|^2}\label{eq:final}
+\end{align}
+where
+\begin{align}
+  \hat{\mb{r}} &= \sin\theta\cos\phi\hat{\mb{i}} + \sin\theta\sin\phi\hat{\mb{j}} + \cos\theta\hat{\mb{k}}\label{eq:r_coords}\ \ \ \text{is the dummy integration vector,}\\
+  \hat{\bs{\mu}}_{\text{em}} &= \sin\Theta\cos\Phi\hat{\mb{i}} + \sin\Theta\sin\Phi\hat{\mb{j}} + \cos\Theta\hat{\mb{k}}\ \ \ \text{is the emission dipole moment,}\label{eq:mu_coords}\\
+  \tilde{\mb{R}}(\mh{r}) &= \begin{bmatrix} \cos\theta\cos^2\phi + \sin^2\phi & (\cos\theta -1)\sin\phi\cos\phi & -\sin\theta\cos\phi\\ (\cos\theta - 1)\sin\phi\cos\phi & \cos\theta\sin^2\phi + \cos^2\phi & -\sin\theta\sin\phi \\ \sin\theta\cos\phi& \sin\theta\sin\phi & \cos\theta \end{bmatrix}\label{eq:matrix}\\&\text{is the rotation matrix that models the objective lens,}\\
+  \hat{\mb{P}}_{\text{det}} &= \cos\phi_{\text{det}}\hat{\mb{i}} + \sin\phi_{\text{det}}\hat{\mb{j}}\ \ \ \text{is the pass axis of the polarizer,}\\
+  \Omega &= \{(\phi, \theta)|\phi \in (0,2\pi]\ \text{and}\ \theta\in (0, \alpha]\}\ \ \ \text{is the cap of the sphere collected by the objective.}
+\end{align}
+
+\begin{align}
+  \eta_{\text{exc},\phi_{\text{pol}}}(\Theta, \Phi, \alpha) &= D\{A + B\sin^{2}{\Theta} + C\sin^{2}{\Theta} \cos{[2 (\Phi - \phi_{\text{pol}}})]\}.
+\end{align}
+\begin{subequations}
+\begin{align}
+  A &= \frac{1}{4} - \frac{3}{8} \cos{\alpha } + \frac{1}{8} \cos^{3}{\alpha }\\
+  B &= \frac{3}{16} \cos{\alpha } - \frac{3}{16} \cos^{3}{\alpha }\\
+  C &= \frac{7}{32} - \frac{3}{32} \cos{\alpha } - \frac{3}{32} \cos^{2}{\alpha } - \frac{1}{32} \cos^{3}{\alpha}\\
+  D &= \frac{4}{3(1 - \cos\alpha)}.
+\end{align}%
+\end{subequations}
 
 \subsection{Detection Model}
+\begin{align}
+  \eta_{\text{det}}(\Theta, \alpha) = 2(A + B\sin^2\Theta)
+\end{align}
 
@@
 
+\begin{figure}[htbp]
+\centering\includegraphics[width=\textwidth]{single-frame}
+\caption{Representative examples of the forward model for single frame
+  microscopes.\newline \newline \textbf{Columns left to right:} (1) schematics of
+  single frame microscopes where the solid line encloses the illumination
+  solid angle, the dashed line encloses the detection solid angle, and the arrow
+  indicates the transmission axis of the illumination polarizer; (2) the excitation
+  efficiency see equation XX; (3) the detection efficiency see equation X; (4)
+  the total efficiency see equation X.\newline \newline \textbf{Rows top to
+    bottom:} (1) epi-illumination (NA = 1.1) with $x$-polarized light and
+  epi-detection (NA = 1.1); (2) orthogonal illumination (NA = 0.8) and detection
+  (NA = 0.8); (3) oblique illumination (NA = 0.8) and detection (NA = 0.8).}
+\end{figure}
+
 \subsection{Microscope Designs}
+Widefield illumination vs light sheet illumination.
+
+- Light sheet has low illumination. 
+- Restricted to orthogonal detection.
 
@@
   \begin{bmatrix}
-    \frac{\partial^2 I}{\partial \Theta^2}&\frac{\partial I}{\partial \Theta}\frac{\partial I}{\partial \Phi}\\
-    \frac{\partial I}{\partial \Theta}\frac{\partial I}{\partial \Phi}&\frac{\partial^2 I}{\partial \Phi^2}\\    
+    \frac{\partial I}{\partial \Theta}\frac{\partial I}{\partial \Theta}&\frac{\partial I}{\partial \Theta}\frac{\partial I}{\partial \Phi}\\\\
+    \frac{\partial I}{\partial \Theta}\frac{\partial I}{\partial \Phi}&\frac{\partial I}{\partial \Phi}\frac{\partial I}{\partial \Phi}\\    
   \end{bmatrix}\\
@@
 \caption{\textbf{Left:} Schematic of a single-arm four-frame epi-illumination
-  microscope with NA = 0.8 and $n = 1.33$. The solid and dashed lines enclose
-  the illumination and detection solid angles, respectively, and the arrows
-  indicate the illumination polarization orientations. \textbf{Right:} Solid
+  microscope with NA = 0.8 and $n = 1.33$.  \textbf{Right:} Solid
   angle uncertainty for the same microscope when $I_{\text{tot}} = 1000$
@@
 
+Can't use $c_v$ because it doesn't converge.
+
+High variance around equator and poles.
+
 \subsection{Two-Arm Designs}
+Widefield vs light sheet.
+
+Widefield NA vs angle/.
+
+Sweep asymmetry for full NA light sheet.
+
 \subsection{Three-Arm Designs}
@@
 \section*{Acknowledgments}
-TODO
+We thank Hari Shroff and Sean Rose for helpful conversations.
 

### 2017-07-07 paper/figures/single-frame.py

@@ -5,23 +5,24 @@ import matplotlib.pyplot as plt
 import os; import time; start = time.time(); print('Running...')
 
 # Main input parameters
-n_pts = 1000
-pols = 3*[0]
-ill_axes = 3*[0]
-det_axes = 3*[0]
-ill_nas = [0.9, 1.1, 1.3]
-det_nas = [0.9, 1.1, 1.3]
-n_rows = 4
-n_cols = len(det_axes)
+n_pts = 10000
+pols = [0, np.pi/4, np.pi/2]
+ill_axes = [0, np.pi/4, np.pi/4]
+det_axes = [0, -np.pi/4, np.pi]
+ill_nas = [1.1, 0.8, 0.8]
+det_nas = [1.1, 0.8, 0.8]
+n_rows = len(det_axes)
+n_cols = 4
 inch_fig = 5
-dpi = 250
-col_labels = ['NA = 0.9', 'NA = 1.1', 'NA = 1.3']
-row_labels = ['Geometry', r'Excitation Efficiency $\eta_{\text{exc}}$', r'Detection Efficiency $\eta_{\text{det}}$', r'Total Efficiency $\eta_{\text{tot}}$']
+dpi = 300
+
+col_labels = ['Geometry', r'Excitation Efficiency $\eta_{\text{exc}}$', r'Detection Efficiency $\eta_{\text{det}}$', r'Total Efficiency $\eta_{\text{tot}}$']
+row_labels = 3*['']
 
 # Generate axes
 size = (inch_fig*n_cols, inch_fig*n_rows)
 fig, axs = plt.subplots(n_rows, n_cols, figsize=size)
-plt.subplots_adjust(wspace=0.2, hspace=-0.3)
+plt.subplots_adjust(wspace=0.1, hspace=-0.1)
 if len(pols) == 1:
     axs = np.expand_dims(axs, 1)
 caxs = util.generate_caxs(axs)
@@ -42,16 +43,16 @@ for i, (pol, det_axis, ill_axis, det_na, ill_na) in enumerate(zip(pols, det_axes
     m = microscope.Microscope(illuminator=ill, detector=det, max_photons=1)
     
     # Plot scene and efficiencies
-    m.plot_excitation_efficiency(n=n_pts, my_ax=axs[1,i], my_cax=caxs[1,i],
+    m.plot_excitation_efficiency(n=n_pts, my_ax=axs[i,1], my_cax=caxs[i,1],
                                  color_min=0, color_max=1.0)
[…]

### 2017-07-13 dipsim/util.py
Polish symmetric widefield plot and add to paper.

@@ -210,6 +210,10 @@ def draw_scene(scene_string, filename='out.png', my_ax=None, dpi=500,
     subprocess.call(['convert', '-density', str(dpi), '-units', 'PixelsPerInch', 'temp.pdf', 'temp.png'])
 
     im = mpimg.imread('temp.png')
+
+    # Chop top of im to make it square and fix asy error
+    im = im[int(im.shape[1]*0.075):,:,:]
+    
     f = plt.figure(figsize=(5, 5), frameon=False)
     local_ax = plt.axes([0, 0, 1, 1]) # x, y, width, height
     if my_ax == None:

### 2017-07-13 paper/figures/symmetric-widefield.py

@@ -7,32 +7,44 @@ import os; import time; start = time.time(); print('Running...')
 import matplotlib.gridspec as gridspec
 
 # Main input parameters
-col_labels = ['Geometry', r'$\sigma_{\Omega}$', 'Mean$\{\sigma_{\Omega}\}$', 'STD$\{\sigma_{\Omega}\}$']
-n_pts = 100 # Points on sphere
-n_pts_sphere = 10000 # Points on sphere
-n_grid_pts = 13
+col_labels = ['Geometry (NA = 0.6, $\\beta=80{}^{\circ}$)', r'$\sigma_{\Omega}$ [sr]', 'Mean${}_{0-99}$$\{\sigma_{\Omega}\}$ [sr]', 'STD${}_{0-99}$$\{\sigma_{\Omega}\}$ [sr]']
+fig_labels = ['a)', 'b)', 'c)', 'd)']
+n_pts = 500 # Points on sphere
+n_pts_sphere = 25000 # Points on sphere
+n_grid_pts = 21
 n_rows, n_cols = 1, len(col_labels)
 inch_fig = 5
 dpi = 300
+percentile = 99
 
 # Setup figure and axes
-fig = plt.figure(figsize=(4.5*inch_fig, inch_fig))
-gs0 = gridspec.GridSpec(1, 2, wspace=0.125)
-gs00 = gridspec.GridSpecFromSubplotSpec(1, 3, subplot_spec=gs0[0], width_ratios=[1, 1, 0.05], wspace=0.05)
-gs01 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[1], width_ratios=[1, 1], wspace=0.3)
-gs010 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs01[0], width_ratios=[1, 0.05], wspace=0.1)
-gs011 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs01[1], width_ratios=[1, 0.05], wspace=0.1)
+fig = plt.figure(figsize=(2.2*inch_fig, 2*inch_fig))
+gs0 = gridspec.GridSpec(2, 2, wspace=0.4, hspace=0.1)
+gs00 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[0,0], width_ratios=[1, 0.05], wspace=0.1)
[…]

### 2017-07-16 dipsim/multiframe.py
Asymmetric light sheet plot.

@@ -31,8 +31,13 @@ class MultiFrameMicroscope:
                                               illum_type=ill_types[i],
                                               na=ill_nas[i], n=n_samp,
                                               phi_pol=phi)
-                
-                m.append(microscope.Microscope(illuminator=ill, detector=det, max_photons=max_photons, color=colors[i]))
+
+                if type(max_photons) == list:
+                    max_photon = max_photons[i]
+                else:
+                    max_photon = max_photons
+                    
+                m.append(microscope.Microscope(illuminator=ill, detector=det, max_photons=max_photon, color=colors[i]))
 
         self.microscopes = m
         self.noise_model = stats.NoiseModel(self.calc_total_intensities, **kwargs)

### 2017-07-16 paper/figures/asymmetric-double.py

@@ -1,0 +1,220 @@
+from dipsim import multiframe, util
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+import matplotlib.patches as patches
+import os; import time; start = time.time(); print('Running...')
+import matplotlib.gridspec as gridspec
+
+# Main input parameters
+col_labels = ['Geometry\n($\\alpha_{1}$ = 60${}^{\circ}$, Dose${}_1$/Dose${}_2$ = 10)', r'$\sigma_{\Omega}$ [sr]', 'Mean${}_{0-99}$$\{\sigma_{\Omega}\}$ [sr]', 'STD${}_{0-99}${$\{\sigma_{\Omega}\}$ [sr]']
+fig_labels = ['a)', 'b)', 'c)', 'd)']
+n_pts = 1000 # Points on sphere
+n_pts_sphere = 50000 # Points on sphere
+n_grid_pts = 21
+inch_fig = 5
+dpi = 300
+percentile = 99
+
+# Setup figure and axes
+fig = plt.figure(figsize=(2.4*inch_fig, 2*inch_fig))
+gs0 = gridspec.GridSpec(2, 2, wspace=0.3, hspace=0.1)
+gs00 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[0,0], width_ratios=[1, 0.05], wspace=0.45)
+gs10 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[1,0], width_ratios=[1, 0.05], wspace=0.45)
+gs01 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[0,1], width_ratios=[1, 0.05], wspace=0.45)
+gs11 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[1,1], width_ratios=[1, 0.05], wspace=0.45)
+ax0 = plt.subplot(gs00[0])
+ax1 = plt.subplot(gs01[0])
+ax2 = plt.subplot(gs10[0])
+ax3 = plt.subplot(gs11[0])
+cax0 = plt.subplot(gs00[1]); cax0.axis('off');
+cax1 = plt.subplot(gs01[1])
+cax2 = plt.subplot(gs10[1])
+cax3 = plt.subplot(gs11[1])
+
+for ax, col_label, fig_label  in zip([ax0, ax1, ax2, ax3], col_labels, fig_labels):
[…]

### 2017-07-17 paper/figures/triple-arm.py
Triple view figure.

@@ -1,0 +1,164 @@
+from dipsim import multiframe, util
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+import matplotlib.patches as patches
+import os; import time; start = time.time(); print('Running...')
+import matplotlib.gridspec as gridspec
+
+# Main input parameters
+col_labels = ['Geometry\n(NA${}_{\\textrm{upper}}$ = 0.6, NA${}_{\\textrm{lower}}$ = 1.1)', r'$\sigma_{\Omega}$ [sr]', 'Mean${}_{0-99}$$\{\sigma_{\Omega}\}$ [sr]', 'STD${}_{0-99}$$\{\sigma_{\Omega}\}$ [sr]']
+fig_labels = ['a)', 'b)', 'c)', 'd)']
+n_pts = 100 # Points on sphere
+n_pts_sphere = 50000 # Points on sphere
+n_grid_pts = 25
+inch_fig = 5
+dpi = 300
+percentile = 99
+
+# Setup figure and axes
+fig = plt.figure(figsize=(2.2*inch_fig, 2*inch_fig))
+gs0 = gridspec.GridSpec(2, 2, wspace=0.4, hspace=0.1)
+gs00 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[0,0], width_ratios=[1, 0.05], wspace=0.1)
+gs10 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[1,0], width_ratios=[1, 0.05], wspace=0.1)
+gs01 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[0,1], width_ratios=[1, 0.05], wspace=0.1)
+gs11 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[1,1], width_ratios=[1, 0.05], wspace=0.1)
+ax0 = plt.subplot(gs00[0])
+ax1 = plt.subplot(gs01[0])
+ax2 = plt.subplot(gs10[0])
+ax3 = plt.subplot(gs11[0])
+cax0 = plt.subplot(gs00[1]); cax0.axis('off');
+cax1 = plt.subplot(gs01[1])
+cax2 = plt.subplot(gs10[1])
+cax3 = plt.subplot(gs11[1])
+
+for ax, col_label, fig_label  in zip([ax0, ax1, ax2, ax3], col_labels, fig_labels):
[…]

### 2017-07-18 paper/paper/paper.tex
Polish symmetric widefield plot and add to paper.
Asymmetric light sheet plot.
Triple view figure.
Paper edits. Methods and results section drafts.

--- 2017-07-06 paper/paper/paper.tex
@@
 \usepackage[OE]{express}
+\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
+\providecommand{\mb}[1]{\mathbf{#1}}
+\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
+\providecommand{\bs}[1]{\boldsymbol{#1}} 
+\providecommand{\intinf}{\int_{-\infty}^{\infty}}
 
 \begin{document}
-\title{Determining Single-Molecule Orientation With Polarized Illumination:
-  Instrument Design and Forward Model}
+\title{Single-Molecule Orientation Determination With Polarized Illumination:
+  Modeling and Microscope Design Comparison}
 
-\author{Talon Chandler,\authormark{1} Author Two,\authormark{2,*} and Author Three\authormark{2,3}}
+\author{Talon Chandler,\authormark{1,*} Shalin Mehta,\authormark{2} Rudolf Oldenbourg,\authormark{2, 3} and Patrick J. La Rivi\`ere\authormark{1}}
 
-\address{\authormark{1}Peer Review, Publications Department, The Optical Society, 2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
-\authormark{2}Publications Department, The Optical Society, 2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
-\authormark{3}Currently with the Department of Electronic Journals, The Optical Society, 2010 Massachusetts Avenue NW, Washington, DC 20036, USA}
-
-\email{talonchandler@talonchandler.com} %% email address is required
+\address{\authormark{1}University of Chicago, Department of Radiology, Chicago, IL, USA.\\
+\authormark{2}Marine Biological Laboratory, Bell Center for Regenerative Medicine, Woods Hole, MA, USA.\\
+\authormark{3}Brown University, Department of Physics, Providence, RI, USA.}
+\email{*talonchandler@talonchandler.com} %% email address is required
 
@@
 \begin{abstract*}
-Abstract here. 
+  TODO
 \end{abstract*}
@@
 %%%%%%%%%%%%%%%%%%%%%%% References %%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{thebibliography}{99}
-\bibitem{gallo99} K. Gallo and G. Assanto, ``All-optical diode based on second-harmonic generation in an asymmetric waveguide,'' \josab {\bfseries 16}(2), 267--269 (1999).
-\end{thebibliography}
+%% Comment before submission
+\bibliography{paper}
+\bibliographystyle{osajnl}
+
+%% Copy .bbl and uncomment these before submission
+% \begin{thebibliography}{99}
+% \bibitem{gallo99} K. Gallo and G. Assanto, ``All-optical diode based on second-harmonic generation in an asymmetric waveguide,'' \josab {\bfseries 16}(2), 267--269 (1999).
+% \end{thebibliography}
 
 \section{Introduction}
+TODO:
+  \begin{itemize}
+  \item General paragraph about the dipole orientation imaging.
+  \item Current state of the art: scanning, detection polarization modulation,
+    geometries
+  \item We propose single-molecule orientation determination with polarized
+    illumination. Compare with others. Highly parallel imaging for live cells.
+  \item Section by section summary. 
+  \end{itemize}
+% Single molecule orientation determination is important because
+
+% A variety of techniques have been tested to determine the orientation of
+% single molecules:
+
+% polarized detection [Fourkas], shape analysis [Moerner], other things from
+% Moerner review.
+
+% We propose polarized illumination as an complementary method to polarized
+% detection. We get the speed (no shape analysis required). We don't waste any
+% dose (polarized light throws away light dose).
+
+% In this work we make three contributions:
+% \begin{itemize}
+% \item we develop the polarized illumination forward model for a broad class of
+%   microscopes (section 2.X-2.X)
+% \item we develop metrics to compare the performance of polarized illumination
+%   microscopes 
+% \item we use these metrics to compare microscopes design and make recommendations
+%   for polarized illumination microscopes. 
+% \end{itemize}
 
 \section{Methods}
-Illumination forward model, CRLB, dispersion index
+In this section we will develop a method to compare polarized illumination
+microscope designs. In sections \ref{excitation}--\ref{forward} we develop the
+forward model for single-frame microscopes with a fixed illumination and
+detection orientation. In section \ref{designs} we list a variety of ways that
+single-frame microscopes can be combined to create experimentally realizable
+multi-frame microscopes. Finally, in section \ref{metrics} we develop metrics
+that we will use to compare multi-frame microscopes.
+
+\subsection{Excitation Efficiency}\label{excitation}
+Consider polarized K\"ohler illumination incident on a single fluorophore at the
+focal point with the optical axis aligned along the $\mh{z}$ axis. In this section we
+will calculate the \emph{excitation efficiency} in this geometry---the fraction
+of the incident power that excites the fluorophore.
+
+We start with the unit electric field transmitted by the polarizer,
+\begin{align}
+ \hat{\mb{E}}_{\text{exc}} = \cos\phi_{\text{exc}}\hat{\mb{x}} +
+\sin\phi_{\text{exc}}\hat{\mb{y}},\label{eq:excfield}
+\end{align}
+where $\phi_{\text{exc}}$ is the angle between the transmission axis of the
+excitation polarizer and the $\mh{x}$ axis. Next, we model the action of the
+illumination lens by multiplying the electric field vector by a position
+dependent rotation matrix,
+  \begin{align}
+  \mb{R}(\mh{r}) &= \begin{bmatrix} \cos\theta\cos^2\phi + \sin^2\phi & (\cos\theta -1)\sin\phi\cos\phi & -\sin\theta\cos\phi\\ (\cos\theta - 1)\sin\phi\cos\phi & \cos\theta\sin^2\phi + \cos^2\phi & -\sin\theta\sin\phi \\ \sin\theta\cos\phi& \sin\theta\sin\phi & \cos\theta \end{bmatrix}
+  \end{align}
+  where
+  \begin{align}
+   \hat{\mb{r}} = \sin\theta\cos\phi\hat{\mb{x}} + \sin\theta\sin\phi\hat{\mb{y}} + \cos\theta\hat{\mb{z}} 
+  \end{align}
+  is a unit vector that points from the origin to the point where the incident
+  electric field intersects the Gaussian sphere of the condenser lens. Next, we
+  find the power that excites the fluorophore by taking the dot product of the
+  rotated incident field $\mb{R}(\mh{r})\mh{E}_{\text{exc}}$ and the excitation
+  dipole moment
+\begin{align}
+  \hat{\bs{\mu}}_{\text{exc}} = \sin\Theta\cos\Phi\hat{\mb{x}} + \sin\Theta\sin\Phi\hat{\mb{y}} + \cos\Theta\hat{\mb{z}}
+\end{align}
+then taking its modulus squared. If the incident fields are spatially incoherent
+then we can find the total power that excites the fluorophore by integrating over the
+illuminated region of the Gaussian sphere $\Omega$ where
+\begin{align}
+  \Omega &= \{(\phi, \theta)\ |\ \phi \in (0,2\pi], \theta\in [0, \alpha]\} \\
+  \alpha &= \arcsin\left(\frac{\text{NA}}{n}\right). \label{eq:alpha}
+\end{align}
+
+Finally, we find the fraction of the incident power that excites the fluorophore by
+dividing the total excitation power by the total incident power. The complete expression for the excitation efficiency in vector notation is 
+\begin{align}
+  \eta_{\text{exc}} = \frac{\int_{\Omega}d\mh{r}|\hat{\bs{\mu}}_{\text{exc}}\cdot\mb{R}(\mh{r})\mh{E}_{\text{exc}}|^2}{\int_{\Omega}d\mh{r}}\label{eq:exc}. 
+\end{align}
+
+We can substitute equations \ref{eq:excfield}--\ref{eq:alpha} into equation
+\ref{eq:exc} and simplify to express the excitation efficiency in scalar notation as
+\begin{align}
+  \eta_{\text{exc}} &= D\{A + B\sin^{2}{\Theta} + C\sin^{2}{\Theta} \cos{[2 (\Phi - \phi_{\text{exc}}})]\}\label{eq:scalarexc}
+\end{align}
+where
+\begin{subequations}
+\begin{align}
+  A &= \frac{1}{4} - \frac{3}{8} \cos{\alpha } + \frac{1}{8} \cos^{3}{\alpha }\\
+  B &= \frac{3}{16} \cos{\alpha } - \frac{3}{16} \cos^{3}{\alpha }\\
+  C &= \frac{7}{32} - \frac{3}{32} \cos{\alpha } - \frac{3}{32} \cos^{2}{\alpha } - \frac{1}{32} \cos^{3}{\alpha}\\
+  D &= \frac{4}{3(1 - \cos\alpha)}.
+\end{align}\label{eq:coefficients}%
+\end{subequations}
+
+At the beginning of this section we assumed that the fluorophore was located at
+the focal point of the condenser. If the condenser is aplanatic then we can
+relax this assumption and the above expressions are valid for fluorophores anywhere
+in the focal plane.
+
+We also assumed that we are using incoherent K\"ohler illumination. If we
+illuminate the back aperture with a weakly focused laser beam and scan the laser
+beam slowly compared to the coherence time of the laser, the fluorophore will be
+excited as if it were excited by a single plane wave. Therefore, we can find the
+excitation efficiency of weakly focused scanned laser illumination by taking the
+limit of equation \ref{eq:scalarexc} as $\alpha \rightarrow 0$ which gives
+\begin{align*}
+  \eta_{\text{exc}} &= \sin^2\Theta\cos^2(\Phi - \phi_{\text{exc}}).
+\end{align*}
+
+\subsection{Detection Efficiency}\label{detection}
+Fourkas calculated the detection efficiency of a single fluorophore when an
+objective with a polarizer is aligned along the $\mh{z}$ axis
+\cite{fourkas2001}. We use his expressions to calculate the detection efficiency
+without a polarizer as
+\begin{align}
+  \eta_{\text{det}} = 2(A + B\sin^2\Theta). \label{eq:scalardet}
+\end{align}
+Note that the detection efficiency only depends on $\Theta$, not $\Phi$, because
+there is no detection polarizer. Also note that $A$ and $B$ in equation
+\ref{eq:coefficients} are a factor of $\frac{3}{2}$ larger than the expressions
+given by Fourkas. We found that Fourkas' expressions were incorrectly
+normalized, and the extra factor of $\frac{3}{2}$ corrects the error.
+
+\subsection{Oblique Illumination and Detection}\label{oblique}
+In sections \ref{excitation} and \ref{detection} we assumed that both the
+illumination and detection objectives had $\mh{z}$ aligned optical axes. To extend
+the forward model to oblique optical axes, we express the dipole orientation in
+rotated coordinates using the following expressions
+\begin{align}
+    \theta' &= \arccos\left(\sin\psi\cos\Phi\sin\theta + \cos\psi\cos\theta\right)\label{eq:thetap}\\
+  \Phi' &=
+          \begin{cases}
+            \arccos\left(\frac{\cos\psi\cos\Phi\sin\theta - \sin\psi\cos\theta}{\sqrt{1 - (\sin\psi\cos\Phi\sin\theta + \cos\psi\cos\theta)^2}}\right) \ \ \ &0 \leq \Phi < \pi  \\
+            -\arccos\left(\frac{\cos\psi\cos\Phi\sin\theta - \sin\psi\cos\theta}{\sqrt{1 - (\sin\psi\cos\Phi\sin\theta + \cos\psi\cos\theta)^2}}\right) \ \ \ &-\pi \leq \Phi < 0\\
+          \end{cases}
+\end{align}
+where $\psi$ is the angle that the optical axis makes with the $\mh{z}$ axis in
+the $\mh{x}$ direction.
+
+\subsection{Single-Frame Microscopes}\label{forward}
+The detected intensity is the product of the excitation efficiency, the
+detection efficiency, and the total intensity emitted by the fluorophore if the
+excitation and detection efficiency were 1 
+\begin{align}
+  I = I_{\text{tot}}\eta_{\text{exc}}\eta_{\text{det}}\label{eq:forward_frame}.
+\end{align}
+Equation \ref{eq:forward_frame} is the forward model for a \emph{single-frame
+  microscope}. Figure \ref{fig:single-frame} shows three representative examples of
+single frame microscopes.
+
+\begin{figure}[htbp]
+\centering\includegraphics[width=\textwidth]{single-frame}
+\caption{Representative examples of single-frame microscopes.\newline \newline
+  \textbf{Columns left to right:} 1) schematics of single-frame microscopes
+  where the solid line encloses the illumination solid angle, the dashed line
+  encloses the detection solid angle, and the arrow indicates the transmission
+  axis of the illumination polarizer; 2) the excitation efficiency, see equation
+  \ref{eq:scalarexc}; 3) the detection efficiency, see equation
+  \ref{eq:scalardet}; (4) the total efficiency, the product of the detection and
+  excitation efficiencies.\newline \newline \textbf{Rows top to bottom:} 1)
+  epi-illumination (NA = 1.1) with $x$-polarized light and epi-detection (NA =
+  1.1); 2) orthogonal illumination (NA = 0.8) and detection (NA = 0.8); 3)
+  oblique illumination (NA = 0.8) and detection (NA = 0.8).}
+  \label{fig:single-frame}
+\end{figure}
+
+The intensity measured by a single-frame microscope does not give us enough
+information to reconstruct the 3D orientation of a single fluorophore. Next, we
+will consider combining several single-frame microscopes to create
+\emph{multi-frame microscopes.}
+
+\subsection{Multi-Frame Microscopes}\label{designs}
+One way to collect multiple frames is to add a universal compensator to the
+illumination arm and rapidly select the incident polarization by changin
+$\phi_{\text{exc}}$ \cite{oldenbourg1995}. All of the multi-frame microscopes we
+will consider in this paper use a universal compensator with four polarization
+settings separated by 45${}^{\circ}$.
+
+We also consider designs that use extra detection and illumination arms. Wu et
+al. have implemented a microscope that uses orthogonal objectives that can act
+as both illumination and detection arms \cite{wu2013}. Wu et al. has also added
+a third view to their microscopes to further improve resolution
+\cite{wu2016}. We will explore the possibility of using these geometries for
+orientation determination of fluorophores.
+
+\subsection{Evaluation Metrics}\label{metrics}
+Our goal is to evaluate the ability of a microscope design to estimate the
+parameters $\Theta$ and $\Phi$ from intensity data.
+
+For each microscope design and fluorophore orientation we calculate the Fisher
+information matrix. If the intensity is Poisson distributed the
+Fisher information matrix is given by
+\begin{align}
+  \mb{F} &= \sum_{k=1}^N \frac{1}{I_k}
+  \begin{bmatrix}
+    \frac{\partial I_k}{\partial \Theta}\frac{\partial I_k}{\partial \Theta}&\frac{\partial I_k}{\partial \Theta}\frac{\partial I_k}{\partial \Phi}\\\\
+    \frac{\partial I_k}{\partial \Theta}\frac{\partial I_k}{\partial \Phi}&\frac{\partial I_k}{\partial \Phi}\frac{\partial I_k}{\partial \Phi}\\    
+  \end{bmatrix}
+\end{align}
+where $I_k$ is the forward model for the $k$th frame of an $N$ frame microscope.
+
+A common way to evaluate the ability to estimate the parameters $\Theta$ and
+$\Phi$ from the data is to calculate the Cramer-Rao lower bound (CRLB) for each
+parameter \cite{kay1993}. The CRLBs are given by the diagonal elements of the
+inverse of the Fisher information matrix and they give the minimum variance of
+unbiased estimators for each parameter. Some authors use the product of the
+CRLBs multiplied by the Jacobian determinant to find the area of uncertainty in
+parameter space \cite{agrawal2012}.
+
+CRLBs and the associated area of uncertainty are parametrization dependent---if
+we choose a different coordinate system then the values will change
+significantly. We would like to compare microscope designs without choosing a
+parametrization.
+
+Instead of the taking the product of the CRLBs, we use
+\begin{align}
+  \sigma_{\Omega} &= \sin\Theta\sqrt{\text{det}\{\mb{F}^{-1}\}}
+\end{align}
+as our evaluation metric. We take the determinant of the inverse Fisher
+information matrix---a parametrization independent value often called the
+\emph{generalized variance} \cite{anderson1958}---take the square root, then
+multiply by the Jacobian determinant, $\sin\Theta$. We call $\sigma_{\Omega}$ the
+\emph{solid-angle uncertainty} because it has units of steradians and is a
+measure of the uncertainty of the orientation parameters.
+
+We calculated $\sigma_{\Omega}$ on 100,000 approximately equally spaced points
+on the unit sphere for each microscope design. A desirable microscope design
+will have a small solid-angle uncertainty that is uniform for all fluorophore
+orientations. To summarize and compare microscope designs we use the mean and
+standard deviation of the solid-angle uncertainty at the sample points. Because
+the solid-angle uncertainty admits unbounded values, we ignore the highest
+percentile of the sample points and take the mean and standard deviation of the
+0-99th percentile.
 
 \section{Results}
-\subsection{One-Arm Designs}
+\subsection{One-Arm Designs}\label{one-arm}
+\begin{itemize}
+\item Figure \ref{fig:single-arm} shows the results when we used a single
+  objective to illuminate and detect.
+\item We swept through the illumination and detection NA while keeping the
+  incident power ($I_{\text{tot}}$) constant and found that the lowest mean and
+  standard deviation of the solid-angle uncertainty occurs with a small
+  illumination NA and a large detection NA.
+\item A small illumination NA maximizes the ``contrast'' in the excitation
+  efficiency, while a large detection NA maximizes the ``contrast'' in the
+  detection efficiency. 
+\end{itemize}
+
 \begin{figure}[htbp]
-\centering\includegraphics[width=\textwidth]{single-frame}
-\caption{Caption 123}
+\centering\includegraphics[width=\textwidth]{single-arm}
+\caption{Epi-illumination microscope with varying NA and underfilling of the
+  back aperture. a) Schematic of a single-arm four-frame epi-illumination
+  microscope. b) Solid angle uncertainty for the microscope in a) when
+  $I_{\text{tot}} = 1000$ photons. c) Mean of the solid-angle uncertainty as a
+  function of illumination and detection NA. d) Standard deviation of
+  solid-angle uncertainty as a function of illumination and detection NA. The
+  microscope in a) and b) is indicated by a cross in c) and d).}
+\label{fig:single-arm}
 \end{figure}
@@
 \subsection{Two-Arm Designs}
+\begin{itemize}
+\item Figure \ref{fig:symmetric-widefield} shows the results when we used a
+  symmetric widefield design. Both objectives serve as a an illumination and
+  detection NA.
+\item We swept through the NA and the angle between the objectives while
+  considering steric constraints and while keeping the incident power constant. 
+  We found that the lowest mean and standard deviation of the solid-angle
+  uncertainty occurs with orthogonal arms with the highest possible NA. 
+\end{itemize}
+
+\begin{figure}[htbp]
+\centering\includegraphics[width=\textwidth]{symmetric-widefield}
+\caption{Symmetric widefield designs with varying NA and angle between objectives. See the Fig. \ref{fig:single-arm} caption for an explanation of a)-d).}
+\label{fig:symmetric-widefield}
+\end{figure}
+
+\begin{itemize}
+\item Figure \ref{fig:double-arm} shows the results when we used a
+  symmetric widefield design and varied the illumination and detection NA. 
+\item Similar to section \ref{one-arm}, we found that the best microscopes occur
+  with a large detection NA and a small illumination NA (like light sheet illumination).
+\end{itemize}
+
+\begin{figure}[htbp]
+\centering\includegraphics[width=\textwidth]{double-arm}
+\caption{Symmetric widefield designs with varying NA and underfilling of the
+  back aperture. See the caption for Fig. \ref{fig:single-arm} for an
+  explanation of a)-d).}
+\label{fig:double-arm}
+\end{figure}
+
+\begin{itemize}
+\item Figure \ref{fig:asymmetric-double} shows the results when we considered
+  asymmetric light-sheet designs. We kept the two arms orthogonal and used light-sheet
+  illumination. We swept the NA asymmetry and the dose asymmetry and found that the
+  best results occur for symmetric designs. 
+\item There is a tradeoff between solid-angle uncertainty mean and variance. You
+  can slightly reduce the mean or variance by using an asymmetric design, but
+  you pay for it. Reducing the mean increases the variance and vice-versa.
+\item The best dual arm microscope design is a symmetric light-sheet
+  illumination microscope with the largest possible detection NA.
+\end{itemize}
+
+\begin{figure}[htbp]
+\centering\includegraphics[width=\textwidth]{asymmetric-double}
+\caption{Asymmetric light-sheet illumination designs with varying NA and
+    dose asymmetry. See the Fig. \ref{fig:single-arm} caption for an explanation of a)-d).}
+\label{fig:asymmetric-double}
+\end{figure}
+
 \subsection{Three-Arm Designs}
+\begin{itemize}
+\item Figure \ref{fig:triple-arm} shows the results when we considered
+  symmetric light-sheet designs with a lower detection arm. We swept
+  through the NA of the upper arms and NA of the lower arm.
+\item The results show that the best designs have the largest possible detection
+  upper and lower NA.
+\item Note to Patrick: I used the same efficiencies for the upper and lower
+  arms. I'm still thinking/unsure about the factor of $\sim$1/9 due to rolling
+  shutter on the lower objective. If we do end up using this factor it will
+  weight the importance of a large NA on the upper objectives.
+\end{itemize}
+
+\begin{figure}[htbp]
+\centering\includegraphics[width=\textwidth]{triple-arm}
+\caption{Symmetric light-sheet illumination with lower detection arm and
+    varying upper and lower NA. See the Fig. \ref{fig:single-arm} caption for
+    an explanation of a)-d).}
+\label{fig:triple-arm}
+\end{figure}
 
 \section{Discussion}
+TODO
+% Fourkas' expressions underpredict the fraction of the total power collected by
+% an objective with a polarizer by a constant fraction of
+% $\frac{3}{2}$. Therefore, using Fourkas' expressions to find $I_{\text{tot}}$
+% will result in overpredictions of $I_{\text{tot}}$, but the incorrect
+% expressions will not cause any error in the predictions of molecular
+% orientation.
+
+% Lu and Bout \cite{lu2008} study the effect of noise on the reconstruction of single fluorophores using Fourkas' expressions.
+Their main results are unaffected by the correction introduced in these notes, but because these notes predict a higher intensity than Fourkas, Lu and Bout overpredict the effect of noise in their model.
+
+% As noted in \cite{lieb2004}, Fourkas' model only considers single molecules in
+% homogeneous environments.
 
 \section{Conclusion}
-After proofreading the manuscript, compress your .TEX manuscript file and all
-figures (which should be in EPS format, or PDF format if you are using
-PDF-\LaTeX) in a ZIP, TAR or TAR-GZIP package. Prism, OSA's article tracking
-system, will process in \LaTeX mode by default but will use PDF-\LaTeX if PDF
-figure files are detected. Note: TAR or TAR-GZIP is no longer required. All
-files must be referenced at the root level (e.g., file \texttt{figure-1.eps},
-not \texttt{/myfigs/figure-1.eps}). If there is video or other multimedia, the
-associated files should be uploaded separately.
+TODO
 
@@
 TODO
+%We thank Hari Shroff and Sean Rose for helpful conversations.
 
 \section*{Disclosures}
-The authors declare that there are no conflicts of interest related to this article.
+TODO
+%The authors declare that there are no conflicts of interest related to this article.
 

### 2017-07-20 paper/paper/paper.tex
Changed mean and std to median and MAD.

@@
 will have a small solid-angle uncertainty that is uniform for all fluorophore
-orientations. To summarize and compare microscope designs we use the mean and
-standard deviation of the solid-angle uncertainty at the sample points. Because
-the solid-angle uncertainty admits unbounded values, we ignore the highest
-percentile of the sample points and take the mean and standard deviation of the
-0-99th percentile.
+orientations. To summarize and compare microscope designs we use the median and
+median absolute deviation (MAD) of the solid-angle uncertainty.
 
@@
 \item We swept through the illumination and detection NA while keeping the
-  incident power ($I_{\text{tot}}$) constant and found that the lowest mean and
-  standard deviation of the solid-angle uncertainty occurs with a small
+  incident power ($I_{\text{tot}}$) constant and found that the lowest median and
+  MAD of the solid-angle uncertainty occurs with a small
   illumination NA and a large detection NA.
@@
   microscope. b) Solid angle uncertainty for the microscope in a) when
-  $I_{\text{tot}} = 1000$ photons. c) Mean of the solid-angle uncertainty as a
-  function of illumination and detection NA. d) Standard deviation of
-  solid-angle uncertainty as a function of illumination and detection NA. The
-  microscope in a) and b) is indicated by a cross in c) and d).}
+  $I_{\text{tot}} = 1000$ photons. c) Median of the solid-angle uncertainty as a
+  function of illumination and detection NA. d) MAD of solid-angle uncertainty
+  as a function of illumination and detection NA. The microscope in a) and b) is
+  indicated by a cross in c) and d).}
 \label{fig:single-arm}
@@
   considering steric constraints and while keeping the incident power constant.
-  We found that the lowest mean and standard deviation of the solid-angle
-  uncertainty occurs with orthogonal arms with the highest possible NA. 
+  We found that the lowest median of the solid-angle uncertainty occurs the largest
+  possible NA. We also found the increasing NA always lowers the MAD, but orthogonal
+  arms are not always best. At large NA, it is advantageous to move the objectives
+  as close or as far apart within steric constraints. 
 \end{itemize}
@@
   best results occur for symmetric designs. 
-\item There is a tradeoff between solid-angle uncertainty mean and variance. You
+\item There is a tradeoff between solid-angle uncertainty median and MAD. You
   can slightly reduce the mean or variance by using an asymmetric design, but
-  you pay for it. Reducing the mean increases the variance and vice-versa.
+  you pay for it. Reducing the median increases the MAD and vice-versa.
 \item The best dual arm microscope design is a symmetric light-sheet
@@
   weight the importance of a large NA on the upper objectives.
+\item There is a small dip in the MAD when the upper and lower objectives have
+  the same NA, but it is still better to use the largest possible lower
+  NA. 
 \end{itemize}

### 2017-07-20 paper/figures/asymmetric-double.py

@@ -7,14 +7,13 @@ import os; import time; start = time.time(); print('Running...')
 import matplotlib.gridspec as gridspec
 
 # Main input parameters
-col_labels = ['Geometry\n($\\alpha_{1}$ = 60${}^{\circ}$, Dose${}_1$/Dose${}_2$ = 10)', r'$\sigma_{\Omega}$ [sr]', 'Mean${}_{0-99}$$\{\sigma_{\Omega}\}$ [sr]', 'STD${}_{0-99}${$\{\sigma_{\Omega}\}$ [sr]']
+col_labels = ['Geometry\n($\\alpha_{1}$ = 60${}^{\circ}$, Dose${}_1$/Dose${}_2$ = 10)', r'$\sigma_{\Omega}$ [sr]', 'Median$\{\sigma_{\Omega}\}$ [sr]', 'MAD$\{\sigma_{\Omega}\}$ [sr]']
 fig_labels = ['a)', 'b)', 'c)', 'd)']
-n_pts = 1000 # Points on sphere
+n_pts = 1500 # Points on sphere
 n_pts_sphere = 50000 # Points on sphere
-n_grid_pts = 21
+n_grid_pts = 25
 inch_fig = 5
 dpi = 300
-percentile = 99
 
 # Setup figure and axes
 fig = plt.figure(figsize=(2.4*inch_fig, 2*inch_fig))
@@ -59,7 +58,7 @@ def is_feasible(pt):
 pts_list = [pt for pt in pts if is_feasible(pt)]
 pts = np.array(pts_list).T
 
-# Calculate mean and variance for each point
+# Calculate med and mad for each point
 def calc_stats(param):
     na1 = n*np.sin(np.deg2rad(param[1]))
     na2 = n*np.sin(np.pi/2 - np.arcsin(na1/n))
@@ -77,16 +76,17 @@ def calc_stats(param):
                                       n_pts=n_pts, max_photons=[dose1, dose2], n_samp=1.33)
 
     exp.calc_estimation_stats()
-    data = exp.sa_uncert[exp.sa_uncert < np.percentile(exp.sa_uncert, percentile)]
-    return np.mean(data), np.std(data)
+    data = exp.sa_uncert
+    med = np.median(data)
+    return med, np.median(np.abs(data - med))
 
-mean = []
-std = []
+med = []
+mad = []
 for i, pt in enumerate(pts.T):
     print('Calculating microscope '+str(i+1)+'/'+str(pts.shape[1]))
     x = calc_stats(pt)
[…]

### 2017-07-20 paper/figures/double-arm.py

@@ -7,14 +7,13 @@ import os; import time; start = time.time(); print('Running...')
 import matplotlib.gridspec as gridspec
 
 # Main input parameters
-col_labels = ['Geometry\n(NA${}_{\\textrm{ill}}$ = 0.6, NA${}_{\\textrm{det}}$ = 0.8)', r'$\sigma_{\Omega}$ [sr]', 'Mean${}_{0-99}$$\{\sigma_{\Omega}\}$ [sr]', 'STD${}_{0-99}$$\{\sigma_{\Omega}\}$ [sr]']
+col_labels = ['Geometry\n(NA${}_{\\textrm{ill}}$ = 0.6, NA${}_{\\textrm{det}}$ = 0.8)', r'$\sigma_{\Omega}$ [sr]', 'Median$\{\sigma_{\Omega}\}$ [sr]', 'MAD$\{\sigma_{\Omega}\}$ [sr]']
 fig_labels = ['a)', 'b)', 'c)', 'd)']
-n_pts = 100 # Points on sphere
-n_pts_sphere = 25000 # Points on sphere
+n_pts = 500 # Points on sphere
+n_pts_sphere = 50000 # Points on sphere
 n_grid_pts = 25
 inch_fig = 5
 dpi = 300
-percentile = 99
 
 # Setup figure and axes
 fig = plt.figure(figsize=(2.2*inch_fig, 2*inch_fig))
@@ -58,7 +57,7 @@ def is_feasible(pt):
 pts_list = [pt for pt in pts if is_feasible(pt)]
 pts = np.array(pts_list).T
 
-# Calculate mean and variance for each point
+# Calculate med and mad for each point
 def calc_stats(param):
     na_ill = param[0]
     na_det = param[1]
@@ -69,16 +68,17 @@ def calc_stats(param):
                                       n_pts=n_pts, max_photons=500, n_samp=1.33)
 
     exp.calc_estimation_stats()
-    data = exp.sa_uncert[exp.sa_uncert < np.percentile(exp.sa_uncert, percentile)]
-    return np.mean(data), np.std(data)
+    data = exp.sa_uncert
+    med = np.median(data)
+    return med, np.median(np.abs(data - med))
 
-mean = []
-std = []
+med = []
+mad = []
 for i, pt in enumerate(pts.T):
     print('Calculating microscope '+str(i+1)+'/'+str(pts.shape[1]))
     x = calc_stats(pt)
-    mean.append(x[0])
-    std.append(x[1])
+    med.append(x[0])
[…]

### 2017-07-20 paper/figures/single-arm.py

@@ -7,14 +7,13 @@ import os; import time; start = time.time(); print('Running...')
 import matplotlib.gridspec as gridspec
 
 # Main input parameters
-col_labels = ['Geometry\n(NA${}_{\\textrm{ill}}$ = 0.6, NA${}_{\\textrm{det}}$ = 0.8)', r'$\sigma_{\Omega}$ [sr]', 'Mean${}_{0-99}$$\{\sigma_{\Omega}\}$ [sr]', 'STD${}_{0-99}$$\{\sigma_{\Omega}\}$ [sr]']
+col_labels = ['Geometry\n(NA${}_{\\textrm{ill}}$ = 0.6, NA${}_{\\textrm{det}}$ = 0.8)', r'$\sigma_{\Omega}$ [sr]', 'Median$\{\sigma_{\Omega}\}$ [sr]', 'MAD$\{\sigma_{\Omega}\}$ [sr]']
 fig_labels = ['a)', 'b)', 'c)', 'd)']
-n_pts = 100 # Points on sphere
-n_pts_sphere = 25000 # Points on sphere
+n_pts = 1000 # Points on sphere
+n_pts_sphere = 50000 # Points on sphere
 n_grid_pts = 25
 inch_fig = 5
 dpi = 300
-percentile = 99
 
 # Setup figure and axes
 fig = plt.figure(figsize=(2.2*inch_fig, 2*inch_fig))
@@ -58,7 +57,7 @@ def is_feasible(pt):
 pts_list = [pt for pt in pts if is_feasible(pt)]
 pts = np.array(pts_list).T
 
-# Calculate mean and variance for each point
+# Calculate med and mad for each point
 def calc_stats(param):
     na_ill = param[0]
     na_det = param[1]
@@ -69,16 +68,17 @@ def calc_stats(param):
                                       n_pts=n_pts, max_photons=1000, n_samp=1.33)
 
     exp.calc_estimation_stats()
-    data = exp.sa_uncert[exp.sa_uncert < np.percentile(exp.sa_uncert, percentile)]
-    return np.mean(data), np.std(data)
+    data = exp.sa_uncert
+    med = np.median(data)
+    return med, np.median(np.abs(data - med))
 
-mean = []
-std = []
+med = []
+mad = []
 for i, pt in enumerate(pts.T):
     print('Calculating microscope '+str(i+1)+'/'+str(pts.shape[1]))
     x = calc_stats(pt)
-    mean.append(x[0])
[…]

### 2017-07-20 paper/figures/symmetric-widefield.py

@@ -7,15 +7,14 @@ import os; import time; start = time.time(); print('Running...')
 import matplotlib.gridspec as gridspec
 
 # Main input parameters
-col_labels = ['Geometry (NA = 0.6, $\\beta=80{}^{\circ}$)', r'$\sigma_{\Omega}$ [sr]', 'Mean${}_{0-99}$$\{\sigma_{\Omega}\}$ [sr]', 'STD${}_{0-99}$$\{\sigma_{\Omega}\}$ [sr]']
+col_labels = ['Geometry (NA = 0.6, $\\beta=80{}^{\circ}$)', r'$\sigma_{\Omega}$ [sr]', 'Median$\{\sigma_{\Omega}\}$ [sr]', 'MAD$\{\sigma_{\Omega}\}$ [sr]']
 fig_labels = ['a)', 'b)', 'c)', 'd)']
 n_pts = 500 # Points on sphere
-n_pts_sphere = 25000 # Points on sphere
+n_pts_sphere = 50000 # Points on sphere
 n_grid_pts = 21
 n_rows, n_cols = 1, len(col_labels)
 inch_fig = 5
 dpi = 300
-percentile = 99
 
 # Setup figure and axes
 fig = plt.figure(figsize=(2.2*inch_fig, 2*inch_fig))
@@ -40,7 +39,6 @@ for ax, col_label, fig_label  in zip([ax0, ax1, ax2, ax3], col_labels, fig_label
     ax.annotate(fig_label, xy=(0,0), xytext=(0, 1.05), textcoords='axes fraction',
                 va='center', ha='center', fontsize=14, annotation_clip=False)
     
-
 for ax in [ax0, ax1, ax2, ax3]:
     ax.tick_params(axis='both', labelsize=14)
 for cax in [cax1, cax2, cax3]:
@@ -64,7 +62,7 @@ def is_feasible(pt):
 pts_list = [pt for pt in pts if is_feasible(pt)]
 pts = np.array(pts_list).T
 
-# Calculate mean and variance for each point
+# Calculate med and mad for each point
 def calc_stats(param):
     na = param[0]
     angle = param[1]
@@ -75,16 +73,17 @@ def calc_stats(param):
                                       n_pts=n_pts, max_photons=500, n_samp=1.33)
 
     exp.calc_estimation_stats()
-    data = exp.sa_uncert[exp.sa_uncert < np.percentile(exp.sa_uncert, percentile)]
-    return np.mean(data), np.std(data)
+    data = exp.sa_uncert
[…]

### 2017-07-20 paper/figures/triple-arm.py

@@ -7,14 +7,13 @@ import os; import time; start = time.time(); print('Running...')
 import matplotlib.gridspec as gridspec
 
 # Main input parameters
-col_labels = ['Geometry\n(NA${}_{\\textrm{upper}}$ = 0.6, NA${}_{\\textrm{lower}}$ = 1.1)', r'$\sigma_{\Omega}$ [sr]', 'Mean${}_{0-99}$$\{\sigma_{\Omega}\}$ [sr]', 'STD${}_{0-99}$$\{\sigma_{\Omega}\}$ [sr]']
+col_labels = ['Geometry\n(NA${}_{\\textrm{upper}}$ = 0.6, NA${}_{\\textrm{lower}}$ = 1.1)', r'$\sigma_{\Omega}$ [sr]', 'Median$\{\sigma_{\Omega}\}$ [sr]', 'MAD$\{\sigma_{\Omega}\}$ [sr]']
 fig_labels = ['a)', 'b)', 'c)', 'd)']
-n_pts = 100 # Points on sphere
+n_pts = 1000 # Points on sphere
 n_pts_sphere = 50000 # Points on sphere
 n_grid_pts = 25
 inch_fig = 5
 dpi = 300
-percentile = 99
 
 # Setup figure and axes
 fig = plt.figure(figsize=(2.2*inch_fig, 2*inch_fig))
@@ -57,7 +56,7 @@ def is_feasible(pt):
 pts_list = [pt for pt in pts if is_feasible(pt)]
 pts = np.array(pts_list).T
 
-# Calculate mean and variance for each point
+# Calculate med and mad for each point
 def calc_stats(param):
     na_upper = param[0]
     na_lower = param[1]
@@ -68,16 +67,17 @@ def calc_stats(param):
                                     n_pts=n_pts, max_photons=500, n_samp=1.33)
 
     exp.calc_estimation_stats()
-    data = exp.sa_uncert[exp.sa_uncert < np.percentile(exp.sa_uncert, percentile)]
-    return np.mean(data), np.std(data)
+    data = exp.sa_uncert
+    med = np.median(data)
+    return med, np.median(np.abs(data - med))
 
-mean = []
-std = []
+med = []
+mad = []
 for i, pt in enumerate(pts.T):
     print('Calculating microscope '+str(i+1)+'/'+str(pts.shape[1]))
     x = calc_stats(pt)
-    mean.append(x[0])
-    std.append(x[1])
+    med.append(x[0])
[…]

### 2017-08-02 paper/paper/paper.tex
Paper draft to Patrick.

--- 2017-07-07 paper/paper/paper.tex
@@
 \usepackage[OE]{express}
+\usepackage{float}
 \providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
@@
 \begin{document}
-\title{Polarized Illumination For Determining Single-Molecule Orientation:
-  Modeling and Design Comparison}
+\title{Single molecule orientation determination with multiview polarized
+  illumination: modeling and microscope design}
 
-\author{Talon Chandler,\authormark{1,*} Shalin Mehta,\authormark{2} Rudolf Oldenbourg,\authormark{2, 3} and Patrick J. La Rivi\`ere\authormark{1}}
+\author{Talon Chandler,\authormark{1,*} Shalin Mehta,\authormark{2} Rudolf
+  Oldenbourg,\authormark{2, 3} and Patrick J. La Rivi\`ere\authormark{1}}
 
@@
 \begin{abstract*}
-Abstract here. 
+  We investigate the use of polarized illumination in multiview microscopes for
+  determining the orientation of single molecules. First, we relate the
+  orientation of single molecules to measurable intensities in multiview
+  microscopes and develop an information theoretic metric---the solid angle
+  uncertainty---to compare the ability of multiview microscopes to estimate the
+  orientation of single molecules. Next, we compare a broad class of microscopes
+  using this metric---single- and dual-view microscopes with varying
+  illumination polarization, illumination numerical aperture (NA), detection NA,
+  obliquity, asymmetry, and exposure. We find that multiview microscopes are
+  better at determining the orientation of single molecules than single-view
+  microscopes. We also find that choosing a small illumination NA and a large
+  detection NA are good design choices, multiview widefield microscopes can
+  benefit from oblique illumination and detection, and asymmetric NA microscopes
+  can benefit from exposure asymmetry.
 \end{abstract*}
 
-\ocis{(180.0180) Microscopy; (260.5430) Polarization; (110.0110) Imaging systems; (180.2520) Fluorescence microscopy; (180.6900) Three-dimensional microscopy.}
+\ocis{(110.0110) Imaging systems; (180.2520) Fluorescence microscopy; (180.0180)
+  Microscopy; (180.6900) Three-dimensional microscopy; (130.5440)
+  Polarization-selective devices; (260.5430) Polarization.}
 
 %%%%%%%%%%%%%%%%%%%%%%% References %%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{thebibliography}{99}
-\bibitem{gallo99} K. Gallo and G. Assanto, ``All-optical diode based on second-harmonic generation in an asymmetric waveguide,'' \josab {\bfseries 16}(2), 267--269 (1999).
-\end{thebibliography}
+%% Comment before submission
+\bibliography{paper}
+\bibliographystyle{osajnl}
+
+%% Copy .bbl and uncomment these before submission
+% \begin{thebibliography}{99}
+% \bibitem{gallo99} K. Gallo and G. Assanto, ``All-optical diode based on second-harmonic generation in an asymmetric waveguide,'' \josab {\bfseries 16}(2), 267--269 (1999).
+% \end{thebibliography}
 
 \section{Introduction}
-Single molecule orientation determination is important because
+The orientation of single molecules is a valuable reporter for biological
+processes. By imaging biological samples with fluorescent probes that are fixed
+to known structures, researchers have studied the dynamics of motor proteins
+\cite{peterman2001, forkey2003}, DNA \cite{backer2016}, actin \cite{mehta2016},
+septin \cite{demay2011, mcquilken2017}, and membranes \cite{anantharam2010}. The
+number of techniques available to biologists for measuring the orientation of
+single molecules in live cells is continually expanding. We follow a recent
+review \cite{backlund2014} and compare three major categories of single molecule
+orientation determination techniques:
 
-A variety of techniques have been tested to determine the orientation of
-single molecules:
+\textbf{Spatial variation of emission} techniques image single molecules and use
+the spatial distribution of intensity in the back focal plane \cite{lieb2004} or
+in the image plane \cite{backer2014} to estimate their orientation. The best
+techniques in this category use defocusing or phase masks in the back focal
+plane to encode orientation and position information in the image
+\cite{agrawal2012}. These techniques can estimate the orientation of
+molecules in all orientations and only require a single frame, but they
+require complex reconstruction algorithms, expensive high numerical aperture
+(NA) optics, and sensitive detectors.
 
-polarized detection [Fourkas], shape analysis [Moerner], other things from
-Moerner review.
+\textbf{Spatial variation of illumination polarization} techniques vary the
+polarization and intensity in the illumination path to excite molecules in
+specific orientations in the focal volume \cite{debarre2004}. Although these
+techniques can accurately estimate the orientation of molecules in all
+orientations, they require scanning which makes these techniques too slow for
+live cell imaging.
 
-We propose polarized illumination as an complementary method to polarized
-detection. We get the speed (no shape analysis required). We don't waste any
-dose (polarized light throws away light dose).
+\textbf{Polarized illumination or detection} techniques exploit the anisotropic
+absorption and emission pattern of single molecules to estimate their
+orientation. By measuring the total intensity in the detector plane during
+several polarization orientations on the illumination or detection path, the
+orientation of the molecule can be estimated. These techniques are easy to
+implement---changing the illumination or detection polarization is simple with a
+polarization splitter \cite{mehta2016} or universal compensator
+\cite{oldenbourg1995}---and the reconstruction methods are straightforward
+\cite{fourkas2001, mehta2016, backer2016}. The main drawback is speed---these
+techniques require the collection of several polarization frames to find the
+orientation of a single molecule. Despite the requirement for multiple frames,
+polarized illumination is the method of choice for live cell imaging because of
+its good combination of speed and ease of implementation. For these reasons,
+polarized illumination is the method we consider and extend in this paper.
 
-In this work we make three contributions:
-\begin{itemize}
-\item we develop the polarized illumination forward model for a broad class of
-  microscopes (section 2.X-2.X)
-\item we develop metrics to compare the performance of polarized illumination
-  microscopes 
-\item we use these metrics to compare microscopes design and make recommendations
-  for polarized illumination microscopes. 
-\end{itemize}
+All single molecule orientation determination methods suffer from some degree of
+anisotropic orientation uncertainty, i.e., molecules in some orientations
+cannot be oriented as well as molecules in others. In some cases the
+orientation cannot be determined at all---the forward model can be degenerate
+for specific dipole orientations \cite{fourkas2001, lu2008}. We feel that the
+importance of isotropic orientation uncertainty has been underappreciated. To
+our knowledge, the only authors who have considered anisotropic orientation
+uncertainty have only analyzed a small set of molecule orientations intead of
+all orientations \cite{agrawal2012}. Our view is that an ideal technique for
+determining the orientation of single molecules can reconstruct the
+orientation with a small and nearly uniform uncertainty for all molecule
+orientations.
 
-\section{Methods}
-\subsection{Illumination Model}
+Recently, there has been increasing interest in multiview microscopy techniques
+for biological imaging \cite{wu2013, keller2015, wu2016, wu2017}. Multiview
+microscopes address two key limitations of single view microscopes. First,
+multiview microscopes can achieve nearly isotropic resolution compared to the
+poor axial resolution of single view microscopes. Second, multiview microscopes
+can easily incorporate selective illumination to reduce phototoxicity. For
+example, compare light-sheet illumination to wide-field illumination where
+most of the illumination light excites out of focus molecules. These
+advantages make multiview microscopes good candidates for imaging live
+biological specimens for long periods.
+
+In this work we explore the use of polarized illumination in multiview
+microscopes for determining the orientation of single molecules. Existing
+multiview microscopes can easily be outfitted with fast-switching polarizers
+that do not degrade image quality, so polarized illumination is a natural way to
+augment multiview microscopes for measuring the orientation of single
+molecules. Furthermore, multiview microscopes can achieve nearly isotropic
+resolution while delivering selective illumination. We hypothesize that
+multiview microscopes will also provide a small and uniform orientation
+uncertainty for single molecules in all orientations. In section \ref{methods}
+we develop the required theory for polarized illumination microscopy and develop
+information theoretic metrics to compare multiview microscopes. In section
+\ref{results} we compare the results for a wide range of multiview microscope
+designs. Finally, in section \ref{discussion} we discuss the results and their
+impact on polarized multiview microscope design.
+
+\section{Methods}\label{methods}
+In this section we develop a method to compare multiview polarized illumination
+microscopes for the task of reconstructing molecular orientations. In sections
+\ref{excitation}--\ref{forward} we develop the forward model for single-frame
+microscopes with a fixed illumination and detection configuration. In section
+\ref{designs} we list a variety of ways that single-frame microscopes can be
+combined to create experimentally realizable multi-frame microscopes. Finally,
+in section \ref{metrics} we develop metrics that we will use to compare
+multi-frame microscopes.
+
+\subsection{Absorption Efficiency}\label{excitation}
+Consider polarized widefield K\"ohler illumination incident on a single
+molecule at the focal point with the optical axis along the $\mh{z}$ axis. In
+this section we will calculate the \emph{absorption efficiency} in this
+geometry---the fraction of the incident power that is absorbed by the molecule.
+
+We start with the unit electric field transmitted by the polarizer,
 \begin{align}
-  \eta_{\text{det},\phi_{\text{det}}}  = \frac{\int_{\Omega}d\mh{r}\ \left|\mh{P}_{\text{det}}\cdot\tilde{\mb{R}}(\mh{r})(\mh{r}\times\hat{\bs{\mu}}_{\text{em}}\times\mh{r})\right|^2}{\int_{\mathbb{S}^2}d\mh{r}\left|\mh{r}\times\hat{\bs{\mu}}_{\text{em}}\times\mh{r}\right|^2}\label{eq:final}
+ \hat{\mb{E}}_{\text{exc}} = \cos\phi_{\text{exc}}\hat{\mb{x}} +
+\sin\phi_{\text{exc}}\hat{\mb{y}},\label{eq:excfield}
+\end{align}
+where $\phi_{\text{exc}}$ is the angle between the transmission axis of the
+excitation polarizer and the $\mh{x}$ axis. Next, we model the action of the an
+ideal, aplanatic, polarization preserving illumination objective by multiplying
+the electric field vector by a direction dependent rotation matrix,
+  \begin{align}
+  \mb{R}(\mh{r}) &= \begin{bmatrix} \cos\theta\cos^2\phi + \sin^2\phi & (\cos\theta -1)\sin\phi\cos\phi & -\sin\theta\cos\phi\\ (\cos\theta - 1)\sin\phi\cos\phi & \cos\theta\sin^2\phi + \cos^2\phi & -\sin\theta\sin\phi \\ \sin\theta\cos\phi& \sin\theta\sin\phi & \cos\theta \end{bmatrix}
   \end{align}
@@
   \begin{align}
-  \hat{\mb{r}} &= \sin\theta\cos\phi\hat{\mb{i}} + \sin\theta\sin\phi\hat{\mb{j}} + \cos\theta\hat{\mb{k}}\label{eq:r_coords}\ \ \ \text{is the dummy integration vector,}\\
-  \hat{\bs{\mu}}_{\text{em}} &= \sin\Theta\cos\Phi\hat{\mb{i}} + \sin\Theta\sin\Phi\hat{\mb{j}} + \cos\Theta\hat{\mb{k}}\ \ \ \text{is the emission dipole moment,}\label{eq:mu_coords}\\
-  \tilde{\mb{R}}(\mh{r}) &= \begin{bmatrix} \cos\theta\cos^2\phi + \sin^2\phi & (\cos\theta -1)\sin\phi\cos\phi & -\sin\theta\cos\phi\\ (\cos\theta - 1)\sin\phi\cos\phi & \cos\theta\sin^2\phi + \cos^2\phi & -\sin\theta\sin\phi \\ \sin\theta\cos\phi& \sin\theta\sin\phi & \cos\theta \end{bmatrix}\label{eq:matrix}\\&\text{is the rotation matrix that models the objective lens,}\\
-  \hat{\mb{P}}_{\text{det}} &= \cos\phi_{\text{det}}\hat{\mb{i}} + \sin\phi_{\text{det}}\hat{\mb{j}}\ \ \ \text{is the pass axis of the polarizer,}\\
-  \Omega &= \{(\phi, \theta)|\phi \in (0,2\pi]\ \text{and}\ \theta\in (0, \alpha]\}\ \ \ \text{is the cap of the sphere collected by the objective.}
+   \hat{\mb{r}} = \sin\theta\cos\phi\hat{\mb{x}} + \sin\theta\sin\phi\hat{\mb{y}} + \cos\theta\hat{\mb{z}} 
+  \end{align}
+  is a unit vector that points from the origin to the point where the incident
+  electric field intersects the Gaussian sphere of the illumination
+  objective. Next, we find the power that excites the molecule by taking the
+  dot product of the rotated incident field $\mb{R}(\mh{r})\mh{E}_{\text{exc}}$
+  and the absorption dipole moment
+\begin{align}
+  \hat{\bs{\mu}}_{\text{abs}} = \sin\Theta\cos\Phi\hat{\mb{x}} + \sin\Theta\sin\Phi\hat{\mb{y}} + \cos\Theta\hat{\mb{z}}
+\end{align}
+then taking its modulus squared. If the incident fields are spatially incoherent
+then we can find the total power that excites the molecule by integrating over the
+illuminated region of the Gaussian sphere $\Omega$ where
+\begin{align}
+  \Omega &= \{\phi, \theta\ |\ 0 < \phi \leq 2\pi, 0 < \theta \leq \alpha\}, \\
+  \alpha &= \arcsin\left(\frac{\text{NA}}{n}\right), \label{eq:alpha}
+\end{align}
+NA is the illumination numerical aperture, and $n$ is the index of refraction of
+the medium.
+
+Finally, we find the fraction of the incident power absorbed by the molecule
+by dividing the total absorbed power by the total incident power. The complete
+expression for the absorption efficiency in vector notation is
+\begin{align}
+  \eta_{\text{abs}} = \frac{\int_{\Omega}d\mh{r}|\hat{\bs{\mu}}_{\text{abs}}\cdot\mb{R}(\mh{r})\mh{E}_{\text{exc}}|^2}{\int_{\Omega}d\mh{r}}\label{eq:abs}. 
 \end{align}
 
+We substitute equations \ref{eq:excfield}--\ref{eq:alpha} into equation
+\ref{eq:abs}, evaluate the integrals, and simplify to express the absorption
+efficiency in scalar notation as
 \begin{align}
-  \eta_{\text{exc},\phi_{\text{pol}}}(\Theta, \Phi, \alpha) &= D\{A + B\sin^{2}{\Theta} + C\sin^{2}{\Theta} \cos{[2 (\Phi - \phi_{\text{pol}}})]\}.
+  \eta_{\text{abs}} &= D\{A + B\sin^{2}{\Theta} + C\sin^{2}{\Theta} \cos{[2 (\Phi - \phi_{\text{exc}}})]\}\label{eq:scalarabs}
 \end{align}
+where
 \begin{subequations}
@@
   D &= \frac{4}{3(1 - \cos\alpha)}.
-\end{align}%
+\end{align}\label{eq:coefficients}%
 \end{subequations}
 
-\subsection{Detection Model}
+At the beginning of this section we assumed that the molecule was located at
+the focal point of the condenser. If the illumination objective is aplanatic
+then we can relax this assumption and the above expressions are valid for
+molecules anywhere in the focal plane.
+
+We also assumed that we used incoherent K\"ohler illumination. If we illuminate
+the back aperture with a weakly focused laser beam and scan the laser beam
+slowly---if the scan velocity is much less than the beam waist divided by the
+coherence time---we can ignore longitudinal excitation and the molecule will
+be excited as if a single plane was incident. Therefore, we can find the
+absorption efficiency of a molecule under weakly focused scanned laser
+illumination by taking the limit of equation \ref{eq:scalarabs} as
+$\alpha \rightarrow 0$ giving
 \begin{align}
-  \eta_{\text{det}}(\Theta, \alpha) = 2(A + B\sin^2\Theta)
+  \eta_{\text{abs}} &= \sin^2\Theta\cos^2(\Phi - \phi_{\text{exc}})
 \end{align}
+which is recognizable as Malus' law generalized to three dimensions. Notice that
+the absorption efficiency can take its a maximum range of values
+$\eta_{\text{abs}} \in [0, 1]$ when $\alpha=0$.
 
-\subsection{Complete Forward Model}
+\subsection{Detection Efficiency}\label{detection}
+Fourkas calculated the detection efficiency of a single molecule when an
+objective with a polarizer is aligned along the $\mh{z}$ axis
+\cite{fourkas2001}. We use his expressions to calculate the detection efficiency
+without a polarizer as
 \begin{align}
-  I = I_{\text{tot}}\eta_{\text{exc}}\eta_{\text{det}}
+  \eta_{\text{det}} = 2(A + B\sin^2\Theta). \label{eq:scalardet}
 \end{align}
+The detection efficiency only depends on $\Theta$, not $\Phi$, because there is
+no detection polarizer. We have assumed that the absorption dipole moment is
+equal to the emission dipole moment.
+
+Note that $A$ and $B$ in equation \ref{eq:coefficients} are a factor of
+$\frac{3}{2}$ larger than the expressions given by Fourkas. We found that
+Fourkas' expressions were incorrectly normalized (the limit of
+$\eta_{\text{det}}$ as $\alpha\rightarrow \pi/2$ should be 1), and the extra
+factor of $\frac{3}{2}$ corrects the error. This correction does not affect
+Fourkas' orientation reconstruction but it does mean that Fourkas' algorithm
+under predicts the total emitted intensity by a factor of $\frac{3}{2}$.
+
+Also note that the detection efficiency can take its a maximum range of values
+when $B$ takes its maximum value, i.e. when
+$\alpha=\arccos\left(\frac{1}{\sqrt{3}}\right) \approx 54.7^{\circ}$. When $n = 1.33$,
+this condition corresponds to $\text{NA} \approx 1.1$.
+
+\subsection{Oblique Illumination and Detection}\label{oblique}
+In sections \ref{excitation} and \ref{detection} we assumed that both the
+illumination and detection objectives had $\mh{z}$ aligned optical axes. To extend
+the forward model to oblique optical axes, we express the dipole orientation in
+rotated coordinates using the following expressions
+\begin{align}
+    \Theta' &= \arccos\left(\sin\psi\cos\Phi\sin\Theta + \cos\psi\cos\Theta\right)\label{eq:thetap}\\
+  \Phi' &=
+          \begin{cases}
+            \arccos\left(\dfrac{\cos\psi\cos\Phi\sin\Theta - \sin\psi\cos\Theta}{\sqrt{1 - (\sin\psi\cos\Phi\sin\Theta + \cos\psi\cos\Theta)^2}}\right) \ \ \ &0 \leq \Phi < \pi  \\
+            -\arccos\left(\dfrac{\cos\psi\cos\Phi\sin\Theta - \sin\psi\cos\Theta}{\sqrt{1 - (\sin\psi\cos\Phi\sin\Theta + \cos\psi\cos\Theta)^2}}\right) \ \ \ &-\pi \leq \Phi < 0\\
+          \end{cases}
+\end{align}
+where $\psi$ is the angle of a right handed rotation about the $\mh{y}$ that maps the $\mh{z}$ axis onto the new optical axis.
+
+\subsection{Single-Frame Microscopes}\label{forward}
+The detected intensity is the product of the the sample exposure
+$I_{\text{tot}}$, the absorption efficiency, and the detection efficiency
+\begin{align}
+  I = I_{\text{tot}}\eta_{\text{abs}}\eta_{\text{det}}\label{eq:forward_frame}.
+\end{align}
+Equation \ref{eq:forward_frame} is the forward model for a \emph{single-frame
+  microscope}. Figure \ref{fig:single-frame} shows the efficiencies from three
+representative examples of single-frame microscopes.
+
+The intensity measured by a single-frame microscope does not give us enough
+information to reconstruct the orientation of a single molecule. Next, we
+will consider combining several single-frame microscopes to create
+\emph{multi-frame microscopes.}
+
+Note our distinction between \emph{frames} and \emph{views}. Frames refer to
+single intensity measurements with fixed illumination, detection, and
+polarization geometries. Views refers to the detection objectives, and multiview
+microscopes use multiple detection objectives.
+
+\begin{figure}[H]
+\centering\includegraphics[width=\textwidth]{single-frame}
+\caption{Representative examples of single-frame microscopes.\newline \newline
+  \textbf{Columns left to right:} 1) schematics of single-frame microscopes
+  where the solid line encloses the illumination solid angle, the dashed line
+  encloses the detection solid angle, and the arrow indicates the transmission
+  axis of the illumination polarizer; 2) the absorption efficiency, see equation
+  \ref{eq:scalarabs}; 3) the detection efficiency, see equation
+  \ref{eq:scalardet}; (4) the total efficiency, the product of the absorption and
+  detection efficiencies.\newline \newline \textbf{Rows top to bottom:} 1)
+  epi-illumination ($\text{NA} = 1.1$) with $x$-polarized light and
+  epi-detection (NA = 1.1); 2) orthogonal illumination ($\text{NA} = 0.8$) and
+  detection ($\text{NA} = 0.8$); 3) oblique illumination ($\text{NA} = 0.8$) and
+  detection ($\text{NA} = 0.8$). All simulations use $n=1.33$.}
+  \label{fig:single-frame}
+\end{figure}
+
+\subsection{Multi-Frame Microscopes}\label{designs}
+One way to collect multiple frames is to add a universal compensator to the
+illumination arm and rapidly select the incident polarization by changing
+$\phi_{\text{exc}}$ \cite{oldenbourg1995}. All of the multi-frame microscopes we
+will consider in this paper use four frames per view with the illumination
+polarization separated by 45${}^{\circ}$.
+
+We also consider several dual-view designs. We evaluate two-view designs that
+allow for illumination and detection from both objectives (see \cite{wu2013} for
+a dual-view design). In all cases we consider the effect of varying the
+illumination and detection numerical aperture. We also consider asymmetric NA
+designs (see \cite{wu2017} for an asymmetric NA design) and the effect of
+sample exposure.
+
+All multiview microscopes are subject to steric constraints---the objectives
+must not collide with the cover slip or each other. We only consider microscope
+designs that meet these criteria.
+
+\subsection{Evaluation Metrics}\label{metrics}
+Our goal is to evaluate the ability of a microscope design to estimate the
+orientation of a single molecule (via parameters $\Theta$ and $\Phi$) from
+intensity data.
+
+For each microscope design and molecule orientation we calculate the Fisher
+information matrix. If the intensity is Poisson distributed the
+Fisher information matrix is given by
+\begin{align}
+  \mb{F} &= \sum_{k=1}^N \frac{1}{I_k}
+  \begin{bmatrix}
+    \dfrac{\partial I_k}{\partial \Theta}\dfrac{\partial I_k}{\partial \Theta}&\dfrac{\partial I_k}{\partial \Theta}\dfrac{\partial I_k}{\partial \Phi}\\\\
+    \dfrac{\partial I_k}{\partial \Theta}\dfrac{\partial I_k}{\partial \Phi}&\dfrac{\partial I_k}{\partial \Phi}\dfrac{\partial I_k}{\partial \Phi}\\    
+  \end{bmatrix}
+\end{align}
+where $I_k$ is the intensity measured in the $k$th frame of an $N$ frame
+microscope.
+
+A common way to evaluate the ability to estimate the parameters $\Theta$ and
+$\Phi$ from the data is to calculate the Cramer-Rao lower bound (CRLB) for each
+parameter \cite{kay1993}. The CRLBs are given by the diagonal elements of the
+inverse of the Fisher information matrix, and they give the minimum variance of
+an unbiased estimator for each parameter. Agrawal et. al. used the square root
+of the product of the CRLBs multiplied by the Jacobian determinant,
+$\sin\Theta\sqrt{\mb{F}^{-1}_{0,0}\mb{F}^{-1}_{1,1}}$, to find the area of
+uncertainty in parameter space \cite{agrawal2012}.
+
+CRLBs and the associated area of uncertainty are parametrization dependent,
+i.e., if we choose a different coordinate system these metrics will change. We
+would like to compare microscope designs without choosing a parametrization, so
+instead we use
+\begin{align}
+  \sigma_{\Omega} &\equiv \sin\Theta\sqrt{\text{det}\{\mb{F}^{-1}\}}
+\end{align}
+as our evaluation metric. We take the determinant of the inverse Fisher
+information matrix---a parametrization independent value called the
+\emph{generalized variance} \cite{anderson1958}---take the square root, then
+multiply by the Jacobian determinant, $\sin\Theta$. We call $\sigma_{\Omega}$
+the \emph{solid-angle uncertainty} because it has units of steradians and is a
+parametrization independent measure of the orientation uncertainty.
+
+For each multi-frame microscope we calculate $\sigma_{\Omega}$ at 10,000
+approximately equally spaced points on the unit sphere. A desirable microscope
+design will have a small solid-angle uncertainty that is uniform for all
+molecule orientations. The most straightforward way to find the location and
+scale of the solid-angle uncertainty is to use the mean and variance,
+respectively. Because the solid-angle uncertainty admits extremely large values
+that can change the mean and variance dramatically depending on the sample
+points, we use the median and median absolute deviation (MAD)---the median of
+the data's absolute difference from the median---as robust replacements for the
+mean and variance.
+
+To compare multi-frame microscopes with different numbers of frames fairly, we
+kept the total exposure to the sample constant by choosing
+$I_{\text{tot}} = \frac{4000}{N}$. Note that $I_{\text{tot}}$ is a measure of
+exposure to the sample, not the detected intensity or the total intensity
+emitted by the molecule.
+
+\section{Results}\label{results}
+Figure \ref{fig:single-arm} shows our results for single-view designs where a
+single objective is used for illumination and detection. We swept through the
+illumination and detection NA while keeping the sample exposure constant, and we
+found that the lowest median and MAD of the solid-angle uncertainty occurs with
+a small illumination NA and a large detection NA where
+$\alpha \approx 57.4^{\circ}$. A small illumination NA maximizes the range of
+absorption efficiencies, while a large detection NA maximizes the number of
+detected photons and the range of the detection efficiencies. Note the relative
+importance of illumination and detection NAs---reducing the illumination NA (an
+inexpensive modification) improves the solid angle uncertainty much less than
+increasing the detection NA (an expensive modification). Also note from Figure
+\ref{fig:single-arm}b) that single-view designs suffer from high orientation
+uncertainty when the molecules are oriented along the optical axis (because the
+molecules are not efficiently excited) and in the transverse plane (because it
+is ambiguous whether the molecule is oriented above or below the transverse
+plane).
 
 \begin{figure}[htbp]
-\centering\includegraphics[width=\textwidth]{single-frame}
-\caption{Representative examples of the forward model for single frame
-  microscopes.\newline \newline \textbf{Columns left to right:} (1) schematics of
-  single frame microscopes where the solid line encloses the illumination
-  solid angle, the dashed line encloses the detection solid angle, and the arrow
-  indicates the transmission axis of the illumination polarizer; (2) the excitation
-  efficiency see equation XX; (3) the detection efficiency see equation X; (4)
-  the total efficiency see equation X.\newline \newline \textbf{Rows top to
-    bottom:} (1) epi-illumination (NA = 1.1) with $x$-polarized light and
-  epi-detection (NA = 1.1); (2) orthogonal illumination (NA = 0.8) and detection
-  (NA = 0.8); (3) oblique illumination (NA = 0.8) and detection (NA = 0.8).}
+\centering\includegraphics[width=\textwidth]{single-arm}
+\caption{Single-view microscope with varying illumination and detection
+  NA. a) Schematic of a single-arm four-frame epi-illumination microscope. The
+  red arrows indicate the four illumination polarization orientations---one for
+  each frame. b) Solid angle uncertainty for the microscope in a). c) Median of
+  the solid-angle uncertainty as a function of illumination and detection NA. d)
+  MAD of solid-angle uncertainty as a function of illumination and detection
+  NA. The microscope in a) and b) is indicated by a cross in c) and d).}
+\label{fig:single-arm}
 \end{figure}
 
-\subsection{Microscope Designs}
-Widefield illumination vs light sheet illumination.
+Figures \ref{fig:symmetric-widefield} and \ref{fig:double-arm} show our results
+for dual-view designs. We illuminate from one objective and detect from the
+other for four polarization frames, then we repeat the polarization frames with
+the same objectives in reversed roles. In Figure \ref{fig:symmetric-widefield}
+we sweep through the NA of both objectives and the angle between the objectives
+while considering steric constraints. Note how adding a second view removes the
+high uncertainty region in the transverse plane. The dual-view microscope still
+has high uncertainty regions near the optical axes, but the uncertainty is
+reduced compared to the single-view microscope. We find that the lowest median
+of the solid-angle uncertainty occurs with the largest possible NA
+objectives. We also find that increasing NA always lowers the MAD, but
+orthogonal arms are not always best. At large NA, it is advantageous to move the
+objectives together or against the cover slip. Given the high uncertainty in the
+transverse plane for a single view shown in Figure \ref{fig:single-arm}b), we
+can see why oblique views perform better than orthogonal views---oblique views
+are complementary because the regions of high orientation of uncertainty from
+each view do not overlap.
 
-- Light sheet has low illumination. 
-- Restricted to orthogonal detection.
-
-\subsection{Evaluation Metrics}
-\begin{align}
-  F &= \sum_{i=0}^N \frac{1}{I}
-  \begin{bmatrix}
-    \frac{\partial I}{\partial \Theta}\frac{\partial I}{\partial \Theta}&\frac{\partial I}{\partial \Theta}\frac{\partial I}{\partial \Phi}\\\\
-    \frac{\partial I}{\partial \Theta}\frac{\partial I}{\partial \Phi}&\frac{\partial I}{\partial \Phi}\frac{\partial I}{\partial \Phi}\\    
-  \end{bmatrix}\\
-  \sigma_{\Omega} &= \frac{\sin\Theta}{\sqrt{\text{det}\{F\}}}
-\end{align}
-We call $\sigma_{\Omega}$ the \emph{solid-angle uncertainty}. 
-We sampled $\sigma_{\Omega}$ on 100,000 approximately equally
-spaced points on the unit sphere and calculated the \emph{coefficient of variation}
-$c_v$ of these points where
-\begin{align}
-  c_v = \frac{\sigma_{\sigma_{\Omega}}}{\mu_{\sigma_{\Omega}}}
-\end{align}
-
-\section{Results}
-\subsection{One-Arm Designs}
 \begin{figure}[htbp]
-\centering\includegraphics[width=\textwidth]{single-arm}
-\caption{\textbf{Left:} Schematic of a single-arm four-frame epi-illumination
-  microscope with NA = 0.8 and $n = 1.33$.  \textbf{Right:} Solid
-  angle uncertainty for the same microscope when $I_{\text{tot}} = 1000$
-  photons. Note the regions of high uncertainty near the equator and poles
-  as noted in XXX.}
+\centering\includegraphics[width=\textwidth]{symmetric-widefield}
+\caption{Dual-view symmetric widefield designs with varying NA and angle between
+  objectives. First we illuminate with the first objective (red solid) and
+  detect from the second objective (red dashed). Then we illuminate from the
+  second objective (blue solid) and detect from the first objective (blue
+  dashed). See the Fig. \ref{fig:single-arm} caption for an explanation of
+  b)-d).}
+\label{fig:symmetric-widefield}
 \end{figure}
 
-Can't use $c_v$ because it doesn't converge.
+Figure \ref{fig:double-arm} shows our results when we used a symmetric widefield
+design and varied the illumination and detection NA. Our results are similar to
+the single view case in Figure \ref{fig:single-arm}. We find that the best
+designs use a small illumination NA and a large detection NA where
+$\alpha \approx 57.4^{\circ}$. This means that dual view light sheet illumination
+geometries are an excellent choice for uniformly reconstructing the orientation
+of single molecules.
 
-High variance around equator and poles.
+\begin{figure}[htbp]
+\centering\includegraphics[width=\textwidth]{double-arm}
+\caption{Symmetric widefield designs with varying illumination and detection NA. See the
+  caption for Fig. \ref{fig:symmetric-widefield} for an explanation of a) and
+  the caption for Fig. \ref{fig:single-arm} for an explanation of b)-d).}
+\label{fig:double-arm}
+\end{figure}
 
-\subsection{Two-Arm Designs}
-Widefield vs light sheet.
+Figure \ref{fig:asymmetric-double} shows our results for asymmetric light sheet
+illumination designs. We kept the two views orthogonal and used light sheet
+illumination on both sides while sweeping through the NA asymmetry and the
+sample exposure asymmetry while keeping the total sample exposure constant. We
+found that symmetric designs are at a local minimum of the median and MAD of the
+solid-angle uncertainty. We found that at extreme asymmetries where the exposure
+from the low NA arm is much larger than the exposure from the high NA arm
+(top-left and bottom-right corners of Figures \ref{fig:asymmetric-double}c) and
+d)) the median and MAD are comparable to the symmetric light-sheet
+microscope. This means that if a very high NA objective is available and a low
+NA objective can provide light sheet illumination, this microscope will perform
+comparably with a dual-view microscope with two objectives of medium NA. Figure
+\ref{fig:asymmetric-double} also shows that designs with slightly asymmetric NA
+can trade off a low solid-angle uncertainty median for a low solid-angle
+uncertainty MAD by changing the sample exposure ratio.
 
-Widefield NA vs angle/.
+\begin{figure}[htbp]
+  \centering\includegraphics[width=\textwidth]{asymmetric-double}
+  \caption{Asymmetric light-sheet illumination designs with varying NA and
+    sample exposure asymmetry. See the Fig. \ref{fig:symmetric-widefield}
+    caption for an explanation of a) (note that the illumination NA is small so
+    the illumination is represented by a dot instead of a solid line) and the
+    Fig. \ref{fig:single-arm} caption for an explanation of b)-d).}
+\label{fig:asymmetric-double}
+\end{figure}
 
-Sweep asymmetry for full NA light sheet.
+% Figure \ref{fig:triple-arm} shows the results when we considered symmetric
+% light-sheet designs with a lower detection arm. We swept through the NA of the
+% upper arms and NA of the lower arm. The results show that the best designs have
+% the largest possible detection upper and lower NA. Note to Patrick: I used the
+% same efficiencies for the upper and lower arms. I'm still thinking/unsure about
+% the factor of $\sim$1/9 due to rolling shutter on the lower objective. If we do
+% end up using this factor it will weight the importance of a large NA on the
+% upper objectives. There is a small dip in the MAD when the upper and lower
+% objectives have the same NA, but it is still better to use the largest possible
+% lower NA.
 
-\subsection{Three-Arm Designs}
+% \begin{figure}[htbp]
+% \centering\includegraphics[width=\textwidth]{triple-arm}
+% \caption{Symmetric light-sheet illumination with lower detection arm and
+%     varying upper and lower NA. See the Fig. \ref{fig:single-arm} caption for
+%     an explanation of a)-d).}
+% \label{fig:triple-arm}
+% \end{figure}
 
-\section{Discussion}
+\section{Discussion and Conclusions}\label{discussion}
+% Fourkas' expressions underpredict the fraction of the total power collected by
+% an objective with a polarizer by a constant fraction of
+% $\frac{3}{2}$. Therefore, using Fourkas' expressions to find $I_{\text{tot}}$
+% will result in overpredictions of $I_{\text{tot}}$, but the incorrect
+% expressions will not cause any error in the predictions of molecular
+% orientation.
 
-\section{Conclusion}
+% Lu and Bout \cite{lu2008} study the effect of noise on the reconstruction of single fluorophores using Fourkas' expressions.
+Their main results are unaffected by the correction introduced in these notes, but because these notes predict a higher intensity than Fourkas, Lu and Bout overpredict the effect of noise in their model.
+
+In this work we have developed a model to predict the intensities measured by
+polarized illumination microscopes in a wide range of geometries, developed
+metrics to measure the ability of a microscope design to estimate the
+orientation of single molecules, and used these metrics to compare a wide
+range of microscope designs. Our main result is a short list of design
+heuristics that can be used to design polarized illumination single molecule
+orientation microscopes:
+\begin{itemize}
+\item All microscopes benefit from a small illumination NA and a large detection
+  NA with $\alpha \approx 57.4^{\circ}$.
+\item Dual-view microscopes outperform single-view microscopes, and dual-view
+  microscopes remove the high uncertainty region in the transverse plane of
+  single-view microscopes.
+\item High NA dual-view microscopes are best when the two views are
+  oblique. Oblique views are complementary because their high uncertainty
+  transverse planes do not overlap.
+\item Symmetric dual-view light-sheet illumination microscopes provide a good
+  compromise between cost, solid-angle uncertainty median, and solid-angle
+  uncertainty MAD. Asymmetric dual-view light-sheet illumination microscopes
+  can improve the solid-angle uncertainty median or MAD by changing the sample
+  exposure ratio between the two views.
+\end{itemize}
+
+Our results are limited in several ways. First, we only consider single
+molecules in homogeneous environments. We ignore light reflected from the cover
+slip and any other inhomogeneities in the sample. If we considered light
+reflected from the cover slip we expect that the results would put a lower
+weighting on large detection NAs because we would collect more light. Second, we
+only considered ideal, polarization preserving, aplanatic objectives. In
+practice, these conditions are not always satisfied and power from high NA rays
+is reduced by a factor of $\frac{1}{\cos\Theta}$ \cite{nov2006}. Therefore, our
+results are most accurate for low NA rays and get progressively worse for high
+NA rays, although we our major conclusions would not change. The main advantage
+of ignoring the power reduction of high NA rays is the tractability of the
+problem. We have provided approximate closed form solutions for the illumination
+and detection efficiencies that have allowed us to draw useful design
+conclusions.
+
+To compare the microscopes in this paper we used a fixed sample exposure. We
+think that this is the fairest way to compare microscopes with different numbers
+of frames. However, in cases where sample exposure and time are not an issue it
+may be more useful to compare microscopes with the same detection exposure per
+frame. This comparison would favor multiple frame microscopes even more that the
+results in this paper, because equal detection exposure per frame would improve
+photon counting statistics.
+
+We have considered single- and dual-view microscopes in a variety of cases, but
+more multiview microscopes can be analyzed with the framework we have
+developed. Wu et. al. have developed a three view design with a third objection
+below the cover slip\cite{wu2016}. Light-field detection schemes are also a type
+of multiview microscope\cite{levoy2006}, and we plan to extend our analysis to
+this case in future work.
+
+Although we have only considered single molecule orientation determination in
+this work, polarized illumination techniques can be applied to find the
+distribution of ensembles of molecules \cite{mehta2016, backer2016}. In future
+work we will investigate the ability of multiview polarized illumination
+microscopes to estimate the orientation and distribution of ensembles of
+molecules.
+
+We conclude that multiview microscopes are useful tools for determining the
+orientation of single molecules. Using simple design heuristics, we can design
+polarized illumination multiview microscopes that can determine the orientation
+of single molecules with a small and uniform orientation uncertainty.
 
@@
 \section*{Acknowledgments}
-We thank Hari Shroff and Sean Rose for helpful conversations.
+We thank Hari Shroff for helpful conversations.
 

### 2017-08-02 paper/figures/asymmetric-double.py

@@ -7,9 +7,9 @@ import os; import time; start = time.time(); print('Running...')
 import matplotlib.gridspec as gridspec
 
 # Main input parameters
-col_labels = ['Geometry\n($\\alpha_{1}$ = 60${}^{\circ}$, Dose${}_1$/Dose${}_2$ = 10)', r'$\sigma_{\Omega}$ [sr]', 'Median$\{\sigma_{\Omega}\}$ [sr]', 'MAD$\{\sigma_{\Omega}\}$ [sr]']
+col_labels = ['Geometry\n($\\alpha_{1}$ = 60${}^{\circ}$,\n Exposure${}_1$/Exposure${}_2$ = 5)', r'$\sigma_{\Omega}$ [sr]', 'Median$\{\sigma_{\Omega}\}$ [sr]', 'MAD$\{\sigma_{\Omega}\}$ [sr]']
 fig_labels = ['a)', 'b)', 'c)', 'd)']
-n_pts = 1500 # Points on sphere
+n_pts = 500 # Points on sphere
 n_pts_sphere = 50000 # Points on sphere
 n_grid_pts = 25
 inch_fig = 5
@@ -112,7 +112,7 @@ def plot_2d_regions(ax, cax, pts, data, special_pt=(-1,-1)):
                     va='center', ha='center', fontsize=fontsize,
                     annotation_clip=False, rotation=rotation, zorder=13)
 
-    my_annotate(ax, 'Dose${}_1$/Dose${}_2$', (0.5, -0.13), fontsize=14)
+    my_annotate(ax, 'Exposure${}_1$/Exposure${}_2$', (0.5, -0.13), fontsize=14)
     my_annotate(ax, '$\\alpha_{1}$', (-0.18, 0.5), fontsize=14, rotation=90)
     my_annotate(ax, '$\\alpha_{2}$', (1.17, 0.5), fontsize=14, rotation=-90)
 
@@ -157,7 +157,7 @@ na1 = 1.33*np.sin(np.deg2rad(alpha1))
 na2 = n*np.sin(np.pi/2 - np.arcsin(na1/n))
 theta1 = np.pi/2 - np.arcsin(na1/n)
 theta2 = -(np.pi/2 - np.arcsin(na2/n))
-dose_rat_plot = 1
+dose_rat_plot = 0.6985
 dose_ratio = 10**dose_rat_plot
 total_photons = 500
 dose2 = total_photons/(1 + dose_ratio)
@@ -205,7 +205,7 @@ scene_string += line_string
[…]

### 2017-08-02 paper/figures/single-frame.py

@@ -16,7 +16,7 @@ n_cols = 4
 inch_fig = 5
 dpi = 300
 
-col_labels = ['Geometry', r'Excitation Efficiency $\eta_{\text{exc}}$', r'Detection Efficiency $\eta_{\text{det}}$', r'Total Efficiency $\eta_{\text{tot}}$']
+col_labels = ['Geometry', r'Absorption Efficiency $\eta_{\textrm{abs}}$', r'Detection Efficiency $\eta_{\textrm{det}}$', r'Total Efficiency $\eta_{\textrm{tot}}$']
 row_labels = 3*['']
 
 # Generate axes

### 2017-08-02 paper/figures/symmetric-widefield.py

@@ -97,6 +97,10 @@ def plot_2d_regions(ax, cax, pts, data, special_pt=(-1,-1)):
     ax.yaxis.set_major_locator(FixedLocator([0, 45, 90, 135, 180]))
     ax.yaxis.set_major_formatter(FuncFormatter(degrees))
 
+    from matplotlib.ticker import FuncFormatter, FixedLocator
+    ax.set_yticks([0, 0.25, 0.5, 0.75, 1.0, 1.33])
+    ax.set_yticklabels(['0', '0.25', '0.5', '0.75', '1.0', '1.33'])
+
     # Annotation
     def my_annotate(ax, annotation, xy, fontsize=9, rotation=0):
         ax.annotate(annotation, xy=(0,0), xytext=xy, textcoords='axes fraction',

### 2017-08-07 paper/paper/paper.tex
Patrick's edits.

@@
 \usepackage{float}
+\usepackage{booktabs}
 \providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
@@
 \begin{document}
-\title{Single molecule orientation determination with multiview polarized
+\title{Single-molecule orientation determination with multiview polarized
   illumination: modeling and microscope design}
 
-\author{Talon Chandler,\authormark{1,*} Shalin Mehta,\authormark{2} Rudolf
-  Oldenbourg,\authormark{2, 3} and Patrick J. La Rivi\`ere\authormark{1}}
-
-\address{\authormark{1}University of Chicago, Department of Radiology, Chicago, IL, USA.\\
-  \authormark{2}Marine Biological Laboratory, Bell Center for Regenerative Medicine, Woods Hole, MA, USA.\\
-  \authormark{3}Brown University, Department of Physics, Providence, RI, USA.}
+\author{Talon Chandler,\authormark{1,*} Shalin Mehta,\authormark{2} Hari Shroff,\authormark{3,4} Rudolf Oldenbourg,\authormark{2, 5} and Patrick J. La Rivi\`ere\authormark{1, 4}}
+\address{\authormark{1}University of Chicago, Department of Radiology, Chicago, Illinois 60637, USA\\
+  \authormark{2}Marine Biological Laboratory, Bell Center for Regenerative Medicine, Woods Hole, Massachusetts, USA\\
+  \authormark{3}Section on High Resolution Optical Imaging, National Institute
+  of Biomedical Imaging and Bioengineering, National Institutes of Health,
+  Bethesda, Maryland 20892, USA\\
+  \authormark{4}Whitman Center, Marine Biological Laboratory, Woods Hole,
+  Massachusetts 02543, USA\\
+  \authormark{5}Brown University, Department of Physics, Providence, Rhode
+  Island 02912, USA}
 \email{*talonchandler@talonchandler.com} %% email address is required
@@
   microscopes. We also find that choosing a small illumination NA and a large
-  detection NA are good design choices, multiview widefield microscopes can
-  benefit from oblique illumination and detection, and asymmetric NA microscopes
-  can benefit from exposure asymmetry.
+  detection NA are good design choices, that multiview widefield microscopes can
+  benefit from oblique illumination and detection, and that asymmetric NA
+  microscopes can benefit from exposure asymmetry.
 \end{abstract*}
@@
 number of techniques available to biologists for measuring the orientation of
-single molecules in live cells is continually expanding. We follow a recent
-review \cite{backlund2014} and compare three major categories of single molecule
+single molecules in live cells is continually expanding. A recent review
+\cite{backlund2014} identified three major categories of single molecule
 orientation determination techniques:
@@
 absorption and emission pattern of single molecules to estimate their
-orientation. By measuring the total intensity in the detector plane during
-several polarization orientations on the illumination or detection path, the
+orientation. By measuring the total intensity in the detector plane while
+varying polarizer orientations on the illumination or detection path, the
 orientation of the molecule can be estimated. These techniques are easy to
@@
 \cite{oldenbourg1995}---and the reconstruction methods are straightforward
-\cite{fourkas2001, mehta2016, backer2016}. The main drawback is speed---these
-techniques require the collection of several polarization frames to find the
-orientation of a single molecule. Despite the requirement for multiple frames,
-polarized illumination is the method of choice for live cell imaging because of
-its good combination of speed and ease of implementation. For these reasons,
-polarized illumination is the method we consider and extend in this paper.
+\cite{fourkas2001, mehta2016, backer2016}. The main drawback is that these
+techniques require several polarization frames to estimate the orientation of a
+single molecule. Despite the requirement for multiple frames, polarized
+illumination or detection methods are good choices for live cell imaging because
+of their good combination of speed and ease of implementation.
+
+The choice remains---is it best to put polarizers on the illumination path or
+the detection path? Placing a polarizer in the illumination path lengthens
+acquisition time because the polarizer needs to be changed for each polarization
+frame, but polarized illumination provides the largest intensity differences
+between molecules in different orientations. Placing polarizers in the detection
+path doesn't slow down acquisition (multiple polarization frames can be
+collected at once with a polarization splitter), but polarized detection
+provides a smaller difference in intensity between molecules in different
+orientations than polarized illumination. We are most interested in finding the
+orientation of molecules with high precision, so polarized illumination is the
+method we consider and extend in this paper.
 
 All single molecule orientation determination methods suffer from some degree of
-anisotropic orientation uncertainty, i.e., molecules in some orientations
-cannot be oriented as well as molecules in others. In some cases the
-orientation cannot be determined at all---the forward model can be degenerate
-for specific dipole orientations \cite{fourkas2001, lu2008}. We feel that the
-importance of isotropic orientation uncertainty has been underappreciated. To
-our knowledge, the only authors who have considered anisotropic orientation
-uncertainty have only analyzed a small set of molecule orientations intead of
-all orientations \cite{agrawal2012}. Our view is that an ideal technique for
-determining the orientation of single molecules can reconstruct the
-orientation with a small and nearly uniform uncertainty for all molecule
-orientations.
+anisotropic orientation uncertainty, i.e., some molecular orientations cannot be
+determined as precisely as others. In some cases the orientation cannot be
+determined at all---the forward model can be degenerate for specific dipole
+orientations \cite{fourkas2001, lu2008}. We feel that the importance of
+isotropic orientation uncertainty has been underappreciated. To our knowledge,
+the only authors who have considered anisotropic orientation uncertainty have
+only analyzed a small set of molecular orientations instead of all orientations
+\cite{agrawal2012}. Our view is that an ideal technique for determining the
+orientation of single molecules can reconstruct the orientation with a small and
+nearly uniform uncertainty for all molecule orientations.
 
@@
 for biological imaging \cite{wu2013, keller2015, wu2016, wu2017}. Multiview
-microscopes address two key limitations of single view microscopes. First,
+microscopes offer two major advantages over single-view microscopes. First,
 multiview microscopes can achieve nearly isotropic resolution compared to the
-poor axial resolution of single view microscopes. Second, multiview microscopes
-can easily incorporate selective illumination to reduce phototoxicity. For
-example, compare light-sheet illumination to wide-field illumination where
-most of the illumination light excites out of focus molecules. These
-advantages make multiview microscopes good candidates for imaging live
-biological specimens for long periods.
+poor axial resolution of single-view microscopes. Second, multiview microscopes
+can use light-sheet illumination to reduce phototoxicity without requiring
+dedicated illumination optics. For example, if two orthogonal objectives are
+focused on the same point, then both objectives can alternate roles as the
+light-sheet illumination path and as the detection path. Light-sheet
+illumination offers a major reduction of phototoxicity because a light sheet
+only illuminates in focus regions while single-view microscopes illuminate out
+of focus regions. Together, these advantages make multiview microscopes good
+candidates for imaging live biological specimens for long periods.
 
@@
 
+Note our distinction between a \emph{frame} and a \emph{view}. A frame refers to
+a single intensity measurement with a fixed polarizer setting, illumination
+path, and detection path. Acquiring additional frames entails varying the
+polarizer setting, illumination path, or detection path. A view refers to a
+detection path, and multiview microscopes use multiple detection paths.
+
+We use roman type for scalars, e.g., $\phi$, $\theta$; bold lowercase type for
+vectors, e.g., $\mb{r}, \bs{\mu}$; hats for unit vectors, e.g.,
+$\mh{r}, \hat{\bs{\mu}}$; and bold capital type for matrices, e.g.,
+$\mb{R}, \mb{F}$.
+
 \subsection{Absorption Efficiency}\label{excitation}
-Consider polarized widefield K\"ohler illumination incident on a single
-molecule at the focal point with the optical axis along the $\mh{z}$ axis. In
-this section we will calculate the \emph{absorption efficiency} in this
-geometry---the fraction of the incident power that is absorbed by the molecule.
+In this section we will calculate the \emph{absorption efficiency} of a single
+molecule---the fraction of the incident power that is absorbed by the
+molecule. Our approach is inspired by Fourkas \cite{fourkas2001}, but here we
+calculate the absorption efficiency instead of the detection efficiency.
 
-We start with the unit electric field transmitted by the polarizer,
+\begin{figure}[H]
+\centering\includegraphics[width=\textwidth]{frames}
+\caption{Coordinate systems for a) the absorption dipole moment
+  $\hat{\bs{\mu}}_{\text{abs}}$, b) the polarizer transmission axis
+  $\hat{\mb{e}}_{\text{exc}}$, and c) the dummy integration vector
+  $\hat{\mb{r}}$.}
+\label{fig:coordinates}
+\end{figure}
+
+We consider a single molecule with an absorption dipole moment
+$\hat{\bs{\mu}}_{\text{abs}}$ that we express in spherical coordinates $\Theta$
+and $\Phi$ (see Figure \ref{fig:coordinates}a)) as
 \begin{align}
- \hat{\mb{E}}_{\text{exc}} = \cos\phi_{\text{exc}}\hat{\mb{x}} +
+  \hat{\bs{\mu}}_{\text{abs}}(\Theta, \Phi) = \sin\Theta\cos\Phi\hat{\mb{x}} + \sin\Theta\sin\Phi\hat{\mb{y}} + \cos\Theta\hat{\mb{z}} \label{eq:mu}.
+\end{align}
+We place the molecule at the focal point of an ideal, aplanatic,
+polarization-preserving objective with the optical axis aligned with the
+$\mh{z}$ axis. Finally, we place a polarizer behind the objective with a
+variable transmission axis $\hat{\mb{e}}_{\text{exc}}$ that we express as
+\begin{align}
+ \hat{\mb{e}}_{\text{exc}}(\phi_{\text{exc}}) = \cos\phi_{\text{exc}}\hat{\mb{x}} +
 \sin\phi_{\text{exc}}\hat{\mb{y}},\label{eq:excfield}
@@
 where $\phi_{\text{exc}}$ is the angle between the transmission axis of the
-excitation polarizer and the $\mh{x}$ axis. Next, we model the action of the an
-ideal, aplanatic, polarization preserving illumination objective by multiplying
-the electric field vector by a direction dependent rotation matrix,
+polarizer and the positive $\mh{x}$ axis (see Figure
+\ref{fig:coordinates}b)). We illuminate the polarizer with a beam of collimated
+light propagating along the optical axis, so the electric field incident on the
+objective is parallel with the transmission axis of the polarizer.
+
+We denote the illuminated region of the objective by $\Omega$. If we use
+unit vectors $\hat{\mb{r}}$ expressed in spherical coordinates $\theta$ and
+$\phi$ (see Figure \ref{fig:coordinates}c)) as
   \begin{align}
-  \mb{R}(\mh{r}) &= \begin{bmatrix} \cos\theta\cos^2\phi + \sin^2\phi & (\cos\theta -1)\sin\phi\cos\phi & -\sin\theta\cos\phi\\ (\cos\theta - 1)\sin\phi\cos\phi & \cos\theta\sin^2\phi + \cos^2\phi & -\sin\theta\sin\phi \\ \sin\theta\cos\phi& \sin\theta\sin\phi & \cos\theta \end{bmatrix}
+   \hat{\mb{r}}(\theta, \phi) = \sin\theta\cos\phi\hat{\mb{x}} + \sin\theta\sin\phi\hat{\mb{y}} + \cos\theta\hat{\mb{z}}, 
   \end{align}
-  where
+  then the region $\Omega$ can be expressed as
   \begin{align}
-   \hat{\mb{r}} = \sin\theta\cos\phi\hat{\mb{x}} + \sin\theta\sin\phi\hat{\mb{y}} + \cos\theta\hat{\mb{z}} 
+  \Omega = \{\phi, \theta\ |\ 0 < \phi \leq 2\pi, 0 < \theta \leq \alpha\},
   \end{align}
-  is a unit vector that points from the origin to the point where the incident
-  electric field intersects the Gaussian sphere of the illumination
-  objective. Next, we find the power that excites the molecule by taking the
-  dot product of the rotated incident field $\mb{R}(\mh{r})\mh{E}_{\text{exc}}$
-  and the absorption dipole moment
+  where $\alpha$ is the angle between the optical axis and the most oblique
+  illuminating ray. Equivalently, $\alpha$ can be expressed in terms of NA using
   \begin{align}
-  \hat{\bs{\mu}}_{\text{abs}} = \sin\Theta\cos\Phi\hat{\mb{x}} + \sin\Theta\sin\Phi\hat{\mb{y}} + \cos\Theta\hat{\mb{z}}
+    \text{NA} = n\sin\alpha, 
   \end{align}
-then taking its modulus squared. If the incident fields are spatially incoherent
-then we can find the total power that excites the molecule by integrating over the
-illuminated region of the Gaussian sphere $\Omega$ where
+  where $n$ is the index of refraction of the sample medium. Note that the NA of
+  the objective is always greater than or equal to the illumination NA because
+  we can under fill the back aperture of the objective.
+
+  The objective focuses the collimated illumination by applying a position-dependent
+  rotation to the electric field which we model by multiplying the incident electric
+  field $\hat{\mb{e}}$ by a position-dependent rotation matrix
   \begin{align}
-  \Omega &= \{\phi, \theta\ |\ 0 < \phi \leq 2\pi, 0 < \theta \leq \alpha\}, \\
-  \alpha &= \arcsin\left(\frac{\text{NA}}{n}\right), \label{eq:alpha}
-\end{align}
-NA is the illumination numerical aperture, and $n$ is the index of refraction of
-the medium.
-
-Finally, we find the fraction of the incident power absorbed by the molecule
-by dividing the total absorbed power by the total incident power. The complete
-expression for the absorption efficiency in vector notation is
-\begin{align}
-  \eta_{\text{abs}} = \frac{\int_{\Omega}d\mh{r}|\hat{\bs{\mu}}_{\text{abs}}\cdot\mb{R}(\mh{r})\mh{E}_{\text{exc}}|^2}{\int_{\Omega}d\mh{r}}\label{eq:abs}. 
+  \mb{R}(\mh{r}) &= \begin{bmatrix} \cos\theta\cos^2\phi + \sin^2\phi & (\cos\theta -1)\sin\phi\cos\phi & -\sin\theta\cos\phi\\ (\cos\theta - 1)\sin\phi\cos\phi & \cos\theta\sin^2\phi + \cos^2\phi & -\sin\theta\sin\phi \\ \sin\theta\cos\phi& \sin\theta\sin\phi & \cos\theta \end{bmatrix} \label{eq:rot_mat}.
   \end{align}
 
-We substitute equations \ref{eq:excfield}--\ref{eq:alpha} into equation
+  To find the absorption efficiency $\eta_{\text{abs}}$ of a single molecule if
+  it were illuminated by a single ray, we take the unit incident electric field
+  $\hat{\mb{e}}$, take the dot product with the absorption dipole moment, then
+  take the modulus squared
+    \begin{align}
+      \eta_{\text{abs}} &= |\hat{\bs{\mu}}_{\text{abs}}\cdot\mh{e}_{\text{exc}}|^2. \label{eq:singleray}
+    \end{align}
+
+    To find the absorption efficiency of a single molecule under incoherent
+    focused light, we integrate over all rays and divide by the total incident
+    power which gives the vector expression
+\begin{align}
+  \eta_{\text{abs}} = \frac{\int_{\Omega}d\mh{r}|\hat{\bs{\mu}}_{\text{abs}}\cdot\mb{R}(\mh{r})\mh{e}_{\text{exc}}|^2}{\int_{\Omega}d\mh{r}}\label{eq:abs}. 
+\end{align}
+
+We substitute equations \ref{eq:mu}--\ref{eq:rot_mat} into equation
 \ref{eq:abs}, evaluate the integrals, and simplify to express the absorption
@@
 
-At the beginning of this section we assumed that the molecule was located at
-the focal point of the condenser. If the illumination objective is aplanatic
-then we can relax this assumption and the above expressions are valid for
-molecules anywhere in the focal plane.
+At the beginning of this section we assumed that the molecule was located at the
+focal point of the condenser. If the illumination objective is aplanatic and we
+use K\"ohler illumination, then we can relax this assumption and the above
+expressions are valid for molecules anywhere in the focal plane.
 
-We also assumed that we used incoherent K\"ohler illumination. If we illuminate
-the back aperture with a weakly focused laser beam and scan the laser beam
-slowly---if the scan velocity is much less than the beam waist divided by the
-coherence time---we can ignore longitudinal excitation and the molecule will
-be excited as if a single plane was incident. Therefore, we can find the
-absorption efficiency of a molecule under weakly focused scanned laser
-illumination by taking the limit of equation \ref{eq:scalarabs} as
-$\alpha \rightarrow 0$ giving
+We also assumed that we used incoherent illumination, but we can extend these
+results to light-sheet illumination that uses a scanned coherent laser. If we
+illuminate the back aperture with a weakly focused laser beam and scan the laser
+beam slowly---if the scan velocity is much less than the beam waist radius
+divided by the coherence time---we can ignore longitudinal excitation and the
+molecule will be excited as if a single plane wave was incident. Therefore, we
+can find the absorption efficiency of a molecule under weakly focused scanned
+laser illumination by taking the limit of equation \ref{eq:scalarabs} as
+$\alpha \rightarrow 0$ (or by plugging equation \ref{eq:mu} and
+\ref{eq:excfield} into equation \ref{eq:singleray}) giving
 \begin{align}
-  \eta_{\text{abs}} &= \sin^2\Theta\cos^2(\Phi - \phi_{\text{exc}})
+  \eta_{\text{abs}} &= \sin^2\Theta\cos^2(\Phi - \phi_{\text{exc}}),
 \end{align}
-which is recognizable as Malus' law generalized to three dimensions. Notice that
-the absorption efficiency can take its a maximum range of values
-$\eta_{\text{abs}} \in [0, 1]$ when $\alpha=0$.
+which is recognizable as Malus' law generalized to three dimensions. This means
+that light-sheet illumination is approximately equivalent to widefield
+illumination with a low NA. Notice that the absorption efficiency can take its
+maximum range of values ($\eta_{\text{abs}} \in [0, 1]$) when $\alpha=0$.
 
@@
 
-Also note that the detection efficiency can take its a maximum range of values
+Also note that the detection efficiency can take its maximum range of values
 when $B$ takes its maximum value, i.e. when
-$\alpha=\arccos\left(\frac{1}{\sqrt{3}}\right) \approx 54.7^{\circ}$. When $n = 1.33$,
-this condition corresponds to $\text{NA} \approx 1.1$.
+$\alpha=\arccos\left(\frac{1}{\sqrt{3}}\right) \approx 54.7^{\circ}$.
 
@@
 In sections \ref{excitation} and \ref{detection} we assumed that both the
-illumination and detection objectives had $\mh{z}$ aligned optical axes. To extend
-the forward model to oblique optical axes, we express the dipole orientation in
+illumination and detection objectives had $\mh{z}$ aligned optical axes, i.e.,
+the same objective was used for illumination and detection. To extend the
+forward model to oblique optical axes which allows for non-coincident
+illumination and detection objectives, we express the dipole orientation in
 rotated coordinates using the following expressions
@@
 \end{align}
-where $\psi$ is the angle of a right handed rotation about the $\mh{y}$ that maps the $\mh{z}$ axis onto the new optical axis.
+where $\psi$ is the angle of a right handed rotation about the $\mh{y}$ axis
+that maps the $\mh{z}$ axis onto the new optical axis.
 
@@
 
-Note our distinction between \emph{frames} and \emph{views}. Frames refer to
-single intensity measurements with fixed illumination, detection, and
-polarization geometries. Views refers to the detection objectives, and multiview
-microscopes use multiple detection objectives.
-
 \begin{figure}[H]
@@
   \ref{eq:scalarabs}; 3) the detection efficiency, see equation
-  \ref{eq:scalardet}; (4) the total efficiency, the product of the absorption and
-  detection efficiencies.\newline \newline \textbf{Rows top to bottom:} 1)
-  epi-illumination ($\text{NA} = 1.1$) with $x$-polarized light and
-  epi-detection (NA = 1.1); 2) orthogonal illumination ($\text{NA} = 0.8$) and
-  detection ($\text{NA} = 0.8$); 3) oblique illumination ($\text{NA} = 0.8$) and
-  detection ($\text{NA} = 0.8$). All simulations use $n=1.33$.}
+  \ref{eq:scalardet}; (4) the total efficiency, the product of the absorption
+  and detection efficiencies.\newline \newline \textbf{Rows top to bottom:} 1)
+  coincident illumination ($\text{NA} = 1.1$ with $x$-polarized light) and
+  detection (NA = 1.1); 2) non-coincident orthogonal illumination
+  ($\text{NA} = 0.8$) and detection ($\text{NA} = 0.8$); 3) non-coincident
+  135${}^{\circ}$-separated illumination ($\text{NA} = 0.8$) and detection
+  ($\text{NA} = 0.8$). All simulations use $n=1.33$.}
   \label{fig:single-frame}
@@
 designs (see \cite{wu2017} for an asymmetric NA design) and the effect of
-sample exposure.
+asymmetric sample exposures. Note that we use the word \emph{symmetry} to refer
+to dual-view designs with objectives that are identical and the word
+\emph{asymmetry} to refer to dual-view designs with objectives that have a
+different NA or sample exposure.
 
@@
 
-For each microscope design and molecule orientation we calculate the Fisher
-information matrix. If the intensity is Poisson distributed the
-Fisher information matrix is given by
+A common way to evaluate the ability to estimate the parameters $\Theta$ and
+$\Phi$ from the data is to calculate the Cramer-Rao lower bound (CRLB) for each
+parameter \cite{kay1993}. The CRLBs are given by the diagonal elements of the
+inverse of the Fisher information matrix, and they give the minimum variance of
+an unbiased estimator for each parameter. For each microscope design and
+molecule orientation we calculate the Fisher information matrix. If the
+intensity is Poisson distributed the Fisher information matrix is given by
 \begin{align}
@@
 \end{align}
-where $I_k$ is the intensity measured in the $k$th frame of an $N$ frame
-microscope.
-
-A common way to evaluate the ability to estimate the parameters $\Theta$ and
-$\Phi$ from the data is to calculate the Cramer-Rao lower bound (CRLB) for each
-parameter \cite{kay1993}. The CRLBs are given by the diagonal elements of the
-inverse of the Fisher information matrix, and they give the minimum variance of
-an unbiased estimator for each parameter. Agrawal et. al. used the square root
-of the product of the CRLBs multiplied by the Jacobian determinant,
+where $I_k$ is the intensity measured in the $k$th frame of an $N$-frame
+microscope. Agrawal et. al. used the square root of the product of the CRLBs
+multiplied by the Jacobian determinant,
 $\sin\Theta\sqrt{\mb{F}^{-1}_{0,0}\mb{F}^{-1}_{1,1}}$, to find the area of
@@
 \end{align}
-as our evaluation metric. We take the determinant of the inverse Fisher
-information matrix---a parametrization independent value called the
+as our evaluation metric. We calculate the determinant of the inverse Fisher
+information matrix---a parametrization-independent value called the
 \emph{generalized variance} \cite{anderson1958}---take the square root, then
@@
 scale of the solid-angle uncertainty is to use the mean and variance,
-respectively. Because the solid-angle uncertainty admits extremely large values
-that can change the mean and variance dramatically depending on the sample
-points, we use the median and median absolute deviation (MAD)---the median of
-the data's absolute difference from the median---as robust replacements for the
-mean and variance.
+respectively. However, because the solid-angle uncertainty admits extremely
+large values that can change the mean and variance dramatically depending on the
+sample points, we use the median and median absolute deviation (MAD)---the
+median of the data's absolute difference from the median---as robust
+alternatives to the mean and variance.
 
@@
 importance of illumination and detection NAs---reducing the illumination NA (an
-inexpensive modification) improves the solid angle uncertainty much less than
-increasing the detection NA (an expensive modification). Also note from Figure
+inexpensive modification requiring under filling the back aperture) improves the
+solid angle uncertainty much less than increasing the detection NA (an expensive
+modification requiring a higher NA objective). Also note in Figure
 \ref{fig:single-arm}b) that single-view designs suffer from high orientation
-uncertainty when the molecules are oriented along the optical axis (because the
-molecules are not efficiently excited) and in the transverse plane (because it
-is ambiguous whether the molecule is oriented above or below the transverse
-plane).
+uncertainty when the molecules are oriented along the optical axis (the
+molecules are not efficiently excited), near the transverse plane (it is
+ambiguous whether the molecule is oriented above or below the transverse plane),
+and near the polarizer orientations (it is ambiguous whether the molecule is on
+either side of the polarizer orientation).
 
@@
   the solid-angle uncertainty as a function of illumination and detection NA. d)
-  MAD of solid-angle uncertainty as a function of illumination and detection
+  MAD of the solid-angle uncertainty as a function of illumination and detection
   NA. The microscope in a) and b) is indicated by a cross in c) and d).}
@@
 Figures \ref{fig:symmetric-widefield} and \ref{fig:double-arm} show our results
-for dual-view designs. We illuminate from one objective and detect from the
-other for four polarization frames, then we repeat the polarization frames with
-the same objectives in reversed roles. In Figure \ref{fig:symmetric-widefield}
-we sweep through the NA of both objectives and the angle between the objectives
-while considering steric constraints. Note how adding a second view removes the
-high uncertainty region in the transverse plane. The dual-view microscope still
-has high uncertainty regions near the optical axes, but the uncertainty is
-reduced compared to the single-view microscope. We find that the lowest median
-of the solid-angle uncertainty occurs with the largest possible NA
-objectives. We also find that increasing NA always lowers the MAD, but
-orthogonal arms are not always best. At large NA, it is advantageous to move the
-objectives together or against the cover slip. Given the high uncertainty in the
-transverse plane for a single view shown in Figure \ref{fig:single-arm}b), we
-can see why oblique views perform better than orthogonal views---oblique views
-are complementary because the regions of high orientation of uncertainty from
-each view do not overlap.
+for dual-view symmetric widefield designs. We illuminate from one objective and
+detect from the other for four polarization frames, then we repeat the
+polarization frames with the same objectives in reversed roles. In Figure
+\ref{fig:symmetric-widefield} we sweep through the NA of both objectives and the
+angle between the objectives while considering steric constraints. Note how
+adding a second view removes the high uncertainty region in the transverse
+plane. The dual-view microscope still has high uncertainty regions near the
+optical axes, but the uncertainty is reduced by almost two orders of magnitude
+compared to the single-view microscope. We find that the lowest median of the
+solid-angle uncertainty occurs with the largest possible NA objectives. We also
+find that increasing NA always lowers the MAD, but orthogonal arms are not
+always best. At large NA, it is advantageous to move the objectives together or
+against the cover slip. Given the high uncertainty in the transverse plane for a
+single-view shown in Figure \ref{fig:single-arm}b), we can see why oblique views
+perform better than orthogonal views---oblique views are complementary because
+the regions of high orientation uncertainty from each view do not overlap.
 
@@
 \caption{Dual-view symmetric widefield designs with varying NA and angle between
-  objectives. First we illuminate with the first objective (red solid) and
-  detect from the second objective (red dashed). Then we illuminate from the
-  second objective (blue solid) and detect from the first objective (blue
-  dashed). See the Fig. \ref{fig:single-arm} caption for an explanation of
-  b)-d).}
+  objectives. a) Schematic of the microscope. We illuminate with the first objective
+  (red solid) and detect from the second objective (red dashed). Then we
+  illuminate from the second objective (blue solid) and detect from the first
+  objective (blue dashed). b) Solid angle uncertainty for the microscope in
+  a). c) Median of the solid-angle uncertainty as a function of NA and the angle
+  between the objectives. d) MAD of the solid-angle uncertainty as a function of NA
+  and the angle between the objectives. The microscope in a) and b) is indicated
+  by a cross in c) and d). e) Median of the solid-angle uncertainty as a
+  function of the angle between the objectives when NA=0.6. f) MAD of the
+  solid-angle uncertainty as a function of the angle between the objectives when
+  NA=0.6. The profile in e) and f) is taken along the green line in c) and d).}
 \label{fig:symmetric-widefield}
@@
 
-Figure \ref{fig:double-arm} shows our results when we used a symmetric widefield
-design and varied the illumination and detection NA. Our results are similar to
-the single view case in Figure \ref{fig:single-arm}. We find that the best
-designs use a small illumination NA and a large detection NA where
-$\alpha \approx 57.4^{\circ}$. This means that dual view light sheet illumination
-geometries are an excellent choice for uniformly reconstructing the orientation
-of single molecules.
+Figure \ref{fig:double-arm} shows our results when we used a dual-view symmetric
+orthogonal widefield design and varied the illumination and detection NA. Our
+results are similar to the single-view case in Figure \ref{fig:single-arm}. We
+find that the best designs use a small illumination NA (like light-sheet
+illumination) and a large detection NA where $\alpha \approx 57.4^{\circ}$. This
+means that dual-view light-sheet illumination geometries are an excellent choice
+for uniformly reconstructing the orientation of single molecules.
 
@@
 \centering\includegraphics[width=\textwidth]{double-arm}
-\caption{Symmetric widefield designs with varying illumination and detection NA. See the
-  caption for Fig. \ref{fig:symmetric-widefield} for an explanation of a) and
-  the caption for Fig. \ref{fig:single-arm} for an explanation of b)-d).}
+\caption{Dual-view symmetric orthogonal widefield designs with varying
+  illumination and detection NA. a) Schematic of the microscope. b) Solid angle
+  uncertainty for the microscope in a). c) Median of the solid-angle uncertainty
+  as a function of illumination and detection NA. d) MAD of the solid-angle
+  uncertainty as a function of illumination and detection NA. The microscope in
+  a) and b) is indicated by a cross in c) and d).}
 \label{fig:double-arm}
@@
 
-Figure \ref{fig:asymmetric-double} shows our results for asymmetric light sheet
-illumination designs. We kept the two views orthogonal and used light sheet
-illumination on both sides while sweeping through the NA asymmetry and the
-sample exposure asymmetry while keeping the total sample exposure constant. We
-found that symmetric designs are at a local minimum of the median and MAD of the
-solid-angle uncertainty. We found that at extreme asymmetries where the exposure
-from the low NA arm is much larger than the exposure from the high NA arm
-(top-left and bottom-right corners of Figures \ref{fig:asymmetric-double}c) and
-d)) the median and MAD are comparable to the symmetric light-sheet
-microscope. This means that if a very high NA objective is available and a low
-NA objective can provide light sheet illumination, this microscope will perform
-comparably with a dual-view microscope with two objectives of medium NA. Figure
-\ref{fig:asymmetric-double} also shows that designs with slightly asymmetric NA
-can trade off a low solid-angle uncertainty median for a low solid-angle
-uncertainty MAD by changing the sample exposure ratio.
+Figure \ref{fig:asymmetric-double} shows our results for dual-view asymmetric
+light-sheet illumination designs. We used light-sheet illumination on both sides
+and kept the objectives orthogonal so that the light sheet from one objective
+illuminates the focal plane of the other objective. We made the objectives
+as large as possible so that both objectives would touch the cover slip and each
+other ($\alpha_1 + \alpha_2 = 90$ in Figure \ref{fig:asymmetric-double}a)). We
+swept through the NA asymmetry and the sample exposure asymmetry while keeping
+the total sample exposure constant, and found that symmetric designs are at a
+local minimum of the median and MAD of the solid-angle uncertainty. We also
+found that at extreme asymmetries where the exposure from the low NA arm is much
+larger than the exposure from the high NA arm (top-left and bottom-right corners
+of Figures \ref{fig:asymmetric-double}c) and d)) the median and MAD are
+comparable to the symmetric light-sheet microscope. This means that if a very
+high NA objective is available and a low NA objective can provide light sheet
+illumination, this microscope will perform comparably with a dual-view
+microscope with two objectives of medium NA. Figure \ref{fig:asymmetric-double}
+also shows that designs with slightly asymmetric NA can trade off a low
+solid-angle uncertainty median for a low solid-angle uncertainty MAD by changing
+the sample exposure ratio.
 
@@
   \centering\includegraphics[width=\textwidth]{asymmetric-double}
-  \caption{Asymmetric light-sheet illumination designs with varying NA and
-    sample exposure asymmetry. See the Fig. \ref{fig:symmetric-widefield}
-    caption for an explanation of a) (note that the illumination NA is small so
-    the illumination is represented by a dot instead of a solid line) and the
-    Fig. \ref{fig:single-arm} caption for an explanation of b)-d).}
+  \caption{Dual-view asymmetric light-sheet illumination designs with varying NA and
+    sample exposure asymmetry. a) Schematic of the microscope. b) Solid angle
+    uncertainty for the microscope in a). c) Median of the solid-angle
+    uncertainty as a function of NA and sample exposure asymmetry. d) MAD of the
+    solid-angle uncertainty as a function of NA and sample exposure
+    asymmetry. The microscope in a) and b) is indicated by a cross in c) and
+    d).}
 \label{fig:asymmetric-double}
@@
 
-% Figure \ref{fig:triple-arm} shows the results when we considered symmetric
-% light-sheet designs with a lower detection arm. We swept through the NA of the
-% upper arms and NA of the lower arm. The results show that the best designs have
-% the largest possible detection upper and lower NA. Note to Patrick: I used the
-% same efficiencies for the upper and lower arms. I'm still thinking/unsure about
-% the factor of $\sim$1/9 due to rolling shutter on the lower objective. If we do
-% end up using this factor it will weight the importance of a large NA on the
-% upper objectives. There is a small dip in the MAD when the upper and lower
-% objectives have the same NA, but it is still better to use the largest possible
-% lower NA.
+Table 1 shows a summary of our evaluation metrics for the best microscopes in
+each class. The single-view widefield microscope with a large illumination NA
+and high detection NA has the lowest solid-angle uncertainty median but the
+largest max. The dual-view oblique symmetric widefield with an intermediate NA
+performs reasonably well on all three metrics, while the dual-view orthogonal
+symmetric light sheet with the largest possible detection NA performs very well
+on all three metrics.
 
-% \begin{figure}[htbp]
-% \centering\includegraphics[width=\textwidth]{triple-arm}
-% \caption{Symmetric light-sheet illumination with lower detection arm and
-%     varying upper and lower NA. See the Fig. \ref{fig:single-arm} caption for
-%     an explanation of a)-d).}
-% \label{fig:triple-arm}
-% \end{figure}
+\begin{table}[ht!]
+\centering
+\caption{Comparison of the best designs in each class of microscope. All values
+  are in steradians.}
+\begin{tabular}{llll}
+  \toprule
+Microscope Type&Max\{$\sigma_{\Omega}$\}&Median\{$\sigma_{\Omega}$\}&MAD\{$\sigma_{\Omega}$\}\\ \midrule
+Single-view widefield (NA${}_\textrm{ill}$=0, NA${}_\textrm{det}$=1.1)&3.04$\times 10^{0}$&8.20$\times 10^{-4}$&1.40$\times 10^{-4}$ \\ 
+Dual-view oblique symmetric widefield (NA=0.6, $\beta$=53${}^{\circ}$)&1.35$\times 10^{-1}$&4.49$\times 10^{-3}$&9.17$\times 10^{-4}$\\
+Dual-view orthogonal symmetric light-sheet (NA=0.94)&3.13$\times 10^{-3}$&1.28$\times 10^{-3}$&1.14$\times 10^{-4}$\\
+\bottomrule
+\end{tabular}
+\end{table}
 
 \section{Discussion and Conclusions}\label{discussion}
-% Fourkas' expressions underpredict the fraction of the total power collected by
-% an objective with a polarizer by a constant fraction of
-% $\frac{3}{2}$. Therefore, using Fourkas' expressions to find $I_{\text{tot}}$
-% will result in overpredictions of $I_{\text{tot}}$, but the incorrect
-% expressions will not cause any error in the predictions of molecular
-% orientation.
-
-% Lu and Bout \cite{lu2008} study the effect of noise on the reconstruction of single fluorophores using Fourkas' expressions.
-Their main results are unaffected by the correction introduced in these notes, but because these notes predict a higher intensity than Fourkas, Lu and Bout overpredict the effect of noise in their model.
-
 In this work we have developed a model to predict the intensities measured by
@@
 range of microscope designs. Our main result is a short list of design
-heuristics that can be used to design polarized illumination single molecule
+heuristics that can be used to design polarized-illumination single-molecule
 orientation microscopes:
@@
   single-view microscopes.
-\item High NA dual-view microscopes are best when the two views are
+\item High NA dual-view widefield microscopes are best when the two views are
   oblique. Oblique views are complementary because their high uncertainty
   transverse planes do not overlap.
-\item Symmetric dual-view light-sheet illumination microscopes provide a good
+\item Symmetric NA dual-view orthogonal light-sheet microscopes provide a good
   compromise between cost, solid-angle uncertainty median, and solid-angle
-  uncertainty MAD. Asymmetric dual-view light-sheet illumination microscopes
+  uncertainty MAD. Asymmetric NA dual-view orthogonal light-sheet microscopes
   can improve the solid-angle uncertainty median or MAD by changing the sample
@@
 only considered ideal, polarization preserving, aplanatic objectives. In
-practice, these conditions are not always satisfied and power from high NA rays
-is reduced by a factor of $\frac{1}{\cos\Theta}$ \cite{nov2006}. Therefore, our
-results are most accurate for low NA rays and get progressively worse for high
-NA rays, although we our major conclusions would not change. The main advantage
-of ignoring the power reduction of high NA rays is the tractability of the
-problem. We have provided approximate closed form solutions for the illumination
-and detection efficiencies that have allowed us to draw useful design
-conclusions.
+practice, these conditions are not always satisfied and the power from high NA
+rays is apodized by a factor of $\frac{n}{\cos\Theta}$
+\cite{nov2006}. Therefore, our results are most accurate for low NA rays and get
+progressively worse for high NA rays, although our major conclusions would not
+change if we included this apodization factor. The main advantage of ignoring
+apodization is the tractability of the problem---we have provided approximate
+closed-form solutions for the illumination and detection efficiencies that have
+allowed us to draw useful design conclusions.
 
@@
 more multiview microscopes can be analyzed with the framework we have
-developed. Wu et. al. have developed a three view design with a third objection
+developed. Wu et. al. have developed a three-view design with a third objective
 below the cover slip\cite{wu2016}. Light-field detection schemes are also a type
@@
 
-Although we have only considered single molecule orientation determination in
+Although we have only considered single-molecule orientation determination in
 this work, polarized illumination techniques can be applied to find the
@@
 molecules.
+
+We have only considered polarized illumination. In the introduction
+we discussed why we prefer polarized illumination to polarized detection, but
+there is information to be gained by adding polarized detection to the
+techniques discussed in this paper. Adding a polarization splitter to the
+detection arm can increase the information available for reconstructing the
+orientation of a single molecule without increasing the acquisition time.
 
@@
 
-\section*{Acknowledgments}
-We thank Hari Shroff for helpful conversations.
-
 \section*{Disclosures}

### 2017-08-07 dipsim/util.py

@@ -41,7 +41,7 @@ def fibonacci_sphere(n):
     return np.vstack((theta, phi)).T
 
 def plot_sphere(filename=None, directions=None, data=None, interact=False,
-                color_norm='linear', color_min=0, color_max=None,
+                color_norm='linear', color_min=0, color_max=None, tick_pos=None,
                 gamma=0.25, color_map='coolwarm', linthresh=1e-3,
                 my_ax=None, my_cax=None, dpi=500, vis_px=1000,
                 save_file=False):
@@ -102,6 +102,9 @@ def plot_sphere(filename=None, directions=None, data=None, interact=False,
             draw_axis(ax)
             cmap = ax.imshow(im, interpolation='none', cmap=color_map, norm=norm)
             f.colorbar(cmap, cax=cax, orientation='vertical')
+
+        if tick_pos != None:
+            cax.yaxis.set_ticklabels(['$10^{-3}$']+8*['']+['$10^{-2}$'])            
             
         # Save
         if save_file:

### 2017-08-07 paper/figures/asymmetric-double.py

@@ -205,7 +205,7 @@ scene_string += line_string
 util.draw_scene(scene_string, my_ax=ax0, dpi=dpi)
 util.plot_sphere(directions=exp.directions, data=exp.sa_uncert,
                  color_norm='log', linthresh=1e-4,
-                 color_max=1e-2, color_min=1e-3,
+                 color_max=1e-2, color_min=1e-3, tick_pos='e',
                  my_ax=ax1, my_cax=cax1)
     
 # Plots last two columns

### 2017-08-07 paper/figures/compare-microscopes.py

@@ -5,59 +5,35 @@ import matplotlib.pyplot as plt
 import os; import time; start = time.time(); print('Running...')
 
 # Main input parameters
-row_labels = ['Geometry', r'$\sigma_{\Omega} = \sqrt{\textrm{det}\{I^{-1}\}}\sin\Theta$', r'Fractional Histogram']
-col_labels = ['One-View Wide-Field \n0.8 NA', 'Orthogonal Two-View Wide-Field \n0.8/0.8 NA', 'Two-View Wide-Field \n0.8/0.8 NA', 'Symmetric Two-View Light-Sheet \n0.8/0.8 NA', 'Asymmetric Two-View Light-Sheet \n0.71/1.1 NA', 'Three-View Wide-Field \n0.8/0.8/1.2 NA', 'Three-View Light-Sheet \n0.8/0.8/1.2 NA']
+row_labels = ['Geometry', r'$\sigma_{\Omega}$']
+col_labels = ['Single-view (NA${}_\\textrm{ill}$=0, NA${}_\\textrm{det}$=1.1', 'Dual-view oblique symmetric widefield (NA=0.6, $\beta$=53${}^{\circ}$', 'Dual-view orthogonal symmetric light-sheet (NA=0.94)']
 
-n_pts = 50000
+n_pts = 10000
 n_cols = len(col_labels)
-n_rows = 3
+n_rows = 2
 inch_fig = 5
-dpi = 400
+dpi = 200
 
 m1 = multiframe.MultiFrameMicroscope(ill_thetas=[0], det_thetas=[0],
-                                    ill_nas=[0.8], det_nas=[0.8],
-                                    ill_types=['wide'], det_types=['lens'],
+                                    ill_nas=[0.8], det_nas=[1.1],
+                                    ill_types=['sheet'], det_types=['lens'],
                                     colors=['(1,0,0)'], n_frames=4,
                                     n_pts=n_pts, max_photons=1000, n_samp=1.33)
 
-m2 = multiframe.MultiFrameMicroscope(ill_thetas=[np.pi/4, -np.pi/4], det_thetas=[-np.pi/4, np.pi/4],
-                                    ill_nas=2*[0.8], det_nas=2*[0.8],
+angle = np.pi/2 - np.arcsin(0.6/1.33)
+m2 = multiframe.MultiFrameMicroscope(ill_thetas=[angle, -angle], det_thetas=[-angle, angle],
+                                    ill_nas=2*[0.6], det_nas=2*[0.6],
[…]

### 2017-08-07 paper/figures/double-arm.py

@@ -43,8 +43,8 @@ for cax in [cax1, cax2, cax3]:
     cax.tick_params(axis='both', labelsize=14)
 
 # Calculate a list of points to sample in region
-n = 1.33
-NA_ill = np.linspace(0, 1.33, num=1000)
+n = 0.94
+NA_ill = np.linspace(0, 0.94, num=1000)
 NA_det = NA_ill
 pts = np.mgrid[n/n_grid_pts/2:n-n/n_grid_pts/2:n_grid_pts*1j,n/n_grid_pts/2:n-n/n_grid_pts/2:n_grid_pts*1j].reshape((2, n_grid_pts**2)).T.tolist()
 
@@ -86,10 +86,10 @@ def plot_2d_regions(ax, cax, pts, data, special_pt=(-1,-1)):
 
     # Set y ticks
     from matplotlib.ticker import FuncFormatter, FixedLocator
-    ax.set_yticks([0, 0.25, 0.5, 0.75, 1.0, 1.33])
-    ax.set_yticklabels(['0', '0.25', '0.5', '0.75', '1.0', '1.33'])
-    ax.set_xticks([0, 0.25, 0.5, 0.75, 1.0, 1.33])
-    ax.set_xticklabels(['0', '0.25', '0.5', '0.75', '1.0', '1.33'])
+    ax.set_yticks([0, 0.25, 0.5, 0.75, 0.94])
+    ax.set_yticklabels(['0', '0.25', '0.5', '0.75', '0.94'])
+    ax.set_xticks([0, 0.25, 0.5, 0.75, 0.94])
+    ax.set_xticklabels(['0', '0.25', '0.5', '0.75', '0.94'])
 
     # Annotation
     def my_annotate(ax, annotation, xy, fontsize=9, rotation=0):
@@ -103,7 +103,7 @@ def plot_2d_regions(ax, cax, pts, data, special_pt=(-1,-1)):
     # Calculate colors
     color_map='coolwarm'
     color_norm='log'
-    color_min=1e-5
+    color_min=1e-4
     color_max=1e-1
     if color_norm == 'linear':
         norm = matplotlib.colors.Normalize(vmin=color_min, vmax=color_max)
@@ -134,7 +134,7 @@ def plot_2d_regions(ax, cax, pts, data, special_pt=(-1,-1)):
 
     # Mask around lines
     ax.fill_between(NA_ill, NA_det, 0, color='white', zorder=2)
[…]

### 2017-08-07 paper/figures/frames/frames.py

@@ -1,0 +1,41 @@
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import matplotlib.image as mpimg
+import subprocess    
+
+inch_fig = 3.3
+fig_labels = ['a)', 'b)', 'c)']
+
+# Setup figure and axes
+fig = plt.figure(figsize=(3*inch_fig, 1*inch_fig))
+gs0 = gridspec.GridSpec(1, 3, wspace=0, hspace=0)
+ax0 = plt.subplot(gs0[0])
+ax1 = plt.subplot(gs0[1])
+ax2 = plt.subplot(gs0[2])
+
+for ax, fig_label  in zip([ax0, ax1, ax2], fig_labels):
+    ax.annotate(fig_label, xy=(0,0), xytext=(0, 0.95), textcoords='axes fraction',
+                va='bottom', ha='center', fontsize=14, annotation_clip=False)
+
+def plot_img(name, ax):
+    subprocess.call(['asy', name+'.asy'])
+    subprocess.call(['convert', '-density', str(300), '-units', 'PixelsPerInch', name+'.pdf', name+'.png'])
+    im = mpimg.imread(name+'.png')
+    ax.spines['right'].set_color('none')
+    ax.spines['left'].set_color('none')
+    ax.spines['top'].set_color('none')
+    ax.spines['bottom'].set_color('none')
+    ax.xaxis.set_ticks_position('none')
+    ax.yaxis.set_ticks_position('none')
+    ax.xaxis.set_ticklabels([])
+    ax.yaxis.set_ticklabels([])
+    ax.imshow(im, interpolation='none')
+
+
+plot_img('frame_a', ax0)
+plot_img('frame_b', ax1)
+plot_img('frame_c', ax2)
+    
+print('Saving final figure.')    
+fig.savefig('../../paper/frames.pdf', dpi=250)
+    

### 2017-08-07 paper/figures/symmetric-widefield.py

@@ -7,39 +7,43 @@ import os; import time; start = time.time(); print('Running...')
 import matplotlib.gridspec as gridspec
 
 # Main input parameters
-col_labels = ['Geometry (NA = 0.6, $\\beta=80{}^{\circ}$)', r'$\sigma_{\Omega}$ [sr]', 'Median$\{\sigma_{\Omega}\}$ [sr]', 'MAD$\{\sigma_{\Omega}\}$ [sr]']
-fig_labels = ['a)', 'b)', 'c)', 'd)']
-n_pts = 500 # Points on sphere
+col_labels = ['Geometry (NA = 0.6, $\\beta=80{}^{\circ}$)', r'$\sigma_{\Omega}$ [sr]', 'Median$\{\sigma_{\Omega}\}$ [sr]', 'MAD$\{\sigma_{\Omega}\}$ [sr]', '', '']
+fig_labels = ['a)', 'b)', 'c)', 'd)', 'e)', 'f)']
+n_pts = 5000 #Points on sphere
 n_pts_sphere = 50000 # Points on sphere
 n_grid_pts = 21
+n_line_pts = 50
 n_rows, n_cols = 1, len(col_labels)
 inch_fig = 5
 dpi = 300
 
 # Setup figure and axes
-fig = plt.figure(figsize=(2.2*inch_fig, 2*inch_fig))
-gs0 = gridspec.GridSpec(2, 2, wspace=0.4, hspace=0.1)
+fig = plt.figure(figsize=(2.2*inch_fig, 3.0*inch_fig))
+gs0 = gridspec.GridSpec(3, 2, wspace=0.4, hspace=0.2)
 gs00 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[0,0], width_ratios=[1, 0.05], wspace=0.1)
 gs10 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[1,0], width_ratios=[1, 0.05], wspace=0.1)
 gs01 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[0,1], width_ratios=[1, 0.05], wspace=0.1)
 gs11 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[1,1], width_ratios=[1, 0.05], wspace=0.1)
[…]

### 2017-08-11 paper/paper/paper.tex
Patrick and Rudolf's second round of edits.
1) Revised justification for using polarized illumination in the introduction.
2) Removed all mentions of frames---referred to intensity measurements instead.
3) New Figure 1d) to clarify the geometry and positions of vectors.
4) Uniform color scale for all figures.
5) Rewording.

@@
 \begin{document}
-\title{Single-molecule orientation determination with multiview polarized
+\title{Single-fluorophore orientation determination with multiview polarized
   illumination: modeling and microscope design}
@@
 \address{\authormark{1}University of Chicago, Department of Radiology, Chicago, Illinois 60637, USA\\
-  \authormark{2}Marine Biological Laboratory, Bell Center for Regenerative Medicine, Woods Hole, Massachusetts, USA\\
+  \authormark{2}Marine Biological Laboratory, Bell Center, Woods Hole, Massachusetts, USA\\
   \authormark{3}Section on High Resolution Optical Imaging, National Institute
@@
   We investigate the use of polarized illumination in multiview microscopes for
-  determining the orientation of single molecules. First, we relate the
-  orientation of single molecules to measurable intensities in multiview
-  microscopes and develop an information theoretic metric---the solid angle
+  determining the orientation of single fluorescent molecules. First, we relate
+  the orientation of single molecules to measurable intensities in multiview
+  microscopes and develop an information-theoretic metric---the solid-angle
   uncertainty---to compare the ability of multiview microscopes to estimate the
@@
 \section{Introduction}
-The orientation of single molecules is a valuable reporter for biological
-processes. By imaging biological samples with fluorescent probes that are fixed
-to known structures, researchers have studied the dynamics of motor proteins
-\cite{peterman2001, forkey2003}, DNA \cite{backer2016}, actin \cite{mehta2016},
-septin \cite{demay2011, mcquilken2017}, and membranes \cite{anantharam2010}. The
-number of techniques available to biologists for measuring the orientation of
-single molecules in live cells is continually expanding. A recent review
-\cite{backlund2014} identified three major categories of single molecule
-orientation determination techniques:
+The orientation of single molecules is a valuable reporter for
+biological processes. By imaging biological samples with fluorescent probes that
+are fixed to known structures, researchers have studied the dynamics of motor
+proteins \cite{peterman2001, forkey2003}, DNA \cite{backer2016}, actin
+\cite{mehta2016}, septin \cite{demay2011, mcquilken2017}, and membranes
+\cite{anantharam2010}. The number of techniques available to biologists for
+measuring the orientation of single molecules in live cells is continually
+expanding. A recent review \cite{backlund2014} identified three major categories
+of single-molecule orientation determination techniques:
 
-\textbf{Spatial variation of emission} techniques image single molecules and use
-the spatial distribution of intensity in the back focal plane \cite{lieb2004} or
+\textbf{Spatial/angular variation of emission} techniques image single molecules
+and use the distribution of intensity in the back focal plane \cite{lieb2004} or
 in the image plane \cite{backer2014} to estimate their orientation. The best
@@
 plane to encode orientation and position information in the image
-\cite{agrawal2012}. These techniques can estimate the orientation of
-molecules in all orientations and only require a single frame, but they
+\cite{agrawal2012}. These techniques can estimate the orientation of molecules
+in all orientations and only require a single intensity measurement, but they
 require complex reconstruction algorithms, expensive high numerical aperture
@@
 
-\textbf{Polarized illumination or detection} techniques exploit the anisotropic
-absorption and emission pattern of single molecules to estimate their
-orientation. By measuring the total intensity in the detector plane while
-varying polarizer orientations on the illumination or detection path, the
-orientation of the molecule can be estimated. These techniques are easy to
-implement---changing the illumination or detection polarization is simple with a
-polarization splitter \cite{mehta2016} or universal compensator
-\cite{oldenbourg1995}---and the reconstruction methods are straightforward
-\cite{fourkas2001, mehta2016, backer2016}. The main drawback is that these
-techniques require several polarization frames to estimate the orientation of a
-single molecule. Despite the requirement for multiple frames, polarized
-illumination or detection methods are good choices for live cell imaging because
-of their good combination of speed and ease of implementation.
+\textbf{Polarized illumination or detection} techniques vary the polarization of
+light in the illumination or detection paths and exploit the anisotropic
+absorption and emission patterns of single molecules to estimate their
+orientation. By illuminating and detecting from the entire sample, not just the
+focal volume, these techniques can estimate the orientation of many molecules in
+parallel without scanning. These techniques are easy to implement---changing the
+illumination or detection polarization is simple with a polarization splitter
+\cite{mehta2016} or universal compensator/polarizer \cite{shribak2003}---and the
+reconstruction methods are straightforward \cite{fourkas2001, mehta2016,
+  backer2016}. The main drawback is that these techniques require several
+intensity measurements under different polarization orientations to estimate the
+orientation of a single molecule. Despite the requirement for multiple
+measurements, polarized illumination or detection methods are good choices for
+live cell imaging because of their combination of speed and ease of
+implementation.
 
-The choice remains---is it best to put polarizers on the illumination path or
-the detection path? Placing a polarizer in the illumination path lengthens
-acquisition time because the polarizer needs to be changed for each polarization
-frame, but polarized illumination provides the largest intensity differences
-between molecules in different orientations. Placing polarizers in the detection
-path doesn't slow down acquisition (multiple polarization frames can be
-collected at once with a polarization splitter), but polarized detection
-provides a smaller difference in intensity between molecules in different
-orientations than polarized illumination. We are most interested in finding the
-orientation of molecules with high precision, so polarized illumination is the
-method we consider and extend in this paper.
+To develop a polarized illumination or detection technique we need to choose a
+method for varying the polarization and place it in the illumination or
+detection path. The simplest and most inexpensive way to vary the polarization
+is with a division-of-time (DoT) polarizer using a rotating polarizer or a
+universal compensator/polarizer. Other polarization methods
+(divison-of-amplitude, division-of-focal-plane) are more expensive and require
+additional alignment or interpolation \cite{tyo2006}. The simplicity of DoT
+polarization comes at the expense of temporal resolution---intensity
+measurements are made one after the other which increases the acquisition
+time. If we place the DoT polarizer in the detection path, we will block
+valuable photons from reaching the detector which unnecessarily increases
+exposure to the sample \cite{demay2011}. By placing the DoT polarizer in the
+illumination arm we minimize exposure to the sample. We focus on DoT polarized
+illumination in this paper because of its combination of simplicity and low
+sample exposure.
 
@@
 the only authors who have considered anisotropic orientation uncertainty have
-only analyzed a small set of molecular orientations instead of all orientations
-\cite{agrawal2012}. Our view is that an ideal technique for determining the
-orientation of single molecules can reconstruct the orientation with a small and
-nearly uniform uncertainty for all molecule orientations.
+only analyzed a subset of molecular orientations \cite{agrawal2012}. Our view is
+that an ideal technique for determining the orientation of single molecules can
+reconstruct the orientation with a small and nearly uniform uncertainty for all
+molecule orientations.
 
@@
 multiview microscopes can achieve nearly isotropic resolution compared to the
-poor axial resolution of single-view microscopes. Second, multiview microscopes
-can use light-sheet illumination to reduce phototoxicity without requiring
-dedicated illumination optics. For example, if two orthogonal objectives are
-focused on the same point, then both objectives can alternate roles as the
-light-sheet illumination path and as the detection path. Light-sheet
-illumination offers a major reduction of phototoxicity because a light sheet
-only illuminates in focus regions while single-view microscopes illuminate out
-of focus regions. Together, these advantages make multiview microscopes good
-candidates for imaging live biological specimens for long periods.
+anisotropic resolution of single-view microscopes whose axial resolution is
+reduced by at least a factor of two compared to lateral resolution. Second,
+multiview microscopes can use light-sheet illumination to reduce phototoxicity
+without requiring dedicated illumination optics. For example, if two orthogonal
+objectives are focused on the same point, then the objectives can alternate
+roles as the light-sheet illumination path and as the detection
+path. Light-sheet illumination offers a major reduction of phototoxicity because
+a light sheet only illuminates in-focus regions while other widefield and
+confocal approaches inevitably illuminate out-of-focus regions. Together, these
+advantages make multiview microscopes good candidates for imaging live
+biological specimens for long periods.
 
@@
 molecules. Furthermore, multiview microscopes can achieve nearly isotropic
-resolution while delivering selective illumination. We hypothesize that
-multiview microscopes will also provide a small and uniform orientation
-uncertainty for single molecules in all orientations. In section \ref{methods}
-we develop the required theory for polarized illumination microscopy and develop
-information theoretic metrics to compare multiview microscopes. In section
-\ref{results} we compare the results for a wide range of multiview microscope
-designs. Finally, in section \ref{discussion} we discuss the results and their
-impact on polarized multiview microscope design.
+resolution while delivering selective illumination. We find that multiview
+microscopes also provide a small and uniform orientation uncertainty for single
+molecules in all orientations. In section \ref{methods} we develop the theory
+required to model polarized illumination microscopes, and we develop metrics to
+compare microscope designs. In section \ref{results} we compare the results for
+a wide range of multiview microscope designs. Finally, in section
+\ref{discussion} we discuss the results and their impact on polarized multiview
+microscope design.
 
@@
 microscopes for the task of reconstructing molecular orientations. In sections
-\ref{excitation}--\ref{forward} we develop the forward model for single-frame
-microscopes with a fixed illumination and detection configuration. In section
-\ref{designs} we list a variety of ways that single-frame microscopes can be
-combined to create experimentally realizable multi-frame microscopes. Finally,
-in section \ref{metrics} we develop metrics that we will use to compare
-multi-frame microscopes.
-
-Note our distinction between a \emph{frame} and a \emph{view}. A frame refers to
-a single intensity measurement with a fixed polarizer setting, illumination
-path, and detection path. Acquiring additional frames entails varying the
-polarizer setting, illumination path, or detection path. A view refers to a
-detection path, and multiview microscopes use multiple detection paths.
+\ref{excitation}--\ref{forward} we develop a model for the intensity collected
+with a fixed illumination and detection configuration as a function of the
+molecule orientation. In section \ref{designs} we list a variety of ways to
+combine these intensity measurements into experimentally realizable polarized
+illumination microscopes. Finally, in section \ref{metrics} we develop metrics
+that we will use to compare these microscopes.
 
@@
   $\hat{\mb{e}}_{\text{exc}}$, and c) the dummy integration vector
-  $\hat{\mb{r}}$.}
+  $\hat{\mb{r}}$. d) The scene consists of a single molecule at the focal point
+  of an objective lens with a polarizer in the aperture plane.}
 \label{fig:coordinates}
@@
 \end{align}
-We place the molecule at the focal point of an ideal, aplanatic,
-polarization-preserving objective with the optical axis aligned with the
-$\mh{z}$ axis. Finally, we place a polarizer behind the objective with a
-variable transmission axis $\hat{\mb{e}}_{\text{exc}}$ that we express as
+We place the molecule in the focal plane of an ideal, aplanatic,
+polarization-preserving objective lens with its optical axis aligned with the
+$\mh{z}$ axis. Next, we place a polarizer in the aperture plane with a variable
+transmission axis $\hat{\mb{p}}_{\text{exc}}$ that we express as
 \begin{align}
- \hat{\mb{e}}_{\text{exc}}(\phi_{\text{exc}}) = \cos\phi_{\text{exc}}\hat{\mb{x}} +
+ \hat{\mb{p}}_{\text{exc}}(\phi_{\text{exc}}) = \cos\phi_{\text{exc}}\hat{\mb{x}} +
 \sin\phi_{\text{exc}}\hat{\mb{y}},\label{eq:excfield}
@@
 polarizer and the positive $\mh{x}$ axis (see Figure
-\ref{fig:coordinates}b)). We illuminate the polarizer with a beam of collimated
-light propagating along the optical axis, so the electric field incident on the
-objective is parallel with the transmission axis of the polarizer.
-
-We denote the illuminated region of the objective by $\Omega$. If we use
-unit vectors $\hat{\mb{r}}$ expressed in spherical coordinates $\theta$ and
-$\phi$ (see Figure \ref{fig:coordinates}c)) as
+\ref{fig:coordinates}b)). Finally, we place a spatially incoherent light source
+(or its image) in the aperture plane to establish widefield illumination. In
+this geometry each point in the sample is illuminated by a range of angles that
+can be changed by adjusting the radius of the stop in the aperture plane. We
+denote the set of illuminated angles by $\Omega$ (see Figure
+\ref{fig:coordinates}d)). If we use unit vectors $\hat{\mb{r}}$ expressed in
+spherical coordinates $\theta$ and $\phi$ (see Figure \ref{fig:coordinates}c))
+as
   \begin{align}
@@
   \end{align}
-  then the region $\Omega$ can be expressed as
+  then the set of angles $\Omega$ can be expressed as
   \begin{align}
@@
   the objective is always greater than or equal to the illumination NA because
-  we can under fill the back aperture of the objective.
+  we can underfill the aperture plane.
 
-  The objective focuses the collimated illumination by applying a position-dependent
-  rotation to the electric field which we model by multiplying the incident electric
-  field $\hat{\mb{e}}$ by a position-dependent rotation matrix
+  The objective focuses the light by applying a position-dependent rotation to
+  the electric field which we model by multiplying the incident electric field
+  $\hat{\mb{p}}_{\text{exc}}$ by a position-dependent rotation matrix
   \begin{align}
@@
   it were illuminated by a single ray, we take the unit incident electric field
-  $\hat{\mb{e}}$, take the dot product with the absorption dipole moment, then
-  take the modulus squared
+  $\hat{\mb{p}}_{\text{exc}}$, take the dot product with the absorption dipole
+  moment, then take the modulus squared
     \begin{align}
-      \eta_{\text{abs}} &= |\hat{\bs{\mu}}_{\text{abs}}\cdot\mh{e}_{\text{exc}}|^2. \label{eq:singleray}
+      \eta_{\text{abs}}^{\text{ray}} &= |\hat{\bs{\mu}}_{\text{abs}}\cdot\mh{p}_{\text{exc}}|^2. \label{eq:singleray}
     \end{align}
 
-    To find the absorption efficiency of a single molecule under incoherent
-    focused light, we integrate over all rays and divide by the total incident
-    power which gives the vector expression
+    To find the absorption efficiency of a single molecule under widefield
+    illumination, we integrate over all illumination rays and divide by the
+    total incident power which gives the vector expression
 \begin{align}
-  \eta_{\text{abs}} = \frac{\int_{\Omega}d\mh{r}|\hat{\bs{\mu}}_{\text{abs}}\cdot\mb{R}(\mh{r})\mh{e}_{\text{exc}}|^2}{\int_{\Omega}d\mh{r}}\label{eq:abs}. 
+  \eta_{\text{abs}} = \frac{\int_{\Omega}d\mh{r}|\hat{\bs{\mu}}_{\text{abs}}\cdot\mb{R}(\mh{r})\mh{p}_{\text{exc}}|^2}{\int_{\Omega}d\mh{r}}\label{eq:abs}. 
 \end{align}
@@
 
-At the beginning of this section we assumed that the molecule was located at the
-focal point of the condenser. If the illumination objective is aplanatic and we
-use K\"ohler illumination, then we can relax this assumption and the above
-expressions are valid for molecules anywhere in the focal plane.
-
-We also assumed that we used incoherent illumination, but we can extend these
-results to light-sheet illumination that uses a scanned coherent laser. If we
-illuminate the back aperture with a weakly focused laser beam and scan the laser
-beam slowly---if the scan velocity is much less than the beam waist radius
-divided by the coherence time---we can ignore longitudinal excitation and the
-molecule will be excited as if a single plane wave was incident. Therefore, we
-can find the absorption efficiency of a molecule under weakly focused scanned
-laser illumination by taking the limit of equation \ref{eq:scalarabs} as
+We can extend these results to light-sheet illumination created by scanning a
+laser. If we shine a narrow collimated laser beam onto a galvanometer in the
+aperture plane and scan the galvanometer, we create a weakly focused light sheet
+in the sample. If we underfill the aperture plane and scan the beam in the
+sample with a velocity much less than the beam waist radius divided by the
+coherence time, we can ignore longitudinal excitation and molecules in the
+sample will be excited as if a single plane wave was incident. Therefore, we can
+find the absorption efficiency of a molecule under weakly focused scanned laser
+illumination by taking the limit of equation \ref{eq:scalarabs} as
 $\alpha \rightarrow 0$ (or by plugging equation \ref{eq:mu} and
@@
 \begin{align}
-  \eta_{\text{abs}} &= \sin^2\Theta\cos^2(\Phi - \phi_{\text{exc}}),
+  \eta_{\text{abs}}^{\text{ray}} &= \sin^2\Theta\cos^2(\Phi - \phi_{\text{exc}}),
 \end{align}
 which is recognizable as Malus' law generalized to three dimensions. This means
-that light-sheet illumination is approximately equivalent to widefield
-illumination with a low NA. Notice that the absorption efficiency can take its
-maximum range of values ($\eta_{\text{abs}} \in [0, 1]$) when $\alpha=0$.
+that light-sheet illumination is approximately equivalent to low-NA widefield
+illumination with respect to the molecular orientations they excite. Notice that
+the absorption efficiency can take its maximum range of values
+($\eta_{\text{abs}} \in [0, 1]$) when $\alpha=0$.
 
@@
 Fourkas' orientation reconstruction but it does mean that Fourkas' algorithm
-under predicts the total emitted intensity by a factor of $\frac{3}{2}$.
+underpredicts the total emitted intensity by a factor of $\frac{3}{2}$.
 
@@
 forward model to oblique optical axes which allows for non-coincident
-illumination and detection objectives, we express the dipole orientation in
-rotated coordinates using the following expressions
+illumination and detection objectives, we express the dipole orientation in a
+rotated coordinate frame using the following expressions
 \begin{align}
@@
 
-\subsection{Single-Frame Microscopes}\label{forward}
+\subsection{Intensity Measurements}\label{forward}
 The detected intensity is the product of the the sample exposure
@@
 \end{align}
-Equation \ref{eq:forward_frame} is the forward model for a \emph{single-frame
-  microscope}. Figure \ref{fig:single-frame} shows the efficiencies from three
-representative examples of single-frame microscopes.
+Equation \ref{eq:forward_frame} is the forward model for a single intensity
+measurement. Figure \ref{fig:single-frame} shows the efficiencies in three
+representative geometries.
 
-The intensity measured by a single-frame microscope does not give us enough
-information to reconstruct the orientation of a single molecule. Next, we
-will consider combining several single-frame microscopes to create
-\emph{multi-frame microscopes.}
+A single intensity measurement does not give us enough information to
+reconstruct the orientation of a single molecule. Next, we will consider
+combining several intensity measurements to create complete polarized
+illumination microscopes.
 
@@
 \centering\includegraphics[width=\textwidth]{single-frame}
-\caption{Representative examples of single-frame microscopes.\newline \newline
-  \textbf{Columns left to right:} 1) schematics of single-frame microscopes
-  where the solid line encloses the illumination solid angle, the dashed line
-  encloses the detection solid angle, and the arrow indicates the transmission
-  axis of the illumination polarizer; 2) the absorption efficiency, see equation
-  \ref{eq:scalarabs}; 3) the detection efficiency, see equation
-  \ref{eq:scalardet}; (4) the total efficiency, the product of the absorption
-  and detection efficiencies.\newline \newline \textbf{Rows top to bottom:} 1)
-  coincident illumination ($\text{NA} = 1.1$ with $x$-polarized light) and
-  detection (NA = 1.1); 2) non-coincident orthogonal illumination
-  ($\text{NA} = 0.8$) and detection ($\text{NA} = 0.8$); 3) non-coincident
-  135${}^{\circ}$-separated illumination ($\text{NA} = 0.8$) and detection
-  ($\text{NA} = 0.8$). All simulations use $n=1.33$.}
+\caption{Representative examples of single intensity measurements.\newline
+  \newline \textbf{Columns left to right:} 1) schematics where the solid line
+  encloses the illumination solid angle, the dashed line encloses the detection
+  solid angle, and the arrow indicates the transmission axis of the illumination
+  polarizer; 2) the absorption efficiency, see equation \ref{eq:scalarabs}; 3)
+  the detection efficiency, see equation \ref{eq:scalardet}; (4) the total
+  efficiency, the product of the absorption and detection efficiencies.\newline
+  \newline \textbf{Rows top to bottom:} 1) coincident illumination
+  ($\text{NA} = 1.1$ with $x$-polarized light) and detection (NA = 1.1); 2)
+  non-coincident orthogonal illumination ($\text{NA} = 0.8$) and detection
+  ($\text{NA} = 0.8$); 3) non-coincident 135${}^{\circ}$-separated illumination
+  ($\text{NA} = 0.8$) and detection ($\text{NA} = 0.8$). All simulations use
+  $n=1.33$.}
   \label{fig:single-frame}
@@
 
-\subsection{Multi-Frame Microscopes}\label{designs}
-One way to collect multiple frames is to add a universal compensator to the
-illumination arm and rapidly select the incident polarization by changing
-$\phi_{\text{exc}}$ \cite{oldenbourg1995}. All of the multi-frame microscopes we
-will consider in this paper use four frames per view with the illumination
-polarization separated by 45${}^{\circ}$.
+\subsection{Polarized Illumination Microscope Configurations}\label{designs}
+One way to collect multiple intensity measurements is to add a universal
+compensator/polarizer to the illumination arm and rapidly select the incident
+polarization by changing $\phi_{\text{exc}}$ \cite{shribak2003}. All of the
+microscopes we will consider in this paper use four polarizations per
+illumination path with the polarization orientations separated by
+45${}^{\circ}$.
 
-We also consider several dual-view designs. We evaluate two-view designs that
+We also consider several dual-view designs. We evaluate dual-view designs that
 allow for illumination and detection from both objectives (see \cite{wu2013} for
@@
 asymmetric sample exposures. Note that we use the word \emph{symmetry} to refer
-to dual-view designs with objectives that are identical and the word
+to dual-view designs with identical objectives and sample exposures and the word
 \emph{asymmetry} to refer to dual-view designs with objectives that have a
@@
 orientation of a single molecule (via parameters $\Theta$ and $\Phi$) from
-intensity data.
-
-A common way to evaluate the ability to estimate the parameters $\Theta$ and
-$\Phi$ from the data is to calculate the Cramer-Rao lower bound (CRLB) for each
-parameter \cite{kay1993}. The CRLBs are given by the diagonal elements of the
-inverse of the Fisher information matrix, and they give the minimum variance of
-an unbiased estimator for each parameter. For each microscope design and
-molecule orientation we calculate the Fisher information matrix. If the
-intensity is Poisson distributed the Fisher information matrix is given by
+intensity data. A common way to evaluate the ability to estimate the parameters
+$\Theta$ and $\Phi$ from the data is to calculate the Cram\'er-Rao lower bound
+(CRLB) for each parameter \cite{kay1993}. The CRLBs are given by the diagonal
+elements of the inverse of the Fisher information matrix, and they give the
+minimum variance of an unbiased estimator for each parameter. For each
+microscope design and molecule orientation we calculate the Fisher information
+matrix. We assume that the detected intensities are Poisson distributed, so the
+Fisher information matrix is given by
 \begin{align}
@@
 \end{align}
-where $I_k$ is the intensity measured in the $k$th frame of an $N$-frame
-microscope. Agrawal et. al. used the square root of the product of the CRLBs
-multiplied by the Jacobian determinant,
+where $I_k$ is the $k$th intensity measurement in a microscope that uses $N$
+intensity measurements. Agrawal et. al. used the square root of the product of
+the CRLBs multiplied by the Jacobian determinant,
 $\sin\Theta\sqrt{\mb{F}^{-1}_{0,0}\mb{F}^{-1}_{1,1}}$, to find the area of
@@
 
-For each multi-frame microscope we calculate $\sigma_{\Omega}$ at 10,000
+For each microscope we calculate $\sigma_{\Omega}$ at 10,000
 approximately equally spaced points on the unit sphere. A desirable microscope
@@
 
-To compare multi-frame microscopes with different numbers of frames fairly, we
-kept the total exposure to the sample constant by choosing
+To compare microscopes with different numbers of intensity measurements fairly,
+we kept the total exposure to the sample constant by choosing
 $I_{\text{tot}} = \frac{4000}{N}$. Note that $I_{\text{tot}}$ is a measure of
@@
 importance of illumination and detection NAs---reducing the illumination NA (an
-inexpensive modification requiring under filling the back aperture) improves the
-solid angle uncertainty much less than increasing the detection NA (an expensive
-modification requiring a higher NA objective). Also note in Figure
+inexpensive modification requiring underfilling the aperture plane) improves the
+solid-angle uncertainty much less than increasing the detection NA (an expensive
+modification requiring a higher-NA objective). Also note in Figure
 \ref{fig:single-arm}b) that single-view designs suffer from high orientation
 uncertainty when the molecules are oriented along the optical axis (the
-molecules are not efficiently excited), near the transverse plane (it is
+molecules are not efficiently excited or detected), near the transverse plane (it is
 ambiguous whether the molecule is oriented above or below the transverse plane),
@@
 \centering\includegraphics[width=\textwidth]{single-arm}
-\caption{Single-view microscope with varying illumination and detection
-  NA. a) Schematic of a single-arm four-frame epi-illumination microscope. The
+\caption{Single-view microscope with varying illumination and detection NA. a)
+  Schematic of a single-view four-polarization epi-illumination microscope. The
   red arrows indicate the four illumination polarization orientations---one for
-  each frame. b) Solid angle uncertainty for the microscope in a). c) Median of
-  the solid-angle uncertainty as a function of illumination and detection NA. d)
-  MAD of the solid-angle uncertainty as a function of illumination and detection
-  NA. The microscope in a) and b) is indicated by a cross in c) and d).}
+  each intensity measurement. b) Solid-angle uncertainty for the microscope in
+  a). c) Median of the solid-angle uncertainty as a function of illumination and
+  detection NA. d) MAD of the solid-angle uncertainty as a function of
+  illumination and detection NA. The microscope in a) and b) is indicated by a
+  cross in c) and d).}
 \label{fig:single-arm}
@@
 for dual-view symmetric widefield designs. We illuminate from one objective and
-detect from the other for four polarization frames, then we repeat the
-polarization frames with the same objectives in reversed roles. In Figure
+detect from the other for four polarization orientations, then we repeat the
+polarization orientations with the same objectives in reversed roles. In Figure
 \ref{fig:symmetric-widefield} we sweep through the NA of both objectives and the
@@
   illuminate from the second objective (blue solid) and detect from the first
-  objective (blue dashed). b) Solid angle uncertainty for the microscope in
+  objective (blue dashed). b) Solid-angle uncertainty for the microscope in
   a). c) Median of the solid-angle uncertainty as a function of NA and the angle
@@
 results are similar to the single-view case in Figure \ref{fig:single-arm}. We
-find that the best designs use a small illumination NA (like light-sheet
-illumination) and a large detection NA where $\alpha \approx 57.4^{\circ}$. This
-means that dual-view light-sheet illumination geometries are an excellent choice
-for uniformly reconstructing the orientation of single molecules.
+find that the best designs use a small illumination NA and a large detection NA
+where $\alpha \approx 57.4^{\circ}$. As we discussed in Section
+\ref{excitation}, low-NA widefield illumination is approximately equivalent to
+light-sheet illumination for the task of estimating the orientation of
+molecules. Therefore, the preference for low-NA widefield illumination also
+implies that dual-view light-sheet illumination geometries are an excellent
+choice for uniformly reconstructing the orientation of single molecules.
 
@@
 \caption{Dual-view symmetric orthogonal widefield designs with varying
-  illumination and detection NA. a) Schematic of the microscope. b) Solid angle
+  illumination and detection NA. a) Schematic of the microscope. b) Solid-angle
   uncertainty for the microscope in a). c) Median of the solid-angle uncertainty
@@
 and kept the objectives orthogonal so that the light sheet from one objective
-illuminates the focal plane of the other objective. We made the objectives
-as large as possible so that both objectives would touch the cover slip and each
-other ($\alpha_1 + \alpha_2 = 90$ in Figure \ref{fig:asymmetric-double}a)). We
-swept through the NA asymmetry and the sample exposure asymmetry while keeping
+illuminates the focal plane of the other objective. Then, we swept through the
+NA of one objective while maximizing the NA of the other objective so that both
+objectives would touch the cover slip and each other
+($\alpha_1 + \alpha_2 = 90^{\circ}$ in Figure \ref{fig:asymmetric-double}a)). We
+swept through the NA asymmetry and the sample-exposure asymmetry while keeping
 the total sample exposure constant, and found that symmetric designs are at a
@@
   \caption{Dual-view asymmetric light-sheet illumination designs with varying NA and
-    sample exposure asymmetry. a) Schematic of the microscope. b) Solid angle
+    sample exposure asymmetry. a) Schematic of the microscope. b) Solid-angle
     uncertainty for the microscope in a). c) Median of the solid-angle
@@
 
-Table 1 shows a summary of our evaluation metrics for the best microscopes in
-each class. The single-view widefield microscope with a large illumination NA
-and high detection NA has the lowest solid-angle uncertainty median but the
-largest max. The dual-view oblique symmetric widefield with an intermediate NA
-performs reasonably well on all three metrics, while the dual-view orthogonal
-symmetric light sheet with the largest possible detection NA performs very well
-on all three metrics.
+Table 1 shows a summary of our evaluation metrics for microscopes in each
+class. The single-view widefield microscope with a large illumination NA and
+high detection NA has the lowest solid-angle uncertainty median but the largest
+max. The dual-view oblique symmetric widefield with an intermediate NA performs
+reasonably well on all three metrics, while the dual-view orthogonal symmetric
+light-sheet microscopes with large NA perform very well on all three metrics.
 
@@
 \centering
-\caption{Comparison of the best designs in each class of microscope. All values
-  are in steradians.}
+\caption{Comparison of designs in each class of microscope. All values are in
+  steradians.}
 \begin{tabular}{llll}
@@
 Microscope Type&Max\{$\sigma_{\Omega}$\}&Median\{$\sigma_{\Omega}$\}&MAD\{$\sigma_{\Omega}$\}\\ \midrule
-Single-view widefield (NA${}_\textrm{ill}$=0, NA${}_\textrm{det}$=1.1)&3.04$\times 10^{0}$&8.20$\times 10^{-4}$&1.40$\times 10^{-4}$ \\ 
+Single-view widefield (NA${}_\textrm{ill}$=0, NA${}_\textrm{det}$=1.1)&3.04$\times 10^{0}$&8.20$\times 10^{-4}$&1.48$\times 10^{-4}$ \\ 
   Dual-view oblique symmetric widefield (NA=0.6, $\beta$=53${}^{\circ}$)&1.35$\times 10^{-1}$&4.49$\times 10^{-3}$&9.17$\times 10^{-4}$\\
+Dual-view orthogonal symmetric light-sheet (NA=0.8)&6.23$\times 10^{-3}$&1.79$\times 10^{-3}$&1.65$\times 10^{-4}$\\  
 Dual-view orthogonal symmetric light-sheet (NA=0.94)&3.13$\times 10^{-3}$&1.28$\times 10^{-3}$&1.14$\times 10^{-4}$\\
@@
 weighting on large detection NAs because we would collect more light. Second, we
-only considered ideal, polarization preserving, aplanatic objectives. In
+only considered ideal, aplanatic, polarization-preserving objectives. In
 practice, these conditions are not always satisfied and the power from high NA
@@
 closed-form solutions for the illumination and detection efficiencies that have
-allowed us to draw useful design conclusions.
+allowed us to draw useful design conclusions. Third, we have used a Poisson
+noise model which ignores noise introduced by the detector. Detector noise can
+be a large fraction of the total noise at low intensities, so including detector
+noise would increase the informational value of high intensity
+measurements. Therefore, we expect that including the effects of detector noise
+would favor microscopes with a large detection NA. Finally, we have focused on
+the task of estimating the orientation of molecules and ignored the equally
+important task of estimating their location. The task of jointly estimating the
+orientation and location of single molecules from multiple views is a topic for
+future work.
 
@@
 think that this is the fairest way to compare microscopes with different numbers
-of frames. However, in cases where sample exposure and time are not an issue it
-may be more useful to compare microscopes with the same detection exposure per
-frame. This comparison would favor multiple frame microscopes even more that the
-results in this paper, because equal detection exposure per frame would improve
-photon counting statistics.
+of intensity measurements. However, in cases where sample exposure and time are
+not an issue it may be more useful to compare microscopes with the same total
+detection exposure per measurement. This comparison would favor microscopes with
+many intensity measurements even more that the results in this paper because
+equal detection exposure would improve photon counting statistics.
 
@@
 work we will investigate the ability of multiview polarized illumination
-microscopes to estimate the orientation and distribution of ensembles of
-molecules.
+microscopes to estimate the orientation distribution of ensembles of molecules.
 

### 2017-08-11 dipsim/util.py

@@ -41,7 +41,7 @@ def fibonacci_sphere(n):
     return np.vstack((theta, phi)).T
 
 def plot_sphere(filename=None, directions=None, data=None, interact=False,
-                color_norm='linear', color_min=0, color_max=None, tick_pos=None,
+                color_norm='linear', color_min=0, color_max=None,
                 gamma=0.25, color_map='coolwarm', linthresh=1e-3,
                 my_ax=None, my_cax=None, dpi=500, vis_px=1000,
                 save_file=False):
@@ -103,9 +103,6 @@ def plot_sphere(filename=None, directions=None, data=None, interact=False,
             cmap = ax.imshow(im, interpolation='none', cmap=color_map, norm=norm)
             f.colorbar(cmap, cax=cax, orientation='vertical')
 
-        if tick_pos != None:
-            cax.yaxis.set_ticklabels(['$10^{-3}$']+8*['']+['$10^{-2}$'])            
-            
         # Save
         if save_file:
             f.savefig(filename, dpi=dpi)

### 2017-08-11 paper/figures/asymmetric-double.py

@@ -120,7 +120,7 @@ def plot_2d_regions(ax, cax, pts, data, special_pt=(-1,-1)):
     color_map='coolwarm'
     color_norm='log'
     color_min=1e-4
-    color_max=1e-1
+    color_max=1e1
     if color_norm == 'linear':
         norm = matplotlib.colors.Normalize(vmin=color_min, vmax=color_max)
     elif color_norm == 'log':
@@ -205,7 +205,7 @@ scene_string += line_string
 util.draw_scene(scene_string, my_ax=ax0, dpi=dpi)
 util.plot_sphere(directions=exp.directions, data=exp.sa_uncert,
                  color_norm='log', linthresh=1e-4,
-                 color_max=1e-2, color_min=1e-3, tick_pos='e',
+                 color_max=1e1, color_min=1e-4,
                  my_ax=ax1, my_cax=cax1)
     
 # Plots last two columns

### 2017-08-11 paper/figures/compare-microscopes.py

@@ -6,7 +6,7 @@ import os; import time; start = time.time(); print('Running...')
 
 # Main input parameters
 row_labels = ['Geometry', r'$\sigma_{\Omega}$']
-col_labels = ['Single-view (NA${}_\\textrm{ill}$=0, NA${}_\\textrm{det}$=1.1', 'Dual-view oblique symmetric widefield (NA=0.6, $\beta$=53${}^{\circ}$', 'Dual-view orthogonal symmetric light-sheet (NA=0.94)']
+col_labels = ['Single-view (NA${}_\\textrm{ill}$=0, NA${}_\\textrm{det}$=1.1', 'Dual-view oblique symmetric widefield (NA=0.6, $\beta$=53${}^{\circ}$', 'Dual-view orthogonal symmetric light-sheet (NA=0.8)', 'Dual-view orthogonal symmetric light-sheet (NA=0.94)']
 
 n_pts = 10000
 n_cols = len(col_labels)
@@ -28,12 +28,19 @@ m2 = multiframe.MultiFrameMicroscope(ill_thetas=[angle, -angle], det_thetas=[-an
                                     n_pts=n_pts, max_photons=500, n_samp=1.33)
 
 m3 = multiframe.MultiFrameMicroscope(ill_thetas=[np.pi/4, -np.pi/4], det_thetas=[-np.pi/4, np.pi/4],
+                                    ill_nas=2*[0], det_nas=2*[0.8],
+                                    ill_types=2*['sheet'], det_types=2*['lens'],
+                                    colors=['(1,0,0)','(0,0,1)'], n_frames=4,
+                                    n_pts=n_pts, max_photons=500, n_samp=1.33)
+
+
+m4 = multiframe.MultiFrameMicroscope(ill_thetas=[np.pi/4, -np.pi/4], det_thetas=[-np.pi/4, np.pi/4],
                                     ill_nas=2*[0], det_nas=2*[1.33*np.sin(np.pi/4)],
                                     ill_types=2*['sheet'], det_types=2*['lens'],
                                     colors=['(1,0,0)','(0,0,1)'], n_frames=4,
                                     n_pts=n_pts, max_photons=500, n_samp=1.33)
 
-experiments = [m1, m2, m3]
+experiments = [m1, m2, m3, m4]
 
 # Generate axes
 size = (inch_fig*n_cols, inch_fig*n_rows)

### 2017-08-11 paper/figures/double-arm.py

@@ -104,7 +104,7 @@ def plot_2d_regions(ax, cax, pts, data, special_pt=(-1,-1)):
     color_map='coolwarm'
     color_norm='log'
     color_min=1e-4
-    color_max=1e-1
+    color_max=1e1
     if color_norm == 'linear':
         norm = matplotlib.colors.Normalize(vmin=color_min, vmax=color_max)
     elif color_norm == 'log':
@@ -152,7 +152,7 @@ scene_string = exp.scene_string()
 util.draw_scene(scene_string, my_ax=ax0, dpi=dpi)
 util.plot_sphere(directions=exp.directions, data=exp.sa_uncert,
                  color_norm='log', linthresh=1e-4,
-                 color_min=None, color_max=None,
+                 color_min=1e-4, color_max=1e1,
                  my_ax=ax1, my_cax=cax1)
     
 # Plots last two columns

### 2017-08-11 paper/figures/frames/frames.py

@@ -4,21 +4,22 @@ import matplotlib.image as mpimg
 import subprocess    
 
 inch_fig = 3.3
-fig_labels = ['a)', 'b)', 'c)']
+fig_labels = ['a)', 'b)', 'c)', 'd)']
 
 # Setup figure and axes
-fig = plt.figure(figsize=(3*inch_fig, 1*inch_fig))
-gs0 = gridspec.GridSpec(1, 3, wspace=0, hspace=0)
+fig = plt.figure(figsize=(4*inch_fig, 1*inch_fig))
+gs0 = gridspec.GridSpec(1, 4, wspace=0, hspace=0)
 ax0 = plt.subplot(gs0[0])
 ax1 = plt.subplot(gs0[1])
 ax2 = plt.subplot(gs0[2])
+ax3 = plt.subplot(gs0[3])
 
-for ax, fig_label  in zip([ax0, ax1, ax2], fig_labels):
+for ax, fig_label  in zip([ax0, ax1, ax2, ax3], fig_labels):
     ax.annotate(fig_label, xy=(0,0), xytext=(0, 0.95), textcoords='axes fraction',
                 va='bottom', ha='center', fontsize=14, annotation_clip=False)
 
 def plot_img(name, ax):
-    subprocess.call(['asy', name+'.asy'])
+    # subprocess.call(['asy', name+'.asy'])
     subprocess.call(['convert', '-density', str(300), '-units', 'PixelsPerInch', name+'.pdf', name+'.png'])
     im = mpimg.imread(name+'.png')
     ax.spines['right'].set_color('none')
@@ -35,6 +36,7 @@ def plot_img(name, ax):
 plot_img('frame_a', ax0)
 plot_img('frame_b', ax1)
 plot_img('frame_c', ax2)
+plot_img('frame_d', ax3)
     
 print('Saving final figure.')    
 fig.savefig('../../paper/frames.pdf', dpi=250)

### 2017-08-11 paper/figures/single-arm.py

@@ -104,7 +104,7 @@ def plot_2d_regions(ax, cax, pts, data, special_pt=(-1,-1)):
     color_map='coolwarm'
     color_norm='log'
     color_min=1e-4
-    color_max=1e0
+    color_max=1e1
     if color_norm == 'linear':
         norm = matplotlib.colors.Normalize(vmin=color_min, vmax=color_max)
     elif color_norm == 'log':
@@ -151,7 +151,7 @@ scene_string = exp.scene_string()
 util.draw_scene(scene_string, my_ax=ax0, dpi=dpi)
 util.plot_sphere(directions=exp.directions, data=exp.sa_uncert,
                  color_norm='log', linthresh=1e-4,
-                 color_min=1e-3, color_max=1e1,
+                 color_min=1e-4, color_max=1e1,
                  my_ax=ax1, my_cax=cax1)
     
 # Plots last two columns

### 2017-08-11 paper/figures/symmetric-widefield.py

@@ -121,8 +121,8 @@ def plot_2d_regions(ax, cax, pts, data, special_pt=(-1,-1),
     # Calculate colors
     color_map='coolwarm'
     color_norm='log'
-    color_min=1e-3
-    color_max=1e0
+    color_min=1e-4
+    color_max=1e1
     if color_norm == 'linear':
         norm = matplotlib.colors.Normalize(vmin=color_min, vmax=color_max)
     elif color_norm == 'log':
@@ -162,7 +162,7 @@ def plot_2d_regions(ax, cax, pts, data, special_pt=(-1,-1),
     ax.set(xlim=[0, 1.33], ylim=[0, 180])
 
 # Plot 1D region
-def plot_1d_regions(ax, pts, data, special_pt=(-1,-1), y_pos=None, y_lim=None):
+def plot_1d_regions(ax, pts, data, special_pt=(-1,-1), y_pos=None, y_lim=None, xtitle=None):
     # Set y ticks
     from matplotlib.ticker import FuncFormatter, FixedLocator
     def degrees(x, pos):
@@ -181,7 +181,7 @@ def plot_1d_regions(ax, pts, data, special_pt=(-1,-1), y_pos=None, y_lim=None):
                     annotation_clip=False, rotation=rotation, zorder=13)
 
     my_annotate(ax, '$\\beta$, Angle Between Objectives', (0.5, -0.12), fontsize=14)
-    my_annotate(ax, 'Median$\{\sigma_{\Omega}\}$ [sr]', (-0.18, 0.5), fontsize=14, rotation=90)
+    my_annotate(ax, xtitle, (-0.18, 0.5), fontsize=14, rotation=90)
 
     ax.set(xlim=[53, 127], ylim=y_lim)
     ax.plot(pts, data, '-g')
@@ -211,7 +211,7 @@ scene_string += arc_string
 util.draw_scene(scene_string, my_ax=ax0, dpi=dpi)
 util.plot_sphere(directions=exp.directions, data=exp.sa_uncert,
                  color_norm='log', linthresh=1e-4,
-                 color_min=None, color_max=None,
+                 color_min=1e-4, color_max=1e1,
                  my_ax=ax1, my_cax=cax1)
 
 # Find profile points
[…]

### 2017-08-13 paper/paper/paper.tex
Hari's edits.

@@
   microscopes. We also find that choosing a small illumination NA and a large
-  detection NA are good design choices, that multiview widefield microscopes can
-  benefit from oblique illumination and detection, and that asymmetric NA
-  microscopes can benefit from exposure asymmetry.
+  detection NA are good design choices, that multiview K\"ohler illumination
+  microscopes can benefit from oblique illumination and detection, and that
+  asymmetric NA microscopes can benefit from exposure asymmetry.
 \end{abstract*}
@@
 techniques can accurately estimate the orientation of molecules in all
-orientations, they require scanning which makes these techniques too slow for
-live cell imaging.
+orientations, they require scanning which makes acquisition time a concern in
+live-cell imaging.
 
@@
 time. If we place the DoT polarizer in the detection path, we will block
-valuable photons from reaching the detector which unnecessarily increases
-exposure to the sample \cite{demay2011}. By placing the DoT polarizer in the
-illumination arm we minimize exposure to the sample. We focus on DoT polarized
-illumination in this paper because of its combination of simplicity and low
-sample exposure.
+valuable photons from reaching the detector which leads to unnecessary exposure
+to the sample \cite{demay2011}. By placing the DoT polarizer in the illumination
+arm we minimize exposure to the sample. We focus on DoT polarized illumination
+in this paper because of its combination of simplicity and low sample exposure.
 
@@
 multiview microscopes can use light-sheet illumination to reduce phototoxicity
-without requiring dedicated illumination optics. For example, if two orthogonal
+without requiring a dedicated illumination path. For example, if two orthogonal
 objectives are focused on the same point, then the objectives can alternate
@@
 polarizer and the positive $\mh{x}$ axis (see Figure
-\ref{fig:coordinates}b)). Finally, we place a spatially incoherent light source
-(or its image) in the aperture plane to establish widefield illumination. In
-this geometry each point in the sample is illuminated by a range of angles that
-can be changed by adjusting the radius of the stop in the aperture plane. We
-denote the set of illuminated angles by $\Omega$ (see Figure
+\ref{fig:coordinates}b)). Finally, we place a spatially incoherent and spatially
+uniform light source (or its image) in the aperture plane to establish K\"ohler
+illumination. In this geometry each point in the sample is illuminated by a
+range of angles that can be changed by adjusting the radius of the stop in the
+aperture plane. We denote the set of illuminated angles by $\Omega$ (see Figure
 \ref{fig:coordinates}d)). If we use unit vectors $\hat{\mb{r}}$ expressed in
@@
 
-    To find the absorption efficiency of a single molecule under widefield
+    To find the absorption efficiency of a single molecule under K\"ohler
     illumination, we integrate over all illumination rays and divide by the
@@
 sample with a velocity much less than the beam waist radius divided by the
-coherence time, we can ignore longitudinal excitation and molecules in the
-sample will be excited as if a single plane wave was incident. Therefore, we can
-find the absorption efficiency of a molecule under weakly focused scanned laser
-illumination by taking the limit of equation \ref{eq:scalarabs} as
-$\alpha \rightarrow 0$ (or by plugging equation \ref{eq:mu} and
-\ref{eq:excfield} into equation \ref{eq:singleray}) giving
+coherence time, we can ignore the excitation caused by fields aligned with the
+optical axis. In this case, molecules in the sample will be excited as if a
+single plane wave was incident. Therefore, we can find the absorption efficiency
+of a molecule under weakly focused scanned laser illumination by taking the
+limit of equation \ref{eq:scalarabs} as $\alpha \rightarrow 0$ (or by plugging
+equation \ref{eq:mu} and \ref{eq:excfield} into equation \ref{eq:singleray})
+giving
 \begin{align}
@@
 which is recognizable as Malus' law generalized to three dimensions. This means
-that light-sheet illumination is approximately equivalent to low-NA widefield
+that light-sheet illumination is approximately equivalent to low-NA K\"ohler
 illumination with respect to the molecular orientations they excite. Notice that
@@
 measurement. Figure \ref{fig:single-frame} shows the efficiencies in three
-representative geometries.
-
-A single intensity measurement does not give us enough information to
-reconstruct the orientation of a single molecule. Next, we will consider
-combining several intensity measurements to create complete polarized
-illumination microscopes.
+representative geometries. Note that in all cases there are multiple
+orientations that give the same intensity measurement. This means that a single
+intensity measurement does not give us enough information to reconstruct the
+orientation of a single molecule. Next, we will consider combining several
+intensity measurements to create complete polarized illumination microscopes.
 
@@
 \centering\includegraphics[width=\textwidth]{single-frame}
-\caption{Representative examples of single intensity measurements.\newline
-  \newline \textbf{Columns left to right:} 1) schematics where the solid line
-  encloses the illumination solid angle, the dashed line encloses the detection
-  solid angle, and the arrow indicates the transmission axis of the illumination
-  polarizer; 2) the absorption efficiency, see equation \ref{eq:scalarabs}; 3)
-  the detection efficiency, see equation \ref{eq:scalardet}; (4) the total
-  efficiency, the product of the absorption and detection efficiencies.\newline
-  \newline \textbf{Rows top to bottom:} 1) coincident illumination
-  ($\text{NA} = 1.1$ with $x$-polarized light) and detection (NA = 1.1); 2)
-  non-coincident orthogonal illumination ($\text{NA} = 0.8$) and detection
-  ($\text{NA} = 0.8$); 3) non-coincident 135${}^{\circ}$-separated illumination
-  ($\text{NA} = 0.8$) and detection ($\text{NA} = 0.8$). All simulations use
-  $n=1.33$.}
+\caption{Representative examples of single intensity measurements. Black dots
+  indicate where the Cartesian unit vectors intersect the unit sphere. \newline \newline \textbf{Columns left to right:} 1) schematics where
+  the solid line encloses the illumination solid angle, the dashed line encloses
+  the detection solid angle, and the arrow indicates the transmission axis of
+  the illumination polarizer; 2) the absorption efficiency, see equation
+  \ref{eq:scalarabs}; 3) the detection efficiency, see equation
+  \ref{eq:scalardet}; (4) the total efficiency, the product of the absorption
+  and detection efficiencies.\newline \newline \textbf{Rows top to bottom:} 1)
+  coincident illumination ($\text{NA} = 1.1$ with $x$-polarized light) and
+  detection (NA = 1.1); 2) non-coincident orthogonal illumination
+  ($\text{NA} = 0.8$) and detection ($\text{NA} = 0.8$); 3) non-coincident
+  135${}^{\circ}$-separated illumination ($\text{NA} = 0.8$) and detection
+  ($\text{NA} = 0.8$). All simulations use $n=1.33$.}
   \label{fig:single-frame}
@@
 Figures \ref{fig:symmetric-widefield} and \ref{fig:double-arm} show our results
-for dual-view symmetric widefield designs. We illuminate from one objective and
+for dual-view symmetric K\"ohler designs. We illuminate from one objective and
 detect from the other for four polarization orientations, then we repeat the
@@
 \centering\includegraphics[width=\textwidth]{symmetric-widefield}
-\caption{Dual-view symmetric widefield designs with varying NA and angle between
+\caption{Dual-view symmetric K\"ohler designs with varying NA and angle between
   objectives. a) Schematic of the microscope. We illuminate with the first objective
@@
   solid-angle uncertainty as a function of the angle between the objectives when
-  NA=0.6. The profile in e) and f) is taken along the green line in c) and d).}
+  NA=0.6. The profile in e) and f) is taken along the orange line in c) and d).}
 \label{fig:symmetric-widefield}
@@
 Figure \ref{fig:double-arm} shows our results when we used a dual-view symmetric
-orthogonal widefield design and varied the illumination and detection NA. Our
+orthogonal K\"ohler design and varied the illumination and detection NA. Our
 results are similar to the single-view case in Figure \ref{fig:single-arm}. We
@@
 where $\alpha \approx 57.4^{\circ}$. As we discussed in Section
-\ref{excitation}, low-NA widefield illumination is approximately equivalent to
+\ref{excitation}, low-NA K\"ohler illumination is approximately equivalent to
 light-sheet illumination for the task of estimating the orientation of
-molecules. Therefore, the preference for low-NA widefield illumination also
+molecules. Therefore, the preference for low-NA K\"ohler illumination also
 implies that dual-view light-sheet illumination geometries are an excellent
@@
 \centering\includegraphics[width=\textwidth]{double-arm}
-\caption{Dual-view symmetric orthogonal widefield designs with varying
+\caption{Dual-view symmetric orthogonal K\"ohler designs with varying
   illumination and detection NA. a) Schematic of the microscope. b) Solid-angle
@@
 local minimum of the median and MAD of the solid-angle uncertainty. We also
-found that at extreme asymmetries where the exposure from the low NA arm is much
-larger than the exposure from the high NA arm (top-left and bottom-right corners
+found that at extreme asymmetries where the exposure from the low-NA arm is much
+larger than the exposure from the high-NA arm (top-left and bottom-right corners
 of Figures \ref{fig:asymmetric-double}c) and d)) the median and MAD are
 comparable to the symmetric light-sheet microscope. This means that if a very
-high NA objective is available and a low NA objective can provide light sheet
+high-NA objective is available and a low-NA objective can provide light sheet
 illumination, this microscope will perform comparably with a dual-view
-microscope with two objectives of medium NA. Figure \ref{fig:asymmetric-double}
-also shows that designs with slightly asymmetric NA can trade off a low
-solid-angle uncertainty median for a low solid-angle uncertainty MAD by changing
-the sample exposure ratio.
+microscope with two medium-NA objectives. Note that this asymmetric exposure
+configuration maximizes the use of the high-NA detection objective---if a
+high-NA objective is available it should be used for detection as much as
+possible. Figure \ref{fig:asymmetric-double} also shows that designs with
+slightly asymmetric NA can trade off a low solid-angle uncertainty median for a
+low solid-angle uncertainty MAD by changing the sample exposure ratio.
 
@@
 Table 1 shows a summary of our evaluation metrics for microscopes in each
-class. The single-view widefield microscope with a large illumination NA and
+class. The single-view K\"ohler microscope with a small illumination NA and
 high detection NA has the lowest solid-angle uncertainty median but the largest
-max. The dual-view oblique symmetric widefield with an intermediate NA performs
+max. The dual-view oblique symmetric K\"ohler with an intermediate NA performs
 reasonably well on all three metrics, while the dual-view orthogonal symmetric
@@
 Microscope Type&Max\{$\sigma_{\Omega}$\}&Median\{$\sigma_{\Omega}$\}&MAD\{$\sigma_{\Omega}$\}\\ \midrule
-Single-view widefield (NA${}_\textrm{ill}$=0, NA${}_\textrm{det}$=1.1)&3.04$\times 10^{0}$&8.20$\times 10^{-4}$&1.48$\times 10^{-4}$ \\ 
-  Dual-view oblique symmetric widefield (NA=0.6, $\beta$=53${}^{\circ}$)&1.35$\times 10^{-1}$&4.49$\times 10^{-3}$&9.17$\times 10^{-4}$\\
+Single-view K\"ohler (NA${}_\textrm{ill}$=0, NA${}_\textrm{det}$=1.1)&3.04$\times 10^{0}$&8.20$\times 10^{-4}$&1.48$\times 10^{-4}$ \\ 
+  Dual-view oblique symmetric K\"ohler (NA=0.6, $\beta$=53${}^{\circ}$)&1.35$\times 10^{-1}$&4.49$\times 10^{-3}$&9.17$\times 10^{-4}$\\
 Dual-view orthogonal symmetric light-sheet (NA=0.8)&6.23$\times 10^{-3}$&1.79$\times 10^{-3}$&1.65$\times 10^{-4}$\\  
@@
   single-view microscopes.
-\item High NA dual-view widefield microscopes are best when the two views are
+\item High-NA dual-view K\"ohler microscopes are best when the two views are
   oblique. Oblique views are complementary because their high uncertainty
@@
 only considered ideal, aplanatic, polarization-preserving objectives. In
-practice, these conditions are not always satisfied and the power from high NA
+practice, these conditions are not always satisfied and the power from high-NA
 rays is apodized by a factor of $\frac{n}{\cos\Theta}$
-\cite{nov2006}. Therefore, our results are most accurate for low NA rays and get
-progressively worse for high NA rays, although our major conclusions would not
+\cite{nov2006}. Therefore, our results are most accurate for low-NA rays and get
+progressively worse for high-NA rays, although our major conclusions would not
 change if we included this apodization factor. The main advantage of ignoring
@@
 \section*{Funding}
-TODO
+Support for this work was provided by the Intramural Research Programs of the
+National Institute of Biomedical Imaging and Bioengineering.
 

### 2017-08-13 paper/figures/symmetric-widefield.py

@@ -113,7 +113,7 @@ def plot_2d_regions(ax, cax, pts, data, special_pt=(-1,-1),
                     annotation_clip=False, rotation=rotation, zorder=13)
 
     my_annotate(ax, 'NA', (0.5, -0.12), fontsize=14)
-    my_annotate(ax, '$\\beta$, Angle Between Objectives', (-0.18, 0.5), fontsize=14, rotation=90)
+    my_annotate(ax, '$\\beta$, Angle Between Objectives', (-0.25, 0.5), fontsize=14, rotation=90)
     my_annotate(ax, 'Objectives collide\nwith cover slip', (0.65, 0.85), fontsize=14)
     my_annotate(ax, 'Objectives collide\nwith each other', (0.65, 0.15), fontsize=14)
     my_annotate(ax, 'Feasible', (0.3, 0.5), fontsize=14)
@@ -141,7 +141,7 @@ def plot_2d_regions(ax, cax, pts, data, special_pt=(-1,-1),
     sc = ax.scatter(pts[0,:], pts[1,:], c=data, s=0, cmap=cmap, norm=norm,
                     marker='s', lw=0)
     
-    ax.plot([line_pt0[0], line_pt1[0]], [line_pt0[1], line_pt1[1]], '-g')
+    ax.plot([line_pt0[0], line_pt1[0]], [line_pt0[1], line_pt1[1]], '-', color='darkmagenta', lw=3, zorder=1)
     ax.plot(special_pt[0], special_pt[1], 'kx', markersize=5)
 
 
@@ -172,7 +172,7 @@ def plot_1d_regions(ax, pts, data, special_pt=(-1,-1), y_pos=None, y_lim=None, x
 
     from matplotlib.ticker import FuncFormatter, FixedLocator
     ax.set_yticks(y_pos)
-    ax.set_yticklabels(["{:.0e}".format(x).replace('e-0', 'e-') for x in y_pos])
+    ax.set_yticklabels(["{:.1e}".format(x).replace('e-0', 'e-') for x in y_pos])
 
     # Annotation
     def my_annotate(ax, annotation, xy, fontsize=9, rotation=0):
@@ -181,10 +181,10 @@ def plot_1d_regions(ax, pts, data, special_pt=(-1,-1), y_pos=None, y_lim=None, x
                     annotation_clip=False, rotation=rotation, zorder=13)
 
[…]

### 2017-08-13 paper/paper/paper.tex
Hari's second round of edits.

@@
 multiview microscopes can use light-sheet illumination to reduce phototoxicity
-without requiring a dedicated illumination path. For example, if two orthogonal
-objectives are focused on the same point, then the objectives can alternate
-roles as the light-sheet illumination path and as the detection
-path. Light-sheet illumination offers a major reduction of phototoxicity because
-a light sheet only illuminates in-focus regions while other widefield and
-confocal approaches inevitably illuminate out-of-focus regions. Together, these
-advantages make multiview microscopes good candidates for imaging live
-biological specimens for long periods.
+while reusing parts of the optical path for both illumination and detection. For
+example, if two orthogonal objectives are focused on the same point, then the
+objectives can alternate roles as the light-sheet illumination path and as the
+detection path. Light-sheet illumination offers a major reduction of
+phototoxicity because a light sheet only illuminates in-focus regions while
+other widefield and confocal approaches inevitably illuminate out-of-focus
+regions. Together, these advantages make multiview microscopes good candidates
+for imaging live biological specimens for long periods.
 
@@
 The detected intensity is the product of the the sample exposure
-$I_{\text{tot}}$, the absorption efficiency, and the detection efficiency
+$I_{\text{exp}}$, the absorption efficiency, and the detection efficiency
 \begin{align}
-  I = I_{\text{tot}}\eta_{\text{abs}}\eta_{\text{det}}\label{eq:forward_frame}.
+  I = I_{\text{exp}}\eta_{\text{abs}}\eta_{\text{det}}\label{eq:forward_frame}.
 \end{align}
@@
 we kept the total exposure to the sample constant by choosing
-$I_{\text{tot}} = \frac{4000}{N}$. Note that $I_{\text{tot}}$ is a measure of
+$I_{\text{exp}} = \frac{4000}{N}$. Note that $I_{\text{exp}}$ is a measure of
 exposure to the sample, not the detected intensity or the total intensity

### 2017-08-18 paper/paper/paper.tex
Shalin's edits.

@@
 and use the distribution of intensity in the back focal plane \cite{lieb2004} or
-in the image plane \cite{backer2014} to estimate their orientation. The best
-techniques in this category use defocusing or phase masks in the back focal
-plane to encode orientation and position information in the image
-\cite{agrawal2012}. These techniques can estimate the orientation of molecules
-in all orientations and only require a single intensity measurement, but they
-require complex reconstruction algorithms, expensive high numerical aperture
-(NA) optics, and sensitive detectors.
+in the image plane \cite{backer2014} to estimate their
+orientation. Information-theoretically optimal techniques in this category use
+defocusing or phase masks in the back focal plane to encode orientation and
+position information in the image \cite{agrawal2012}. These techniques can
+estimate the orientation of molecules in all orientations and only require a
+single intensity measurement, but they require complex reconstruction
+algorithms, expensive high numerical aperture (NA) optics, and sensitive
+detectors.
 
@@
 We can extend these results to light-sheet illumination created by scanning a
-laser. If we shine a narrow collimated laser beam onto a galvanometer in the
+laser. If we shine a narrow, collimated, laser beam onto a galvanometer in the
 aperture plane and scan the galvanometer, we create a weakly focused light sheet
-in the sample. If we underfill the aperture plane and scan the beam in the
-sample with a velocity much less than the beam waist radius divided by the
-coherence time, we can ignore the excitation caused by fields aligned with the
-optical axis. In this case, molecules in the sample will be excited as if a
-single plane wave was incident. Therefore, we can find the absorption efficiency
-of a molecule under weakly focused scanned laser illumination by taking the
-limit of equation \ref{eq:scalarabs} as $\alpha \rightarrow 0$ (or by plugging
-equation \ref{eq:mu} and \ref{eq:excfield} into equation \ref{eq:singleray})
-giving
+in the sample. If we scan the beam in the sample with a velocity much less than
+the beam waist radius divided by the coherence time, then we can ignore the
+coherence effects of scanning---the light sheet is effectively composed of many
+incoherent beams. If we also ensure that the laser beam is weakly focused, we
+can ignore the excitation caused by the small fields aligned with the optical
+axis. Under slowly-scanned and weakly-focused light-sheet illumination,
+molecules in the sample will be excited as if a single plane wave was
+incident. Therefore, we can find the absorption efficiency of a molecule in this
+case by taking the limit of equation \ref{eq:scalarabs} as
+$\alpha \rightarrow 0$ (or by plugging equation \ref{eq:mu} and
+\ref{eq:excfield} into equation \ref{eq:singleray}) giving
 \begin{align}
@@
 
-For each microscope we calculate $\sigma_{\Omega}$ at 10,000
-approximately equally spaced points on the unit sphere. A desirable microscope
-design will have a small solid-angle uncertainty that is uniform for all
-molecule orientations. The most straightforward way to find the location and
-scale of the solid-angle uncertainty is to use the mean and variance,
-respectively. However, because the solid-angle uncertainty admits extremely
-large values that can change the mean and variance dramatically depending on the
-sample points, we use the median and median absolute deviation (MAD)---the
-median of the data's absolute difference from the median---as robust
-alternatives to the mean and variance.
+For each microscope we calculate $\sigma_{\Omega}$ at 10,000 approximately
+equally spaced points on the unit sphere. A desirable microscope design will
+have a small solid-angle uncertainty that is uniform for all molecule
+orientations. The most straightforward way to find the location and scale of the
+solid-angle uncertainty is to use the mean and variance, respectively. However,
+the solid-angle uncertainty can take extremely large values, e.g., when a
+fluorescent dipole is aligned with the optical axis of a single excitation
+objective. Such large values can change the mean and variance dramatically
+depending on the sample points, so we use the median and median absolute
+deviation (MAD)---the median of the data's absolute difference from the
+median---as robust alternatives to the mean and variance.
 
@@
 a small illumination NA and a large detection NA where
-$\alpha \approx 57.4^{\circ}$. A small illumination NA maximizes the range of
-absorption efficiencies, while a large detection NA maximizes the number of
-detected photons and the range of the detection efficiencies. Note the relative
-importance of illumination and detection NAs---reducing the illumination NA (an
-inexpensive modification requiring underfilling the aperture plane) improves the
-solid-angle uncertainty much less than increasing the detection NA (an expensive
+$\alpha \approx 65^{\circ}$ (see Figure \ref{fig:single-arm}d) at
+NA${}_{\text{det}} \approx 1.2$). A small illumination NA maximizes the range of
+absorption efficiencies, while a large detection NA optimizes the number of
+detected photons and the range of the detection efficiencies. The optimal
+detection NA is slightly larger than $\alpha = 57.4^{\circ}$ because increasing
+the detection NA reduces the effect of shot noise. Note the relative importance
+of illumination and detection NAs---reducing the illumination NA (an inexpensive
+modification requiring underfilling the aperture plane) improves the solid-angle
+uncertainty much less than increasing the detection NA (an expensive
 modification requiring a higher-NA objective). Also note in Figure
@@
 uncertainty when the molecules are oriented along the optical axis (the
-molecules are not efficiently excited or detected), near the transverse plane (it is
-ambiguous whether the molecule is oriented above or below the transverse plane),
-and near the polarizer orientations (it is ambiguous whether the molecule is on
-either side of the polarizer orientation).
+molecules are not efficiently excited or detected), near the transverse plane
+(it is ambiguous whether the molecule is oriented above or below the transverse
+plane), and near the polarizer orientations (it is ambiguous whether the
+molecule is on either side of the polarizer orientation).
 
@@
   detection NA. d) MAD of the solid-angle uncertainty as a function of
-  illumination and detection NA. The microscope in a) and b) is indicated by a
-  cross in c) and d).}
+  illumination and detection NA. The microscope configuration in a) and b) is
+  indicated by a cross in c) and d).}
 \label{fig:single-arm}
@@
 find that increasing NA always lowers the MAD, but orthogonal arms are not
-always best. At large NA, it is advantageous to move the objectives together or
+always optimal. At large NA, it is advantageous to move the objectives together or
 against the cover slip. Given the high uncertainty in the transverse plane for a
@@
 \caption{Dual-view symmetric K\"ohler designs with varying NA and angle between
-  objectives. a) Schematic of the microscope. We illuminate with the first objective
-  (red solid) and detect from the second objective (red dashed). Then we
-  illuminate from the second objective (blue solid) and detect from the first
+  objectives. a) Schematic of the microscope. We illuminate with the first
+  objective (red solid) and detect from the second objective (red dashed). Then
+  we illuminate from the second objective (blue solid) and detect from the first
   objective (blue dashed). b) Solid-angle uncertainty for the microscope in
   a). c) Median of the solid-angle uncertainty as a function of NA and the angle
-  between the objectives. d) MAD of the solid-angle uncertainty as a function of NA
-  and the angle between the objectives. The microscope in a) and b) is indicated
-  by a cross in c) and d). e) Median of the solid-angle uncertainty as a
-  function of the angle between the objectives when NA=0.6. f) MAD of the
-  solid-angle uncertainty as a function of the angle between the objectives when
-  NA=0.6. The profile in e) and f) is taken along the orange line in c) and d).}
+  between the objectives. d) MAD of the solid-angle uncertainty as a function of
+  NA and the angle between the objectives. The microscope configuration in a)
+  and b) is indicated by a cross in c) and d). e) Median of the solid-angle
+  uncertainty as a function of the angle between the objectives when NA=0.6. f)
+  MAD of the solid-angle uncertainty as a function of the angle between the
+  objectives when NA=0.6. The profile in e) and f) is taken along the purple
+  line in c) and d).}
 \label{fig:symmetric-widefield}
@@
 results are similar to the single-view case in Figure \ref{fig:single-arm}. We
-find that the best designs use a small illumination NA and a large detection NA
-where $\alpha \approx 57.4^{\circ}$. As we discussed in Section
-\ref{excitation}, low-NA K\"ohler illumination is approximately equivalent to
-light-sheet illumination for the task of estimating the orientation of
-molecules. Therefore, the preference for low-NA K\"ohler illumination also
-implies that dual-view light-sheet illumination geometries are an excellent
-choice for uniformly reconstructing the orientation of single molecules.
+find that optimal designs use a small illumination NA and a large detection
+NA. As we discussed in Section \ref{excitation}, low-NA K\"ohler illumination is
+approximately equivalent to light-sheet illumination for the task of estimating
+the orientation of molecules. Therefore, the preference for low-NA K\"ohler
+illumination also implies that dual-view light-sheet illumination geometries are
+an excellent choice for uniformly reconstructing the orientation of single
+molecules.
 
@@
   as a function of illumination and detection NA. d) MAD of the solid-angle
-  uncertainty as a function of illumination and detection NA. The microscope in
-  a) and b) is indicated by a cross in c) and d).}
+  uncertainty as a function of illumination and detection NA. The microscope
+  configuration in a) and b) is indicated by a cross in c) and d).}
 \label{fig:double-arm}
@@
   \centering\includegraphics[width=\textwidth]{asymmetric-double}
-  \caption{Dual-view asymmetric light-sheet illumination designs with varying NA and
-    sample exposure asymmetry. a) Schematic of the microscope. b) Solid-angle
-    uncertainty for the microscope in a). c) Median of the solid-angle
-    uncertainty as a function of NA and sample exposure asymmetry. d) MAD of the
+  \caption{Dual-view asymmetric light-sheet illumination designs with varying NA
+    and sample exposure asymmetry. a) Schematic of the microscope. b)
+    Solid-angle uncertainty for the microscope in a). c) Median of the
     solid-angle uncertainty as a function of NA and sample exposure
-    asymmetry. The microscope in a) and b) is indicated by a cross in c) and
-    d).}
+    asymmetry. d) MAD of the solid-angle uncertainty as a function of NA and
+    sample exposure asymmetry. The microscope configuration in a) and b) is
+    indicated by a cross in c) and d).}
 \label{fig:asymmetric-double}
@@
 \begin{itemize}
-\item All microscopes benefit from a small illumination NA and a large detection
-  NA with $\alpha \approx 57.4^{\circ}$.
+\item Single-view microscopes benefit from a small illumination NA and a large
+  detection NA with $\alpha \approx 65^{\circ}$.
 \item Dual-view microscopes outperform single-view microscopes, and dual-view
@@
   single-view microscopes.
-\item High-NA dual-view K\"ohler microscopes are best when the two views are
+\item High-NA dual-view K\"ohler microscopes are optimal when the two views are
   oblique. Oblique views are complementary because their high uncertainty

### 2017-08-18 paper/figures/asymmetric-double.py

@@ -113,8 +113,8 @@ def plot_2d_regions(ax, cax, pts, data, special_pt=(-1,-1)):
                     annotation_clip=False, rotation=rotation, zorder=13)
 
     my_annotate(ax, 'Exposure${}_1$/Exposure${}_2$', (0.5, -0.13), fontsize=14)
-    my_annotate(ax, '$\\alpha_{1}$', (-0.18, 0.5), fontsize=14, rotation=90)
-    my_annotate(ax, '$\\alpha_{2}$', (1.17, 0.5), fontsize=14, rotation=-90)
+    my_annotate(ax, '$\\alpha_{1}$, Objective 1 Half Angle', (-0.18, 0.5), fontsize=14, rotation=90)
+    my_annotate(ax, '$\\alpha_{2}$, Objective 2 Half Angle', (1.17, 0.5), fontsize=14, rotation=-90)
 
     # Calculate colors
     color_map='coolwarm'

### 2017-08-22 paper/paper/paper.tex
Compare SNR. Affiliation updates.

@@
 
-\author{Talon Chandler,\authormark{1,*} Shalin Mehta,\authormark{2} Hari Shroff,\authormark{3,4} Rudolf Oldenbourg,\authormark{2, 5} and Patrick J. La Rivi\`ere\authormark{1, 4}}
+\author{Talon Chandler,\authormark{1,*} Shalin Mehta,\authormark{1,2,3} Hari
+  Shroff,\authormark{4,5} Rudolf Oldenbourg,\authormark{2, 6} and Patrick J. La
+  Rivi\`ere\authormark{1, 5}}
 \address{\authormark{1}University of Chicago, Department of Radiology, Chicago, Illinois 60637, USA\\
-  \authormark{2}Marine Biological Laboratory, Bell Center, Woods Hole, Massachusetts, USA\\
-  \authormark{3}Section on High Resolution Optical Imaging, National Institute
+  \authormark{2}Marine Biological Laboratory, Bell Center, Woods Hole, Massachusetts 02543, USA\\
+  \authormark{3}Chan Zuckerberg Biohub, San Francisco, California 94158, USA\\
+  \authormark{4}Section on High Resolution Optical Imaging, National Institute
   of Biomedical Imaging and Bioengineering, National Institutes of Health,
   Bethesda, Maryland 20892, USA\\
-  \authormark{4}Whitman Center, Marine Biological Laboratory, Woods Hole,
+  \authormark{5}Marine Biological Laboratory, Whitman Center, Woods Hole,
   Massachusetts 02543, USA\\
-  \authormark{5}Brown University, Department of Physics, Providence, Rhode
+  \authormark{6}Brown University, Department of Physics, Providence, Rhode
   Island 02912, USA}
@@
 estimate the orientation of molecules in all orientations and only require a
-single intensity measurement, but they require complex reconstruction
+single frame of intensity measurements, but they require complex reconstruction
 algorithms, expensive high numerical aperture (NA) optics, and sensitive
-detectors.
+detectors. Furthermore, these techniques spread the signal from a single
+molecule over a large area on the detector, so they suffer from a poor
+signal-to-noise ratio \cite{toprak2006}.
 
@@
 reconstruction methods are straightforward \cite{fourkas2001, mehta2016,
-  backer2016}. The main drawback is that these techniques require several
-intensity measurements under different polarization orientations to estimate the
-orientation of a single molecule. Despite the requirement for multiple
-measurements, polarized illumination or detection methods are good choices for
-live cell imaging because of their combination of speed and ease of
-implementation.
+  backer2016}. Also, these techniques keep the signal from a single molecule
+localized on the detector, so they have a signal-to-noise ratio advantage over
+defocusing and phase masking techniques. The main drawback is that these
+techniques require several intensity measurements under different polarization
+orientations to estimate the orientation of a single molecule. Despite the
+requirement for multiple measurements, polarized illumination or detection
+methods are good choices for live cell imaging because of their combination of
+speed and ease of implementation.
 

### 2017-08-22 dipsim/multiframe.py
Error ellipse plots.

@@ -49,13 +49,13 @@ class MultiFrameMicroscope:
         return np.array(I)
     
     def calc_estimation_stats(self):
-        sphere_dx = np.arccos(1 - 2/self.n_pts) # avg. half angle betweeen n points on sphere
+        self.sphere_dx = np.arccos(1 - 2/self.n_pts) # avg. half angle betweeen n points on sphere
         self.directions = util.fibonacci_sphere(self.n_pts)
 
         # Fisher information matrix
         fi_list = []
         for direction in self.directions:
-            fi_list.append(self.noise_model.calc_fi(direction, 2*[sphere_dx]))
+            fi_list.append(self.noise_model.calc_fi(direction, 2*[self.sphere_dx]))
         self.fi = np.array(fi_list)
 
         # Solid angle uncertainty
@@ -69,4 +69,30 @@ class MultiFrameMicroscope:
         asy_string = ''
         for m in self.microscopes:
             asy_string += m.scene_string()
+        return asy_string
+
+    def ellipse_string(self, n_pts=20):
+        pts = util.fibonacci_sphere(n_pts)
+        asy_string = ''
+        for pt in pts:
+            fi = self.noise_model.calc_fi(pt, 2*[self.sphere_dx])
+            fi_inv = np.linalg.pinv(fi)
+            sigx = np.sqrt(fi_inv[0,0])
+            sigxy = np.sqrt(pt[0])*fi_inv[1,0]
+            sigy = np.sqrt(pt[0])*np.sqrt(fi_inv[1,1])
+            a = np.sqrt(0.5*(sigx**2+sigy**2) + np.sqrt(0.25*((sigx**2-sigy**2)**2) + sigxy**2))
+            b = np.sqrt(0.5*(sigx**2+sigy**2) - np.sqrt(0.25*((sigx**2-sigy**2)**2) + sigxy**2))
+            theta = 0.5*np.arctan2(2*sigxy,sigx**2-sigy**2)
+            
+            ellipse_string = 'ellipse(Theta, Phi, a1, b1, theta, false, (0,0,0));\n'
+            ellipse_string  = ellipse_string.replace('Theta', str(pt[0]))
+            ellipse_string  = ellipse_string.replace('Phi', str(pt[1]))
+            ellipse_string  = ellipse_string.replace('a1', str(a))
+            ellipse_string  = ellipse_string.replace('b1', str(b))
+            ellipse_string  = ellipse_string.replace('theta', str(theta))
[…]

### 2017-08-22 dipsim/util.py

@@ -178,14 +178,31 @@ def draw_scene(scene_string, filename='out.png', my_ax=None, dpi=500,
       }
     }
 
+    void ellipse(real Theta, real Phi, real a, real b, real theta, bool dash, triple color) {
+      triple normal = expi(Theta, Phi);
+      real a_scaled = a/max(a, b);
+      real b_scaled = b/max(a, b);      
+      path3 mycircle = rotate(degrees(Phi), Z)*rotate(degrees(Theta), Y)*shift(Z)*rotate(degrees(theta), Z)*scale(a_scaled, b_scaled, 1)*circle(c=O, r=0.05, normal=Z);
+      if (dash) {
+        draw(mycircle, p=linetype(new real[] {8,8}, offset=xpart(color))+rgb(xpart(color), ypart(color), zpart(color)));
+      } else {
+        draw(mycircle, p=rgb(xpart(color), ypart(color), zpart(color)));
+      }
+    }
+
     void mydot(real Theta, triple color) {
       triple normal = expi(Theta, 0);
       dot(normal, p=rgb(xpart(color), ypart(color), zpart(color)));
     }
 
-    void arrow(real Theta, real Phi_Pol, triple color) {
-      draw(rotate(Theta, Y)*rotate(Phi_Pol, Z)*(Z--(Z+0.2*X)), p=rgb(xpart(color), ypart(color), zpart(color)), arrow=Arrow3(emissive(rgb(xpart(color), ypart(color), zpart(color)))));
-      draw(rotate(Theta, Y)*rotate(Phi_Pol, Z)*(Z--(Z-0.2*X)), p=rgb(xpart(color), ypart(color), zpart(color)), arrow=Arrow3(emissive(rgb(xpart(color), ypart(color), zpart(color)))));
+    void arrow(real Theta, real Phi_Pol, triple color, bool dash) {
+      if (dash) {
+        draw(rotate(Theta, Y)*rotate(Phi_Pol, Z)*(Z--(Z+0.2*X)), p=linetype(new real[] {4,4}, offset=xpart(color))+rgb(xpart(color), ypart(color), zpart(color)), arrow=Arrow3(emissive(rgb(xpart(color), ypart(color), zpart(color)))));
[…]

### 2017-08-22 notes/2017-08-22-error-ellipse/figures/dual-arm.py

@@ -1,0 +1,76 @@
+from dipsim import multiframe, util
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+import matplotlib.patches as patches
+import os; import time; start = time.time(); print('Running...')
+import matplotlib.gridspec as gridspec
+
+# Main input parameters
+col_labels = ['Geometry\n(NA${}_{\\textrm{1}}$ = 0.8, NA${}_{\\textrm{2}}$ = 0.8)', 'Uncertainty Ellipses', r'$\sigma_{\Omega}$ [sr]']
+fig_labels = ['a)', 'b)', 'c)']
+n_pts = 100 # Points on sphere
+n_pts_sphere = 50000 # Points on sphere
+n_grid_pts = 5
+inch_fig = 5
+dpi = 300
+
+# Setup figure and axes
+fig = plt.figure(figsize=(3.2*inch_fig, 1*inch_fig))
+gs0 = gridspec.GridSpec(1, 3, wspace=0.2, hspace=0.1)
+gs00 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[0,0], width_ratios=[1, 0.05], wspace=0.1)
+gs10 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[0,1], width_ratios=[1, 0.05], wspace=0.1)
+gs20 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[0,2], width_ratios=[1, 0.05], wspace=0.1)
+
+ax0 = plt.subplot(gs00[0])
+ax1 = plt.subplot(gs10[0])
+ax2 = plt.subplot(gs20[0])
+cax2 = plt.subplot(gs20[1])
+
+for ax, col_label, fig_label  in zip([ax0, ax1, ax2], col_labels, fig_labels):
+    ax.annotate(col_label, xy=(0,0), xytext=(0.5, 1.05), textcoords='axes fraction',
+                va='bottom', ha='center', fontsize=14, annotation_clip=False)
+    ax.annotate(fig_label, xy=(0,0), xytext=(0, 1.05), textcoords='axes fraction',
+                va='bottom', ha='center', fontsize=14, annotation_clip=False)
+    
+for ax in [ax0, ax1, ax2]:
+    ax.tick_params(axis='both', labelsize=14)
+for cax in [cax2]:
+    cax.tick_params(axis='both', labelsize=14)
+
+# Calculate a list of points to sample in region
+n = 1.33
[…]

### 2017-08-22 notes/2017-08-22-error-ellipse/figures/single-arm.py

@@ -1,0 +1,77 @@
+from dipsim import multiframe, util
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+import matplotlib.patches as patches
+import os; import time; start = time.time(); print('Running...')
+import matplotlib.gridspec as gridspec
+
+# Main input parameters
+col_labels = ['Geometry\n(NA${}_{\\textrm{ill}}$ = 0.6, NA${}_{\\textrm{det}}$ = 0.8)', 'Uncertainty Ellipses', r'$\sigma_{\Omega}$ [sr]']
+fig_labels = ['a)', 'b)', 'c)']
+n_pts = 100 # Points on sphere
+n_pts_sphere = 50000 # Points on sphere
+n_grid_pts = 5
+inch_fig = 5
+dpi = 300
+
+# Setup figure and axes
+fig = plt.figure(figsize=(3.2*inch_fig, 1*inch_fig))
+gs0 = gridspec.GridSpec(1, 3, wspace=0.2, hspace=0.1)
+gs00 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[0,0], width_ratios=[1, 0.05], wspace=0.1)
+gs10 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[0,1], width_ratios=[1, 0.05], wspace=0.1)
+gs20 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[0,2], width_ratios=[1, 0.05], wspace=0.1)
+
+ax0 = plt.subplot(gs00[0])
+ax1 = plt.subplot(gs10[0])
+ax2 = plt.subplot(gs20[0])
+cax2 = plt.subplot(gs20[1])
+
+for ax, col_label, fig_label  in zip([ax0, ax1, ax2], col_labels, fig_labels):
+    ax.annotate(col_label, xy=(0,0), xytext=(0.5, 1.05), textcoords='axes fraction',
+                va='bottom', ha='center', fontsize=14, annotation_clip=False)
+    ax.annotate(fig_label, xy=(0,0), xytext=(0, 1.05), textcoords='axes fraction',
+                va='bottom', ha='center', fontsize=14, annotation_clip=False)
+    
+for ax in [ax0, ax1, ax2]:
+    ax.tick_params(axis='both', labelsize=14)
+for cax in [cax2]:
+    cax.tick_params(axis='both', labelsize=14)
+
+# Calculate a list of points to sample in region
+n = 1.33
[…]

### 2017-08-23 paper/paper/paper.tex
Shalin "present address"
Fixed ellipse bug. Ellipse plots in paper with explanations.

@@
   \authormark{2}Marine Biological Laboratory, Bell Center, Woods Hole, Massachusetts 02543, USA\\
-  \authormark{3}Chan Zuckerberg Biohub, San Francisco, California 94158, USA\\
+  \authormark{3}(present address) Chan Zuckerberg Biohub, San Francisco, California 94158, USA\\
   \authormark{4}Section on High Resolution Optical Imaging, National Institute
@@
 
+We also use the Fisher information matrix to calculate uncertainty ellipses for
+each microscope and fluorophore orientation. The general equation of an ellipse
+with axis radii $a$ and $b$ and rotation angle $\gamma$ is
+\begin{align}
+  \frac{(x\cos\gamma + y\sin\gamma)^2}{a^2} + \frac{(x\cos\gamma - y\sin\gamma)^2}{b^2} =1. 
+\end{align}
+We calculate the ellipse parameters in terms of the Fisher information matrix
+using the following set of equations
+\begin{align}
+  \sigma_x &= \sqrt{\mb{F}^{-1}_{0,0}} \label{eq:first}\\
+  \sigma_y &= \sin\Theta\sqrt{\mb{F}^{-1}_{1,1}} \label{eq:second}\\ 
+  \sigma_{xy} &= \sin\Theta\sqrt{\mb{F}^{-1}_{1,0}} \label{eq:third}\\  
+  a^2 &= \frac{\sigma_x^2 + \sigma_y^2}{2} + \sqrt{\frac{(\sigma_x^2 - \sigma_y^2)^2}{4} + \sigma_{xy}^2}\\
+  b^2 &= \frac{\sigma_x^2 + \sigma_y^2}{2} - \sqrt{\frac{(\sigma_x^2 - \sigma_y^2)^2}{4} + \sigma_{xy}^2}\\
+  \gamma &= \frac{1}{2}\text{arctan}\left(\frac{2\sigma_{xy}}{\sigma_x^2 - \sigma_y^2}\right). \label{eq:last}
+\end{align}
+Equations \ref{eq:first}--\ref{eq:last} are identical to those found in
+\cite{coe2009} with an additional factor of $\sin\Theta$ on equations
+\ref{eq:second} and \ref{eq:third} to account for the non-Euclidean geometry of
+the sphere. Instead of plotting ellipses with widely varying radii on the same
+figure, we chose a constant major axis radius. This means that the uncertainty
+ellipses indicate the relative size of uncertainty in different directions, not
+the absolute size of the uncertainty.
+
+We used NumPy \cite{walt2011} for numerical computation, SymPy \cite{meurer2017}
+for symbolic computation, Matplotlib \cite{hunter2007} for high-level plotting,
+VisPy \cite{vispy} for 3D rendered plots, and Asymptote \cite{bowman2008} for 3D line plots.
+
 \section{Results}\label{results}
@@
 a small illumination NA and a large detection NA where
-$\alpha \approx 65^{\circ}$ (see Figure \ref{fig:single-arm}d) at
+$\alpha \approx 65^{\circ}$ (see Figure \ref{fig:single-arm}e) at
 NA${}_{\text{det}} \approx 1.2$). A small illumination NA maximizes the range of
@@
 modification requiring a higher-NA objective). Also note in Figure
-\ref{fig:single-arm}b) that single-view designs suffer from high orientation
+\ref{fig:single-arm}c) that single-view designs suffer from high orientation
 uncertainty when the molecules are oriented along the optical axis (the
@@
 (it is ambiguous whether the molecule is oriented above or below the transverse
-plane), and near the polarizer orientations (it is ambiguous whether the
-molecule is on either side of the polarizer orientation).
+plane, see Figure \ref{fig:single-arm}b)), and near the polarizer orientations
+(it is ambiguous whether the molecule is on either side of the polarizer
+orientation, see Figure \ref{fig:single-arm}b)).
 
@@
   red arrows indicate the four illumination polarization orientations---one for
-  each intensity measurement. b) Solid-angle uncertainty for the microscope in
-  a). c) Median of the solid-angle uncertainty as a function of illumination and
-  detection NA. d) MAD of the solid-angle uncertainty as a function of
-  illumination and detection NA. The microscope configuration in a) and b) is
-  indicated by a cross in c) and d).}
+  each intensity measurement. b) Uncertainty ellipses for the microscope in
+  a). The ellipses indicate the relative size of the uncertainty in different
+  directions, not the absolute size of the uncertainty. c) Solid-angle
+  uncertainty for the microscope in a). The solid-angle uncertainty is a measure
+  of the absolute size of the uncertainty in all directions. d) Median of the
+  solid-angle uncertainty as a function of illumination and detection NA. e) MAD
+  of the solid-angle uncertainty as a function of illumination and detection
+  NA. The microscope configuration in a), b), and c) is indicated by a cross in
+  d) and e).}
 \label{fig:single-arm}
@@
 against the cover slip. Given the high uncertainty in the transverse plane for a
-single-view shown in Figure \ref{fig:single-arm}b), we can see why oblique views
+single-view shown in Figure \ref{fig:single-arm}c), we can see why oblique views
 perform better than orthogonal views---oblique views are complementary because
@@
   we illuminate from the second objective (blue solid) and detect from the first
-  objective (blue dashed). b) Solid-angle uncertainty for the microscope in
-  a). c) Median of the solid-angle uncertainty as a function of NA and the angle
-  between the objectives. d) MAD of the solid-angle uncertainty as a function of
-  NA and the angle between the objectives. The microscope configuration in a)
-  and b) is indicated by a cross in c) and d). e) Median of the solid-angle
-  uncertainty as a function of the angle between the objectives when NA=0.6. f)
-  MAD of the solid-angle uncertainty as a function of the angle between the
-  objectives when NA=0.6. The profile in e) and f) is taken along the purple
-  line in c) and d).}
+  objective (blue dashed). b) Uncertainty ellipses for the microscope in a). The
+  ellipses indicate the relative size of the uncertainty in different
+  directions, not the absolute size of the uncertainty. c) Solid-angle
+  uncertainty for the microscope in a). The solid-angle uncertainty is a measure
+  of the absolute size of the uncertainty in all directions. d) Median of the
+  solid-angle uncertainty as a function of NA and the angle between the
+  objectives. e) MAD of the solid-angle uncertainty as a function of NA and the
+  angle between the objectives. The microscope configuration in a), b), and c)
+  is indicated by a cross in d) and e). f) Median of the solid-angle uncertainty
+  as a function of the angle between the objectives when NA=0.6. g) MAD of the
+  solid-angle uncertainty as a function of the angle between the objectives when
+  NA=0.6. The profile in f) and g) is taken along the purple line in d) and e).}
 \label{fig:symmetric-widefield}
@@
 \caption{Dual-view symmetric orthogonal K\"ohler designs with varying
-  illumination and detection NA. a) Schematic of the microscope. b) Solid-angle
-  uncertainty for the microscope in a). c) Median of the solid-angle uncertainty
-  as a function of illumination and detection NA. d) MAD of the solid-angle
-  uncertainty as a function of illumination and detection NA. The microscope
-  configuration in a) and b) is indicated by a cross in c) and d).}
+  illumination and detection NA. a) Schematic of the microscope. b) Uncertainty
+  ellipses for the microscope in a). The ellipses indicate the relative size of
+  the uncertainty in different directions, not the absolute size of the
+  uncertainty. c) Solid-angle uncertainty for the microscope in a). The
+  solid-angle uncertainty is a measure of the absolute size of the uncertainty
+  in all directions. d) Median of the solid-angle uncertainty as a function of
+  illumination and detection NA. e) MAD of the solid-angle uncertainty
+  as a function of illumination and detection NA. The microscope configuration
+  in a), b), and c) is indicated by a cross in d) and e).}
 \label{fig:double-arm}
@@
 larger than the exposure from the high-NA arm (top-left and bottom-right corners
-of Figures \ref{fig:asymmetric-double}c) and d)) the median and MAD are
+of Figures \ref{fig:asymmetric-double}d) and e)) the median and MAD are
 comparable to the symmetric light-sheet microscope. This means that if a very
@@
     and sample exposure asymmetry. a) Schematic of the microscope. b)
-    Solid-angle uncertainty for the microscope in a). c) Median of the
-    solid-angle uncertainty as a function of NA and sample exposure
-    asymmetry. d) MAD of the solid-angle uncertainty as a function of NA and
-    sample exposure asymmetry. The microscope configuration in a) and b) is
-    indicated by a cross in c) and d).}
+    Uncertainty ellipses for the microscope in a). The ellipses indicate the
+    relative size of the uncertainty in different directions, not the absolute
+    size of the uncertainty. c) Solid-angle uncertainty for the microscope in
+    a). The solid-angle uncertainty is a measure of the absolute size of the
+    uncertainty in all directions. d) Median of the solid-angle uncertainty as a
+    function of NA and sample exposure asymmetry. e) MAD of the solid-angle
+    uncertainty as a function of NA and sample exposure asymmetry. The
+    microscope configuration in a), b), and c) is indicated by a cross in d) and e).}
 \label{fig:asymmetric-double}

### 2017-08-23 dipsim/multiframe.py

@@ -12,32 +12,37 @@ class MultiFrameMicroscope:
     NoiseModel.
     """
     def __init__(self, ill_thetas, det_thetas, ill_nas, det_nas, ill_types,
-                 det_types, colors, n_frames=4, n_pts=1000, max_photons=1000,
-                 n_samp=1.33, **kwargs):
+                 det_types, colors, n_frames=4, n_det_frames=1,
+                 n_pts=1000, max_photons=1000, n_samp=1.33, **kwargs):
+        
         self.n_frames = n_frames
+        self.n_det_frames = n_det_frames
         self.n_pts = n_pts
 
         m = []
         for i, det_theta in enumerate(det_thetas):
-            # Specify detection
-            det = detector.Detector(theta_optical_axis=det_thetas[i],
-                                    det_type=det_types[i],
-                                    na=det_nas[i], n=n_samp)
+            # Cycle through detection polarizations
+            for k in range(self.n_det_frames):
+                phi_det = np.pi*k/self.n_det_frames
+                det = detector.Detector(theta_optical_axis=det_thetas[i],
+                                        det_type=det_types[i],
+                                        na=det_nas[i], n=n_samp,
+                                        phi_pol=phi_det)
 
-            # Cycle through polarizations
-            for n in range(self.n_frames):
-                phi = np.pi*n/self.n_frames
-                ill = illuminator.Illuminator(theta_optical_axis=ill_thetas[i],
-                                              illum_type=ill_types[i],
-                                              na=ill_nas[i], n=n_samp,
-                                              phi_pol=phi)
+                # Cycle through illumination polarizations
+                for n in range(self.n_frames):
+                    phi_ill = np.pi*n/self.n_frames
+                    ill = illuminator.Illuminator(theta_optical_axis=ill_thetas[i],
+                                                  illum_type=ill_types[i],
+                                                  na=ill_nas[i], n=n_samp,
+                                                  phi_pol=phi_ill)
 
-                if type(max_photons) == list:
-                    max_photon = max_photons[i]
-                else:
-                    max_photon = max_photons
-                    
[…]

### 2017-08-23 paper/figures/asymmetric-double.py

@@ -7,8 +7,8 @@ import os; import time; start = time.time(); print('Running...')
 import matplotlib.gridspec as gridspec
 
 # Main input parameters
-col_labels = ['Geometry\n($\\alpha_{1}$ = 60${}^{\circ}$,\n Exposure${}_1$/Exposure${}_2$ = 5)', r'$\sigma_{\Omega}$ [sr]', 'Median$\{\sigma_{\Omega}\}$ [sr]', 'MAD$\{\sigma_{\Omega}\}$ [sr]']
-fig_labels = ['a)', 'b)', 'c)', 'd)']
+col_labels = ['Geometry\n($\\alpha_{1}$ = 60${}^{\circ}$,\n Exposure${}_1$/Exposure${}_2$ = 5)', 'Uncertainty Ellipses', r'$\sigma_{\Omega}$ [sr]', 'Median$\{\sigma_{\Omega}\}$ [sr]', 'MAD$\{\sigma_{\Omega}\}$ [sr]']
+fig_labels = ['a)', 'b)', 'c)', 'd)', 'e)']
 n_pts = 500 # Points on sphere
 n_pts_sphere = 50000 # Points on sphere
 n_grid_pts = 25
@@ -16,30 +16,30 @@ inch_fig = 5
 dpi = 300
 
 # Setup figure and axes
-fig = plt.figure(figsize=(2.4*inch_fig, 2*inch_fig))
-gs0 = gridspec.GridSpec(2, 2, wspace=0.3, hspace=0.1)
-gs00 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[0,0], width_ratios=[1, 0.05], wspace=0.45)
-gs10 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[1,0], width_ratios=[1, 0.05], wspace=0.45)
-gs01 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[0,1], width_ratios=[1, 0.05], wspace=0.45)
-gs11 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[1,1], width_ratios=[1, 0.05], wspace=0.45)
-ax0 = plt.subplot(gs00[0])
-ax1 = plt.subplot(gs01[0])
-ax2 = plt.subplot(gs10[0])
-ax3 = plt.subplot(gs11[0])
[…]

### 2017-08-23 paper/figures/double-arm.py

@@ -7,8 +7,8 @@ import os; import time; start = time.time(); print('Running...')
 import matplotlib.gridspec as gridspec
 
 # Main input parameters
-col_labels = ['Geometry\n(NA${}_{\\textrm{ill}}$ = 0.6, NA${}_{\\textrm{det}}$ = 0.8)', r'$\sigma_{\Omega}$ [sr]', 'Median$\{\sigma_{\Omega}\}$ [sr]', 'MAD$\{\sigma_{\Omega}\}$ [sr]']
-fig_labels = ['a)', 'b)', 'c)', 'd)']
+col_labels = ['Geometry\n(NA${}_{\\textrm{ill}}$ = 0.6, NA${}_{\\textrm{det}}$ = 0.8)', 'Uncertainty Ellipses', r'$\sigma_{\Omega}$ [sr]', 'Median$\{\sigma_{\Omega}\}$ [sr]', 'MAD$\{\sigma_{\Omega}\}$ [sr]']
+fig_labels = ['a)', 'b)', 'c)', 'd)', 'e)']
 n_pts = 500 # Points on sphere
 n_pts_sphere = 50000 # Points on sphere
 n_grid_pts = 25
@@ -17,29 +17,29 @@ dpi = 300
 
 # Setup figure and axes
 fig = plt.figure(figsize=(2.2*inch_fig, 2*inch_fig))
-gs0 = gridspec.GridSpec(2, 2, wspace=0.4, hspace=0.1)
-gs00 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[0,0], width_ratios=[1, 0.05], wspace=0.1)
-gs10 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[1,0], width_ratios=[1, 0.05], wspace=0.1)
-gs01 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[0,1], width_ratios=[1, 0.05], wspace=0.1)
-gs11 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[1,1], width_ratios=[1, 0.05], wspace=0.1)
-ax0 = plt.subplot(gs00[0])
-ax1 = plt.subplot(gs01[0])
-ax2 = plt.subplot(gs10[0])
-ax3 = plt.subplot(gs11[0])
-cax0 = plt.subplot(gs00[1]); cax0.axis('off');
[…]

### 2017-08-23 paper/figures/single-arm.py

@@ -7,8 +7,8 @@ import os; import time; start = time.time(); print('Running...')
 import matplotlib.gridspec as gridspec
 
 # Main input parameters
-col_labels = ['Geometry\n(NA${}_{\\textrm{ill}}$ = 0.6, NA${}_{\\textrm{det}}$ = 0.8)', r'$\sigma_{\Omega}$ [sr]', 'Median$\{\sigma_{\Omega}\}$ [sr]', 'MAD$\{\sigma_{\Omega}\}$ [sr]']
-fig_labels = ['a)', 'b)', 'c)', 'd)']
+col_labels = ['Geometry\n(NA${}_{\\textrm{ill}}$ = 0.6, NA${}_{\\textrm{det}}$ = 0.8)', 'Uncertainty Ellipses', r'$\sigma_{\Omega}$ [sr]', 'Median$\{\sigma_{\Omega}\}$ [sr]', 'MAD$\{\sigma_{\Omega}\}$ [sr]']
+fig_labels = ['a)', 'b)', 'c)', 'd)', 'e)']
 n_pts = 1000 # Points on sphere
 n_pts_sphere = 50000 # Points on sphere
 n_grid_pts = 25
@@ -17,29 +17,29 @@ dpi = 300
 
 # Setup figure and axes
 fig = plt.figure(figsize=(2.2*inch_fig, 2*inch_fig))
-gs0 = gridspec.GridSpec(2, 2, wspace=0.4, hspace=0.1)
-gs00 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[0,0], width_ratios=[1, 0.05], wspace=0.1)
-gs10 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[1,0], width_ratios=[1, 0.05], wspace=0.1)
-gs01 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[0,1], width_ratios=[1, 0.05], wspace=0.1)
-gs11 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[1,1], width_ratios=[1, 0.05], wspace=0.1)
-ax0 = plt.subplot(gs00[0])
-ax1 = plt.subplot(gs01[0])
-ax2 = plt.subplot(gs10[0])
-ax3 = plt.subplot(gs11[0])
[…]

### 2017-08-23 paper/figures/symmetric-widefield.py

@@ -7,8 +7,8 @@ import os; import time; start = time.time(); print('Running...')
 import matplotlib.gridspec as gridspec
 
 # Main input parameters
-col_labels = ['Geometry (NA = 0.6, $\\beta=80{}^{\circ}$)', r'$\sigma_{\Omega}$ [sr]', 'Median$\{\sigma_{\Omega}\}$ [sr]', 'MAD$\{\sigma_{\Omega}\}$ [sr]', '', '']
-fig_labels = ['a)', 'b)', 'c)', 'd)', 'e)', 'f)']
+col_labels = ['Geometry\n (NA = 0.6, $\\beta=80{}^{\circ}$)', 'Uncertainty Ellipses', r'$\sigma_{\Omega}$ [sr]', 'Median$\{\sigma_{\Omega}\}$ [sr]', 'MAD$\{\sigma_{\Omega}\}$ [sr]', '', '']
+fig_labels = ['a)', 'b)', 'c)', 'd)', 'e)', 'f)', 'g)']
 n_pts = 5000 #Points on sphere
 n_pts_sphere = 50000 # Points on sphere
 n_grid_pts = 21
@@ -18,34 +18,38 @@ inch_fig = 5
 dpi = 300
 
 # Setup figure and axes
-fig = plt.figure(figsize=(2.2*inch_fig, 3.0*inch_fig))
-gs0 = gridspec.GridSpec(3, 2, wspace=0.4, hspace=0.2)
-gs00 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[0,0], width_ratios=[1, 0.05], wspace=0.1)
-gs10 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[1,0], width_ratios=[1, 0.05], wspace=0.1)
-gs01 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[0,1], width_ratios=[1, 0.05], wspace=0.1)
-gs11 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[1,1], width_ratios=[1, 0.05], wspace=0.1)
-gs20 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[2,0], width_ratios=[1, 0.05], wspace=0.1)
[…]

### 2017-08-28 paper/paper/paper.tex
Include symmetries in paper and stress fixed dipoles.

@@
 
-We consider a single molecule with an absorption dipole moment
+We consider a single molecule with a fixed absorption dipole moment
 $\hat{\bs{\mu}}_{\text{abs}}$ that we express in spherical coordinates $\Theta$
@@
             -\arccos\left(\dfrac{\cos\psi\cos\Phi\sin\Theta - \sin\psi\cos\Theta}{\sqrt{1 - (\sin\psi\cos\Phi\sin\Theta + \cos\psi\cos\Theta)^2}}\right) \ \ \ &-\pi \leq \Phi < 0\\
-          \end{cases}
+          \end{cases}\label{eq:phip}
 \end{align}
@@
 
-We used NumPy \cite{walt2011} for numerical computation, SymPy \cite{meurer2017}
-for symbolic computation, Matplotlib \cite{hunter2007} for high-level plotting,
-VisPy \cite{vispy} for 3D rendered plots, and Asymptote \cite{bowman2008} for 3D line plots.
+The solid-angle uncertainty and the uncertainty ellipse are good local measures
+of orientation uncertainty, but they are insensitive to degeneracies with
+distant dipole orientations. For example, all of the intensity measurements in
+this paper are identical under inversion of the dipole orientation
+($I(\Theta, \Phi) = I(\Theta + \pi, \Phi)$ in equations \ref{eq:scalarabs},
+\ref{eq:scalardet}, and \ref{eq:forward_frame}), but the solid-angle uncertainty
+and the uncertainty ellipse give no indication of this degeneracy. For this
+reason we also use the number of degeneracies as a measure of a micrsocope's
+ability to estimate the orientation of a dipole from intensity measurements. A
+small number of degeneracies is important for a microscope to find the
+orientation of a molecule---a twofold degenerate microscope can find the
+orientation of a molecule down to the hemisphere while a fourfold symmetric
+microscope can only find the orientation down to the quadrant.
+
+To calculate and display these evaluation metrics we use NumPy \cite{walt2011}
+for numerical computation, SymPy \cite{meurer2017} for symbolic computation,
+Matplotlib \cite{hunter2007} for high-level plotting, VisPy \cite{vispy} for 3D
+rendered plots, and Asymptote \cite{bowman2008} for 3D line plots.
 
@@
 orientation, see Figure \ref{fig:single-arm}b)).
+
+The single-view microscope in Figure \ref{fig:single-arm} is fourfold degenerate
+($I(\Theta, \Phi) = I(\Theta + \pi, \Phi) = I(\Theta, \Phi + \pi) = I(\Theta +
+\pi, \Phi+\pi)$ in equations \ref{eq:scalarabs}, \ref{eq:scalardet}, and
+\ref{eq:forward_frame}). In addition to the usual inversion degeneracy,
+single-view microscopes suffer from a degeneracy when a dipole is reflected
+through the transverse plane. Evidence of this degeneracy is found in Figures
+\ref{fig:single-arm}b) and \ref{fig:single-arm}c) which show the ambiguity of
+whether the molecule is oriented above or below the transverse plane. Near the
+transverse plane this ambiguity is evident in the high solid-angle uncertainty,
+but far from the transverse plane evidence of the ambiguity is lost. Adding a
+second view and taking eight intensity measurements instead of four removes this
+degeneracy (see equations \ref{eq:thetap} and \ref{eq:phip}), so the dual-view
+microscopes we consider in this paper are only twofold degenerate. This means
+that while dual-view microscopes can find the orientation of a molecule up to
+the hemisphere, while single-view microscopes can only find the orientation of a
+molecule up to the quadrant.
 
@@
 find that increasing NA always lowers the MAD, but orthogonal arms are not
-always optimal. At large NA, it is advantageous to move the objectives together or
-against the cover slip. Given the high uncertainty in the transverse plane for a
-single-view shown in Figure \ref{fig:single-arm}c), we can see why oblique views
-perform better than orthogonal views---oblique views are complementary because
-the regions of high orientation uncertainty from each view do not overlap.
+always optimal. At large NA, it is advantageous to move the objectives together
+or against the cover slip. Given the high uncertainty in the transverse plane
+for a single-view shown in Figure \ref{fig:single-arm}c), we can see why oblique
+views perform better than orthogonal views---oblique views are complementary
+because the regions of high orientation uncertainty from each view do not
+overlap.
 
@@
 Table 1 shows a summary of our evaluation metrics for microscopes in each
-class. The single-view K\"ohler microscope with a small illumination NA and
-high detection NA has the lowest solid-angle uncertainty median but the largest
-max. The dual-view oblique symmetric K\"ohler with an intermediate NA performs
-reasonably well on all three metrics, while the dual-view orthogonal symmetric
-light-sheet microscopes with large NA perform very well on all three metrics.
+class. The single-view K\"ohler microscope with a small illumination NA and high
+detection NA has the lowest solid-angle uncertainty median but the largest max
+and fourfold symmetry. The dual-view oblique symmetric K\"ohler with an
+intermediate NA performs reasonably well on all three metrics, while the
+dual-view orthogonal symmetric light-sheet microscopes with large NA perform
+very well on all three metrics.
 
@@
 \centering
-\caption{Comparison of designs in each class of microscope. All values are in
-  steradians.}
-\begin{tabular}{llll}
+\caption{Comparison of designs in each class of microscope. All solid-angle
+  uncertainty statistics are in steradians.}
+\begin{tabular}{lllll}
   \toprule
-Microscope Type&Max\{$\sigma_{\Omega}$\}&Median\{$\sigma_{\Omega}$\}&MAD\{$\sigma_{\Omega}$\}\\ \midrule
-Single-view K\"ohler (NA${}_\textrm{ill}$=0, NA${}_\textrm{det}$=1.1)&3.04$\times 10^{0}$&8.20$\times 10^{-4}$&1.48$\times 10^{-4}$ \\ 
-  Dual-view oblique symmetric K\"ohler (NA=0.6, $\beta$=53${}^{\circ}$)&1.35$\times 10^{-1}$&4.49$\times 10^{-3}$&9.17$\times 10^{-4}$\\
-Dual-view orthogonal symmetric light-sheet (NA=0.8)&6.23$\times 10^{-3}$&1.79$\times 10^{-3}$&1.65$\times 10^{-4}$\\  
-Dual-view orthogonal symmetric light-sheet (NA=0.94)&3.13$\times 10^{-3}$&1.28$\times 10^{-3}$&1.14$\times 10^{-4}$\\
+
+  Microscope Type&$n$-fold&Max\{$\sigma_{\Omega}$\}&Median\{$\sigma_{\Omega}$\}&MAD\{$\sigma_{\Omega}$\}\\ 
+    &Degeneracy&&&\\\midrule
+Single-view K\"ohler (NA${}_\textrm{ill}$=0, NA${}_\textrm{det}$=1.1)&4&3.04$\times 10^{0}$&8.20$\times 10^{-4}$&1.48$\times 10^{-4}$ \\ 
+  Dual-view oblique symmetric K\"ohler (NA=0.6, $\beta$=53${}^{\circ}$)&2&1.35$\times 10^{-1}$&4.49$\times 10^{-3}$&9.17$\times 10^{-4}$\\
+Dual-view orthogonal symmetric light-sheet (NA=0.8)&2&6.23$\times 10^{-3}$&1.79$\times 10^{-3}$&1.65$\times 10^{-4}$\\  
+Dual-view orthogonal symmetric light-sheet (NA=0.94)&2&3.13$\times 10^{-3}$&1.28$\times 10^{-3}$&1.14$\times 10^{-4}$\\
 \bottomrule
@@
   detection NA with $\alpha \approx 65^{\circ}$.
-\item Dual-view microscopes outperform single-view microscopes, and dual-view
-  microscopes remove the high uncertainty region in the transverse plane of
-  single-view microscopes.
+\item Dual-view microscopes outperform single-view microscopes along several
+  metrics. Dual-view microscopes have fewer degeneracies, lower median and MAD
+  solid-angle uncertainty, and no high solid-angle uncertainty region in the
+  transverse plane.
 \item High-NA dual-view K\"ohler microscopes are optimal when the two views are
@@
 
-Although we have only considered single-molecule orientation determination in
-this work, polarized illumination techniques can be applied to find the
-distribution of ensembles of molecules \cite{mehta2016, backer2016}. In future
-work we will investigate the ability of multiview polarized illumination
-microscopes to estimate the orientation distribution of ensembles of molecules.
+Although we have only considered the orientation determination of fixed single
+molecules, polarized illumination techniques can be used to study the
+orientation distribution of ensembles of dynamic molecules \cite{mehta2016,
+  backer2016}. In future work we will investigate the ability of multiview
+polarized illumination microscopes to estimate the orientation distribution of
+ensembles of dynamic molecules.
 

### 2017-09-01 dipsim/multiframe.py
Polarization offset option.

@@ -13,9 +13,11 @@ class MultiFrameMicroscope:
     """
     def __init__(self, ill_thetas, det_thetas, ill_nas, det_nas, ill_types,
                  det_types, colors, n_frames=4, n_det_frames=1,
-                 n_pts=1000, max_photons=1000, n_samp=1.33, **kwargs):
+                 frame_offsets=None, n_pts=1000, max_photons=1000, n_samp=1.33,
+                 **kwargs):
         
         self.n_frames = n_frames
+        self.frame_offsets = frame_offsets
         self.n_det_frames = n_det_frames
         self.n_pts = n_pts
 
@@ -32,6 +34,10 @@ class MultiFrameMicroscope:
                 # Cycle through illumination polarizations
                 for n in range(self.n_frames):
                     phi_ill = np.pi*n/self.n_frames
+                    if self.frame_offsets != None:
+                        if self.frame_offsets[i]:
+                            phi_ill += np.pi/(2*self.n_frames)
+                            
                     ill = illuminator.Illuminator(theta_optical_axis=ill_thetas[i],
                                                   illum_type=ill_types[i],
                                                   na=ill_nas[i], n=n_samp,
@@ -82,6 +88,8 @@ class MultiFrameMicroscope:
         for pt in pts:
             fi = self.noise_model.calc_fi(pt, 2*[self.sphere_dx])
             fi_inv = np.linalg.pinv(fi)
+            if fi_inv[0,0] < 0:
+                fi_inv[0,0] = 0
             sigx = np.sqrt(fi_inv[0,0])
             sigxy = np.sin(pt[0])*fi_inv[1,0]
             sigy = np.sin(pt[0])*np.sqrt(fi_inv[1,1])

### 2017-09-01 notes/2017-08-31-polarized-detection/report/report.tex
Polarized detection notes.

\documentclass[11pt]{article}

%%%%%%%%%%%%
% Packages %
%%%%%%%%%%%%
\hyphenpenalty=10000
\usepackage{array,multirow}
\usepackage{tocloft}
\renewcommand\cftsecleader{\cftdotfill{\cftdotsep}}
\def\undertilde#1{\mathord{\vtop{\ialign{##\crcr
$\hfil\displaystyle{#1}\hfil$\crcr\noalign{\kern1.5pt\nointerlineskip}
$\hfil\tilde{}\hfil$\crcr\noalign{\kern1.5pt}}}}}
\usepackage{cleveref}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{epstopdf}
\usepackage{braket}
\usepackage{upgreek}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{amssymb,latexsym,amsmath,gensymb}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{float}
\usepackage{enumitem}
\usepackage{pdflscape}
\usepackage{url}
\usepackage{tikz, calc}
\usetikzlibrary{shapes.geometric, arrows, calc}
\tikzstyle{norm} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{arrow} = [thick, ->, >=stealth]

\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
\providecommand{\mb}[1]{\mathbf{#1}}
\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
\providecommand{\bs}[1]{\boldsymbol{#1}} 
\providecommand{\intinf}{\int_{-\infty}^{\infty}}
\providecommand{\fig}[4]{
  % filename, width, caption, label
\begin{figure}[h]
 \captionsetup{width=1.0\linewidth}
 \centering
 \includegraphics[width = #2\textwidth]{#1}
 \caption{#3}
 \label{fig:#4}
\end{figure}
}

\newcommand{\tensor}[1]{\overset{\text{\tiny$\leftrightarrow$}}{\mb{#1}}}
\newcommand{\tunderbrace}[2]{\underbrace{#1}_{\textstyle#2}}
\providecommand{\figs}[7]{
  % filename1, filename2, caption1, caption2, label1, label2, shift
\begin{figure}[H]
\centering
\begin{minipage}[b]{.45\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#1}
  \captionsetup{justification=justified, singlelinecheck=true}
  \caption{#3}
  \label{fig:#5}
\end{minipage}
\hspace{2em}
\begin{minipage}[b]{.45\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#2}
  \vspace{#7em}
  \captionsetup{justification=justified}
  \caption{#4}
  \label{fig:#6}
\end{minipage}
\end{figure}
}
\makeatletter

\providecommand{\code}[1]{
\begin{center}
\lstinputlisting{#1}
\end{center}
}

\newcommand{\crefrangeconjunction}{--}
%%%%%%%%%%%
% Spacing %
%%%%%%%%%%%
% Margins
\usepackage[
top    = 1.5cm,
bottom = 1.5cm,
left   = 1.5cm,
right  = 1.5cm]{geometry}

% Indents, paragraph space
%\usepackage{parskip} 
\setlength{\parskip}{1.5ex}

% Section spacing
\usepackage{titlesec}
\titlespacing*{\title}
{0pt}{0ex}{0ex}
\titlespacing*{\section}
{0pt}{0ex}{0ex}
\titlespacing*{\subsection}
{0pt}{0ex}{0ex}
\titlespacing*{\subsubsection}
{0pt}{0ex}{0ex}

% Line spacing
\linespread{1.1}

%%%%%%%%%%%%
% Document %
%%%%%%%%%%%%
\begin{document}
\title{\vspace{-2.5em} More Design Studies of Multiview Polarized \\Illumination
  and/or Detection Microscopy \vspace{-1em}} \author{Talon
  Chandler}% and Patrick La Rivi\`ere}
\date{\vspace{-1em}\today\vspace{-1em}}
\maketitle
\section{Introduction}
In these notes I will use the model and metrics we established in the working
paper draft to study three more design questions:
\begin{itemize}
\item What is the effect of using fewer than four illumination polarization orientation?
\item Do polarized illumination or polarized detection microscopes provide
  better orientation reconstructions?
\item How do microscopes with polarizers on both the illumination and detection
  paths compare to microscopes with polarizers on only the illumination or
  detection paths.
\end{itemize}

\section{Methods}
Figure \ref{fig:single-frame} shows the efficiencies for single measurements
taken with polarized illumination and/or detection microscopes. These results
were created with Fourkas' expressions on the detection side and my expressions
from the paper on the illumination side. Notice that the dashed arrows indicate
polarized detectors. Also notice that the detection efficiency range is only
0-0.25 instead of 0-0.5 in the paper because the polarizer on the detector
blocks half of the photons.

Unless otherwise stated, all of the designs I will compare in these notes will
use NA${}_{\text{ill}} = 0$ and NA${}_{\text{det}} = 0.8$, and all 2-view
microscopes will use orthogonal views.

All designs use the same total sample exposure. This means that
$I_{\text{exp}} = 4000/N$ where $N$ is the number of intensity
measurements. Note that $N$ includes polarized detection orientations as well as
polarized illumination orientations.

\fig{../figures/single-frame.pdf}{1.0}{Representative examples of single
  intensity measurements. Black dots indicate where the Cartesian unit vectors
  intersect the unit sphere. \newline \newline \textbf{Columns left to right:}
  1) schematics where the solid line encloses the illumination solid angle, the
  dashed line encloses the detection solid angle, the solid arrow indicates the
  transmission axis of the illumination polarizer, and the dashed arrow
  indicates the transmission axis of the detection polarizer; 2) the absorption
  efficiency; 3) the detection efficiency; 4) the total efficiency, the product
  of the absorption and detection efficiencies.\newline \newline \textbf{Rows
    top to bottom:} 1) coincident illumination ($\text{NA} = 0.8$ with
  $x$-polarized light) and detection (NA = 1.1); 2) non-coincident orthogonal
  illumination ($\text{NA} = 0.8$) and detection ($\text{NA} = 0.8$); 3)
  non-coincident 135${}^{\circ}$-separated illumination ($\text{NA} = 0.8$) and
  detection ($\text{NA} = 0.8$). All simulations use $n=1.33$.}{single-frame}

\section{Results \& Discussion}
Table 1 shows the results for microscopes with varying polarization on the
detection arm, the illumination arm, and both arms; varying numbers of views;
and varying numbers of polarization orientations per view.

\fig{../figures/compare-illumination.pdf}{0.6}{Comparing polarized illumination
  microscope designs. Each row in this figure corresponds to a row in the top
  third of Table 1.}{compare-illumination}
\fig{../figures/compare-detection.pdf}{0.6}{Comparing polarized detection
  microscope designs. Each row in this figure corresponds to a row in the middle
  third of Table 1.}{compare-illumination}
\fig{../figures/compare-both.pdf}{0.6}{Comparing polarized illumination and
  detection microscope designs. Each row in this figure corresponds to a row in
  the bottom third of Table 1. The illumination and detection polarization
  arrows in the schematic are offset by a few degrees for visualization purposes
  only---in the simulations the orientations are
  identical.}{compare-both}

\subsection{Minimum Number of Illumination Polarizations}
In the paper draft we only considered microscopes with four illumination
polarizations per view. Figure \ref{fig:compare-illumination} shows a comparison
of several polarized illumination microscopes and the top third of Table 1 shows
the summary statistics.

We find that for low-NA illumination with a single view there is no advantage to
using more than three polarization orientations. Two polarization orientations
is too few because the system suffers from an extra symmetry. Four polarization
orientations adds no extra information, and in a real system each polarized
illumination orientation requires extra switching and data processing time.

We also find that for low-NA illumination with two views that two polarization
orientations per view is sufficient for an orientation reconstruction.

Recall that these results use a Poisson noise model. I expect that with a
Poisson+Gaussian model that fewer polarization frames would be even more
advantageous because each intensity measurement will collect more photons and
each measurement would avoid the noise floor.

Figures \ref{fig:ill-pol} and \ref{fig:ill-pol-ortho} show how the number of
polarization frames affects the reconstruction when the illumination NA is
larger than 0. For large illumination NA, more polarized illumination
orientations offers slightly better reconstructions. As the illumination NA
decreases, the benefit of increasing the number of polarization orientations
vanishes.

\fig{../figures/ill-polarization.pdf}{1.0}{Single-view K\"ohler illumination
  microscope with varying number of illumination polarization orientations and
  illumination NA. b) Solid-angle uncertainty for the microscope in a). c)
  Median of the solid-angle uncertainty as a function of illumination
  polarization orientation and illumination NA. d) MAD of the solid-angle
  uncertainty as a function of illumination polarization orientation and
  illumination NA. The microscope in a) and b) is indicated by a cross in c) and
  d).}{ill-pol}

\fig{../figures/ill-polarization-ortho.pdf}{1.0}{Dual-view symmetric K\"ohler
  illumination microscope with varying number of illumination polarization
  orientations and illumination NA. b) Solid-angle uncertainty for the
  microscope in a). c) Median of the solid-angle uncertainty as a function of
  illumination polarization orientation and illumination NA. d) MAD of the
  solid-angle uncertainty as a function of illumination polarization orientation
  and illumination NA. The microscope in a) and b) is indicated by a cross in c)
  and d).}{ill-pol-ortho}

\subsection{Polarized Illumination vs. Polarized Detection}
In the paper we only considered polarized illumination microscopes and ignored
polarized detection microscopes. Figure \ref{fig:compare-illumination} shows a
comparison of several polarized detection microscopes and the middle third of
Table 1 shows the summary statistics.

Polarized illumination microscopes outperform polarized detection microscopes
with the same number of frames. The main difference between polarized detection
and polarized illumination is the size of the intensity signals for the same
sample exposure. Polarized detection signals are smaller than polarized
illumination signals because the polarizer blocks half of the photons from the
detector, so polarized detection signals suffer from a relative uncertainty.

The preference for polarized illumination would be even stronger if we used a
Poisson+Gaussian noise model because larger intensity measurements would contain
more information about the fluorophore orientation.

Note that these results only apply if we split the sample exposure equally
between intensity measurements. If a lossless beam splitter is available, then
adding the beam splitter is a ``free'' source of information about the
fluorophore orientation.

\section{Polarized Illumination and Detection}
Figure \ref{fig:compare-both} shows a comparison of several polarized
illumination+detection microscopes, and the bottom third of Table
1 shows the summary statistics.

Polarized illumination outperforms the polarized illumination+detection
microscopes. Some of the polarizer orientation combinations do not give much
information about the fluorophore orientation, so they are relatively useless
measurements. This results shows that more measurements are not always
beneficial.

Once again, these results only apply if we split the sample exposure equally
between intensity measurements. Adding a lossless beam splitter is still a good
choice if you have the option.

\begin{table}[ht!]
\centering
\caption{Comparison of designs in each class of microscope. All solid-angle
  uncertainty statistics are in steradians. }
\begin{tabular}{lllllll}
  \toprule
  Polarized Ill.&Number of &Polarization Settings&$n$-fold&Max\{$\sigma_{\Omega}$\}&Median\{$\sigma_{\Omega}$\}&MAD\{$\sigma_{\Omega}$\}\\ 
  or Det.&Views&Per View&Degeneracy&&&\\\midrule
  Illumination&1&2&8&5.10$\times 10^{00}$&1.65$\times 10^{-3}$&3.77$\times 10^{-4}$ \\ 
                &&3&4&5.10$\times 10^{00}$&1.65$\times 10^{-3}$&3.79$\times 10^{-4}$ \\
                &&4&4&5.10$\times 10^{00}$&1.65$\times 10^{-3}$&3.77$\times 10^{-4}$ \\\cline{2-7}  
                &2&2&2&6.23$\times 10^{-3}$&1.79$\times 10^{-3}$&1.66$\times 10^{-4}$\\
                &&3&2&1.24$\times 10^{-2}$&1.79$\times 10^{-3}$&1.66$\times 10^{-4}$\\  
                &&4&2&6.23$\times 10^{-3}$&1.79$\times 10^{-3}$&1.65$\times 10^{-4}$\\\hline  
  Detection&1&2&8&9.16$\times 10^{+2}$&2.62$\times 10^{-3}$&1.08$\times 10^{-3}$\\
                &&3&4&1.30$\times 10^{+1}$&2.38$\times 10^{-3}$&7.62$\times 10^{-4}$\\
                &&4&4&1.30$\times 10^{+1}$&2.28$\times 10^{-3}$&6.55$\times 10^{-4}$\\\cline{2-7}      
                &2&2&8&2.41$\times 10^{+1}$&3.20$\times 10^{-3}$&8.01$\times 10^{-4}$\\
                &&3&2&8.02$\times 10^{-2}$&2.97$\times 10^{-3}$&7.32$\times 10^{-4}$\\  
                &&4&2&8.08$\times 10^{-2}$&3.06$\times 10^{-3}$&7.96$\times 10^{-4}$\\\hline
  Both&1&2 ill. $\times$ 2 det. = 4&8&7.22$\times 10^{00}$&2.66$\times 10^{-3}$&6.63$\times 10^{-4}$\\
                &&3 ill. $\times$ 3 det. = 9&4&7.21$\times 10^{00}$&2.63$\times 10^{-3}$&6.19$\times 10^{-4}$\\
                &&2 ill. $\times$ 4 det. = 8&4&7.19$\times 10^{00}$&2.62$\times 10^{-3}$&6.04$\times 10^{-4}$\\\cline{2-7}      
                &2&2 ill. $\times$ 2 det. = 4&2&7.04$\times 10^{-3}$&2.67$\times 10^{-3}$&3.76$\times 10^{-4}$\\
                &&3 ill. $\times$ 3 det. = 9&2&1.72$\times 10^{-2}$&2.66$\times 10^{-3}$&4.84$\times 10^{-4}$\\
                &&2 ill. $\times$ 4 det. = 8&2&7.49$\times 10^{-3}$&2.70$\times 10^{-3}$&5.37$\times 10^{-4}$\\    
  \bottomrule
\end{tabular}
\end{table}

\end{document}

### 2017-09-01 notes/2017-08-31-polarized-detection/figures/compare-both.py

@@ -1,0 +1,91 @@
+from dipsim import util
+from dipsim import multiframe
+import numpy as np
+import matplotlib.pyplot as plt
+import os; import time; start = time.time(); print('Running...')
+
+# Main input parameters
+col_labels = ['Geometry', 'Uncertainty Ellipse', r'$\sigma_{\Omega}$']
+row_labels = 6*['']
+
+n_pts = 10000
+n_rows = len(row_labels)
+n_cols = 3
+inch_fig = 5
+dpi = 200
+
+m1 = multiframe.MultiFrameMicroscope(ill_thetas=[0], det_thetas=[0],
+                                    ill_nas=[0.01], det_nas=[0.8],
+                                    ill_types=['sheet'], det_types=['polarized'],
+                                    colors=['(1,0,0)'], n_det_frames=2, n_frames=2,
+                                    n_pts=n_pts, max_photons=4000/4, n_samp=1.33)
+
+m2 = multiframe.MultiFrameMicroscope(ill_thetas=[0], det_thetas=[0],
+                                    ill_nas=[0.01], det_nas=[0.8],
+                                    ill_types=['sheet'], det_types=['polarized'],
+                                    colors=['(1,0,0)'], n_det_frames=3, n_frames=3,
+                                    n_pts=n_pts, max_photons=4000/9, n_samp=1.33)
+
+m3 = multiframe.MultiFrameMicroscope(ill_thetas=[0], det_thetas=[0],
+                                    ill_nas=[0.01], det_nas=[0.8],
+                                    ill_types=['sheet'], det_types=['polarized'],
+                                    colors=['(1,0,0)'], n_det_frames=4, n_frames=2,
+                                    n_pts=n_pts, max_photons=4000/8, n_samp=1.33)
+
+m4 = multiframe.MultiFrameMicroscope(ill_thetas=[np.pi/4, -np.pi/4], det_thetas=[-np.pi/4, np.pi/4],
+                                     ill_nas=2*[0.01], det_nas=2*[0.8],
+                                     ill_types=2*['sheet'], det_types=2*['polarized'],
+                                     colors=['(1,0,0)','(0,0,1)'], n_det_frames=2, n_frames=2,
+                                     n_pts=n_pts, max_photons=4000/8, n_samp=1.33)
+
[…]

### 2017-09-01 notes/2017-08-31-polarized-detection/figures/compare-detection.py

@@ -1,0 +1,91 @@
+from dipsim import util
+from dipsim import multiframe
+import numpy as np
+import matplotlib.pyplot as plt
+import os; import time; start = time.time(); print('Running...')
+
+# Main input parameters
+col_labels = ['Geometry', 'Uncertainty Ellipse', r'$\sigma_{\Omega}$']
+row_labels = 6*['']
+
+n_pts = 10000
+n_rows = len(row_labels)
+n_cols = 3
+inch_fig = 5
+dpi = 200
+
+m1 = multiframe.MultiFrameMicroscope(ill_thetas=[0], det_thetas=[0],
+                                    ill_nas=[0.01], det_nas=[0.8],
+                                    ill_types=['unpolarized'], det_types=['polarized'],
+                                    colors=['(1,0,0)'], n_det_frames=2, n_frames=1,
+                                    n_pts=n_pts, max_photons=4000/2, n_samp=1.33)
+
+m2 = multiframe.MultiFrameMicroscope(ill_thetas=[0], det_thetas=[0],
+                                    ill_nas=[0.01], det_nas=[0.8],
+                                    ill_types=['unpolarized'], det_types=['polarized'],
+                                    colors=['(1,0,0)'], n_det_frames=3, n_frames=1,
+                                    n_pts=n_pts, max_photons=4000/3, n_samp=1.33)
+
+m3 = multiframe.MultiFrameMicroscope(ill_thetas=[0], det_thetas=[0],
+                                    ill_nas=[0.01], det_nas=[0.8],
+                                    ill_types=['unpolarized'], det_types=['polarized'],
+                                    colors=['(1,0,0)'], n_det_frames=4, n_frames=1,
+                                    n_pts=n_pts, max_photons=4000/4, n_samp=1.33)
+
+m4 = multiframe.MultiFrameMicroscope(ill_thetas=[np.pi/4, -np.pi/4], det_thetas=[-np.pi/4, np.pi/4],
+                                     ill_nas=2*[0.01], det_nas=2*[0.8],
+                                     ill_types=2*['unpolarized'], det_types=2*['polarized'],
+                                     colors=['(1,0,0)','(0,0,1)'], n_det_frames=2, n_frames=1,
+                                     n_pts=n_pts, max_photons=4000/4, n_samp=1.33)
+
[…]

### 2017-09-01 notes/2017-08-31-polarized-detection/figures/compare-illumination.py

@@ -1,0 +1,91 @@
+from dipsim import util
+from dipsim import multiframe
+import numpy as np
+import matplotlib.pyplot as plt
+import os; import time; start = time.time(); print('Running...')
+
+# Main input parameters
+col_labels = ['Geometry', 'Uncertainty Ellipse', r'$\sigma_{\Omega}$']
+row_labels = 6*['']
+
+n_pts = 10000
+n_rows = len(row_labels)
+n_cols = 3
+inch_fig = 5
+dpi = 200
+
+m1 = multiframe.MultiFrameMicroscope(ill_thetas=[0], det_thetas=[0],
+                                    ill_nas=[0.01], det_nas=[0.8],
+                                    ill_types=['sheet'], det_types=['lens'],
+                                    colors=['(1,0,0)'], n_frames=2,
+                                    n_pts=n_pts, max_photons=4000/2, n_samp=1.33)
+
+m2 = multiframe.MultiFrameMicroscope(ill_thetas=[0], det_thetas=[0],
+                                    ill_nas=[0.01], det_nas=[0.8],
+                                    ill_types=['sheet'], det_types=['lens'],
+                                    colors=['(1,0,0)'], n_frames=3,
+                                    n_pts=n_pts, max_photons=4000/3, n_samp=1.33)
+
+m3 = multiframe.MultiFrameMicroscope(ill_thetas=[0], det_thetas=[0],
+                                    ill_nas=[0.01], det_nas=[0.8],
+                                    ill_types=['sheet'], det_types=['lens'],
+                                    colors=['(1,0,0)'], n_frames=4,
+                                    n_pts=n_pts, max_photons=4000/4, n_samp=1.33)
+
+m4 = multiframe.MultiFrameMicroscope(ill_thetas=[np.pi/4, -np.pi/4], det_thetas=[-np.pi/4, np.pi/4],
+                                     ill_nas=2*[0.01], det_nas=2*[0.8],
+                                     ill_types=2*['sheet'], det_types=2*['lens'],
+                                     colors=['(1,0,0)','(0,0,1)'], n_frames=2,
+                                     n_pts=n_pts, max_photons=4000/4, n_samp=1.33)
+
+m5 = multiframe.MultiFrameMicroscope(ill_thetas=[np.pi/4, -np.pi/4], det_thetas=[-np.pi/4, np.pi/4],
[…]

### 2017-09-01 notes/2017-08-31-polarized-detection/figures/det-polarization-ortho.py

@@ -1,0 +1,163 @@
+from dipsim import multiframe, util
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+import matplotlib.patches as patches
+import os; import time; start = time.time(); print('Running...')
+import matplotlib.gridspec as gridspec
+
+# Main input parameters
+col_labels = ['Geometry (NA${}_{\\textrm{ill}}$ = 0.2, NA${}_{\\textrm{det}}$ = 0.8\n 3 detection polarizations)', 'Uncertainty Ellipse', r'$\sigma_{\Omega}$ [sr]', 'Median$\{\sigma_{\Omega}\}$ [sr]', 'MAD$\{\sigma_{\Omega}\}$ [sr]']
+fig_labels = ['a)', 'b)', 'c)', 'd)', 'e)']
+n_pts = 1000 # Points on sphere
+n_pts_sphere = 20000 # Points on sphere
+n_grid_pts = 10
+inch_fig = 5
+dpi = 300
+
+# Setup figure and axes
+fig = plt.figure(figsize=(2.2*inch_fig, 2*inch_fig))
+gs0 = gridspec.GridSpec(2, 1, wspace=0, hspace=0.2, height_ratios=[0.9,1])
+gs_up = gridspec.GridSpecFromSubplotSpec(1, 4, subplot_spec=gs0[0], width_ratios=[1, 1, 1, 0.06], wspace=0.1)
+gs_middle = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[1], width_ratios=[1, 1], wspace=0.4)
+gs_middle_left = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs_middle[0], width_ratios=[1, 0.05], wspace=0.1)
+gs_middle_right = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs_middle[1], width_ratios=[1, 0.05], wspace=0.1)
+
+ax0 = plt.subplot(gs_up[0])
+ax1 = plt.subplot(gs_up[1])
+ax2 = plt.subplot(gs_up[2])
+cax2 = plt.subplot(gs_up[3])
+ax3 = plt.subplot(gs_middle_left[0])
+cax3 = plt.subplot(gs_middle_left[1])
+ax4 = plt.subplot(gs_middle_right[0])
+cax4 = plt.subplot(gs_middle_right[1])
+
[…]

### 2017-09-01 notes/2017-08-31-polarized-detection/figures/det-polarization.py

@@ -1,0 +1,161 @@
+from dipsim import multiframe, util
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+import matplotlib.patches as patches
+import os; import time; start = time.time(); print('Running...')
+import matplotlib.gridspec as gridspec
+
+# Main input parameters
+col_labels = ['Geometry (NA${}_{\\textrm{ill}}$ = 0.2, NA${}_{\\textrm{det}}$ = 0.8\n 3 detection polarizations)', 'Uncertainty Ellipse', r'$\sigma_{\Omega}$ [sr]', 'Median$\{\sigma_{\Omega}\}$ [sr]', 'MAD$\{\sigma_{\Omega}\}$ [sr]']
+fig_labels = ['a)', 'b)', 'c)', 'd)', 'e)']
+n_pts = 1000 # Points on sphere
+n_pts_sphere = 20000 # Points on sphere
+n_grid_pts = 10
+inch_fig = 5
+dpi = 300
+
+fig = plt.figure(figsize=(2.2*inch_fig, 2*inch_fig))
+gs0 = gridspec.GridSpec(2, 1, wspace=0, hspace=0.2, height_ratios=[0.9,1])
+gs_up = gridspec.GridSpecFromSubplotSpec(1, 4, subplot_spec=gs0[0], width_ratios=[1, 1, 1, 0.06], wspace=0.1)
+gs_middle = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[1], width_ratios=[1, 1], wspace=0.4)
+gs_middle_left = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs_middle[0], width_ratios=[1, 0.05], wspace=0.1)
+gs_middle_right = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs_middle[1], width_ratios=[1, 0.05], wspace=0.1)
+
+ax0 = plt.subplot(gs_up[0])
+ax1 = plt.subplot(gs_up[1])
+ax2 = plt.subplot(gs_up[2])
+cax2 = plt.subplot(gs_up[3])
+ax3 = plt.subplot(gs_middle_left[0])
+cax3 = plt.subplot(gs_middle_left[1])
+ax4 = plt.subplot(gs_middle_right[0])
+cax4 = plt.subplot(gs_middle_right[1])
+
[…]

### 2017-09-01 notes/2017-08-31-polarized-detection/figures/ill-polarization-ortho.py

@@ -1,0 +1,164 @@
+from dipsim import multiframe, util
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+import matplotlib.patches as patches
+import os; import time; start = time.time(); print('Running...')
+import matplotlib.gridspec as gridspec
+
+# Main input parameters
+col_labels = ['Geometry (NA${}_{\\textrm{ill}}$ = 0.2, NA${}_{\\textrm{det}}$ = 0.8\n 3 illumination polarizations)', 'Uncertainty Ellipse', r'$\sigma_{\Omega}$ [sr]', 'Median$\{\sigma_{\Omega}\}$ [sr]', 'MAD$\{\sigma_{\Omega}\}$ [sr]']
+fig_labels = ['a)', 'b)', 'c)', 'd)', 'e)']
+n_pts = 1000 # Points on sphere
+n_pts_sphere = 20000 # Points on sphere
+n_grid_pts = 10
+inch_fig = 5
+dpi = 300
+
+# Setup figure and axes
+fig = plt.figure(figsize=(2.2*inch_fig, 2*inch_fig))
+gs0 = gridspec.GridSpec(2, 1, wspace=0, hspace=0.2, height_ratios=[0.9,1])
+gs_up = gridspec.GridSpecFromSubplotSpec(1, 4, subplot_spec=gs0[0], width_ratios=[1, 1, 1, 0.06], wspace=0.1)
+gs_middle = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[1], width_ratios=[1, 1], wspace=0.4)
+gs_middle_left = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs_middle[0], width_ratios=[1, 0.05], wspace=0.1)
+gs_middle_right = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs_middle[1], width_ratios=[1, 0.05], wspace=0.1)
+
+ax0 = plt.subplot(gs_up[0])
+ax1 = plt.subplot(gs_up[1])
+ax2 = plt.subplot(gs_up[2])
+cax2 = plt.subplot(gs_up[3])
+ax3 = plt.subplot(gs_middle_left[0])
+cax3 = plt.subplot(gs_middle_left[1])
+ax4 = plt.subplot(gs_middle_right[0])
+cax4 = plt.subplot(gs_middle_right[1])
+
[…]

### 2017-09-01 notes/2017-08-31-polarized-detection/figures/ill-polarization-two.py

@@ -1,0 +1,162 @@
+from dipsim import multiframe, util
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+import matplotlib.patches as patches
+import os; import time; start = time.time(); print('Running...')
+import matplotlib.gridspec as gridspec
+
+# Main input parameters
+col_labels = ['Geometry (NA${}_{\\textrm{ill}}$ = 0.2, NA${}_{\\textrm{det}}$ = 0.8\n 3 illumination polarizations)', 'Uncertainty Ellipses', r'$\sigma_{\Omega}$ [sr]', 'Median$\{\sigma_{\Omega}\}$ [sr]', 'MAD$\{\sigma_{\Omega}\}$ [sr]']
+fig_labels = ['a)', 'b)', 'c)', 'd)', 'e)']
+n_pts = 1000 # Points on sphere
+n_pts_sphere = 20000 # Points on sphere
+n_grid_pts = 10
+inch_fig = 5
+dpi = 300
+
+# Setup figure and axes
+fig = plt.figure(figsize=(2.2*inch_fig, 2*inch_fig))
+gs0 = gridspec.GridSpec(2, 1, wspace=0, hspace=0.2, height_ratios=[0.9,1])
+gs_up = gridspec.GridSpecFromSubplotSpec(1, 4, subplot_spec=gs0[0], width_ratios=[1, 1, 1, 0.06], wspace=0.1)
+gs_middle = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[1], width_ratios=[1, 1], wspace=0.4)
+gs_middle_left = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs_middle[0], width_ratios=[1, 0.05], wspace=0.1)
+gs_middle_right = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs_middle[1], width_ratios=[1, 0.05], wspace=0.1)
+
+ax0 = plt.subplot(gs_up[0])
+ax1 = plt.subplot(gs_up[1])
+ax2 = plt.subplot(gs_up[2])
+cax2 = plt.subplot(gs_up[3])
+ax3 = plt.subplot(gs_middle_left[0])
+cax3 = plt.subplot(gs_middle_left[1])
+ax4 = plt.subplot(gs_middle_right[0])
+cax4 = plt.subplot(gs_middle_right[1])
+
[…]

### 2017-09-01 notes/2017-08-31-polarized-detection/figures/ill-polarization.py

@@ -1,0 +1,162 @@
+from dipsim import multiframe, util
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+import matplotlib.patches as patches
+import os; import time; start = time.time(); print('Running...')
+import matplotlib.gridspec as gridspec
+
+# Main input parameters
+col_labels = ['Geometry (NA${}_{\\textrm{ill}}$ = 0.2, NA${}_{\\textrm{det}}$ = 0.8\n 3 illumination polarizations)', 'Uncertainty Ellipses', r'$\sigma_{\Omega}$ [sr]', 'Median$\{\sigma_{\Omega}\}$ [sr]', 'MAD$\{\sigma_{\Omega}\}$ [sr]']
+fig_labels = ['a)', 'b)', 'c)', 'd)', 'e)']
+n_pts = 1000 # Points on sphere
+n_pts_sphere = 20000 # Points on sphere
+n_grid_pts = 10
+inch_fig = 5
+dpi = 300
+
+# Setup figure and axes
+fig = plt.figure(figsize=(2.2*inch_fig, 2*inch_fig))
+gs0 = gridspec.GridSpec(2, 1, wspace=0, hspace=0.2, height_ratios=[0.9,1])
+gs_up = gridspec.GridSpecFromSubplotSpec(1, 4, subplot_spec=gs0[0], width_ratios=[1, 1, 1, 0.06], wspace=0.1)
+gs_middle = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs0[1], width_ratios=[1, 1], wspace=0.4)
+gs_middle_left = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs_middle[0], width_ratios=[1, 0.05], wspace=0.1)
+gs_middle_right = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=gs_middle[1], width_ratios=[1, 0.05], wspace=0.1)
+
+ax0 = plt.subplot(gs_up[0])
+ax1 = plt.subplot(gs_up[1])
+ax2 = plt.subplot(gs_up[2])
+cax2 = plt.subplot(gs_up[3])
+ax3 = plt.subplot(gs_middle_left[0])
+cax3 = plt.subplot(gs_middle_left[1])
+ax4 = plt.subplot(gs_middle_right[0])
+cax4 = plt.subplot(gs_middle_right[1])
+
[…]

### 2017-09-01 notes/2017-08-31-polarized-detection/figures/single-frame.py

@@ -1,0 +1,67 @@
+from dipsim import multiframe, util, detector, illuminator, microscope, util
+import numpy as np
+import matplotlib.pyplot as plt
+
+import os; import time; start = time.time(); print('Running...')
+
+# Main input parameters
+n_pts = 1000
+ill_types = ['unpolarized', 'unpolarized', 'wide']
+ill_pols = [0, np.pi/4, np.pi/2]
+det_pols = [0, np.pi/4, np.pi/2]
+ill_axes = [0, -np.pi/4, np.pi/4]
+det_axes = [0, np.pi/4, np.pi]
+ill_nas = [0.8, 0.8, 0.8]
+det_nas = [1.0, 0.8, 0.8]
+n_rows = len(det_axes)
+n_cols = 4
+inch_fig = 5
+dpi = 300
+
+col_labels = ['Geometry', r'Absorption Efficiency $\eta_{\textrm{abs}}$', r'Detection Efficiency $\eta_{\textrm{det}}$', r'Total Efficiency $\eta_{\textrm{tot}}$']
+row_labels = 3*['']
+
+# Generate axes
+size = (inch_fig*n_cols, inch_fig*n_rows)
+fig, axs = plt.subplots(n_rows, n_cols, figsize=size)
+plt.subplots_adjust(wspace=0.1, hspace=-0.1)
+if len(ill_pols) == 1:
+    axs = np.expand_dims(axs, 1)
+caxs = util.generate_caxs(axs)
+
+# Compute and plot on axes
+for i, (ill_type, det_pol, ill_pol, det_axis, ill_axis, det_na, ill_na) in enumerate(zip(ill_types, det_pols, ill_pols, det_axes, ill_axes, det_nas, ill_nas)):
+    print('Computing microscope: ' + str(i))
+
+    # Create microscope
+    ill = illuminator.Illuminator(illum_type=ill_type,
+                                  theta_optical_axis=ill_axis,
+                                  na=ill_na,
+                                  phi_pol=ill_pol)
+    
+    det = detector.Detector(det_type='polarized',
+                            theta_optical_axis=np.array(det_axis),
+                            na=det_na,
+                            phi_pol=det_pol)
+                                  
+    m = microscope.Microscope(illuminator=ill, detector=det, max_photons=1)
+    
+    # Plot scene and efficiencies
+    m.plot_excitation_efficiency(n=n_pts, my_ax=axs[i,1], my_cax=caxs[i,1],
[…]

### 2017-09-02 paper/paper/paper.tex
Rudolf's edits. No Kohler, proportionality, several wording upgrades

@@
   microscopes. We also find that choosing a small illumination NA and a large
-  detection NA are good design choices, that multiview K\"ohler illumination
-  microscopes can benefit from oblique illumination and detection, and that
-  asymmetric NA microscopes can benefit from exposure asymmetry.
+  detection NA are good design choices, that multiview microscopes can benefit
+  from oblique illumination and detection, and that asymmetric NA microscopes
+  can benefit from exposure asymmetry.
 \end{abstract*}
@@
 microscopes also provide a small and uniform orientation uncertainty for single
-molecules in all orientations. In section \ref{methods} we develop the theory
-required to model polarized illumination microscopes, and we develop metrics to
-compare microscope designs. In section \ref{results} we compare the results for
-a wide range of multiview microscope designs. Finally, in section
-\ref{discussion} we discuss the results and their impact on polarized multiview
-microscope design.
+molecules in all orientations. We also find that single-view microscopes have
+symmetries that can be broken by adding an second view. This reduction of
+degeneracy allows multiview microscopes to uniquely determine the orientation of
+single molecules.
+
+In section \ref{methods} we develop the theory required to model polarized
+illumination microscopes, and we develop metrics to compare microscope
+designs. In section \ref{results} we compare the results for a wide range of
+multiview microscope designs. Finally, in section \ref{discussion} we discuss
+the results and their impact on polarized multiview microscope design.
 
@@
 \ref{fig:coordinates}b)). Finally, we place a spatially incoherent and spatially
-uniform light source (or its image) in the aperture plane to establish K\"ohler
-illumination. In this geometry each point in the sample is illuminated by a
-range of angles that can be changed by adjusting the radius of the stop in the
+uniform light source (or its image) in the aperture plane to illuminate the
+focal plane. In this geometry each point in the sample is illuminated by a range
+of angles that can be changed by adjusting the radius of the stop in the
 aperture plane. We denote the set of illuminated angles by $\Omega$ (see Figure
@@
 
-    To find the absorption efficiency of a single molecule under K\"ohler
-    illumination, we integrate over all illumination rays and divide by the
-    total incident power which gives the vector expression
+    To find the absorption efficiency of a single molecule, we integrate over
+    all illumination rays and divide by the total incident power which gives the
+    vector expression
 \begin{align}
@@
 in the sample. If we scan the beam in the sample with a velocity much less than
-the beam waist radius divided by the coherence time, then we can ignore the
-coherence effects of scanning---the light sheet is effectively composed of many
-incoherent beams. If we also ensure that the laser beam is weakly focused, we
-can ignore the excitation caused by the small fields aligned with the optical
-axis. Under slowly-scanned and weakly-focused light-sheet illumination,
-molecules in the sample will be excited as if a single plane wave was
-incident. Therefore, we can find the absorption efficiency of a molecule in this
-case by taking the limit of equation \ref{eq:scalarabs} as
+the beam waist radius divided by the coherence time, then the coherence of the
+laser is rendered ineffectual by the scanning---the light sheet is effectively
+composed of many incoherent beams. If we also ensure that the laser beam is
+weakly focused, we can ignore the excitation caused by the small fields aligned
+with the optical axis. Under slowly-scanned and weakly-focused light-sheet
+illumination, molecules in the sample will be excited as if a single plane wave
+was incident. Therefore, we can find the absorption efficiency of a molecule in
+this case by taking the limit of equation \ref{eq:scalarabs} as
 $\alpha \rightarrow 0$ (or by plugging equation \ref{eq:mu} and
@@
 which is recognizable as Malus' law generalized to three dimensions. This means
-that light-sheet illumination is approximately equivalent to low-NA K\"ohler
-illumination with respect to the molecular orientations they excite. Notice that
-the absorption efficiency can take its maximum range of values
+that light-sheet illumination is approximately equivalent to low-NA illumination
+with respect to the molecular orientations they excite. Notice that the
+absorption efficiency can take its maximum range of values
 ($\eta_{\text{abs}} \in [0, 1]$) when $\alpha=0$.
@@
 \subsection{Detection Efficiency}\label{detection}
-Fourkas calculated the detection efficiency of a single molecule when an
-objective with a polarizer is aligned along the $\mh{z}$ axis
-\cite{fourkas2001}. We use his expressions to calculate the detection efficiency
-without a polarizer as
+Fourkas calculated the detection efficiency of a single molecule when an ideal
+objective with its optical axis aligned with the $\mh{z}$ axis is focused on the
+molecule and a polarizer is placed in the aperture plane\cite{fourkas2001}. We
+use his expressions to calculate the detection efficiency without a polarizer as
 \begin{align}
@@
 Fourkas' expressions were incorrectly normalized (the limit of
-$\eta_{\text{det}}$ as $\alpha\rightarrow \pi/2$ should be 1), and the extra
-factor of $\frac{3}{2}$ corrects the error. This correction does not affect
-Fourkas' orientation reconstruction but it does mean that Fourkas' algorithm
-underpredicts the total emitted intensity by a factor of $\frac{3}{2}$.
+$\eta_{\text{det}}$ as $\alpha\rightarrow \pi/2$ should be $\frac{1}{2}$), and
+the extra factor of $\frac{3}{2}$ corrects the error. This correction does not
+affect Fourkas' orientation reconstruction, but it does mean that Fourkas'
+algorithm underpredicts the total emitted intensity by a factor of
+$\frac{3}{2}$.
 
@@
 when $B$ takes its maximum value, i.e. when
-$\alpha=\arccos\left(\frac{1}{\sqrt{3}}\right) \approx 54.7^{\circ}$.
+$\alpha=\arccos\left(\frac{1}{\sqrt{3}}\right) \approx 54.7^{\circ}$. When
+$\alpha = 90$ the detection efficiency is $\frac{1}{2}$ for all fluorophore
+orientations and when $\alpha = 0$ the detection efficiency is zero for all
+fluorophore orientations, so we expect there to be an intermediate value of
+$\alpha$ where the detection efficiency takes its maximum range.
 
@@
 illumination and detection objectives had $\mh{z}$ aligned optical axes, i.e.,
-the same objective was used for illumination and detection. To extend the
-forward model to oblique optical axes which allows for non-coincident
-illumination and detection objectives, we express the dipole orientation in a
-rotated coordinate frame using the following expressions
+we used epi- or transmitted light illumination. To extend the forward model to
+oblique optical axes which allows for non-coincident illumination and detection
+objectives, we express the dipole orientation in a rotated coordinate frame
+using the following expressions
 \begin{align}
@@
 \subsection{Intensity Measurements}\label{forward}
-The detected intensity is the product of the the sample exposure
+The detected intensity is proportional to the product of the sample exposure
 $I_{\text{exp}}$, the absorption efficiency, and the detection efficiency
 \begin{align}
-  I = I_{\text{exp}}\eta_{\text{abs}}\eta_{\text{det}}\label{eq:forward_frame}.
+  I \propto I_{\text{exp}}\eta_{\text{abs}}\eta_{\text{det}}\label{eq:forward_frame},
 \end{align}
-Equation \ref{eq:forward_frame} is the forward model for a single intensity
-measurement. Figure \ref{fig:single-frame} shows the efficiencies in three
-representative geometries. Note that in all cases there are multiple
-orientations that give the same intensity measurement. This means that a single
-intensity measurement does not give us enough information to reconstruct the
-orientation of a single molecule. Next, we will consider combining several
-intensity measurements to create complete polarized illumination microscopes.
+where the proportionality depends on the exposure time, the molecular quantum
+efficiencies, and other constant factors. Equation \ref{eq:forward_frame} is the
+forward model for a single intensity measurement. Figure
+\ref{fig:single-frame} shows the efficiencies in three representative
+geometries. Note that in all cases there are multiple orientations that give the
+same intensity measurement. This means that a single intensity measurement does
+not give us enough information to reconstruct the orientation of a single
+molecule. Next, we will consider combining several intensity measurements to
+create complete polarized illumination microscopes.
 
@@
 Figures \ref{fig:symmetric-widefield} and \ref{fig:double-arm} show our results
-for dual-view symmetric K\"ohler designs. We illuminate from one objective and
+for dual-view symmetric designs. We illuminate from one objective and
 detect from the other for four polarization orientations, then we repeat the
@@
 \centering\includegraphics[width=\textwidth]{symmetric-widefield}
-\caption{Dual-view symmetric K\"ohler designs with varying NA and angle between
+\caption{Dual-view symmetric designs with varying NA and angle between
   objectives. a) Schematic of the microscope. We illuminate with the first
@@
 Figure \ref{fig:double-arm} shows our results when we used a dual-view symmetric
-orthogonal K\"ohler design and varied the illumination and detection NA. Our
-results are similar to the single-view case in Figure \ref{fig:single-arm}. We
-find that optimal designs use a small illumination NA and a large detection
-NA. As we discussed in Section \ref{excitation}, low-NA K\"ohler illumination is
+orthogonal design and varied the illumination and detection NA. Our results are
+similar to the single-view case in Figure \ref{fig:single-arm}. We find that
+optimal designs use a small illumination NA and a large detection NA. As we
+discussed in Section \ref{excitation}, low-NA illumination is
 approximately equivalent to light-sheet illumination for the task of estimating
-the orientation of molecules. Therefore, the preference for low-NA K\"ohler
+the orientation of molecules. Therefore, the preference for low-NA
 illumination also implies that dual-view light-sheet illumination geometries are
@@
 \centering\includegraphics[width=\textwidth]{double-arm}
-\caption{Dual-view symmetric orthogonal K\"ohler designs with varying
+\caption{Dual-view symmetric orthogonal designs with varying
   illumination and detection NA. a) Schematic of the microscope. b) Uncertainty
@@
 Table 1 shows a summary of our evaluation metrics for microscopes in each
-class. The single-view K\"ohler microscope with a small illumination NA and high
+class. The single-view microscope with a small illumination NA and high
 detection NA has the lowest solid-angle uncertainty median but the largest max
-and fourfold symmetry. The dual-view oblique symmetric K\"ohler with an
+and fourfold symmetry. The dual-view oblique symmetric with an
 intermediate NA performs reasonably well on all three metrics, while the
@@
     &Degeneracy&&&\\\midrule
-Single-view K\"ohler (NA${}_\textrm{ill}$=0, NA${}_\textrm{det}$=1.1)&4&3.04$\times 10^{0}$&8.20$\times 10^{-4}$&1.48$\times 10^{-4}$ \\ 
-  Dual-view oblique symmetric K\"ohler (NA=0.6, $\beta$=53${}^{\circ}$)&2&1.35$\times 10^{-1}$&4.49$\times 10^{-3}$&9.17$\times 10^{-4}$\\
+Single-view (NA${}_\textrm{ill}$=0, NA${}_\textrm{det}$=1.1)&4&3.04$\times 10^{0}$&8.20$\times 10^{-4}$&1.48$\times 10^{-4}$ \\ 
+  Dual-view oblique symmetric (NA=0.6, $\beta$=53${}^{\circ}$)&2&1.35$\times 10^{-1}$&4.49$\times 10^{-3}$&9.17$\times 10^{-4}$\\
 Dual-view orthogonal symmetric light-sheet (NA=0.8)&2&6.23$\times 10^{-3}$&1.79$\times 10^{-3}$&1.65$\times 10^{-4}$\\  
@@
   transverse plane.
-\item High-NA dual-view K\"ohler microscopes are optimal when the two views are
+\item High-NA dual-view microscopes are optimal when the two views are
   oblique. Oblique views are complementary because their high uncertainty

### 2017-09-05 paper/paper/paper.tex
Round of small edits from all authors.

@@
 %% Comment before submission
-\bibliography{paper}
-\bibliographystyle{osajnl}
+% \bibliography{paper}
+% \bibliographystyle{osajnl}
 
@@
 % \end{thebibliography}
+
+\begin{thebibliography}{10}
+\newcommand{\enquote}[1]{``#1''}
+
+\bibitem{peterman2001}
+E.~J. Peterman, H.~Sosa, L.~S. Goldstein, and W.~E. Moerner, \enquote{Polarized
+  fluorescence microscopy of individual and many kinesin motors bound to
+  axonemal microtubules,} Biophys. J. \textbf{81}, 2851--2863 (2001).
+
+\bibitem{forkey2003}
+J.~N. Forkey, M.~E. Quinlan, M.~A. Shaw, J.~E.~T. Corrie, and Y.~E. Goldman,
+  \enquote{Three-dimensional structural dynamics of myosin {V} by
+  single-molecule fluorescence polarization,} \nat \textbf{422}, 399--404
+  (2003).
+
+\bibitem{backer2016}
+A.~S. Backer, M.~Lee, and W.~E. Moerner, \enquote{Enhanced {DNA} imaging using
+  super-resolution microscopy and simultaneous single-molecule orientation
+  measurements,} in \enquote{Conference on Lasers and Electro-Optics,}
+  (Optical Society of America, 2016), p. JTh4B.4.
+
+\bibitem{mehta2016}
+S.~B. Mehta, M.~McQuilken, P.~J. La~Rivi\`ere, P.~Occhipinti, A.~Verma,
+  R.~Oldenbourg, A.~S. Gladfelter, and T.~Tani, \enquote{Dissection of
+  molecular assembly dynamics by tracking orientation and position of single
+  molecules in live cells,} Proc. Natl. Acad. Sci. U.S.A. \textbf{113},
+  E6352--E6361 (2016).
+
+\bibitem{demay2011}
+B.~DeMay, N.~Noda, A.~Gladfelter, and R.~Oldenbourg, \enquote{Rapid and
+  quantitative imaging of excitation polarized fluorescence reveals ordered
+  septin dynamics in live yeast,} Biophys. J. \textbf{101}, 985--994 (2011).
+
+\bibitem{mcquilken2017}
+M.~McQuilken, M.~S. Jentzsch, A.~Verma, S.~B. Mehta, R.~Oldenbourg, and A.~S.
+  Gladfelter, \enquote{Analysis of septin reorganization at cytokinesis using
+  polarized fluorescence microscopy,} Front. Cell Dev. Biol. \textbf{5}, 42
+  (2017).
+
+\bibitem{anantharam2010}
+A.~Anantharam, B.~Onoa, R.~H. Edwards, R.~W. Holz, and D.~Axelrod,
+  \enquote{Localized topological changes of the plasma membrane upon exocytosis
+  visualized by polarized {TIRFM},} J. Cell Biol. \textbf{188}, 415--428
+  (2010).
+
+\bibitem{backlund2014}
+M.~P. Backlund, M.~D. Lew, A.~S. Backer, S.~J. Sahl, and W.~E. Moerner,
+  \enquote{The role of molecular dipole orientation in single-molecule
+  fluorescence microscopy and implications for super-resolution imaging,}
+  ChemPhysChem \textbf{15}, 587--599 (2014).
+
+\bibitem{lieb2004}
+M.~A. Lieb, J.~M. Zavislan, and L.~Novotny, \enquote{Single-molecule
+  orientations determined by direct emission pattern imaging,} J. Opt. Soc. Am.
+  B \textbf{21}, 1210--1215 (2004).
+
+\bibitem{backer2014}
+A.~S. Backer and W.~E. Moerner, \enquote{Extending single-molecule microscopy
+  using optical {F}ourier processing,} J. Phys. Chem. B \textbf{118},
+  8313--8329 (2014).
+
+\bibitem{agrawal2012}
+A.~Agrawal, S.~Quirin, G.~Grover, and R.~Piestun, \enquote{Limits of 3{D}
+  dipole localization and orientation estimation for single-molecule imaging:
+  towards {G}reen's tensor engineering,} Opt. Express \textbf{20}, 26667--26680
+  (2012).
+
+\bibitem{toprak2006}
+E.~Toprak, J.~Enderlein, S.~Syed, S.~A. McKinney, R.~G. Petschek, T.~Ha, Y.~E.
+  Goldman, and P.~R. Selvin, \enquote{Defocused orientation and position
+  imaging ({DOPI}) of myosin {V},} Proc. Natl. Acad. Sci. U.S.A. \textbf{103}, 6495--6499 (2006).
+
+\bibitem{debarre2004}
+A.~D{\'e}barre, R.~Jaffiol, C.~Julien, D.~Nutarelli, A.~Richard,
+  P.~Tch{\'e}nio, F.~Chaput, and J.-P. Boilot, \enquote{Quantitative
+  determination of the 3{D} dipole orientation of single molecules,} Eur. Phys.
+  J. D. \textbf{28}, 67--77 (2004).
+
+\bibitem{shribak2003}
+M.~Shribak and R.~Oldenbourg, \enquote{Techniques for fast and sensitive
+  measurements of two-dimensional birefringence distributions,} Appl. Opt.
+  \textbf{42}, 3009--3017 (2003).
+
+\bibitem{fourkas2001}
+J.~T. Fourkas, \enquote{Rapid determination of the three-dimensional
+  orientation of single molecules,} Opt. Lett. \textbf{26}, 211--213 (2001).
+
+\bibitem{tyo2006}
+J.~S. Tyo, D.~L. Goldstein, D.~B. Chenault, and J.~A. Shaw, \enquote{Review of
+  passive imaging polarimetry for remote sensing applications,} Appl. Opt.
+  \textbf{45}, 5453--5469 (2006).
+
+\bibitem{lu2008}
+C.-Y. Lu and D.~A.~V. Bout, \enquote{Analysis of orientational dynamics of
+  single fluorophore trajectories from three-angle polarization experiments,}
+  J. Chem. Phys. \textbf{128}, 244501 (2008).
+
+\bibitem{wu2013}
+Y.~Wu, P.~Wawrzusin, J.~Senseney, R.~S. Fischer, R.~Christensen, A.~Santella,
+  A.~G. York, P.~W. Winter, C.~M. Waterman, Z.~Bao, D.~A. Colon-Ramos,
+  M.~McAuliffe, and H.~Shroff, \enquote{Spatially isotropic four-dimensional
+  imaging with dual-view plane illumination microscopy,} Nat. Biotechnol.
+  \textbf{31}, 1032--1038 (2013).
+
+\bibitem{keller2015}
+R.~K. Chhetri, F.~Amat, Y.~Wan, B.~Hockendorf, W.~C. Lemon, and P.~J. Keller,
+  \enquote{Whole-animal functional and developmental imaging with isotropic
+  spatial resolution,} Nat. Methods \textbf{12}, 1171--1178 (2015).
+
+\bibitem{wu2016}
+Y.~Wu, P.~Chandris, P.~W. Winter, E.~Y. Kim, V.~Jaumouill{\'e}, A.~Kumar,
+  M.~Guo, J.~M. Leung, C.~Smith, I.~Rey-Suarez, H.~Liu, C.~M. Waterman, K.~S.
+  Ramamurthi, P.~J. La~Rivi\`ere, and H.~Shroff, \enquote{Simultaneous
+  multiview capture and fusion improves spatial resolution in wide-field and
+  light-sheet microscopy,} Optica \textbf{3}, 897--910 (2016).
+
+\bibitem{wu2017}
+Y.~Wu, A.~Kumar, C.~Smith, E.~Ardiel, P.~Chandris, R.~Christensen,
+  I.~Rey-Suarez, M.~Guo, H.~Vishwasrao, J.~Chen, J.~Tang, A.~Upadhyaya,
+  P.~La~Rivi\`ere, and H.~Shroff, \enquote{Reflective imaging improves
+  resolution, speed, and collection efficiency in light sheet microscopy,}
+  bioRxiv  (2017).
+
+\bibitem{kay1993}
+S.~M. Kay, \emph{Fundamentals of statistical signal processing} (Prentice Hall
+  PTR, 1993).
+
+\bibitem{anderson1958}
+T.~W. Anderson, \emph{An introduction to multivariate statistical analysis}
+  (Wiley New York, 1958).
+
+\bibitem{coe2009}
+D.~Coe, \enquote{Fisher matrices and confidence ellipses: a quick-start guide
+  and software,} arXiv preprint arXiv:0906.4123  (2009).
+
+\bibitem{walt2011}
+S.~van~der Walt, S.~C. Colbert, and G.~Varoquaux, \enquote{The {N}um{P}y array:
+  a structure for efficient numerical computation,} CoRR \textbf{abs/1102.1523}
+  (2011).
+
+\bibitem{meurer2017}
+A.~Meurer, C.~P. Smith, M.~Paprocki, O.~\v{C}ert\'{i}k, S.~B. Kirpichev,
+  M.~Rocklin, A.~Kumar, S.~Ivanov, J.~K. Moore, S.~Singh, T.~Rathnayake,
+  S.~Vig, B.~E. Granger, R.~P. Muller, F.~Bonazzi, H.~Gupta, S.~Vats,
+  F.~Johansson, F.~Pedregosa, M.~J. Curry, A.~R. Terrel, v.~Rou\v{c}ka,
+  A.~Saboo, I.~Fernando, S.~Kulal, R.~Cimrman, and A.~Scopatz,
+  \enquote{Sym{P}y: symbolic computing in {P}ython,} PeerJ Computer Science
+  \textbf{3}, e103 (2017).
+
+\bibitem{hunter2007}
+J.~D. Hunter, \enquote{Matplotlib: A {2D} graphics environment,} Computing In
+  Science \& Engineering \textbf{9}, 90--95 (2007).
+
+\bibitem{vispy}
+L.~Campagnola, A.~Klein, E.~Larson, A.~Griffiths, C.~Rossant, N.~P. Rougier,
+  I.~Zaid, M.~S. Suraj, A.~Taylor, sylm21, M.~F. Kaptan, A.~J. Champandard,
+  karlcz, nippoo, Siddharth, X.~Gao, C.~GESTES, bdurin, J.~D. Reaver,
+  R.~McMullen, FiachAntaw, T.~Mansencal, M.~Dartiailh, StephenWest, J.~Kieffer,
+  G.~Baty, R.~Peters, macduff111, E.~S. de~Andrade, and D.~Rufat,
+  \enquote{vispy: Version 0.4.0,}  (2015).
+
+\bibitem{bowman2008}
+J.~C. Bowman and A.~Hammerlindl, \enquote{Asymptote: A vector graphics
+  language,} TUGBOAT \textbf{29}, 288--294 (2008).
+
+\bibitem{nov2006}
+L.~Novotny and B.~Hecht, \emph{Principles of Nano-Optics} (Cambridge University
+  Press, 2006).
+
+\bibitem{levoy2006}
+M.~Levoy, R.~Ng, A.~Adams, M.~Footer, and M.~Horowitz, \enquote{Light field
+  microscopy,} in \enquote{ACM SIGGRAPH 2006 Papers,}  (ACM, New York, NY, USA,
+  2006), SIGGRAPH '06, pp. 924--934.
+
+\end{thebibliography}
 
@@
 absorption and emission patterns of single molecules to estimate their
-orientation. By illuminating and detecting from the entire sample, not just the
-focal volume, these techniques can estimate the orientation of many molecules in
+orientation. By illuminating and detecting from the focal plane, not just the
+focal point, these techniques can estimate the orientation of many molecules in
 parallel without scanning. These techniques are easy to implement---changing the
@@
 molecules in all orientations. We also find that single-view microscopes have
-symmetries that can be broken by adding an second view. This reduction of
+symmetries that can be broken by adding a second view. This reduction of
 degeneracy allows multiview microscopes to uniquely determine the orientation of
@@
 Fourkas' expressions were incorrectly normalized (the limit of
-$\eta_{\text{det}}$ as $\alpha\rightarrow \pi/2$ should be $\frac{1}{2}$), and
+$\eta_{\text{det}}$ as $\alpha\rightarrow \frac{\pi}{2}$ should be $\frac{1}{2}$), and
 the extra factor of $\frac{3}{2}$ corrects the error. This correction does not
@@
 $\alpha=\arccos\left(\frac{1}{\sqrt{3}}\right) \approx 54.7^{\circ}$. When
-$\alpha = 90$ the detection efficiency is $\frac{1}{2}$ for all fluorophore
-orientations and when $\alpha = 0$ the detection efficiency is zero for all
-fluorophore orientations, so we expect there to be an intermediate value of
-$\alpha$ where the detection efficiency takes its maximum range.
+$\alpha = \frac{\pi}{2}$ the detection efficiency is $\frac{1}{2}$ for all
+fluorophore orientations and when $\alpha = 0$ the detection efficiency is zero
+for all fluorophore orientations, so we expect there to be an intermediate value
+of $\alpha$ where the detection efficiency takes its maximum range. This means
+that the maximum modulation of detected intensities will occur when we choose a
+detection objective with $\alpha \approx 54.7^{\circ}$.
 
@@
 illumination and detection objectives had $\mh{z}$ aligned optical axes, i.e.,
-we used epi- or transmitted light illumination. To extend the forward model to
-oblique optical axes which allows for non-coincident illumination and detection
-objectives, we express the dipole orientation in a rotated coordinate frame
-using the following expressions
+we assumed traditional epi- or transmitted light illumination. To extend the
+forward model to oblique optical axes which allows for non-coincident
+illumination and detection objectives, we express the dipole orientation in a
+rotated coordinate frame using the following expressions
 \begin{align}
@@
 efficiencies, and other constant factors. Equation \ref{eq:forward_frame} is the
-forward model for a single intensity measurement. Figure
-\ref{fig:single-frame} shows the efficiencies in three representative
-geometries. Note that in all cases there are multiple orientations that give the
-same intensity measurement. This means that a single intensity measurement does
-not give us enough information to reconstruct the orientation of a single
-molecule. Next, we will consider combining several intensity measurements to
-create complete polarized illumination microscopes.
+forward model for a single intensity measurement. Figure \ref{fig:single-frame}
+shows the efficiencies in three representative geometries. Note that in all
+cases there are multiple orientations that give the same intensity
+measurement. This means that a single intensity measurement does not give us
+enough information to reconstruct the orientation of a single molecule. Next, we
+will consider combining several intensity measurements to create polarized
+illumination microscopes that can reconstruct the orientation of single
+molecules.
 
@@
 equally spaced points on the unit sphere. A desirable microscope design will
-have a small solid-angle uncertainty that is uniform for all molecule
+have a small solid-angle uncertainty that is uniform for all molecular
 orientations. The most straightforward way to find the location and scale of the
@@
 and the uncertainty ellipse give no indication of this degeneracy. For this
-reason we also use the number of degeneracies as a measure of a micrsocope's
+reason we also use the number of degeneracies as a measure of a microscope's
 ability to estimate the orientation of a dipole from intensity measurements. A
@@
 orientation of a molecule---a twofold degenerate microscope can find the
-orientation of a molecule down to the hemisphere while a fourfold symmetric
+orientation of a molecule down to the hemisphere while a fourfold degenerate
 microscope can only find the orientation down to the quadrant.
@@
 \pi, \Phi+\pi)$ in equations \ref{eq:scalarabs}, \ref{eq:scalardet}, and
-\ref{eq:forward_frame}). In addition to the usual inversion degeneracy,
-single-view microscopes suffer from a degeneracy when a dipole is reflected
-through the transverse plane. Evidence of this degeneracy is found in Figures
-\ref{fig:single-arm}b) and \ref{fig:single-arm}c) which show the ambiguity of
-whether the molecule is oriented above or below the transverse plane. Near the
-transverse plane this ambiguity is evident in the high solid-angle uncertainty,
-but far from the transverse plane evidence of the ambiguity is lost. Adding a
-second view and taking eight intensity measurements instead of four removes this
-degeneracy (see equations \ref{eq:thetap} and \ref{eq:phip}), so the dual-view
-microscopes we consider in this paper are only twofold degenerate. This means
-that while dual-view microscopes can find the orientation of a molecule up to
-the hemisphere, while single-view microscopes can only find the orientation of a
-molecule up to the quadrant.
+\ref{eq:forward_frame}). The first twofold degeneracy is due to the usual
+inversion degeneracy---the dipole is an axis rather than a direction. A second
+twofold degeneracy is present in single-view designs---the microscope cannot
+distinguish between dipoles reflected through the transverse plane. Evidence of
+the second degeneracy is found in Figures \ref{fig:single-arm}b) and
+\ref{fig:single-arm}c) which show the ambiguity of whether the molecule is
+oriented above or below the transverse plane. Near the transverse plane this
+ambiguity is evident in the high solid-angle uncertainty, but far from the
+transverse plane evidence of the ambiguity is lost. Adding a second view and
+taking eight intensity measurements instead of four removes this degeneracy (see
+equations \ref{eq:thetap} and \ref{eq:phip}), so the dual-view microscopes we
+consider in this paper are only twofold degenerate. This means that while
+dual-view microscopes can find the orientation of a molecule up to the
+hemisphere, single-view microscopes can only find the orientation of a molecule
+up to the quadrant.
 
@@
 detection NA has the lowest solid-angle uncertainty median but the largest max
-and fourfold symmetry. The dual-view oblique symmetric with an
+and fourfold degeneracy. The dual-view oblique symmetric microscope with an
 intermediate NA performs reasonably well on all three metrics, while the
@@
 \section*{Funding}
-Support for this work was provided by the Intramural Research Programs of the
-National Institute of Biomedical Imaging and Bioengineering.
+TC was supported by a University of Chicago Biological Sciences Division
+Graduate Fellowship. SM and RO were supported by NIH grant R01GM114274. HS was
+supported by the Intramural Research Programs of the National Institute of
+Biomedical Imaging and Bioengineering.
 

### 2017-09-05 paper/paper/paper.tex
Clarifying Rudolf's change.

@@
 when $B$ takes its maximum value, i.e. when
-$\alpha=\arccos\left(\frac{1}{\sqrt{3}}\right) \approx 54.7^{\circ}$. When
-$\alpha = \frac{\pi}{2}$ the detection efficiency is $\frac{1}{2}$ for all
-fluorophore orientations and when $\alpha = 0$ the detection efficiency is zero
-for all fluorophore orientations, so we expect there to be an intermediate value
-of $\alpha$ where the detection efficiency takes its maximum range. This means
+$\alpha=\arccos\left(\frac{1}{\sqrt{3}}\right) \approx 54.7^{\circ}$. This means
 that the maximum modulation of detected intensities will occur when we choose a

### 2017-09-18 paper/paper/paper.tex
Patrick's funding sources.
Add Hari's funding.
Rudolf and Talon's paper edits re: our results compared to Fourkas

@@
 requirement for multiple measurements, polarized illumination or detection
-methods are good choices for live cell imaging because of their combination of
+methods are good choices for live-cell imaging because of their combination of
 speed and ease of implementation.
@@
 universal compensator/polarizer. Other polarization methods
-(divison-of-amplitude, division-of-focal-plane) are more expensive and require
+(division-of-amplitude, division-of-focal-plane) are more expensive and require
 additional alignment or interpolation \cite{tyo2006}. The simplicity of DoT
@@
 isotropic orientation uncertainty has been underappreciated. To our knowledge,
-the only authors who have considered anisotropic orientation uncertainty have
-only analyzed a subset of molecular orientations \cite{agrawal2012}. Our view is
-that an ideal technique for determining the orientation of single molecules can
-reconstruct the orientation with a small and nearly uniform uncertainty for all
-molecule orientations.
+the only authors who have considered anisotropic orientation uncertainty in 3D
+have only analyzed a subset of molecular orientations \cite{agrawal2012}. Our
+view is that an ideal technique for determining the orientation of single
+molecules can reconstruct the orientation with a small and nearly uniform
+uncertainty for all molecular orientations.
 
@@
 focal plane. In this geometry each point in the sample is illuminated by a range
-of angles that can be changed by adjusting the radius of the stop in the
+of angles that can be changed by adjusting the radius of the diaphragm in the
 aperture plane. We denote the set of illuminated angles by $\Omega$ (see Figure
@@
 \subsection{Detection Efficiency}\label{detection}
-Fourkas calculated the detection efficiency of a single molecule when an ideal
-objective with its optical axis aligned with the $\mh{z}$ axis is focused on the
-molecule and a polarizer is placed in the aperture plane\cite{fourkas2001}. We
-use his expressions to calculate the detection efficiency without a polarizer as
+In this section we will calculate the \emph{detection efficiency}---the fraction
+of the power emitted by a single molecule that we detect. Fourkas calculated the
+detection efficiency when an ideal objective with its optical axis aligned with
+the $\mh{z}$ axis is focused on a single molecule and a polarizer is placed in
+the aperture plane \cite{fourkas2001}. He found that
+\begin{align}
+  \eta_{\text{det}}^{\text{pol}} = A + B\sin^2\Theta + C\sin^{2}{\Theta} \cos{[2 (\Phi - \phi_{\text{det}}})], \label{eq:det_pol}
+\end{align}
+where $\phi_{\text{det}}$ is the angle between the transmission axis of the
+detection polarizer and the positive $\mh{x}$ axis.
+
+To calculate the detection efficiency without a polarizer, we use Equation
+\ref{eq:det_pol} and take the sum of polarized detection efficiencies with
+orthogonal polarizer orientations to find
 \begin{align}
@@
 \end{align}
-The detection efficiency only depends on $\Theta$, not $\Phi$, because there is
-no detection polarizer. We have assumed that the absorption dipole moment is
-equal to the emission dipole moment.
+In this case the detection efficiency only depends on $\Theta$, not $\Phi$,
+because there is no detection polarizer. We have assumed that the absorption
+dipole moment is equal to the emission dipole moment.
 
-Note that $A$ and $B$ in equation \ref{eq:coefficients} are a factor of
+Note that $A$, $B$, and $C$ in equation \ref{eq:coefficients} are a factor of
 $\frac{3}{2}$ larger than the expressions given by Fourkas. We found that
@@
 \subsection{Intensity Measurements}\label{forward}
-The detected intensity is proportional to the product of the sample exposure
-$I_{\text{exp}}$, the absorption efficiency, and the detection efficiency
+The detected intensity is proportional to the product of the sample irradiance
+$I_{\text{in}}$, the absorption efficiency, and the detection efficiency
 \begin{align}
-  I \propto I_{\text{exp}}\eta_{\text{abs}}\eta_{\text{det}}\label{eq:forward_frame},
+  I \propto I_{\text{in}}\eta_{\text{abs}}\eta_{\text{det}}\label{eq:forward_frame},
 \end{align}
@@
 
-CRLBs and the associated area of uncertainty are parametrization dependent,
+CRLBs and the associated area of uncertainty are parameterization dependent,
 i.e., if we choose a different coordinate system these metrics will change. We
-would like to compare microscope designs without choosing a parametrization, so
+would like to compare microscope designs without choosing a parameterization, so
 instead we use
@@
 as our evaluation metric. We calculate the determinant of the inverse Fisher
-information matrix---a parametrization-independent value called the
+information matrix---a parameterization-independent value called the
 \emph{generalized variance} \cite{anderson1958}---take the square root, then
@@
 the \emph{solid-angle uncertainty} because it has units of steradians and is a
-parametrization independent measure of the orientation uncertainty.
+parameterization independent measure of the orientation uncertainty.
 
@@
 To compare microscopes with different numbers of intensity measurements fairly,
-we kept the total exposure to the sample constant by choosing
-$I_{\text{exp}} = \frac{4000}{N}$. Note that $I_{\text{exp}}$ is a measure of
-exposure to the sample, not the detected intensity or the total intensity
+we kept the irradiance incident on the sample constant by choosing
+$I_{\text{in}} = \frac{4000}{N}$. Note that $I_{\text{in}}$ is a measure of
+irradiance to the sample, not the detected intensity or the total intensity
 emitted by the molecule.
@@
 single objective is used for illumination and detection. We swept through the
-illumination and detection NA while keeping the sample exposure constant, and we
-found that the lowest median and MAD of the solid-angle uncertainty occurs with
-a small illumination NA and a large detection NA where
+illumination and detection NA while keeping the sample irradiance constant, and
+we found that the lowest median and MAD of the solid-angle uncertainty occurs
+with a small illumination NA and a large detection NA where
 $\alpha \approx 65^{\circ}$ (see Figure \ref{fig:single-arm}e) at
@@
 swept through the NA asymmetry and the sample-exposure asymmetry while keeping
-the total sample exposure constant, and found that symmetric designs are at a
+the total sample irradiance constant, and found that symmetric designs are at a
 local minimum of the median and MAD of the solid-angle uncertainty. We also
@@
 Our results are limited in several ways. First, we only consider single
-molecules in homogeneous environments. We ignore light reflected from the cover
-slip and any other inhomogeneities in the sample. If we considered light
-reflected from the cover slip we expect that the results would put a lower
-weighting on large detection NAs because we would collect more light. Second, we
-only considered ideal, aplanatic, polarization-preserving objectives. In
-practice, these conditions are not always satisfied and the power from high-NA
-rays is apodized by a factor of $\frac{n}{\cos\Theta}$
-\cite{nov2006}. Therefore, our results are most accurate for low-NA rays and get
-progressively worse for high-NA rays, although our major conclusions would not
-change if we included this apodization factor. The main advantage of ignoring
-apodization is the tractability of the problem---we have provided approximate
-closed-form solutions for the illumination and detection efficiencies that have
-allowed us to draw useful design conclusions. Third, we have used a Poisson
-noise model which ignores noise introduced by the detector. Detector noise can
-be a large fraction of the total noise at low intensities, so including detector
-noise would increase the informational value of high intensity
+molecules in homogeneous environments whose absorption and emission dipoles are
+parallel. We ignore light reflected from the cover slip and any other
+inhomogeneities in the sample. If we considered light reflected from the cover
+slip we expect that the results would put a lower weighting on large detection
+NAs because we would collect more light. Second, we only considered ideal,
+aplanatic, polarization-preserving objectives. In practice, these conditions are
+not always satisfied. For example, the power from high-NA rays is usually
+apodized by a factor of $\frac{n}{\cos\Theta}$ \cite{nov2006}. Therefore, our
+results are most accurate for paraxial rays and get progressively worse for
+non-paraxial rays, although our major conclusions would not change if we
+included this apodization factor. The main advantage of ignoring apodization is
+the tractability of the problem---we have provided approximate closed-form
+solutions for the illumination and detection efficiencies that have allowed us
+to draw useful design conclusions. Third, we have used a Poisson noise model
+which ignores noise introduced by the detector. Detector noise can be a large
+fraction of the total noise at low intensities, so including detector noise
+would increase the informational value of high intensity
 measurements. Therefore, we expect that including the effects of detector noise
@@
 
-To compare the microscopes in this paper we used a fixed sample exposure. We
-think that this is the fairest way to compare microscopes with different numbers
-of intensity measurements. However, in cases where sample exposure and time are
-not an issue it may be more useful to compare microscopes with the same total
-detection exposure per measurement. This comparison would favor microscopes with
-many intensity measurements even more that the results in this paper because
-equal detection exposure would improve photon counting statistics.
+To compare the microscopes in this paper we used a fixed irradiance incident on
+the sample. We think that this is the fairest way to compare microscopes with
+different numbers of intensity measurements. However, in cases where sample
+irradiance and time are not an issue it may be more useful to compare
+microscopes with the same total detection irradiance per measurement. This
+comparison would favor microscopes with many intensity measurements even more
+that the results in this paper because equal detection irradiance would improve
+photon counting statistics.
 
 We have considered single- and dual-view microscopes in a variety of cases, but
-more multiview microscopes can be analyzed with the framework we have
+other multiview microscopes can be analyzed with the framework we have
 developed. Wu et. al. have developed a three-view design with a third objective
@@
 TC was supported by a University of Chicago Biological Sciences Division
-Graduate Fellowship. SM and RO were supported by NIH grant R01GM114274. HS was
-supported by the Intramural Research Programs of the National Institute of
-Biomedical Imaging and Bioengineering.
+Graduate Fellowship. SM and RO were supported by NIH grant R01GM114274. HS and
+PL were supported by Marine Biological Laboratory Whitman Center Fellowships.
+HS was supported by the Intramural Research Programs of the National Institute
+of Biomedical Imaging and Bioengineering. PL was supported by NIH grant
+R01EB017293.
 

### 2017-09-18 paper/paper/paper.tex
"total sample irradiance" -> "sample irradiance"

@@
 swept through the NA asymmetry and the sample-exposure asymmetry while keeping
-the total sample irradiance constant, and found that symmetric designs are at a
-local minimum of the median and MAD of the solid-angle uncertainty. We also
-found that at extreme asymmetries where the exposure from the low-NA arm is much
-larger than the exposure from the high-NA arm (top-left and bottom-right corners
-of Figures \ref{fig:asymmetric-double}d) and e)) the median and MAD are
-comparable to the symmetric light-sheet microscope. This means that if a very
-high-NA objective is available and a low-NA objective can provide light sheet
+the sample irradiance constant, and found that symmetric designs are at a local
+minimum of the median and MAD of the solid-angle uncertainty. We also found that
+at extreme asymmetries where the exposure from the low-NA arm is much larger
+than the exposure from the high-NA arm (top-left and bottom-right corners of
+Figures \ref{fig:asymmetric-double}d) and e)) the median and MAD are comparable
+to the symmetric light-sheet microscope. This means that if a very high-NA
+objective is available and a low-NA objective can provide light sheet
 illumination, this microscope will perform comparably with a dual-view

### 2017-09-19 notes/2017-09-19-ensemble-efficiencies/report/report.tex
Orientation forward model w/ notes.

\documentclass[11pt]{article}

%%%%%%%%%%%%
% Packages %
%%%%%%%%%%%%
\hyphenpenalty=10000
\usepackage{tocloft}
\renewcommand\cftsecleader{\cftdotfill{\cftdotsep}}
\def\undertilde#1{\mathord{\vtop{\ialign{##\crcr
$\hfil\displaystyle{#1}\hfil$\crcr\noalign{\kern1.5pt\nointerlineskip}
$\hfil\tilde{}\hfil$\crcr\noalign{\kern1.5pt}}}}}
\usepackage{cleveref}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{epstopdf}
\usepackage{braket}
\usepackage{upgreek}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{amssymb,latexsym,amsmath,gensymb}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{float}
\usepackage{enumitem}
\usepackage{pdflscape}
\usepackage{url}
\usepackage{tikz, calc}
\usetikzlibrary{shapes.geometric, arrows, calc}
\tikzstyle{norm} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{arrow} = [thick, ->, >=stealth]

\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
\providecommand{\mb}[1]{\mathbf{#1}}
\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
\providecommand{\bs}[1]{\boldsymbol{#1}} 
\providecommand{\intinf}{\int_{-\infty}^{\infty}}
\providecommand{\fig}[4]{
  % filename, width, caption, label
\begin{figure}[h]
 \captionsetup{width=1.0\linewidth}
 \centering
 \includegraphics[width = #2\textwidth]{#1}
 \caption{#3}
 \label{fig:#4}
\end{figure}
}

\newcommand{\tensor}[1]{\overset{\text{\tiny$\leftrightarrow$}}{\mb{#1}}}
\newcommand{\tunderbrace}[2]{\underbrace{#1}_{\textstyle#2}}
\providecommand{\figs}[7]{
  % filename1, filename2, caption1, caption2, label1, label2, shift
\begin{figure}[H]
\centering
\begin{minipage}[b]{.45\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#1}
  \captionsetup{justification=justified, singlelinecheck=true}
  \caption{#3}
  \label{fig:#5}
\end{minipage}
\hspace{2em}
\begin{minipage}[b]{.45\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#2}
  \vspace{#7em}
  \captionsetup{justification=justified}
  \caption{#4}
  \label{fig:#6}
\end{minipage}
\end{figure}
}
\makeatletter

\providecommand{\code}[1]{
\begin{center}
\lstinputlisting{#1}
\end{center}
}

\newcommand{\crefrangeconjunction}{--}
%%%%%%%%%%%
% Spacing %
%%%%%%%%%%%
% Margins
\usepackage[
top    = 1.5cm,
bottom = 1.5cm,
left   = 1.5cm,
right  = 1.5cm]{geometry}

% Indents, paragraph space
%\usepackage{parskip}
\setlength{\parskip}{1.5ex}

% Section spacing
\usepackage{titlesec}
\titlespacing*{\title}
{0pt}{0ex}{0ex}
\titlespacing*{\section}
{0pt}{0ex}{0ex}
\titlespacing*{\subsection}
{0pt}{0ex}{0ex}
\titlespacing*{\subsubsection}
{0pt}{0ex}{0ex}

% Line spacing
\linespread{1.1}

%%%%%%%%%%%%
% Document %
%%%%%%%%%%%%
\begin{document}
\title{\vspace{-2.5em} Excitation and Detection Efficiency of Ensembles of
  Molecules Under Polarized Illumination \vspace{-1em}} \author{Talon
  Chandler}% and Patrick La Rivi\`ere}
\date{\vspace{-1em}\today\vspace{-1em}}
\maketitle
\section{Introduction}
In these notes I will calculate the excitation and detection efficiencies of
ensembles of molecules as a function of polarizer orientation and microscope
geometry. This work extends the results in the 2017-06-14 notes from single
molecules to ensembles of molecules.

We start by extending the excitation and detection efficiencies to arbitrary
distributions of molecules. Next, we review possible orientation distribution
functions that we can expect real ensembles to follow. Finally, we see how
orientation distributions affect the intensities measured by microscopes.

\section{Excitation and Detection Efficiency for Arbitrary Distributions of Molecules}
Previously we showed that the excitation efficiency of a single molecule under
incoherent uniform angular illumination is given by
\begin{align}
  \eta_{\text{abs}}^{\text{single}} &= D\{A + B\sin^{2}{\Theta} + C\sin^{2}{\Theta} \cos{[2 (\Phi - \phi_{\text{exc}}})]\}\label{eq:scalarabs}
\end{align}
where
\begin{subequations}
\begin{align}
  A &= \frac{1}{4} - \frac{3}{8} \cos{\alpha } + \frac{1}{8} \cos^{3}{\alpha }\\
  B &= \frac{3}{16} \cos{\alpha } - \frac{3}{16} \cos^{3}{\alpha }\\
  C &= \frac{7}{32} - \frac{3}{32} \cos{\alpha } - \frac{3}{32} \cos^{2}{\alpha } - \frac{1}{32} \cos^{3}{\alpha}\\
  D &= \frac{4}{3(1 - \cos\alpha)}.
\end{align}\label{eq:coefficients}%
\end{subequations}

If an ensemble of independent fluorescent molecules is illuminated then the
absorption efficiency of the ensemble is
\begin{align}
  \eta_{\text{abs}} = \int_{\mathbb{S}^2}d\mh{r} f(\mh{r}) \eta_{\text{abs}}^{\text{single}} \label{eq:ensemble}%
\end{align} 
where $f(\mh{r})$ is the orientation probability distribution function of the
fluorophores. For equation \ref{eq:ensemble} to be true the fluorophores must
absorb light independently---no homo-FRET or coherence effects.

Similarly, the detection efficiency of an ensemble is
\begin{align}
  \eta_{\text{det}} = \int_{\mathbb{S}^2}d\mh{r} f(\mh{r}) \eta_{\text{det}}^{\text{single}} \label{eq:ensemble_det}%
\end{align} 
which is only true if the fluorophores emit light independently. I haven't given
the full expression for $\eta_{\text{det}}^{\text{single}}$ here---see the paper
or equation 2 with $D = 1$.

Note that equations \ref{eq:ensemble} and \ref{eq:ensemble_det} are much more
computationally expensive than the single molecule efficiencies that I've
implemented previously. The double integrals do not have closed form solutions,
so I am using a numerical integration scheme---Gausse-Legendre adaptive
quadrature methods implemented in QUADPACK and called from \texttt{scipy}. These
methods allow us to easily trade speed for accuracy.

\section{Axial Distributions On The Sphere}
In this section we will review several orientation probability distribution functions.

The \textbf{Von Mises-Fisher distribution} is given by
\begin{align*}
  f(\mh{r}; \bs{\hat{\mu}}, \kappa) = \frac{\sqrt{\kappa}}{(2\pi{})^{3/2}I_{1/2}(\kappa)}\text{exp}\{\kappa \bs{\hat{\mu}}^T\mh{r}\}
\end{align*}
where $I_{1/2}$ denotes the modified Bessel function of the first kind. The Von
Mises-Fisher distribution is rotationally symmetric about $\bs{\hat{\mu}}$, but
it is not antipodally symmetric. Therefore, the Von Mises-Fisher distribution is
not an appropriate choice for modeling axial data.

\fig{../figures/vonmises.pdf}{1.0}{Von Mises-Fisher distributions with constant $\bs{\hat{\mu}}$ and varying $\kappa$. }{vonmises}

The \textbf{Watson distribution} \cite{sra2007modeling, jupp} is given by
\begin{align*}
  f(\mh{r}; \bs{\hat{\mu}}, \kappa) = \frac{1}{4\pi{}_1F_1\left(\frac{1}{2}, \frac{3}{2}, \kappa\right)}\text{exp}\{\kappa (\bs{\hat{\mu}}^T\mh{r})^2\}
\end{align*}
where ${}_1F_1$ denotes a confluent hypergeometric function. The Watson
distribution is antipodally symmetric ($f(\mh{r}) = f(-\mh{r})$) and
rotationally symmetric about $\bs{\hat{\mu}}$. The parameter $\kappa$ is a
concentration parameter---when $\kappa$ is large and positive the distribution
is concentrated near $\mu$, when $\kappa$ is large and negative the distribution
is concentrated near the great circle orthogonal to $\bs{\hat{\mu}}$, and when
$\kappa$ is small the distribution is uniform. $\bs{\hat{\mu}}$ is a unit vector
which can be specified using two angles, so the Watson distribution is a 3
parameter distribution.

The Watson distribution is (1) antipodally symmetric, (2) rotationally symmetric
about a single axis, and (3) smooth, so we think it is a good distribution to
describe distributions of molecules in biological samples. The Watson
distribution has also been used as a model for paleomagnetic data (orientations
of the magnetic axis in rocks) \cite{Love2007}, MRI diffusion tensor imaging
data \cite{Jespersen2012}, and microphone array directionality data
\cite{alex}. We will use the Watson distribution as a model for fluorophores for
the remaining sections of these notes.

\fig{../figures/watson.pdf}{1.0}{Watson distributions with constant
  $\bs{\hat{\mu}}$ and varying $\kappa$. }{watson}

For completeness, the \textbf{Bingham distribution} \cite{sra2007modeling, jupp}
is a generalized version of the Watson distribution that allows for
distributions that are not rotationally symmetric about a single axis. The
Bingham distribution is given by
\begin{align}
  f(\mh{r}; \bs{\hat{\mu}}, \kappa) = \frac{1}{4\pi{}_1F_1\left(\frac{1}{2}, \frac{3}{2}, \mb{K}\right)}\text{exp}\{\mh{r}^T\mb{K}\mh{r}\}
\end{align}
where ${}_1F_1$ is denotes a confluent hypergeometric function with a matrix
argument. The matrix $\mb{K}$ has three free parameters (two eigenvalues and a
choice of principal axis), so the Bingham distribution is a 5 parameter
distribution.

\section{Forward Model For Watson Distributed Molecules}
The detected intensity is proportional to the number of molecules in the
ensemble $N$, the absorption efficiency $\eta_{\text{abs}}$, and the detection
efficiency $\eta_{\text{det}}$
\begin{align}
  I \propto N\eta_{\text{abs}}\eta_{\text{det}}. 
\end{align}
Our goal is to estimate the parameters of the ensemble distribution
($\Theta$, $\Phi$, $\kappa$) and the number of molecules in the distribution $N$
from the intensity measurements. If we can estimate the number of photons each
fluorophore emits (from a datasheet or a separate experiment), then we can
estimate $N$ directly. If this information is not available, we will estimate a
number that is directly proportional to $N$.

Figure \ref{fig:single-frame} shows representative examples of intensity
measurements collected from Watson-distributed fluorophores. Note how the total
efficiency pattern changes as $\kappa$ changes. For large $\kappa$, we recover
the case of single fluorophores and there is a concentrated set of orientations
with high total efficiency. As $\kappa$ decreases, the set of orientations with
high total efficiency grows until $\kappa = 0$ and the total efficiency is
independent of $\bs{\hat{\mu}}$ (this is because the mean orientation is
meaningless for a uniform distribution). When $\kappa < 0$ the total efficiency
is high in a ring of orientations.

\fig{../figures/single-frame.pdf}{1.0}{Representative examples of ensemble
  intensity measurements. Black dots indicate where the Cartesian unit vectors
  intersect the unit sphere. \newline \newline \textbf{Columns left to right:}
  1) fluorophore distributions with $\bs{\hat{\mu}} = \mb{\hat{z}}$ and varying
  $\kappa$; 2) microscope geometry schematics---all cases use two 0.8 NA
  objectives in a diSPIM geometry 3) the absorption efficiency as a function of
  $\bs{\hat{\mu}}$, see equation \ref{eq:ensemble}; 3) the detection efficiency
  as a function of $\bs{\hat{\mu}}$, see equation \ref{eq:ensemble_det}; (4) the
  total efficiency, the product of the absorption and detection
  efficiencies. \textbf{Note:} The last three columns do not use the same
  scale.}{single-frame}

\section{Computation Time Estimates}
A function evaluation for the intensity collected from a single fluorophore
\textbf{currently} takes $\sim$ 0.1 ms.

To calculate the intensity collected from a distribution of fluorophores we need
to repeat this calculation for each fluorophore orientation. Empirically,
$\sim 500$ orientations bounds the error on the integral below 1\%, so the full
calculation takes $\sim 50$ms.

We're planning to implement a Fisher scoring algorithm that acts on the data
from each voxel to estimate 4 parameters: the orientation distribution
($\Theta$, $\Phi$, $\kappa$) and the number of fluorophores $N$. The algorithm
(1) computes the Fisher information matrix at a starting guess which takes 5
function evaluations (a central point and 4 derivative directions) which takes
$\sim 0.25\ \text{s}$, (2) updates its guess (cheap), then (3) recomputes the Fisher
information matrix until convergence. It's difficult to guess the number of
iterations we will require for convergence, but let's say we can get it down to
10 iterations ($\sim$ 2.5 s).

The data set we took this summer has $\sim 150\times 150\times 800 \approx 2\e{7}$
useful voxels which will take $\sim5\e{7} \text{s} = 1.5\ \text{y}$.

There are still plenty of places we can save time. In approximate order of ease:
\begin{enumerate}
\item Move the function evaluations from Python to C. This shouldn't take much
  more than a day, and forums report a 3-10$\times$ improvement.
\item Optimize the integration method to reduce the number of function
  evaluations. Empirically, when $\kappa$ is small we need fewer that 500
  function evaluations for the integration to have a 1\% error.
\item Bin voxels ($2\times2\times2$ binning gives $8\times$ speedup,
  $3\times3\times3$ gives $27\times$ speedup).
\item Process voxels in parallel on the GPU (10-1000$\times$ speedup)
\end{enumerate}

All of these improvements together will (conservatively) put the reconstruction time at $\sim 1\ \text{day}$, and likely much faster.
\bibliography{report}{}
\bibliographystyle{unsrt}

\end{document}

### 2017-09-19 dipsim/detector.py

@@ -1,5 +1,7 @@
 import numpy as np
 import dipsim.util as util
+import dipsim.fluorophore as flu
+from scipy import integrate, special
 
 class Detector:
     """A Detector is specified by its optical axis, numerical aperture, and
@@ -14,15 +16,29 @@ class Detector:
         self.alpha = np.arcsin(self.na/self.n)
         self.phi_pol = phi_pol
 
-    def calc_collection_efficiency(self, fluorophore):
-        theta = util.theta_prime(fluorophore.mu_em[0], fluorophore.mu_em[1], self.theta_optical_axis)
-        phi = util.phi_prime(fluorophore.mu_abs[0], fluorophore.mu_abs[1], self.theta_optical_axis)        
-        A = (1.0/4.0) - (3.0/8.0)*np.cos(self.alpha) + (1.0/8.0)*(np.cos(self.alpha)**3)
-        B = (3.0/16.0)*np.cos(self.alpha) - (3.0/16.0)*(np.cos(self.alpha)**3)
-        C = (7.0/32.0) - (3.0/32.0)*np.cos(self.alpha) - (3.0/32.0)*(np.cos(self.alpha)**2) - (1.0/32.0)*(np.cos(self.alpha)**3)
-        if self.det_type=='lens':
-            return 2*(A + B*(np.sin(theta)**2))
-        elif self.det_type=='polarized':
-            return A + B*(np.sin(theta)**2) + C*(np.sin(theta)**2)*np.cos(2*(phi - self.phi_pol))
-        elif self.det_type=='4pi':
-            return 1.0
+    def calc_collection_efficiency(self, fluorophore, epsrel=1e-2):
+        if fluorophore.kappa == None:
+            theta = util.theta_prime(fluorophore.mu_em[0], fluorophore.mu_em[1], self.theta_optical_axis)
+            phi = util.phi_prime(fluorophore.mu_abs[0], fluorophore.mu_abs[1], self.theta_optical_axis)        
+            A = (1.0/4.0) - (3.0/8.0)*np.cos(self.alpha) + (1.0/8.0)*(np.cos(self.alpha)**3)
+            B = (3.0/16.0)*np.cos(self.alpha) - (3.0/16.0)*(np.cos(self.alpha)**3)
+            C = (7.0/32.0) - (3.0/32.0)*np.cos(self.alpha) - (3.0/32.0)*(np.cos(self.alpha)**2) - (1.0/32.0)*(np.cos(self.alpha)**3)
+            if self.det_type=='lens':
[…]

### 2017-09-19 dipsim/fluorophore.py

@@ -2,13 +2,17 @@ import numpy as np
 
 class Fluorophore:
     """A single fluorophore is specified by its 3D position, (unit) absorption
-    dipole moment (theta, phi), and (unit) emission dipole moment (theta, phi).
+    dipole moment (theta, phi), and (unit) emission dipole moment (theta, phi),
+    and ensemble distribution (kappa = None means single molecule. kappa != None
+    means Watson distribution).
 
     """
     def __init__(self, position=np.array([0, 0, 0]),
                  mu_abs=np.array([0, 0]),
-                 mu_em=np.array([0, 0])):
+                 mu_em=np.array([0, 0]),
+                 kappa=None):
         self.position = position
         self.mu_abs = mu_abs        
         self.mu_em = mu_em
         self.mu_ind = 0
+        self.kappa = kappa

### 2017-09-19 dipsim/illuminator.py

@@ -1,5 +1,7 @@
 import numpy as np
 import dipsim.util as util
+import dipsim.fluorophore as flu
+from scipy import integrate, special
 
 class Illuminator:
     """An Illuminator is specified by its illumination type (wide, sheet),
@@ -15,18 +17,29 @@ class Illuminator:
         self.alpha = np.arcsin(self.na/self.n)
         self.phi_pol = phi_pol
 
-    def calc_excitation_efficiency(self, fluorophore):
-        theta = util.theta_prime(fluorophore.mu_abs[0], fluorophore.mu_abs[1], self.theta_optical_axis)
-        phi = util.phi_prime(fluorophore.mu_abs[0], fluorophore.mu_abs[1], self.theta_optical_axis)
-        A = (1.0/4.0) - (3.0/8.0)*np.cos(self.alpha) + (1.0/8.0)*(np.cos(self.alpha)**3)
-        B = (3.0/16.0)*np.cos(self.alpha) - (3.0/16.0)*(np.cos(self.alpha)**3)
-        C = (7.0/32.0) - (3.0/32.0)*np.cos(self.alpha) - (3.0/32.0)*(np.cos(self.alpha)**2) - (1.0/32.0)*(np.cos(self.alpha)**3)        
-        D = 4.0/(3.0*(1.0 - np.cos(self.alpha)))
-        if self.illum_type == 'wide':
-            return D*(A + B*(np.sin(theta)**2) + C*(np.sin(theta)**2)*np.cos(2*(phi - self.phi_pol)))
-        elif self.illum_type == 'sheet':
-            return (np.sin(theta)*np.cos(phi - self.phi_pol))**2
-        elif self.illum_type == 'unpolarized':
-            return 2*D*(A + B*(np.sin(theta)**2))
+    def calc_excitation_efficiency(self, fluorophore, epsrel=1e-2):
+        if fluorophore.kappa == None:
+            theta = util.theta_prime(fluorophore.mu_abs[0], fluorophore.mu_abs[1], self.theta_optical_axis)
+            phi = util.phi_prime(fluorophore.mu_abs[0], fluorophore.mu_abs[1], self.theta_optical_axis)
+            A = (1.0/4.0) - (3.0/8.0)*np.cos(self.alpha) + (1.0/8.0)*(np.cos(self.alpha)**3)
+            B = (3.0/16.0)*np.cos(self.alpha) - (3.0/16.0)*(np.cos(self.alpha)**3)
[…]

### 2017-09-19 dipsim/microscope.py

@@ -21,40 +21,41 @@ class Microscope:
     def calc_intensity(self, direction):
         return self.max_photons*self.calc_sensitivity(direction)
     
-    def calc_sensitivity(self, direction):
-        flu = fluorophore.Fluorophore(mu_abs=direction, mu_em=direction)
-        excite = self.illuminator.calc_excitation_efficiency(flu)
-        collect = self.detector.calc_collection_efficiency(flu)        
+    def calc_sensitivity(self, direction, kappa=None, epsrel=1e-2):
+        flu = fluorophore.Fluorophore(mu_abs=direction, mu_em=direction, kappa=kappa)
+        excite = self.illuminator.calc_excitation_efficiency(flu, epsrel=epsrel)
+        collect = self.detector.calc_collection_efficiency(flu, epsrel=epsrel)        
         return excite*collect
 
-    def plot_sensitivity(self, filename='', n=50, **kwargs):
+    def plot_sensitivity(self, filename='', n=50, kappa=None, **kwargs):
         directions = util.fibonacci_sphere(n)
         print('Generating data for microscope: '+filename)
-        I = np.apply_along_axis(self.calc_sensitivity, 1, directions)
+        I = np.apply_along_axis(self.calc_sensitivity, 1, directions, kappa=kappa)
         print('Plotting data for microscope: '+filename)
         util.plot_sphere(filename, directions=directions, data=I, **kwargs)
 
-    def calc_excitation_efficiency(self, direction):
-        flu = fluorophore.Fluorophore(mu_abs=direction, mu_em=direction)
+    def calc_excitation_efficiency(self, direction, kappa=None):
+        flu = fluorophore.Fluorophore(mu_abs=direction, mu_em=direction, kappa=kappa)
         I = self.illuminator.calc_excitation_efficiency(flu)
         return I
     
-    def plot_excitation_efficiency(self, filename='', n=50, **kwargs):
+    def plot_excitation_efficiency(self, filename='', n=50, kappa=None, **kwargs):
         directions = util.fibonacci_sphere(n)
         print('Generating data for microscope: '+filename)
-        I = np.apply_along_axis(self.calc_excitation_efficiency, 1, directions)
+        I = np.apply_along_axis(self.calc_excitation_efficiency, 1, directions, kappa=kappa)
+
         print('Plotting data for microscope: '+filename)
         util.plot_sphere(filename, directions=directions, data=I, **kwargs)
 
-    def calc_collection_efficiency(self, direction):
[…]

### 2017-09-19 dipsim/util.py

@@ -154,16 +154,17 @@ def generate_caxs(axs):
     return np.array(caxs).reshape(axs.shape)
 
 def draw_scene(scene_string, filename='out.png', my_ax=None, dpi=500,
-               save_file=False):
+               save_file=False, chop=True):
                 
     asy_string = """
     import three;
+    import graph3;
     settings.outformat = "pdf";
     settings.prc = true;
     settings.embed= true;
     settings.render=16;
 
-    size(6cm,0);
+    size(6cm,6cm);
     currentprojection = orthographic(1, 1, 1);
 
     void circle(real Theta, real Alpha, bool dash, triple color) {
@@ -205,15 +206,38 @@ def draw_scene(scene_string, filename='out.png', my_ax=None, dpi=500,
       }
     }
 
-    // Sphere
-    draw(unitsphere, surfacepen=material(diffusepen=white+opacity(0.1), emissivepen=grey, specularpen=white));
+    void watson(real Theta, real Phi, real kappa, real x, real y, real z) {
+     int n_phi = 10;
+     int n_theta = 10;
 
-    // Draw points on sphere
-    dotfactor = 7;
-    dot(X); 
-    dot(Y); 
+     real max_radius = 0;
+     if(kappa > 0){
+       max_radius = exp(kappa);
+     }
+     else{
+       max_radius = 1.0;
+     }
 
-    circle(0, pi/2, false, (0, 0, 0));
+     for(int i=0; i <= n_theta; ++i) {
+       real Theta_i = pi*i/n_theta;
+       real weight = exp(kappa*(cos(Theta_i)**2))/max_radius;     
+       path3 mycircle = circle(c=Z*weight*cos(Theta_i), r=weight*sin(Theta_i));
+       draw(shift((x, y, z))*rotate(angle=degrees(Phi), u=O, v=Z)*rotate(angle=degrees(Theta), u=O, v=Y)*mycircle);
+     }
+
+     triple f(real t) {
+       real weight = exp(kappa*(cos(t)**2))/max_radius;
+       return (0, weight*sin(t), weight*cos(t));
+     }
+     path3 phi_path = graph(f, 0, 2pi, operator ..);
+
+     for(int i=0; i <= n_phi; ++i) {
+       real Phi_i = 2*pi*i/n_theta;
[…]

### 2017-09-19 notes/2017-09-19-ensemble-efficiencies/figures/benchmarks.py

@@ -1,0 +1,33 @@
+from dipsim import multiframe, util, detector, illuminator, microscope, util
+import numpy as np
+import matplotlib.pyplot as plt
+
+import os; import time;  print('Running...')
+
+# Compute and plot on axes
+pol = np.pi/4
+det_axis = np.pi/4
+ill_axis = -np.pi/4
+det_na = 0.8
+ill_na = 0.8
+kappa = 1
+
+# Create microscope
+ill = illuminator.Illuminator(illum_type='wide',
+                              theta_optical_axis=ill_axis,
+                              na=ill_na,
+                              phi_pol=pol)
+
+det = detector.Detector(theta_optical_axis=np.array(det_axis),
+                        na=det_na)
+
+m = microscope.Microscope(illuminator=ill, detector=det, max_photons=100)
+
+# Calculate total efficiency
+epsrel = 1000
+kappas = [None, 0]#np.arange(-10, 10, 1)
+for kappa in kappas:
+    start = time.time();
+    x = m.calc_sensitivity(direction=(0,0), kappa=kappa, epsrel=epsrel)
+    print('kappa = '+str(kappa)+'\t'+str(np.round(x, 5))+'\t'+str(np.round(time.time() - start, 5)))
+#os.system('say "done"')

### 2017-09-19 notes/2017-09-19-ensemble-efficiencies/figures/single-frame.py

@@ -1,0 +1,69 @@
+from dipsim import multiframe, util, detector, illuminator, microscope, util
+import numpy as np
+import matplotlib.pyplot as plt
+
+import os; import time; start = time.time(); print('Running...')
+
+# Main input parameters
+n_pts = 250
+pols = 4*[np.pi/4]
+ill_axes = 4*[np.pi/4]
+det_axes = 4*[-np.pi/4]
+ill_nas = 4*[0.8]
+det_nas = 4*[0.8]
+kappas = [-3, 0, 3, 6]
+n_rows = len(det_axes)
+n_cols = 5
+inch_fig = 5
+dpi = 300
+
+col_labels = ['Fluorophore Geometry', 'Microscope Geometry', r'Absorption Efficiency $\eta_{\textrm{abs}}$', r'Detection Efficiency $\eta_{\textrm{det}}$', r'Total Efficiency $\eta_{\textrm{tot}}$']
+row_labels = ['$\kappa = -3$', '$\kappa = 0$', '$\kappa = 3$', '$\kappa = 6$']
+
+# Generate axes
+size = (inch_fig*n_cols, inch_fig*n_rows)
+fig, axs = plt.subplots(n_rows, n_cols, figsize=size)
+plt.subplots_adjust(wspace=0.1, hspace=-0.1)
+if len(pols) == 1:
+    axs = np.expand_dims(axs, 1)
+caxs = util.generate_caxs(axs)
+
+# Compute and plot on axes
+for i, (pol, det_axis, ill_axis, det_na, ill_na, kappa) in enumerate(zip(pols, det_axes, ill_axes, det_nas, ill_nas, kappas)):
+    print('Computing microscope: ' + str(i))
+
+    # Create microscope
+    ill = illuminator.Illuminator(illum_type='wide',
+                                  theta_optical_axis=ill_axis,
+                                  na=ill_na,
+                                  phi_pol=pol)
+    
+    det = detector.Detector(theta_optical_axis=np.array(det_axis),
+                            na=det_na)
+                                  
+    m = microscope.Microscope(illuminator=ill, detector=det, max_photons=1)
+    
+    # Plot scene and efficiencies
+    m.plot_excitation_efficiency(n=n_pts, kappa=kappa, my_ax=axs[i,2], my_cax=caxs[i,2],
+                                 color_min=0, color_max=0.75)
[…]

### 2017-09-19 notes/2017-09-19-ensemble-efficiencies/figures/vonmises.py

@@ -1,0 +1,29 @@
+from mpl_toolkits.mplot3d import Axes3D
+import matplotlib.pyplot as plt
+import numpy as np
+import matplotlib.gridspec as gridspec
+import scipy.special as special
+
+inch_fig = 3
+f, axs = plt.subplots(nrows=1, ncols=7, figsize=(7*inch_fig, inch_fig), subplot_kw={'projection':'3d'})
+plt.subplots_adjust(wspace=-0.7)
+kappas = [0.1, 0.5, 1.0, 2.0, 3.0, 5.0, 7.0]
+for ax, kappa in zip(axs, kappas):
+    ax.set_aspect("equal")
+    
+    phi, theta = np.mgrid[0:2*np.pi:20j, 0:np.pi:30j]
+    norm = np.sqrt(kappa)/(((2*np.pi)**(1.5))*special.iv(0.5, kappa))
+    weight = np.exp(kappa*np.cos(theta))
+    x = norm*weight*np.cos(phi)*np.sin(theta)
+    y = norm*weight*np.sin(phi)*np.sin(theta)
+    z = norm*weight*np.cos(theta)
+    ax.plot_wireframe(x, y, z - 0.7, color="k", lw=0.1)
+
+    ax.set_xlim(-0.9, 0.9)
+    ax.set_ylim(-0.9, 0.9)
+    ax.set_zlim(-0.9, 0.9)    
+    ax.set_axis_off()
+    ax.azim = 0
+    ax.elev = 30
+    ax.set_title('$\kappa = '+str(kappa)+'$', y=0.9, fontsize=20)
+f.savefig('vonmises.pdf')

### 2017-09-19 notes/2017-09-19-ensemble-efficiencies/figures/watson.py

@@ -1,0 +1,30 @@
+from mpl_toolkits.mplot3d import Axes3D
+import matplotlib.pyplot as plt
+import numpy as np
+import matplotlib.gridspec as gridspec
+import scipy.special as special
+
+
+inch_fig = 3
+f, axs = plt.subplots(nrows=1, ncols=7, figsize=(7*inch_fig, inch_fig), subplot_kw={'projection':'3d'})
+plt.subplots_adjust(wspace=-0.7)
+kappas = [-3, -2, -.5, 0, .5, 2, 3]
+for ax, kappa in zip(axs, kappas):
+    ax.set_aspect("equal")
+    
+    phi, theta = np.mgrid[0:2*np.pi:20j, 0:np.pi:30j]
+    norm = 1.0/(4*np.pi*special.hyp1f1(0.5, 1.5, kappa))
+    weight = np.exp(kappa*(np.cos(theta)**2))
+    x = norm*weight*np.cos(phi)*np.sin(theta)
+    y = norm*weight*np.sin(phi)*np.sin(theta)
+    z = norm*weight*np.cos(theta)
+    ax.plot_wireframe(x, y, z, color="k", lw=0.1)
+
+    ax.set_xlim(-0.3, 0.3)
+    ax.set_ylim(-0.3, 0.3)
+    ax.set_zlim(-0.3, 0.3)    
+    ax.set_axis_off()
+    ax.azim = 0
+    ax.elev = 30
+    ax.set_title('$\kappa = '+str(kappa)+'$', y=0.9, fontsize=20)
+f.savefig('watson.pdf')

### 2017-09-20 notes/2017-09-19-ensemble-efficiencies/report/report.tex
Clarified that notes only consider polarized illumination.

@@
   Chandler}% and Patrick La Rivi\`ere}
-\date{\vspace{-1em}\today\vspace{-1em}}
+\date{\vspace{-1em}September 19, 2017\\ (Last Updated: \today)\vspace{-1em}}
 \maketitle
@@
 \end{align} 
-which is only true if the fluorophores emit light independently. I haven't given
-the full expression for $\eta_{\text{det}}^{\text{single}}$ here---see the paper
-or equation 2 with $D = 1$.
+which is only true if the fluorophores emit light independently. If the detector
+uses a polarizer then $\eta_{\text{det}}^{\text{single}}$ is given by equation 2 with $D = 1$ (see Fourkas \cite{fourkas2001}). If the detector does not use a polarizer then
+\begin{align}
+  \eta_{\text{det}}^{\text{single}} = 2(A + B\sin^2\Theta).
+\end{align}
 
@@
   $\kappa$; 2) microscope geometry schematics---all cases use two 0.8 NA
-  objectives in a diSPIM geometry 3) the absorption efficiency as a function of
-  $\bs{\hat{\mu}}$, see equation \ref{eq:ensemble}; 3) the detection efficiency
-  as a function of $\bs{\hat{\mu}}$, see equation \ref{eq:ensemble_det}; (4) the
-  total efficiency, the product of the absorption and detection
+  objectives in a diSPIM geometry with polarized illumination and unpolarized
+  detection 3) the absorption efficiency as a function of $\bs{\hat{\mu}}$, see
+  equation \ref{eq:ensemble}; 3) the detection efficiency as a function of
+  $\bs{\hat{\mu}}$, see equation \ref{eq:ensemble_det}; (4) the total
+  efficiency, the product of the absorption and detection
   efficiencies. \textbf{Note:} The last three columns do not use the same

### 2017-09-27 paper/paper/paper.tex
Change most instances of "molecule" to "dipole".

@@
   We investigate the use of polarized illumination in multiview microscopes for
-  determining the orientation of single fluorescent molecules. First, we relate
-  the orientation of single molecules to measurable intensities in multiview
-  microscopes and develop an information-theoretic metric---the solid-angle
-  uncertainty---to compare the ability of multiview microscopes to estimate the
-  orientation of single molecules. Next, we compare a broad class of microscopes
-  using this metric---single- and dual-view microscopes with varying
-  illumination polarization, illumination numerical aperture (NA), detection NA,
-  obliquity, asymmetry, and exposure. We find that multiview microscopes are
-  better at determining the orientation of single molecules than single-view
-  microscopes. We also find that choosing a small illumination NA and a large
-  detection NA are good design choices, that multiview microscopes can benefit
-  from oblique illumination and detection, and that asymmetric NA microscopes
-  can benefit from exposure asymmetry.
+  determining the orientation of single-molecule fluorescence transition
+  dipoles. First, we relate the orientation of single dipoles to measurable
+  intensities in multiview microscopes and develop an information-theoretic
+  metric---the solid-angle uncertainty---to compare the ability of multiview
+  microscopes to estimate the orientation of single dipoles. Next, we compare
+  a broad class of microscopes using this metric---single- and dual-view
+  microscopes with varying illumination polarization, illumination numerical
+  aperture (NA), detection NA, obliquity, asymmetry, and exposure. We find that
+  multi-view microscopes can measure all dipole orientations, while the
+  orientations measurable with single-view microscopes is halved because of
+  symmetries in the detection process. We also find that choosing a small
+  illumination NA and a large detection NA are good design choices, that
+  multiview microscopes can benefit from oblique illumination and detection, and
+  that asymmetric NA microscopes can benefit from exposure asymmetry.
 \end{abstract*}
@@
 \section{Introduction}
-The orientation of single molecules is a valuable reporter for
-biological processes. By imaging biological samples with fluorescent probes that
-are fixed to known structures, researchers have studied the dynamics of motor
-proteins \cite{peterman2001, forkey2003}, DNA \cite{backer2016}, actin
-\cite{mehta2016}, septin \cite{demay2011, mcquilken2017}, and membranes
-\cite{anantharam2010}. The number of techniques available to biologists for
-measuring the orientation of single molecules in live cells is continually
-expanding. A recent review \cite{backlund2014} identified three major categories
-of single-molecule orientation determination techniques:
+The orientation of single-molecule fluorescence transition dipoles is a valuable
+reporter for biological processes. By imaging biological samples with
+fluorescent probes that are fixed to known structures, researchers have studied
+the dynamics of motor proteins \cite{peterman2001, forkey2003}, DNA
+\cite{backer2016}, actin \cite{mehta2016}, septin \cite{demay2011,
+  mcquilken2017}, and membranes \cite{anantharam2010}. The number of techniques
+available to biologists for measuring the orientation of single dipoles in live
+cells is continually expanding. A recent review \cite{backlund2014} identified
+three major categories of single-dipole orientation determination techniques:
 
@@
 and use the distribution of intensity in the back focal plane \cite{lieb2004} or
-in the image plane \cite{backer2014} to estimate their
+in the image plane \cite{backer2014} to estimate their dipole
 orientation. Information-theoretically optimal techniques in this category use
@@
 position information in the image \cite{agrawal2012}. These techniques can
-estimate the orientation of molecules in all orientations and only require a
+estimate the orientation of dipoles in all orientations and only require a
 single frame of intensity measurements, but they require complex reconstruction
@@
 \textbf{Spatial variation of illumination polarization} techniques vary the
-polarization and intensity in the illumination path to excite molecules in
+polarization and intensity in the illumination path to excite dipoles in
 specific orientations in the focal volume \cite{debarre2004}. Although these
-techniques can accurately estimate the orientation of molecules in all
+techniques can accurately estimate the orientation of dipoles in all
 orientations, they require scanning which makes acquisition time a concern in
@@
 light in the illumination or detection paths and exploit the anisotropic
-absorption and emission patterns of single molecules to estimate their
+absorption and emission patterns of single molecules to estimate their dipole
 orientation. By illuminating and detecting from the focal plane, not just the
-focal point, these techniques can estimate the orientation of many molecules in
+focal point, these techniques can estimate the orientation of many dipoles in
 parallel without scanning. These techniques are easy to implement---changing the
@@
 techniques require several intensity measurements under different polarization
-orientations to estimate the orientation of a single molecule. Despite the
+orientations to estimate the orientation of a single dipole. Despite the
 requirement for multiple measurements, polarized illumination or detection
@@
 
-All single molecule orientation determination methods suffer from some degree of
-anisotropic orientation uncertainty, i.e., some molecular orientations cannot be
+All single-dipole orientation determination methods suffer from some degree of
+anisotropic orientation uncertainty, i.e., some dipole orientations cannot be
 determined as precisely as others. In some cases the orientation cannot be
@@
 the only authors who have considered anisotropic orientation uncertainty in 3D
-have only analyzed a subset of molecular orientations \cite{agrawal2012}. Our
-view is that an ideal technique for determining the orientation of single
-molecules can reconstruct the orientation with a small and nearly uniform
-uncertainty for all molecular orientations.
+have only analyzed a subset of dipole orientations \cite{agrawal2012}. Our view
+is that an ideal technique for determining the orientation of single dipoles can
+reconstruct the orientation with a small and nearly uniform uncertainty for all
+dipole orientations.
 
@@
 In this work we explore the use of polarized illumination in multiview
-microscopes for determining the orientation of single molecules. Existing
+microscopes for determining the orientation of single dipoles. Existing
 multiview microscopes can easily be outfitted with fast-switching polarizers
@@
 augment multiview microscopes for measuring the orientation of single
-molecules. Furthermore, multiview microscopes can achieve nearly isotropic
+dipoles. Furthermore, multiview microscopes can achieve nearly isotropic
 resolution while delivering selective illumination. We find that multiview
 microscopes also provide a small and uniform orientation uncertainty for single
-molecules in all orientations. We also find that single-view microscopes have
+dipoles in all orientations. We also find that single-view microscopes have
 symmetries that can be broken by adding a second view. This reduction of
 degeneracy allows multiview microscopes to uniquely determine the orientation of
-single molecules.
+single dipoles.
 
@@
 In this section we develop a method to compare multiview polarized illumination
-microscopes for the task of reconstructing molecular orientations. In sections
+microscopes for the task of reconstructing dipole orientations. In sections
 \ref{excitation}--\ref{forward} we develop a model for the intensity collected
 with a fixed illumination and detection configuration as a function of the
-molecule orientation. In section \ref{designs} we list a variety of ways to
+dipole orientation. In section \ref{designs} we list a variety of ways to
 combine these intensity measurements into experimentally realizable polarized
@@
 In this section we will calculate the \emph{absorption efficiency} of a single
-molecule---the fraction of the incident power that is absorbed by the
-molecule. Our approach is inspired by Fourkas \cite{fourkas2001}, but here we
+dipole---the fraction of the incident power that is absorbed by the
+dipole. Our approach is inspired by Fourkas \cite{fourkas2001}, but here we
 calculate the absorption efficiency instead of the detection efficiency.
@@
 
-  To find the absorption efficiency $\eta_{\text{abs}}$ of a single molecule if
+  To find the absorption efficiency $\eta_{\text{abs}}$ of a single dipole if
   it were illuminated by a single ray, we take the unit incident electric field
@@
 
-    To find the absorption efficiency of a single molecule, we integrate over
+    To find the absorption efficiency of a single dipole, we integrate over
     all illumination rays and divide by the total incident power which gives the
@@
 \end{subequations}
+Note that the factor $D$ keeps the specimen irradiance constant for any value of
+$\alpha$.
 
@@
 with the optical axis. Under slowly-scanned and weakly-focused light-sheet
-illumination, molecules in the sample will be excited as if a single plane wave
-was incident. Therefore, we can find the absorption efficiency of a molecule in
+illumination, dipoles in the sample will be excited as if a single plane wave
+was incident. Therefore, we can find the absorption efficiency of a dipole in
 this case by taking the limit of equation \ref{eq:scalarabs} as
@@
 that light-sheet illumination is approximately equivalent to low-NA illumination
-with respect to the molecular orientations they excite. Notice that the
+with respect to the dipole orientations they excite. Notice that the
 absorption efficiency can take its maximum range of values
@@
 In this section we will calculate the \emph{detection efficiency}---the fraction
-of the power emitted by a single molecule that we detect. Fourkas calculated the
+of the power emitted by a single dipole that we detect. Fourkas calculated the
 detection efficiency when an ideal objective with its optical axis aligned with
-the $\mh{z}$ axis is focused on a single molecule and a polarizer is placed in
+the $\mh{z}$ axis is focused on a single dipole and a polarizer is placed in
 the aperture plane \cite{fourkas2001}. He found that
@@
 measurement. This means that a single intensity measurement does not give us
-enough information to reconstruct the orientation of a single molecule. Next, we
+enough information to reconstruct the orientation of a single dipole. Next, we
 will consider combining several intensity measurements to create polarized
 illumination microscopes that can reconstruct the orientation of single
-molecules.
+dipoles.
 
@@
 Our goal is to evaluate the ability of a microscope design to estimate the
-orientation of a single molecule (via parameters $\Theta$ and $\Phi$) from
+orientation of a single dipole (via parameters $\Theta$ and $\Phi$) from
 intensity data. A common way to evaluate the ability to estimate the parameters
@@
 minimum variance of an unbiased estimator for each parameter. For each
-microscope design and molecule orientation we calculate the Fisher information
+microscope design and dipole orientation we calculate the Fisher information
 matrix. We assume that the detected intensities are Poisson distributed, so the
@@
 equally spaced points on the unit sphere. A desirable microscope design will
-have a small solid-angle uncertainty that is uniform for all molecular
+have a small solid-angle uncertainty that is uniform for all dipole
 orientations. The most straightforward way to find the location and scale of the
@@
 We also use the Fisher information matrix to calculate uncertainty ellipses for
-each microscope and fluorophore orientation. The general equation of an ellipse
+each microscope and dipole orientation. The general equation of an ellipse
 with axis radii $a$ and $b$ and rotation angle $\gamma$ is
@@
 small number of degeneracies is important for a microscope to find the
-orientation of a molecule---a twofold degenerate microscope can find the
-orientation of a molecule down to the hemisphere while a fourfold degenerate
+orientation of a dipole---a twofold degenerate microscope can find the
+orientation of a dipole down to the hemisphere while a fourfold degenerate
 microscope can only find the orientation down to the quadrant.
@@
 \ref{fig:single-arm}c) that single-view designs suffer from high orientation
-uncertainty when the molecules are oriented along the optical axis (the
-molecules are not efficiently excited or detected), near the transverse plane
-(it is ambiguous whether the molecule is oriented above or below the transverse
+uncertainty when the dipoles are oriented along the optical axis (the
+dipoles are not efficiently excited or detected), near the transverse plane
+(it is ambiguous whether the dipole is oriented above or below the transverse
 plane, see Figure \ref{fig:single-arm}b)), and near the polarizer orientations
-(it is ambiguous whether the molecule is on either side of the polarizer
+(it is ambiguous whether the dipole is on either side of the polarizer
 orientation, see Figure \ref{fig:single-arm}b)).
@@
 the second degeneracy is found in Figures \ref{fig:single-arm}b) and
-\ref{fig:single-arm}c) which show the ambiguity of whether the molecule is
+\ref{fig:single-arm}c) which show the ambiguity of whether the dipole is
 oriented above or below the transverse plane. Near the transverse plane this
@@
 consider in this paper are only twofold degenerate. This means that while
-dual-view microscopes can find the orientation of a molecule up to the
-hemisphere, single-view microscopes can only find the orientation of a molecule
+dual-view microscopes can find the orientation of a dipole up to the
+hemisphere, single-view microscopes can only find the orientation of a dipole
 up to the quadrant.
@@
 approximately equivalent to light-sheet illumination for the task of estimating
-the orientation of molecules. Therefore, the preference for low-NA
+the orientation of dipoles. Therefore, the preference for low-NA
 illumination also implies that dual-view light-sheet illumination geometries are
 an excellent choice for uniformly reconstructing the orientation of single
-molecules.
+dipoles.
 
@@
 metrics to measure the ability of a microscope design to estimate the
-orientation of single molecules, and used these metrics to compare a wide
+orientation of single dipoles, and used these metrics to compare a wide
 range of microscope designs. Our main result is a short list of design
-heuristics that can be used to design polarized-illumination single-molecule
+heuristics that can be used to design polarized-illumination single-dipole
 orientation microscopes:
@@
 would favor microscopes with a large detection NA. Finally, we have focused on
-the task of estimating the orientation of molecules and ignored the equally
+the task of estimating the orientation of dipoles and ignored the equally
 important task of estimating their location. The task of jointly estimating the
-orientation and location of single molecules from multiple views is a topic for
+orientation and location of single dipoles from multiple views is a topic for
 future work.
@@
 Although we have only considered the orientation determination of fixed single
-molecules, polarized illumination techniques can be used to study the
-orientation distribution of ensembles of dynamic molecules \cite{mehta2016,
-  backer2016}. In future work we will investigate the ability of multiview
-polarized illumination microscopes to estimate the orientation distribution of
-ensembles of dynamic molecules.
+dipoles, polarized illumination techniques can be used to study the orientation
+distribution of ensembles of dynamic dipoles \cite{mehta2016, backer2016}. In
+future work we will investigate the ability of multiview polarized illumination
+microscopes to estimate the orientation distribution of ensembles of dynamic
+dipoles.
 
@@
 detection arm can increase the information available for reconstructing the
-orientation of a single molecule without increasing the acquisition time.
+orientation of a single dipole without increasing the acquisition time.
 
 We conclude that multiview microscopes are useful tools for determining the
-orientation of single molecules. Using simple design heuristics, we can design
+orientation of single dipoles. Using simple design heuristics, we can design
 polarized illumination multiview microscopes that can determine the orientation
-of single molecules with a small and uniform orientation uncertainty.
+of single dipoles with a small and uniform orientation uncertainty.
 

### 2017-09-29 dipsim/detector.py
Fix fluorophore distribution error. Integrate intensities, not eff.

@@ -1,7 +1,5 @@
 import numpy as np
 import dipsim.util as util
-import dipsim.fluorophore as flu
-from scipy import integrate, special
 
 class Detector:
     """A Detector is specified by its optical axis, numerical aperture, and
@@ -16,29 +14,16 @@ class Detector:
         self.alpha = np.arcsin(self.na/self.n)
         self.phi_pol = phi_pol
 
-    def calc_collection_efficiency(self, fluorophore, epsrel=1e-2):
-        if fluorophore.kappa == None:
-            theta = util.theta_prime(fluorophore.mu_em[0], fluorophore.mu_em[1], self.theta_optical_axis)
-            phi = util.phi_prime(fluorophore.mu_abs[0], fluorophore.mu_abs[1], self.theta_optical_axis)        
-            A = (1.0/4.0) - (3.0/8.0)*np.cos(self.alpha) + (1.0/8.0)*(np.cos(self.alpha)**3)
-            B = (3.0/16.0)*np.cos(self.alpha) - (3.0/16.0)*(np.cos(self.alpha)**3)
-            C = (7.0/32.0) - (3.0/32.0)*np.cos(self.alpha) - (3.0/32.0)*(np.cos(self.alpha)**2) - (1.0/32.0)*(np.cos(self.alpha)**3)
-            if self.det_type=='lens':
-                return 2*(A + B*(np.sin(theta)**2))
-            elif self.det_type=='polarized':
-                return A + B*(np.sin(theta)**2) + C*(np.sin(theta)**2)*np.cos(2*(phi - self.phi_pol))
-            elif self.det_type=='4pi':
-                return 1.0
-        else:
-            def int_func(theta, phi):
-                theta_p = fluorophore.mu_em[0]
-                phi_p = fluorophore.mu_em[1]
-                eta_det_single = self.calc_collection_efficiency(flu.Fluorophore(mu_abs=(theta, phi), mu_em=(theta,phi)))
-                norm = 1.0/(4*np.pi*special.hyp1f1(0.5, 1.5, fluorophore.kappa))
-                weight = np.exp(fluorophore.kappa*(np.dot(util.tp2xyz((theta, phi)), util.tp2xyz((theta_p, phi_p)))**2))
-                jacobian = np.sin(theta)
-                return jacobian*norm*weight*eta_det_single
[…]

### 2017-09-29 dipsim/fluorophore.py

@@ -2,17 +2,12 @@ import numpy as np
 
 class Fluorophore:
     """A single fluorophore is specified by its 3D position, (unit) absorption
-    dipole moment (theta, phi), and (unit) emission dipole moment (theta, phi),
-    and ensemble distribution (kappa = None means single molecule. kappa != None
-    means Watson distribution).
-
+    dipole moment (theta, phi), and (unit) emission dipole moment (theta, phi).
     """
     def __init__(self, position=np.array([0, 0, 0]),
                  mu_abs=np.array([0, 0]),
-                 mu_em=np.array([0, 0]),
-                 kappa=None):
+                 mu_em=np.array([0, 0])):
         self.position = position
         self.mu_abs = mu_abs        
         self.mu_em = mu_em
         self.mu_ind = 0
-        self.kappa = kappa

### 2017-09-29 dipsim/illuminator.py

@@ -1,7 +1,5 @@
 import numpy as np
 import dipsim.util as util
-import dipsim.fluorophore as flu
-from scipy import integrate, special
 
 class Illuminator:
     """An Illuminator is specified by its illumination type (wide, sheet),
@@ -17,29 +15,18 @@ class Illuminator:
         self.alpha = np.arcsin(self.na/self.n)
         self.phi_pol = phi_pol
 
-    def calc_excitation_efficiency(self, fluorophore, epsrel=1e-2):
-        if fluorophore.kappa == None:
-            theta = util.theta_prime(fluorophore.mu_abs[0], fluorophore.mu_abs[1], self.theta_optical_axis)
-            phi = util.phi_prime(fluorophore.mu_abs[0], fluorophore.mu_abs[1], self.theta_optical_axis)
-            A = (1.0/4.0) - (3.0/8.0)*np.cos(self.alpha) + (1.0/8.0)*(np.cos(self.alpha)**3)
-            B = (3.0/16.0)*np.cos(self.alpha) - (3.0/16.0)*(np.cos(self.alpha)**3)
-            C = (7.0/32.0) - (3.0/32.0)*np.cos(self.alpha) - (3.0/32.0)*(np.cos(self.alpha)**2) - (1.0/32.0)*(np.cos(self.alpha)**3)        
-            D = 4.0/(3.0*(1.0 - np.cos(self.alpha)))
-            if self.illum_type == 'wide':
-                return D*(A + B*(np.sin(theta)**2) + C*(np.sin(theta)**2)*np.cos(2*(phi - self.phi_pol)))
-            elif self.illum_type == 'sheet':
-                return (np.sin(theta)*np.cos(phi - self.phi_pol))**2
-            elif self.illum_type == 'unpolarized':
-                return 2*D*(A + B*(np.sin(theta)**2))
-            else:
-                print("Warning: invalid illum_type")
+    def calc_excitation_efficiency(self, fluorophore):
+        theta = util.theta_prime(fluorophore.mu_abs[0], fluorophore.mu_abs[1], self.theta_optical_axis)
+        phi = util.phi_prime(fluorophore.mu_abs[0], fluorophore.mu_abs[1], self.theta_optical_axis)
+        A = (1.0/4.0) - (3.0/8.0)*np.cos(self.alpha) + (1.0/8.0)*(np.cos(self.alpha)**3)
[…]

### 2017-09-29 dipsim/microscope.py

@@ -3,6 +3,7 @@ import matplotlib.pyplot as plt
 import functools
 import vispy
 from dipsim import util, fluorophore, visuals
+from scipy import integrate, special
 
 class Microscope:
     """
@@ -18,44 +19,55 @@ class Microscope:
         self.max_photons = max_photons
         self.color = color
 
-    def calc_intensity(self, direction):
-        return self.max_photons*self.calc_sensitivity(direction)
-    
-    def calc_sensitivity(self, direction, kappa=None, epsrel=1e-2):
-        flu = fluorophore.Fluorophore(mu_abs=direction, mu_em=direction, kappa=kappa)
-        excite = self.illuminator.calc_excitation_efficiency(flu, epsrel=epsrel)
-        collect = self.detector.calc_collection_efficiency(flu, epsrel=epsrel)        
-        return excite*collect
+    def calc_intensity(self, direction, kappa=None, epsrel=1e-2):
+        fluo = fluorophore.Fluorophore(mu_abs=direction, mu_em=direction)        
+        if kappa==None:
+            excite = self.illuminator.calc_excitation_efficiency(fluo)
+            collect = self.detector.calc_collection_efficiency(fluo)
+            return self.max_photons*excite*collect
+        else:
+            def int_func(theta, phi):
+                theta_p = fluo.mu_em[0]
+                phi_p = fluo.mu_em[1]
+                int_single = self.calc_intensity((theta, phi), kappa=None)
+                norm = 1.0/(4*np.pi*special.hyp1f1(0.5, 1.5, kappa))
+                weight = np.exp(kappa*(np.dot(util.tp2xyz((theta, phi)), util.tp2xyz((theta_p, phi_p)))**2))
+                jacobian = np.sin(theta)
+                return jacobian*norm*weight*int_single
+            #return integrate.dblquad(int_func, 0, 2*np.pi, lambda x: 0, lambda y: np.pi, epsrel=epsrel, epsabs=0)[0]            
+            integral = integrate.nquad(int_func, [[0, np.pi], [0, 2*np.pi]], opts={'epsrel':epsrel, 'epsabs':0, 'limit':1}, full_output=True)
+            print(integral)
+            return integral[0]
 
-    def plot_sensitivity(self, filename='', n=50, kappa=None, **kwargs):
+    def plot_intensity(self, filename='', n=50, kappa=None, **kwargs):
         directions = util.fibonacci_sphere(n)
         print('Generating data for microscope: '+filename)
[…]

### 2017-09-29 dipsim/util.py

@@ -40,6 +40,15 @@ def fibonacci_sphere(n):
     phi = np.mod((np.pi*(3.0 - np.sqrt(5.0)))*np.arange(n), 2*np.pi) - np.pi
     return np.vstack((theta, phi)).T
 
+def sphere_profile(n):
+    # Returns equally spaced points on a profile around the unit sphere in
+    # spherical coordinates. Only on the equator for now. 
+    z = np.linspace(1 - 1/n, -1 + 1/n, num=n)
+    theta = n*[np.pi/2]
+    phi = np.linspace(-np.pi, np.pi, num=n)
+    return np.vstack((theta, phi)).T
+
+
 def plot_sphere(filename=None, directions=None, data=None, interact=False,
                 color_norm='linear', color_min=0, color_max=None,
                 gamma=0.25, color_map='coolwarm', linthresh=1e-3,

### 2017-09-29 notes/2017-09-19-ensemble-efficiencies/figures/single-frame.py

@@ -13,11 +13,11 @@ ill_nas = 4*[0.8]
 det_nas = 4*[0.8]
 kappas = [-3, 0, 3, 6]
 n_rows = len(det_axes)
-n_cols = 5
+n_cols = 3
 inch_fig = 5
 dpi = 300
 
-col_labels = ['Fluorophore Geometry', 'Microscope Geometry', r'Absorption Efficiency $\eta_{\textrm{abs}}$', r'Detection Efficiency $\eta_{\textrm{det}}$', r'Total Efficiency $\eta_{\textrm{tot}}$']
+col_labels = ['Fluorophore Geometry', 'Microscope Geometry', '$I$ (Arbitrary Units)']
 row_labels = ['$\kappa = -3$', '$\kappa = 0$', '$\kappa = 3$', '$\kappa = 6$']
 
 # Generate axes
@@ -44,12 +44,8 @@ for i, (pol, det_axis, ill_axis, det_na, ill_na, kappa) in enumerate(zip(pols, d
     m = microscope.Microscope(illuminator=ill, detector=det, max_photons=1)
     
     # Plot scene and efficiencies
-    m.plot_excitation_efficiency(n=n_pts, kappa=kappa, my_ax=axs[i,2], my_cax=caxs[i,2],
-                                 color_min=0, color_max=0.75)
-    m.plot_collection_efficiency(n=n_pts, kappa=kappa, my_ax=axs[i,3], my_cax=caxs[i,3],
-                                 color_min=0, color_max=0.25)
-    m.plot_sensitivity(n=n_pts, kappa=kappa, my_ax=axs[i,4], my_cax=caxs[i,4],
-                                 color_min=0, color_max=0.15)
+    m.plot_intensity(n=n_pts, kappa=kappa, my_ax=axs[i,2], my_cax=caxs[i,2],
+                     color_min=0, color_max=0.06)
 
     scene_string = m.scene_string()
     util.draw_scene(scene_string, my_ax=axs[i,1], dpi=dpi)

### 2017-09-29 notes/2017-09-28-anisotropy/report/report.tex
Anisotropy verification notes.

\documentclass[11pt]{article}

%%%%%%%%%%%%
% Packages %
%%%%%%%%%%%%
\hyphenpenalty=10000
\usepackage{tocloft}
\renewcommand\cftsecleader{\cftdotfill{\cftdotsep}}
\def\undertilde#1{\mathord{\vtop{\ialign{##\crcr
$\hfil\displaystyle{#1}\hfil$\crcr\noalign{\kern1.5pt\nointerlineskip}
$\hfil\tilde{}\hfil$\crcr\noalign{\kern1.5pt}}}}}
\usepackage{cleveref}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{epstopdf}
\usepackage{braket}
\usepackage{upgreek}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{amssymb,latexsym,amsmath,gensymb}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{float}
\usepackage{enumitem}
\usepackage{pdflscape}
\usepackage{url}
\usepackage{tikz, calc}
\usetikzlibrary{shapes.geometric, arrows, calc}
\tikzstyle{norm} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{arrow} = [thick, ->, >=stealth]

\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
\providecommand{\mb}[1]{\mathbf{#1}}
\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
\providecommand{\bs}[1]{\boldsymbol{#1}} 
\providecommand{\intinf}{\int_{-\infty}^{\infty}}
\providecommand{\fig}[4]{
  % filename, width, caption, label
\begin{figure}[h]
 \captionsetup{width=1.0\linewidth}
 \centering
 \includegraphics[width = #2\textwidth]{#1}
 \caption{#3}
 \label{fig:#4}
\end{figure}
}

\newcommand{\tensor}[1]{\overset{\text{\tiny$\leftrightarrow$}}{\mb{#1}}}
\newcommand{\tunderbrace}[2]{\underbrace{#1}_{\textstyle#2}}
\providecommand{\figs}[7]{
  % filename1, filename2, caption1, caption2, label1, label2, shift
\begin{figure}[H]
\centering
\begin{minipage}[b]{.45\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#1}
  \captionsetup{justification=justified, singlelinecheck=true}
  \caption{#3}
  \label{fig:#5}
\end{minipage}
\hspace{2em}
\begin{minipage}[b]{.45\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#2}
  \vspace{#7em}
  \captionsetup{justification=justified}
  \caption{#4}
  \label{fig:#6}
\end{minipage}
\end{figure}
}
\makeatletter

\providecommand{\code}[1]{
\begin{center}
\lstinputlisting{#1}
\end{center}
}

\newcommand{\crefrangeconjunction}{--}
%%%%%%%%%%%
% Spacing %
%%%%%%%%%%%
% Margins
\usepackage[
top    = 1.5cm,
bottom = 1.5cm,
left   = 1.5cm,
right  = 1.5cm]{geometry}

% Indents, paragraph space
%\usepackage{parskip}
\setlength{\parskip}{1.5ex}

% Section spacing
\usepackage{titlesec}
\titlespacing*{\title}
{0pt}{0ex}{0ex}
\titlespacing*{\section}
{0pt}{0ex}{0ex}
\titlespacing*{\subsection}
{0pt}{0ex}{0ex}
\titlespacing*{\subsubsection}
{0pt}{0ex}{0ex}

% Line spacing
\linespread{1.1}

%%%%%%%%%%%%
% Document %
%%%%%%%%%%%%
\begin{document}
\title{\vspace{-2.5em} Verifying The Polarized Fluorescence Microscopy Model Using Anisotropy \vspace{-1em}} \author{Talon
  Chandler}% and Patrick La Rivi\`ere}
\date{\vspace{-1em}\today\vspace{-1em}}
\maketitle
\section{Introduction}
Chemists often use the \textit{anisotropy} of a sample to characterize
rotational dynamics and the transition moments of molecules. In these notes I
will use the polarized fluorescence microscopy model from previous notes to
calculate the anisotropy of several fluorophore distributions. I will reproduce
known results to verify the model. This work is useful for verifying that our
model is correct and for relating our work to published results.

\section{Measuring The Anisotropy}
The anisotropy, $r$, of a sample is measured by (1) exciting a sample with a
low-NA beam of linearly polarized light, (2) placing a low-NA detection arm
orthogonal to the illumination arm, (3) placing a linear polarizer in the
detection arm, (4) measuring the intensity with the detection polarizer parallel
($I_\parallel$) and perpendicular ($I_\perp$) to the illumination polarization,
(5) calculating the anisotropy using
\begin{align}
  r = \frac{I_\parallel - I_\perp}{I_\parallel + 2I_\perp}.
\end{align}
See Figure 1 for a schematic of the $I_\parallel$ and $I_\perp$ measurements.
\fig{../figures/schematic.pdf}{0.7}{Schematics for anisotropy
  measurements. Solid (dashed) arrows show the illumination (detection)
  polarization orientation.
  $\text{NA}_{\text{ill}} = \text{NA}_{\text{det}} = 0$.}{schematic}

Lakowicz \cite{lak} gives several known anisotropies for special dipole
distributions. We assume that the excitation and emission dipole moments are
collinear. If the dipole moment orientations are
\begin{itemize}
\item completely aligned and parallel to the excitation polarization
  $\rightarrow r = 1$
\item completely aligned and perpendicular to the excitation polarization
  $\rightarrow r = -0.5$
\item uniformly distributed $\rightarrow r=0.4$.
\end{itemize}

\section{Calculating The Anisotropy With The Polarized Fluorescence Microscopy Model}
We can calculate the intensities measured by the experiments in Figure 1 using
\begin{align}
  I \propto \int_{\mathbb{S}^2}d\mh{r} f(\mh{r}; \hat{\bs{\mu}}, \kappa) \eta_{\text{exc}}(\mh{r}) \eta_{\text{det}}(\mh{r}) \label{eq:ensemble_det}%
\end{align}
where $f(\mh{r}; \hat{\bs{\mu}}, \kappa)$ is the Watson distribution with
central orientation $\hat{\bs{\mu}}$ and concentration parameter $\kappa$,
$\eta_{\text{exc}}$ is the excitation efficiency of a single fluorophore and
$\eta_{\text{det}}$ is the detection efficiency of a single fluorophore. See
previous note sets and the paper for the efficiency expressions.

Figures \ref{fig:results} and \ref{fig:profile} show the intensities and
anisotropy of the experiment in Figure 1 as a function of fluorophore
distribution ($\hat{\bs{\mu}}$ and $\kappa$). The results match the special
cases mentioned in Lakowicz.  When the fluorophores are completely aligned
($\kappa = \infty$) and parallel to the excitation polarization ($\mh{y}$-axis)
then $r = 1$ (see bottom right sphere in Figure \ref{fig:results} or right plot
in Figure \ref{fig:profile}). When the fluorophores are completely aligned
($\kappa = \infty$) and perpendicular to the excitation polarization
($\mh{x}$-axis) then $r = -0.5$ (see bottom right sphere in Figure
\ref{fig:results} or right plot in Figure \ref{fig:profile}). Finally, when the
fluorophores are uniformly distributed ($\kappa = 0$) then $r = 0.4$ (see Figure
\ref{fig:results} bottom row second from the left or the right plot in Figure
\ref{fig:profile}).

\fig{../figures/anisotropy-kappa.pdf}{1.0}{Intensities and anisotropy as a
  function of fluorophore distribution. \textbf{Rows:} 1) $I_\perp$ see Figure
  1; 2) $I_\parallel$ see Figure 1; 3) Anisotropy see Equation
  1. \textbf{Columns:} Varying concentration parameter of the Watson
  distribution. Note that $\kappa = \infty$ corresponds to a single fluorophore
  or a perfectly concentrated ensemble of fluorophores.}{results}

\fig{../figures/anisotropy-profile.pdf}{1.0}{Profiles of the spheres in Figure 2
  in the $x-y$ plane. $\phi$ is the azimuth angle measured from the $+x$-axis in
  the $x-y$ plane. \textbf{Columns:} 1) $I_\perp$ see Figure 1; 2) $I_\parallel$
  see Figure 1; 3) Anisotropy see Equation 1. \textbf{Colors:} Varying
  concentration parameter of the Watson distribution.}{profile}

\bibliography{report}{}
\bibliographystyle{unsrt}

\end{document}

### 2017-09-29 notes/2017-09-28-anisotropy/figures/anisotropy-kappa.py

@@ -1,0 +1,58 @@
+from dipsim import multiframe, util, detector, illuminator, microscope, util
+import dipsim.fluorophore as flu
+import numpy as np
+import matplotlib.pyplot as plt
+
+import os; import time; start = time.time(); print('Running...')
+
+# Main input parameters
+n_pts = 1000
+kappas = [-3, 0, 3, None]
+n_cols = len(kappas)
+n_rows = 3
+inch_fig = 5
+dpi = 300
+
+col_labels = ['$\kappa = -3$', '$\kappa = 0$', '$\kappa = 3$', '$\kappa = \infty$']
+row_labels = ['$I_\perp$', '$I_\parallel$', '$r = \\frac{I_\parallel - I_\perp}{I_\parallel + 2I_\perp}$']
+
+# Generate axes
+size = (inch_fig*n_cols, inch_fig*n_rows)
+fig, axs = plt.subplots(n_rows, n_cols, figsize=size)
+plt.subplots_adjust(wspace=0.1, hspace=-0.1)
+caxs = util.generate_caxs(axs)
+
+# Compute and plot on axes
+for i, kappa in enumerate(kappas):
+    print('Computing microscope: ' + str(i))
+
+    # y-polarized input from x optical axis
+    ill = illuminator.Illuminator(illum_type='sheet', theta_optical_axis=np.pi/2, phi_pol=np.pi/2)
+    # perp: x-polarized detection
+    # par: y-polarized detection
+    det_perp = detector.Detector(det_type='polarized', na=0.01, phi_pol=0)
+    det_par = detector.Detector(det_type='polarized', na=0.01, phi_pol=np.pi/2)
+    m_perp = microscope.Microscope(illuminator=ill, detector=det_perp, max_photons=1e6)
+    m_par = microscope.Microscope(illuminator=ill, detector=det_par, max_photons=1e6)
+    test_perp = m_perp.calc_intensity((0,0), kappa=0)
+    test_par = m_par.calc_intensity((0,0), kappa=0)    
+    directions = util.fibonacci_sphere(n_pts)
+    I_perp = np.apply_along_axis(m_perp.calc_intensity, 1, directions, kappa=kappa)
+    I_par = np.apply_along_axis(m_par.calc_intensity, 1, directions, kappa=kappa)
+    
+    r = (I_par - I_perp)/(I_par + 2*I_perp)
+        
+    util.plot_sphere('', directions=directions, data=I_perp, my_ax=axs[0,i], my_cax=caxs[0,i],
[…]

### 2017-09-29 notes/2017-09-28-anisotropy/figures/anisotropy-profile.py

@@ -1,0 +1,75 @@
+from dipsim import multiframe, util, detector, illuminator, microscope, util
+import numpy as np
+import matplotlib.pyplot as plt
+
+import os; import time; start = time.time(); print('Running...')
+
+# Main input parameters
+n_pts = 200
+kappas = [-3, 0, 3, None]
+n_cols = 3
+n_rows = 1
+inch_fig = 5
+dpi = 300
+col_labels = ['$I_\perp$', '$I_\parallel$', '$r = \\frac{I_\parallel - I_\perp}{I_\parallel + 2I_\perp}$']
+
+# Generate axes
+size = (inch_fig*n_cols, 0.7*inch_fig*n_rows)
+fig, axs = plt.subplots(n_rows, n_cols, figsize=size)
+plt.subplots_adjust(wspace=0.3, hspace=-0.1)
+#caxs = util.generate_caxs(axs)
+
+# Compute and plot on axes
+for i, kappa in enumerate(kappas):
+    print('Computing microscope: ' + str(i))
+
+    # y-polarized input from x optical axis
+    ill = illuminator.Illuminator(illum_type='sheet', theta_optical_axis=np.pi/2, phi_pol=np.pi/2)
+    # perp: x-polarized detection
+    # par: y-polarized detection
+    det_perp = detector.Detector(det_type='polarized', na=0.006, phi_pol=0)
+    det_par = detector.Detector(det_type='polarized', na=0.006, phi_pol=np.pi/2)
+    m_perp = microscope.Microscope(illuminator=ill, detector=det_perp, max_photons=1e6)
+    m_par = microscope.Microscope(illuminator=ill, detector=det_par, max_photons=1e6)
+    
+    directions = util.sphere_profile(n_pts)
+    I_perp = np.apply_along_axis(m_perp.calc_intensity, 1, directions, kappa=kappa)
+    I_par = np.apply_along_axis(m_par.calc_intensity, 1, directions, kappa=kappa)
+    r = (I_par - I_perp)/(I_par + 2*I_perp)
+    
+    phi = directions[:,1]
+    if kappa == None:
+        k_string = '$\infty$'
+    else:
+        k_string = str(kappa)
+    axs[0].plot(np.rad2deg(phi), I_perp, '-')
+    axs[1].plot(np.rad2deg(phi), I_par, '-', label='$\kappa = \ $'+k_string)
[…]

### 2017-09-29 notes/2017-09-28-anisotropy/figures/anisotropy.py

@@ -1,0 +1,38 @@
+from dipsim import multiframe, util, detector, illuminator, microscope, util
+import numpy as np
+import matplotlib.pyplot as plt
+
+import os; import time; start = time.time(); print('Running...')
+
+# Compute anisotropy
+def anisotropy(phi):
+    print('Computing microscope: ')
+    # y-polarized input from x optical axis
+    ill = illuminator.Illuminator(illum_type='sheet', theta_optical_axis=np.pi/2, phi_pol=np.pi/2)
+    # perp: x-polarized detection
+    # par: y-polarized detection
+    det_perp = detector.Detector(det_type='polarized', na=0.001, phi_pol=np.pi/2)
+    det_par = detector.Detector(det_type='polarized', na=0.001, phi_pol=0)
+    m_perp = microscope.Microscope(illuminator=ill, detector=det_perp, max_photons=1e10)
+    m_par = microscope.Microscope(illuminator=ill, detector=det_par, max_photons=1e10)
+    I_perp = m_perp.calc_intensity((np.pi/2, phi), kappa=0.0, epsrel=1e-8)
+    I_par = m_par.calc_intensity((np.pi/2, phi), kappa=0.0, epsrel=1e-8)
+    r = (I_par - I_perp)/(I_par + 2*I_perp)
+    print(I_perp, I_par, r)
+    if I_par + 2*I_perp == 0:
+        print("*")
+        return 0.0
+    else:
+        return r
+
+f, ax = plt.subplots(1, 1, figsize=(5, 5))
+x = list(np.linspace(-np.pi, np.pi, 100))
+v_aniso = np.vectorize(anisotropy)
+r_vect = v_aniso(x) 
+ax.plot(x, r_vect, '-k')
+ax.set_xlabel('$\phi$')
+ax.set_ylabel('$r$')
+f.savefig('anisotropy.pdf', dpi=300)
+
+print('Total time: '+str(np.round(time.time() - start, 2)))
+os.system('say "done"')

### 2017-09-29 notes/2017-09-28-anisotropy/figures/schematic.py

@@ -1,0 +1,28 @@
+from dipsim import multiframe, util, detector, illuminator, microscope, util
+import numpy as np
+import matplotlib.pyplot as plt
+
+import os; import time; start = time.time(); print('Running...')
+
+print('Computing microscope: ')
+# y-polarized input from x optical axis
+ill = illuminator.Illuminator(illum_type='sheet', theta_optical_axis=np.pi/2, phi_pol=np.pi/2)
+# perp: x-polarized detection
+# par: y-polarized detection
+det_perp = detector.Detector(det_type='polarized', na=0.001, phi_pol=0)
+det_par = detector.Detector(det_type='polarized', na=0.001, phi_pol=np.pi/2)
+m_perp = microscope.Microscope(illuminator=ill, detector=det_perp, max_photons=1e10)
+m_par = microscope.Microscope(illuminator=ill, detector=det_par, max_photons=1e10)
+
+f, axs = plt.subplots(1, 2, figsize=(10, 5))
+util.draw_scene(m_perp.scene_string(), my_ax=axs[0], dpi=300)
+util.draw_scene(m_par.scene_string(), my_ax=axs[1], dpi=300)
+axs[0].annotate('$I_\perp$', xy=(0,0), xytext=(0.5, 1.05), textcoords='axes fraction',
+                va='bottom', ha='center', fontsize=14, annotation_clip=False)
+axs[1].annotate('$I_\parallel$', xy=(0,0), xytext=(0.5, 1.05), textcoords='axes fraction',
+            va='bottom', ha='center', fontsize=14, annotation_clip=False)
+
+f.savefig('schematic.pdf', dpi=300)
+
+print('Total time: '+str(np.round(time.time() - start, 2)))
+os.system('say "done"')

### 2017-10-03 notes/2017-10-03-split-aperture-coordinate-change/report/report.tex
Split aperture coordinate change algorithm.

\documentclass[11pt]{article}

%%%%%%%%%%%%
% Packages %
%%%%%%%%%%%%

\hyphenpenalty=10000
\usepackage{tocloft}
\renewcommand\cftsecleader{\cftdotfill{\cftdotsep}}
\def\undertilde#1{\mathord{\vtop{\ialign{##\crcr
$\hfil\displaystyle{#1}\hfil$\crcr\noalign{\kern1.5pt\nointerlineskip}
$\hfil\tilde{}\hfil$\crcr\noalign{\kern1.5pt}}}}}
\usepackage{cleveref}
\usepackage{xcolor}
\usepackage[colorlinks = true,
            linkcolor = blue,
            urlcolor  = blue,
            citecolor = blue,
            anchorcolor = blue]{hyperref}
\usepackage{epstopdf}
\usepackage{braket}
\usepackage{upgreek}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{amssymb,latexsym,amsmath,gensymb}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{float}
\usepackage{enumitem}
\usepackage{pdflscape}
\usepackage{url}
\usepackage{tikz, calc}
\usetikzlibrary{shapes.geometric, arrows, calc}
\tikzstyle{norm} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{arrow} = [thick, ->, >=stealth]
\newcommand{\argmin}{\arg\!\min}
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
\providecommand{\mb}[1]{\mathbf{#1}}
\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
\providecommand{\bs}[1]{\boldsymbol{#1}} 
\providecommand{\intinf}{\int_{-\infty}^{\infty}}
\providecommand{\fig}[4]{
  % filename, width, caption, label
\begin{figure}[h]
 \captionsetup{width=1.0\linewidth}
 \centering
 \includegraphics[width = #2\textwidth]{#1}
 \caption{#3}
 \label{fig:#4}
\end{figure}
}

\newcommand{\tensor}[1]{\overset{\text{\tiny$\leftrightarrow$}}{\mb{#1}}}
\newcommand{\tunderbrace}[2]{\underbrace{#1}_{\textstyle#2}}
\providecommand{\figs}[7]{
  % filename1, filename2, caption1, caption2, label1, label2, shift
\begin{figure}[H]
\centering
\begin{minipage}[b]{.45\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#1}
  \captionsetup{justification=justified, singlelinecheck=true}
  \caption{#3}
  \label{fig:#5}
\end{minipage}
\hspace{2em}
\begin{minipage}[b]{.45\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#2}
  \vspace{#7em}
  \captionsetup{justification=justified}
  \caption{#4}
  \label{fig:#6}
\end{minipage}
\end{figure}
}
\makeatletter

\providecommand{\code}[1]{
\begin{center}
\lstinputlisting{#1}
\end{center}
}

\newcommand{\crefrangeconjunction}{--}
%%%%%%%%%%%
% Spacing %
%%%%%%%%%%%
% Margins
\usepackage[
top    = 1.5cm,
bottom = 1.5cm,
left   = 1.5cm,
right  = 1.5cm]{geometry}

% Indents, paragraph space
%\usepackage{parskip}
\setlength{\parskip}{1.5ex}

% Section spacing
\usepackage{titlesec}
\titlespacing*{\title}
{0pt}{0ex}{0ex}
\titlespacing*{\section}
{0pt}{0ex}{0ex}
\titlespacing*{\subsection}
{0pt}{0ex}{0ex}
\titlespacing*{\subsubsection}
{0pt}{0ex}{0ex}

% Line spacing
\linespread{1.1}

%%%%%%%%%%%%
% Document %
%%%%%%%%%%%%
\begin{document}
\title{\vspace{-2.5em} Ad-Hoc Coordinate Change Algorithm For The Split-Aperture\\
  Polarization Microscope \vspace{-1em}} \author{Talon
  Chandler}% and Patrick La Rivi\`ere}
\date{\vspace{-1em}\today\vspace{-1em}}
\maketitle
\section{Introduction}
The split-aperture polarization microscope can measure the azimuth orientation
of a single dipole in two rotated views. These notes describe an algorithm for
converting noise-corrupted azimuth orientations in two rotated views into the
azimuth and inclination orientation. We define the coordinates, outline the
algorithm, then show the error in the estimate with noise-free and noise-corrupted
measurements.

\section{Coordinates}
\begin{figure*}[h]
 \captionsetup{width=1.0\linewidth}
 \centering
 \begin{subfigure}[t]{0.33\textwidth}
   \centering
   \includegraphics[width = 1.0\textwidth]{../figures/frame_a.pdf}
   \caption{$\theta$ and $\phi$ coordinate definitions.}
   \label{fig:frames_a}
 \end{subfigure}%
 ~
 \begin{subfigure}[t]{0.33\textwidth}
   \centering
   \includegraphics[width = 1.0\textwidth]{../figures/frame_b.pdf}
   \caption{$\phi'$ coordinate definition in the $+\alpha$ rotated frame.}
   \label{fig:frames_b}
 \end{subfigure}%
 ~
 \begin{subfigure}[t]{0.33\textwidth}
   \centering
   \includegraphics[width = 1.0\textwidth]{../figures/frame_c.pdf}
   \caption{$\phi''$ coordinate definition in the $-\alpha$ rotated frame.}
   \label{fig:frames_c}
 \end{subfigure}%
 \caption{Coordinate definitions.}
 \label{fig:frames}
\end{figure*}

Figure 1 shows one way of defining the inclination and azimuth angle in the
laboratory frame ($\theta$, $\phi$) and the azimuth angles in the rotated views
($\phi'$, $\phi''$). The two rotated frames are given by right-handed rotations
about the $+\mh{y}$ axis by an angle $\alpha$.

In
\href{https://github.com/talonchandler/dipsim/blob/master/notes/2017-06-09-rotated-frames/report/report.pdf}{previous
  notes} I have derived the relationships between the angles in Figure 1. We can
reuse those results and write:
\begin{align}
  \phi' &=
          \begin{cases}
            \arccos\left(\frac{\cos\alpha\cos\phi\sin\theta - \sin\alpha\cos\theta}{\sqrt{1 - (\sin\alpha\cos\phi\sin\theta + \cos\alpha\cos\theta)^2}}\right) \ \ \ &0 \leq \phi < \pi\\
            -\arccos\left(\frac{\cos\alpha\cos\phi\sin\theta - \sin\alpha\cos\theta}{\sqrt{1 - (\sin\alpha\cos\phi\sin\theta + \cos\alpha\cos\theta)^2}}\right) \ \ \ &-\pi \leq \phi < 0\\
            \end{cases}\\
  \phi'' &=
          \begin{cases}
            \arccos\left(\frac{\cos\alpha\cos\phi\sin\theta + \sin\alpha\cos\theta}{\sqrt{1 - (-\sin\alpha\cos\phi\sin\theta + \cos\alpha\cos\theta)^2}}\right) \ \ \ &0 \leq \phi < \pi\\
            -\arccos\left(\frac{\cos\alpha\cos\phi\sin\theta + \sin\alpha\cos\theta}{\sqrt{1 - (-\sin\alpha\cos\phi\sin\theta + \cos\alpha\cos\theta)^2}}\right) \ \ \ &-\pi \leq \phi < 0\\            
          \end{cases}.
\end{align}

\section{Procedure}
Equations 1 and 2 take $\theta$ and $\phi$ as input and give $\phi'$ and
$\phi''$ as output. We can rewrite equations 1 and 2 in operator form as
\begin{align}
  \mb{g} = \mathcal{H}\mb{f}
\end{align}
where $\mb{g} = [\phi', \phi'']$, $\mb{f} = [\theta, \phi]$, and $\mathcal{H}$
denotes the operation in equations 1 and 2. If there is noise introduced in the
measurement process then the model becomes
\begin{align}
  \mb{g} = \mathcal{H}\mb{f} + \mb{n}
\end{align}
where $\mb{n}$ is a noise vector.

Our task is to estimate $\mb{f}$ from noise-corrupted measurements $\mb{g}$. To
do so, we solve the following problem
\begin{align}
  \tilde{\mb{f}} = \argmin_{\mb{f}} ||\mb{g} - \mathcal{H}\mb{f}||_2 \label{eq:problem}
\end{align}
where $||\cdot||_2$ denotes the 2-norm. The search space of $\mb{f}$ is small
enough that a brute-force search is appropriate. First, we evaluate
$\mathcal{H}\mb{f}$ on a predetermined set of points to construct a lookup
table. When we take a measurement $\mb{g}$, we compute
$||\mb{g} - \mathcal{H}\mb{f}||_2$ at every point in the lookup table. Finally,
we choose the $\mb{f}$ that minimizes $||\mb{g} - \mathcal{H}\mb{f}||_2$ and
call that our estimate $\tilde{\mb{f}}$.

The samples used to construct the lookup table should be equally spaced if we
expect the lookup table procedure to introduce an equal amount of error into our
estimates. Note that $\mb{f}$ is a unit direction ($\mb{f}\in \mathbb{S}^2$) so
if we equally space the components of $\mb{f}$ ($\theta$ and $\phi$) most of the
sample points will be concentrated near the $\mb{z}$ axis. Instead of
constructing a lookup table with equally spaced $\theta$ and $\phi$ points, I
used approximately equally spaced points found using the
\href{http://stackoverflow.com/a/26127012/5854689}{Fibonnaci sphere algorithm}.

\section{Results}
I used $\alpha = 30^{\circ}$ for all tests.

To test the algorithm I chose 2000 equally spaced values of $\mb{f}$, calculated
$\mb{g} = \mathcal{H}\mb{f}$, then used the algorithm with a 500,000 point
lookup table to find $\tilde{\mb{f}}$. Finally, I found the error
$\mb{e} = |\tilde{\mb{f}} - \mb{f}|$ and plotted the individual components of
$\mb{e}$ in Figure 2.

The 2000 points I tested are not the same points as the lookup table points, so
the errors shown in Figure 2 are errors introduced by the lookup table
procedure. The error is different for different points because $\mathcal{H}$ is
nonlinear and because the points are varying distances from the lookup table
points. Each point in Figure 2 took $\sim 2$ ms to compute. We can trade speed
for accuracy in this algorithm by changing the number of points in the lookup
table.

I also performed the same test with noise-corrupted data. I chose 2000 equally
spaced values of $\mb{f}$, calculated $\mb{g} = \mathcal{H}\mb{f} + \mb{n}$
where $\mb{n}$ is a vector of Gaussian noise with $\mu = 0$ and
$\sigma = 2^\circ$, then used the algorithm with a 500,000 point lookup table to
find $\tilde{\mb{f}}$. Again, I found the error
$\mb{e} = |\tilde{\mb{f}} - \mb{f}|$ and plotted the individual components of
$\mb{e}$ in Figure 3.

\section{Limitations}
Using a precomputed lookup table will always introduce a varying amount of error
for different dipole orientations. If we want to have a constant error for all
dipole orientations, we could investigate iterative solutions that will stop
after a convergence criteria is reached.
\pagebreak

This model assumes that
\begin{itemize}[nosep]
\item both views use a narrow detection NA (no D-shaped aperture effects)
\item the azimuth angles measured from both views are equally weighted
\item the noise on the azimuth angles is normally distributed
\end{itemize}

\fig{../figures/recon-error.pdf}{.9}{Error introduced by the algorithm with
  noise-free data.}{results}

\fig{../figures/recon-error-gauss.pdf}{.9}{Error in the estimate with
  noise-corrupted data. I assumed that the two azimuth angle measurements had normally distributed noise with $\mu = 0$ and $\sigma = 2^{\circ}$.}{results}
\end{document}

### 2017-10-03 notes/2017-10-03-split-aperture-coordinate-change/figures/phiab2thetaphi.py

@@ -1,0 +1,82 @@
+import numpy as np
+import util
+import matplotlib.pyplot as plt
+import os
+
+def fibonacci_sphere(n):
+    # Returns "equally" spaced points on a unit sphere in spherical coordinates.
+    # http://stackoverflow.com/a/26127012/5854689
+    z = np.linspace(1 - 1/n, -1 + 1/n, num=n) 
+    theta = np.arccos(z)
+    phi = np.mod((np.pi*(3.0 - np.sqrt(5.0)))*np.arange(n), 2*np.pi) - np.pi
+    return np.vstack((theta, phi)).T
+
+def phi_prime(theta, phi, alpha=np.pi/6):
+    # Returns phi coordinate in a frame rotated by a right-handed rotation by
+    # angle alpha about the positive y-axis. 
+    if theta == 0:
+        return 0
+    num = np.cos(alpha)*np.cos(phi)*np.sin(theta) - np.sin(alpha)*np.cos(theta)
+    den = np.sqrt(1 - (np.sin(alpha)*np.cos(phi)*np.sin(theta) + np.cos(alpha)*np.cos(theta))**2)
+    if phi < np.pi and phi > 0:
+        return np.arccos(num/den)
+    elif phi < 0 and phi > -np.pi:
+        return -np.arccos(num/den)
+    else:
+        return 0
+
+def thetaphi2phiab(theta, phi, alpha=np.pi/6):
+    # Converts theta, phi coordinates into phi_a, phi_b coordinates
+    phi_a = phi_prime(theta, phi, alpha=alpha)
+    phi_b = phi_prime(theta, phi, alpha=-alpha)
+    return np.array([phi_a, phi_b])
+
+def generate_LUTs(n=1e4, alpha=np.pi/6):
+    # Generates LUTs in both coordinate systems
+    thetaphiLUT = fibonacci_sphere(n)
+    phiabLUT = []
+    for direction in thetaphiLUT:
+        phiabLUT.append(thetaphi2phiab(*direction, alpha=alpha))
+    phiabLUT = np.array(phiabLUT)
+    return thetaphiLUT, phiabLUT
+
+def phiab2thetaphi(phia, phib, thetaphiLUT=None, phiabLUT=None):
+    # Converts phi_a, phi_b coordinates into theta, phi using lookup tables.
+    if phiabLUT is None:
+        thetaphiLUT, phiabLUT = generate_LUTs()
+    target = np.array([phia, phib])
[…]

### 2017-10-03 notes/2017-10-03-split-aperture-coordinate-change/figures/util.py

@@ -1,0 +1,95 @@
+from dipsim import visuals
+import numpy as np
+import subprocess
+import matplotlib
+import matplotlib.pyplot as plt
+import matplotlib.image as mpimg
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+import vispy
+
+def plot_sphere(filename=None, directions=None, data=None, interact=False,
+                color_norm='linear', color_min=0, color_max=None,
+                gamma=0.25, color_map='coolwarm', linthresh=1e-3,
+                my_ax=None, my_cax=None, dpi=500, vis_px=1000,
+                save_file=False):
+
+    # Setup viewing window
+    vispy.use('glfw')
+    canvas = vispy.scene.SceneCanvas(keys='interactive', bgcolor='white',
+                               size=(vis_px, vis_px), show=interact)
+    my_cam = vispy.scene.cameras.turntable.TurntableCamera(fov=0, azimuth=135,
+                                                           scale_factor=2.05)
+    view = canvas.central_widget.add_view(camera=my_cam)
+
+    # Plot dots
+    dots = vispy.scene.visuals.Markers(parent=view.scene)
+    dots.antialias = False
+    dots.set_data(pos=np.array([[1.01,0,0],[0,1.01,0],[0,0,1.01]]),
+                  edge_color='black', face_color='black', size=vis_px/50)
+
+    if color_max == None:
+        color_max = np.max(data)
+    
+    # Calculate colors
+    if color_norm == 'linear':
+        norm = matplotlib.colors.Normalize(vmin=color_min, vmax=color_max)
+    elif color_norm == 'log':
+        norm = matplotlib.colors.LogNorm(vmin=color_min, vmax=color_max)
+    elif color_norm == 'linlog':
+        norm = matplotlib.colors.SymLogNorm(linthresh=linthresh, vmin=-color_max, vmax=color_max)
+    elif color_norm == 'power':
+        norm = matplotlib.colors.PowerNorm(gamma=gamma, vmin=data.min(), vmax=data.max())
+
+    norm_data = norm(data).data
+    norm_data = np.expand_dims(norm_data, 1)    
+    cmap = matplotlib.cm.get_cmap(color_map)
+    colors = np.apply_along_axis(cmap, 1, norm_data)
+    
+    # Plot sphere
+    sphere = visuals.MySphere(parent=view.scene, radius=1.0,
+                              directions=directions, colors=colors)
[…]

### 2017-10-16 notes/2017-10-03-split-aperture-coordinate-change/report/report.tex
Split aperture test data reconstruction.

@@
   Chandler}% and Patrick La Rivi\`ere}
-\date{\vspace{-1em}\today\vspace{-1em}}
+\date{\vspace{-1em}October 3, 2017\\(Last Updated: \today)\vspace{-1em}}
 \maketitle
@@
 algorithm, then show the error in the estimate with noise-free and noise-corrupted
-measurements.
+measurements. Finally, we test the algorithm on a set of single-molecule data.
 
@@
 
-\section{Results}
+\section{Accuracy Results}
 I used $\alpha = 30^{\circ}$ for all tests.
@@
   noise-corrupted data. I assumed that the two azimuth angle measurements had normally distributed noise with $\mu = 0$ and $\sigma = 2^{\circ}$.}{results}
+\pagebreak
+
+\section{Test Data Results}
+Tomomi supplied five frames of test data in a \texttt{.xlsx} file. Each frame
+had $\sim$40 single molecules with its x-y position and its azimuthal angle from
+each view.
+
+I used the algorithm detailed above to calculate the coordinates in the
+laboratory frame (see Figure 1a)) then plotted the orientations and positions of
+the molecules in Figures 4-8. The origin is marked by a dot in the top corner. The
+data was supplied without units, so the scale bar may be in units of $\mu$m or
+pixels.
+
+\fig{../test-analysis/output/frame01.pdf}{.4}{Frame 1.}{results}
+\fig{../test-analysis/output/frame02.pdf}{.4}{Frame 2.}{results}
+\fig{../test-analysis/output/frame03.pdf}{.4}{Frame 3.}{results}
+\fig{../test-analysis/output/frame04.pdf}{.4}{Frame 4.}{results}
+\fig{../test-analysis/output/frame05.pdf}{.4}{Frame 5.}{results}
+
 \end{document}

### 2017-10-16 notes/2017-10-03-split-aperture-coordinate-change/test-analysis/phiab2thetaphi.py

@@ -1,0 +1,87 @@
+import numpy as np
+import util
+import matplotlib.pyplot as plt
+import os
+
+def fibonacci_sphere(n):
+    # Returns "equally" spaced points on a unit sphere in spherical coordinates.
+    # http://stackoverflow.com/a/26127012/5854689
+    z = np.linspace(1 - 1/n, -1 + 1/n, num=n) 
+    theta = np.arccos(z)
+    phi = np.mod((np.pi*(3.0 - np.sqrt(5.0)))*np.arange(n), 2*np.pi) - np.pi
+    return np.vstack((theta, phi)).T
+
+def phi_prime(theta, phi, alpha=np.pi/6):
+    # Returns phi coordinate in a frame rotated by a right-handed rotation by
+    # angle alpha about the positive y-axis. 
+    if theta == 0:
+        return 0
+    num = np.cos(alpha)*np.cos(phi)*np.sin(theta) - np.sin(alpha)*np.cos(theta)
+    den = np.sqrt(1 - (np.sin(alpha)*np.cos(phi)*np.sin(theta) + np.cos(alpha)*np.cos(theta))**2)
+    if phi < np.pi and phi > 0:
+        return np.arccos(num/den)
+    elif phi < 0 and phi > -np.pi:
+        return -np.arccos(num/den)
+    else:
+        return 0
+
+def thetaphi2phiab(theta, phi, alpha=np.pi/6):
+    # Converts theta, phi coordinates into phi_a, phi_b coordinates
+    phi_a = phi_prime(theta, phi, alpha=alpha)
+    phi_b = phi_prime(theta, phi, alpha=-alpha)
+    return np.array([phi_a, phi_b])
+
+def generate_LUTs(n=1e4, alpha=np.pi/6):
+    # Generates LUTs in both coordinate systems
+    thetaphiLUT = fibonacci_sphere(n)
+    phiabLUT = []
+    for direction in thetaphiLUT:
+        phiabLUT.append(thetaphi2phiab(*direction, alpha=alpha))
+    phiabLUT = np.array(phiabLUT)
+    return thetaphiLUT, phiabLUT
+
+def phiab2thetaphi(phia, phib, thetaphiLUT=None, phiabLUT=None):
+    # Converts phi_a, phi_b coordinates into theta, phi using lookup tables.
+    if phiabLUT is None:
+        thetaphiLUT, phiabLUT = generate_LUTs()
+    target = np.array([phia, phib])
[…]

### 2017-10-16 notes/2017-10-03-split-aperture-coordinate-change/test-analysis/plotthetaphi.py

@@ -1,0 +1,22 @@
+import numpy as np
+import util
+import glob
+files = glob.glob("./output/*.csv")
+
+for file in files:
+    print(file)
+    data = np.genfromtxt(file, delimiter=',', skip_header=1,
+                         names=['R-view x (um)', 'R-view y (um)', 'Theta [z polar angle] (rad)', 'Phi [x azimuth angle] (rad)'])
+
+    # Create asymptote string
+    asy_string = ''
+    for entry in data:
+        temp_str = 'dip_arrow(x_pos, y_pos, theta, phi, 10, (0,0,0));'
+        temp_str = temp_str.replace('x_pos', str(entry['Rview_x_um']))
+        temp_str = temp_str.replace('y_pos', str(entry['Rview_y_um']))
+        temp_str = temp_str.replace('theta', str(entry['Theta_z_polar_angle_rad']))
+        temp_str = temp_str.replace('phi', str(entry['Phi_x_azimuth_angle_rad']))        
+        asy_string += temp_str
+    
+    # Generate asymptote plot
+    util.draw_scene(asy_string, filename=file.replace('.csv', '.pdf'), save_file=True)

### 2017-10-16 notes/2017-10-03-split-aperture-coordinate-change/test-analysis/util.py

@@ -1,0 +1,236 @@
+from dipsim import visuals
+import numpy as np
+import subprocess
+import matplotlib
+import matplotlib.pyplot as plt
+import matplotlib.image as mpimg
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+import vispy
+
+def plot_sphere(filename=None, directions=None, data=None, interact=False,
+                color_norm='linear', color_min=0, color_max=None,
+                gamma=0.25, color_map='coolwarm', linthresh=1e-3,
+                my_ax=None, my_cax=None, dpi=500, vis_px=1000,
+                save_file=False):
+
+    # Setup viewing window
+    vispy.use('glfw')
+    canvas = vispy.scene.SceneCanvas(keys='interactive', bgcolor='white',
+                               size=(vis_px, vis_px), show=interact)
+    my_cam = vispy.scene.cameras.turntable.TurntableCamera(fov=0, azimuth=135,
+                                                           scale_factor=2.05)
+    view = canvas.central_widget.add_view(camera=my_cam)
+
+    # Plot dots
+    dots = vispy.scene.visuals.Markers(parent=view.scene)
+    dots.antialias = False
+    dots.set_data(pos=np.array([[1.01,0,0],[0,1.01,0],[0,0,1.01]]),
+                  edge_color='black', face_color='black', size=vis_px/50)
+
+    if color_max == None:
+        color_max = np.max(data)
+    
+    # Calculate colors
+    if color_norm == 'linear':
+        norm = matplotlib.colors.Normalize(vmin=color_min, vmax=color_max)
+    elif color_norm == 'log':
+        norm = matplotlib.colors.LogNorm(vmin=color_min, vmax=color_max)
+    elif color_norm == 'linlog':
+        norm = matplotlib.colors.SymLogNorm(linthresh=linthresh, vmin=-color_max, vmax=color_max)
+    elif color_norm == 'power':
+        norm = matplotlib.colors.PowerNorm(gamma=gamma, vmin=data.min(), vmax=data.max())
+
+    norm_data = norm(data).data
+    norm_data = np.expand_dims(norm_data, 1)    
+    cmap = matplotlib.cm.get_cmap(color_map)
[…]

### 2017-10-25 dipsim/cythonlib/calc_efficiencies.pyx
Cython speedup for efficiency evals. Likelihood function evals.

@@ -1,0 +1,30 @@
+#cython: boundscheck=False, wraparound=False, nonecheck=False
+from libc.math cimport sin, cos, acos, sqrt, M_PI
+
+def cy_calc_collection_efficiency(double theta_lab, double phi_lab, double theta_optical_axis, double alpha):
+    cdef double theta_opt, A, B
+    theta_opt = acos(sin(theta_optical_axis)*cos(phi_lab)*sin(theta_lab) + cos(theta_optical_axis)*cos(theta_lab))
+    A = 0.25 - 0.375*cos(alpha) + 0.125*(cos(alpha)**3)
+    B = 0.1875*(cos(alpha) - (cos(alpha)**3))
+    return 2.0*(A + B*(sin(theta_opt)**2))
+
+def cy_calc_excitation_efficiency(double theta_lab, double phi_lab, double phi_pol, double theta_optical_axis):
+    cdef double theta_opt, num, den, acos_arg
+    theta_opt = acos(sin(theta_optical_axis)*cos(phi_lab)*sin(theta_lab) + cos(theta_optical_axis)*cos(theta_lab))
+    num = cos(theta_optical_axis)*cos(phi_lab)*sin(theta_lab) - sin(theta_optical_axis)*cos(theta_lab)
+    den = sqrt(1 - (sin(theta_optical_axis)*cos(phi_lab)*sin(theta_lab) + cos(theta_optical_axis)*cos(theta_lab))**2)
+    acos_arg = num/den
+
+    # Prevent nan from floating point errors
+    if acos_arg > 1.0:
+        acos_arg = 1.0
+    if acos_arg < -1.0:
+        acos_arg = -1.0
+
+    # Check sign
+    if phi_lab < M_PI and phi_lab > 0:
+        phi_opt = acos(acos_arg)
+    else:
+        phi_opt = -acos(acos_arg)
+        
+    return (sin(theta_opt)*cos(phi_opt - phi_pol))**2

### 2017-10-25 dipsim/cythonlib/setup.py

@@ -1,0 +1,13 @@
+from distutils.core import setup
+from distutils.extension import Extension
+from Cython.Distutils import build_ext
+
+ext_modules=[ Extension("calc_efficiencies",
+              ["calc_efficiencies.pyx"],
+              libraries=["m"],
+              extra_compile_args = ["-ffast-math"])]
+
+setup(
+  name = "calc_efficiencies",
+  cmdclass = {"build_ext": build_ext},
+  ext_modules = ext_modules)

### 2017-10-25 dipsim/microscope.py

@@ -5,6 +5,7 @@ import vispy
 from dipsim import util, visuals, detector, illuminator
 from dipsim import fluorophore as flu
 from scipy import integrate, special, interpolate
+from dipsim.cythonlib.calc_efficiencies import cy_calc_collection_efficiency, cy_calc_excitation_efficiency
 
 class Microscope:
     """
@@ -26,6 +27,7 @@ class Microscope:
         self.fluos_LUT = None
         self.excite_bspl = None
         self.collect_bspl = None
+        self.cython_flag = True
     
     def calc_intensity(self, fluorophore=flu.Fluorophore(), epsrel=1e-2):
         # Calculate the intensity measured by this microscope when a given
@@ -35,8 +37,12 @@ class Microscope:
             if self.precompute_flag:
                 excite, collect = self.efficiency_from_LUT(fluorophore)
             else:
-                excite = self.illuminator.calc_excitation_efficiency(fluorophore)
-                collect = self.detector.calc_collection_efficiency(fluorophore)
+                if self.cython_flag:
+                    excite = cy_calc_excitation_efficiency(fluorophore.theta, fluorophore.phi, self.illuminator.phi_pol, self.illuminator.theta_optical_axis)
+                    collect = cy_calc_collection_efficiency(fluorophore.theta, fluorophore.phi, self.detector.theta_optical_axis, self.detector.alpha)
+                else:
+                    excite = self.illuminator.calc_excitation_efficiency(fluorophore)
+                    collect = self.detector.calc_collection_efficiency(fluorophore)
             return self.max_photons*fluorophore.c*excite*collect
         else:
             # For fluorophore distributions (this function calls itself!)

### 2017-10-25 notes/2017-10-10-voxel-reconstruction/figures/plot-likelihood.py

@@ -5,10 +5,11 @@ import os; import time; start = time.time(); print('Running...')
 import matplotlib.gridspec as gridspec
 
 # Setup k and c sweep
-kappas = [10 , 100, np.inf]
+kappas = [-10, 0, 10, np.inf]
 cs = [0.1, 1, 2]
 col_labels = ['$\kappa$ = ' + str(x) for x in kappas]
 row_labels = ['c = ' + str(x) for x in cs]
+col_labels[-1] = '$\kappa = \infty$'
 
 # Generate axes
 inch_fig = 5
@@ -47,13 +48,13 @@ for i, kappa in enumerate(kappas):
         print(i, j)
         fluo = fluorophore.Fluorophore(theta=0.01, phi=0.01)
         data = exp.calc_intensities(fluo)
-        n_pts = 50
+        n_pts = 500
         directions = util.fibonacci_sphere(n_pts)
         input_points = np.hstack([directions, kappa*np.ones([n_pts, 1]), c*np.ones([n_pts, 1])])
         likelihoods = [exp.noise_model.loglikelihood(x, data=data) for x in input_points]
         util.plot_sphere(directions=directions, data=likelihoods,
                          my_ax=axs[j,i], my_cax=caxs[j,i],
-                         color_min=1e1, color_max=1e3)
+                         color_min=0, color_max=7000)
 
 # Label axes and save
 util.label_rows_and_cols(axs, row_labels, col_labels, row_pos=(-0.2, 0.5))

### 2017-10-25 notes/2017-10-25-group-meeting/group_meeting.tex
Oct 25 group meeting. Likelihood functions.

% Created 2017-10-25 Wed 12:35
\documentclass[presentation]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{amssymb}
\usepackage{hyperref}
\tolerance=1000
\usepackage{graphicx}
\usetheme{simple}
\usecolortheme{}
\usefonttheme{serif}
\useinnertheme{}
\useoutertheme{}
\author{Talon Chandler}
\date{October 25, 2017}
\title{Progress Report On 3D Orientation Determination}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs 25.3.1 (Org mode 8.2.10)}}
\begin{document}

\maketitle
\begin{frame}[label=sec-1]{Distributions of fluorophores}
\begin{center}
  \includegraphics[width=1.0\textwidth, interpolate=true]{figs/watson.pdf}\\
\end{center}
\end{frame}

\begin{frame}[label=sec-2]{Log-likelihood as a function of the estimate orientation ($\Theta$, $\Phi$)}
\begin{center}
True orientation: $\Theta = 0, \Phi = 0$\\
  \includegraphics[width=0.55\textwidth, interpolate=true]{figs/likelihood2.pdf}\\
\end{center}
\end{frame}

\begin{frame}[label=sec-3]{Log-likelihood as a function of the fluorophore distribution ($\Theta$, $\Phi$, $\kappa$, $c$)}
\begin{center}
True orientation: $\Theta = 0, \Phi = 0, \kappa = \infty, c = 1$\\
  \includegraphics[width=0.9\textwidth, interpolate=true]{figs/likelihood-dist.pdf}\\
\end{center}
\end{frame}

\begin{frame}[label=sec-4]{Log-likelihood as a function of the fluorophore distribution ($\Theta$, $\Phi$, $\kappa$, $c$)}
\begin{center}
True orientation: $\Theta = 0, \Phi = 0, \kappa = 10, c = 1$\\
  \includegraphics[width=0.9\textwidth, interpolate=true]{figs/likelihood-dist2.pdf}\\
\end{center}
\end{frame}
\begin{frame}[label=sec-5]{Discussion}
\begin{itemize}
\item $\Theta$ and $\Phi$ are bounded
\item $\kappa$ and $c$ are unbounded...estimate log($\kappa$) and log($c$) instead?
\end{itemize}
\end{frame}
% Emacs 25.3.1 (Org mode 8.2.10)
\end{document}

### 2017-11-08 paper/paper/paper.tex
Optics express reviewer comments and revisions.

@@
 \usepackage{booktabs}
+\usepackage{menukeys}
 \providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
@@
 \begin{document}
+\section*{Reviewer Comments and Author Responses}\hypertarget{test}{}
+
+This document shows the {\color{OE}reviewers' original comments}, our responses,
+and the complete paper marked up with the latest revisions. If this document is
+viewed as a .pdf, the reader can use the
+\hyperlink{test}{{\color{urlblue}links}} in each of our responses to view the
+section of the paper that has been changed. After viewing the part of the paper
+that was changed, the reader can use the back button ($\cmd + [$ in macOS
+Preview.app) to return to the comments and responses.
+
+\section*{Reviewer 1 Comments and Responses}
+
+{\color{OE} Chandler et al. present an exhaustive study of the potential of
+  polarization- dependent multiview microscopy for the determination of the
+  orientation of individual fluorophores (or any other nano-object with dipolar
+  absorption and emission characteristics). Based on earlier work by Fourkas
+  (Ref. 15), the authors establish a quantitative description of
+  single-fluorophore microscopy that is general enough to cover both
+  polarization-dependent excitation and detection, as well non-collinear
+  excitation and detection paths and specialized modalities such as light-sheet
+  illumination. The authors thus show that an astute choice of the experimental
+  configuration allows to determine the 3D orientation of the fluorophore with
+  high precision. The latter point is quantified with an information-theoretic
+  treatment of the solid-angle uncertainty and illustrated with simulations of
+  the resulting performance as a function of several experimental
+  parameters. The manuscript is rounded off with a clear discussion of the
+  implications for the experimenter, which also takes into account practical
+  experimental constraints such as space requirements and the availability of
+  high-NA objectives.}
+
+{\color{OE} The manuscript is well-written, thorough and clear, and the
+  illustrations are carefully designed to convey the most important conclusions
+  in graphical form.  Given the widespread interest in the accurate
+  determination of the orientation of single fluorophores, I would expect this
+  work to be interesting to many readers, as it carefully addresses both
+  theoretical and practical issues related to this topic. I myself have enjoyed
+  reading the manuscript and following the reasoning of the authors, and I can
+  wholeheartedly recommend publication of this
+  nice work in Optics Express.}
+
+{\color{OE} I have a few remarks that the authors may wish to take into account
+  to make some
+  (very minor) revisions:}
+
+{\color{OE} As far as I know, the correct way to cite the second author of Ref. 17 is D.
+A. Vanden Bout. ("Vanden" is part of his last name and should not be
+abbreviated.)}\\
+
+\hyperlink{vanden}{\color{urlblue} Corrected.} Thank you for pointing this out.\\
+
+{\color{OE} The rotation matrix of Eq. 6 is identical to the one given by
+  Fourkas in Ref.  15. This might surprise readers, because the matrix of
+  Fourkas describes the transformation of rays emanating from the focus and
+  being refracted by the objective towards the detection polarizer, while Eq. 6
+  deals with excitation light being focused, i.e. traveling the reverse optical
+  path as far as the objective is concerned. Making the corresponding
+  adjustments in Fourkas's treatment, I find that all $\sin\theta$ terms change their
+  sign, which affects four matrix elements in Eq. 6. However, using this
+  modified matrix results in the exact same of the expressions in Eq. 10, so
+  this issue has no practical consequences. It may nevertheless be worthwhile to
+  briefly comment on it in order to eliminate a potential source of confusion.}\\
+
+Excellent point. We've changed the signs of the $\sin\theta$ terms in Eq. 6 and \hyperlink{signs}{\color{urlblue} added a sentence} comparing our rotation matrix with Fourkas'.\\
+
+{\color{OE} The expressions for $A$, $B$ and $C$ in Eq. 10 all contain a factor
+  of (1-$\cos\alpha$), which can be canceled with the same term in the
+  denominator of $D$. I assume the authors have chosen the slightly more
+  complicated form of Eq. 10 for didactic purposes (explicitly showing the
+  irradiance-conserving factor) and to facilitate comparison with Fourkas's
+  results in the subsequent discussion, but it may be worthwhile to mention the
+  possibility of further simplification before starting
+  numerical calculations.}\\
+
+Very good point. We've \hyperlink{dfactor}{\color{urlblue} added a sentence}
+that mentions that these expressions can be simplified.\\
+
+{\color{OE} I think the authors raise an important point when they note that
+  Fourkas has incorrectly normalized his expressions for the detected
+  intensities. Maybe the authors can briefly comment on whether the underlying
+  issue is really only the choice of normalization constant or if this
+  discrepancy points to a more fundamental reasoning error that would require an
+  angle-dependent correction term.}\\
+
+Fourkas' issue is with incorrect normalization only---there is no
+angle-dependent correction required. We've \hyperlink{normalization}{\color{urlblue}
+  modified two sentences} to make this clearer.\\
+
+{\color{OE} If one sets $a=b=r$ in Eq. 19 then one does not recover the
+  $\gamma$-independent equation of a circle. It seems to me that $x$ and $y$
+  have to be exchanged in the
+  numerator of the second term of the left-hand side of Eq. 19.}\\
+
+Excellent catch. Thank you. We've \hyperlink{ellipse}{\color{urlblue}
+  corrected the equation} and confirmed that when
+$a=b=r$ we recover the equation of a circle. We also confirmed that this was a
+transcription error only, so it did not affect the ellipse plots throughout the
+paper.
+
+\section*{Reviewer 2 Comments and Responses}
+
+{\color{OE} The work by Chandler et al presents an elegant derivation of the
+  expressions that characterise the intensity of an single-molecule level
+  micrograph containing dipoles with varied orientations, obtained with
+  instruments offering multiple views of the same sample. This relates chiefly
+  to Light Sheet Microscopes.}
+
+{\color{OE} The derivations in paragraphs 2.2 and 2.3 are scholastic, although
+  they provide a correction to a previous work by Fourkas.}
+
+{\color{OE} Figures 4-6 provide the most interesting information to the
+  reader, as they describe the theoretical performance of distinct experimental
+  arrangements.}
+
+{\color{OE} Surprisingly, no mention is made of the temporal evolution of the
+  dipoles. In other words, while these results apply to a fixed condition, it
+  would be interesting to discuss a metric concerning the advantages, if any,
+  offered by multi view microscopes to capture the temporal dependence of dipole
+  orientation evolution.}\\
+
+We thank the reviewer for a valuable comment. We neglected to comment on the
+ability of our designs to image the temporal evolution of a dipole, so we have
+\hyperlink{temporal}{\color{urlblue} added several sentences} to address this
+omission. In particular, we highlight that the brightness of the fluorophores is
+the main factor that limits temporal resolution in these designs. We also
+mention that the comparison we carried out in the manuscript only compare microscopes
+with identical temporal resolution because we use a fixed sample irradiance.\\
+
+{\color{OE} While this work provides an interesting and novel perspective into
+  the problem, what is missing is an experimental validation of the author's
+  analytical derivations. This would make the manuscript truly valuable, by
+  providing a comparison with experiments where the orientation of the dipoles
+  can be fixed, or the rate of change be controlled (i.e. by controlling the
+  viscosity). In this respect, an obvious application would be steady state
+  anisotropy.}
+
+{\color{OE} We therefore recommend the authors to resubmit the manuscript
+  including an experimental section where they test their predictions on at
+  least one sample.}\\
+
+We thank the reviewer for the suggestion, but we think that including
+experimental results would significantly change the scope of this manuscript. We
+wrote this paper to serve as a design guide for those building multiview
+polarized light microscopes, and we don't think that experimental results would
+help the paper fulfill this role. Also, many of the experimental predictions
+made in this paper have already been predicted by Fourkas \cite{fourkas2001} and
+verified by others \cite{lu2008}. For these reasons, we have decided to resubmit
+the manuscript without experimental results.
+
+\section*{Pre-Production Review Comments and Responses}
+
+{\color{OE} Be sure to include and correct the year on your copyright line to
+  the current year.}\\
+
+\hyperlink{copyright}{\color{urlblue} Added.}\\
+
+{\color{OE} For book reference publishers with "Press" as part of publisher
+  name, e.g., "Cambridge U. Press," delete "Press" and spell out "University."}\\
+
+\hyperlink{press}{\color{urlblue} Corrected.}\\
+
+{\color{OE} Do not include the city in parentheses with the publisher and year
+  for book citations.}\\
+
+\hyperlink{wiley}{\color{urlblue} Changed ``Wiley New York'' to ``Wiley''.}\\
+
+{\color{OE} When you cite a part of a figure in the text, include parentheses
+  around the subfigure letter: Fig. 1(b).}\\
+
+Changed throughout and marked in blue.\\
+
+{\color{OE} Do not enclose your figure numbers in parentheses. Parentheses are
+  reserved for parts of a figure [Fig. 2(a)] and equations [Eq. (4)].}\\
+
+Changed throughout and marked in blue.\\
+
+{\color{OE} Abbreviate the single word "equation" to "Eq." and the plural word
+  "equations" to "Eqs." throughout the text, except for when either word appears
+  at the beginning of a sentence.}\\
+
+Changed throughout and marked in blue.\\
+
+{\color{OE} Include parentheses around the equation number in the text:
+  Eq. (2), or Eqs. (6)--(8).}\\
+
+Changed throughout.\\
+
+{\color{OE} When citing multiple equations in your text, it should be
+  formatted like this: Eqs. (1) and (2) or Eqs. (1)--(3).}\\
+
+Changed throughout.\\
+
+{\color{OE} Include the figure abbreviation each time you cite one in your
+  text. Instead of having "(a) and (b)", each time it should be "Fig. 2(a) and
+  2(b)".}\\
+
+Changed throughout and marked in blue.\\
+
+{\color{OE} If your paper has funding from the U.S. Department of Energy (DOE)
+  or any of its sub-agencies, you are required to deposit a copy of your
+  accepted manuscript to the DOE PAGES system. For more information, see
+  https://www.osti.gov/elink/am-submission.jsp.}\\
+
+No funding from the DOE.\\
+
+{\color{OE} Please confirm that all of your funding sources have been added
+  into your Prism submission in the Fees and Funding step.}\\
+
+Confirmed. \\
+
+{\color{OE} Please make sure to include all of your funding. See the style
+  guide for more details at
+  https://www.osapublishing.org/submit/templates/default.cfm}\\
+
+Revised the funding section to match the template and mentioned individual funding
+sources in the acknowledgments.
+
+\section*{Other Notes}
+
+\hyperlink{typo}{\color{urlblue} Corrected a typo} in the caption of Fig. 1 from $\hat{\mb{e}}_{\text{exc}}$ to
+$\hat{\mb{p}}_{\text{exc}}$ for consistency with the rest of the paper.
+
+\pagebreak
+
 \title{Single-fluorophore orientation determination with multiview polarized
@@
 %%%%%%%%%%%%%%%%%%% abstract and OCIS codes %%%%%%%%%%%%%%%%
-\begin{abstract*}
+\begin{abstract}
   We investigate the use of polarized illumination in multiview microscopes for
@@
   that asymmetric NA microscopes can benefit from exposure asymmetry.
-\end{abstract*}
+\end{abstract}
+\hypertarget{copyright}{}
 
@@
 \bibitem{lu2008}
-C.-Y. Lu and D.~A.~V. Bout, \enquote{Analysis of orientational dynamics of
+C.-Y. Lu and D.~A.~\hypertarget{vanden}{{\color{urlblue} Vanden}} Bout, \enquote{Analysis of orientational dynamics of
   single fluorophore trajectories from three-angle polarization experiments,}
@@
   T.~W. Anderson, \emph{An introduction to multivariate statistical analysis}
-  (Wiley New York, 1958).
+(\hypertarget{wiley}{{\color{urlblue}
+Wiley}}, 1958).
 
@@
 \bibitem{nov2006}
-L.~Novotny and B.~Hecht, \emph{Principles of Nano-Optics} (Cambridge University
-  Press, 2006).
+L.~Novotny and B.~Hecht, \emph{Principles of Nano-Optics} (\hypertarget{press}
+  {{\color{urlblue}
+Cambridge University}}, 2006).
 
@@
 
-\section{Introduction}
+\section{Introduction}\label{sec:intro}
 The orientation of single-molecule fluorescence transition dipoles is a valuable
@@
   $\hat{\bs{\mu}}_{\text{abs}}$, b) the polarizer transmission axis
-  $\hat{\mb{e}}_{\text{exc}}$, and c) the dummy integration vector
+   $\hat{\mb{p}}_{\text{exc}}$, and c) the dummy integration vector
   $\hat{\mb{r}}$. d) The scene consists of a single molecule at the focal point
   of an objective lens with a polarizer in the aperture plane.}
+\hypertarget{typo}{}
 \label{fig:coordinates}
@@
 $\hat{\bs{\mu}}_{\text{abs}}$ that we express in spherical coordinates $\Theta$
-and $\Phi$ (see Figure \ref{fig:coordinates}a)) as
+and $\Phi$ {\color{urlblue}[}see {\color{urlblue}Fig.} \ref{fig:coordinates}{\color{urlblue}(a)}{\color{urlblue}]} as
 \begin{align}
@@
 where $\phi_{\text{exc}}$ is the angle between the transmission axis of the
-polarizer and the positive $\mh{x}$ axis (see Figure
-\ref{fig:coordinates}b)). Finally, we place a spatially incoherent and spatially
+polarizer and the positive $\mh{x}$ axis {\color{urlblue}[}see {\color{urlblue}Fig.}
+\ref{fig:coordinates}{\color{urlblue}(b)}{\color{urlblue}]}. Finally, we place a spatially incoherent and spatially
 uniform light source (or its image) in the aperture plane to illuminate the
@@
 of angles that can be changed by adjusting the radius of the diaphragm in the
-aperture plane. We denote the set of illuminated angles by $\Omega$ (see Figure
-\ref{fig:coordinates}d)). If we use unit vectors $\hat{\mb{r}}$ expressed in
-spherical coordinates $\theta$ and $\phi$ (see Figure \ref{fig:coordinates}c))
+aperture plane. We denote the set of illuminated angles by $\Omega$ {\color{urlblue}[}see {\color{urlblue}Fig.}
+\ref{fig:coordinates}{\color{urlblue}(d)}{\color{urlblue}]}. If we use unit vectors $\hat{\mb{r}}$ expressed in
+spherical coordinates $\theta$ and $\phi$ {\color{urlblue}[}see {\color{urlblue}Fig.} \ref{fig:coordinates}{\color{urlblue}(c)}{\color{urlblue}]}
 as
@@
   \begin{align}
-  \mb{R}(\mh{r}) &= \begin{bmatrix} \cos\theta\cos^2\phi + \sin^2\phi & (\cos\theta -1)\sin\phi\cos\phi & -\sin\theta\cos\phi\\ (\cos\theta - 1)\sin\phi\cos\phi & \cos\theta\sin^2\phi + \cos^2\phi & -\sin\theta\sin\phi \\ \sin\theta\cos\phi& \sin\theta\sin\phi & \cos\theta \end{bmatrix} \label{eq:rot_mat}.
-  \end{align}
+  \mb{R}(\mh{r}) &= \begin{bmatrix} \cos\theta\cos^2\phi + \sin^2\phi & (\cos\theta -1)\sin\phi\cos\phi & \sin\theta\cos\phi\\ (\cos\theta - 1)\sin\phi\cos\phi & \cos\theta\sin^2\phi + \cos^2\phi & \sin\theta\sin\phi \\ -\sin\theta\cos\phi& -\sin\theta\sin\phi & \cos\theta \end{bmatrix} \label{eq:rot_mat}.
+  \end{align}
+  \hypertarget{signs}
+  {{\color{urlblue}Note the difference of signs between Eq.
+  (\ref{eq:rot_mat}) and the rotation matrix given by
+    Fourkas---our matrix rotates rays propagating towards the sample while
+    Fourkas' matrix rotates rays propagating away from the sample.}}
 
@@
 
-We substitute equations \ref{eq:mu}--\ref{eq:rot_mat} into equation
-\ref{eq:abs}, evaluate the integrals, and simplify to express the absorption
-efficiency in scalar notation as
+We substitute {\color{urlblue}Eqs.} (\ref{eq:mu})--(\ref{eq:rot_mat}) into
+{\color{urlblue}Eq.} (\ref{eq:abs}), evaluate the integrals, and simplify to
+express the absorption efficiency in scalar notation as
 \begin{align}
@@
 Note that the factor $D$ keeps the specimen irradiance constant for any value of
-$\alpha$.
+$\alpha$. \hypertarget{dfactor}{{\color{urlblue} We have kept $D$ as a
+    multiplicative factor in Eq. (\ref{eq:scalarabs}) to facilitate comparison
+    with Fourkas' results, but Eq. (\ref{eq:scalarabs}) can be simplified
+    further by distributing $D$.}}
 
@@
 was incident. Therefore, we can find the absorption efficiency of a dipole in
-this case by taking the limit of equation \ref{eq:scalarabs} as
-$\alpha \rightarrow 0$ (or by plugging equation \ref{eq:mu} and
-\ref{eq:excfield} into equation \ref{eq:singleray}) giving
+this case by taking the limit of {\color{urlblue}Eq.} (\ref{eq:scalarabs}) as
+$\alpha \rightarrow 0$ {\color{urlblue}[}or by plugging {\color{urlblue}Eqs.}
+(\ref{eq:mu}) and (\ref{eq:excfield}) into {\color{urlblue}Eq.}
+(\ref{eq:singleray}){\color{urlblue}]} giving
 \begin{align}
@@
 
-To calculate the detection efficiency without a polarizer, we use Equation
-\ref{eq:det_pol} and take the sum of polarized detection efficiencies with
+To calculate the detection efficiency without a polarizer, we use {\color{urlblue}Eq.}
+(\ref{eq:det_pol}) and take the sum of polarized detection efficiencies with
 orthogonal polarizer orientations to find
@@
 
-Note that $A$, $B$, and $C$ in equation \ref{eq:coefficients} are a factor of
-$\frac{3}{2}$ larger than the expressions given by Fourkas. We found that
-Fourkas' expressions were incorrectly normalized (the limit of
-$\eta_{\text{det}}$ as $\alpha\rightarrow \frac{\pi}{2}$ should be $\frac{1}{2}$), and
-the extra factor of $\frac{3}{2}$ corrects the error. This correction does not
-affect Fourkas' orientation reconstruction, but it does mean that Fourkas'
-algorithm underpredicts the total emitted intensity by a factor of
-$\frac{3}{2}$.
+Note that $A$, $B$, and $C$ in {\color{urlblue}Eq.} (\ref{eq:coefficients}) are a
+factor of $\frac{3}{2}$ larger than the expressions given by Fourkas. We found
+that Fourkas' expressions were incorrectly normalized (the limit of
+$\eta_{\text{det}}$ as $\alpha\rightarrow \frac{\pi}{2}$ should be
+$\frac{1}{2}$), and the extra factor of $\frac{3}{2}$ corrects the
+error. \hypertarget{normalization}{{\color{urlblue} The extra factor of
+    $\frac{3}{2}$ is a constant multiplicative factor, so the}} correction
+{\color{urlblue} is not angle-dependent which means that it} does not affect
+Fourkas' orientation reconstruction. {\color{urlblue}However, the correction}
+does mean that Fourkas' algorithm underpredicts the total emitted intensity by a
+factor of $\frac{3}{2}$.
 
@@
 where the proportionality depends on the exposure time, the molecular quantum
-efficiencies, and other constant factors. Equation \ref{eq:forward_frame} is the
+efficiencies, and other constant factors. Equation (\ref{eq:forward_frame}) is the
 forward model for a single intensity measurement. Figure \ref{fig:single-frame}
@@
   the detection solid angle, and the arrow indicates the transmission axis of
-  the illumination polarizer; 2) the absorption efficiency, see equation
-  \ref{eq:scalarabs}; 3) the detection efficiency, see equation
-  \ref{eq:scalardet}; (4) the total efficiency, the product of the absorption
+  the illumination polarizer; 2) the absorption efficiency, see {\color{urlblue}Eq.}
+  (\ref{eq:scalarabs}); 3) the detection efficiency, see {\color{urlblue}Eq.}
+  (\ref{eq:scalardet}); 4) the total efficiency, the product of the absorption
   and detection efficiencies.\newline \newline \textbf{Rows top to bottom:} 1)
@@
 with axis radii $a$ and $b$ and rotation angle $\gamma$ is
-\begin{align}
-  \frac{(x\cos\gamma + y\sin\gamma)^2}{a^2} + \frac{(x\cos\gamma - y\sin\gamma)^2}{b^2} =1. 
-\end{align}
+\hypertarget{ellipse}{{\color{urlblue} 
+\begin{align}
+  \frac{(x\cos\gamma + y\sin\gamma)^2}{a^2} + \frac{(x\sin\gamma - y\cos\gamma)^2}{b^2} =1. 
+\end{align}}}%
 We calculate the ellipse parameters in terms of the Fisher information matrix
@@
 \end{align}
-Equations \ref{eq:first}--\ref{eq:last} are identical to those found in
-\cite{coe2009} with an additional factor of $\sin\Theta$ on equations
-\ref{eq:second} and \ref{eq:third} to account for the non-Euclidean geometry of
+Equations (\ref{eq:first})--(\ref{eq:last}) are identical to those found in
+\cite{coe2009} with an additional factor of $\sin\Theta$ on {\color{urlblue}Eqs.}
+(\ref{eq:second}) and (\ref{eq:third}) to account for the non-Euclidean geometry of
 the sphere. Instead of plotting ellipses with widely varying radii on the same
@@
 this paper are identical under inversion of the dipole orientation
-($I(\Theta, \Phi) = I(\Theta + \pi, \Phi)$ in equations \ref{eq:scalarabs},
-\ref{eq:scalardet}, and \ref{eq:forward_frame}), but the solid-angle uncertainty
+{\color{urlblue}[}$I(\Theta, \Phi) = I(\Theta + \pi, \Phi)$ in {\color{urlblue}Eqs.} (\ref{eq:scalarabs}),
+(\ref{eq:scalardet}), and (\ref{eq:forward_frame}){\color{urlblue}]}, but the solid-angle uncertainty
 and the uncertainty ellipse give no indication of this degeneracy. For this
@@
 with a small illumination NA and a large detection NA where
-$\alpha \approx 65^{\circ}$ (see Figure \ref{fig:single-arm}e) at
-NA${}_{\text{det}} \approx 1.2$). A small illumination NA maximizes the range of
+$\alpha \approx 65^{\circ}$ {\color{urlblue}[}see {\color{urlblue}Fig.} \ref{fig:single-arm}{\color{urlblue}(e)} at
+NA${}_{\text{det}} \approx 1.2${\color{urlblue}]}. A small illumination NA maximizes the range of
 absorption efficiencies, while a large detection NA optimizes the number of
@@
 uncertainty much less than increasing the detection NA (an expensive
-modification requiring a higher-NA objective). Also note in Figure
-\ref{fig:single-arm}c) that single-view designs suffer from high orientation
+modification requiring a higher-NA objective). Also note in {\color{urlblue}Fig.}
+\ref{fig:single-arm}{\color{urlblue}(c)} that single-view designs suffer from high orientation
 uncertainty when the dipoles are oriented along the optical axis (the
@@
 (it is ambiguous whether the dipole is oriented above or below the transverse
-plane, see Figure \ref{fig:single-arm}b)), and near the polarizer orientations
-(it is ambiguous whether the dipole is on either side of the polarizer
-orientation, see Figure \ref{fig:single-arm}b)).
-
-The single-view microscope in Figure \ref{fig:single-arm} is fourfold degenerate
-($I(\Theta, \Phi) = I(\Theta + \pi, \Phi) = I(\Theta, \Phi + \pi) = I(\Theta +
-\pi, \Phi+\pi)$ in equations \ref{eq:scalarabs}, \ref{eq:scalardet}, and
-\ref{eq:forward_frame}). The first twofold degeneracy is due to the usual
+plane, see {\color{urlblue}Fig.} \ref{fig:single-arm}{\color{urlblue}(b)}, and near the polarizer orientations
+{\color{urlblue}[}it is ambiguous whether the dipole is on either side of the polarizer
+orientation, see {\color{urlblue}Fig.} \ref{fig:single-arm}{\color{urlblue}(b)}{\color{urlblue}]}.
+
+The single-view microscope in {\color{urlblue}Fig.} \ref{fig:single-arm} is
+fourfold degenerate
+{\color{urlblue}[}$I(\Theta, \Phi) = I(\Theta + \pi, \Phi) = I(\Theta, \Phi + \pi) = I(\Theta +
+\pi, \Phi+\pi)$ in {\color{urlblue}Eqs.} (\ref{eq:scalarabs}), (\ref{eq:scalardet}),
+and (\ref{eq:forward_frame}){\color{urlblue}]}. The first twofold degeneracy is due to the usual
 inversion degeneracy---the dipole is an axis rather than a direction. A second
@@
 distinguish between dipoles reflected through the transverse plane. Evidence of
-the second degeneracy is found in Figures \ref{fig:single-arm}b) and
-\ref{fig:single-arm}c) which show the ambiguity of whether the dipole is
-oriented above or below the transverse plane. Near the transverse plane this
-ambiguity is evident in the high solid-angle uncertainty, but far from the
-transverse plane evidence of the ambiguity is lost. Adding a second view and
-taking eight intensity measurements instead of four removes this degeneracy (see
-equations \ref{eq:thetap} and \ref{eq:phip}), so the dual-view microscopes we
-consider in this paper are only twofold degenerate. This means that while
-dual-view microscopes can find the orientation of a dipole up to the
-hemisphere, single-view microscopes can only find the orientation of a dipole
-up to the quadrant.
+the second degeneracy is found in {\color{urlblue}Figs.}
+\ref{fig:single-arm}{\color{urlblue}(b)} and
+\ref{fig:single-arm}{\color{urlblue}(c)} which show the ambiguity of whether the
+dipole is oriented above or below the transverse plane. Near the transverse
+plane this ambiguity is evident in the high solid-angle uncertainty, but far
+from the transverse plane evidence of the ambiguity is lost. Adding a second
+view and taking eight intensity measurements instead of four removes this
+degeneracy {\color{urlblue}[}see {\color{urlblue}Eqs.} (\ref{eq:thetap}) and (\ref{eq:phip}){\color{urlblue}]}, so the dual-view
+microscopes we consider in this paper are only twofold degenerate. This means
+that while dual-view microscopes can find the orientation of a dipole up to the
+hemisphere, single-view microscopes can only find the orientation of a dipole up
+to the quadrant.
 
@@
 detect from the other for four polarization orientations, then we repeat the
-polarization orientations with the same objectives in reversed roles. In Figure
+polarization orientations with the same objectives in reversed roles. In {\color{urlblue}Fig.}
 \ref{fig:symmetric-widefield} we sweep through the NA of both objectives and the
@@
 or against the cover slip. Given the high uncertainty in the transverse plane
-for a single-view shown in Figure \ref{fig:single-arm}c), we can see why oblique
+for a single-view shown in {\color{urlblue}Fig.} \ref{fig:single-arm}{\color{urlblue}(c)}), we can see why oblique
 views perform better than orthogonal views---oblique views are complementary
@@
 orthogonal design and varied the illumination and detection NA. Our results are
-similar to the single-view case in Figure \ref{fig:single-arm}. We find that
+similar to the single-view case in {\color{urlblue}Fig.} \ref{fig:single-arm}. We find that
 optimal designs use a small illumination NA and a large detection NA. As we
@@
 objectives would touch the cover slip and each other
-($\alpha_1 + \alpha_2 = 90^{\circ}$ in Figure \ref{fig:asymmetric-double}a)). We
-swept through the NA asymmetry and the sample-exposure asymmetry while keeping
-the sample irradiance constant, and found that symmetric designs are at a local
+{\color{urlblue}[}$\alpha_1 + \alpha_2 = 90^{\circ}$ in {\color{urlblue}Fig.}
+\ref{fig:asymmetric-double}{\color{urlblue}(a)}{\color{urlblue}]}. We swept
+through the NA asymmetry and the sample-exposure asymmetry while keeping the
+sample irradiance constant, and found that symmetric designs are at a local
 minimum of the median and MAD of the solid-angle uncertainty. We also found that
 at extreme asymmetries where the exposure from the low-NA arm is much larger
-than the exposure from the high-NA arm (top-left and bottom-right corners of
-Figures \ref{fig:asymmetric-double}d) and e)) the median and MAD are comparable
-to the symmetric light-sheet microscope. This means that if a very high-NA
-objective is available and a low-NA objective can provide light sheet
-illumination, this microscope will perform comparably with a dual-view
-microscope with two medium-NA objectives. Note that this asymmetric exposure
-configuration maximizes the use of the high-NA detection objective---if a
-high-NA objective is available it should be used for detection as much as
-possible. Figure \ref{fig:asymmetric-double} also shows that designs with
-slightly asymmetric NA can trade off a low solid-angle uncertainty median for a
-low solid-angle uncertainty MAD by changing the sample exposure ratio.
+than the exposure from the high-NA arm {\color{urlblue}[}top-left and
+bottom-right corners of {\color{urlblue}Figs.}
+\ref{fig:asymmetric-double}{\color{urlblue}(d)} and {\color{urlblue}6(e)]} the
+median and MAD are comparable to the symmetric light-sheet microscope. This
+means that if a very high-NA objective is available and a low-NA objective can
+provide light sheet illumination, this microscope will perform comparably with a
+dual-view microscope with two medium-NA objectives. Note that this asymmetric
+exposure configuration maximizes the use of the high-NA detection objective---if
+a high-NA objective is available it should be used for detection as much as
+possible. Figure \ref{fig:asymmetric-double} also shows that
+designs with slightly asymmetric NA can trade off a low solid-angle uncertainty
+median for a low solid-angle uncertainty MAD by changing the sample exposure
+ratio.
 
@@
 
+\hypertarget{temporal}{{\color{urlblue} We also chose to compare microscopes
+    with a fixed irradiance incident on the sample so that all of the
+    microscopes had the same temporal resolution. Given a fixed irradiance, the
+    main factor that limits the temporal resolution of an orientation microscope
+    is the brightness of the fluorophore (the rate of photon emission from the
+    fluorophore). The orientation of bright fluorophores can be determined more
+    quickly than dim fluorophores at the same level of accuracy because bright
+    fluorophores generate measurements with a higher signal-to-noise ratio.}}
+
 We have considered single- and dual-view microscopes in a variety of cases, but
@@
 \section*{Funding}
+National Institute of Health (NIH) (R01GM114274, R01EB017293).
+
+\section*{Acknowledgments}
 TC was supported by a University of Chicago Biological Sciences Division
-Graduate Fellowship. SM and RO were supported by NIH grant R01GM114274. HS and
-PL were supported by Marine Biological Laboratory Whitman Center Fellowships.
-HS was supported by the Intramural Research Programs of the National Institute
-of Biomedical Imaging and Bioengineering. PL was supported by NIH grant
-R01EB017293.
+Graduate Fellowship. HS and PL were supported by Marine Biological Laboratory
+Whitman Center Fellowships. HS was supported by the Intramural Research
+Programs of the National Institute of Biomedical Imaging and Bioengineering.
 

### 2017-11-09 paper/paper/paper.tex
Revised temporal resolution response. See Hari's Nov 9 email.

@@
 \usepackage{menukeys}
+\usepackage[normalem]{ulem}
 \providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
@@
 
-We thank the reviewer for a valuable comment. We neglected to comment on the
-ability of our designs to image the temporal evolution of a dipole, so we have
-\hyperlink{temporal}{\color{urlblue} added several sentences} to address this
-omission. In particular, we highlight that the brightness of the fluorophores is
-the main factor that limits temporal resolution in these designs. We also
-mention that the comparison we carried out in the manuscript only compare microscopes
-with identical temporal resolution because we use a fixed sample irradiance.\\
+We thank the reviewer for this valuable comment. We neglected to comment on the
+ability of our designs to image dipoles that are rotating during the measurement
+time, so we have \hyperlink{temporal}{\color{urlblue} added a paragraph} to
+address this omission. We highlight that the modeling in this work only applies
+to fixed dipoles, so estimating the orientation of rotating dipoles using these
+methods will, in general, lead to incorrect estimates. However, if the dipole
+rotates through a small solid angle during the measurement time, the models used
+in this paper will give a slightly biased estimate of the average orientation of
+the dipole. The precision results in this paper for fixed dipoles apply to
+rotating dipoles if the dipoles rotate through a small solid angle during the
+measurement time. If the dipoles rotate through a large solid angle during the
+measurement time, then further modeling is required to
+accurately estimate the dipole orientation.\\
 
@@
 
-\hypertarget{temporal}{{\color{urlblue} We also chose to compare microscopes
-    with a fixed irradiance incident on the sample so that all of the
-    microscopes had the same temporal resolution. Given a fixed irradiance, the
-    main factor that limits the temporal resolution of an orientation microscope
-    is the brightness of the fluorophore (the rate of photon emission from the
-    fluorophore). The orientation of bright fluorophores can be determined more
-    quickly than dim fluorophores at the same level of accuracy because bright
-    fluorophores generate measurements with a higher signal-to-noise ratio.}}
+% \hypertarget{temporal}{{\color{urlblue} We also chose to compare microscopes
+%     with a fixed irradiance incident on the sample so that all of the
+%     microscopes had the same temporal resolution. Given a fixed irradiance, the
+%     main factor that limits the temporal resolution of an orientation microscope
+%     is the brightness of the fluorophore (the rate of photon emission from the
+%     fluorophore). The orientation of bright fluorophores can be determined more
+%     quickly than dim fluorophores at the same level of accuracy because bright
+%     fluorophores generate measurements with a higher signal-to-noise ratio.}}
 
@@
 
-Although we have only considered the orientation determination of fixed single
-dipoles, polarized illumination techniques can be used to study the orientation
-distribution of ensembles of dynamic dipoles \cite{mehta2016, backer2016}. In
-future work we will investigate the ability of multiview polarized illumination
+\hypertarget{temporal}{{\color{urlblue} The models in this paper only consider
+    estimating the orientation of fixed dipoles, but real dipoles change their
+    orientation during measurement. Therefore, it is necessary to comment on the
+    applicability of our results to dynamic dipoles. In general, using the
+    models in this paper to estimate the orientation of rotating dipoles will
+    lead to incorrect results. However, if a dipole rotates through a small
+    solid angle during the measurement, the models in this paper will give a
+    precise and slightly biased estimate of the average orientation of the
+    dipole. The results in this paper apply to the precision of our estimates
+    for the average dipole orientation of rotating dipoles, but the direction
+    and size of the bias has not been characterized in detail. In future work we
+    will model dynamic dipoles and attempt to estimate the average dipole
+    orientation and the dipole's orientation distribution. Single-view polarized
+    illumination techniques have already been used to study the orientation
+    distribution of ensembles of dynamic dipoles \cite{mehta2016, backer2016},
+    but multiview polarized illumination microscopes have not been investigated
+    for this purpose.}}
+
+{\color{red}\sout{Although we have only considered the orientation determination
+    of fixed single dipoles, polarized illumination techniques can be used to
+    study the orientation distribution of ensembles of dynamic dipole. In future
+    work we will investigate the ability of multiview polarized illumination
     microscopes to estimate the orientation distribution of ensembles of dynamic
-dipoles.
+    dipoles.}}
 

### 2017-11-13 paper/paper/paper.tex
Patrick's edits to responses.

@@
 in this paper will give a slightly biased estimate of the average orientation of
-the dipole. The precision results in this paper for fixed dipoles apply to
-rotating dipoles if the dipoles rotate through a small solid angle during the
-measurement time. If the dipoles rotate through a large solid angle during the
+the dipole. If the dipoles rotate through a large solid angle during the
 measurement time, then further modeling is required to
@@
     solid angle during the measurement, the models in this paper will give a
-    precise and slightly biased estimate of the average orientation of the
-    dipole. The results in this paper apply to the precision of our estimates
-    for the average dipole orientation of rotating dipoles, but the direction
-    and size of the bias has not been characterized in detail. In future work we
-    will model dynamic dipoles and attempt to estimate the average dipole
-    orientation and the dipole's orientation distribution. Single-view polarized
-    illumination techniques have already been used to study the orientation
-    distribution of ensembles of dynamic dipoles \cite{mehta2016, backer2016},
-    but multiview polarized illumination microscopes have not been investigated
-    for this purpose.}}
+    slightly biased estimate of the average orientation of the dipole. The
+    direction and size of the bias has not been characterized in detail. In
+    future work we will model dynamic dipoles and attempt to estimate the
+    average dipole orientation and the dipole's orientation
+    distribution. Single-view polarized illumination techniques have already
+    been used to study the orientation distribution of ensembles of dynamic
+    dipoles \cite{mehta2016, backer2016}, but multiview polarized illumination
+    microscopes have not been investigated for this purpose.}}
 

### 2017-11-13 paper/paper/paper.tex
Remove comments and redlines. Submission copy.

@@
 \begin{document}
-\section*{Reviewer Comments and Author Responses}\hypertarget{test}{}
-
-This document shows the {\color{OE}reviewers' original comments}, our responses,
-and the complete paper marked up with the latest revisions. If this document is
-viewed as a .pdf, the reader can use the
-\hyperlink{test}{{\color{urlblue}links}} in each of our responses to view the
-section of the paper that has been changed. After viewing the part of the paper
-that was changed, the reader can use the back button ($\cmd + [$ in macOS
-Preview.app) to return to the comments and responses.
-
-\section*{Reviewer 1 Comments and Responses}
-
-{\color{OE} Chandler et al. present an exhaustive study of the potential of
-  polarization- dependent multiview microscopy for the determination of the
-  orientation of individual fluorophores (or any other nano-object with dipolar
-  absorption and emission characteristics). Based on earlier work by Fourkas
-  (Ref. 15), the authors establish a quantitative description of
-  single-fluorophore microscopy that is general enough to cover both
-  polarization-dependent excitation and detection, as well non-collinear
-  excitation and detection paths and specialized modalities such as light-sheet
-  illumination. The authors thus show that an astute choice of the experimental
-  configuration allows to determine the 3D orientation of the fluorophore with
-  high precision. The latter point is quantified with an information-theoretic
-  treatment of the solid-angle uncertainty and illustrated with simulations of
-  the resulting performance as a function of several experimental
-  parameters. The manuscript is rounded off with a clear discussion of the
-  implications for the experimenter, which also takes into account practical
-  experimental constraints such as space requirements and the availability of
-  high-NA objectives.}
-
-{\color{OE} The manuscript is well-written, thorough and clear, and the
-  illustrations are carefully designed to convey the most important conclusions
-  in graphical form.  Given the widespread interest in the accurate
-  determination of the orientation of single fluorophores, I would expect this
-  work to be interesting to many readers, as it carefully addresses both
-  theoretical and practical issues related to this topic. I myself have enjoyed
-  reading the manuscript and following the reasoning of the authors, and I can
-  wholeheartedly recommend publication of this
-  nice work in Optics Express.}
-
-{\color{OE} I have a few remarks that the authors may wish to take into account
-  to make some
-  (very minor) revisions:}
-
-{\color{OE} As far as I know, the correct way to cite the second author of Ref. 17 is D.
-A. Vanden Bout. ("Vanden" is part of his last name and should not be
-abbreviated.)}\\
-
-\hyperlink{vanden}{\color{urlblue} Corrected.} Thank you for pointing this out.\\
-
-{\color{OE} The rotation matrix of Eq. 6 is identical to the one given by
-  Fourkas in Ref.  15. This might surprise readers, because the matrix of
-  Fourkas describes the transformation of rays emanating from the focus and
-  being refracted by the objective towards the detection polarizer, while Eq. 6
-  deals with excitation light being focused, i.e. traveling the reverse optical
-  path as far as the objective is concerned. Making the corresponding
-  adjustments in Fourkas's treatment, I find that all $\sin\theta$ terms change their
-  sign, which affects four matrix elements in Eq. 6. However, using this
-  modified matrix results in the exact same of the expressions in Eq. 10, so
-  this issue has no practical consequences. It may nevertheless be worthwhile to
-  briefly comment on it in order to eliminate a potential source of confusion.}\\
-
-Excellent point. We've changed the signs of the $\sin\theta$ terms in Eq. 6 and \hyperlink{signs}{\color{urlblue} added a sentence} comparing our rotation matrix with Fourkas'.\\
-
-{\color{OE} The expressions for $A$, $B$ and $C$ in Eq. 10 all contain a factor
-  of (1-$\cos\alpha$), which can be canceled with the same term in the
-  denominator of $D$. I assume the authors have chosen the slightly more
-  complicated form of Eq. 10 for didactic purposes (explicitly showing the
-  irradiance-conserving factor) and to facilitate comparison with Fourkas's
-  results in the subsequent discussion, but it may be worthwhile to mention the
-  possibility of further simplification before starting
-  numerical calculations.}\\
-
-Very good point. We've \hyperlink{dfactor}{\color{urlblue} added a sentence}
-that mentions that these expressions can be simplified.\\
-
-{\color{OE} I think the authors raise an important point when they note that
-  Fourkas has incorrectly normalized his expressions for the detected
-  intensities. Maybe the authors can briefly comment on whether the underlying
-  issue is really only the choice of normalization constant or if this
-  discrepancy points to a more fundamental reasoning error that would require an
-  angle-dependent correction term.}\\
-
-Fourkas' issue is with incorrect normalization only---there is no
-angle-dependent correction required. We've \hyperlink{normalization}{\color{urlblue}
-  modified two sentences} to make this clearer.\\
-
-{\color{OE} If one sets $a=b=r$ in Eq. 19 then one does not recover the
-  $\gamma$-independent equation of a circle. It seems to me that $x$ and $y$
-  have to be exchanged in the
-  numerator of the second term of the left-hand side of Eq. 19.}\\
-
-Excellent catch. Thank you. We've \hyperlink{ellipse}{\color{urlblue}
-  corrected the equation} and confirmed that when
-$a=b=r$ we recover the equation of a circle. We also confirmed that this was a
-transcription error only, so it did not affect the ellipse plots throughout the
-paper.
-
-\section*{Reviewer 2 Comments and Responses}
-
-{\color{OE} The work by Chandler et al presents an elegant derivation of the
-  expressions that characterise the intensity of an single-molecule level
-  micrograph containing dipoles with varied orientations, obtained with
-  instruments offering multiple views of the same sample. This relates chiefly
-  to Light Sheet Microscopes.}
-
-{\color{OE} The derivations in paragraphs 2.2 and 2.3 are scholastic, although
-  they provide a correction to a previous work by Fourkas.}
-
-{\color{OE} Figures 4-6 provide the most interesting information to the
-  reader, as they describe the theoretical performance of distinct experimental
-  arrangements.}
-
-{\color{OE} Surprisingly, no mention is made of the temporal evolution of the
-  dipoles. In other words, while these results apply to a fixed condition, it
-  would be interesting to discuss a metric concerning the advantages, if any,
-  offered by multi view microscopes to capture the temporal dependence of dipole
-  orientation evolution.}\\
-
-We thank the reviewer for this valuable comment. We neglected to comment on the
-ability of our designs to image dipoles that are rotating during the measurement
-time, so we have \hyperlink{temporal}{\color{urlblue} added a paragraph} to
-address this omission. We highlight that the modeling in this work only applies
-to fixed dipoles, so estimating the orientation of rotating dipoles using these
-methods will, in general, lead to incorrect estimates. However, if the dipole
-rotates through a small solid angle during the measurement time, the models used
-in this paper will give a slightly biased estimate of the average orientation of
-the dipole. If the dipoles rotate through a large solid angle during the
-measurement time, then further modeling is required to
-accurately estimate the dipole orientation.\\
-
-{\color{OE} While this work provides an interesting and novel perspective into
-  the problem, what is missing is an experimental validation of the author's
-  analytical derivations. This would make the manuscript truly valuable, by
-  providing a comparison with experiments where the orientation of the dipoles
-  can be fixed, or the rate of change be controlled (i.e. by controlling the
-  viscosity). In this respect, an obvious application would be steady state
-  anisotropy.}
-
-{\color{OE} We therefore recommend the authors to resubmit the manuscript
-  including an experimental section where they test their predictions on at
-  least one sample.}\\
-
-We thank the reviewer for the suggestion, but we think that including
-experimental results would significantly change the scope of this manuscript. We
-wrote this paper to serve as a design guide for those building multiview
-polarized light microscopes, and we don't think that experimental results would
-help the paper fulfill this role. Also, many of the experimental predictions
-made in this paper have already been predicted by Fourkas \cite{fourkas2001} and
-verified by others \cite{lu2008}. For these reasons, we have decided to resubmit
-the manuscript without experimental results.
-
-\section*{Pre-Production Review Comments and Responses}
-
-{\color{OE} Be sure to include and correct the year on your copyright line to
-  the current year.}\\
-
-\hyperlink{copyright}{\color{urlblue} Added.}\\
-
-{\color{OE} For book reference publishers with "Press" as part of publisher
-  name, e.g., "Cambridge U. Press," delete "Press" and spell out "University."}\\
-
-\hyperlink{press}{\color{urlblue} Corrected.}\\
-
-{\color{OE} Do not include the city in parentheses with the publisher and year
-  for book citations.}\\
-
-\hyperlink{wiley}{\color{urlblue} Changed ``Wiley New York'' to ``Wiley''.}\\
-
-{\color{OE} When you cite a part of a figure in the text, include parentheses
-  around the subfigure letter: Fig. 1(b).}\\
-
-Changed throughout and marked in blue.\\
-
-{\color{OE} Do not enclose your figure numbers in parentheses. Parentheses are
-  reserved for parts of a figure [Fig. 2(a)] and equations [Eq. (4)].}\\
-
-Changed throughout and marked in blue.\\
-
-{\color{OE} Abbreviate the single word "equation" to "Eq." and the plural word
-  "equations" to "Eqs." throughout the text, except for when either word appears
-  at the beginning of a sentence.}\\
-
-Changed throughout and marked in blue.\\
-
-{\color{OE} Include parentheses around the equation number in the text:
-  Eq. (2), or Eqs. (6)--(8).}\\
-
-Changed throughout.\\
-
-{\color{OE} When citing multiple equations in your text, it should be
-  formatted like this: Eqs. (1) and (2) or Eqs. (1)--(3).}\\
-
-Changed throughout.\\
-
-{\color{OE} Include the figure abbreviation each time you cite one in your
-  text. Instead of having "(a) and (b)", each time it should be "Fig. 2(a) and
-  2(b)".}\\
-
-Changed throughout and marked in blue.\\
-
-{\color{OE} If your paper has funding from the U.S. Department of Energy (DOE)
-  or any of its sub-agencies, you are required to deposit a copy of your
-  accepted manuscript to the DOE PAGES system. For more information, see
-  https://www.osti.gov/elink/am-submission.jsp.}\\
-
-No funding from the DOE.\\
-
-{\color{OE} Please confirm that all of your funding sources have been added
-  into your Prism submission in the Fees and Funding step.}\\
-
-Confirmed. \\
-
-{\color{OE} Please make sure to include all of your funding. See the style
-  guide for more details at
-  https://www.osapublishing.org/submit/templates/default.cfm}\\
-
-Revised the funding section to match the template and mentioned individual funding
-sources in the acknowledgments.
-
-\section*{Other Notes}
-
-\hyperlink{typo}{\color{urlblue} Corrected a typo} in the caption of Fig. 1 from $\hat{\mb{e}}_{\text{exc}}$ to
-$\hat{\mb{p}}_{\text{exc}}$ for consistency with the rest of the paper.
-
-\pagebreak
-
 \title{Single-fluorophore orientation determination with multiview polarized
@@
 \end{abstract}
-\hypertarget{copyright}{}
 
@@
 \bibitem{lu2008}
-C.-Y. Lu and D.~A.~\hypertarget{vanden}{{\color{urlblue} Vanden}} Bout, \enquote{Analysis of orientational dynamics of
+C.-Y. Lu and D.~A.~Vanden Bout, \enquote{Analysis of orientational dynamics of
   single fluorophore trajectories from three-angle polarization experiments,}
@@
   T.~W. Anderson, \emph{An introduction to multivariate statistical analysis}
-(\hypertarget{wiley}{{\color{urlblue}
-Wiley}}, 1958).
+(Wiley, 1958).
 
@@
 \bibitem{nov2006}
-L.~Novotny and B.~Hecht, \emph{Principles of Nano-Optics} (\hypertarget{press}
-  {{\color{urlblue}
-Cambridge University}}, 2006).
+L.~Novotny and B.~Hecht, \emph{Principles of Nano-Optics} (Cambridge University, 2006).
 
@@
 $\hat{\bs{\mu}}_{\text{abs}}$ that we express in spherical coordinates $\Theta$
-and $\Phi$ {\color{urlblue}[}see {\color{urlblue}Fig.} \ref{fig:coordinates}{\color{urlblue}(a)}{\color{urlblue}]} as
+and $\Phi$ [see Fig. \ref{fig:coordinates}(a)] as
 \begin{align}
@@
 where $\phi_{\text{exc}}$ is the angle between the transmission axis of the
-polarizer and the positive $\mh{x}$ axis {\color{urlblue}[}see {\color{urlblue}Fig.}
-\ref{fig:coordinates}{\color{urlblue}(b)}{\color{urlblue}]}. Finally, we place a spatially incoherent and spatially
-uniform light source (or its image) in the aperture plane to illuminate the
-focal plane. In this geometry each point in the sample is illuminated by a range
-of angles that can be changed by adjusting the radius of the diaphragm in the
-aperture plane. We denote the set of illuminated angles by $\Omega$ {\color{urlblue}[}see {\color{urlblue}Fig.}
-\ref{fig:coordinates}{\color{urlblue}(d)}{\color{urlblue}]}. If we use unit vectors $\hat{\mb{r}}$ expressed in
-spherical coordinates $\theta$ and $\phi$ {\color{urlblue}[}see {\color{urlblue}Fig.} \ref{fig:coordinates}{\color{urlblue}(c)}{\color{urlblue}]}
-as
+polarizer and the positive $\mh{x}$ axis [see Fig.
+\ref{fig:coordinates}(b)]. Finally, we place a spatially incoherent and
+spatially uniform light source (or its image) in the aperture plane to
+illuminate the focal plane. In this geometry each point in the sample is
+illuminated by a range of angles that can be changed by adjusting the radius of
+the diaphragm in the aperture plane. We denote the set of illuminated angles by
+$\Omega$ [see Fig. \ref{fig:coordinates}(d)]. If we use unit vectors
+$\hat{\mb{r}}$ expressed in spherical coordinates $\theta$ and $\phi$ [see
+Fig. \ref{fig:coordinates}{(c)] as
   \begin{align}
@@
   \end{align}
-  \hypertarget{signs}
-  {{\color{urlblue}Note the difference of signs between Eq.
+  Note the difference of signs between Eq.
   (\ref{eq:rot_mat}) and the rotation matrix given by
     Fourkas---our matrix rotates rays propagating towards the sample while
-    Fourkas' matrix rotates rays propagating away from the sample.}}
+    Fourkas' matrix rotates rays propagating away from the sample.
 
@@
 
-We substitute {\color{urlblue}Eqs.} (\ref{eq:mu})--(\ref{eq:rot_mat}) into
-{\color{urlblue}Eq.} (\ref{eq:abs}), evaluate the integrals, and simplify to
+We substitute Eqs. (\ref{eq:mu})--(\ref{eq:rot_mat}) into
+Eq. (\ref{eq:abs}), evaluate the integrals, and simplify to
 express the absorption efficiency in scalar notation as
@@
 Note that the factor $D$ keeps the specimen irradiance constant for any value of
-$\alpha$. \hypertarget{dfactor}{{\color{urlblue} We have kept $D$ as a
-    multiplicative factor in Eq. (\ref{eq:scalarabs}) to facilitate comparison
-    with Fourkas' results, but Eq. (\ref{eq:scalarabs}) can be simplified
-    further by distributing $D$.}}
+$\alpha$. We have kept $D$ as a multiplicative factor in
+Eq. (\ref{eq:scalarabs}) to facilitate comparison with Fourkas' results, but
+Eq. (\ref{eq:scalarabs}) can be simplified further by distributing $D$.
 
@@
 was incident. Therefore, we can find the absorption efficiency of a dipole in
-this case by taking the limit of {\color{urlblue}Eq.} (\ref{eq:scalarabs}) as
-$\alpha \rightarrow 0$ {\color{urlblue}[}or by plugging {\color{urlblue}Eqs.}
-(\ref{eq:mu}) and (\ref{eq:excfield}) into {\color{urlblue}Eq.}
-(\ref{eq:singleray}){\color{urlblue}]} giving
+this case by taking the limit of Eq. (\ref{eq:scalarabs}) as
+$\alpha \rightarrow 0$ [or by plugging Eqs. (\ref{eq:mu}) and
+(\ref{eq:excfield}) into Eq. (\ref{eq:singleray})] giving
 \begin{align}
@@
 
-To calculate the detection efficiency without a polarizer, we use {\color{urlblue}Eq.}
+To calculate the detection efficiency without a polarizer, we use Eq.
 (\ref{eq:det_pol}) and take the sum of polarized detection efficiencies with
@@
 
-Note that $A$, $B$, and $C$ in {\color{urlblue}Eq.} (\ref{eq:coefficients}) are a
-factor of $\frac{3}{2}$ larger than the expressions given by Fourkas. We found
-that Fourkas' expressions were incorrectly normalized (the limit of
+Note that $A$, $B$, and $C$ in Eq. (\ref{eq:coefficients}) are a factor of
+$\frac{3}{2}$ larger than the expressions given by Fourkas. We found that
+Fourkas' expressions were incorrectly normalized (the limit of
 $\eta_{\text{det}}$ as $\alpha\rightarrow \frac{\pi}{2}$ should be
-$\frac{1}{2}$), and the extra factor of $\frac{3}{2}$ corrects the
-error. \hypertarget{normalization}{{\color{urlblue} The extra factor of
-    $\frac{3}{2}$ is a constant multiplicative factor, so the}} correction
-{\color{urlblue} is not angle-dependent which means that it} does not affect
-Fourkas' orientation reconstruction. {\color{urlblue}However, the correction}
-does mean that Fourkas' algorithm underpredicts the total emitted intensity by a
-factor of $\frac{3}{2}$.
+$\frac{1}{2}$), and the extra factor of $\frac{3}{2}$ corrects the error. The
+extra factor of $\frac{3}{2}$ is a constant multiplicative factor, so the
+correction is not angle-dependent which means that it does not affect Fourkas'
+orientation reconstruction. However, the correction does mean that Fourkas'
+algorithm underpredicts the total emitted intensity by a factor of
+$\frac{3}{2}$.
 
@@
 \caption{Representative examples of single intensity measurements. Black dots
-  indicate where the Cartesian unit vectors intersect the unit sphere. \newline \newline \textbf{Columns left to right:} 1) schematics where
-  the solid line encloses the illumination solid angle, the dashed line encloses
-  the detection solid angle, and the arrow indicates the transmission axis of
-  the illumination polarizer; 2) the absorption efficiency, see {\color{urlblue}Eq.}
-  (\ref{eq:scalarabs}); 3) the detection efficiency, see {\color{urlblue}Eq.}
-  (\ref{eq:scalardet}); 4) the total efficiency, the product of the absorption
-  and detection efficiencies.\newline \newline \textbf{Rows top to bottom:} 1)
-  coincident illumination ($\text{NA} = 1.1$ with $x$-polarized light) and
-  detection (NA = 1.1); 2) non-coincident orthogonal illumination
-  ($\text{NA} = 0.8$) and detection ($\text{NA} = 0.8$); 3) non-coincident
-  135${}^{\circ}$-separated illumination ($\text{NA} = 0.8$) and detection
-  ($\text{NA} = 0.8$). All simulations use $n=1.33$.}
+  indicate where the Cartesian unit vectors intersect the unit sphere. \newline
+  \newline \textbf{Columns left to right:} 1) schematics where the solid line
+  encloses the illumination solid angle, the dashed line encloses the detection
+  solid angle, and the arrow indicates the transmission axis of the illumination
+  polarizer; 2) the absorption efficiency, see Eq.  (\ref{eq:scalarabs}); 3) the
+  detection efficiency, see Eq.  (\ref{eq:scalardet}); 4) the total efficiency,
+  the product of the absorption and detection efficiencies.\newline \newline
+  \textbf{Rows top to bottom:} 1) coincident illumination ($\text{NA} = 1.1$
+  with $x$-polarized light) and detection (NA = 1.1); 2) non-coincident
+  orthogonal illumination ($\text{NA} = 0.8$) and detection ($\text{NA} = 0.8$);
+  3) non-coincident 135${}^{\circ}$-separated illumination ($\text{NA} = 0.8$)
+  and detection ($\text{NA} = 0.8$). All simulations use $n=1.33$.}
   \label{fig:single-frame}
@@
 with axis radii $a$ and $b$ and rotation angle $\gamma$ is
-\hypertarget{ellipse}{{\color{urlblue} 
 \begin{align}
   \frac{(x\cos\gamma + y\sin\gamma)^2}{a^2} + \frac{(x\sin\gamma - y\cos\gamma)^2}{b^2} =1. 
-\end{align}}}%
+\end{align}%
 We calculate the ellipse parameters in terms of the Fisher information matrix
@@
 Equations (\ref{eq:first})--(\ref{eq:last}) are identical to those found in
-\cite{coe2009} with an additional factor of $\sin\Theta$ on {\color{urlblue}Eqs.}
+\cite{coe2009} with an additional factor of $\sin\Theta$ on Eqs.
 (\ref{eq:second}) and (\ref{eq:third}) to account for the non-Euclidean geometry of
@@
 this paper are identical under inversion of the dipole orientation
-{\color{urlblue}[}$I(\Theta, \Phi) = I(\Theta + \pi, \Phi)$ in {\color{urlblue}Eqs.} (\ref{eq:scalarabs}),
-(\ref{eq:scalardet}), and (\ref{eq:forward_frame}){\color{urlblue}]}, but the solid-angle uncertainty
-and the uncertainty ellipse give no indication of this degeneracy. For this
-reason we also use the number of degeneracies as a measure of a microscope's
-ability to estimate the orientation of a dipole from intensity measurements. A
-small number of degeneracies is important for a microscope to find the
-orientation of a dipole---a twofold degenerate microscope can find the
+[$I(\Theta, \Phi) = I(\Theta + \pi, \Phi)$ in Eqs. (\ref{eq:scalarabs}),
+(\ref{eq:scalardet}), and (\ref{eq:forward_frame})], but the
+solid-angle uncertainty and the uncertainty ellipse give no indication of this
+degeneracy. For this reason we also use the number of degeneracies as a measure
+of a microscope's ability to estimate the orientation of a dipole from intensity
+measurements. A small number of degeneracies is important for a microscope to
+find the orientation of a dipole---a twofold degenerate microscope can find the
 orientation of a dipole down to the hemisphere while a fourfold degenerate
@@
 with a small illumination NA and a large detection NA where
-$\alpha \approx 65^{\circ}$ {\color{urlblue}[}see {\color{urlblue}Fig.} \ref{fig:single-arm}{\color{urlblue}(e)} at
-NA${}_{\text{det}} \approx 1.2${\color{urlblue}]}. A small illumination NA maximizes the range of
+$\alpha \approx 65^{\circ}$ [see Fig. \ref{fig:single-arm}(e) at
+NA${}_{\text{det}} \approx 1.2$]. A small illumination NA maximizes the range of
 absorption efficiencies, while a large detection NA optimizes the number of
@@
 uncertainty much less than increasing the detection NA (an expensive
-modification requiring a higher-NA objective). Also note in {\color{urlblue}Fig.}
-\ref{fig:single-arm}{\color{urlblue}(c)} that single-view designs suffer from high orientation
-uncertainty when the dipoles are oriented along the optical axis (the
-dipoles are not efficiently excited or detected), near the transverse plane
+modification requiring a higher-NA objective). Also note in
+Fig. \ref{fig:single-arm}(c) that single-view designs suffer from high
+orientation uncertainty when the dipoles are oriented along the optical axis
+(the dipoles are not efficiently excited or detected), near the transverse plane
 (it is ambiguous whether the dipole is oriented above or below the transverse
-plane, see {\color{urlblue}Fig.} \ref{fig:single-arm}{\color{urlblue}(b)}, and near the polarizer orientations
-{\color{urlblue}[}it is ambiguous whether the dipole is on either side of the polarizer
-orientation, see {\color{urlblue}Fig.} \ref{fig:single-arm}{\color{urlblue}(b)}{\color{urlblue}]}.
-
-The single-view microscope in {\color{urlblue}Fig.} \ref{fig:single-arm} is
-fourfold degenerate
-{\color{urlblue}[}$I(\Theta, \Phi) = I(\Theta + \pi, \Phi) = I(\Theta, \Phi + \pi) = I(\Theta +
-\pi, \Phi+\pi)$ in {\color{urlblue}Eqs.} (\ref{eq:scalarabs}), (\ref{eq:scalardet}),
-and (\ref{eq:forward_frame}){\color{urlblue}]}. The first twofold degeneracy is due to the usual
+plane, see Fig. \ref{fig:single-arm}(b), and near the polarizer orientations [it
+is ambiguous whether the dipole is on either side of the polarizer orientation,
+see Fig. \ref{fig:single-arm}(b)].
+
+The single-view microscope in Fig. \ref{fig:single-arm} is fourfold degenerate
+[$I(\Theta, \Phi) = I(\Theta + \pi, \Phi) = I(\Theta, \Phi + \pi) = I(\Theta +
+\pi, \Phi+\pi)$ in Eqs. (\ref{eq:scalarabs}), (\ref{eq:scalardet}), and
+(\ref{eq:forward_frame})]. The first twofold degeneracy is due to the usual
 inversion degeneracy---the dipole is an axis rather than a direction. A second
@@
 distinguish between dipoles reflected through the transverse plane. Evidence of
-the second degeneracy is found in {\color{urlblue}Figs.}
-\ref{fig:single-arm}{\color{urlblue}(b)} and
-\ref{fig:single-arm}{\color{urlblue}(c)} which show the ambiguity of whether the
-dipole is oriented above or below the transverse plane. Near the transverse
-plane this ambiguity is evident in the high solid-angle uncertainty, but far
-from the transverse plane evidence of the ambiguity is lost. Adding a second
-view and taking eight intensity measurements instead of four removes this
-degeneracy {\color{urlblue}[}see {\color{urlblue}Eqs.} (\ref{eq:thetap}) and (\ref{eq:phip}){\color{urlblue}]}, so the dual-view
-microscopes we consider in this paper are only twofold degenerate. This means
-that while dual-view microscopes can find the orientation of a dipole up to the
-hemisphere, single-view microscopes can only find the orientation of a dipole up
-to the quadrant.
+the second degeneracy is found in Figs. \ref{fig:single-arm}(b) and
+\ref{fig:single-arm}(c) which show the ambiguity of whether the dipole is
+oriented above or below the transverse plane. Near the transverse plane this
+ambiguity is evident in the high solid-angle uncertainty, but far from the
+transverse plane evidence of the ambiguity is lost. Adding a second view and
+taking eight intensity measurements instead of four removes this degeneracy
+[see Eqs. (\ref{eq:thetap}) and
+(\ref{eq:phip})], so the dual-view microscopes we consider in
+this paper are only twofold degenerate. This means that while dual-view
+microscopes can find the orientation of a dipole up to the hemisphere,
+single-view microscopes can only find the orientation of a dipole up to the
+quadrant.
 
@@
 Figures \ref{fig:symmetric-widefield} and \ref{fig:double-arm} show our results
-for dual-view symmetric designs. We illuminate from one objective and
-detect from the other for four polarization orientations, then we repeat the
-polarization orientations with the same objectives in reversed roles. In {\color{urlblue}Fig.}
+for dual-view symmetric designs. We illuminate from one objective and detect
+from the other for four polarization orientations, then we repeat the
+polarization orientations with the same objectives in reversed roles. In Fig.
 \ref{fig:symmetric-widefield} we sweep through the NA of both objectives and the
@@
 or against the cover slip. Given the high uncertainty in the transverse plane
-for a single-view shown in {\color{urlblue}Fig.} \ref{fig:single-arm}{\color{urlblue}(c)}), we can see why oblique
+for a single-view shown in Fig. \ref{fig:single-arm}(c), we can see why oblique
 views perform better than orthogonal views---oblique views are complementary
@@
 orthogonal design and varied the illumination and detection NA. Our results are
-similar to the single-view case in {\color{urlblue}Fig.} \ref{fig:single-arm}. We find that
+similar to the single-view case in Fig. \ref{fig:single-arm}. We find that
 optimal designs use a small illumination NA and a large detection NA. As we
@@
 objectives would touch the cover slip and each other
-{\color{urlblue}[}$\alpha_1 + \alpha_2 = 90^{\circ}$ in {\color{urlblue}Fig.}
-\ref{fig:asymmetric-double}{\color{urlblue}(a)}{\color{urlblue}]}. We swept
-through the NA asymmetry and the sample-exposure asymmetry while keeping the
-sample irradiance constant, and found that symmetric designs are at a local
+[$\alpha_1 + \alpha_2 = 90^{\circ}$ in Fig. \ref{fig:asymmetric-double}(a)]. We
+swept through the NA asymmetry and the sample-exposure asymmetry while keeping
+the sample irradiance constant, and found that symmetric designs are at a local
 minimum of the median and MAD of the solid-angle uncertainty. We also found that
 at extreme asymmetries where the exposure from the low-NA arm is much larger
-than the exposure from the high-NA arm {\color{urlblue}[}top-left and
-bottom-right corners of {\color{urlblue}Figs.}
-\ref{fig:asymmetric-double}{\color{urlblue}(d)} and {\color{urlblue}6(e)]} the
-median and MAD are comparable to the symmetric light-sheet microscope. This
-means that if a very high-NA objective is available and a low-NA objective can
-provide light sheet illumination, this microscope will perform comparably with a
-dual-view microscope with two medium-NA objectives. Note that this asymmetric
-exposure configuration maximizes the use of the high-NA detection objective---if
-a high-NA objective is available it should be used for detection as much as
-possible. Figure \ref{fig:asymmetric-double} also shows that
-designs with slightly asymmetric NA can trade off a low solid-angle uncertainty
-median for a low solid-angle uncertainty MAD by changing the sample exposure
-ratio.
+than the exposure from the high-NA arm [top-left and bottom-right corners of
+Figs. \ref{fig:asymmetric-double}(d) and 6(e)] the median and MAD are comparable
+to the symmetric light-sheet microscope. This means that if a very high-NA
+objective is available and a low-NA objective can provide light sheet
+illumination, this microscope will perform comparably with a dual-view
+microscope with two medium-NA objectives. Note that this asymmetric exposure
+configuration maximizes the use of the high-NA detection objective---if a
+high-NA objective is available it should be used for detection as much as
+possible. Figure \ref{fig:asymmetric-double} also shows that designs with
+slightly asymmetric NA can trade off a low solid-angle uncertainty median for a
+low solid-angle uncertainty MAD by changing the sample exposure ratio.
 
@@
 
-\hypertarget{temporal}{{\color{urlblue} The models in this paper only consider
-    estimating the orientation of fixed dipoles, but real dipoles change their
-    orientation during measurement. Therefore, it is necessary to comment on the
-    applicability of our results to dynamic dipoles. In general, using the
-    models in this paper to estimate the orientation of rotating dipoles will
-    lead to incorrect results. However, if a dipole rotates through a small
-    solid angle during the measurement, the models in this paper will give a
-    slightly biased estimate of the average orientation of the dipole. The
-    direction and size of the bias has not been characterized in detail. In
-    future work we will model dynamic dipoles and attempt to estimate the
-    average dipole orientation and the dipole's orientation
-    distribution. Single-view polarized illumination techniques have already
-    been used to study the orientation distribution of ensembles of dynamic
+The models in this paper only consider estimating the orientation of fixed
+dipoles, but real dipoles change their orientation during
+measurement. Therefore, it is necessary to comment on the applicability of our
+results to dynamic dipoles. In general, using the models in this paper to
+estimate the orientation of rotating dipoles will lead to incorrect
+results. However, if a dipole rotates through a small solid angle during the
+measurement, the models in this paper will give a slightly biased estimate of
+the average orientation of the dipole. The direction and size of the bias has
+not been characterized in detail. In future work we will model dynamic dipoles
+and attempt to estimate the average dipole orientation and the dipole's
+orientation distribution. Single-view polarized illumination techniques have
+already been used to study the orientation distribution of ensembles of dynamic
 dipoles \cite{mehta2016, backer2016}, but multiview polarized illumination
-    microscopes have not been investigated for this purpose.}}
-
-{\color{red}\sout{Although we have only considered the orientation determination
-    of fixed single dipoles, polarized illumination techniques can be used to
-    study the orientation distribution of ensembles of dynamic dipole. In future
-    work we will investigate the ability of multiview polarized illumination
-    microscopes to estimate the orientation distribution of ensembles of dynamic
-    dipoles.}}
+microscopes have not been investigated for this purpose.
 

### 2017-11-27 paper/paper/paper.tex
Copy edits for final version.

@@
 \author{Talon Chandler,\authormark{1,*} Shalin Mehta,\authormark{1,2,3} Hari
-  Shroff,\authormark{4,5} Rudolf Oldenbourg,\authormark{2, 6} and Patrick J. La
-  Rivi\`ere\authormark{1, 5}}
+  Shroff,\authormark{4,5} Rudolf Oldenbourg,\authormark{2,6} and Patrick J. La
+  Rivi\`ere\authormark{1,5}}
 \address{\authormark{1}University of Chicago, Department of Radiology, Chicago, Illinois 60637, USA\\
@@
 
-\subsection{Absorption Efficiency}\label{excitation}
+\subsection{Absorption efficiency}\label{excitation}
 In this section we will calculate the \emph{absorption efficiency} of a single
@@
 
-\subsection{Detection Efficiency}\label{detection}
+\subsection{Detection efficiency}\label{detection}
 In this section we will calculate the \emph{detection efficiency}---the fraction
@@
 
-\subsection{Oblique Illumination and Detection}\label{oblique}
+\subsection{Oblique illumination and detection}\label{oblique}
 In sections \ref{excitation} and \ref{detection} we assumed that both the
@@
 
-\subsection{Intensity Measurements}\label{forward}
+\subsection{Intensity measurements}\label{forward}
 The detected intensity is proportional to the product of the sample irradiance
@@
 
-\subsection{Polarized Illumination Microscope Configurations}\label{designs}
+\subsection{Polarized illumination microscope configurations}\label{designs}
 One way to collect multiple intensity measurements is to add a universal
@@
 
-\subsection{Evaluation Metrics}\label{metrics}
+\subsection{Evaluation metrics}\label{metrics}
 Our goal is to evaluate the ability of a microscope design to estimate the
@@
 
+% \begin{table}[ht!]
+% \centering
+% \caption{Comparison of designs in each class of microscope. All solid-angle
+%   uncertainty statistics are in steradians.}
+% \begin{tabular}{lllll}
+%   \toprule
+
+%   Microscope Type&$n$-fold&Max\{$\sigma_{\Omega}$\}&Median\{$\sigma_{\Omega}$\}&MAD\{$\sigma_{\Omega}$\}\\ 
+%     &Degeneracy&&&\\\midrule
+% Single-view (NA${}_\textrm{ill}$=0, NA${}_\textrm{det}$=1.1)&4&3.04$\times 10^{0}$&8.20$\times 10^{-4}$&1.48$\times 10^{-4}$ \\ 
+%   Dual-view oblique symmetric (NA=0.6, $\beta$=53${}^{\circ}$)&2&1.35$\times 10^{-1}$&4.49$\times 10^{-3}$&9.17$\times 10^{-4}$\\
+% Dual-view orthogonal symmetric light-sheet (NA=0.8)&2&6.23$\times 10^{-3}$&1.79$\times 10^{-3}$&1.65$\times 10^{-4}$\\  
+% Dual-view orthogonal symmetric light-sheet (NA=0.94)&2&3.13$\times 10^{-3}$&1.28$\times 10^{-3}$&1.14$\times 10^{-4}$\\
+% \bottomrule
+% \end{tabular}
+% \end{table}
+
 \begin{table}[ht!]
@@
   \toprule
-
-  Microscope Type&$n$-fold&Max\{$\sigma_{\Omega}$\}&Median\{$\sigma_{\Omega}$\}&MAD\{$\sigma_{\Omega}$\}\\ 
-    &Degeneracy&&&\\\midrule
-Single-view (NA${}_\textrm{ill}$=0, NA${}_\textrm{det}$=1.1)&4&3.04$\times 10^{0}$&8.20$\times 10^{-4}$&1.48$\times 10^{-4}$ \\ 
-  Dual-view oblique symmetric (NA=0.6, $\beta$=53${}^{\circ}$)&2&1.35$\times 10^{-1}$&4.49$\times 10^{-3}$&9.17$\times 10^{-4}$\\
-Dual-view orthogonal symmetric light-sheet (NA=0.8)&2&6.23$\times 10^{-3}$&1.79$\times 10^{-3}$&1.65$\times 10^{-4}$\\  
-Dual-view orthogonal symmetric light-sheet (NA=0.94)&2&3.13$\times 10^{-3}$&1.28$\times 10^{-3}$&1.14$\times 10^{-4}$\\
+  Microscope Type&Single-view&Dual-view&Dual-view&Dual-view\\ 
+    &NA${}_\textrm{ill}$=0&oblique&orthogonal&orthogonal \\
+    &NA${}_\textrm{det}$=1.1&symmetric&symmetric&symmetric\\
+    &&NA=0.6&light-sheet&light-sheet\\  
+    &&$\beta$=53${}^{\circ}$&NA=0.8&NA=0.94\\\midrule
+
+  $n$-fold Degeneracy&4&2&2&2\\
+  Max\{$\sigma_{\Omega}$\}&3.04$\times 10^{0}$&1.35$\times 10^{-1}$&$6.23\times 10^{-3}$&3.13$\times 10^{-3}$\\
+  Median\{$\sigma_{\Omega}$\}&8.20$\times 10^{-4}$&4.49$\times 10^{-3}$&$1.79\times 10^{-3}$&1.28$\times 10^{-3}$\\
+  MAD\{$\sigma_{\Omega}$\}&1.48$\times 10^{-4}$&9.17$\times 10^{-4}$&$1.65\times 10^{-4}$&1.14$\times 10^{-4}$\\
 \bottomrule
@@
 
-\section{Discussion and Conclusions}\label{discussion}
+\section{Discussion and conclusions}\label{discussion}
 In this work we have developed a model to predict the intensities measured by

### 2017-11-29 slides/2017-10-11-group-meeting/group_meeting.tex
Move slides to new folder. 2017-11-29 slides.

% Created 2017-10-11 Wed 11:15
\documentclass[presentation]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{amssymb}
\usepackage{hyperref}
\tolerance=1000
\usepackage{graphicx}
\usetheme{simple}
\usecolortheme{}
\usefonttheme{serif}
\useinnertheme{}
\useoutertheme{}
\author{Talon Chandler}
\date{October 11, 2017}
\title{Progress Report On Orientation Determination With Maximum Likelihood Methods}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs 25.3.1 (Org mode 8.2.10)}}
\begin{document}

\maketitle
\begin{frame}[label=sec-1]{Microscope Geometry}
\begin{center}
  \includegraphics[width=0.5\textwidth, interpolate=true]{figs/geometry.pdf}\\
  1.1 and 0.71 NA\\
  8 Poisson-distributed intensity measurements\\
  Goal: Estimate the orientation of the dipole.
\end{center}
\end{frame}
\begin{frame}[label=sec-2]{Fisher Scoring Algorithm}
\begin{center}
\begin{align*}
  \vec{\theta} = \{\Theta, \Phi\} &\rightarrow \text{Parameters}\\
  \vec{X}  &\rightarrow \text{Data (8 intensity measurements)}\\
  \textbf{F}(\vec{\theta}) &\rightarrow \text{Fisher information matrix}\\
  \vec{V}(\vec{\theta}, \vec{X}) = \frac{\partial}{\partial \vec{\theta}} \log L(\vec{\theta}, \vec{X}) &\rightarrow \text{Score}
\end{align*}
\begin{align*}
  \vec{\theta}_{i+1} = \vec{\theta}_{i} + \textbf{F}^{-1}(\vec{\theta})\vec{V}(\vec{\theta}, \vec{X})
\end{align*}
\end{center}
\end{frame}
\begin{frame}[label=sec-3]{Fisher Scoring Run 1}
\begin{center}
True orientation: $\Theta = \pi/2, \Phi = \pi/4$\\
Starting orientation: $\Theta = \pi/3, \Phi = \pi/3$
  \includegraphics[width=1.0\textwidth, interpolate=true]{figs/recon-history1.pdf}
\end{center}
\end{frame}
\begin{frame}[label=sec-4]{Fisher Scoring Run 2}
\begin{center}
True orientation: $\Theta = \pi/2, \Phi = \pi/4$\\
Starting orientation: $\Theta = \pi/3, \Phi = \pi/2$
  \includegraphics[width=1.0\textwidth, interpolate=true]{figs/recon-history2.pdf}
\end{center}
\end{frame}
\begin{frame}[label=sec-5]{Fisher Scoring Run 3}
\begin{center}
True orientation: $\Theta = \pi/2, \Phi = \pi/4$\\
Starting orientation: $\Theta = 0, \Phi = 0$
  \includegraphics[width=1.0\textwidth, interpolate=true]{figs/recon-history3.pdf}
\end{center}
\end{frame}
\begin{frame}[label=sec-6]{Log-likelihood as a function of the estimate orientation}
  \begin{center}
  True orientation: $\Theta = \pi/2, \Phi = \pi/4$\\
    \includegraphics[width=0.55\textwidth, interpolate=true]{figs/likelihood.pdf}\\
Gradient descent isn't going to cut it.\\
I need to be careful with derivatives on curved spaces. 
  \end{center}
\end{frame}

\begin{frame}[label=sec-7]{Log-likelihood as a function of the estimate orientation}
\begin{center}
True orientation: $\Theta = 0, \Phi = 0$\\
  \includegraphics[width=0.55\textwidth, interpolate=true]{figs/likelihood2.pdf}\\
\end{center}
\end{frame}

\begin{frame}[label=sec-8]{Brute Force Optimization}
Choose number of points (100). \\
Plot error in degrees (great circle arc) between truth and estimate as a function of truth. \\
\begin{center}
  \includegraphics[width=0.55\textwidth, interpolate=true]{figs/angle-error-brute.pdf}\\
\end{center}
\end{frame}
\begin{frame}[label=sec-9]{Particle Swarm Optimization}
Choose starting number of points (25). \\
    Choose max number of function evaluations (100). \\
    Plot error in degrees (great circle arc) between truth and estimate as a function of truth. \\
\begin{center}
  \includegraphics[width=0.55\textwidth, interpolate=true]{figs/angle-error.pdf}\\
\end{center}
\end{frame}
% Emacs 25.3.1 (Org mode 8.2.10)
\end{document}

### 2017-11-29 slides/2017-10-18-group-meeting/group_meeting.tex
Move slides to new folder. 2017-11-29 slides.

--- 2017-10-25 notes/2017-10-25-group-meeting/group_meeting.tex
@@
-% Created 2017-10-25 Wed 12:35
+% Created 2017-10-18 Wed 16:10
 \documentclass[presentation]{beamer}
@@
 \author{Talon Chandler}
-\date{October 25, 2017}
+\date{October 18, 2017}
 \title{Progress Report On 3D Orientation Determination}
@@
 \maketitle
-\begin{frame}[label=sec-1]{Distributions of fluorophores}
-\begin{center}
-  \includegraphics[width=1.0\textwidth, interpolate=true]{figs/watson.pdf}\\
-\end{center}
-\end{frame}
-
-\begin{frame}[label=sec-2]{Log-likelihood as a function of the estimate orientation ($\Theta$, $\Phi$)}
+\begin{frame}[label=sec-1]{Log-likelihood as a function of the estimate orientation}
 \begin{center}
@@
 
-\begin{frame}[label=sec-3]{Log-likelihood as a function of the fluorophore distribution ($\Theta$, $\Phi$, $\kappa$, $c$)}
+\begin{frame}[label=sec-2]{}
 \begin{center}
-True orientation: $\Theta = 0, \Phi = 0, \kappa = \infty, c = 1$\\
-  \includegraphics[width=0.9\textwidth, interpolate=true]{figs/likelihood-dist.pdf}\\
+  \includegraphics[width=0.9\textwidth, interpolate=true]{figs/sketch1.png}\\
 \end{center}
@@
 
-\begin{frame}[label=sec-4]{Log-likelihood as a function of the fluorophore distribution ($\Theta$, $\Phi$, $\kappa$, $c$)}
+\begin{frame}[label=sec-3]{}
 \begin{center}
-True orientation: $\Theta = 0, \Phi = 0, \kappa = 10, c = 1$\\
-  \includegraphics[width=0.9\textwidth, interpolate=true]{figs/likelihood-dist2.pdf}\\
+  \includegraphics[width=0.9\textwidth, interpolate=true]{figs/sketch2.png}\\
 \end{center}
 \end{frame}
-\begin{frame}[label=sec-5]{Discussion}
+
+\begin{frame}[label=sec-4]{}
+\begin{center}
+  \includegraphics[width=0.9\textwidth, interpolate=true]{figs/sketch3.png}\\
+\end{center}
+\end{frame}
+
+\begin{frame}[label=sec-5]{Workarounds}
 \begin{itemize}
-\item $\Theta$ and $\Phi$ are bounded
-\item $\kappa$ and $c$ are unbounded...estimate log($\kappa$) and log($c$) instead?
+\item Block half of one aperture
+\item Add a third view (bottom view, microlenses)
+\item Assume fluorophores are never oriented in one of the degenerate directions
 \end{itemize}
+\end{frame}
+\begin{frame}[label=sec-6]{}
+\begin{center}
+  \includegraphics[width=0.9\textwidth, interpolate=true]{figs/2dpol}\\
+\end{center}
+\end{frame}
+\begin{frame}[label=sec-7]{}
+\begin{center}
+  \includegraphics[width=0.9\textwidth, interpolate=true]{figs/3dpol}\\
+\end{center}
+\end{frame}
+
+\begin{frame}[label=sec-8]{}
+\begin{center}
+  \includegraphics[width=0.9\textwidth, interpolate=true]{figs/data}\\
+\end{center}
+\end{frame}
+
+\begin{frame}[label=sec-9]{Reconstructed 3D orientation from split-aperture instant FluoPolScope}
+\begin{center}
+  \includegraphics[width=0.7\textwidth, interpolate=true]{figs/frame01}\\
+\end{center}
 \end{frame}

### 2017-11-29 slides/2017-11-29-group-meeting/group_meeting.tex
Move slides to new folder. 2017-11-29 slides.

% Created 2017-11-29 Wed 11:57
\documentclass[presentation]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{amssymb}
\usepackage{hyperref}
\tolerance=1000
\usepackage{graphicx} \DeclareMathOperator{\argmin}{argmin}
\usetheme{simple}
\usecolortheme{}
\usefonttheme{serif}
\useinnertheme{}
\useoutertheme{}
\author{Talon Chandler}
\date{November 29, 2017}
\title{Faster Techniques For 3D Orientation Reconstruction}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs 25.3.1 (Org mode 8.2.10)}}
\begin{document}

\maketitle
\begin{frame}[label=sec-1]{Background: Real Projective Plane $\mathbb{RP}^2$}
\begin{itemize}
\item Set of all infinite lines through the origin
\item Two dimensional space
\item Not a vector space! 
\item Single dipole orientations live in the real projective plane 
\item The sphere $\mathbb{S}^2$ is a \textit{double cover} of $\mathbb{RP}^2$ 
\item The sphere $\mathbb{S}^2$ can be \textit{embedded} in 3D Euclidean space $\mathbb{R}^3$ 
\end{itemize}
\end{frame}
\begin{frame}[label=sec-2]{Watson Distribution}
\begin{center}
  \includegraphics[width=1.0\textwidth, interpolate=true]{figs/watson.pdf}\\
\end{center}
\begin{align*}
  &f(\mh{r}; \bs{\hat{\mu}}, \kappa) = \frac{1}{4\pi{}_1F_1\left(\frac{1}{2}, \frac{3}{2}, \kappa\right)}\text{exp}\{\kappa (\bs{\hat{\mu}}^T\mh{r})^2\}\\
  &r \in \mathbb{RP}^2, \hat{\mu} \in \mathbb{RP}^2, \kappa \in \mathbb{R}
\end{align*}
\begin{itemize}
\item Expensive special function ${}_1F_1$
\item Can't take integrals or derivatives wrt $\kappa$
\item Can't take intensity integrals wrt $\hat{\mu}$
\end{itemize}
\end{frame}

\begin{frame}[label=sec-3]{Central Angular Gaussian Distribution}
\begin{align*}
  &f(\mh{r}; \mathbf{A}) = \frac{r^T\mathbf{A}r}{|\mathbf{A}|}\\
  &r \in \mathbb{RP}^2\\
  &\mathbf{A}\ \text{is a}\ 3\times 3\ \text{positive definite matrix}
\end{align*}
\begin{itemize}
\item Projection of a Gaussian in $\mathbb{R}^3$ into $\mathbb{RP}^2$
\item $\mathbf{A}$ defines a ellipsoid.
\item Too general for us. We want rotational symmetry. 
\end{itemize}
\end{frame}
\begin{frame}[label=sec-4]{Spheroid Distribution}
\vspace{-1em}
\begin{center}
  \includegraphics[width=1.0\textwidth, interpolate=true]{figs/spheroid.pdf}
\end{center}
\vspace{-4em}
\begin{align*}
  f(\mh{r}; \mathbf{A}) = \frac{r^T\mathbf{A}r}{|\mathbf{A}|}
\end{align*}
\vspace{-1em}
\begin{itemize}
\item $\mathbf{A}$ is a 3 $\times$ 3 positive definite matrix with 3 orthonormal evecs.
\item The first evec is pointed along the symmetry axis $\hat{\mu} \in \mathbb{RP}^2$
\item The other two evecs are have the same eigenvalue. 
\item The ratio of the symmetry evalue to orthogonal evalues is $\kappa^2$. 
\item Cheap to compute. Can take derivatives and integrals.
\item $\kappa$ has an easy interpretation---ratio of symmetry axis radius to orthogonal axis radius
\item Matches Rudolf's previous work with birefringent materials. We're comfortable reasoning about spheroids. 
\end{itemize}
\end{frame}

\begin{frame}[label=sec-5]{Next Topic: Optimization Problem}
\begin{align*}
f&: \mathbb{S}^2 \rightarrow \mathbb{R}\\ \\
x^* &= \underset{x\in \mathbb{S}^2}{\mathrm{argmin}}\ f(x)
\end{align*}
\begin{itemize}
\item What is the gradient of $f$?\\
\item Use spherical coordinates ($\theta$, $\phi$). $\nabla f \stackrel{?}{=} \left(\frac{\partial f}{\partial \theta}, \frac{\partial f}{\partial \phi})$
\item Step size is different at different points on the sphere.
\item Gradient depends on the coordinate choice. 
\item Trouble at the poles!
\end{itemize}
\end{frame}
\begin{frame}[label=sec-6]{Instead optimize on a manifold}
\begin{center}
  \includegraphics[width=0.4\textwidth, interpolate=true]{figs/book.png}\\
\end{center}
\end{frame}
\begin{frame}[label=sec-7]{Rough Recipe For Optimizing On A Manifold}
\begin{enumerate}
\item Let $f:\mathbb{S}^2 \rightarrow \mathbb{R}$ be the function we want to optimize. 
\item {Embed the manifold in Euclidean space and define a new smooth function $\bar{f}: \mathbb{R}^3 \rightarrow \mathbb{R}$. If we 
constr$\bar{f}$ to $\mathbb{S}^2$ then we recover $f$.}
\item Choose an initial guess on $\mathbb{S}^2$. 
\item Take the usual Euclidean gradient of $\bar{f}$. $\nabla \bar{f} = \left(\frac{\partial f}{\partial x}, \frac{\partial f}{\partial y}, \frac{\partial f}{\partial z})$
\item Project the Euclidean gradient onto the tangent space of $\mathbb{S}^2$. This projected gradient is called the \textit{Riemannian gradient}. 
\item Update your guess by moving along the Riemannian gradient. 
\item \textit{Retract} your new guess from the tangent space back onto $\mathbb{S}^2$. 
\item Repeat from step 4.
\end{enumerate}
\end{frame}
\begin{frame}[label=sec-8]{}
\begin{center}
  \includegraphics[width=1.0\textwidth, interpolate=true]{figs/tangent.pdf}\\
\end{center}
\end{frame}
\begin{frame}[label=sec-9]{Summary + Work In Progress}
\begin{itemize}
\item A central angular Gaussian distribution with rotational symmetry gives the spheroid distribution.
\item Spheroid distributions are easier to handle than Watson distributions.
\item I'm working on integrating the spheroid distribution analytically which will yield a much faster forward model.
\item Optimizing on an embedded manifold allows us to calculate gradients correctly.
\item Multiple seed gradient methods will be much faster than the gradient-free particle swarm methods I've been using.
\item Initial seed could be generated with Rudolf's proposed change of coordinates.
\end{itemize}
\end{frame}
% Emacs 25.3.1 (Org mode 8.2.10)
\end{document}

### 2017-11-29 slides/2017-11-29-group-meeting/figs/spheroid.py

@@ -1,0 +1,37 @@
+from mpl_toolkits.mplot3d import Axes3D
+import matplotlib.pyplot as plt
+import numpy as np
+import matplotlib.gridspec as gridspec
+import scipy.special as special
+
+
+inch_fig = 3
+f, axs = plt.subplots(nrows=1, ncols=7, figsize=(7*inch_fig, inch_fig), subplot_kw={'projection':'3d'})
+plt.subplots_adjust(wspace=-0.7)
+kappas = [0.1, 0.25, 0.5, 1, 2, 4, 10]
+for ax, kappa in zip(axs, kappas):
+    ax.set_aspect("equal")
+    
+    phi, theta = np.mgrid[0:2*np.pi:15j, 0:np.pi:20j]
+    if kappa == 1:
+        norm = 4*np.pi
+    elif kappa < 1:
+        z = np.sqrt(1 - kappa**2)
+        norm = 2*np.pi*(1 + (kappa**2/(2*z))*np.log((1+z)/(1-z)))
+    elif kappa > 1:
+        z = np.sqrt(kappa**2 - 1)        
+        norm = 2*np.pi*(1 + (kappa**2/z)*np.arcsin(z/kappa))
+    weight = 1.5
+    x = (1.0/norm)*weight*np.cos(phi)*np.sin(theta)
+    y = (1.0/norm)*weight*np.sin(phi)*np.sin(theta)
+    z = (1.0/norm)*weight*np.cos(theta)*kappa
+    ax.plot_wireframe(x, y, z, color="k", lw=0.1)
+
+    ax.set_xlim(-0.3, 0.3)
+    ax.set_ylim(-0.3, 0.3)
+    ax.set_zlim(-0.3, 0.3)    
+    ax.set_axis_off()
+    ax.azim = 0
+    ax.elev = 30
+    ax.set_title('$\kappa = '+str(kappa)+'$', y=0.9, fontsize=20)
+f.savefig('spheroid.pdf')

### 2017-11-29 slides/2017-11-29-group-meeting/scratch/calc_rot.py

@@ -1,0 +1,12 @@
+import sympy.matrices.dense
+from sympy import Symbol, simplify
+theta = Symbol("theta")
+phi = Symbol("phi")
+Rytheta = sympy.dense.rot_axis2(-theta)
+Rzphi = sympy.dense.rot_axis3(-phi)
+
+print(Rytheta)
+print(Rzphi)
+print(simplify(Rzphi*Rytheta))
+
+

### 2017-12-06 slides/2017-12-06-group-meeting/group_meeting.tex
Dec 6 group meeting.

% Created 2017-12-06 Wed 11:31
\documentclass[presentation]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{amssymb}
\usepackage{hyperref}
\tolerance=1000
\usepackage{graphicx} \DeclareMathOperator{\argmin}{argmin}
\usetheme{simple}
\usecolortheme{}
\usefonttheme{serif}
\useinnertheme{}
\useoutertheme{}
\author{Talon Chandler}
\date{December 6, 2017}
\title{Update On 3D Orientation Reconstruction}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs 25.3.1 (Org mode 8.2.10)}}
\begin{document}

\maketitle
\begin{frame}[label=sec-1]{Spheroid Distribution}
\vspace{-1em}
\begin{center}
  \includegraphics[width=1.0\textwidth, interpolate=true]{figs/spheroid.pdf}
\end{center}
\vspace{-4em}
\begin{align*}
  f(\vec{r}; \mathbf{A}) &= \frac{r^T\mathbf{A}^{-1}r}{|\mathbf{A}|}\\
  f(\vec{r}; \vec{r}', \kappa) &= \frac{\kappa}{S(\kappa)\sqrt{1+(\vec{r}\cdot\vec{r}')^2\left(\frac{1}{\kappa^2} - 1\right)}}\\
  S(\kappa) &= 2\pi\left(1 + \frac{\kappa^2}{2\sqrt{1-\kappa^2}}\log\left[\frac{1+\sqrt{1-\kappa^2}}{1-\sqrt{1-\kappa^2}}\right]\right)
\end{align*}
\vspace{-1em}
\begin{itemize}
\end{itemize}
\end{frame}
\begin{frame}[label=sec-2]{Forward Model: Single diSPIM Frame}
\begin{align*}
  \text{Measured Intensity:   }& I = \int_{\mathbb{S}^2}d\vec{r}\ I^s(\vec{r})f(\vec{r}; \vec{r}', \kappa)\\ \\
  \text{Single Fluorophore:   }& I^s(\vec{r}) = 2[A+B(1 - \cos^2\phi\sin^2\theta)]\times \\ & \hspace{7em}\sin^2\theta\cos^2(\phi - \phi_{\text{pol}})\\ \\
  \text{Spheroid Distribution:   }& f(\vec{r}; \vec{r}', \kappa) = \frac{\kappa}{S(\kappa)\sqrt{1+(\vec{r}\cdot\vec{r}')^2\left(\frac{1}{\kappa^2} - 1\right)}}
\end{align*}
\begin{itemize}
\item No luck integrating in closed form ;(
\end{itemize}
\end{frame}
\begin{frame}[label=sec-3]{Integrating Faster With Spherical Quadrature}
\begin{center}
  \includegraphics[width=0.7\textwidth, interpolate=true]{figs/quadrature.png}
  \raisebox{-10pt}{\makebox[0pt][r]{\footnotesize Beentjes, 2012}}
\end{center}
\begin{itemize}
\item Instead of evaluating integrals one after the other and using Gaussian quadrature (left), evaluate the integral all at once using Lebedev quadrature (right)
\item How many grid points do I need to use to accurately integrate?\\ To within 1\%? Perfectly? $\rightarrow$
\end{itemize}
\end{frame}
\begin{frame}[label=sec-4]{Harmonic Analysis On The Sphere}
\begin{align*}
  \text{Measured Intensity:   }& I = \int_{\mathbb{S}^2}d\vec{r}\ I^s(\vec{r})f(\vec{r})
\end{align*}
\begin{itemize}
\item Rewrite $I^s(\vec{r})$ and $f(\vec{r})$ in terms of orthogonal spherical harmonics\\ \\
\item Expensive integral becomes a cheap dot product of coefficients\\ \\
\item With measurements under different polarization settings we get $\vec{g} = \mathcal{H}\vec{f}$ where $\vec{g}$ is a vector of intensity measurements, $\vec{f}$ is a vector of spherical harmonic coefficients of the fluorophore distribution, and the rows of $\mathcal{H}$ are the spherical harmonic coefficients of $I^s(\vec{r})$ for each measurement \\ 
\item Null space of $\mathcal{H}$ corresponds to symmetries/degeneracies
\end{itemize}
\end{frame}
\begin{frame}[label=sec-5]{Harmonic Analysis}
\begin{itemize}
\item Harmonic analysis on $\mathbb{R}^n \rightarrow$ $n$-D Fourier Transform\\
\item Harmonic analysis on $\mathbb{S}^1$ (circle) $\rightarrow$ Fourier Series\\
\item Harmonic analysis on $\mathbb{S}^2$ (sphere) $\rightarrow$ 2D Fourier Series
\end{itemize}
\end{frame}
\begin{frame}[label=sec-6]{Spherical Harmonics}
\begin{center}
  \includegraphics[width=0.9\textwidth, interpolate=true]{figs/harmonics.png}
  \raisebox{-10pt}{\makebox[0pt][r]{\footnotesize Jarosz, 2008}}
\end{center}
\end{frame}

\begin{frame}[label=sec-7]{Single View diSPIM Frame}
\begin{center}
  \includegraphics[width=0.7\textwidth, interpolate=true]{figs/harmonics_supp.png}
  \raisebox{-10pt}{\makebox[0pt][r]{\footnotesize Jarosz, 2008}}
\end{center}
\begin{itemize}
\item Non-zero coefficients for single view diSPIM frame shown in red boxes
\end{itemize}
\end{frame}
\begin{frame}[label=sec-8]{Early Conclusions + Upcoming Work}
\begin{itemize}
\item Odd bands aren't antipodally symmetric so they aren't present in the sample.
\item Zero coefficients on $y_2^{\pm 1}$. Corresponds to degeneracy? Investigating. 
\item Adding polarizers on the detection side with give us access to 4th order band. Still investigating degeneracy. 
\item A single fluorophore has non-zero components in all even bands (sort of like FT of $\delta(x)$ is 1). We can only measure the first few bands which constrains the accuracy of our estimates. 
\item Now we're estimating spherical harmonic coefficients instead of parameters of a hypothesized distribution. That's all of the info we can access. 
\item Upcoming---reconstruction in terms of matrix inversion! 
\end{itemize}
\end{frame}
% Emacs 25.3.1 (Org mode 8.2.10)
\end{document}

### 2017-12-12 harmonic-paper/paper/paper.tex
Harmonic analysis notes draft.

\documentclass[10pt]{article}
\usepackage{float}
\usepackage{booktabs}
\usepackage{menukeys}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[normalem]{ulem}
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
\providecommand{\mb}[1]{\mathbf{#1}}
\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
\providecommand{\bs}[1]{\boldsymbol{#1}} 
\providecommand{\intinf}{\int_{-\infty}^{\infty}}
\providecommand{\mbb}[1]{\mathbb{#1}}
\linespread{1.3}

\usepackage[
top    = 1.5cm,
bottom = 1.5cm,
left   = 1.5cm,
right  = 1.5cm]{geometry}

\begin{document}
\title{\vspace{-2.5em} Harmonic Analysis of Fluorophore Orientation Measurement Systems \vspace{-1em}}
\maketitle
% \author{Talon Chandler,\authormark{1,*} Shalin Mehta,\authormark{1,2,3} Hari
%   Shroff,\authormark{4,5} Rudolf Oldenbourg,\authormark{2,6} and Patrick J. La
%   Rivi\`ere\authormark{1,5}}
% \address{\authormark{1}University of Chicago, Department of Radiology, Chicago, Illinois 60637, USA\\
%   \authormark{2}Marine Biological Laboratory, Bell Center, Woods Hole, Massachusetts 02543, USA\\
%   \authormark{3}(present address) Chan Zuckerberg Biohub, San Francisco, California 94158, USA\\
%   \authormark{4}Section on High Resolution Optical Imaging, National Institute
%   of Biomedical Imaging and Bioengineering, National Institutes of Health,
%   Bethesda, Maryland 20892, USA\\
%   \authormark{5}Marine Biological Laboratory, Whitman Center, Woods Hole,
%   Massachusetts 02543, USA\\
%   \authormark{6}Brown University, Department of Physics, Providence, Rhode
%   Island 02912, USA}
% \email{*talonchandler@talonchandler.com} %% email address is required

%%%%%%%%%%%%%%%%%%% abstract and OCIS codes %%%%%%%%%%%%%%%%

\begin{abstract} 
TODO 
\end{abstract}

% \ocis{(110.0110) Imaging systems; (180.2520) Fluorescence microscopy; (180.0180)
%   Microscopy; (180.6900) Three-dimensional microscopy; (130.5440)
%   Polarization-selective devices; (260.5430) Polarization.}

%%%%%%%%%%%%%%%%%%%%%%% References %%%%%%%%%%%%%%%%%%%%%%%%%
% Comment before submission
 % \bibliography{paper}
 % \bibliographystyle{osajnl}

% Copy .bbl and uncomment these before submission

% \begin{thebibliography}{10}
% \newcommand{\enquote}[1]{``#1''}
% \end{thebibliography}

\section{Introduction}\label{sec:intro}
TODO

\section{Theory}
We use $\mh{r}$ and $(\theta, \phi)$ interchangeably to represent points on the
unit sphere $\mbb{S}^2$ using the parameterization
\begin{align}
\mh{r} = (\sin\theta\cos\phi, \sin\theta\sin\phi, \cos\theta).
\end{align}
Whenever possible we use the notation and vocabulary of Barrett and Myers
\cite{barrett}.

\subsection{Laplace series}
Consider a square-integrable function $f: \mbb{S}^2 \rightarrow \mbb{R}$ that
maps points on the unit sphere to the real numbers. We can express $f$ as a
weighted sum of spherical harmonics called the Laplace series \cite{arfken,
  sloan}
\begin{align}
  f(\theta, \phi) &= \sum_{l=0}^{\infty}\sum_{m=-l}^{l}F_l^m y_l^m(\theta, \phi) \label{eq:double}
\end{align}
where 
\begin{align}
  y_l^m(\theta, \phi) =
  \begin{cases}
    \sqrt{2}K_l^m\cos(m\phi)P_l^m(\cos\theta), & m > 0\\
    K_l^0P_l^0(\cos\theta), & m = 0\\
    \sqrt{2}K_l^m\sin(-m\phi)P_l^{-m}(\cos\theta), & m < 0\\
  \end{cases}
\end{align}
\begin{align}
  K_l^m = \sqrt{\frac{(2l+1)}{4\pi}\frac{(l-|m|)!}{(l+|m|)!}}
\end{align}
and $P_l^m(x)$ are the associated Legendre polynomials. We have chosen to use
the real spherical harmonics because we are representing real functions
only. For convenience we will rewrite equation \ref{eq:double} using a single
index $i = l(l+1) + m$ which gives
\begin{align}
  f(\theta, \phi) &= \sum_{i=0}^{\infty}F_iy_i(\theta, \phi). \label{eq:concrete}
\end{align}
We will also rewrite the functions $f$ and $y_i$ as vectors using a single index
to represent points on the sphere which gives
\begin{align}
  \mb{f} = \sum_{i=0}^{\infty}F_i\mb{y}_i. \label{eq:concrete2}
\end{align}
where $\mb{f} \in \mbb{R}^{\infty\times 1}$ and $\mb{y}_i \in \mbb{R}^{\infty\times 1}$

Finally, we express equation \ref{eq:concrete2} in its final form using the
notation
\begin{align}
 \mb{f} = \mb{Y}^T\mb{F} \label{eq:abstract}
\end{align}
where $\cdot^T$ denotes a transpose, $\mb{F}\in\mbb{R}^{\infty\times 1}$ is an
ordered vector of the Laplace series coefficients, and
$\mb{Y}\in\mbb{R}^{\infty \times \infty}$ is a matrix with each column
consisting of a spherical harmonic $\mb{y}_i$. Note that equation
\ref{eq:abstract} is just a reindexed version of equation \ref{eq:double}.

A key property of the spherical harmonics is that they form an orthonormal basis
on the sphere---that is
\begin{align}
  \mb{Y}^T\mb{Y} = \mb{Y}\mb{Y}^T = \mb{I} \label{eq:ortho}
\end{align}
where $\mb{I}$ is the identity matrix. We can exploit this orthogonality to
compute the Laplace coefficients $\mb{F}$ for a given function $\mb{f}$. If we
left-multiply both sides of equation \ref{eq:abstract} by $\mb{Y}$ and use
equation \ref{eq:ortho} to simplify we get
\begin{align}
  \mb{F} = \mb{Y}\mb{f}. 
\end{align}
We call $\mb{Y}$ the \textit{Laplace expansion operator} because it acts on a
function $\mb{f}$ and generates its Laplace coefficients $\mb{F}$. Note that the
Laplace series of a function on $\mbb{S}^2$ is analogous to the Fourier series
of a function on $\mbb{S}^1$. When we expand a function into a Laplace or
Fourier series, we are expressing the function in a convenient orthonormal
basis.

\subsection{Harmonic analysis of orientation measurements}
Consider a single fluorophore with a fixed dipole axis oriented along a
direction $\mh{r}$. To measure the orientation of the dipole axis we set up an
illumination system to excite the fluorophore---a light source and a combination
of lenses and polarizers---and a detection system to detect the light that the
fluorophore emits as it relaxes---a detector and a combination of lenses and
polarizers. We are deliberately considering very general orientation
measurement systems at this stage.

We can model the complete measurement system using a function
$h(\mh{r}): \mbb{S}^2 \rightarrow \mbb{R}_+$ that maps the dipole orientation to
measured intensity values. We call $h(\mh{r})$ the \textit{point response
  function} of the measurement system. To find the point response function we
can model the measurement system mathematically, or we can build the measurement
system and take measurements as we manipulate the orientation of a single
fluorophore.

Single fluorophores with fixed dipole axes only appear in a very small set of
samples---samples with few fluorophores fixed to immobile structures or samples
with photoactivatable fluorophores. Most samples of interest contain many
rotating fluorophores, and we would like to measure the quantity and orientation
distribution of these fluorophores. Let
$f(\mh{r}): \mbb{S}^2 \rightarrow \mbb{R}_+$ be a function that maps a direction
(a point on the sphere) to half the number of fluorophores pointing along that
direction. We call $f$ the \textit{orientation distribution
  function}. For a single fluorophore oriented along $\mh{r}_0$ the orientation
distribution function is
\begin{align}
  f_{\text{single}}(\mh{r}) = \frac{1}{2}\left[\delta(\mh{r} - \mh{r}_0) + \delta(\mh{r} + \mh{r}_0)\right]
\end{align}
because dipoles are antipodally symmetric. When we add fluorophores to the measurement
volume we add the individual orientation distribution functions
\begin{align}
  f_{N}(\mh{r}) = \sum_{i=0}^N \frac{1}{2}\left[\delta(\mh{r} - \mh{r}_i) + \delta(\mh{r} + \mh{r}_i)\right]
\end{align}
As the number of fluorophores increases the orientation distribution function
approaches a continuous function.

The total intensity measured by the system is given by multiplying the point
response function with the orientation distribution function and integrating
over the sphere
\begin{align}
  g = \int_{\mathbb{S}^2}d\mh{r}\ h(\mh{r})f(\mh{r}). \label{eq:forward}
\end{align}

Equation \ref{eq:forward} assumes that the measurement system is
\textit{linear}---if $\int_{\mathbb{S}^2}d\mh{r}\ h(\mh{r})f_1(\mh{r}) = g_1$
and $\int_{\mathbb{S}^2}d\mh{r}\ h(\mh{r})f_2(\mh{r}) = g_2$ then
$\int_{\mathbb{S}^2}d\mh{r}\ h(\mh{r})[af_1(\mh{r}) + bf_2(\mh{r})] =
ag_1+bg_2$. This requirement is satisfied in most experimental cases as long as
the fluorophores and detectors are not saturated. Note that this requirement
does not mean that $h$ and $f$ have to be linear functions.

We simplify our notation by reindexing $h$ and $f$ and rewriting equation
\ref{eq:forward} as
\begin{align}
  g = \mb{h}^T\mb{f}. 
\end{align}
If we expand $\mb{h}$ and $\mb{f}$ into their respective Laplace series using equation
\ref{eq:abstract} then
\begin{align}
  g = (\mb{Y}^T\mb{H})^T(\mb{Y}^T\mb{F})
\end{align}
where $\mb{H} \in \mbb{R}^{\infty\times 1}$ and
$\mb{F} \in \mbb{R}^{\infty \times 1}$ are vectors made of the Laplace
coefficients of $\mb{h}$ and $\mb{f}$, respectively. Exploiting the
orthogonality of the spherical harmonics (equation \ref{eq:ortho}) gives
\begin{align}
  g = \mb{H}^T\mb{F} \label{eq:dot}
\end{align}
Equation \ref{eq:dot} gives us a fundamental insight into fluorophore
orientation measurement systems---the measured intensity is a sum of the Laplace
coefficients of the orientation distribution function weighted by the Laplace
coefficients of the point response function.

Most orientation measurement systems make multiple intensity measurements with
different point response functions by using multiple detectors or by changing
the geometry of the measurement system. We can represent an orientation
measurement system that takes $N$ measurements using the compact notation
\begin{align}
  \mb{g} = \mb{\Psi}\mb{F} \label{eq:fw_mat}
\end{align}
where $\mb{g} \in \mathbb{R}^N_+$ is a vector of the intensity measurements,
and $\mb{\Psi} \in \mbb{R}^{N\times \infty}$ is the \textit{system matrix}
where the $n$th row of the system matrix consists of the Laplace coefficients of
the $n$th point response function. We can also rewrite equation \ref{eq:fw_mat}
as
\begin{align}
  \mb{g} = \mathcal{H}\mb{f}
\end{align}
where $\mathcal{H} \equiv \mb{\Psi}\mb{Y}^T \in \mbb{R}^{N\times \infty}$ is the 
continuous-to-discrete forward operator of the orientation measurement system.

In general, equation \ref{eq:fw_mat} is a product of an $N\times \infty$ system
matrix with a $\infty \times 1$ vector. Luckily, all orientation measurements
systems are \textit{band-limited}---the system matrix only consists of $M$
non-zero columns which simplifies equation \ref{eq:fw_mat} to a product of an
$N\times M$ system matrix with an $M \times 1$ vector.

In practice the intensity measurements made in equation \ref{eq:fw_mat} are
corrupted by Poisson noise and background counts, so the complete forward model is
given by
\begin{align}
  \mb{g} \sim \text{Pois}(\mathcal{H}\mb{f} + \mb{b})
\end{align}
where $\mb{b} \in \mbb{R}_+^N$ is a vector of background intensity
measurements.

\subsection{Analysis and design}
We can use the formalism developed in the previous section to characterize and
optimize an orientation imaging system. The first task for the designer is to
find the point response function for each intensity measurement
$\{h_1, h_2, \ldots h_M\}$ via modeling or direct measurement. The second task
is to find the Laplace series coefficients for the point response functions and
use them to assemble the system matrix $\mb{\Psi}$. The third task is to
calculate the \textit{crosstalk matrix} defined as
\begin{align}
  \mb{B} \equiv \mb{\Psi}^T\mb{\Psi}.
\end{align}
The crosstalk matrix is an infinite positive semidefinite matrix that
characterizes the aliasing properties of the measurement system. The
off-diagonal entries $\beta_{i,j}$ measure the degree of aliasing between the
$i$th and $j$th Laplace components---a zero indicates that the components are
orthogonal, and a one indicates that the components are linearly dependent. The
diagonal entries $\beta_{i,i}$ indicate how efficiently the $i$th Laplace
component is transferred by the measurement system---a zero indicates that the
component cannot be recovered from the data, and a one indicates that the data
is efficiently transmitted. The ideal crosstalk matrix is the identity
matrix---such a system could transmit all Laplace components without aliasing
and could be used to reconstruct all possible fluorophore distributions.

The final task of the designer is to optimize the system based on the
information in the crosstalk matrix. The designer may want to add extra
measurements so that extra Laplace components are transmitted by the system. The
designer may also have access to other parameters---polarizer orientations,
numerical aperture, illumination and detection geometry---that may be optimized
to extend and diagonalize the crosstalk matrix.

\subsection{Reconstruction methods}
Our goal is to estimate the orientation distribution function from a set of
intensity measurements $\mb{g}$. If nothing is known about the orientation
distribution function and the noise on the measurements is small, then we can
use the minimum-norm least-squares (MNLS) estimate
\begin{align}
  \hat{\mb{F}}_\text{MNLS} = \mb{\Psi}^+(\mb{g}-\mb{b}) \label{eq:mnls_est}
\end{align}
where $\mathbf{\Psi}^+$ is the Moore-Penrose pseudoinverse of
$\mb{\Psi}$. Once we have an estimate of the Laplace coefficients, we can
find the estimated orientation distribution function with
\begin{align}
  \hat{\mb{f}}_\text{MNLS} = \mb{Y}^T\hat{\mb{F}}_\text{MNLS}. \label{eq:mnls}
\end{align}

The MNLS algorithm is very efficient, but it amplifies noise when $\mb{\Psi}$
has small singular values. In low-noise applications this is not an issue, but
in many applications amplifying noise is not acceptable.

A more generally applicable algorithm is the Richardson-Lucy (RL) algorithm with
updates given by
\begin{align}
  \hat{\mb{F}}^{(i+1)} = \text{diag}(\mathbf{H}^T\textbf{1})^{-1}\text{diag}\left[\mb{\Psi}^T\text{diag}(\mb{\Psi}\hat{\mb{F}}^{(i)} + \mb{b})^{-1}\mb{g}\right]\hat{\mb{F}}^{(i)}. \label{eq:rl_est}
\end{align}
Once the RL algorithm has converged to an estimate of the Laplace coefficients
$\hat{\mb{F}}_\text{RL}$, we can find the estimated orientation distribution function
with
\begin{align}
  \hat{\mb{f}}_\text{RL} = \mb{Y}^T\hat{\mb{F}}_\text{RL}. \label{eq:rl}
\end{align}

\subsection{Reconstruction with priors (in progress)}

We would also like the ability to incorporate priors into our reconstruction
algorithms. For example, we may know that we are measuring a single dipole, or
we may know that the orientation distribution function is rotationally
symmetric. All such priors can be formulated as a restriction on the set of
possible orientation distributions. Our goal is to find a way to map our
estimates of the measurable Laplace coefficients to a member of the set of
possible orientation measurements.

Let $\{\mb{z}_1, \mb{z}_2,\ldots \mb{z}_p\}$ denote the \textit{prior set}---the
(possibly infinite) set of orientation distribution functions that could be
present in the sample. We can assemble the set into the rows of a matrix
$\mb{z}\in\mbb{R}^{p\times\infty}$ and apply the Laplace expansion
operator to obtain a matrix $\mb{Z}\in\mbb{R}^{p\times\infty}$ with the
Laplace coefficients of each prior in each row
\begin{align}
  \mb{Z} = \mb{Y}\mb{z}.
\end{align}
When $\mb{Z}$ acts on vectors of Laplace coefficients it returns the expansion
of those Laplace coefficients in terms of the Laplace expansion of members of
the prior set. Therefore, if we have an estimate of the directly measurable
Laplace coefficients $\hat{\mb{F}}$ (from equation \ref{eq:mnls_est} or
\ref{eq:rl_est}), we can estimate the corresponding Laplace coefficients in the
prior set by multiplication with the matrix $\mb{Z}^T\mb{Z}$
\begin{align}
  \hat{\mb{F}}_{\text{prior}} = \mb{Z}^T\mb{Z}\hat{\mb{F}}. \label{eq:prior}
\end{align}
Multiplication by $\mb{Z}^T\mb{Z}$ amounts to projecting the Laplace
coefficients onto the nearest member of the prior set. As usual, we can find
the estimated orientation distribution function with
\begin{align}
  \hat{\mb{f}}_{\text{prior}} = \mb{Y}^T\hat{\mb{F}}_{\text{prior}}.
\end{align}

We need to choose our prior set carefully for a given measurement system. If we
choose a prior set that is too large for a measurement system, then multiple
members of the prior set will give rise to the same intensity
measurements---they will be indistinguishable. We call members of the prior set
that give rise to the same measurements \textit{degenerate}.

\subsection{Single-molecule prior (in progress)}
\textbf{I strongly suspect that} if we can measure all of the spherical
harmonics in a single band and we have a single-molecule prior, then we can
reconstruct the orientation of a single molecule in any orientation. I'm still
developing a proper argument for this.

From \cite{barrett} section 6.7.7: ``$\mbb{SO}(3)$...has a set of irreducible
representations usually denoted by an index $l$ and the $l$th irreducible
representation has $2l+1$ dimensions.'' These irreducible representations
correspond to the rows in Figure XXX.

\subsection{Rotational symmetry prior (in progress)}

\subsection{Cram\'er-Rao lower bounds (in progress)}
In the previous sections we have broken down the reconstruction into three
steps
\begin{enumerate}
\item Estimate the measurable Laplace coefficients using equation
  \ref{eq:mnls_est} or equation \ref{eq:rl_est}.
\item If applicable, use equation \ref{eq:prior} to incorporate priors.
\item Use the Laplace coefficients to calculate the orientation distribution
  function.
\end{enumerate}
If our goal is to estimate the orientation distribution function from intensity
measurements, then we would like to know how noise on our measurements affects
the estimate of the orientation distribution function.

\section{Methods (in progress)}
To demonstrate the value of using harmonic analysis for designing orientation
measurement systems we will work through several examples of orientation
measurement systems.

For each system we will calculate the point response functions, the system matrix,
and the crosstalk matrix. Next, we will optimize the orientation measurement system
and make measurement recommendations. Finally, we will analyze the ability of
the systems to measure the orientation of single fluorophores.

\section{Results (in progress)}
\subsection{Point detectors}
Consider an excited dipole being measured by a single point detector. The point
response function for this measurement system is given by
\begin{align}
  h_1(\theta, \phi) = \sin^2\theta
\end{align}
where $\theta$ is the angle between the dipole axis and the detector. We can
expand the point response function in a Laplace series given by
\begin{align}
  h_1(\theta, \phi) = \frac{2\sqrt{\pi}}{3}y_0^0(\theta, \phi) + \frac{4\sqrt{5\pi}}{15}y_2^0(\theta, \phi).
\end{align}

In other words, a small single detector measures a linear combination of two
spherical harmonics of the orientation distribution function. As expected, we
cannot learn much about the orientation distribution function from this
measurement unless we have a prior that either (1) restricts the orientation
distribution functions to one of the spherical harmonics $y_0^0$ or $y_2^0$ or
(2) restricts the orientation distribution function to set that directly maps to
either $y_0^0$ or $y_2^0$.

If we add two more point detector along orthogonal directions their point response
functions and Laplace expansions are
\begin{align}
  h_2(\theta, \phi) &= \cos^2\theta + \sin^2\theta\sin^2\phi = \frac{4\sqrt{\pi}}{3}y_0^0(\theta, \phi) - \frac{\sqrt{30\pi}}{15}y_2^{-2}(\theta, \phi) + \frac{2\sqrt{5\pi}}{15}y_2^{0}(\theta, \phi) - \frac{\sqrt{30\pi}}{15}y_2^2(\theta, \phi)\\
  h_3(\theta, \phi) &= \cos^2\theta + \sin^2\theta\sin^2\phi = \frac{4\sqrt{\pi}}{3}y_0^0(\theta, \phi) + \frac{\sqrt{30\pi}}{15}y_2^{-2}(\theta, \phi) + \frac{2\sqrt{5\pi}}{15}y_2^{0}(\theta, \phi) + \frac{\sqrt{30\pi}}{15}y_2^2(\theta, \phi)
\end{align}
These detectors add two more measurements, but they also introduce two more
unknowns.

\textbf{Still investigating} I suspect that we need $>=6$ detectors along
independent axes to create an invertible system that can reconstruct the
orientation of single molecules in any orientation. Single detectors measure
linear combinations of spherical harmonics in the $l=0$ (one harmonic) and $l=2$
(five harmonics) bands for a total of six harmonics.

\subsection{Single view microscopes}
\subsubsection{Epi-illumination and epi-detection}
Point response functions for a wide range of microscopes have been calculated in
\cite{chandler}. First, consider a epi-illumination microscope with polarized
illumination and epi-detection. The point response function is
\begin{align}
  h_{\text{epi-epi}}(\theta, \phi) = 2D\{A+B\sin^2\theta + C\sin^2\theta\cos[2(\phi - \phi_{\text{exc}})]\}(A + B\sin^2\theta)
\end{align}
where $A, B, C, D$ are constants depending on the numerical aperture and $\phi_{\text{exc}}$ is the excitation polarization angle. The Laplace expansion is given by
\begin{align*}
  h_{\text{epipol-epi}}(\theta, \phi) = &\frac{2\sqrt{\pi}D}{15}(15A^2 + 20AB + 5AC + 8B^2 + 4BC)y_0^1(\theta, \phi)\\
                                     &-\frac{\sqrt{30\pi}CD}{105}(7iA\sin(2\phi_{\text{exc}}) - 7A\cos(2\phi_{\text{exc}}) + 6iB\sin(2\phi_{\text{exc}}) - 6B\cos(2\phi_{\text{exc}}))y_2^{-2}(\theta, \phi)\\
                                     &-\frac{2\sqrt{5\pi}D}{105}(28AB + 7AC + 16B^2 + 8BC)y_2^0(\theta, \phi)\\
                                     &+\frac{\sqrt{30\pi}CD}{105}(7iA\sin(2\phi_{\text{exc}}) - 7A\cos(2\phi_{\text{exc}}) + 6iB\sin(2\phi_{\text{exc}}) - 6B\cos(2\phi_{\text{exc}}))y_2^{-2}(\theta, \phi)\\
                                     &-\frac{2\sqrt{10\pi}BCD}{105}e^{-2i\phi_{\text{exc}}}y_4^{-2}(\theta, \phi)\\
                                     &+\frac{8\sqrt{\pi}BD}{105}(2B + C)y_4^{0}(\theta, \phi)\\
                                     &-\frac{2\sqrt{10\pi}BCD}{105}e^{-2i\phi_{\text{exc}}}y_4^{-2}(\theta, \phi)\\
\end{align*}
We can see that adding illumination polarizers extends our measurements into the
$l=4$ band.

I am surprised to see complex constants show up here. This could indicate an issue
with my integration program or integrating around poles?

\subsubsection{Orthogonal illumination and detection}
The point response function for low-NA polarized illumination (light sheet
illumination) with an orthogonal detector is
\begin{align}
  h_{\text{orthopol-epi}} = 2(\sin^2\theta\cos^2(\phi - \phi_{\text{exc}}))(A + B(1 - \cos^2\phi\sin^2\theta)).
\end{align}
It's Laplace expansion is
\begin{align*}
  h_{\text{orthopol-epi}} = &\cdots y_0^0(\theta, \phi)\\
                            &+ \cdots y_2^{-2}(\theta, \phi)
                            + \cdots y_2^0(\theta, \phi)
                            + \cdots y_2^2(\theta, \phi)\\
                            &+ \cdots y_4^{-4}(\theta, \phi)
                            + \cdots y_4^{-2}(\theta, \phi)
                            + \cdots y_4^0(\theta, \phi)
                            + \cdots y_4^2(\theta, \phi)
                            + \cdots y_4^4(\theta, \phi)
  \end{align*}

\subsection{Dual view microscopes}
\subsubsection{Epi-illumination and epi-detection}
\subsubsection{Orthogonal illumination and detection}
\section{Discussion and conclusions (in progress)}
\begin{itemize}
\item Single detectors measure members of the $l=0$ and $l=2$ bands. Changing the axis of the detector changes what members of the $l=2$ band are measured. 
\item Polarized illumination with a single detector measures members of the $l=0$, $l=2$, and $l=4$ bands.
\item \textbf{In progress} Polarized illumination and detection measures members of the $l=0$, $l=2$, $l=4$, and $l=6$ bands. 
\end{itemize}

\section*{Funding}
% National Institute of Health (NIH) (R01GM114274, R01EB017293).

\section*{Acknowledgments}
% TC was supported by a University of Chicago Biological Sciences Division
% Graduate Fellowship. HS and PL were supported by Marine Biological Laboratory
% Whitman Center Fellowships. HS was supported by the Intramural Research
% Programs of the National Institute of Biomedical Imaging and Bioengineering.

\section*{Disclosures}
% The authors declare that there are no conflicts of interest related to this article.

% Comment before submission
\bibliography{paper}{}
\bibliographystyle{unsrt}
%\bibliographystyle{default}

\end{document}

### 2017-12-12 harmonic-paper/calculations/single-view.py

@@ -1,0 +1,54 @@
+import numpy as np
+from mpmath import *
+from sympy import *
+from sympy.matrices.dense import *
+import sympy.functions.special.spherical_harmonics as sh
+import functools
+
+print("Working...")
+
+# Symbols
+theta = Symbol('theta')
+phi = Symbol('phi')
+phi_pol = Symbol('phi_pol')
+A = Symbol('A')
+B = Symbol('B')
+C = Symbol('C')
+D = Symbol('D')
+
+# Takes an expression and returns list of coefficients in (l, m, coeff) tuples.
+def ylm_coeffs(model, max_l=4):
+    coeffs = []
+    for n in range(0, max_l+2, 2):
+        complex_coeffs = []
+        for m in range(-n, n+1):
+            print("Integrating: "+ str(n) + ', ' + str(m))            
+            ynm = simplify(expand_trig(sh.Ynm(n, m, theta, phi).expand(func=True).rewrite(cos)))
+            theta_int = integrate(expand(sin(theta)*ynm*model), (theta, 0, pi)) # theta integral
+            final_int = simplify(integrate(expand_trig(theta_int), (phi, 0, 2*pi))) # phi integral
+            complex_coeffs.append(final_int)
+        coeffs.append(c2r_coeffs(complex_coeffs))
+    return np.array(coeffs)
+
+# Convert coefficients of complex Ylm to coefficients of real ylm
+def c2r_coeffs(coeffs):
+    real_coeffs = []
+    mid_ind = floor(len(coeffs)/2)
+    for i, coeff in enumerate(coeffs):
+        if i < mid_ind:
+            real_coeffs.append((I/sqrt(2))*(coeffs[i] + coeffs[-i-1]))
+        elif i == mid_ind:
+            real_coeffs.append(coeffs[i])
+        elif i > mid_ind:
+            real_coeffs.append((1/sqrt(2))*(-coeffs[i] + coeffs[-i-1]))
+    return coeffs
+
+#model = cos(theta)**2 + sin(theta)**2*sin(phi)**2 # x detector
+model = cos(theta)**2 + sin(theta)**2*cos(phi)**2 # y detector
+#model = sin(theta)**2 # z detector
+
+#model =  D*(A + B*(sin(theta)**2) + C*(sin(theta)**2)*(cos(phi - phi_pol)**2))*(A + B*(sin(theta)**2)) # epipol-epi
[…]

### 2017-12-20 harmonic-paper/calculations/dispim-coeffs.py
December 20 group meeting.

@@ -1,0 +1,53 @@
+import numpy as np
+from mpmath import *
+from sympy import *
+from sympy.matrices.dense import *
+import sympy.functions.special.spherical_harmonics as sh
+import functools
+
+print("Working...")
+
+# Symbols
+theta = Symbol('theta', real=True)
+phi = Symbol('phi', real=True)
+Theta = Symbol('Theta', real=True)
+Phi = Symbol('Phi', real=True)
+
+phi_pol = Symbol('phi_pol', real=True)
+A = Symbol('A')
+B = Symbol('B')
+C = Symbol('C')
+D = Symbol('D')
+
+# Takes an expression and returns list of coefficients in (l, m, coeff) tuples.
+def ylm_coeffs(model, max_l=4):
+    coeffs = []
+    for n in range(0, max_l+2, 2):
+        coeff_row = []        
+        for m in range(-n, n+1):
+            print("Integrating: "+ str(n) + ', ' + str(m))            
+            ynm = conjugate(sh.Ynm(n, m, theta, phi).expand(func=True))
+            theta_int = integrate(expand(sin(theta)*ynm*model), (theta, 0, pi)) # theta integral
+            final_int = simplify(integrate(expand_trig(theta_int), (phi, 0, 2*pi))) # phi integral
+            coeff_row.append(final_int)
+        coeffs.append(coeff_row)
+    return np.array(coeffs)
+
+# Other models
+#model1 = 1 - sin(theta)**2*cos(phi)**2 # x detector
+#model1 = 1 - sin(theta)**2*sin(phi)**2 # y detector
+#model1 = sin(theta)**2 # z detector
+#model1 = 1 - (sin(Theta)*cos(Phi)*sin(theta)*cos(phi) + sin(Theta)*sin(Phi)*sin(theta)*sin(phi) + cos(Theta)*cos(theta))**2 # Theta, Phi detector
+#model2 = 1
+#model =  D*(A + B*(sin(theta)**2) + C*(sin(theta)**2)*(cos(phi - phi_pol)**2))*(A + B*(sin(theta)**2)) # epipol-epi
+
+# diSPIM model
+model1 = (sin(theta)**2)*(cos(phi - phi_pol)**2)*2*(A + B*(cos(theta)**2 + sin(theta)**2*sin(phi)**2)) # orthopol-epi (illumination from z - detection on x)
[…]

### 2017-12-20 harmonic-paper/calculations/single-view.py

@@ -1,54 +1,0 @@
-import numpy as np
-from mpmath import *
-from sympy import *
-from sympy.matrices.dense import *
-import sympy.functions.special.spherical_harmonics as sh
-import functools
-
-print("Working...")
-
-# Symbols
-theta = Symbol('theta')
-phi = Symbol('phi')
-phi_pol = Symbol('phi_pol')
-A = Symbol('A')
-B = Symbol('B')
-C = Symbol('C')
-D = Symbol('D')
-
-# Takes an expression and returns list of coefficients in (l, m, coeff) tuples.
-def ylm_coeffs(model, max_l=4):
-    coeffs = []
-    for n in range(0, max_l+2, 2):
-        complex_coeffs = []
-        for m in range(-n, n+1):
-            print("Integrating: "+ str(n) + ', ' + str(m))            
-            ynm = simplify(expand_trig(sh.Ynm(n, m, theta, phi).expand(func=True).rewrite(cos)))
-            theta_int = integrate(expand(sin(theta)*ynm*model), (theta, 0, pi)) # theta integral
-            final_int = simplify(integrate(expand_trig(theta_int), (phi, 0, 2*pi))) # phi integral
-            complex_coeffs.append(final_int)
-        coeffs.append(c2r_coeffs(complex_coeffs))
-    return np.array(coeffs)
-
-# Convert coefficients of complex Ylm to coefficients of real ylm
-def c2r_coeffs(coeffs):
-    real_coeffs = []
-    mid_ind = floor(len(coeffs)/2)
-    for i, coeff in enumerate(coeffs):
-        if i < mid_ind:
-            real_coeffs.append((I/sqrt(2))*(coeffs[i] + coeffs[-i-1]))
-        elif i == mid_ind:
-            real_coeffs.append(coeffs[i])
-        elif i > mid_ind:
-            real_coeffs.append((1/sqrt(2))*(-coeffs[i] + coeffs[-i-1]))
-    return coeffs
-
-#model = cos(theta)**2 + sin(theta)**2*sin(phi)**2 # x detector
-model = cos(theta)**2 + sin(theta)**2*cos(phi)**2 # y detector
-#model = sin(theta)**2 # z detector
-
-#model =  D*(A + B*(sin(theta)**2) + C*(sin(theta)**2)*(cos(phi - phi_pol)**2))*(A + B*(sin(theta)**2)) # epipol-epi
[…]
